### 2015-04-28 inst/examples/synchrony-example.Rmd
Example for get_synchrony function.

# Checking `get_community_synchrony`
A quick check to make sure the function is working as expected.

The `get_community_synchrony` function requires several other packages...
```{r load_libs, echo=TRUE}
library(communitySynchrony)
library(plyr)
library(reshape2)
library(synchrony)
```

### Bring in data
I'm going to use the Kansas data as a test first, then move to the Idaho data.
```{r get_data, echo=TRUE}
site <- "Kansas"
spp_list <- c("BOCU","BOHI","SCSC")
num_spp <- length(spp_list)
ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  ks_data <- rbind(ks_data, spp_data)
} #end species looping for raw data
ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
```

Chengjin identified several quad-years that need to be removed due to dominance of BOCU.
```{r remove_plots, echo=TRUE}
tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
tmp2<-which(ks_data$quad_data=="q27")
tmp3<-which(ks_data$quad=="q28")
tmp4<-which(ks_data$quad=="q30")
tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
ks_data<-ks_data[-tmp,]

# exclude the records later than 1968, to keep the same random year effect...
ks_data<-subset(ks_data,year<68)
```

### Run the function
```{r run_it, echo=TRUE}
out <- get_comm_synchrony(ts_data = ks_data)
str(out)
names(out)
```

### 2015-04-28 inst/examples/synchrony-example.Rmd
adds plotting of per capita growth rates

@@
 library(synchrony)
+library(ggplot2)
 ```
@@
 ```
+
+### Plot per capita growth rates
+```{r plotit, echo=TRUE, include=TRUE}
+pgr_data <- out$growth_rates
+pgr_melt <- melt(pgr_data, id.vars = "year")
+ggplot(pgr_melt, aes(x=(year+1900), y=value, color=variable))+
+  geom_line()+
+  geom_point()
+```

### 2015-04-28 inst/examples/synchrony-example.Rmd
more time series plotting checks

@@
 ```
+
+### Try with Idaho
+```{r idaho_try, echo=TRUE}
+site <- "Idaho"
+spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
+num_spp <- length(spp_list)
+id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  id_data <- rbind(id_data, spp_data)
+} #end species looping for raw data
+id_data <- id_data[2:nrow(id_data),] #remove first NA row
+id_data <- subset(id_data, species!="ARTR") #take out the shrub
+
+out <- get_comm_synchrony(ts_data = id_data)
+str(out)
+names(out)
+```
+
+### Plot Idaho per capita growth rates
+```{r plotit_idaho, echo=TRUE, include=TRUE}
+pgr_data <- out$growth_rates
+pgr_melt <- melt(pgr_data, id.vars = "year")
+ggplot(pgr_melt, aes(x=(year+1900), y=value, color=variable))+
+  geom_line()+
+  geom_point()
+```
+
+### Plot Idaho percent cover through time
+```{r perc_cover, echo=TRUE, include=TRUE}
+ggplot(out$percent_cover, aes(x=(year+1900), y=tot_cover, color=species))+
+  geom_line()+
+  geom_point()
+```

### 2015-04-28 R/community_synchrony.R
adds cover time series to output

@@ -78,5 +78,6 @@ get_comm_synchrony <- function(ts_data){
   return(list(stability = stability,
               pgr_synchrony = growth_rate_synchrony,
               abund_synchrony = synch_abundance,
-              growth_rates = obs_gr))
+              growth_rates = obs_gr,
+              percent_cover = ts_agg))
 }

### 2015-04-29 docs/prelim_analysis.Rmd
initial commit of prelim_analysis and docs/ folder

---
output: pdf_document
---
# Preliminary community synchrony analysis

A couple comments from Claire:

1. In the absence of environmental stochasticity, species fluctuations are not necessarily independent. 
For example in a zero sum game with ecologically equivalent species (Hubbell's neutral model), species fluctuate randomly but they are not independent since the sum is constrained. If there are only two species their fluctuations are necessarily negatively correlated, because the increase in one species must be compensated by a decrease in the other. In a non-neutral model where species have stable equilibrium abundance there are also non independent fluctuations due to species interactions. In competitive communities fluctuations would most of the time have negative correlations overall, but sometimes they can be positive overall.

2. I'm not entirely clear here: null with respect to what? Species interactions, demographic stochasticity and environmental stochasticity can all affect synchrony?

3. Could you show the absolute values, either with three bars per plot, or with a black line that is not at zero but at its unperturbed synchrony value? OK, I see the absolute values in table 1, the differences between sites are quite large compared to the effects of demographic and environmental stochasticities - but it should still be visible. Does this suggest that environmental and demographic stochasticities are minor explanatory variables for community synchrony? What is, then? Diversity?

4. In principle, could the signs of the effects of demographic stochasticity and environmental stochasticity be predicted? For demographic stochasticity, see comment by table 1. For environmental stochasticity: could you quantify the synchrony of species responses to the environment from the synchrony of the intercepts?

5. We could compute the synchrony expected under independent fluctuations: $\sum_{i=1}^{N}{\sigma_{i}^2} / (\sum_{i=1}^{N}{\sigma_{i}})^2$. We could expect demographic stochasticity to tend towards this value: the sign of the effect of demographic stochasticity should be expected synchrony under independent fluctuations - observed synchrony; removal should be the opposite - does that work?

### 2015-04-29 docs/prelim_analysis.Rmd
calculates expected and observed synchrony

@@
 
-5.
-We could compute the synchrony expected under independent fluctuations: $\sum_{i=1}^{N}{\sigma_{i}^2} / (\sum_{i=1}^{N}{\sigma_{i}})^2$.
-We could expect demographic stochasticity to tend towards this value: the sign of the effect of demographic stochasticity should be expected synchrony under independent fluctuations - observed synchrony; removal should be the opposite - does that work?
+5. We could compute the synchrony expected under independent fluctuations: 
+$$\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.$$
+We could expect demographic stochasticity to tend towards this value: the sign of the effect of demographic stochasticity should be $\text{E}(\phi_{r}) - \phi_{r}$, where $\phi_{r}$ is the observed synchrony of per capita growth rates; removal should be the opposite - does that work?
+
+## Calculate expected and observed sychrony
+Under the assumption of independent fluctuations in species' per capit growth rates, we can calculate expected synchrony as:
+
+$$\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }$$
+
+where $\sigma_{i}$ is the standard deviation of species $i$'s per capita growth rate through time.
+
+```{r load_libs, echo=FALSE, message=FALSE}
+library(communitySynchrony)
+library(plyr)
+library(reshape2)
+library(synchrony)
+library(ggplot2)
+```
+
+```{r get_synchs, echo=TRUE, include=TRUE}
+site <- "Kansas"
+spp_list <- c("BOCU","BOHI","SCSC")
+num_spp <- length(spp_list)
+ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  ks_data <- rbind(ks_data, spp_data)
+} #end species looping for raw data
+ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
+
+tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
+tmp2<-which(ks_data$quad_data=="q27")
+tmp3<-which(ks_data$quad=="q28")
+tmp4<-which(ks_data$quad=="q30")
+tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
+tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
+tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
+ks_data<-ks_data[-tmp,]
+
+# exclude the records later than 1968, to keep the same random year effect...
+ks_data<-subset(ks_data,year<68)
+
+out <- get_comm_synchrony(ts_data = ks_data)
+obs_synch <- round(as.numeric(out$pgr_synchrony[1]), 2)
+exp_synch <- round(out$pgr_expected_synch_ind_flucts, 2)
+exp_demo_effect <- exp_synch - obs_synch
+print(out$pgr_synchrony)
+print(out$pgr_expected_synch_ind_flucts)
+```
+
+```{r sign_text, echo=FALSE}
+if(sign(exp_demo_effect)==-1){
+  sign_text <- "decrease"
+  remove_text <- "increase"
+}
+if(sign(exp_demo_effect)==1){
+ sign_text <- "increase"
+ remove_text <- "decrease" 
+}
+```
+
+So, $\text{E}(\phi_{r})$ = `r exp_synch` and $\phi_{r}$ = `r obs_synch`.
+Which means the expected effect of demographic stochasticity [$\text{E}(\sigma_{d}) = \text{E}(\phi_{r}) - \phi_{r}$] is `r exp_demo_effect`.
+So, based on the sign of $\text{E}(\sigma_{d})$ we expect demographic stochasticity to `r sign_text` synchrony among species' per capita growth rates.
+Thus, removing demographic stochasticity should `r remove_text` synchrony in our population model simulations.
+
+Here are the results for both Kansas and Idaho:

### 2015-04-29 docs/prelim_analysis.Rmd
fixes species_list bug in function

@@
 library(ggplot2)
+library(xtable)
 ```
 
-```{r get_synchs, echo=TRUE, include=TRUE}
+```{r get_synchs, echo=FALSE, include=TRUE}
 site <- "Kansas"
@@
 Here are the results for both Kansas and Idaho:
+
+```{r idaho_synch, echo=FALSE, include=TRUE}
+site <- "Idaho"
+spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
+num_spp <- length(spp_list)
+id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  id_data <- rbind(id_data, spp_data)
+} #end species looping for raw data
+id_data <- id_data[2:nrow(id_data),] #remove first NA row
+id_data <- subset(id_data, species!="ARTR") #take out the shrub
+
+out_id <- get_comm_synchrony(ts_data = id_data)
+out_ks <- out
+```
+
+```{r synchrony_table, echo=FALSE, include=TRUE, results='asis'}
+get_table_metrics <- function(output){
+  tmp <- output
+  exp_tmp <- round(tmp$pgr_expected_synch_ind_flucts, 2)
+  obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
+  demo_tmp <- exp_tmp - obs_tmp
+  return(c(obs_tmp, exp_tmp, demo_tmp))
+}
+
+table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
+                                  get_table_metrics(out_id)))
+table_data$Site <-c("Kansas", "Idaho")
+table_data <- table_data[,c("Site", "V1", "V2", "V3")]
+colnames(table_data) <- c("Site",
+                          "$\\phi_{r}$", 
+                          "$\\text{E}(\\phi_{r})$", 
+                          "$\\text{E}(\\sigma_{d})$")
+print(xtable(table_data), include.rownames = F, 
+      sanitize.colnames.function = identity)
+```

### 2015-04-29 R/community_synchrony.R

@@ -64,8 +64,8 @@ get_comm_synchrony <- function(ts_data){
   merged_df <- merge(ts_mat, lag_df[,-rm_col], by.x = "year", by.y="lagyear")
   transitions <- nrow(merged_df)
   obs_gr <- matrix(nrow=transitions, ncol=num_spp)
-  name_ids1 <- which(colnames(merged_df) %in% spp_list)
-  name_ids2 <- which(colnames(merged_df) %in% paste(spp_list, "_t0", sep=""))
+  name_ids1 <- which(colnames(merged_df) %in% species_list)
+  name_ids2 <- which(colnames(merged_df) %in% paste(species_list, "_t0", sep=""))
   for(i in 1:transitions){
     obs_gr[i,] <- as.numeric(log(merged_df[i,name_ids1]/merged_df[i,name_ids2]))
   }

### 2015-04-29 R/community_synchrony.R
melts growth rate data for easy plotting

@@ -87,6 +87,9 @@ get_comm_synchrony <- function(ts_data){
   
   
   ##  Output
+  obs_gr <- melt(obs_gr, id.vars = "year")
+  obs_gr <- obs_gr[with(obs_gr, order(year)), ]
+  colnames(obs_gr) <- c("year","species","pgr")
   return(list(stability = stability,
               pgr_synchrony = growth_rate_synchrony,
               pgr_expected_synch_ind_flucts = expected_synchrony_ind_flucts,

### 2015-04-29 docs/prelim_analysis.Rmd
updated with table and plot for all sites

--- 2015-04-29 docs/prelim_analysis.Rmd
@@
 ---
-output: pdf_document
+output:
+  pdf_document:
+    fig_caption: yes
 ---
+
 # Preliminary community synchrony analysis
@@
 library(ggplot2)
+library(xtable)
+library(gridExtra)
+library(RColorBrewer)
 ```
 
-```{r get_synchs, echo=TRUE, include=TRUE}
+```{r get_all_data, echo=FALSE}
 site <- "Kansas"
@@
 
+site <- "Idaho"
+spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
+num_spp <- length(spp_list)
+id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  id_data <- rbind(id_data, spp_data)
+} #end species looping for raw data
+id_data <- id_data[2:nrow(id_data),] #remove first NA row
+id_data <- subset(id_data, species!="ARTR") #take out the shrub
+
+site <- "Montana"
+spp_list <- sort(c("BOGR","HECO","PASM","POSE"))
+num_spp <- length(spp_list)
+mt_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  mt_data <- rbind(mt_data, spp_data)
+} #end species looping for raw data
+mt_data <- mt_data[2:nrow(mt_data),] #remove first NA row
+
+site <- "NewMexico"
+spp_list <- sort(c("BOER","SPFL"))
+num_spp <- length(spp_list)
+nm_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  nm_data <- rbind(nm_data, spp_data)
+} #end species looping for raw data
+nm_data <- nm_data[2:nrow(nm_data),] #remove first NA row
+
+site <- "Arizona"
+spp_list <- sort(c("BOER","BORO"))
+num_spp <- length(spp_list)
+az_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  az_data <- rbind(az_data, spp_data)
+} #end species looping for raw data
+az_data <- az_data[2:nrow(az_data),] #remove first NA row
+```
+
+```{r get_synchs, echo=FALSE, include=TRUE}
 out <- get_comm_synchrony(ts_data = ks_data)
@@
 Here are the results for both Kansas and Idaho:
+
+```{r get_all_synch, echo=FALSE, include=TRUE}
+out_ks <- out
+out_id <- get_comm_synchrony(ts_data = id_data)
+out_mt <- get_comm_synchrony(ts_data = mt_data)
+out_nm <- get_comm_synchrony(ts_data = nm_data)
+out_az <- get_comm_synchrony(ts_data = az_data)
+```
+
+```{r synchrony_table, echo=FALSE, include=TRUE, results='asis'}
+get_table_metrics <- function(output){
+  tmp <- output
+  exp_tmp <- round(tmp$pgr_expected_synch_ind_flucts, 2)
+  obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
+  demo_tmp <- exp_tmp - obs_tmp
+  return(c(obs_tmp, exp_tmp, demo_tmp))
+}
+
+table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
+                                  get_table_metrics(out_id),
+                                  get_table_metrics(out_mt),
+                                  get_table_metrics(out_nm),
+                                  get_table_metrics(out_az)))
+table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
+table_data <- table_data[,c("Site", "V1", "V2", "V3")]
+table_data$spp <- c(length(unique(ks_data$species)),
+                    length(unique(id_data$species)),
+                    length(unique(mt_data$species)),
+                    length(unique(nm_data$species)),
+                    length(unique(az_data$species)))
+colnames(table_data) <- c("Site",
+                          "$\\phi_{r}$", 
+                          "$\\text{E}(\\phi_{r})$", 
+                          "$\\text{E}(\\sigma_{d})$",
+                          "Species richness")
+
+synch_cap <- "Observed and expected synchrony among species' per capita growth rates and expected effect of demographic stochasticity on observed synchrony.
+Species richness values reflect the number of species analyzed from the community, not the actual richness."
+print(xtable(table_data, caption = synch_cap),
+      caption.placement="top",
+      include.rownames = F, 
+      sanitize.colnames.function = identity)
+```
+
+```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site."}
+out_ks$percent_cover$site <- "Kansas"
+out_id$percent_cover$site <- "Idaho"
+out_mt$percent_cover$site <- "Montana"
+out_nm$percent_cover$site <- "New Mexico"
+out_az$percent_cover$site <- "Arizona"
+cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
+                    out_mt$percent_cover, out_nm$percent_cover, 
+                    out_az$percent_cover)
+
+out_ks$growth_rates$site <- "Kansas"
+out_id$growth_rates$site <- "Idaho"
+out_mt$growth_rates$site <- "Montana"
+out_nm$growth_rates$site <- "New Mexico"
+out_az$growth_rates$site <- "Arizona"
+pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
+                  out_mt$growth_rates, out_nm$growth_rates, 
+                  out_az$growth_rates)
+all_spp <- unique(pgr_data$species)
+color_match <- data.frame(species = all_spp,
+                          color = brewer.pal(11,"Paired"))
+pgr_data <- merge(pgr_data, color_match, by = "species")
+cover_data <- merge(cover_data, color_match, by="species")
+
+cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
+                geom_line()+
+                geom_point()+
+                facet_grid(site~., scales = "free_y")+
+                guides(color=FALSE)+
+                ylab("Mean Percent Cover")+
+                xlab("Year")+
+                scale_color_manual(values = brewer.pal(11,"Paired"))
+
+pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
+              geom_line()+
+              geom_point()+
+              facet_grid(site~., scales = "free_y")+
+              guides(color=FALSE)+
+              ylab("Per Capita Growth Rate")+
+              xlab("Year")+
+              scale_color_manual(values = brewer.pal(11,"Paired"))
+full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
+```

### 2015-04-30 R/vital_rate_regression_functions.R
initial commit of vital rate functions

@@ -1,0 +1,12 @@
+##  Functions using INLA to get vital rate regression parameters
+
+#' Estimate growth regression coefficients using INLA
+#' 
+#' @author Andrew Tredennick
+#' @param dataframe Time series dataframe of genet sizes at t0 and t1 with 
+#'                  neighborhood crowding covariate for each observation.
+#' @return Dataframe with regression coefficients.
+
+get_growth_params <- function(dataframe){
+  
+}

### 2015-04-30 R/vital_rate_regression_functions.R
fixes bug in crowding estimation

@@ -5,8 +5,63 @@
 #' @author Andrew Tredennick
 #' @param dataframe Time series dataframe of genet sizes at t0 and t1 with 
 #'                  neighborhood crowding covariate for each observation.
-#' @return Dataframe with regression coefficients.
+#' @param crowd_mat Matrix of crowding indices for each observation. Matrix
+#'                  dimensions are nrow = number of observations,
+#'                  ncol = number of species.
+#' @return Dataframe with named regression coefficients.
 
-get_growth_params <- function(dataframe){
+get_growth_params <- function(dataframe, crowd_mat){
+  D <- dataframe
+  D$logarea.t0 <- log(D$area.t0)
+  D$logarea.t1 <- log(D$area.t1)
+  D$quad <- as.character(D$quad)
   
+  crowd = crowd_mat 
+  crowd[crowd<1e-99]=0 # make really small crowding indices 0
+  
+  library(INLA)
+  # Set up ID variables for INLA random effects
+  D$yearID <- D$year+max(D$year) # for random year offset on intercept
+  D$GroupID <- as.numeric(D$Group)
+  
+  formula <- logarea.t1 ~ logarea.t0+crowd+
+    f(yearID, model="iid", prior="normal",param=c(0,0.001))+
+    f(GroupID, model="iid", prior="normal",param=c(0,0.001))+
+    f(year, logarea.t0, model="iid", prior="normal",param=c(0,0.001))
+  
+  outINLA <- inla(formula, data=D,
+                  family=c("gaussian"), verbose=FALSE,
+                  control.predictor = list(link = 1),
+                  control.compute=list(dic=T,mlik=T),
+                  control.inla = list(h = 1e-6))
+  
+  # Fit variance
+  x <- outINLA$summary.fitted.values$mean #fitted values from INLA
+  y <- (D$logarea.t1-outINLA$summary.fitted.values$mean)^2 #calculates the variance (residuals^2)
+  outVar <- nls(y~a*exp(b*x),start=list(a=1,b=0)) #model the variance as function of fitted size
+  
+  # Collect parameters
+  # random year and group effects
+  params <- as.data.frame(outINLA$summary.random$yearID[,1:2])
+  params <- cbind(params, outINLA$summary.random$year[,2])
[…]

### 2015-04-30 inst/examples/vital_rate_fxns_example.Rmd
initial commit of test for growth regression function

# Test of INLA regression fitting functions
```{r libs}
library(communitySynchrony)
library(INLA)
```

### Format data
```{r get_data}
sppList=sort(c("PSSP","HECO","POSE","ARTR"))
alpha.effect=c(0.014,0.017,0.026,0.018) # for spp in alphabetical order

# Set up list for storage
all_dataframes <- list()
all_crowding <- list()
for(spp in 1:length(sppList)){
  doSpp=sppList[spp]
  growDfile=paste("../../data/Idaho/",doSpp,"/growDnoNA.csv",sep="")
  growD=read.csv(growDfile)
  D <- growD  #subset(growD,allEdge==0)
  D$logarea.t0 <- log(D$area.t0)
  D$logarea.t1 <- log(D$area.t1)
  D$quad <- as.character(D$quad)

  # remove outliers (large plants that obviously do not turn into tiny plants) for ARTR only
  if(doSpp=="ARTR"){
    tmp=which(D$quad=="Q23" & D$year==45 & D$trackID==67)
    tmp=c(tmp,which(D$quad=="Q12" & D$year==55 & D$trackID==25))
    tmp=c(tmp,which(D$quad=="Q26" & D$year==45 & D$trackID==73))
    D=D[-tmp,]
  }else{
    D=D
  }

  # calculate crowding 
  for(i in 1:length(sppList)){
    distDfile=paste("../../data/Idaho/",sppList[i],"/",sppList[i],"_genet_xy.csv",sep="")
    if(i==1){
      distD=read.csv(distDfile)
      distD$nbSpp=sppList[i]  
    }else{
      tmp=read.csv(distDfile)
      tmp$nbSpp=sppList[i] 
      distD=rbind(distD,tmp)
    }
  }

  distD=distD[,c("quad","year","trackID","area","nbSpp","x","y")]
  W=matrix(NA,dim(D)[1],length(sppList))
  for(i in 1:dim(D)[1]){
    tmpD=subset(distD,year==D$year[i] & quad==D$quad[i])
    focal=which(tmpD$trackID==D$trackID[i] & tmpD$nbSpp==doSpp)
    xx=tmpD$x[focal] ; yy=tmpD$y[focal]
    tmpD$distance=sqrt((xx-tmpD$x)^2+(yy-tmpD$y)^2)
    tmpD=subset(tmpD,distance>0)
    if(dim(tmpD)[1]>0){
      for(k in 1:length(sppList)){
        sppI=which(tmpD$nbSpp==sppList[k])
        if(length(sppI)>0){
          W[i,k]=sum(exp(-1*alpha.effect[k]*tmpD$distance[sppI]^2)*tmpD$area[sppI])         
        }else{
          W[i,k]=0
        }
      }
    }else{
      W[i,]=0
    }   
  }

  all_crowding[[spp]] <- W
  all_dataframes[[spp]] <- D
}
names(all_crowding) <- sppList
names(all_dataframes) <- sppList
```

### Call regression function for growth
```{r call_growth_fxn}
growth_params <- list()
for(spp in sppList){
  tmp <- get_growth_params(dataframe = all_dataframes[[spp]],
                           crowd_mat = all_crowding[[spp]])
  growth_params[[spp]] <- tmp
}
```

### 2015-04-30 inst/examples/vital_rate_fxns_example.Rmd
adds plots of yearly intercepts and slopes for check

@@
 library(INLA)
+library(reshape2)
+library(ggplot2)
 ```
@@
 ```
+
+### Plot yearly intercepts and slopes
+```{r plots}
+intercepts <- matrix(ncol=length(sppList), nrow=nrow(growth_params[[1]]))
+slopes <- matrix(ncol=length(sppList), nrow=nrow(growth_params[[1]]))
+i <- 1
+for(spp in sppList){
+  intercepts[,i] <- growth_params[[spp]]$Intercept.yr
+  slopes[,i] <- growth_params[[spp]]$logarea.t0.yr
+  i <- i+1
+}
+colnames(intercepts) <- sppList
+colnames(slopes) <- sppList
+intm <- melt(intercepts)
+slopem <- melt(slopes)
+intm$variable <- "intercept"
+slopem$variable <- "slope"
+combo <- rbind(intm, slopem)
+ggplot(subset(combo, Var2!="ARTR"), aes(x=Var1, y=value, color=Var2))+
+  geom_line()+
+  geom_point()+
+  facet_grid(variable~., scales="free_y")
+```

### 2015-04-30 R/crowding_functions.R
initial commit of crowding estimation function

@@ -1,0 +1,15 @@
+##  Functions to estimate crowding effects on focal species
+
+#' Estimate growth crowding effects on focal species
+#' 
+#' @author Andrew Tredennick
+#' @param site Name of the focal site (character).
+#' @param data_path The directory path leading to the data folder (character).
+#' @param alphas The alpha values for each species at the site, in alphabetical
+#'               order by species.
+#' @return List of crowding matrices, one list entry per species.
+
+estimate_crowding <- function(site, data_path, alphas){
+  species_list <- sort(list.files(data_path))
+  alpha.effect <- alphas
+}

### 2015-04-30 R/crowding_functions.R
adds species number variable

@@ -11,5 +11,7 @@
 
 estimate_crowding <- function(site, data_path, alphas){
   species_list <- sort(list.files(data_path))
+  num_species <- length(species_list)
   alpha.effect <- alphas
+  all_crowding <- list()
 }

### 2015-05-05 docs/prelim_analysis.Rmd
writing IPM description

@@
 ```
+
+# Methods
+## Integral projection models (IPM)
+We built an environmentally and demographically stochastic integral projection model (IPM), where either environmental or demographic stochasticity can be turned off.
+Our IPM follows the specification of Chu and Adler (2015) where the population of species __j__ is a density function $n(u_{j},t)$ giving the density of sized-__u__ genets at time __t__.
+Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$.
+So, the density function for any size __v__ at time $t+1$ is
+
+$$
+n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bold{w_{j}}(u_{j}))n(u_{j},t)
+$$

### 2015-05-06 runscripts/run_ipm_simulations.R
commits prelim ipm simulation results

@@ -21,10 +21,31 @@ for(do_site in site_list){
   Rpars_tmp <- Rpars_all[[do_site]]
   
 #   if(do_site=="Idaho"){
+#     spp_list <- names(Gpars_tmp)
+#     Rp <- as.data.frame(Rpars_tmp)
+#     Rp$species <- c(rep(spp_list, each=4),
+#                     rep(spp_list, each=1),
+#                     rep(spp_list, each=6),
+#                     rep(spp_list, each=1),
+#                     rep(spp_list, each=1),
+#                     rep(spp_list, each=22),
+#                     rep(spp_list, each=412),
+#                     rep(spp_list, each=1),
+#                     rep(spp_list, each=1))
+#     Rp <- subset(Rp, species!="ARTR")
+#     colid <- which(colnames(Rp)=="species")
+#     Rp <- Rp[,-colid]
+#     Rp$species <- c(rep(c("ARTR", "fine", "fine", "fine"), 3),
+#                     rep("fine", (nrow(Rp)-12)))
+#     Rp <- subset(Rp, species=="fine")
+#     colid <- which(colnames(Rp)=="species")
+#     Rpars_tmp <- as.matrix(Rp[,-colid])
+#   }
+#   
+#   if(do_site=="Idaho"){
 #     id <- which(names(Gpars_tmp)=="ARTR")
 #     Gpars_tmp <- Gpars_tmp[-id]
 #     Spars_tmp <- Spars_tmp[-id] 
-# #     Rpars_tmp <- Rpars_tmp[[id]] <- NULL
 #   }
   
   spp_list <- names(Gpars_tmp)
@@ -70,8 +91,8 @@ for(do_site in site_list){
     do_demo_stoch <- do_demo_stoch_vec[stoch]
     n_spp <- Nspp <- length(spp_list)
     A=10000
-    tlimit=50
-    burn_in=10
+    tlimit=2500
+    burn_in=500
     spp_list=spp_list
     Nyrs=Nyrs; constant=do_env_const
     iter_matrix_dims=iter_matrix_dims; max_size=max_size
@@ -93,5 +114,5 @@ for(do_site in site_list){
 } # end site loop
 
 # Save the output
-# saveRDS(output_list, "../results/ipm_simulation_lists.RDS")
+saveRDS(output_list, "../results/ipm_simulation_lists.RDS")
 

### 2015-05-06 runscripts/run_ipm_source.R

@@ -34,6 +34,8 @@ NoOverlap_Inter=FALSE
   
   # loop through species really quick to set low initial densities
   for(i in 1:n_spp) nt[[i]][]=0.001
+  if(do_site=="Idaho")
+    nt[[1]][] <- 0
   new.nt <- nt #set initial density vector to be fed into IPM
   
   # set up matrix to record cover

### 2015-05-06 inst/examples/ipm_sims_example.Rmd
new example checking IPM output

---
output: html_document
---
# Checking IPM simulation output
```{r read_data}
library(plyr)
library(reshape2)
library(ggplot2)
output_list <- readRDS("../../results/ipm_simulation_lists.RDS")
mlist <- melt(output_list)
```

## Look at simulated time series
```{r time_series}
ggplot(mlist, aes(x=Var1, y=(value*100), color=Var2))+
  geom_line()+
  facet_grid(L1~L2, scale="free")
```
Hmm... looks like something is wrong with a Kansas species. Also need to compare these results to the one's from Idaho-only script in Diversity-Stability repo.

## Run through synchrony library
```{r get_synchs}
library(communitySynchrony)
get_comm_synchrony(ts_data = subset(mlist, L2=="ENVDEMO" & L1=="Arizona"))
```

### 2015-05-13 runscripts/call_vitalrate_fxns.R
merges crowding by id and outputs as matrix for regressions

@@ -79,6 +79,8 @@ for(do_site in site_list){
       
     # Get correct crowding matrix
     crowd_growth_now <- crowd_growth[[do_site]][[do_species]]
+    tmpmerge <- merge(D, crowd_growth_now, by.x="X", by.y="xID")
+    crowd_growth_now <- as.matrix(tmpmerge[,c("V1","V2")])
     
     # Run through the function
     tmp <- get_growth_params(dataframe = D,

### 2015-05-13 docs/prelim_analysis.Rmd
adds text on IPM and study species to prelim manuscript

--- 2015-04-29 docs/prelim_analysis.Rmd
@@
 ```
+
+# Methods
+## Study sites and species
+We use the exact same data as presented in Chu and Adler (2015). From their paper:
+
+>Many range experiment stations in the western U.S.
+began mapping permanent quadrats in the early 20th century, and continued annual censuses for decades.
+Here we focus on five long-term data sets (Fig.
+1), four of which have been distributed publicly (Adler et al.
+2007b, Zachmann et al.
+2010, Anderson et al.
+2011, 2012).
+At each study site, all individual plants within each 1-m2 quadrat were identified and mapped annually using a pantograph (Hill 1920; Fig.
+2).
+Mapped polygons represent basal cover for grasses and canopy cover for shrubs (Fig.
+2).
+
+>To fit our population models, we needed to study species common enough to provide a large sample size (i.e.
+the most common species at each site).
+In order to describe species interactions, we needed to select species that co-occurred together across many years.
+Thus, our analysis focused on the niche differences and average fitness differences of the most common, co-occurring species at each site, which implies that the average fitness differences among species could be relatively small.
+To select these species, we first identified species that occurred in at least 20% of years.
+Then we used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
+Based on the degree of aggregation of quadrats on the NMDS plot, we selected quadrats and species for our full analysis.
+We also confirmed these selections with local personnel from each study site.
+Below we further describe each study site and our criteria for quadrat and species selections.
+
+>_Sonoran desert, Arizona_
+
+>The Sonoran desert data set comes from 178 permanent quadrats located on semi-desert grasslands at the Santa Rita Experimental Range (31º50’ N, 110º53’ W; elevation 1150 meters), Arizona (Anderson et al.
+2012).
+Mean annual precipitation was 450 mm with the monthly variation from 5 mm (May) to 107 mm (July), and mean annual temperature was 16 °C with the monthly variation from 8 °C (January) to 26 °C (July) during the sampling period.
+These quadrats were mapped annually from 1915 to 1933 in most cases.
+The dominant flora varies across the broad range of soils and topography that occur within the semi-desert grasslands.
+For our present analysis, we chose 32 quadrats dominated by black grama Bouteloua eriopoda (BOER) and Bouteloua rothrockii (BORO).
+
+>_Sagebrush steppe, Idaho_
+
+>The sagebrush data set comes from the U.S.
+Sheep Experiment Station, currently a US Department of Agriculture Agricultural Research Service field station, located 9.6 km north of Dubois, Idaho (44.2°N, 112.1°W; elevation 1500 meters).
+26 quadrats were established between 1926 and 1932 (Zachmann et al.
+2010).
+During the period of sampling (1926 – 1957), mean annual precipitation was 270 mm with the monthly variation from 16 mm (March) to 42 mm (June), and mean annual temperature was 6 °C with the monthly variation from -8 °C (January) to 21 °C (July).
+The vegetation is dominated by the shrub, Artemisia tripartita (ARTR), and the C3 perennial bunchgrasses Pseudoroegneria spicata (PSSP), Hesperostipa comata (HECO), and Poa secunda (POSE).
+
+>_Southern mixed prairie, Kansas_
+
+>The southern mixed prairie data set represents the pioneering work of Albertson and colleagues (Albertson and Tomanek 1965), who established 51 permanent quadrats inside and outside livestock exclosures near Hays, Kansas (38.8°N, 99.3°W; elevation 650 meters).
+The mean annual precipitation was 580 mm with the monthly variation from 10 mm (January) to 103 mm (June), and mean annual temperature was 12 °C with the monthly variation from -2.0 °C (January) to 27 °C (July).
+These quadrats were mapped from 1932 to 1972 at the end of each growing season (Adler et al.
+2007b).
+Variation in soil depth and texture creates distinct plant communities.
+Many of the quadrats were located on shortgrass communities dominated by Bouteloua gracilis and Buchloë dactyloides.
+We had difficulty working with these species, which can be hard to distinguish in the field in the absence of inflorescences, and were often mapped inconsistently.
+Therefore, we focused on 7 quadrats dominated by Bouteloua curtipendula (BOCU), Bouteloua hirsuta (BOHI), and Schizachyrium scoparium (SCSC).
+
+>_Northern mixed prairie, Montana_
+
+>The northern mixed prairie data set comes from the Fort Keogh Livestock and Range Research Laboratory, another USDA Agricultural Research Service field station.
+The study site is located on alluvial plains near the Tongue River (46o19’N, 105o48’W; elevation 720 meters).
+Mean annual precipitation was 343 mm with the monthly variation from 7 mm (February) to 74 mm (June), and mean annual temperature was 8 °C with the monthly variation from -7 °C (January) to 25 °C (July).
+44 quadrats, distributed across six pastures assigned to one of three grazing intensities, were mapped annually (with some exceptions) from 1932 through 1945 (Anderson et al.
+2011).
+Our analysis includes 19 quadrats in which Bouteloua gracilis (BOGR), Hesperostipa comata (HECO), Pascopyrum smithii (PASM), and Poa secunda (POSE) were prevalent.
+
+>_Chihuahuan desert, New Mexico_
+
+>The last data set consists of 76 permanent quadrats located in Chihuahuan desert plant communities at the Jornada Experimental Range (32.62°N, 106.67°W; elevation 1260 meters), New Mexico.
+Mean annual precipitation was 264 mm with the monthly variation from 6 mm (April) to 48 mm (August), and mean annual temperature was 14 °C with the monthly variation from 4 °C (January) to 26 °C (July).
+Quadrats were established from 1915-1920 and mapping was conducted annually until the 1960’s (it continues on roughly 5 yr intervals to the present).
+Due to the changes in mapping frequency in the 1960’s, and the effects of a severe drought in the 1950’s that triggered a conversion of many quadrats from grassland to shrubland, we analyzed data for the period from 1915 to 1950.
+We selected 48 quadrats in which both Bouteloua eriopoda (BOER) and Sporobolus flexuosus (SPFL) were common (Anderson et al.
+in preparation).
+
+## Analytical approach
+While the observational data allows to directly calculate species synchrony within the community, it is impossible to isolate the contributions of demographic and environmental stochasticity on species synchrony based on the data alone.
+So we now turn to a empirically-based modeling approach where we can turn demographic and environmental stochasticity on/off.
+To our knowledge, Integral Projection Models (IPMs) are the only tools available that allow us to isolate sources of stochasticity.
+For example, individual-based models require demographic stochasticity, making it impossible to isolate the effect of environmental stochasticity alone.
+
+We describe our approach in three steps.
+First, we describe the statistical models for three vitale rates: survival, growth, and recruitment.
+Second, we describe the IPM construction in detail.
+Third, we describe the IPM simulations we conducted to answer our questions.
+
+## Statistical models for vital rates
+
+## Integral projection models (IPM)
+We built an environmentally and demographically stochastic integral projection model (IPM), where either environmental or demographic stochasticity can be turned off.
+Our IPM follows the specification of Chu and Adler (2015) where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_.
+Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$.
+So, the density function for any size _v_ at time $t+1$ is
+
+$$
+n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
+$$
+
+where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal that describes all possible transitions from size $u$ to $v$ and $\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding experienced from all other species by a genet of size $u_j$ and species $j$.
+The integral is evaluated over all possible sizes between predefined lower (_L_) and upper (_U_) size limits that extend beyond the range of observed genet sizes.
+
+The population kernal is defined as the joint contributions of survival (_S_), growth (_G_), and recruitment (_R_):
+
+$$
+k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
+$$
+
+which, said plainly, means we are calculating growth (_G_) for individuals that survive (_S_) from time _t_ to _t+1_ and adding in newly recruited (_R_) individuals of an average sized one-year-old genet for the focal species.
+Our stastical model for recruitment (_R_, described below) returns the number of new recruit produced per quadrat.
+Following previous work, we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM.

### 2015-05-13 docs/prelim_analysis.Rmd
updates to prelim document

@@
   pdf_document:
-    fig_caption: yes
+    fig_caption: true
+    keep_tex: true
+fontsize: 12pt
+geometry: margin=1in
 ---
@@
 
-## Statistical models for vital rates
+## Statistical models for survival, growth, and recruitment
+### Survival and growth
+We modeled survival probability and growht on individual genets as a function of genet size, the crowding experienced by the focal genet from both hetero and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
+Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure).
+
+We follow the approach of Chu and Adler (2015) to estimate crowding, assuming that the crowding experienced by a focal genet depends on distance to each neighbor genet and the neighbor's size, _u_:
+
+$$
+w_{ijm,t} = \sum_k e^{-\delta_{jm}d_{ijkm,t}^{2}}u_{km,t}.
+$$
+
+In the above, $w_{ijm,t}$ is the crowding that genet _i_ of species _j_ in year _t_ experiences from neighbors of species _m_.
+The spatial scale over which species _m_ neighbors exert influence on any genet of species _j_ is determined by $\delta_{jm}$.
+The function is applied for all _k_ genets of species _m_ that neighbor the focal genet at time _t_, and $d_{ijkm,t}$ is the distance between genet _i_ in species _j_ and genet _k_ in species _m_.
+When $k=m$, the effect is intraspecific crowding.
+We use regression-specific (survival and growth) $\delta$ values estimated by Chu and Adler (2015).
 

### 2015-05-13 docs/prelim_analysis.tex
updates to prelim document

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}
\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\begin{document}

\maketitle

\section{Preliminary community synchrony
analysis}\label{preliminary-community-synchrony-analysis}

A couple comments from Claire:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  In the absence of environmental stochasticity, species fluctuations
  are not necessarily independent. For example in a zero sum game with
  ecologically equivalent species (Hubbell's neutral model), species
  fluctuate randomly but they are not independent since the sum is
  constrained. If there are only two species their fluctuations are
  necessarily negatively correlated, because the increase in one species
  must be compensated by a decrease in the other. In a non-neutral model
  where species have stable equilibrium abundance there are also non
  independent fluctuations due to species interactions. In competitive
  communities fluctuations would most of the time have negative
  correlations overall, but sometimes they can be positive overall.
\item
  I'm not entirely clear here: null with respect to what? Species
  interactions, demographic stochasticity and environmental
  stochasticity can all affect synchrony?
\item
  Could you show the absolute values, either with three bars per plot,
  or with a black line that is not at zero but at its unperturbed
  synchrony value? OK, I see the absolute values in table 1, the
  differences between sites are quite large compared to the effects of
  demographic and environmental stochasticities - but it should still be
  visible. Does this suggest that environmental and demographic
  stochasticities are minor explanatory variables for community
  synchrony? What is, then? Diversity?
\item
  In principle, could the signs of the effects of demographic
  stochasticity and environmental stochasticity be predicted? For
  demographic stochasticity, see comment by table 1. For environmental
  stochasticity: could you quantify the synchrony of species responses
  to the environment from the synchrony of the intercepts?
\item
  We could compute the synchrony expected under independent
  fluctuations:
  \[\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.\]
  We could expect demographic stochasticity to tend towards this value:
  the sign of the effect of demographic stochasticity should be
  \(\text{E}(\phi_{r}) - \phi_{r}\), where \(\phi_{r}\) is the observed
  synchrony of per capita growth rates; removal should be the opposite -
  does that work?
\end{enumerate}

\subsection{Calculate expected and observed
sychrony}\label{calculate-expected-and-observed-sychrony}

Under the assumption of independent fluctuations in species' per capit
growth rates, we can calculate expected synchrony as:

\[\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }\]

where \(\sigma_{i}\) is the standard deviation of species \(i\)'s per
capita growth rate through time.

\begin{verbatim}
## Warning: package 'synchrony' was built under R version 3.1.2
\end{verbatim}

\begin{verbatim}
## Community synchrony: 0.543
## Mean pairwise correlation: 0.319
\end{verbatim}

\begin{verbatim}
## [1] 0.3369
\end{verbatim}

So, \(\text{E}(\phi_{r})\) = 0.34 and \(\phi_{r}\) = 0.54. Which means
the expected effect of demographic stochasticity
{[}\(\text{E}(\sigma_{d}) = \text{E}(\phi_{r}) - \phi_{r}\){]} is -0.2.
So, based on the sign of \(\text{E}(\sigma_{d})\) we expect demographic
stochasticity to decrease synchrony among species' per capita growth
rates. Thus, removing demographic stochasticity should increase
synchrony in our population model simulations.

Here are the results for both Kansas and Idaho:

\% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Wed May
13 14:53:49 2015

\begin{table}[ht]
\centering
\caption{Observed and expected synchrony among species' per capita growth rates and expected effect of demographic stochasticity on observed synchrony. Species richness values reflect the number of species analyzed from the community, not the actual richness.} 
\begin{tabular}{lrrrr}
  \hline
Site & $\phi_{r}$ & $\text{E}(\phi_{r})$ & $\text{E}(\sigma_{d})$ & Species richness \\ 
  \hline
Kansas & 0.54 & 0.34 & -0.20 &   3 \\ 
  Idaho & 0.63 & 0.35 & -0.28 &   3 \\ 
  Montana & 0.53 & 0.27 & -0.26 &   4 \\ 
  New Mexico & 0.86 & 0.51 & -0.35 &   2 \\ 
  Arizona & 0.89 & 0.52 & -0.37 &   2 \\ 
   \hline
\end{tabular}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics{prelim_analysis_files/figure-latex/time_series_fig.pdf}
\caption{Observed time series of mean percent cover (averaged over
quadrats for each year) and per capita growth rates for each species at
each site.}
\end{figure}

\section{Methods}\label{methods}

\subsection{Study sites and species}\label{study-sites-and-species}

We use the exact same data as presented in Chu and Adler (2015). From
their paper:

\begin{quote}
Many range experiment stations in the western U.S. began mapping
permanent quadrats in the early 20th century, and continued annual
censuses for decades. Here we focus on five long-term data sets (Fig.
1), four of which have been distributed publicly (Adler et al. 2007b,
Zachmann et al. 2010, Anderson et al. 2011, 2012). At each study site,
all individual plants within each 1-m2 quadrat were identified and
mapped annually using a pantograph (Hill 1920; Fig. 2). Mapped polygons
represent basal cover for grasses and canopy cover for shrubs (Fig. 2).
\end{quote}

\begin{quote}
To fit our population models, we needed to study species common enough
to provide a large sample size (i.e.~the most common species at each
site). In order to describe species interactions, we needed to select
species that co-occurred together across many years. Thus, our analysis
focused on the niche differences and average fitness differences of the
most common, co-occurring species at each site, which implies that the
average fitness differences among species could be relatively small. To
select these species, we first identified species that occurred in at
least 20\% of years. Then we used Nonmetric Multidimensional Scaling
(NMDS) to identify quadrats with a similar composition of these
candidate species. Based on the degree of aggregation of quadrats on the
NMDS plot, we selected quadrats and species for our full analysis. We
also confirmed these selections with local personnel from each study
site. Below we further describe each study site and our criteria for
quadrat and species selections.
\end{quote}

\begin{quote}
\emph{Sonoran desert, Arizona}
\end{quote}

\begin{quote}
The Sonoran desert data set comes from 178 permanent quadrats located on
semi-desert grasslands at the Santa Rita Experimental Range (31º50' N,
110º53' W; elevation 1150 meters), Arizona (Anderson et al. 2012). Mean
annual precipitation was 450 mm with the monthly variation from 5 mm
(May) to 107 mm (July), and mean annual temperature was 16 °C with the
monthly variation from 8 °C (January) to 26 °C (July) during the
sampling period. These quadrats were mapped annually from 1915 to 1933
in most cases. The dominant flora varies across the broad range of soils
and topography that occur within the semi-desert grasslands. For our
present analysis, we chose 32 quadrats dominated by black grama
Bouteloua eriopoda (BOER) and Bouteloua rothrockii (BORO).
\end{quote}

\begin{quote}
\emph{Sagebrush steppe, Idaho}
\end{quote}

\begin{quote}
The sagebrush data set comes from the U.S. Sheep Experiment Station,
currently a US Department of Agriculture Agricultural Research Service
field station, located 9.6 km north of Dubois, Idaho (44.2°N, 112.1°W;
elevation 1500 meters). 26 quadrats were established between 1926 and
1932 (Zachmann et al. 2010). During the period of sampling (1926 --
1957), mean annual precipitation was 270 mm with the monthly variation
from 16 mm (March) to 42 mm (June), and mean annual temperature was 6 °C
with the monthly variation from -8 °C (January) to 21 °C (July). The
vegetation is dominated by the shrub, Artemisia tripartita (ARTR), and
the C3 perennial bunchgrasses Pseudoroegneria spicata (PSSP),
Hesperostipa comata (HECO), and Poa secunda (POSE).
\end{quote}

\begin{quote}
\emph{Southern mixed prairie, Kansas}
\end{quote}

\begin{quote}
The southern mixed prairie data set represents the pioneering work of
Albertson and colleagues (Albertson and Tomanek 1965), who established
51 permanent quadrats inside and outside livestock exclosures near Hays,
Kansas (38.8°N, 99.3°W; elevation 650 meters). The mean annual
precipitation was 580 mm with the monthly variation from 10 mm (January)
to 103 mm (June), and mean annual temperature was 12 °C with the monthly
variation from -2.0 °C (January) to 27 °C (July). These quadrats were
mapped from 1932 to 1972 at the end of each growing season (Adler et al.
2007b). Variation in soil depth and texture creates distinct plant
communities. Many of the quadrats were located on shortgrass communities
dominated by Bouteloua gracilis and Buchloë dactyloides. We had
difficulty working with these species, which can be hard to distinguish
in the field in the absence of inflorescences, and were often mapped
inconsistently. Therefore, we focused on 7 quadrats dominated by
Bouteloua curtipendula (BOCU), Bouteloua hirsuta (BOHI), and
Schizachyrium scoparium (SCSC).
\end{quote}

\begin{quote}
\emph{Northern mixed prairie, Montana}
\end{quote}

\begin{quote}
The northern mixed prairie data set comes from the Fort Keogh Livestock
and Range Research Laboratory, another USDA Agricultural Research
Service field station. The study site is located on alluvial plains near
the Tongue River (46o19'N, 105o48'W; elevation 720 meters). Mean annual
precipitation was 343 mm with the monthly variation from 7 mm (February)
to 74 mm (June), and mean annual temperature was 8 °C with the monthly
variation from -7 °C (January) to 25 °C (July). 44 quadrats, distributed
across six pastures assigned to one of three grazing intensities, were
mapped annually (with some exceptions) from 1932 through 1945 (Anderson
et al. 2011). Our analysis includes 19 quadrats in which Bouteloua
gracilis (BOGR), Hesperostipa comata (HECO), Pascopyrum smithii (PASM),
and Poa secunda (POSE) were prevalent.
\end{quote}

\begin{quote}
\emph{Chihuahuan desert, New Mexico}
\end{quote}

\begin{quote}
The last data set consists of 76 permanent quadrats located in
Chihuahuan desert plant communities at the Jornada Experimental Range
(32.62°N, 106.67°W; elevation 1260 meters), New Mexico. Mean annual
precipitation was 264 mm with the monthly variation from 6 mm (April) to
48 mm (August), and mean annual temperature was 14 °C with the monthly
variation from 4 °C (January) to 26 °C (July). Quadrats were established
from 1915-1920 and mapping was conducted annually until the 1960's (it
continues on roughly 5 yr intervals to the present). Due to the changes
in mapping frequency in the 1960's, and the effects of a severe drought
in the 1950's that triggered a conversion of many quadrats from
grassland to shrubland, we analyzed data for the period from 1915 to
1950. We selected 48 quadrats in which both Bouteloua eriopoda (BOER)
and Sporobolus flexuosus (SPFL) were common (Anderson et al. in
preparation).
\end{quote}

\subsection{Analytical approach}\label{analytical-approach}

While the observational data allows to directly calculate species
synchrony within the community, it is impossible to isolate the
contributions of demographic and environmental stochasticity on species
synchrony based on the data alone. So we now turn to a empirically-based
modeling approach where we can turn demographic and environmental
stochasticity on/off. To our knowledge, Integral Projection Models
(IPMs) are the only tools available that allow us to isolate sources of
stochasticity. For example, individual-based models require demographic
stochasticity, making it impossible to isolate the effect of
environmental stochasticity alone.

We describe our approach in three steps. First, we describe the
statistical models for three vitale rates: survival, growth, and
recruitment. Second, we describe the IPM construction in detail. Third,
we describe the IPM simulations we conducted to answer our questions.

\subsection{Statistical models for survival, growth, and
recruitment}\label{statistical-models-for-survival-growth-and-recruitment}

\subsubsection{Survival and growth}\label{survival-and-growth}

We modeled survival probability and growht on individual genets as a
function of genet size, the crowding experienced by the focal genet from
both hetero and conspecific genets in its neighborhood (described
below), temporal varation among years, and spatial variation among
quadrat groups. Groups are sets of quadrats located in close proximity
within a pasture or grazing exclosure).

We follow the approach of Chu and Adler (2015) to estimate crowding,
assuming that the crowding experienced by a focal genet depends on
distance to each neighbor genet and the neighbor's size, \emph{u}:

\[
w_{ijm,t} = \sum_k e^{-\delta_{jm}d_{ijkm,t}^{2}}u_{km,t}.
\]

In the above, \(w_{ijm,t}\) is the crowding that genet \emph{i} of
species \emph{j} in year \emph{t} experiences from neighbors of species
\emph{m}. The spatial scale over which species \emph{m} neighbors exert
influence on any genet of species \emph{j} is determined by
\(\delta_{jm}\). The function is applied for all \emph{k} genets of
species \emph{m} that neighbor the focal genet at time \emph{t}, and
\(d_{ijkm,t}\) is the distance between genet \emph{i} in species
\emph{j} and genet \emph{k} in species \emph{m}. When \(k=m\), the
effect is intraspecific crowding.

\subsection{Integral projection models
(IPM)}\label{integral-projection-models-ipm}

We built an environmentally and demographically stochastic integral
projection model (IPM), where either environmental or demographic
stochasticity can be turned off. Our IPM follows the specification of
Chu and Adler (2015) where the population of species \emph{j} is a
density function \(n(u_{j},t)\) giving the density of sized-\emph{u}
genets at time \emph{t}. Genet size is on the natural log scale, so that
\(n(u_{j},t)du\) is the number of genets whose area (on the arithmetic
scale) is between \(e^{u_{j}}\) and \(e^{u_{j}+du}\). So, the density
function for any size \emph{v} at time \(t+1\) is

\[
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
\]

where \(k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})\) is the population kernal
that describes all possible transitions from size \(u\) to \(v\) and
\(\bar{\bold{w_{j}}}\) is a vector of estimates of average crowding
experienced from all other species by a genet of size \(u_j\) and
species \(j\). The integral is evaluated over all possible sizes between
predefined lower (\emph{L}) and upper (\emph{U}) size limits that extend
beyond the range of observed genet sizes.

The population kernal is defined as the joint contributions of survival
(\emph{S}), growth (\emph{G}), and recruitment (\emph{R}):

\[
k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
\]

which, said plainly, means we are calculating growth (\emph{G}) for
individuals that survive (\emph{S}) from time \emph{t} to \emph{t+1} and
adding in newly recruited (\emph{R}) individuals of an average sized
one-year-old genet for the focal species. Our stastical model for
recruitment (\emph{R}, described below) returns the number of new
recruit produced per quadrat. Following previous work, we assume that
fecundity increases linearly with size
(\(R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})\))
to incorporate the recruitment function in the spatially-implicit IPM.

\end{document}

### 2015-05-14 inst/examples/ipm_sims_example.Rmd
updates ipm simulations after removing quad-years as per Chu and Adler

@@
 ```
+
 Hmm... looks like something is wrong with a Kansas species. Also need to compare these results to the one's from Idaho-only script in Diversity-Stability repo.
@@
 ```{r get_synchs}
-library(communitySynchrony)
-get_comm_synchrony(ts_data = subset(mlist, L2=="ENVDEMO" & L1=="Arizona"))
+# library(communitySynchrony)
+# get_comm_synchrony(ts_data = subset(mlist, L2=="ENVDEMO" & L1=="Arizona"))
 ```

### 2015-05-18 R/vital_rate_regression_functions.R
changes INLA h-value to 1e-10; fixes #12

@@ -34,7 +34,7 @@ get_growth_params <- function(dataframe, crowd_mat, alpha){
                   family=c("gaussian"), verbose=FALSE,
                   control.predictor = list(link = 1),
                   control.compute=list(dic=T,mlik=T),
-                  control.inla = list(h = 1e-6))
+                  control.inla = list(h = 1e-10))
   
   # Fit variance
   x <- outINLA$summary.fitted.values$mean #fitted values from INLA
@@ -105,7 +105,7 @@ get_survival_params <- function(dataframe, crowd_mat, alpha){
                   family=c("binomial"), verbose=FALSE,
                   control.compute=list(dic=T,mlik=T),
                   control.predictor = list(link = 1),
-                  control.inla = list(h = 1e-6),
+                  control.inla = list(h = 1e-10),
                   Ntrials=rep(1,nrow(D)))
   
   #Collect parameters

### 2015-05-18 inst/examples/ipm_sims_example.Rmd
adds new function to calculate synchrony metrics from ipm output; fixes #4

@@
 library(ggplot2)
+library(synchrony)
 output_list <- readRDS("../../results/ipm_simulation_lists.RDS")
@@
 ```{r get_synchs}
-# library(communitySynchrony)
-# get_comm_synchrony(ts_data = subset(mlist, L2=="ENVDEMO" & L1=="Arizona"))
+library(communitySynchrony)
+colnames(mlist)[1:3] <- c("year", "species", "cover")
+ts_data <- subset(mlist, L2=="ENVDEMO" & L1=="Arizona")[1:3]
+test1 <- get_ipm_synchrony(ts_data = ts_data)
+str(test1)
 ```

### 2015-05-18 R/community_synchrony.R

@@ -97,3 +97,83 @@ get_comm_synchrony <- function(ts_data){
               growth_rates = obs_gr,
               percent_cover = ts_agg))
 }
+
+
+
+
+#' Calculate synchrony and stability from IPM time series
+#' 
+#' @param ts_data A melted dataframe of IPM cover values with three columns: year, species, cover.
+#' @return A list of data frames for synchrony and stability metrics for the community.
+get_ipm_synchrony <- function(ts_data){
+  species_list <- unique(ts_data$species)
+  num_spp <- length(species_list)
+  
+  ## Calculate stability
+  # population stability (mean/sd)
+  stability <- numeric(num_spp+1)
+  obs_vector <- numeric(num_spp)
+  for(i in 1:num_spp){ # loop through species for population stability
+    tmp <- subset(ts_data, species==species_list[i])
+    stability[i] <- mean(tmp$cover)/sd(tmp$cover)
+    obs_vector[i] <- nrow(tmp)
+  } # end species looping for population stability
+
+  # community stability (mean/sd of summed cover)
+  # calculate total cover from the species-level data for each quadrat
+  ts_sum <- ddply(ts_data, .(year), summarise,
+                  all_cover = sum(cover))
+  stability[num_spp+1] <- mean(ts_sum$all_cover)/sd(ts_sum$all_cover)
+  stability <- as.data.frame(stability)
+  stability$level <- c(species_list, "community")
+  stability$num_observations <- c(obs_vector, nrow(ts_sum))
+
+  ##  Calculate synchrony using Loreau and de Mazancourt metric
+  # requires the R package 'synchrony'
+  # Synchrony of abundance
+  ts_mat <- dcast(ts_data, formula = year~species, value.var="cover")
+  ts_mat <- ts_mat[which(complete.cases(ts_mat)==TRUE),]
+  synch_abundance <- community.sync(ts_mat[2:(num_spp+1)]) #Loreau and de Mazancourt 2008 (Am Nat)
+  
+  # caclulate observed growth rates
+  # create lagged data frame to only get observed yearly transitions
+  lag_df <- ts_mat
+  lag_df$lagyear <- lag_df$year+1
[…]

### 2015-05-19 inst/examples/ipm_sims_example.Rmd
adds figure for simulation scenarios and synchrony

@@
 
-Hmm... looks like something is wrong with a Kansas species. Also need to compare these results to the one's from Idaho-only script in Diversity-Stability repo.
-
 ## Run through synchrony library
@@
 ```
+
+## Plot results of simulation scenarios
+```{r results_plot}
+colnames(mlist)[1:3] <- c("year", "species", "cover")
+mlist <- subset(mlist, year>500)
+sites <- unique(mlist$L1)
+sims <- unique(mlist$L2)
+pgr_list <- list()
+for(dosite in sites){
+  tmp_data <- subset(mlist, L1==dosite)
+  if(dosite=="Idaho"){
+    tmp_data <- subset(tmp_data, species!="ARTR")
+  }
+  within_site_list <- list()
+  for(dosim in sims){
+    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
+    tmpsynch <- get_ipm_synchrony(tmpsim)
+    tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
+    within_site_list[[dosim]] <- tmp_pgr_synch
+  }
+  pgr_list[[dosite]] <- within_site_list
+}
+pgr_synch <- melt(pgr_list)
+
+ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
+  geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
+  facet_wrap("L1")+
+  scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
+  guides(fill=FALSE)+
+  scale_y_continuous(limits=c(0,1))+
+  xlab("Source of stochasticity")+
+  ylab("Species synchrony")
+
+```
+
+So all sites except Montana show an increase in species synchrony when environmental variation is included in the model.
+This is probably because they are all perennial grass species in semi-arid grasslands that respond similarly to environmental conditions.

### 2015-05-19 inst/examples/ipm_sims_example.Rmd
adds figure for looking a relation to niche difference

@@
 This is probably because they are all perennial grass species in semi-arid grasslands that respond similarly to environmental conditions.
+
+## Comparison with niche differecnes
+```{r niches}
+niche_diff <- data.frame(site=sites,
+                         ND = c(0.58, 0.59, 0.66, 0.69, 0.68))
+pgr_subs <- subset(pgr_synch, L2!="ENVDEMO")
+pgr_cast <- dcast(pgr_subs, L1~L2, value.var = "value")
+synch_diff <- ddply(pgr_cast, .(L1), summarise,
+                    value = ENV-DEMO)
+niche_synch <- merge(niche_diff, synch_diff, by.x="site", by.y="L1")
+ggplot(niche_synch, aes(x=ND, y=value))+
+  geom_point(size=5)
+```

### 2015-05-20 docs/prelim_analysis.Rmd
updates doc with new results and text

@@
 ---
+# Drivers of species synchrony in real communities
+*Andrew Tredennick, Peter Adler, Claire de Mazancourt, Michele Loreau, and Stephen Ellner*
 
-# Preliminary community synchrony analysis
+Quick summary
+-------------
+1.
+__Question__.
+How (a)synchronous are species population dynamics in natural communities, and what determines the degree of synchrony among species?
+Of particular interest to us, do species' responses to environmental fluction increase or decrease population synchrony?
 
-A couple comments from Claire:
-
-1. In the absence of environmental stochasticity, species fluctuations are not necessarily independent. 
-For example in a zero sum game with ecologically equivalent species (Hubbell's neutral model), species fluctuate randomly but they are not independent since the sum is constrained.
-If there are only two species their fluctuations are necessarily negatively correlated, because the increase in one species must be compensated by a decrease in the other.
-In a non-neutral model where species have stable equilibrium abundance there are also non independent fluctuations due to species interactions.
-In competitive communities fluctuations would most of the time have negative correlations overall, but sometimes they can be positive overall.
-
-2. I'm not entirely clear here: null with respect to what? Species interactions, demographic stochasticity and environmental stochasticity can all affect synchrony?
+2. __Data__. Five longterm demographic datasets from semi-arid North American grasslands. All species coexist stabley, yet are similar -- they are all perennial grasses.
 
 3.
-Could you show the absolute values, either with three bars per plot, or with a black line that is not at zero but at its unperturbed synchrony value?
-OK, I see the absolute values in table 1, the differences between sites are quite large compared to the effects of demographic and environmental stochasticities - but it should still be visible.
-Does this suggest that environmental and demographic stochasticities are minor explanatory variables for community synchrony?
-What is, then?
-Diversity?
+__Analysis__.
+(a) Calculate community synchrony from the observed time series using an established metric, and compare this to expectation from theory.
+(b) Use IPMs with environmental or demographic stochasticity turned on/off to quantify the relative contribution of each to community synchrony.
 
 4.
-In principle, could the signs of the effects of demographic stochasticity and environmental stochasticity be predicted?
-For demographic stochasticity, see comment by table 1.
-For environmental stochasticity: could you quantify the synchrony of species responses to the environment from the synchrony of the intercepts?
+__Initial conclusions__.
+Species' responses to environmental fluctuations tend to synchronize population dynamics: at four of the five sites removing environmental fluctuations decreased community synchrony.
+However, even with environmental fluctuations present, populations were not completely synchronized.
+This implies that niche differences among these species remain strong even as they respond similarly to environmental conditions.
 
-5. We could compute the synchrony expected under independent fluctuations: 
-$$\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.$$
-We could expect demographic stochasticity to tend towards this value: the sign of the effect of demographic stochasticity should be $\text{E}(\phi_{r}) - \phi_{r}$, where $\phi_{r}$ is the observed synchrony of per capita growth rates; removal should be the opposite - does that work?
+Introduction
+------------
+Asynchrony in population dynamics among co-occuring plant species stabilizes ecosystem-level properties like annual biomass production (Hector et al.
+2010, Loreau and de Mazancourt 2013, de Mazancourt et al.
+2013, Hautier et al.
+2014).
+Species-specific responses to environmental conditions, in particular, can stabilize ecosystem properties even as environmental forcings fluctuate through time.
+Such asynchrony of species dynamics in response to the environment has the potential to buffer ecosystems subjected to an increasingly variable climate.
+Thus, the extent to which natural communities exhibit asynchronous dynamics determines, in part, their ability to cope with interannual climate extremes that are projected to become more common.
 
-## Calculate expected and observed sychrony
-Under the assumption of independent fluctuations in species' per capit growth rates, we can calculate expected synchrony as:
+Theory suggests that community synchrony is driven by a combination of internal and external forces.
+Internal to the community, demographic stochasticity can induce asynchronous dynamics when species tend to fluctuate independently in a constant environment (Loreau and de Mazancourt 2008).
+However, species live in non-constant environments that act as strong external drivers of community dynamics.
+Environmental variability can induce asynchronous dynamics when species' responses to environmental conditions are temporally uncorrelated.
+When species respond similarly to environmental conditions, community dynamics become more synchronized (Gonzalez and Loreau 2009, Loreau and de Mazancourt 2008).
 
-$$\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }$$
+It is impossible to draw inference on the relative contribution of demographic and environmental stochasticity on synchrony based on empirical estimates alone, so our analysis involves two steps.
+First, we will use longterm datasets from five North American grasslands (Fig.
+1; Chu and Adler 2015) to calculate community synchrony.
+Second, we will fit multispecies population models to the data for dominant species to evaluate the relative contribution of species' environmental responses and demographic stochasticity to community synchrony.
+Sparse data on sub-dominant species constrained model fitting to dominant species that account for ##% - ##% of local plant abundance at their respective sites.
+Our focus is on temporal trends of per capita growth rates becuase these best reflect the short-term responses of populations to environmental conditions that fluctuate on annual time scales (Loreau and de Mazancourt 2008).
 
-where $\sigma_{i}$ is the standard deviation of species $i$'s per capita growth rate through time.
+What is expected by theory?[^devel]
+---------------------------
+[^devel]: Note that this is still in development...so some of it may be wrong.
+
+Loreau and de Mazancourt (2008) derived a statistic to measure community-wide synchrony of per capita population growth rates:
+
+\begin{equation}
+\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
+\end{equation}
+
+where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
+The statistic is bounded at 0 and 1, where 0 indicates perfect asynchrony and 1 indicates perfect synchrony.
+When applied to a neutral model where species have perfect niche overlap but have independent demographic and environmental fluctuations through time,  expected synchrony ($\phi_r$) is defined as:
+
+\begin{equation}
+\phi_r = \frac{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(SN)}{[(1/S)(\sum_{i}\sqrt{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(N)})]^2}
+\end{equation}
+
+where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
+*S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
+
+Loreau and de Mazancourt show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx 1$ meaning species are expected to fluctuate synchronously.
+At the other limit, where environmental forcing and community regulation are strong relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$ or:
+
+\begin{equation}
+\phi_{r} = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.
+\end{equation}
+
+Based on the above theory, we expect synchrony in a community where only demographic stochasticity is operating on top of independent fluctuations to be $1/S$.
+
+Data and Methods
+----------------
+## Study sites and species
+We use the exact same data as presented in Chu and Adler (2015). From their paper:
+
+>Many range experiment stations in the western U.S.
+began mapping permanent quadrats in the early 20th century, and continued annual censuses for decades.
+Here we focus on five long-term data sets (Fig.
+1), four of which have been distributed publicly (Adler et al.
+2007b, Zachmann et al.
+2010, Anderson et al.
+2011, 2012).
+At each study site, all individual plants within each 1-m2 quadrat were identified and mapped annually using a pantograph (Hill 1920; Fig.
+2).
+Mapped polygons represent basal cover for grasses and canopy cover for shrubs (Fig.
+2).
+
+>To fit our population models, we needed to study species common enough to provide a large sample size (i.e.
+the most common species at each site).
+In order to describe species interactions, we needed to select species that co-occurred together across many years.
+Thus, our analysis focused on the niche differences and average fitness differences of the most common, co-occurring species at each site, which implies that the average fitness differences among species could be relatively small.
+To select these species, we first identified species that occurred in at least 20% of years.
+Then we used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
+Based on the degree of aggregation of quadrats on the NMDS plot, we selected quadrats and species for our full analysis.
+We also confirmed these selections with local personnel from each study site.
+Below we further describe each study site and our criteria for quadrat and species selections.
+
+>_Sonoran desert, Arizona_
+
+>The Sonoran desert data set comes from 178 permanent quadrats located on semi-desert grasslands at the Santa Rita Experimental Range (31º50’ N, 110º53’ W; elevation 1150 meters), Arizona (Anderson et al.
+2012).
+Mean annual precipitation was 450 mm with the monthly variation from 5 mm (May) to 107 mm (July), and mean annual temperature was 16 °C with the monthly variation from 8 °C (January) to 26 °C (July) during the sampling period.
+These quadrats were mapped annually from 1915 to 1933 in most cases.
+The dominant flora varies across the broad range of soils and topography that occur within the semi-desert grasslands.
+For our present analysis, we chose 32 quadrats dominated by black grama Bouteloua eriopoda (BOER) and Bouteloua rothrockii (BORO).
+
+>_Sagebrush steppe, Idaho_
+
+>The sagebrush data set comes from the U.S.
+Sheep Experiment Station, currently a US Department of Agriculture Agricultural Research Service field station, located 9.6 km north of Dubois, Idaho (44.2°N, 112.1°W; elevation 1500 meters).
+26 quadrats were established between 1926 and 1932 (Zachmann et al.
+2010).
+During the period of sampling (1926 – 1957), mean annual precipitation was 270 mm with the monthly variation from 16 mm (March) to 42 mm (June), and mean annual temperature was 6 °C with the monthly variation from -8 °C (January) to 21 °C (July).
+The vegetation is dominated by the shrub, Artemisia tripartita (ARTR), and the C3 perennial bunchgrasses Pseudoroegneria spicata (PSSP), Hesperostipa comata (HECO), and Poa secunda (POSE).
+
+>_Southern mixed prairie, Kansas_
+
+>The southern mixed prairie data set represents the pioneering work of Albertson and colleagues (Albertson and Tomanek 1965), who established 51 permanent quadrats inside and outside livestock exclosures near Hays, Kansas (38.8°N, 99.3°W; elevation 650 meters).
+The mean annual precipitation was 580 mm with the monthly variation from 10 mm (January) to 103 mm (June), and mean annual temperature was 12 °C with the monthly variation from -2.0 °C (January) to 27 °C (July).
+These quadrats were mapped from 1932 to 1972 at the end of each growing season (Adler et al.
+2007b).
+Variation in soil depth and texture creates distinct plant communities.
+Many of the quadrats were located on shortgrass communities dominated by Bouteloua gracilis and Buchloë dactyloides.
+We had difficulty working with these species, which can be hard to distinguish in the field in the absence of inflorescences, and were often mapped inconsistently.
+Therefore, we focused on 7 quadrats dominated by Bouteloua curtipendula (BOCU), Bouteloua hirsuta (BOHI), and Schizachyrium scoparium (SCSC).
+
+>_Northern mixed prairie, Montana_
+
+>The northern mixed prairie data set comes from the Fort Keogh Livestock and Range Research Laboratory, another USDA Agricultural Research Service field station.
+The study site is located on alluvial plains near the Tongue River (46o19’N, 105o48’W; elevation 720 meters).
+Mean annual precipitation was 343 mm with the monthly variation from 7 mm (February) to 74 mm (June), and mean annual temperature was 8 °C with the monthly variation from -7 °C (January) to 25 °C (July).
+44 quadrats, distributed across six pastures assigned to one of three grazing intensities, were mapped annually (with some exceptions) from 1932 through 1945 (Anderson et al.
+2011).
+Our analysis includes 19 quadrats in which Bouteloua gracilis (BOGR), Hesperostipa comata (HECO), Pascopyrum smithii (PASM), and Poa secunda (POSE) were prevalent.
+
+>_Chihuahuan desert, New Mexico_
+
+>The last data set consists of 76 permanent quadrats located in Chihuahuan desert plant communities at the Jornada Experimental Range (32.62°N, 106.67°W; elevation 1260 meters), New Mexico.
+Mean annual precipitation was 264 mm with the monthly variation from 6 mm (April) to 48 mm (August), and mean annual temperature was 14 °C with the monthly variation from 4 °C (January) to 26 °C (July).
+Quadrats were established from 1915-1920 and mapping was conducted annually until the 1960’s (it continues on roughly 5 yr intervals to the present).
+Due to the changes in mapping frequency in the 1960’s, and the effects of a severe drought in the 1950’s that triggered a conversion of many quadrats from grassland to shrubland, we analyzed data for the period from 1915 to 1950.
+We selected 48 quadrats in which both Bouteloua eriopoda (BOER) and Sporobolus flexuosus (SPFL) were common (Anderson et al.
+in preparation).
+
+## Analytical approach
+While the observational data allows us to directly calculate species synchrony within the community, it is impossible to isolate the contributions of demographic and environmental stochasticity on species synchrony based on the data alone.
+So we now turn to a empirically-based modeling approach where we can turn demographic and environmental stochasticity on/off.
+To our knowledge, Integral Projection Models (IPMs) are the only tools available that allow us to isolate sources of stochasticity.
+For example, individual-based models require demographic stochasticity, making it impossible to isolate the effect of environmental stochasticity alone.
+
+We describe our approach in three steps.
+First, we describe the statistical models for three vitale rates: survival, growth, and recruitment.
+Second, we describe the IPM construction in detail.
+Third, we describe the IPM simulations we conducted to answer our questions.
+
+## Statistical models for survival, growth, and recruitment
+### Survival and growth
+We modeled survival probability and growht on individual genets as a function of genet size, the crowding experienced by the focal genet from both heterospecific and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
+Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure).
+
+We follow the approach of Chu and Adler (2015) to estimate crowding, assuming that the crowding experienced by a focal genet depends on distance to each neighbor genet and the neighbor's size, _u_:
+
+\begin{equation}
+w_{ijm,t} = \sum_k e^{-\delta_{jm}d_{ijkm,t}^{2}}u_{km,t}.
+\end{equation}
+
+In the above, $w_{ijm,t}$ is the crowding that genet _i_ of species _j_ in year _t_ experiences from neighbors of species _m_.
+The spatial scale over which species _m_ neighbors exert influence on any genet of species _j_ is determined by $\delta_{jm}$.
+The function is applied for all _k_ genets of species _m_ that neighbor the focal genet at time _t_, and $d_{ijkm,t}$ is the distance between genet _i_ in species _j_ and genet _k_ in species _m_.
+When $k=m$, the effect is intraspecific crowding.
+We use regression-specific (survival and growth) $\delta$ values estimated by Chu and Adler (2015).
+
+### Recruitment
+We model recruitment as a function of plot-level cover, temporal varation among years, and spatial variation among quadrat groups.
+
+## Integral projection models (IPM)
+### Model structure
+We built an environmentally and demographically stochastic integral projection model (IPM), where either environmental or demographic stochasticity can be turned off.
+Our IPM follows the specification of Chu and Adler (2015) where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_.
+Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$.
+So, the density function for any size _v_ at time $t+1$ is
+
+\begin{equation}
+n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
+\end{equation}
+
+where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal that describes all possible transitions from size $u$ to $v$ and $\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding experienced from all other species by a genet of size $u_j$ and species $j$.
+The integral is evaluated over all possible sizes between predefined lower (_L_) and upper (_U_) size limits that extend beyond the range of observed genet sizes.
+
+The population kernal is defined as the joint contributions of survival (_S_), growth (_G_), and recruitment (_R_):
+
+\begin{equation}
+k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
+\end{equation}
+
+which, said plainly, means we are calculating growth (_G_) for individuals that survive (_S_) from time _t_ to _t+1_ and adding in newly recruited (_R_) individuals of an average sized one-year-old genet for the focal species.
+Our stastical model for recruitment (_R_, described below) returns the number of new recruit produced per quadrat.
+Following previous work, we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM.
+
+### Model simulations
+Our interest is in the (de)synchronizing effects of either demographic or environmental stochasticity.
+To look at these effects, we run three types of model simulations: (1) demographic stochasticity only, (2) environmental stochasticity only, and (3) both.
+Environmental stochasticity is implemented by drawing random year effects for the vital rate regressions.
+We can model a constant environment by simply using vital rate regressions with mean coefficient values.
+To include demographic stochasticity we treat survival as a binomial process and recruitment as a poisson process in the IPM[^stoch].
+See Figure 2 for example model runs.
+
+[^stoch]: This is the crux of the matter here, we need to implement the Poisson approximation at the iteration matrix stage.
+
+Results
+-------
+## Observed and expected synchrony
+Our first task was to calculate the observed and expected community synchrony at each of our study sites.
+Observed synchrony ($\phi_r$) is calculate using equation 1 and expected synchrony ($\text{E}(\phi_r)$) is simply $1/S$ or equation 3.
+Expected synchrony under indpendent fluctuations is lower than the observed synchrony at every site (Table 1).
+This implies that demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$).
+Thus, some combination of density-dependence (species interactions) and environmental forcing are synchronizing population dynamics.
+To isolate those effects, we turn to our simulation results from the IPMs.
 
@@
 
-```{r get_synchs, echo=FALSE, include=TRUE}
-out <- get_comm_synchrony(ts_data = ks_data)
-obs_synch <- round(as.numeric(out$pgr_synchrony[1]), 2)
-exp_synch <- round(out$pgr_expected_synch_ind_flucts, 2)
-exp_demo_effect <- exp_synch - obs_synch
-print(out$pgr_synchrony)
-print(out$pgr_expected_synch_ind_flucts)
-```
-
-```{r sign_text, echo=FALSE}
-if(sign(exp_demo_effect)==-1){
-  sign_text <- "decrease"
-  remove_text <- "increase"
-}
-if(sign(exp_demo_effect)==1){
- sign_text <- "increase"
- remove_text <- "decrease" 
-}
-```
-
-So, $\text{E}(\phi_{r})$ = `r exp_synch` and $\phi_{r}$ = `r obs_synch`.
-Which means the expected effect of demographic stochasticity [$\text{E}(\sigma_{d}) = \text{E}(\phi_{r}) - \phi_{r}$] is `r exp_demo_effect`.
-So, based on the sign of $\text{E}(\sigma_{d})$ we expect demographic stochasticity to `r sign_text` synchrony among species' per capita growth rates.
-Thus, removing demographic stochasticity should `r remove_text` synchrony in our population model simulations.
-
-Here are the results for both Kansas and Idaho:
-
 ```{r get_all_synch, echo=FALSE, include=TRUE}
-out_ks <- out
+out_ks <- get_comm_synchrony(ts_data = ks_data)
 out_id <- get_comm_synchrony(ts_data = id_data)
@@
 
-```{r synchrony_table, echo=FALSE, include=TRUE, results='asis'}
+```{r synchrony_table, echo=FALSE, include=TRUE, results='asis', message=FALSE}
 get_table_metrics <- function(output){
@@
       include.rownames = F, 
-      sanitize.colnames.function = identity)
+      sanitize.colnames.function = identity,
+      comment=FALSE)
 ```
@@
 ```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site."}
-out_ks$percent_cover$site <- "Kansas"
-out_id$percent_cover$site <- "Idaho"
-out_mt$percent_cover$site <- "Montana"
-out_nm$percent_cover$site <- "New Mexico"
-out_az$percent_cover$site <- "Arizona"
-cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
-                    out_mt$percent_cover, out_nm$percent_cover, 
-                    out_az$percent_cover)
-
-out_ks$growth_rates$site <- "Kansas"
-out_id$growth_rates$site <- "Idaho"
-out_mt$growth_rates$site <- "Montana"
-out_nm$growth_rates$site <- "New Mexico"
-out_az$growth_rates$site <- "Arizona"
-pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
-                  out_mt$growth_rates, out_nm$growth_rates, 
-                  out_az$growth_rates)
-all_spp <- unique(pgr_data$species)
-color_match <- data.frame(species = all_spp,
-                          color = brewer.pal(11,"Paired"))
-pgr_data <- merge(pgr_data, color_match, by = "species")
-cover_data <- merge(cover_data, color_match, by="species")
-
-cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
-                geom_line()+
-                geom_point()+
-                facet_grid(site~., scales = "free_y")+
-                guides(color=FALSE)+
-                ylab("Mean Percent Cover")+
-                xlab("Year")+
-                scale_color_manual(values = brewer.pal(11,"Paired"))
-
-pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
-              geom_line()+
-              geom_point()+
-              facet_grid(site~., scales = "free_y")+
-              guides(color=FALSE)+
-              ylab("Per Capita Growth Rate")+
-              xlab("Year")+
-              scale_color_manual(values = brewer.pal(11,"Paired"))
-full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
+# out_ks$percent_cover$site <- "Kansas"
+# out_id$percent_cover$site <- "Idaho"
+# out_mt$percent_cover$site <- "Montana"
+# out_nm$percent_cover$site <- "New Mexico"
+# out_az$percent_cover$site <- "Arizona"
+# cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
+#                     out_mt$percent_cover, out_nm$percent_cover, 
+#                     out_az$percent_cover)
+# 
+# out_ks$growth_rates$site <- "Kansas"
+# out_id$growth_rates$site <- "Idaho"
+# out_mt$growth_rates$site <- "Montana"
+# out_nm$growth_rates$site <- "New Mexico"
+# out_az$growth_rates$site <- "Arizona"
+# pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
+#                   out_mt$growth_rates, out_nm$growth_rates, 
+#                   out_az$growth_rates)
+# all_spp <- unique(pgr_data$species)
+# color_match <- data.frame(species = all_spp,
+#                           color = brewer.pal(11,"Paired"))
+# pgr_data <- merge(pgr_data, color_match, by = "species")
+# cover_data <- merge(cover_data, color_match, by="species")
+# 
+# cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
+#                 geom_line()+
+#                 geom_point()+
+#                 facet_grid(site~., scales = "free_y")+
+#                 guides(color=FALSE)+
+#                 ylab("Mean Percent Cover")+
+#                 xlab("Year")+
+#                 scale_color_manual(values = brewer.pal(11,"Paired"))
+# 
+# pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
+#               geom_line()+
+#               geom_point()+
+#               facet_grid(site~., scales = "free_y")+
+#               guides(color=FALSE)+
+#               ylab("Per Capita Growth Rate")+
+#               xlab("Year")+
+#               scale_color_manual(values = brewer.pal(11,"Paired"))
+# full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
 ```
 
-# Methods
-## Study sites and species
-We use the exact same data as presented in Chu and Adler (2015). From their paper:
+## Effect of removing demographic and environmental stochasticity
+At four of our five study areas, the presence of environmental stochasticity via random year effects on vital rate regressions increased community synchrony of per capita population growth rates (Figure 1).
+Only at the Montana site (Northern Mixed Prairie) did incorporating species' responses to environmental conditions decrease community synchrony.
+However, even with similar responses to environmental conditions, community synchrony did not reach 1 at any of the sites.
+Thus, there remains sufficiently strong temporal niche differences to drive the community away from perfect sychrony.
 
->Many range experiment stations in the western U.S.
-began mapping permanent quadrats in the early 20th century, and continued annual censuses for decades.
-Here we focus on five long-term data sets (Fig.
-1), four of which have been distributed publicly (Adler et al.
-2007b, Zachmann et al.
-2010, Anderson et al.
-2011, 2012).
-At each study site, all individual plants within each 1-m2 quadrat were identified and mapped annually using a pantograph (Hill 1920; Fig.
-2).
-Mapped polygons represent basal cover for grasses and canopy cover for shrubs (Fig.
-2).
+*NOTE: we have not dealt with species interactions yet...*
 
->To fit our population models, we needed to study species common enough to provide a large sample size (i.e.
-the most common species at each site).
-In order to describe species interactions, we needed to select species that co-occurred together across many years.
-Thus, our analysis focused on the niche differences and average fitness differences of the most common, co-occurring species at each site, which implies that the average fitness differences among species could be relatively small.
-To select these species, we first identified species that occurred in at least 20% of years.
-Then we used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
-Based on the degree of aggregation of quadrats on the NMDS plot, we selected quadrats and species for our full analysis.
-We also confirmed these selections with local personnel from each study site.
-Below we further describe each study site and our criteria for quadrat and species selections.
+```{r read_data, echo=FALSE}
+output_list <- readRDS("../results/ipm_simulation_lists.RDS")
+mlist <- melt(output_list)
+```
 
->_Sonoran desert, Arizona_
+```{r get_synchs, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site.
+DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6}
+colnames(mlist)[1:3] <- c("year", "species", "cover")
+mlist <- subset(mlist, year>500)
+sites <- unique(mlist$L1)
+sims <- unique(mlist$L2)
+pgr_list <- list()
+for(dosite in sites){
+  tmp_data <- subset(mlist, L1==dosite)
+  if(dosite=="Idaho"){
+    tmp_data <- subset(tmp_data, species!="ARTR")
+  }
+  within_site_list <- list()
+  for(dosim in sims){
+    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
+    tmpsynch <- get_ipm_synchrony(tmpsim)
+    tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
+    within_site_list[[dosim]] <- tmp_pgr_synch
+  }
+  pgr_list[[dosite]] <- within_site_list
+}
+pgr_synch <- melt(pgr_list)
 
->The Sonoran desert data set comes from 178 permanent quadrats located on semi-desert grasslands at the Santa Rita Experimental Range (31º50’ N, 110º53’ W; elevation 1150 meters), Arizona (Anderson et al.
-2012).
-Mean annual precipitation was 450 mm with the monthly variation from 5 mm (May) to 107 mm (July), and mean annual temperature was 16 °C with the monthly variation from 8 °C (January) to 26 °C (July) during the sampling period.
-These quadrats were mapped annually from 1915 to 1933 in most cases.
-The dominant flora varies across the broad range of soils and topography that occur within the semi-desert grasslands.
-For our present analysis, we chose 32 quadrats dominated by black grama Bouteloua eriopoda (BOER) and Bouteloua rothrockii (BORO).
+ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
+  geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
+  facet_wrap("L1", scales="free_x")+
+  scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
+  guides(fill=FALSE)+
+  scale_y_continuous(limits=c(0,1))+
+  xlab("Source of stochasticity")+
+  ylab("Species synchrony")
 
->_Sagebrush steppe, Idaho_
+```
 
->The sagebrush data set comes from the U.S.
-Sheep Experiment Station, currently a US Department of Agriculture Agricultural Research Service field station, located 9.6 km north of Dubois, Idaho (44.2°N, 112.1°W; elevation 1500 meters).
-26 quadrats were established between 1926 and 1932 (Zachmann et al.
-2010).
-During the period of sampling (1926 – 1957), mean annual precipitation was 270 mm with the monthly variation from 16 mm (March) to 42 mm (June), and mean annual temperature was 6 °C with the monthly variation from -8 °C (January) to 21 °C (July).
-The vegetation is dominated by the shrub, Artemisia tripartita (ARTR), and the C3 perennial bunchgrasses Pseudoroegneria spicata (PSSP), Hesperostipa comata (HECO), and Poa secunda (POSE).
+```{r niches, echo=FALSE}
+# niche_diff <- data.frame(site=sites,
+#                          ND = c(0.58, 0.59, 0.66, 0.69, 0.68))
+# pgr_subs <- subset(pgr_synch, L2!="ENVDEMO")
+# pgr_cast <- dcast(pgr_subs, L1~L2, value.var = "value")
+# synch_diff <- ddply(pgr_cast, .(L1), summarise,
+#                     value = ENV-DEMO)
+# niche_synch <- merge(niche_diff, synch_diff, by.x="site", by.y="L1")
+# ggplot(niche_synch, aes(x=ND, y=value))+
+#   geom_point(size=5)+
+#   ylab("Community synchrony")
+```
 
->_Southern mixed prairie, Kansas_
+Theory predicts that demographic stochasticty is the only force that, in isolation, can desynchronize population dynamics.
+However, theory also predicts that demographic stochasticity is likely unable to counteract the synchronizing effect of environmental forces.
+Our results support this prediction in the cases where environmental stochasticity is a synchronizing force (compare ENV and ENVDEMO bars in Figure 1).
 
->The southern mixed prairie data set represents the pioneering work of Albertson and colleagues (Albertson and Tomanek 1965), who established 51 permanent quadrats inside and outside livestock exclosures near Hays, Kansas (38.8°N, 99.3°W; elevation 650 meters).
-The mean annual precipitation was 580 mm with the monthly variation from 10 mm (January) to 103 mm (June), and mean annual temperature was 12 °C with the monthly variation from -2.0 °C (January) to 27 °C (July).
-These quadrats were mapped from 1932 to 1972 at the end of each growing season (Adler et al.
-2007b).
-Variation in soil depth and texture creates distinct plant communities.
-Many of the quadrats were located on shortgrass communities dominated by Bouteloua gracilis and Buchloë dactyloides.
-We had difficulty working with these species, which can be hard to distinguish in the field in the absence of inflorescences, and were often mapped inconsistently.
-Therefore, we focused on 7 quadrats dominated by Bouteloua curtipendula (BOCU), Bouteloua hirsuta (BOHI), and Schizachyrium scoparium (SCSC).
+\newpage
 
->_Northern mixed prairie, Montana_
-
->The northern mixed prairie data set comes from the Fort Keogh Livestock and Range Research Laboratory, another USDA Agricultural Research Service field station.
-The study site is located on alluvial plains near the Tongue River (46o19’N, 105o48’W; elevation 720 meters).
-Mean annual precipitation was 343 mm with the monthly variation from 7 mm (February) to 74 mm (June), and mean annual temperature was 8 °C with the monthly variation from -7 °C (January) to 25 °C (July).
-44 quadrats, distributed across six pastures assigned to one of three grazing intensities, were mapped annually (with some exceptions) from 1932 through 1945 (Anderson et al.
-2011).
-Our analysis includes 19 quadrats in which Bouteloua gracilis (BOGR), Hesperostipa comata (HECO), Pascopyrum smithii (PASM), and Poa secunda (POSE) were prevalent.
-
->_Chihuahuan desert, New Mexico_
-
->The last data set consists of 76 permanent quadrats located in Chihuahuan desert plant communities at the Jornada Experimental Range (32.62°N, 106.67°W; elevation 1260 meters), New Mexico.
-Mean annual precipitation was 264 mm with the monthly variation from 6 mm (April) to 48 mm (August), and mean annual temperature was 14 °C with the monthly variation from 4 °C (January) to 26 °C (July).
-Quadrats were established from 1915-1920 and mapping was conducted annually until the 1960’s (it continues on roughly 5 yr intervals to the present).
-Due to the changes in mapping frequency in the 1960’s, and the effects of a severe drought in the 1950’s that triggered a conversion of many quadrats from grassland to shrubland, we analyzed data for the period from 1915 to 1950.
-We selected 48 quadrats in which both Bouteloua eriopoda (BOER) and Sporobolus flexuosus (SPFL) were common (Anderson et al.
-in preparation).
-
-## Analytical approach
-While the observational data allows to directly calculate species synchrony within the community, it is impossible to isolate the contributions of demographic and environmental stochasticity on species synchrony based on the data alone.
-So we now turn to a empirically-based modeling approach where we can turn demographic and environmental stochasticity on/off.
-To our knowledge, Integral Projection Models (IPMs) are the only tools available that allow us to isolate sources of stochasticity.
-For example, individual-based models require demographic stochasticity, making it impossible to isolate the effect of environmental stochasticity alone.
-
-We describe our approach in three steps.
-First, we describe the statistical models for three vitale rates: survival, growth, and recruitment.
-Second, we describe the IPM construction in detail.
-Third, we describe the IPM simulations we conducted to answer our questions.
-
-## Statistical models for survival, growth, and recruitment
-### Survival and growth
-We modeled survival probability and growht on individual genets as a function of genet size, the crowding experienced by the focal genet from both hetero and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
-Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure).
-
-We follow the approach of Chu and Adler (2015) to estimate crowding, assuming that the crowding experienced by a focal genet depends on distance to each neighbor genet and the neighbor's size, _u_:
-
-$$
-w_{ijm,t} = \sum_k e^{-\delta_{jm}d_{ijkm,t}^{2}}u_{km,t}.
-$$
-
-In the above, $w_{ijm,t}$ is the crowding that genet _i_ of species _j_ in year _t_ experiences from neighbors of species _m_.
-The spatial scale over which species _m_ neighbors exert influence on any genet of species _j_ is determined by $\delta_{jm}$.
-The function is applied for all _k_ genets of species _m_ that neighbor the focal genet at time _t_, and $d_{ijkm,t}$ is the distance between genet _i_ in species _j_ and genet _k_ in species _m_.
-When $k=m$, the effect is intraspecific crowding.
-We use regression-specific (survival and growth) $\delta$ values estimated by Chu and Adler (2015).
-
-## Integral projection models (IPM)
-We built an environmentally and demographically stochastic integral projection model (IPM), where either environmental or demographic stochasticity can be turned off.
-Our IPM follows the specification of Chu and Adler (2015) where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_.
-Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$.
-So, the density function for any size _v_ at time $t+1$ is
-
-$$
-n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
-$$
-
-where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal that describes all possible transitions from size $u$ to $v$ and $\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding experienced from all other species by a genet of size $u_j$ and species $j$.
-The integral is evaluated over all possible sizes between predefined lower (_L_) and upper (_U_) size limits that extend beyond the range of observed genet sizes.
-
-The population kernal is defined as the joint contributions of survival (_S_), growth (_G_), and recruitment (_R_):
-
-$$
-k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
-$$
-
-which, said plainly, means we are calculating growth (_G_) for individuals that survive (_S_) from time _t_ to _t+1_ and adding in newly recruited (_R_) individuals of an average sized one-year-old genet for the focal species.
-Our stastical model for recruitment (_R_, described below) returns the number of new recruit produced per quadrat.
-Following previous work, we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM.
+Additional Figures
+------------------
+```{r time_series, echo=FALSE, fig.cap="Example time series from the simulations."}
+ggplot(mlist, aes(x=year, y=(cover*100), color=species))+
+  geom_line()+
+  facet_grid(L1~L2, scale="free")
+```

### 2015-05-26 inst/examples/ipm_demostoch_notes.Rmd
proposed demographic stochasticity estimation code

---
output: pdf_document
---

Example code for estimating covariance between size-classes. Based on equation 10.2.5 in Steve's book:

$$
Cov[n_{i}(t+1)n_{j}(t+1)] = -h^{2} \sum_{k} n_{k}(t)P(z_{i},z{k})P(z_{j},z{k})
$$

where, in our case (*I think*), $P()$ is the survival$\times$growth kernal.

Here is the proposed code:

```{r covcode, eval=FALSE}
####
##  Define a function that returns P(z_{i},z{k})*P(z_{j},z{k}) for all i,j combos
####
# X is one column (k) of the survivalXgrowth kernal
# pop_vector is the population vector whose elements are actual numbers of individuals
get_cov <- function(X, pop_vector){
  pairs <- expand.grid(X, X)
  pairs$tag <- pairs[,1] - pairs[,2]
  pairs$multi <- pairs[,1]*pairs[,2]*pop_vector
  return(pairs$multi)
}

####
##  Apply the function over each column
####
test <- apply(K.matrix, MARGIN = 2, FUN = "get_cov", 
              pop_vector=(nt[[doSpp]]*h[doSpp]))

####
##  Create a tag for identifying the focal i
####
mat_dim <- sqrt(dim(test)[1])
test <- as.data.frame(test)
test$tag <- rep(c(1:mat_dim), each=mat_dim)

####
##  Loop over each i and calculate the covariance with each j (vectorized)
####
cov_str <- matrix(ncol=mat_dim, nrow=mat_dim)
for(do_i in 1:mat_dim){
  tmp <- subset(test, tag==do_i) #subset out the focal i
  rmtmp <- which(colnames(tmp)=="tag") #get rid of id column
  # Sum over k columns
  cov_str[do_i,] <- (-h[doSpp]^2) * apply(tmp[,-rmtmp], MARGIN = 2, FUN = "sum")
}

####
##  Make the matrix positive definite by setting i=j diagonals to 1
####
diag(cov_str) <- 1

####
##  Feed into Poisson random vector generator...
####
```

This should result in a matrix the same dimensions as $P()$ where each $i,j$ column is the appropriate covariance.

### 2015-06-01 inst/examples/ipm_demostoch_notes.Rmd
adds poisson generator code

@@
 This should result in a matrix the same dimensions as $P()$ where each $i,j$ column is the appropriate covariance.
+
+### Generate correlated Poisson vector
+From Yahav and Shmueli (2007; http://arxiv.org/pdf/0710.5670.pdf), here is a proposal for generating the poisson random vector with the covariance matrix.
+
+```{r poisson, echo=TRUE, eval=TRUE}
+####
+##  Define function
+####
+# Generate a p-dimensional Poisson
+# p = the dimension of the distribution
+# samples = the number of observations
+# R = correlation matrix p X p
+# lambda = rate vector p X 1
+library(MASS)
+GenerateMultivariatePoisson<-function(pD, samples, R, lambda){
+  normal_mu=rep(0, pD)
+  normal = mvrnorm(samples, normal_mu, R)
+  pois = normal
+  p=pnorm(normal)
+  for (s in 1:pD){pois[s]=qpois(p[s], lambda[s])}
+  return(pois)
+}
+
+####
+##  Example
+####
+pD = 2
+samples = 1
+R = matrix(data = c(1,0.1,
+                    0.8,1), ncol=2, nrow=2)
+lambda = c(0.8, 0.3)
+GenerateMultivariatePoisson(pD = pD, samples = samples, R = R, lambda = lambda)
+
+```

### 2015-06-02 inst/examples/ipm_demostoch_notes.Rmd
updates example doc with iteration matrix covariance and poisson estimation

@@
 GenerateMultivariatePoisson(pD = pD, samples = samples, R = R, lambda = lambda)
+```
 
+### Example
+```{r example, echo=TRUE, eval=FALSE}
+####
+##  Define multivariate Poisson functions
+####
+# Function to get all possible pairs for a column of the iteration matrix
+get_pairs <- function(X, pop_vector){
+  pairs <- expand.grid(X, X)
+  pairs$multi <- pairs[,1]*pairs[,2]*pop_vector
+  return(pairs$multi)
+}
+
+# Apply the 'get_pairs' function across all columns of the iteration matrix
+# and calculate equation 10.2.5
+get_cov <- function(K){
+  test <- apply(K, MARGIN = 2, FUN = "get_pairs", 
+                pop_vector=(nt[[doSpp]]))
+  mat_dim <- sqrt(dim(test)[1])
+  test <- as.data.frame(test)
+  test$tag <- rep(c(1:mat_dim), each=mat_dim)
+  cov_str <- matrix(ncol=mat_dim, nrow=mat_dim)
+  for(do_i in 1:mat_dim){
+    tmp <- subset(test, tag==do_i) #subset out the focal i
+    rmtmp <- which(colnames(tmp)=="tag") #get rid of id column
+    # Sum over k columns
+    cov_str[do_i,] <- (-h[doSpp]^2) * apply(tmp[,-rmtmp], MARGIN = 2, FUN = "sum")
+  }
+  diag(cov_str) <- 1
+  return(cov_str)
+}
+
+# Get a correlated Poisson vector based on the covariance and population vector
+GenerateMultivariatePoisson<-function(pD, samples, R, lambda){
+  normal_mu=rep(0, pD)
+  normal = mvrnorm(samples, normal_mu, R)
+  pois = normal
+  p=pnorm(normal)
+  for (s in 1:pD){pois[s]=qpois(p[s], lambda[s])}
+  return(pois)
+}
+
+####
+##  Run the model... (there would be more code here, obviously)
+####
+# IPM code above here...
+# Leading to looping over species...
+# And projecting the population from nt -> new.nt
+for(doSpp in 1:Nspp){  
+    if(cover[doSpp]>0){    
+      # make kernels and project
+      K.matrix=make.K.matrix(v[[doSpp]],WmatG[[doSpp]],WmatS[[doSpp]],
+                             Rpars,rpa,Gpars,Spars,doYear,doSpp)  
+      covmat <- get_cov(K=K.matrix)
+      new.nt[[doSpp]] <- GenerateMultivariatePoisson(pD = length(nt[[doSpp]]),
+                                  samples = 1,
+                                  R = covmat,
+                                  lambda = K.matrix%*%nt[[doSpp]])
+      sizeSave[[doSpp]][,i]=new.nt[[doSpp]]/sum(new.nt[[doSpp]])  
+    }    
+  } # next species
 ```
+
+And this seems to work. Not only do we not get extinctions, but synchrony of population growth ratee is 0.33, which is exactly as expected by theory: 1/S = 1/3 = 0.33.
+
+![Example IPM simulation for Idaho (ARTR set to 0) with demographic stochasticity only.](idaho_demographic_stoch.pdf)

### 2015-06-11 inst/examples/ipm_demostoch_notes.Rmd
adds example in IPM context to demograhpic test

--- 2015-05-26 inst/examples/ipm_demostoch_notes.Rmd
@@
 ---
-output: pdf_document
+output: 
+  pdf_document:
+    fig_caption: yes
 ---
+
+# Demographic stochasticity in IPMs
 
@@
 
-where, in our case (*I think*), $P()$ is the survival$\times$growth kernal.
+where, $P()$ is the survival$\times$growth kernal.
+Note that in the function below, the line `pairs$multi <- pairs[,1]*pairs[,2]*pop_vector` corresponds to $n_{k}(t)P(z_{i},z{k})P(z_{j},z{k})$ for a single column *k*.
 
-Here is the proposed code:
+The function for generating the correlated poisson vector is from http://arxiv.org/pdf/0710.5670.pdf.
+
+### Proposed functions
+Here are the proposed functions to calculate the covariance matrix for *P()* and for generating the correlated poisson vector based on that covariance structure:
 
@@
 # pop_vector is the population vector whose elements are actual numbers of individuals
-get_cov <- function(X, pop_vector){
+
+##  This function expands the P matrix to get all pairs of i and j
+get_pairs <- function(X, pop_vector){
   pairs <- expand.grid(X, X)
-  pairs$tag <- pairs[,1] - pairs[,2]
   pairs$multi <- pairs[,1]*pairs[,2]*pop_vector
@@
 
-####
-##  Apply the function over each column
-####
-test <- apply(K.matrix, MARGIN = 2, FUN = "get_cov", 
-              pop_vector=(nt[[doSpp]]*h[doSpp]))
-
-####
-##  Create a tag for identifying the focal i
-####
+##  This function takes the P.matrix, runs it through `get_pairs`
+##    and then calculates the covariance structure.
+get_cov <- function(MAT){
+  test <- apply(MAT, MARGIN = 2, FUN = "get_pairs", 
+                pop_vector=(nt[[doSpp]]))
   mat_dim <- sqrt(dim(test)[1])
@@
   test$tag <- rep(c(1:mat_dim), each=mat_dim)
-
-####
-##  Loop over each i and calculate the covariance with each j (vectorized)
-####
   cov_str <- matrix(ncol=mat_dim, nrow=mat_dim)
@@
   }
+  diag(cov_str) <- 1 #set diagonals to perfect correlation
+  return(cov_str)
+}
 
 ####
-##  Make the matrix positive definite by setting i=j diagonals to 1
+##  Function for generating correlated poisson with covariance structure
 ####
-diag(cov_str) <- 1
+# p = the dimension of the distribution (length of pop vector)
+# samples = the number of observations
+# R = correlation matrix (p X p)
+# lambda = rate vector (p)
+GenerateMultivariatePoisson<-function(pD, samples, R, lambda){
+  normal_mu=rep(0, pD)
+  normal = mvrnorm(samples, normal_mu, R)
+  pois = normal
+  p=pnorm(normal)
+  for (s in 1:pD){pois[s]=qpois(p[s], lambda[s])}
+  return(pois)
+}
+```
+
+### Example in IPM context
+Here is a little example as it would occur in the IPM.
+In the code below, I am assuming a bit of familiarity, but note that `v` and `u` are both the population vector, `rpa` is the number of recruites per area, and `muWG` and `muWS` are the estimated crowding effects.
+```{r ipm_ex, echo=TRUE, eval=FALSE}
+####
+##  Iteration matrix functions
+####
+# The iteration matrix is decomposed into the survivalXgrowth kernel (P)
+#   and the recruitment kernel (R).
+
+# Get recruitment values from regression
+make.R.values=function(v,u,Rpars,rpa,doYear,doSpp){
+  f(v,u,Rpars,rpa,doSpp)
+}
+
+# Get survival and growth values from regression
+make.P.values <- function(v,u,muWG,muWS, Gpars,Spars,doYear,doSpp){ 
+  S(u,muWS,Spars,doYear,doSpp)*G(v,u,muWG,Gpars,doYear,doSpp) 
+}
+
+# Turn the values into iteration matrix
+make.P.matrix <- function(v,muWG,muWS,Gpars,Spars,doYear,doSpp) {
+  muWG=expandW(v,v,muWG)
+  muWS=expandW(v,v,muWS)
+  P.matrix=outer(v,v,make.P.values,muWG,muWS,Gpars,Spars,doYear,doSpp)
+  return(h[doSpp]*P.matrix)
+} 
+make.R.matrix=function(v,Rpars,rpa,doYear,doSpp) {
+  R.matrix=outer(v,v,make.R.values,Rpars,rpa,doYear,doSpp)
+  return(h[doSpp]*R.matrix)
+}
 
 ####
-##  Feed into Poisson random vector generator...
+##  Pretend we're inside the simulation loop...
 ####
+P.matrix <- make.P.matrix(v[[doSpp]],WmatG[[doSpp]],WmatS[[doSpp]],Gpars,Spars,doYear,doSpp)  
+R.matrix <- make.R.matrix(v[[doSpp]],Rpars,rpa,doYear,doSpp)
+
+# Calculate covariance
+covmat <- get_cov(K=P.matrix)
+# Generate the poisson vector for P contribution the population
+pCont <- GenerateMultivariatePoisson(pD = length(nt[[doSpp]]),
+                                     samples = 1,
+                                     R = covmat,
+                                     lambda = P.matrix%*%nt[[doSpp]])
+
+# Run recruitment matrix through regular poisson for recruitment contribution
+rCont <- rpois(length(nt[[doSpp]]),R.matrix%*%nt[[doSpp]])
+
+# Sum the survivalXgrowth and recruitment contributions for
+#   updated population vector
+next.nt <- pCont+rCont
 ```
 
-This should result in a matrix the same dimensions as $P()$ where each $i,j$ column is the appropriate covariance.
+And...it's slow, but appears to work (figure on next page):
+
+![Example time series with three species fromt the Idaho dataset where only demographic stochasticity is acting (i.e., no temporal variation due to random year effects).](demo_plot.png)

### 2015-07-12 docs/manuscript/manuscript.Rmd
ms intro to organize thoughts

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
##bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
##csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in
---

Drivers of species synchrony in real communities
================================================
### Andrew T. Tredennick, Claire de Mazancourt, Michel Loreau, Stephen P. Ellner, and Peter B. Adler

*Andrew T. Tredennick, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah*

*Claire de Mazancourt, Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, CNRS, 09200 Moulis, France*

*Michel Loreau, Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, CNRS, 09200 Moulis, France*

*Stephen P. Ellber, Department of Ecology and Evolutionary Biology, Cornell University, Ithaca, New York*

*Peter B. Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah*

```{r libraris, include=FALSE}
library(ggplot2)
library(plyr)
library(gridExtra)
library(reshape2)
library(xtable)
library(ggmcmc)
```

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))

```

Abstract
--------
Compensatory dynamics among species is the leading hypothesis for why species richness tends to confer temporal stability via population asynchrony. Theory predicts three processes that can lead to compensatory dynamics: interspecific interactions, species specific response to environmental conditions, and demographic stochasticity. These processes can interact, making it difficult to isolate their relative contributions. Analyses of biodiversity-ecosystem function experiments have shed some light on this subject, but no attempts have been made to examine these processes in natural communities. Here, we provide the first analysis of the drivers of compensatory dynamics in natural systems. We use empirically driven integral projection models to simulate communities of competing perennial grasses from five semi-arid grassland in the western United States. Using the models, we can turn off/on species competition, environmental fluctuations, and demographic stochasticity. We ran a suite of simulations to estimate the contribution of each process to species synchrony. In these natural communities, we find...

Introduction
------------
More biologically diverse ecosystems tend to be more stable through time. There is broad experimental evidence for this proposition and theory -- whether mechanistic, phenomenonological, or statistical -- predicts this trend as well. What remains uncertain are the ultimate mechanisms the lead more diverse ecosystems or communities to be more stable through time. A leading hypothesis is that compensatory dynamics among species stabilizes community biomass through asynchronous population fluctuations. However, compensatory dynamics, if observed, are the result of several interacting processes that are difficult to dissect.

Compensatory dynamics occur when increases in the abundance or biomass of one species are counteracted by decreases in another, or *vice versa*. This can stabilize community biomass by maintaining a relatively unvarying sum of species abundance or biomass through time. Early theory based on species competition for shared resources predicts compensatory dynamics as a result of interspecific interactions: if two species are competing for the same limited and limiting resource, an increase in abundance of one species must be met by a decrease in abundance of the competing species. The equality of amplitude in fluctuations depends on how limited each species is to the shared resource.

Another, non-mutually exclusive, school of thought is that compensatory dynamics result from species specific responses to environmental conditions that vary through time. In this case, species need not directly interact to have asynchronous fluctuations through time. For example, different species may have non-identical growth responses to temperature that result in asynchronous fluctuations as the mean annual temperature varies through time.

Lastly, demographic stochasticity alone can lead to independent species fluctuations in neutral communities. Clearly all three mechanisms can be at play simultaneously, and the challenge is to isolate their individual and interacting effects. Significant progress on this front has been made in recent years thanks to sophisticated analyses of biodiversity-ecosystem function experiments. But these analyses suffer from several limitations. First, it is unclear how readily results from BEF experiments apply to natural communities, where the breadth of diversity-stability trends is much greater. Second, for terrestrial plants, very few experiments have run longer than ten years. This severely limits our ability to detect non-transient processes and long-term process like demographic stochasticity. Third, it is impossible to directly estimate the strength interspecific interactions. Any measure of interspecific interactions from BEF experiments are coarse measures of the difference between species biomass in monoculture versus polyculture that conflate inter and intraspecific competition.

Here, we provide a new analysis of demographic data from five grasslands to overcome some of the aforementioned limitations, and to test theory on species synchrony in natural communtities. We use an empirically based modeling approach to quantify the effect of species specific responses to the environment, interspecific interactions, and demographic stochasticity on compensatory dynamics.

Methods
-------
### Simulations
1. Demographic stochasticity, environmental fluctuations, competition
2. Demographic stochasticity, environmental fluctuations, no competition
3. Demographic stochasticity, no environmental fluctuations, competition
4. No demographic stochasticity, environmental fluctuations, competition
5. Demographic stochasticity, no environmental fluctuations, no competition
6. No demographic stochasticity, environmental fluctuations, no competition
7. No demographic stochasticity, no environmental fluctuations, competition

Results
-------
Fig. 1. multi-panel regression plots: x-axis = synchrony of #1, y-axis = synchrony of each scenario. Only five points, since five sites, but still enough for a correlation.

Fig. 2. CV of cover versus species synchrony in monoculture (no competition)

Tables
------
```{r table1, echo=FALSE, results='asis', message=FALSE, comment=NA}
site_data <- data.frame(Site = c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico"),
                        Biome = c("Sonoran desert", "Sagebrush steppe", 
                                  "Southern mixed prairie", "Northern mixed prairie",
                                  "Chihuahuan desert"),
                        Species_Rich = as.character(c(2,3,3,2,4)),
                        Years = as.character(c(19,32,41,23,36)))
colnames(site_data) <- c("Site", "Biome", "No. of Species", "No. of Time Points")
print(xtable(site_data, caption = "Site descriptions."),type="latex", comment=FALSE, include.rownames=FALSE, caption.placement="top")
```

### 2015-07-12 R/run_ipm.R
adds script to run ipm in monoculture

@@ -16,7 +16,8 @@
 
 run_ipm <- function(A=10000, tlimit=2500, burn_in=500, spp_list,
                     Nyrs, constant=FALSE, iter_matrix_dims, max_size,
-                    Rpars, Spars, Gpars, demographic_stochasticity=FALSE){
+                    Rpars, Spars, Gpars, demographic_stochasticity=FALSE,
+                    spp_interact=TRUE){
   NoOverlap_Inter <- FALSE
   library(IPMdoit)
   library(boot)
@@ -31,12 +32,31 @@ run_ipm <- function(A=10000, tlimit=2500, burn_in=500, spp_list,
   iter_matrix_dims=iter_matrix_dims; max_size=max_size
   Rpars=Rpars; Spars=Spars; Gpars=Gpars
   demographic_stochasticity=do_demo_stoch
+  spp_interact=spp_interact
 
   # Turn off random year effects if constant==TRUE
   if(constant==TRUE){
     Rpars$intcpt.yr=matrix(Rpars$intcpt.mu,Nyrs,Nspp,byrow=T)
     Gpars$intcpt.yr[]=0;Gpars$slope.yr[]=0
     Spars$intcpt.yr[]=0;Spars$slope.yr[]=0    
+  }
+  
+  # Turn off competition if spp_interact==FALSE
+  if(spp_interact==FALSE){
+    rnbtmp <- Rpars$dd
+    rnbtmp[] <- 0
+    diag(rnbtmp) <- diag(Rpars$dd)
+    Rpars$dd <- rnbtmp
+    
+    gnbtmp <- Gpars$nb
+    gnbtmp[] <- 0
+    diag(gnbtmp) <- diag(Gpars$nb)
+    Gpars$nb <- gnbtmp
+    
+    snbtmp <- Spars$nb
+    snbtmp[] <- 0
+    diag(snbtmp) <- diag(Spars$nb)
+    Spars$nb <- snbtmp    
   }
   
   # Build initial vectors and matrices

### 2015-07-12 runscripts/ipm_monoculture.R

@@ -1,0 +1,107 @@
+##  Script to run IPM for monoculture synchrony
+
+rm(list=ls(all=TRUE))
+library(communitySynchrony)
+
+Gpars_all <- readRDS("../results/growth_params_list.RDS")
+Spars_all <- readRDS("../results/surv_params_list.RDS")
+Rpars_all <- readRDS("../results/recruit_parameters.RDS")
+
+site_list <- names(Gpars_all)
+
+output_list <- list()
+
+for(do_site in site_list){
+  Gpars_tmp <- Gpars_all[[do_site]]
+  Spars_tmp <- Spars_all[[do_site]]
+  Rpars_tmp <- Rpars_all[[do_site]]
+  
+  #   if(do_site=="Idaho"){
+  #     spp_list <- names(Gpars_tmp)
+  #     Rp <- as.data.frame(Rpars_tmp)
+  #     Rp$species <- c(rep(spp_list, each=4),
+  #                     rep(spp_list, each=1),
+  #                     rep(spp_list, each=6),
+  #                     rep(spp_list, each=1),
+  #                     rep(spp_list, each=1),
+  #                     rep(spp_list, each=22),
+  #                     rep(spp_list, each=412),
+  #                     rep(spp_list, each=1),
+  #                     rep(spp_list, each=1))
+  #     Rp <- subset(Rp, species!="ARTR")
+  #     colid <- which(colnames(Rp)=="species")
+  #     Rp <- Rp[,-colid]
+  #     Rp$species <- c(rep(c("ARTR", "fine", "fine", "fine"), 3),
+  #                     rep("fine", (nrow(Rp)-12)))
+  #     Rp <- subset(Rp, species=="fine")
+  #     colid <- which(colnames(Rp)=="species")
+  #     Rpars_tmp <- as.matrix(Rp[,-colid])
+  #   }
+  #   
+  #   if(do_site=="Idaho"){
+  #     id <- which(names(Gpars_tmp)=="ARTR")
+  #     Gpars_tmp <- Gpars_tmp[-id]
+  #     Spars_tmp <- Spars_tmp[-id] 
+  #   }
+  
+  spp_list <- names(Gpars_tmp)
+  Nyrs <- nrow(Gpars_tmp[[1]])
+  
+  #Import and format parameters
+  site_path <- paste("../data/", do_site, sep="")
+  Gpars <- format_growth_params(do_site = do_site, species_list = spp_list, 
[…]

### 2015-07-12 runscripts/run_ipm_source.R

@@ -12,12 +12,31 @@ NoOverlap_Inter=FALSE
   iter_matrix_dims=iter_matrix_dims; max_size=max_size
   Rpars=Rpars; Spars=Spars; Gpars=Gpars
   demographic_stochasticity=do_demo_stoch
+  spp_interact = spp_interact
 
   # Turn off random year effects if constant==TRUE
   if(constant==TRUE){
     Rpars$intcpt.yr=matrix(Rpars$intcpt.mu,Nyrs,Nspp,byrow=T)
     Gpars$intcpt.yr[]=0;Gpars$slope.yr[]=0
     Spars$intcpt.yr[]=0;Spars$slope.yr[]=0    
+  }
+
+  # Turn off competition if spp_interact==FALSE
+  if(spp_interact==FALSE){
+    rnbtmp <- Rpars$dd
+    rnbtmp[] <- 0
+    diag(rnbtmp) <- diag(Rpars$dd)
+    Rpars$dd <- rnbtmp
+    
+    gnbtmp <- Gpars$nb
+    gnbtmp[] <- 0
+    diag(gnbtmp) <- diag(Gpars$nb)
+    Gpars$nb <- gnbtmp
+    
+    snbtmp <- Spars$nb
+    snbtmp[] <- 0
+    diag(snbtmp) <- diag(Spars$nb)
+    Spars$nb <- snbtmp    
   }
   
   # Build initial vectors and matrices

### 2015-07-16 inst/examples/demog_stoch_test.Rmd
adds results from demographic stoch test

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{todonotes}

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

---

Comparing IPM with approximate demographic stochasticity to IBM
===============================================================

Our approach to incorporating demographic stochasticity in integral projection models (IPMs) is new, and thus requires testing.
Individually-based models (IBMs) inherently include demographic stochasticity (e.g., "coin flips" for survival of a genet from a Bernouli trial).
IBMs, then, provide the baseline against which to compare an IPM with our approximation to demogrpahic stochasticity.
If the two models produce similar equilibrium cover and species synchrony, then we can conclude our approximation is valid.
Here we use the Idaho dataset as a test case, and simulate communities from the IPM with demographic stochastisticy and the IBM.
We ignore temporal variation due to random year effects to isolate variation from demographic stochasticity alone.

```{r figs, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2.5, fig.cap="Density estimates of equilibrium cover from 2,000 iterations from each model."}
library(ggplot2)
library(plyr)
library(reshape2)
ipm_outs <- readRDS("../../results/idaho_ipm_demogstoch_only.RDS")
ibm_outs <- readRDS("../../results/idaho_ibm_demogstoch_only.RDS")
ipm_outs <- as.data.frame(ipm_outs[501:2500,2:4])
ibm_outs <- as.data.frame(ibm_outs[501:2500,5:7])
ipm_outs$model <- "IPM"
ibm_outs$model <- "IBM"
ipm_outs$timestep <- c(1:2000)
ibm_outs$timestep <- c(1:2000)
colnames(ipm_outs) <- c("HECO", "POSE", "PSSP", "model", "timestep")
colnames(ibm_outs) <- c("HECO", "POSE", "PSSP", "model", "timestep")
outs <- rbind(ipm_outs, ibm_outs)

outd <- melt(outs, id.vars = c("model", "timestep"))

ggplot(outd)+
  geom_line(stat="density", aes(x=value*100, color=model))+
  facet_wrap("variable", scales="free")+
  scale_color_manual(values=c("black", "purple"))+
  xlab("equilibrium cover (%)")+
  theme_bw()
```

```{r synch_calcs, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(synchrony)
library(xtable)
ipm_outs <- readRDS("../../results/idaho_ipm_demogstoch_only.RDS")
ibm_outs <- readRDS("../../results/idaho_ibm_demogstoch_only.RDS")
ipm_mat <- ipm_outs[501:2500, 2:4]
ibm_mat <- ibm_outs[501:2500, 5:7]

####
####  Synchrony of population size (cover)
####
ipm_synch <- as.numeric(unlist(community.sync(ipm_mat)[1]))
ibm_synch <- as.numeric(unlist(community.sync(ibm_mat)[1]))

####
####  Synchrony of population growth rates
####
sppList <- c("HECO", "POSE", "PSSP")
cov_d <- as.data.frame(ipm_mat)
colnames(cov_d) <- sppList
cov_d$year <- c(1:nrow(cov_d))
lag_d <- cov_d
lag_d$lagYear <- lag_d$year+1
colnames(lag_d)[1:3] <- paste(sppList,".t0", sep="")
lag_d <- lag_d[,-4]
all_d <- merge(cov_d, lag_d, by.x = "year", by.y="lagYear")
transitions <- nrow(all_d)
obs_gr <- matrix(nrow=transitions, ncol=length(sppList))
for(i in 1:transitions){
  obs_gr[i,] <- as.numeric(log(all_d[i,2:4]/all_d[i,5:7]))
}
ipm_gr_synch <- as.numeric(unlist(community.sync(obs_gr[,1:3])[1]))

sppList <- c("HECO", "POSE", "PSSP")
cov_d <- as.data.frame(ibm_mat)
colnames(cov_d) <- sppList
cov_d$year <- c(1:nrow(cov_d))
lag_d <- cov_d
lag_d$lagYear <- lag_d$year+1
colnames(lag_d)[1:3] <- paste(sppList,".t0", sep="")
lag_d <- lag_d[,-4]
all_d <- merge(cov_d, lag_d, by.x = "year", by.y="lagYear")
transitions <- nrow(all_d)
obs_gr <- matrix(nrow=transitions, ncol=length(sppList))
for(i in 1:transitions){
  obs_gr[i,] <- as.numeric(log(all_d[i,2:4]/all_d[i,5:7]))
}
ibm_gr_synch <- as.numeric(unlist(community.sync(obs_gr[,1:3])[1]))

####
####  Make table
####
table_data <- data.frame(Model = c("IPM", "IBM"),
                         Cover_Synch = c(round(ipm_synch,2), round(ibm_synch,2)),
                         PGR_Synch = c(round(ipm_gr_synch,2), round(ibm_gr_synch,2)))
colnames(table_data) <- c("Model", "Cover Sychrony", "Growth Rate Synchrony")
print(xtable(table_data, caption = "Species synchrony of percent cover and per capita growth rates."), type="latex", comment=FALSE, include.rownames=FALSE, caption.placement="top")
```

### 2015-07-16 runscripts/run_ipm_idaho_demogstoch.R

@@ -69,7 +69,7 @@ for(do_site in site_list){
     max_size <- c(170,40)
   }
   if(do_site=="Idaho"){
-    iter_matrix_dims <- c(50,75,50,50)
+    iter_matrix_dims <- c(50,75,50,75)
     max_size <- c(3000,202,260,225)
   }
   if(do_site=="Kansas"){

### 2015-07-16 runscripts/run_ipm_source.R

@@ -53,7 +53,7 @@ NoOverlap_Inter=FALSE
   nt <- inits$v
   
   # loop through species really quick to set low initial densities
-  for(i in 1:n_spp) nt[[i]][]=0.1
+  for(i in 1:n_spp) nt[[i]][]=1
   if(do_site=="Idaho")
     nt[[1]][] <- 0
   new.nt <- nt #set initial density vector to be fed into IPM

### 2015-07-16 inst/examples/demog_stoch_test.Rmd
adds update to demog stoch description

@@
 Our approach to incorporating demographic stochasticity in integral projection models (IPMs) is new, and thus requires testing.
-Individually-based models (IBMs) inherently include demographic stochasticity (e.g., "coin flips" for survival of a genet from a Bernouli trial).
+Individually-based models (IBMs) inherently include demographic stochasticity (e.g., "coin flips" for survival of a genet from a Bernoulli trial).
 IBMs, then, provide the baseline against which to compare an IPM with our approximation to demogrpahic stochasticity.
@@
 We ignore temporal variation due to random year effects to isolate variation from demographic stochasticity alone.
+
+Figure 1 compares the distribution of percent cover over 2,000 iterations (after a 500 iteration burn in) for each species from the IBM and the IPM with approximate demographic stochasticity.
+There are some small differences, but overall the approximation leads to similar long term population dynamics when compared to the IBM.
+The IPM and IBM also show similar interannual dynamics, as signified by the very similar synchrony metrics (Table 1). 
+Moreover, species synchrony from models where only demographic stochasticity is acting matches theoretical expectations:  theory predicts synchrony = 1/*S* (where *S* is species richness) when demographic stochasticity is much stronger than the effects of environmental stochasticity and species interactions.
+Indeed, the IPM produces synchrony = 0.3 (1/3 with 3 species) when we exclude random year effects (species interactions are present, but not influential due to small competition coefficients).
 

### 2015-07-16 runscripts/run_ipm_idaho_demogstoch.R
updates demographic stoch structure to match IPMdoit functions

@@ -107,5 +107,5 @@ for(do_site in site_list){
 } # end site loop
 
 # Save the output
-saveRDS(covSave, "../results/idaho_ipm_demogstoch_only.RDS")
+saveRDS(covSave, "../results/idaho_ipm_demogstoch_only_fromPACKAGE.RDS")
 

### 2015-07-16 runscripts/run_ipm_source.R

@@ -53,7 +53,7 @@ NoOverlap_Inter=FALSE
   nt <- inits$v
   
   # loop through species really quick to set low initial densities
-  for(i in 1:n_spp) nt[[i]][]=1
+  for(i in 1:n_spp) nt[[i]][]=0.1
   if(do_site=="Idaho")
     nt[[1]][] <- 0
   new.nt <- nt #set initial density vector to be fed into IPM
@@ -107,16 +107,24 @@ NoOverlap_Inter=FALSE
         # Make kernels and project
         popv=nt[[doSpp]]
         h=inits$h[[doSpp]]
-        K_matrix=make_K_matrix(inits$v[[doSpp]],crowd_list$WmatG[[doSpp]],
-                               crowd_list$WmatS[[doSpp]],
-                               Rpars,recs_per_area,Gpars,Spars,
-                               doYear,doSpp,h=inits$h[[doSpp]],
-                               demo_stoch=demographic_stochasticity,popv=nt[[doSpp]]) 
         if(demographic_stochasticity==FALSE){
+          K_matrix=make_K_matrix(inits$v[[doSpp]],crowd_list$WmatG[[doSpp]],
+                                 crowd_list$WmatS[[doSpp]],
+                                 Rpars,recs_per_area,Gpars,Spars,
+                                 doYear,doSpp,h=inits$h[[doSpp]]) 
           new.nt[[doSpp]]=K_matrix%*%nt[[doSpp]]  
         }
         if(demographic_stochasticity==TRUE){
-          new.nt[[doSpp]]=K_matrix[[1]]+K_matrix[[2]]  
+          P.matrix <- make.P.matrix(inits$v[[doSpp]],crowd_list$WmatG[[doSpp]],
+                                    crowd_list$WmatS[[doSpp]],Gpars,Spars,doYear,doSpp)  
+          R.matrix <- make.R.matrix(inits$v[[doSpp]],Rpars,recs_per_area,doYear,doSpp)  
+          covmat <- get_cov(K=P.matrix*inits$h[[doSpp]])
+          pCont <- GenerateMultivariatePoisson(pD = length(nt[[doSpp]]),
+                                               samples = 1,
+                                               R = covmat,
+                                               lambda = inits$h[[doSpp]]*P.matrix%*%nt[[doSpp]])
[…]

### 2015-09-12 docs/prelim_analysis.Rmd
updated prelim ms with new results

@@
 ---
+layout: 12pt
+header-includes:
+   - \usepackage{lineno}
+   - \linenumbers
+   - \usepackage{setspace}
+   - \doublespacing
+   - \usepackage{todonotes}
+   - \usepackage[document]{ragged2e}
+
+## rmarkdown render options
 output:
@@
 geometry: margin=1in
+
 ---
-# Drivers of species synchrony in real communities
-*Andrew Tredennick, Peter Adler, Claire de Mazancourt, Michele Loreau, and Stephen Ellner*
+# Environmental responses synchronize population \newline dynamics in five semi-arid grasslands
 
-Quick summary
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+
+Andrew T.
+Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, \newline Stephen Ellner\textsuperscript{3}, and Peter B.
+Adler\textsuperscript{1}
+
+\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}
+
+\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}
+
+\textit{\small{\textsuperscript{3}Department of Ecology and Evolutionary Biology, Cornell University, Ithaca, NY 14853, USA}}
+
+\renewcommand*{\thefootnote}{\arabic{footnote}}
+\setcounter{footnote}{0}
+
+Abstract
 -------------
-1.
-__Question__.
-How (a)synchronous are species population dynamics in natural communities, and what determines the degree of synchrony among species?
-Of particular interest to us, do species' responses to environmental fluction increase or decrease population synchrony?
+Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time. Theory
 
-2. __Data__. Five longterm demographic datasets from semi-arid North American grasslands. All species coexist stabley, yet are similar -- they are all perennial grasses.
-
-3.
-__Analysis__.
-(a) Calculate community synchrony from the observed time series using an established metric, and compare this to expectation from theory.
-(b) Use IPMs with environmental or demographic stochasticity turned on/off to quantify the relative contribution of each to community synchrony.
-
-4.
-__Initial conclusions__.
-Species' responses to environmental fluctuations tend to synchronize population dynamics: at four of the five sites removing environmental fluctuations decreased community synchrony.
-However, even with environmental fluctuations present, populations were not completely synchronized.
-This implies that niche differences among these species remain strong even as they respond similarly to environmental conditions.
+__Keywords__
+synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
 
@@
 ```{r read_data, echo=FALSE}
-output_list <- readRDS("../results/ipm_simulation_lists.RDS")
+output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
 mlist <- melt(output_list)
@@
 colnames(mlist)[1:3] <- c("year", "species", "cover")
-mlist <- subset(mlist, year>500)
 sites <- unique(mlist$L1)
@@
 pgr_synch <- melt(pgr_list)
+expected_synch <- table_data[,c(1,3)]
+expected_synch$L2 <- "E(SYNCH)"
+colnames(expected_synch) <- c("L1", "value", "L2")
+expected_synch <- expected_synch[,c("value", "L2", "L1")]
+expected_synch[which(expected_synch[,"L1"]=="New Mexico"),"L1"] <- "NewMexico"
 
-ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
-  geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
-  facet_wrap("L1", scales="free_x")+
-  scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
-  guides(fill=FALSE)+
+pgr_synch <- rbind(pgr_synch, expected_synch)
+
+# ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
+#   geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
+#   facet_wrap("L1", scales="free_x")+
+#   scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
+#   guides(fill=FALSE)+
+#   scale_y_continuous(limits=c(0,1))+
+#   xlab("Source of stochasticity")+
+#   ylab("Species synchrony")
+
+pgrsynch_plot <- ggplot(pgr_synch, aes(x=L2, y=value, color=L1, group=L1))+
+  geom_line()+
+  geom_point(size=5)+
+  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
+                     name = "Site")+
   scale_y_continuous(limits=c(0,1))+
-  xlab("Source of stochasticity")+
-  ylab("Species synchrony")
+  xlab("Simulation")+
+  ylab("Species synchrony")+
+  theme_bw()
 
-```
+####
+####  Compare observed and IPM simulated synchrony
+####
+nsites <- length(sites)
 
-```{r niches, echo=FALSE}
-# niche_diff <- data.frame(site=sites,
-#                          ND = c(0.58, 0.59, 0.66, 0.69, 0.68))
-# pgr_subs <- subset(pgr_synch, L2!="ENVDEMO")
-# pgr_cast <- dcast(pgr_subs, L1~L2, value.var = "value")
-# synch_diff <- ddply(pgr_cast, .(L1), summarise,
-#                     value = ENV-DEMO)
-# niche_synch <- merge(niche_diff, synch_diff, by.x="site", by.y="L1")
-# ggplot(niche_synch, aes(x=ND, y=value))+
-#   geom_point(size=5)+
-#   ylab("Community synchrony")
 ```

### 2015-09-12 docs/prelim_analysis.Rmd
adds new figure comparing IPM and observartion synchrony

@@
 -------------
-Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time. Theory
+Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time.
+While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
+Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence sychronize dynamics, demographic stochasticity desynchronize dynamics, and environmental forcing can either synchronize or desynchronize poplation dynamics depending on the correlation of species-specific environmental responses.
+Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony through time.
+We find that the trophically-similar grass species in our dataset have population dynamics that are synchronized by similar responses to environmental fluctuations.
+Species interactions tend to be weak, so removing interspecific competition did not reduce synchrony.
+The effect of demographic stochasticity depended on the spatial scale of our modeling.
+Though we found that similar responses to environmental fluctuations synchronized species, dynamics were not completly synchronous.
+We conclude that niche differentiation among species is strong enough to reduce the synchronizing effect of coupled density dependence via competition, but this is countered by the synchronizing effect of similar environmental responses.
+The next frontier is to identify the specific environmental drivers of species synchrony to inform predictive efforts.
 
-__Keywords__
-synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
+__Keywords__: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
 
@@
 
+## Effect of removing demographic and environmental stochasticity
+At four of our five study areas, the presence of environmental stochasticity via random year effects on vital rate regressions increased community synchrony of per capita population growth rates (Figure 1).
+Only at the Montana site (Northern Mixed Prairie) did incorporating species' responses to environmental conditions decrease community synchrony.
+However, even with similar responses to environmental conditions, community synchrony did not reach 1 at any of the sites.
+Thus, there remains sufficiently strong temporal niche differences to drive the community away from perfect sychrony.
+
+Theory predicts that demographic stochasticty is the only force that, in isolation, can desynchronize population dynamics.
+However, theory also predicts that demographic stochasticity is likely unable to counteract the synchronizing effect of environmental forces.
+Our results support this prediction in the cases where environmental stochasticity is a synchronizing force (compare ENV and ENVDEMO bars in Figure 1).
+
+\newpage{}
+
+Tables
+------
 ```{r load_libs, echo=FALSE, message=FALSE}
@@
 
-## Effect of removing demographic and environmental stochasticity
-At four of our five study areas, the presence of environmental stochasticity via random year effects on vital rate regressions increased community synchrony of per capita population growth rates (Figure 1).
-Only at the Montana site (Northern Mixed Prairie) did incorporating species' responses to environmental conditions decrease community synchrony.
-However, even with similar responses to environmental conditions, community synchrony did not reach 1 at any of the sites.
-Thus, there remains sufficiently strong temporal niche differences to drive the community away from perfect sychrony.
+\newpage{}
 
-*NOTE: we have not dealt with species interactions yet...*
-
+Figures
+-------
 ```{r read_data, echo=FALSE}
@@
 
-```{r get_synchs, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site.
+```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.", fig.width=6, fig.height=3}
+nyrs <- c(max(out_az$stability$num_observations),
+          max(out_id$stability$num_observations),
+          max(out_ks$stability$num_observations),
+          max(out_mt$stability$num_observations),
+          max(out_nm$stability$num_observations))
+
+boots <- 100
+sites <- sort(table_data$Site)
+sites[5] <- "NewMexico"
+synch <- matrix(NA, boots, length(sites))
+
+for(do_site in 1:length(sites)){
+  full_ipm_outs <- subset(mlist, L2=="ENVINTER")
+  site_ipm <- subset(full_ipm_outs, L1==sites[do_site])
+  if(sites[do_site]=="Idaho") site_ipm <- subset(site_ipm, Var2!="ARTR")
+  species_list <- unique(site_ipm$Var2)
+  for(i in 1:boots){
+    begin_year <- sample(x = 1:(max(site_ipm$Var1)-nyrs[do_site]), 1)
+    end_year <- begin_year+nyrs[1]
+    tmp <- subset(site_ipm, Var1 %in% begin_year:end_year)
+    tmp2 <- dcast(tmp, Var1~Var2, value.var = "value")
+
+    # caclulate observed growth rates
+    # create lagged data frame to only get observed yearly transitions
+    lag_df <- tmp2
+    lag_df$lagyear <- lag_df$Var1+1
+    colnames(lag_df)[2:(num_spp+1)] <- paste0(colnames(lag_df)[2:(num_spp+1)],"_t0")
+
+    # merge the lag df with observed
+    rm_col <- which(colnames(lag_df)=="Var1")
+    merged_df <- merge(tmp2, lag_df[,-rm_col], by.x = "Var1", by.y="lagyear")
+    transitions <- nrow(merged_df)
+    obs_gr <- matrix(nrow=transitions, ncol=num_spp)
+    name_ids1 <- which(colnames(merged_df) %in% species_list)
+    name_ids2 <- which(colnames(merged_df) %in% paste(species_list, "_t0", sep=""))
+    for(ii in 1:transitions){
+      obs_gr[ii,] <- as.numeric(log(merged_df[ii,name_ids1]/merged_df[ii,name_ids2]))
+    }
+    synch[i,do_site] <- as.numeric(community.sync(obs_gr)[1])
+  }
+}
+
+colnames(synch) <- sites
+synch_df <- melt(synch)
+
+obs_pgr_df <- c(as.numeric(out_az$pgr_synchrony[1]),
+                as.numeric(out_id$pgr_synchrony[1]),
+                as.numeric(out_ks$pgr_synchrony[1]),
+                as.numeric(out_mt$pgr_synchrony[1]),
+                as.numeric(out_nm$pgr_synchrony[1]))
+obs_pgr_df <- data.frame(pgr_synch=obs_pgr_df,
+                         Var2=sites)
+
+ggplot()+
+  geom_line(data=synch_df, aes(x=value, color=Var2), stat="density", adjust=1)+
+  geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
+  facet_wrap("Var2")+
+  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
+  guides(color=FALSE)+
+  theme_bw()
+```
+
+```{r figure_2, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site.
 DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6}
@@
 
+print(pgrsynch_plot)
+
 ####
@@
 
-Theory predicts that demographic stochasticty is the only force that, in isolation, can desynchronize population dynamics.
-However, theory also predicts that demographic stochasticity is likely unable to counteract the synchronizing effect of environmental forces.
-Our results support this prediction in the cases where environmental stochasticity is a synchronizing force (compare ENV and ENVDEMO bars in Figure 1).
-
 \newpage
@@
 ------------------
-```{r time_series, echo=FALSE, fig.cap="Example time series from the simulations."}
+```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations."}
 ggplot(mlist, aes(x=year, y=(cover*100), color=species))+

### 2015-10-23 docs/prelim_analysis.Rmd
updates ms with site description

@@
 While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
-Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence sychronize dynamics, demographic stochasticity desynchronize dynamics, and environmental forcing can either synchronize or desynchronize poplation dynamics depending on the correlation of species-specific environmental responses.
+Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence sychronizes dynamics, demographic stochasticity desynchronizes dynamics, and environmental forcing can either synchronize or desynchronize poplation dynamics depending on the correlation of species-specific environmental responses.
 Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony through time. 
@@
 Species interactions tend to be weak, so removing interspecific competition did not reduce synchrony. 
-The effect of demographic stochasticity depended on the spatial scale of our modeling.
-Though we found that similar responses to environmental fluctuations synchronized species, dynamics were not completly synchronous.
+The effect of demographic stochasticity depended on the spatial scale of our modeling; demographic stochasticity was strongest in small populations and was able to modestly counteract the synchronizing effect of the environment.
+Although we found that similar responses to environmental fluctuations synchronized species, dynamics were not completly synchronous. 
 We conclude that niche differentiation among species is strong enough to reduce the synchronizing effect of coupled density dependence via competition, but this is countered by the synchronizing effect of similar environmental responses.
@@
 Species-specific responses to environmental conditions, in particular, can stabilize ecosystem properties even as environmental forcings fluctuate through time. 
-Such asynchrony of species dynamics in response to the environment has the potential to buffer ecosystems subjected to an increasingly variable climate.
-Thus, the extent to which natural communities exhibit asynchronous dynamics determines, in part, their ability to cope with interannual climate extremes that are projected to become more common.
+Such asynchrony of species dynamics in response to the environment has the potential to buffer ecosystems subjected to an increasingly variable envirnmental conditions.
 
@@
 It is impossible to draw inference on the relative contribution of demographic and environmental stochasticity on synchrony based on empirical estimates alone, so our analysis involves two steps.  
-First, we will use longterm datasets from five North American grasslands (Fig.
-1; Chu and Adler 2015) to calculate community synchrony.
-Second, we will fit multispecies population models to the data for dominant species to evaluate the relative contribution of species' environmental responses and demographic stochasticity to community synchrony.
+First, we will use longterm datasets from five North American grasslands (Fig. 1; Chu and Adler 2015) to calculate community synchrony. 
+Second, we will fit multispecies population models to the data for dominant species to evaluate the relative contribution of species' environmental responses, interspecific interactions, and demographic stochasticity to community synchrony.
 Sparse data on sub-dominant species constrained model fitting to dominant species that account for ##% - ##% of local plant abundance at their respective sites. 
@@
 ## Study sites and species
-We use the exact same data as presented in Chu and Adler (2015). From their paper:
+In the early 20th century, several range experiment stations in the western U.S. began mapping permanent quadrats, and conducted annual censuses for decades. 
+We use such long-term demographic data sets from five semiarid grasslands in the western United States (described in detail by Chu and Adler 2015).
+Each study site includes a set of $1-m^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
+The resulting mapped polygons represent basal cover for grasses and canopy cover for shrubs.
+Data come from the Sonoran desert in Arizona (Anderson et al.
+2012), sagebrush steppe in Idaho (Zachmann et al.
+2010), southern mixed prairie in Kansas (Adler et al.
+2007), northern mixed prairie in Montana (Anderson et al.
+2011), and Chihuahuan desert in New Mexico (Anderson et al.
+in preparation, Chu and Adler 2015).
 
->Many range experiment stations in the western U.S.
-began mapping permanent quadrats in the early 20th century, and continued annual censuses for decades.
-Here we focus on five long-term data sets (Fig.
-1), four of which have been distributed publicly (Adler et al.
-2007b, Zachmann et al.
-2010, Anderson et al.
-2011, 2012).
-At each study site, all individual plants within each 1-m2 quadrat were identified and mapped annually using a pantograph (Hill 1920; Fig.
-2).
-Mapped polygons represent basal cover for grasses and canopy cover for shrubs (Fig.
-2).
-
->To fit our population models, we needed to study species common enough to provide a large sample size (i.e.
-the most common species at each site).
-In order to describe species interactions, we needed to select species that co-occurred together across many years.
-Thus, our analysis focused on the niche differences and average fitness differences of the most common, co-occurring species at each site, which implies that the average fitness differences among species could be relatively small.
-To select these species, we first identified species that occurred in at least 20% of years.
-Then we used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
-Based on the degree of aggregation of quadrats on the NMDS plot, we selected quadrats and species for our full analysis.
-We also confirmed these selections with local personnel from each study site.
-Below we further describe each study site and our criteria for quadrat and species selections.
-
->_Sonoran desert, Arizona_
-
->The Sonoran desert data set comes from 178 permanent quadrats located on semi-desert grasslands at the Santa Rita Experimental Range (31º50’ N, 110º53’ W; elevation 1150 meters), Arizona (Anderson et al.
-2012).
-Mean annual precipitation was 450 mm with the monthly variation from 5 mm (May) to 107 mm (July), and mean annual temperature was 16 °C with the monthly variation from 8 °C (January) to 26 °C (July) during the sampling period.
-These quadrats were mapped annually from 1915 to 1933 in most cases.
-The dominant flora varies across the broad range of soils and topography that occur within the semi-desert grasslands.
-For our present analysis, we chose 32 quadrats dominated by black grama Bouteloua eriopoda (BOER) and Bouteloua rothrockii (BORO).
-
->_Sagebrush steppe, Idaho_
-
->The sagebrush data set comes from the U.S.
-Sheep Experiment Station, currently a US Department of Agriculture Agricultural Research Service field station, located 9.6 km north of Dubois, Idaho (44.2°N, 112.1°W; elevation 1500 meters).
-26 quadrats were established between 1926 and 1932 (Zachmann et al.
-2010).
-During the period of sampling (1926 – 1957), mean annual precipitation was 270 mm with the monthly variation from 16 mm (March) to 42 mm (June), and mean annual temperature was 6 °C with the monthly variation from -8 °C (January) to 21 °C (July).
-The vegetation is dominated by the shrub, Artemisia tripartita (ARTR), and the C3 perennial bunchgrasses Pseudoroegneria spicata (PSSP), Hesperostipa comata (HECO), and Poa secunda (POSE).
-
->_Southern mixed prairie, Kansas_
-
->The southern mixed prairie data set represents the pioneering work of Albertson and colleagues (Albertson and Tomanek 1965), who established 51 permanent quadrats inside and outside livestock exclosures near Hays, Kansas (38.8°N, 99.3°W; elevation 650 meters).
-The mean annual precipitation was 580 mm with the monthly variation from 10 mm (January) to 103 mm (June), and mean annual temperature was 12 °C with the monthly variation from -2.0 °C (January) to 27 °C (July).
-These quadrats were mapped from 1932 to 1972 at the end of each growing season (Adler et al.
-2007b).
-Variation in soil depth and texture creates distinct plant communities.
-Many of the quadrats were located on shortgrass communities dominated by Bouteloua gracilis and Buchloë dactyloides.
-We had difficulty working with these species, which can be hard to distinguish in the field in the absence of inflorescences, and were often mapped inconsistently.
-Therefore, we focused on 7 quadrats dominated by Bouteloua curtipendula (BOCU), Bouteloua hirsuta (BOHI), and Schizachyrium scoparium (SCSC).
-
->_Northern mixed prairie, Montana_
-
->The northern mixed prairie data set comes from the Fort Keogh Livestock and Range Research Laboratory, another USDA Agricultural Research Service field station.
-The study site is located on alluvial plains near the Tongue River (46o19’N, 105o48’W; elevation 720 meters).
-Mean annual precipitation was 343 mm with the monthly variation from 7 mm (February) to 74 mm (June), and mean annual temperature was 8 °C with the monthly variation from -7 °C (January) to 25 °C (July).
-44 quadrats, distributed across six pastures assigned to one of three grazing intensities, were mapped annually (with some exceptions) from 1932 through 1945 (Anderson et al.
-2011).
-Our analysis includes 19 quadrats in which Bouteloua gracilis (BOGR), Hesperostipa comata (HECO), Pascopyrum smithii (PASM), and Poa secunda (POSE) were prevalent.
-
->_Chihuahuan desert, New Mexico_
-
->The last data set consists of 76 permanent quadrats located in Chihuahuan desert plant communities at the Jornada Experimental Range (32.62°N, 106.67°W; elevation 1260 meters), New Mexico.
-Mean annual precipitation was 264 mm with the monthly variation from 6 mm (April) to 48 mm (August), and mean annual temperature was 14 °C with the monthly variation from 4 °C (January) to 26 °C (July).
-Quadrats were established from 1915-1920 and mapping was conducted annually until the 1960’s (it continues on roughly 5 yr intervals to the present).
-Due to the changes in mapping frequency in the 1960’s, and the effects of a severe drought in the 1950’s that triggered a conversion of many quadrats from grassland to shrubland, we analyzed data for the period from 1915 to 1950.
-We selected 48 quadrats in which both Bouteloua eriopoda (BOER) and Sporobolus flexuosus (SPFL) were common (Anderson et al.
-in preparation).
+Chu and Adler describe the selection of focal species in each data set, so we offer a brief explanation here.
+Focal species needed to comply with two criteria: (1) be common enough to provide a large sample size for model fitting and (2) co-occur with other focal species often enough to allow us to describe species interactions within the model.
+Thus, Chu and Adler (2015) chose species that occurred in at least 20% of years and then used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
 
@@
 While the observational data allows us to directly calculate species synchrony within the community, it is impossible to isolate the contributions of demographic and environmental stochasticity on species synchrony based on the data alone.
-So we now turn to a empirically-based modeling approach where we can turn demographic and environmental stochasticity on/off.
-To our knowledge, Integral Projection Models (IPMs) are the only tools available that allow us to isolate sources of stochasticity.
-For example, individual-based models require demographic stochasticity, making it impossible to isolate the effect of environmental stochasticity alone.
-
-We describe our approach in three steps.
-First, we describe the statistical models for three vitale rates: survival, growth, and recruitment.
-Second, we describe the IPM construction in detail.
-Third, we describe the IPM simulations we conducted to answer our questions.
+So we now turn to a empirically-based modeling approach where we can turn species' interactions and demographic and environmental stochasticity on/off. 
+We use two models: (1) an environmentally stochastic Integral Projection Model (IPM) and (2) a demographically and environmentally stochastic Individual Based Model (IBM).
+Both models are driven by the same vital rate regressions for survival, growth, and recreuitment.
+We use the IPM to explore the drivers of synchrony assuming a large population where demographic stochasticity is weak or absent.
+We use the IBM to isolate the effect of demographic stochasticity and to examine how demographic and environmental stochasticity interact as population size increases (i.e., as the effect of demographic stochasticity weakens).
+The use of two model types (IPM and IBM) is the only way to isolate the effects of environmental stochasticity (an IPM with no demographic stochasticity) and demographic stochasticity (an IBM with no environmental stochasticity).
 
@@
 ### Survival and growth
-We modeled survival probability and growht on individual genets as a function of genet size, the crowding experienced by the focal genet from both heterospecific and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
+We modeled survival probability and growth on individual genets as a function of genet size, the crowding experienced by the focal genet from both heterospecific and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
 Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure).

### 2015-10-23 docs/prelim_analysis.Rmd
adds in new sims from IBM

@@
 We use such long-term demographic data sets from five semiarid grasslands in the western United States (described in detail by Chu and Adler 2015).
-Each study site includes a set of $1-m^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
+Each study site includes a set of 1-$\text{m}^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
 The resulting mapped polygons represent basal cover for grasses and canopy cover for shrubs.
@@
 ------
-```{r load_libs, echo=FALSE, message=FALSE}
+
+\singlespacing
+
+Table 1: Site descriptions and focal species.
+\footnotesize
+
+| Site Name | Biome | Location (Lat, Lon) | Obs. Years | Species |
+| --------- | ----- | ------------------- | ---------- | ------- |
+| Arizona | Sonoran Desert | 31&deg;50' N, 110&deg;53' W | 1915-1933 | *Bouteloua eriopoda* |
+| | | | | *Bouteloua rothrockii* |
+| Idaho | Sagebrush steppe | 44.2&deg; N, 112.1&deg; W | 1926-1957 | *Artemisia tripartita* |
+| | | | | *Pseudoroegneria spicata* |
+| | | | | *Hesperostipa comata* |
+| | | | | *Poa secunda* |
+| Kansas | Southern mixed prairie | 38.8&deg; N, 99.3&deg; W | 1932-1972 | *Bouteloua curtipendula* |
+| | | | | *Bouteloua hirsuta* |
+| | | | | *Schizachyrium scoparium* |
+| Montana | Northern mixed prairie | 46&deg;19' N, 105&deg;48' W | 1926-1957 | *Bouteloua gracilis* |
+| | | | | *Hesperostipa comata* |
+| | | | | *Pascopyrum smithii* |
+| | | | | *Poa secunda* |
+| New Mexico | Chihuahuan Desert | 32.62&deg; N, 106.67&deg; W | 1915-1950 | *Bouteloua eriopoda* |
+| | | | | *Sporobolus flexuosus* |
+
+```{r load_libs, echo=FALSE, message=FALSE, warning=FALSE}
 library(communitySynchrony)
@@
 -------
-```{r read_data, echo=FALSE}
+```{r read_data, echo=FALSE, eval=F}
 output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
@@
 
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.", fig.width=6, fig.height=3}
+```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.", fig.width=6, fig.height=3, eval=F}
 nyrs <- c(max(out_az$stability$num_observations),
@@
 ```{r figure_2, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site.
-DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6}
+DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6, eval=F}
 colnames(mlist)[1:3] <- c("year", "species", "cover")
@@
 ------------------
-```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations."}
+```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations.", eval=F}
 ggplot(mlist, aes(x=year, y=(cover*100), color=species))+

### 2015-10-26 docs/prelim_analysis.Rmd
updates rec fxn in IBM with sqrt(cover); appears to be working now

@@
 -------
+```{r concept_fig, echo=FALSE, eval=T}
+library(png)
+library(grid)
+img <- readPNG("./prelim_analysis_files/synchrony_conceptual_fig.png")
+grid.raster(img)
+```
+
 ```{r read_data, echo=FALSE, eval=F}

### 2015-10-26 runscripts/ibm_largepop_wrapper.R

@@ -122,8 +122,8 @@ for(constant in constant.vec){
       } # end Arizona do_site
       
       if(do_site == "Idaho"){
-        init.cover <- c(0,1,1,1)       # in percent cover
-        # init.cover <- rep(1, times=Nspp) # in percent cover
+        # init.cover <- c(0,1,1,1)       # in percent cover
+        init.cover <- rep(1, times=Nspp) # in percent cover
         maxSize <- c(8000,500,500,500)   # in centimeters
         minSize <- 0.25                  # in centimeters 
       } # end Idaho do_site
@@ -152,7 +152,7 @@ for(constant in constant.vec){
       ####
       for(expand in expand_vec){
         source("ibm_skeleton.R")
-        matplot(output[,8:11], type="l")
+        matplot(output[,4:7], type="l")
         ####
         ####  Save site output; raw time series
         ####

### 2015-10-26 runscripts/ibm_skeleton.R

@@ -64,7 +64,7 @@ recruit <- function(Rpars,sizes,spp,doYear,lastID,L,expand){
   # calculate recruits
   lambda <- rep(NA,Nspp) # seed production
   for(i in 1:Nspp){
-    lambda[i] <- totArea[i]*exp(Rpars$intcpt.yr[doYear,i]+totArea%*%Rpars$dd[i,])
+    lambda[i] <- totArea[i]*exp(Rpars$intcpt.yr[doYear,i]+sqrt(totArea)%*%Rpars$dd[i,])
   }
   # number of draws from distribution depends on size of landscape
   NN <- rnbinom(length(lambda)*expand^2,mu=lambda,size=Rpars$theta)  

### 2015-10-26 docs/prelim_analysis.Rmd
adds IBM results; updates ms intro

@@
    - \usepackage[document]{ragged2e}
+bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
+csl: components/ecology.csl
 
@@
 ---
-# Environmental responses synchronize population \newline dynamics in five semi-arid grasslands
+# Environmental responses synchronize population \newline dynamics in five semi-arid grasslands, but not completely
 
@@
 ------------
-Asynchrony in population dynamics among co-occuring plant species stabilizes ecosystem-level properties like annual biomass production (Hector et al.
-2010, Loreau and de Mazancourt 2013, de Mazancourt et al.
-2013, Hautier et al.
-2014).
-Species-specific responses to environmental conditions, in particular, can stabilize ecosystem properties even as environmental forcings fluctuate through time. 
-Such asynchrony of species dynamics in response to the environment has the potential to buffer ecosystems subjected to an increasingly variable envirnmental conditions.
+All populations fluctuate through time as a result of density-dependence (inter and intraspecific interactions), demographic stochasticity, and fluctuating environmental forces [@Lande2003; @Loreau2008a; @Mutshinda2009; @Gonzalez2009].
+Population fluctuations play integral roles at the community and ecosystem levels by creating opportunities for coexistence [@Chesson2000a] and by stabilizing aggregate properties like annual biomass production [@Hector2010; @Loreau2013; @DeMazancourt2013a; @Hautier2014].
+In both cases (species coexistence and ecosystem stability), temporal niche differentiation driven by species-specific responses to environmental conditions is necessary to create asynchrounous population dynamics.
+However, detecting and interpreting species asynchrony in natural systems has been plagued by methodological issues, and few have attempted to decompose the drivers of species synchrony.
+Thus, we lack general answers to two fundamental questions that have fascinated ecologists for decades: (1) How (a)synchronous are interacting populations in natural ecosystems?
+and (2) What drives temporal (a)synchrnous dynamics in the real world?
 
-Theory suggests that community synchrony is driven by a combination of internal and external forces. 
-Internal to the community, demographic stochasticity can induce asynchronous dynamics when species tend to fluctuate independently in a constant environment (Loreau and de Mazancourt 2008). 
-However, species live in non-constant environments that act as strong external drivers of community dynamics. 
-Environmental variability can induce asynchronous dynamics when species' responses to environmental conditions are temporally uncorrelated. 
-When species respond similarly to environmental conditions, community dynamics become more synchronized (Gonzalez and Loreau 2009, Loreau and de Mazancourt 2008).
+The seminal study by @Houlahan2007 suggested that asynchronous population dynamics are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
+The conclusion of @Houlahan2007, that abiotic factors synchronize species' abundances, rests on the implicit assumption that species will fluctuate independently in the absence of temporal niche differences.
+Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
+However, the null hypothesis of independent fluctuations is the incorrect starting point from which to judge spceies synchrony.
 
-It is impossible to draw inference on the relative contribution of demographic and environmental stochasticity on synchrony based on empirical estimates alone, so our analysis involves two steps.  
-First, we will use longterm datasets from five North American grasslands (Fig. 1; Chu and Adler 2015) to calculate community synchrony. 
-Second, we will fit multispecies population models to the data for dominant species to evaluate the relative contribution of species' environmental responses, interspecific interactions, and demographic stochasticity to community synchrony.
-Sparse data on sub-dominant species constrained model fitting to dominant species that account for ##% - ##% of local plant abundance at their respective sites. 
-Our focus is on temporal trends of per capita growth rates becuase these best reflect the short-term responses of populations to environmental conditions that fluctuate on annual time scales (Loreau and de Mazancourt 2008).
+All organisms exist in variable environments, and those that coexist on similar resources should be expected to respond to the environment in similar ways.
+Based on such a proposition, @Loreau2008 suggest that the true null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
+Then, the judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
+In such a light, Houlahan et al.'s (2007) results can be viewed very differently.
+However, Houlahan et al.'s (2007) meausure of compensatory dynamics was summed species covariances through time.
+Summed covariances have no upper or lower bound, making it difficult to interpret their results based on deviations from perfect synchrony.
+In the end, the fact that few communities exhibited negative summed covariance yields little insight as to whether communities tend to be sychronized by environmental forces or not.
+
+Aside from issues relating to the null expectation of species synchrony, disentangling the factors that drive species syncrony is difficult.
 
@@
 ---------------------------
-[^devel]: Note that this is still in development...so some of it may be wrong.
 
@@
 -------
-```{r concept_fig, echo=FALSE, eval=T}
+```{r concept_fig, echo=FALSE, fig.cap="Conceptual figure outlining the effects of interspecific interactions and environmental stochasticity in large populations where demographic stochasticity is effectively absent (A) and in populations of varying size where demographic stochasticity is present (B).", eval=T}
 library(png)

### 2015-10-26 runscripts/ibm_largepop_wrapper.R

@@ -152,7 +152,7 @@ for(constant in constant.vec){
       ####
       for(expand in expand_vec){
         source("ibm_skeleton.R")
-        matplot(output[,4:7], type="l")
+        # matplot(output[,4:7], type="l")
         ####
         ####  Save site output; raw time series
         ####

### 2015-10-27 docs/ms_communitySynchrony.Rmd
rms bad IBM runs; updates ms filename and caching

--- 2015-10-26 docs/prelim_analysis.Rmd
@@
    - \doublespacing
-   - \usepackage{todonotes}
    - \usepackage[document]{ragged2e}
@@
     fig_caption: true
-    keep_tex: true
+    keep_tex: false
 fontsize: 12pt
@@
 ---
-# Environmental responses synchronize population \newline dynamics in five semi-arid grasslands, but not completely
+```{r caching, include=FALSE}
+library("methods")
+library("knitr")
+basename <- "ms_communitySynchrony"
+opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
+               cache.path = paste("components/cache/", basename, "/", sep=""))
+opts_chunk$set(cache = 2)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
+               comment = NA, verbose = TRUE, echo=FALSE)
+
+# PDF-based figures
+opts_chunk$set(dev='pdf')
+```
+
+# Environmental responses synchronize population dynamics in five semi-arid grasslands, but not completely
 
@@
 All organisms exist in variable environments, and those that coexist on similar resources should be expected to respond to the environment in similar ways.
-Based on such a proposition, @Loreau2008 suggest that the true null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
+Based on such a proposition, @Loreau2008a suggest that the true null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
 Then, the judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
@@
 Aside from issues relating to the null expectation of species synchrony, disentangling the factors that drive species syncrony is difficult.
+Observational timeseries of species' abundances allows easy quantification of summed covariances [@Houlahan2007] or species synchrony [@Gross2014], but quantifying the relative influence of density-dependence, demographic stochasticity, and environmental stochasticity remains a challenge.
+Indeed, to our knowledge, only one other study has attempted to separate the influences of density-dependence and environmental stochasticity [@Mutshinda2009].
+@Mutshinda2009 used a statistical technique to decompose synchrony of species' abundances calculate from five time series.
+They found that the effect of environmental forcing on temporal dynamics is much stronger than that of interspecific interactions, across all taxa they analyzed [@Mutshinda2009].
+However, to link environmental forcing to species synchrony, @Mutshinda2009 examined the correlations among species' environmental responses.
 
-What is expected by theory?[^devel]
----------------------------
+In this paper, we seek to overcome the limitations of previous studies by explicitly testing theory on all the drivers of species synchrony using empirically-based multi-species population models that include the effects of density-dependence, environmental stochasticity, and demographic stochasticity.
+Our theoretical framework is that of @Loreau2008a, which defines expectations for the effects of density-dependence, environmental stochasticity, and demographic stochasticity on species synchrony.
+We confront theoretical expectations with model-based analyses of community time series from five semi-arid grasslands [@Chu2015].
+Our objectives were to (1) reproduce observed species synchrony using multi-species population models, (2) determine the relative influences of environmental stochasticity and interspecific interactions on species synchrony in large populations immune to demographic stochasticity, and (3) determine how dempographic stochasticity augments the influence of the environment and interspecific interactions across a range of populations sizes.
+To achieve our objectives, we analyzed spatially-implicit integral projection models (IPMs) and spatially-explicit individually-based models (IBMs) fit to demographic data from each our study sites.
+We first show that the models are able to reproduce observed synchrony of species' per capita growth rates and then perturb the models to include/exclude the factors to drive synchrony, as identified by theory.
+We used the IPM to model large populations that are immune to demographic stochasticity to isolate the influence of environmental stochasticity; we used the IBM to model populations where demographic stochasticity is present.
 
-Loreau and de Mazancourt (2008) derived a statistic to measure community-wide synchrony of per capita population growth rates:
+Theoretical expectations
+------------------------
+@Loreau2008a derived a statistic to measure community-wide synchrony of per capita population growth rates:
 
@@
 
+\normalsize
+
+\pagebreak{}
+
 ```{r load_libs, echo=FALSE, message=FALSE, warning=FALSE}
@@
 id_data <- id_data[2:nrow(id_data),] #remove first NA row
-id_data <- subset(id_data, species!="ARTR") #take out the shrub
+# id_data <- subset(id_data, species!="ARTR") #take out the shrub
 
@@
 
-```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site."}
+```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site.", eval=FALSE}
 # out_ks$percent_cover$site <- "Kansas"
@@
 
-\newpage{}
+\pagebreak{}
 
@@
 library(grid)
-img <- readPNG("./prelim_analysis_files/synchrony_conceptual_fig.png")
+img <- readPNG("./components/synchrony_conceptual_fig.png")
 grid.raster(img)
@@
 
-```{r read_data, echo=FALSE, eval=F}
+```{r read_data, echo=FALSE, eval=T}
 output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
@@
 
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.", fig.width=6, fig.height=3, eval=F}
+```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.", fig.width=6, fig.height=6, fig.width=3, eval=T}
 nyrs <- c(max(out_az$stability$num_observations),
@@
   site_ipm <- subset(full_ipm_outs, L1==sites[do_site])
-  if(sites[do_site]=="Idaho") site_ipm <- subset(site_ipm, Var2!="ARTR")
-  species_list <- unique(site_ipm$Var2)
+  colnames(site_ipm) <- c("year", "species", "cover", "L1", "L2")
+  # if(sites[do_site]=="Idaho") site_ipm <- subset(site_ipm, Var2!="ARTR")
+  species_list <- unique(site_ipm$species)
+  num_spp <- length(species_list)
   for(i in 1:boots){
-    begin_year <- sample(x = 1:(max(site_ipm$Var1)-nyrs[do_site]), 1)
-    end_year <- begin_year+nyrs[1]
-    tmp <- subset(site_ipm, Var1 %in% begin_year:end_year)
-    tmp2 <- dcast(tmp, Var1~Var2, value.var = "value")
+    begin_year <- sample(x = 1:(max(site_ipm$year)-nyrs[do_site]), 1)
+    end_year <- begin_year+nyrs[do_site]
+    tmp <- subset(site_ipm, year %in% begin_year:end_year)
+    tmp2 <- dcast(tmp, year~species, value.var = "cover")
 
@@
     lag_df <- tmp2
-    lag_df$lagyear <- lag_df$Var1+1
+    lag_df$lagyear <- lag_df$year+1
     colnames(lag_df)[2:(num_spp+1)] <- paste0(colnames(lag_df)[2:(num_spp+1)],"_t0")
@@
     # merge the lag df with observed
-    rm_col <- which(colnames(lag_df)=="Var1")
-    merged_df <- merge(tmp2, lag_df[,-rm_col], by.x = "Var1", by.y="lagyear")
+    rm_col <- which(colnames(lag_df)=="year")
+    merged_df <- merge(tmp2, lag_df[,-rm_col], by.x = "year", by.y="lagyear")
     transitions <- nrow(merged_df)
@@
 ggplot()+
-  geom_line(data=synch_df, aes(x=value, color=Var2), stat="density", adjust=1)+
+  geom_line(data=synch_df, aes(x=value, color=Var2), stat="density", adjust=2)+
   geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
-  facet_wrap("Var2")+
+  facet_wrap("Var2", ncol=1)+
   scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
@@
 ```{r figure_2, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site.
-DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6, eval=F}
+DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6, eval=T}
 colnames(mlist)[1:3] <- c("year", "species", "cover")
@@
   tmp_data <- subset(mlist, L1==dosite)
-  if(dosite=="Idaho"){
-    tmp_data <- subset(tmp_data, species!="ARTR")
-  }
+#   if(dosite=="Idaho"){
+#     tmp_data <- subset(tmp_data, species!="ARTR")
+#   }
   within_site_list <- list()
@@
 print(pgrsynch_plot)
-
-####
-####  Compare observed and IPM simulated synchrony
-####
-nsites <- length(sites)
-
 ```
 
-\newpage
+```{r ibm_sims_figure, echo=FALSE, fig.cap="Species synchrony from spatially-explicit IBM simulations with (A,C) and without (B,D) environmental fluctuations and with (A,B) and without (C,D) interspecific interactions.
+Points represent replicate IBM simulations (out of 20 possible) in which no species went extinct.
+Lines show linear fits between plot size (proxy for population size) and species synchrony.
+Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).", fig.height=5}
+library(png)
+library(grid)
+img <- readPNG("./components/ibm_sims_fig.png")
+grid.raster(img)
+```
 
-Additional Figures
-------------------
 ```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations.", eval=F}
@@
 ```
+
+\pagebreak{}
+
+References
+----------

### 2015-10-27 devel/plot_ibm_sims.R

@@ -480,8 +480,10 @@ const.nointer.plot <- ggplot(output_df, aes(x=expansion, y=synchrony, color=site
 ####
 ####  Arrange plots and save ---------------------------------------------------
 ####
+# png("../../docs/prelim_analysis_files/ibm_sims_fig.png", height = 10, width = 10, units = "in", res=100)
 out.plot <- grid.arrange(fluct.inter.plot, const.inter.plot, 
                          fluct.nointer.plot, const.nointer.plot,
                          nrow=2, ncol=2)
-
+print(out.plot)
+# dev.off()
 

### 2015-10-27 devel/plot_ipm_sims.R

@@ -1,0 +1,49 @@
+library(communitySynchrony)
+
+output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
+mlist <- melt(output_list)
+colnames(mlist)[1:3] <- c("year", "species", "cover")
+sites <- unique(mlist$L1)
+sims <- unique(mlist$L2)
+pgr_list <- list()
+for(dosite in sites){
+  tmp_data <- subset(mlist, L1==dosite)
+  if(dosite=="Idaho"){
+    tmp_data <- subset(tmp_data, species!="ARTR")
+  }
+  within_site_list <- list()
+  for(dosim in sims){
+    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
+    tmpsynch <- get_ipm_synchrony(tmpsim)
+    tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
+    within_site_list[[dosim]] <- tmp_pgr_synch
+  }
+  pgr_list[[dosite]] <- within_site_list
+}
+pgr_synch <- melt(pgr_list)
+expected_synch <- table_data[,c(1,3)]
+expected_synch$L2 <- "E(SYNCH)"
+colnames(expected_synch) <- c("L1", "value", "L2")
+expected_synch <- expected_synch[,c("value", "L2", "L1")]
+expected_synch[which(expected_synch[,"L1"]=="New Mexico"),"L1"] <- "NewMexico"
+
+pgr_synch <- rbind(pgr_synch, expected_synch)
+
+# ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
+#   geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
+#   facet_wrap("L1", scales="free_x")+
+#   scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
+#   guides(fill=FALSE)+
+#   scale_y_continuous(limits=c(0,1))+
+#   xlab("Source of stochasticity")+
+#   ylab("Species synchrony")
+
+ggplot(pgr_synch, aes(x=L2, y=value, color=L1, group=L1))+
+  geom_line()+
+  geom_point(size=5)+
+  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
+                     name = "Site")+
+  scale_y_continuous(limits=c(0,1))+
+  xlab("Simulation")+
+  ylab("Species synchrony")+
+  theme_bw()

### 2015-10-27 runscripts/ibm_largepop_wrapper.R

@@ -39,7 +39,7 @@ filename.flag <- c("fluctnointer", "fluctinter", "constnointer", "constinter")
 # filename.flag <- c("fluctinter", "constnointer", "constinter")
 
 ## Looping over different landscape sizes
-expand_vec <- c(1,2,3,4,5) # 1 = 1x1 m^2, 2 = 2x2m^2, etc
+expand_vec <- c(1,2) # 1 = 1x1 m^2, 2 = 2x2m^2, etc
 
 
 ####
@@ -50,7 +50,6 @@ Spars_all <- readRDS("../results/surv_params_list.RDS")
 Rpars_all <- readRDS("../results/recruit_parameters.RDS")
 
 site_names <- names(Gpars_all)
-site_names <- c("Arizona", "Idaho", "Kansas", "NewMexico")
 n_sites <- length(site_names)
 
 
@@ -58,116 +57,110 @@ n_sites <- length(site_names)
 ####
 ####  Start looping over simulation types --------------------------------------
 ####
-count <- 1 # counter for filename flags
-for(constant in constant.vec){
-  for(sppinter in sppinter.vec){
+for(i in 1:length(constant.vec)){
+  constant <- constant.vec[i]
+  sppinter <- sppinter.vec[i]
+  filename.flag.current <- filename.flag[i]
     
-    filename.flag.current <- filename.flag[count]
+  ####
+  ####  Start loop over sites ----------------------------------------------------  
+  ####
+  for(do_site in site_names){
+    ##  Get site-specific regression parameters
+    Gpars_site <- Gpars_all[[do_site]]
+    Spars_site <- Spars_all[[do_site]]
+    Rpars_site <- Rpars_all[[do_site]]
+    
+    ##  Define bookkeeping variables
+    sppList <- spp_list <- names(Gpars_site)
+    Nyrs <- nrow(Gpars_site[[1]])
+    Nspp <- length(sppList)
+    site_path <- paste("../data/", do_site, sep="")
+    
+    ##  Formate regression parameters for model
+    Gpars <- format_growth_params(do_site = do_site, species_list = spp_list, 
+                                  Nyrs = Nyrs, Gdata_species = Gpars_site)
+    Spars <- format_survival_params(do_site = do_site, species_list = spp_list, 
+                                    Nyrs = Nyrs, Sdata_species = Spars_site)
+    Rpars <- format_recruitment_params(do_site = do_site, species_list = spp_list, 
+                                       Nyrs = Nyrs, Rdata_species = Rpars_site,
[…]

### 2015-10-27 runscripts/ibm_largepop_wrapper.R
updates ibm run simulation loops

@@ -39,7 +39,7 @@ filename.flag <- c("fluctnointer", "fluctinter", "constnointer", "constinter")
 # filename.flag <- c("fluctinter", "constnointer", "constinter")
 
 ## Looping over different landscape sizes
-expand_vec <- c(1,2) # 1 = 1x1 m^2, 2 = 2x2m^2, etc
+expand_vec <- c(1,2,3,4,5) # 1 = 1x1 m^2, 2 = 2x2m^2, etc
 
 
 ####

### 2015-10-28 docs/ms_communitySynchrony.Rmd
updates ms ibm figure

@@
 __Keywords__: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
+
+<!-- \setlength{\parindent}{5ex} -->
 
@@
 Lines show linear fits between plot size (proxy for population size) and species synchrony.
-Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).", fig.height=5}
+Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).", fig.height=6}
 library(png)

### 2015-11-03 docs/ms_communitySynchrony.Rmd
updates figures

@@
 \setcounter{footnote}{0}
+
+\newpage{}
 
@@
 -------
-```{r concept_fig, echo=FALSE, fig.cap="Conceptual figure outlining the effects of interspecific interactions and environmental stochasticity in large populations where demographic stochasticity is effectively absent (A) and in populations of varying size where demographic stochasticity is present (B).", eval=T}
+```{r concept_fig, echo=FALSE, fig.cap="Conceptual figure outlining the effects of interspecific interactions and environmental stochasticity in large populations where demographic stochasticity is effectively absent (A) and in populations of varying size where demographic stochasticity is present (B).", eval=F}
 library(png)
@@
 
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.", fig.width=6, fig.height=6, fig.width=3, eval=T}
+```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.
+Density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.", fig.width=6, fig.height=6, fig.width=3, eval=T}
 nyrs <- c(max(out_az$stability$num_observations),
@@
 
-```{r figure_2, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site.
-DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6, eval=T}
-colnames(mlist)[1:3] <- c("year", "species", "cover")
-sites <- unique(mlist$L1)
-sims <- unique(mlist$L2)
-pgr_list <- list()
-for(dosite in sites){
-  tmp_data <- subset(mlist, L1==dosite)
-#   if(dosite=="Idaho"){
-#     tmp_data <- subset(tmp_data, species!="ARTR")
-#   }
-  within_site_list <- list()
-  for(dosim in sims){
-    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
-    tmpsynch <- get_ipm_synchrony(tmpsim)
-    tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
-    within_site_list[[dosim]] <- tmp_pgr_synch
-  }
-  pgr_list[[dosite]] <- within_site_list
-}
-pgr_synch <- melt(pgr_list)
-expected_synch <- table_data[,c(1,3)]
-expected_synch$L2 <- "E(SYNCH)"
-colnames(expected_synch) <- c("L1", "value", "L2")
-expected_synch <- expected_synch[,c("value", "L2", "L1")]
-expected_synch[which(expected_synch[,"L1"]=="New Mexico"),"L1"] <- "NewMexico"
+\pagebreak{}
 
-pgr_synch <- rbind(pgr_synch, expected_synch)
-
-# ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
-#   geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
-#   facet_wrap("L1", scales="free_x")+
-#   scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
-#   guides(fill=FALSE)+
-#   scale_y_continuous(limits=c(0,1))+
-#   xlab("Source of stochasticity")+
-#   ylab("Species synchrony")
-
-pgrsynch_plot <- ggplot(pgr_synch, aes(x=L2, y=value, color=L1, group=L1))+
-  geom_line()+
-  geom_point(size=5)+
-  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
-                     name = "Site")+
-  scale_y_continuous(limits=c(0,1))+
-  xlab("Simulation")+
-  ylab("Species synchrony")+
-  theme_bw()
-
-print(pgrsynch_plot)
+```{r figure_2, echo=FALSE, fig.cap="Community synchrony of per capita growth rates (points, solid lines) and percent cover (triangles, dashed lines) from IPM simulations with (INTER) and without (NoINTER) interspecific interactions.
+'F' in x-axis labels refers to a fluctuating environment.
+Errorbars represent the upper and lower 90th quantiles of synchrony from 100 random contiguous 50 year sections from a 2,000 iteration IPM run.
+Points represent the mean synchrony across the 100 sections.", eval=T}
+library(png)
+library(grid)
+img <- readPNG("./components/ipm_pgr_plot.png")
+grid.raster(img)
 ```
 
-```{r ibm_sims_figure, echo=FALSE, fig.cap="Species synchrony from spatially-explicit IBM simulations with (A,C) and without (B,D) environmental fluctuations and with (A,B) and without (C,D) interspecific interactions.
+\pagebreak{}
+
+```{r ibm_sims_figure, echo=FALSE, fig.cap="Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
 Points represent replicate IBM simulations (out of 20 possible) in which no species went extinct.
 Lines show linear fits between plot size (proxy for population size) and species synchrony.
-Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).", fig.height=6}
+Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).
+Simulation type codes: C = constant environment, F = fluctuating environment, INTER = interspecific interactions included, NoINTER = interspecific interactions excluded (set to zero).", eval=T}
 library(png)
@@
 ```
+
+\pagebreak{}
 

### 2015-11-03 devel/plot_ibm_sims.R

@@ -7,7 +7,7 @@
 # Clear workspace 
 rm(list=ls(all=TRUE))
 
-totSims <- 20      # number of simulations per site (1 here since using large landscape)
+totSims <- 20      # number of simulations per site 
 totT <- 100      # time steps of simulation
 burn.in <- 25    # time steps to discard before calculating cover values
 site_colors <- c("grey45", "steelblue", "slateblue4", "darkorange", "purple")
@@ -51,7 +51,7 @@ rm(stringlist)
 
 
 output_df <- data.frame(experiment=NA, site=NA, expansion=NA, 
-                        run=NA, synchrony=NA)
+                        run=NA, pgr_synch=NA, abund_synch=NA)
 
 for(do_exp in unique(files_df$experiment)){
   exp_files <- files_df[which(files_df$experiment==do_exp),"filename"]
@@ -121,17 +121,16 @@ for(do_exp in unique(files_df$experiment)){
             obs_gr[ii,] <- as.numeric(log(mergedts[ii,name_ids1]/mergedts[ii,name_ids2]))
           }
           
-          synch.run[count] <- as.numeric(community.sync(obs_gr)[1])
+          colids <- which(colnames(tmpsynch) %in% paste0("Cov.",species_names))
+          synch.abund <- as.numeric(community.sync(tmpsynch[,colids])[1])
+          synch.pgr <- as.numeric(community.sync(obs_gr)[1])
+          tmpD <- data.frame(experiment=do_exp, site=do_site, expansion=i, 
+                             run=count, pgr_synch=synch.pgr, 
+                             abund_synch=synch.abund)
+          output_df <- rbind(output_df, tmpD)
           count <- count+1
         } # end loop over coexistence runs
         
-        # Create tmp dataframe for rbinding to storage df
-        tmpD <- data.frame(synchrony=synch.run) # synchrony of coexistence runs
-        tmpD$experiment <- do_exp
-        tmpD$site <- do_site # current site
-        tmpD$expansion <- i # current plot size
-        tmpD$run <- c(1:length(runs)) # index for simulation run
-        output_df <- rbind(output_df, tmpD[,c("experiment","site", "expansion", "run", "synchrony")])
       } # end if/then for coexistence runs
       
     }# end plot size sim loop
[…]

### 2015-11-10 docs/ms_communitySynchrony.Rmd
updates ms with intro text

@@
 ------------
+Biodiversity tends to confer ecosystem stability in the face of environmental fluctuations [@Hector2010; @DeMazancourt2013a; @Hautier2014].
+In the absence of a highly dominant and stable species, asynchronous population dynamics (also called compensatory dynamics) among interacting species is the only way for biodiversity to stabilize aggregate community or ecosystem function [@Loreau2013].
+Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@DeMazancourt2013a; @Gross2014; @Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
+Futhermore, analyses of observational time series tend to show synchronous, rather than asynchrounous, population dynamics among species of similar ecology (e.g., compete for and use similar resources) [@Houlahan2007; @Rocha2011; @Vasseur2014].
+Thus, our understanding of the prevelance and drivers of asynchronous population dynamics is murky at best.
+Only by fully understanding the mechanisms that lead to asynchronous, or synchronous, in natural ecosystems can we predict how future biodiversity loss may impact ecosystem functioning [@Loreau2010].
+
 All populations fluctuate through time as a result of density-dependence (inter and intraspecific interactions), demographic stochasticity, and fluctuating environmental forces [@Lande2003; @Loreau2008a; @Mutshinda2009; @Gonzalez2009].
-Population fluctuations play integral roles at the community and ecosystem levels by creating opportunities for coexistence [@Chesson2000a] and by stabilizing aggregate properties like annual biomass production [@Hector2010; @Loreau2013; @DeMazancourt2013a; @Hautier2014].
-In both cases (species coexistence and ecosystem stability), temporal niche differentiation driven by species-specific responses to environmental conditions is necessary to create asynchrounous population dynamics.
-However, detecting and interpreting species asynchrony in natural systems has been plagued by methodological issues, and few have attempted to decompose the drivers of species synchrony.
-Thus, we lack general answers to two fundamental questions that have fascinated ecologists for decades: (1) How (a)synchronous are interacting populations in natural ecosystems?
-and (2) What drives temporal (a)synchrnous dynamics in the real world?
-
-The seminal study by @Houlahan2007 suggested that asynchronous population dynamics are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
+The seminal study by @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
 The conclusion of @Houlahan2007, that abiotic factors synchronize species' abundances, rests on the implicit assumption that species will fluctuate independently in the absence of temporal niche differences.

### 2015-11-11 docs/appendix_stats.Rmd
updates ms intro, methods, and results

--- 2015-11-03 docs/ms_communitySynchrony.Rmd
@@
 ---
-layout: 12pt
-header-includes:
-   - \usepackage{lineno}
-   - \linenumbers
-   - \usepackage{setspace}
-   - \doublespacing
-   - \usepackage[document]{ragged2e}
-bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
-csl: components/ecology.csl
+date: "November 10, 2015"
+output: pdf_document
+---
 
-## rmarkdown render options
-output:
-  pdf_document:
-    fig_caption: true
-    keep_tex: false
-fontsize: 12pt
-geometry: margin=1in
-
----
-```{r caching, include=FALSE}
-library("methods")
-library("knitr")
-basename <- "ms_communitySynchrony"
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 2)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
-               comment = NA, verbose = TRUE, echo=FALSE)
-
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-```
-
-# Environmental responses synchronize population dynamics in five semi-arid grasslands, but not completely
-
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
-
-Andrew T.
-Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, \newline Stephen Ellner\textsuperscript{3}, and Peter B.
-Adler\textsuperscript{1}
-
-\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}
-
-\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}
-
-\textit{\small{\textsuperscript{3}Department of Ecology and Evolutionary Biology, Cornell University, Ithaca, NY 14853, USA}}
-
-\renewcommand*{\thefootnote}{\arabic{footnote}}
-\setcounter{footnote}{0}
-
-\newpage{}
-
-Abstract
--------------
-Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time. 
-While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
-Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence sychronizes dynamics, demographic stochasticity desynchronizes dynamics, and environmental forcing can either synchronize or desynchronize poplation dynamics depending on the correlation of species-specific environmental responses.
-Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony through time. 
-We find that the trophically-similar grass species in our dataset have population dynamics that are synchronized by similar responses to environmental fluctuations. 
-Species interactions tend to be weak, so removing interspecific competition did not reduce synchrony. 
-The effect of demographic stochasticity depended on the spatial scale of our modeling; demographic stochasticity was strongest in small populations and was able to modestly counteract the synchronizing effect of the environment.
-Although we found that similar responses to environmental fluctuations synchronized species, dynamics were not completly synchronous. 
-We conclude that niche differentiation among species is strong enough to reduce the synchronizing effect of coupled density dependence via competition, but this is countered by the synchronizing effect of similar environmental responses.
-The next frontier is to identify the specific environmental drivers of species synchrony to inform predictive efforts.
-
-__Keywords__: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
-
-<!-- \setlength{\parindent}{5ex} -->
-
-Introduction
-------------
-All populations fluctuate through time as a result of density-dependence (inter and intraspecific interactions), demographic stochasticity, and fluctuating environmental forces [@Lande2003; @Loreau2008a; @Mutshinda2009; @Gonzalez2009].
-Population fluctuations play integral roles at the community and ecosystem levels by creating opportunities for coexistence [@Chesson2000a] and by stabilizing aggregate properties like annual biomass production [@Hector2010; @Loreau2013; @DeMazancourt2013a; @Hautier2014].
-In both cases (species coexistence and ecosystem stability), temporal niche differentiation driven by species-specific responses to environmental conditions is necessary to create asynchrounous population dynamics.
-However, detecting and interpreting species asynchrony in natural systems has been plagued by methodological issues, and few have attempted to decompose the drivers of species synchrony.
-Thus, we lack general answers to two fundamental questions that have fascinated ecologists for decades: (1) How (a)synchronous are interacting populations in natural ecosystems?
-and (2) What drives temporal (a)synchrnous dynamics in the real world?
-
-The seminal study by @Houlahan2007 suggested that asynchronous population dynamics are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
-The conclusion of @Houlahan2007, that abiotic factors synchronize species' abundances, rests on the implicit assumption that species will fluctuate independently in the absence of temporal niche differences.
-Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
-However, the null hypothesis of independent fluctuations is the incorrect starting point from which to judge spceies synchrony.
-
-All organisms exist in variable environments, and those that coexist on similar resources should be expected to respond to the environment in similar ways.
-Based on such a proposition, @Loreau2008a suggest that the true null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
-Then, the judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
-In such a light, Houlahan et al.'s (2007) results can be viewed very differently.
-However, Houlahan et al.'s (2007) meausure of compensatory dynamics was summed species covariances through time.
-Summed covariances have no upper or lower bound, making it difficult to interpret their results based on deviations from perfect synchrony.
-In the end, the fact that few communities exhibited negative summed covariance yields little insight as to whether communities tend to be sychronized by environmental forces or not.
-
-Aside from issues relating to the null expectation of species synchrony, disentangling the factors that drive species syncrony is difficult.
-Observational timeseries of species' abundances allows easy quantification of summed covariances [@Houlahan2007] or species synchrony [@Gross2014], but quantifying the relative influence of density-dependence, demographic stochasticity, and environmental stochasticity remains a challenge.
-Indeed, to our knowledge, only one other study has attempted to separate the influences of density-dependence and environmental stochasticity [@Mutshinda2009].
-@Mutshinda2009 used a statistical technique to decompose synchrony of species' abundances calculate from five time series.
-They found that the effect of environmental forcing on temporal dynamics is much stronger than that of interspecific interactions, across all taxa they analyzed [@Mutshinda2009].
-However, to link environmental forcing to species synchrony, @Mutshinda2009 examined the correlations among species' environmental responses.
-
-In this paper, we seek to overcome the limitations of previous studies by explicitly testing theory on all the drivers of species synchrony using empirically-based multi-species population models that include the effects of density-dependence, environmental stochasticity, and demographic stochasticity.
-Our theoretical framework is that of @Loreau2008a, which defines expectations for the effects of density-dependence, environmental stochasticity, and demographic stochasticity on species synchrony.
-We confront theoretical expectations with model-based analyses of community time series from five semi-arid grasslands [@Chu2015].
-Our objectives were to (1) reproduce observed species synchrony using multi-species population models, (2) determine the relative influences of environmental stochasticity and interspecific interactions on species synchrony in large populations immune to demographic stochasticity, and (3) determine how dempographic stochasticity augments the influence of the environment and interspecific interactions across a range of populations sizes.
-To achieve our objectives, we analyzed spatially-implicit integral projection models (IPMs) and spatially-explicit individually-based models (IBMs) fit to demographic data from each our study sites.
-We first show that the models are able to reproduce observed synchrony of species' per capita growth rates and then perturb the models to include/exclude the factors to drive synchrony, as identified by theory.
-We used the IPM to model large populations that are immune to demographic stochasticity to isolate the influence of environmental stochasticity; we used the IBM to model populations where demographic stochasticity is present.
-
-Theoretical expectations
-------------------------
-@Loreau2008a derived a statistic to measure community-wide synchrony of per capita population growth rates:
-
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
-\end{equation}
-
-where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
-The statistic is bounded at 0 and 1, where 0 indicates perfect asynchrony and 1 indicates perfect synchrony.
-When applied to a neutral model where species have perfect niche overlap but have independent demographic and environmental fluctuations through time,  expected synchrony ($\phi_r$) is defined as:
-
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(SN)}{[(1/S)(\sum_{i}\sqrt{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(N)})]^2}
-\end{equation}
-
-where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
-*S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
-
-Loreau and de Mazancourt show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx 1$ meaning species are expected to fluctuate synchronously.
-At the other limit, where environmental forcing and community regulation are strong relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$ or:
-
-\begin{equation}
-\phi_{r} = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.
-\end{equation}
-
-Based on the above theory, we expect synchrony in a community where only demographic stochasticity is operating on top of independent fluctuations to be $1/S$.
-
-Data and Methods
-----------------
-## Study sites and species
-In the early 20th century, several range experiment stations in the western U.S. began mapping permanent quadrats, and conducted annual censuses for decades. 
-We use such long-term demographic data sets from five semiarid grasslands in the western United States (described in detail by Chu and Adler 2015).
-Each study site includes a set of 1-$\text{m}^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
-The resulting mapped polygons represent basal cover for grasses and canopy cover for shrubs.
-Data come from the Sonoran desert in Arizona (Anderson et al.
-2012), sagebrush steppe in Idaho (Zachmann et al.
-2010), southern mixed prairie in Kansas (Adler et al.
-2007), northern mixed prairie in Montana (Anderson et al.
-2011), and Chihuahuan desert in New Mexico (Anderson et al.
-in preparation, Chu and Adler 2015).
-
-Chu and Adler describe the selection of focal species in each data set, so we offer a brief explanation here.
-Focal species needed to comply with two criteria: (1) be common enough to provide a large sample size for model fitting and (2) co-occur with other focal species often enough to allow us to describe species interactions within the model.
-Thus, Chu and Adler (2015) chose species that occurred in at least 20% of years and then used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
-
-## Analytical approach
-While the observational data allows us to directly calculate species synchrony within the community, it is impossible to isolate the contributions of demographic and environmental stochasticity on species synchrony based on the data alone.
-So we now turn to a empirically-based modeling approach where we can turn species' interactions and demographic and environmental stochasticity on/off. 
-We use two models: (1) an environmentally stochastic Integral Projection Model (IPM) and (2) a demographically and environmentally stochastic Individual Based Model (IBM).
-Both models are driven by the same vital rate regressions for survival, growth, and recreuitment.
-We use the IPM to explore the drivers of synchrony assuming a large population where demographic stochasticity is weak or absent.
-We use the IBM to isolate the effect of demographic stochasticity and to examine how demographic and environmental stochasticity interact as population size increases (i.e., as the effect of demographic stochasticity weakens).
-The use of two model types (IPM and IBM) is the only way to isolate the effects of environmental stochasticity (an IPM with no demographic stochasticity) and demographic stochasticity (an IBM with no environmental stochasticity).
+# Appendix X
 
@@
 Following previous work, we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM.
-
-### Model simulations
-Our interest is in the (de)synchronizing effects of either demographic or environmental stochasticity.
-To look at these effects, we run three types of model simulations: (1) demographic stochasticity only, (2) environmental stochasticity only, and (3) both.
-Environmental stochasticity is implemented by drawing random year effects for the vital rate regressions.
-We can model a constant environment by simply using vital rate regressions with mean coefficient values.
-To include demographic stochasticity we treat survival as a binomial process and recruitment as a poisson process in the IPM[^stoch].
-See Figure 2 for example model runs.
-
-[^stoch]: This is the crux of the matter here, we need to implement the Poisson approximation at the iteration matrix stage.
-
-Results
--------
-## Observed and expected synchrony
-Our first task was to calculate the observed and expected community synchrony at each of our study sites.
-Observed synchrony ($\phi_r$) is calculate using equation 1 and expected synchrony ($\text{E}(\phi_r)$) is simply $1/S$ or equation 3.
-Expected synchrony under indpendent fluctuations is lower than the observed synchrony at every site (Table 1).
-This implies that demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$).
-Thus, some combination of density-dependence (species interactions) and environmental forcing are synchronizing population dynamics.
-To isolate those effects, we turn to our simulation results from the IPMs.
-
-## Effect of removing demographic and environmental stochasticity
-At four of our five study areas, the presence of environmental stochasticity via random year effects on vital rate regressions increased community synchrony of per capita population growth rates (Figure 1).
-Only at the Montana site (Northern Mixed Prairie) did incorporating species' responses to environmental conditions decrease community synchrony.
-However, even with similar responses to environmental conditions, community synchrony did not reach 1 at any of the sites.
-Thus, there remains sufficiently strong temporal niche differences to drive the community away from perfect sychrony.
-
-Theory predicts that demographic stochasticty is the only force that, in isolation, can desynchronize population dynamics.
-However, theory also predicts that demographic stochasticity is likely unable to counteract the synchronizing effect of environmental forces.
-Our results support this prediction in the cases where environmental stochasticity is a synchronizing force (compare ENV and ENVDEMO bars in Figure 1).
-
-\newpage{}
-
-Tables
-------
-
-\singlespacing
-
-Table 1: Site descriptions and focal species.
-\footnotesize
-
-| Site Name | Biome | Location (Lat, Lon) | Obs. Years | Species |
-| --------- | ----- | ------------------- | ---------- | ------- |
-| Arizona | Sonoran Desert | 31&deg;50' N, 110&deg;53' W | 1915-1933 | *Bouteloua eriopoda* |
-| | | | | *Bouteloua rothrockii* |
-| Idaho | Sagebrush steppe | 44.2&deg; N, 112.1&deg; W | 1926-1957 | *Artemisia tripartita* |
-| | | | | *Pseudoroegneria spicata* |
-| | | | | *Hesperostipa comata* |
-| | | | | *Poa secunda* |
-| Kansas | Southern mixed prairie | 38.8&deg; N, 99.3&deg; W | 1932-1972 | *Bouteloua curtipendula* |
-| | | | | *Bouteloua hirsuta* |
-| | | | | *Schizachyrium scoparium* |
-| Montana | Northern mixed prairie | 46&deg;19' N, 105&deg;48' W | 1926-1957 | *Bouteloua gracilis* |
-| | | | | *Hesperostipa comata* |
-| | | | | *Pascopyrum smithii* |
-| | | | | *Poa secunda* |
-| New Mexico | Chihuahuan Desert | 32.62&deg; N, 106.67&deg; W | 1915-1950 | *Bouteloua eriopoda* |
-| | | | | *Sporobolus flexuosus* |
-
-\normalsize
-
-\pagebreak{}
-
-```{r load_libs, echo=FALSE, message=FALSE, warning=FALSE}
-library(communitySynchrony)
-library(plyr)
-library(reshape2)
-library(synchrony)
-library(ggplot2)
-library(xtable)
-library(gridExtra)
-library(RColorBrewer)
-```
-
-```{r get_all_data, echo=FALSE}
-site <- "Kansas"
-spp_list <- c("BOCU","BOHI","SCSC")
-num_spp <- length(spp_list)
-ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  ks_data <- rbind(ks_data, spp_data)
-} #end species looping for raw data
-ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
-
-tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
-tmp2<-which(ks_data$quad_data=="q27")
-tmp3<-which(ks_data$quad=="q28")
-tmp4<-which(ks_data$quad=="q30")
-tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
-tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
-tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
-ks_data<-ks_data[-tmp,]
-
-# exclude the records later than 1968, to keep the same random year effect...
-ks_data<-subset(ks_data,year<68)
-
-site <- "Idaho"
-spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
-num_spp <- length(spp_list)
-id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  id_data <- rbind(id_data, spp_data)
-} #end species looping for raw data
-id_data <- id_data[2:nrow(id_data),] #remove first NA row
-# id_data <- subset(id_data, species!="ARTR") #take out the shrub
-
-site <- "Montana"
-spp_list <- sort(c("BOGR","HECO","PASM","POSE"))
-num_spp <- length(spp_list)
-mt_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  mt_data <- rbind(mt_data, spp_data)
-} #end species looping for raw data
-mt_data <- mt_data[2:nrow(mt_data),] #remove first NA row
-
-site <- "NewMexico"
-spp_list <- sort(c("BOER","SPFL"))
-num_spp <- length(spp_list)
-nm_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  nm_data <- rbind(nm_data, spp_data)
-} #end species looping for raw data
-nm_data <- nm_data[2:nrow(nm_data),] #remove first NA row
-
-site <- "Arizona"
-spp_list <- sort(c("BOER","BORO"))
-num_spp <- length(spp_list)
-az_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  az_data <- rbind(az_data, spp_data)
-} #end species looping for raw data
-az_data <- az_data[2:nrow(az_data),] #remove first NA row
-```
-
-```{r get_all_synch, echo=FALSE, include=TRUE}
-out_ks <- get_comm_synchrony(ts_data = ks_data)
-out_id <- get_comm_synchrony(ts_data = id_data)
-out_mt <- get_comm_synchrony(ts_data = mt_data)
-out_nm <- get_comm_synchrony(ts_data = nm_data)
-out_az <- get_comm_synchrony(ts_data = az_data)
-```
-
-```{r synchrony_table, echo=FALSE, include=TRUE, results='asis', message=FALSE}
-get_table_metrics <- function(output){
-  tmp <- output
-  exp_tmp <- round(tmp$pgr_expected_synch_ind_flucts, 2)
-  obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
-  demo_tmp <- exp_tmp - obs_tmp
-  return(c(obs_tmp, exp_tmp, demo_tmp))
-}
-
-table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
-                                  get_table_metrics(out_id),
-                                  get_table_metrics(out_mt),
-                                  get_table_metrics(out_nm),
-                                  get_table_metrics(out_az)))
-table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
-table_data <- table_data[,c("Site", "V1", "V2", "V3")]
-table_data$spp <- c(length(unique(ks_data$species)),
-                    length(unique(id_data$species)),
-                    length(unique(mt_data$species)),
-                    length(unique(nm_data$species)),
-                    length(unique(az_data$species)))
-colnames(table_data) <- c("Site",
-                          "$\\phi_{r}$", 
-                          "$\\text{E}(\\phi_{r})$", 
-                          "$\\text{E}(\\sigma_{d})$",
-                          "Species richness")
-
-synch_cap <- "Observed and expected synchrony among species' per capita growth rates and expected effect of demographic stochasticity on observed synchrony.
-Species richness values reflect the number of species analyzed from the community, not the actual richness."
-print(xtable(table_data, caption = synch_cap),
-      caption.placement="top",
-      include.rownames = F, 
-      sanitize.colnames.function = identity,
-      comment=FALSE)
-```
-
-```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site.", eval=FALSE}
-# out_ks$percent_cover$site <- "Kansas"
-# out_id$percent_cover$site <- "Idaho"
-# out_mt$percent_cover$site <- "Montana"
-# out_nm$percent_cover$site <- "New Mexico"
-# out_az$percent_cover$site <- "Arizona"
-# cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
-#                     out_mt$percent_cover, out_nm$percent_cover, 
-#                     out_az$percent_cover)
-# 
-# out_ks$growth_rates$site <- "Kansas"
-# out_id$growth_rates$site <- "Idaho"
-# out_mt$growth_rates$site <- "Montana"
-# out_nm$growth_rates$site <- "New Mexico"
-# out_az$growth_rates$site <- "Arizona"
-# pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
-#                   out_mt$growth_rates, out_nm$growth_rates, 
-#                   out_az$growth_rates)
-# all_spp <- unique(pgr_data$species)
-# color_match <- data.frame(species = all_spp,
-#                           color = brewer.pal(11,"Paired"))
-# pgr_data <- merge(pgr_data, color_match, by = "species")
-# cover_data <- merge(cover_data, color_match, by="species")
-# 
-# cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
-#                 geom_line()+
-#                 geom_point()+
-#                 facet_grid(site~., scales = "free_y")+
-#                 guides(color=FALSE)+
-#                 ylab("Mean Percent Cover")+
-#                 xlab("Year")+
-#                 scale_color_manual(values = brewer.pal(11,"Paired"))
-# 
-# pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
-#               geom_line()+
-#               geom_point()+
-#               facet_grid(site~., scales = "free_y")+
-#               guides(color=FALSE)+
-#               ylab("Per Capita Growth Rate")+
-#               xlab("Year")+
-#               scale_color_manual(values = brewer.pal(11,"Paired"))
-# full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
-```
-
-\pagebreak{}
-
-Figures
--------
-```{r concept_fig, echo=FALSE, fig.cap="Conceptual figure outlining the effects of interspecific interactions and environmental stochasticity in large populations where demographic stochasticity is effectively absent (A) and in populations of varying size where demographic stochasticity is present (B).", eval=F}
-library(png)
-library(grid)
-img <- readPNG("./components/synchrony_conceptual_fig.png")
-grid.raster(img)
-```
-
-```{r read_data, echo=FALSE, eval=T}
-output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
-mlist <- melt(output_list)
-```
-
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.
-Density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.", fig.width=6, fig.height=6, fig.width=3, eval=T}
-nyrs <- c(max(out_az$stability$num_observations),
-          max(out_id$stability$num_observations),
-          max(out_ks$stability$num_observations),
-          max(out_mt$stability$num_observations),
-          max(out_nm$stability$num_observations))
-
-boots <- 100
-sites <- sort(table_data$Site)
-sites[5] <- "NewMexico"
-synch <- matrix(NA, boots, length(sites))
-
-for(do_site in 1:length(sites)){
-  full_ipm_outs <- subset(mlist, L2=="ENVINTER")
-  site_ipm <- subset(full_ipm_outs, L1==sites[do_site])
-  colnames(site_ipm) <- c("year", "species", "cover", "L1", "L2")
-  # if(sites[do_site]=="Idaho") site_ipm <- subset(site_ipm, Var2!="ARTR")
-  species_list <- unique(site_ipm$species)
-  num_spp <- length(species_list)
-  for(i in 1:boots){
-    begin_year <- sample(x = 1:(max(site_ipm$year)-nyrs[do_site]), 1)
-    end_year <- begin_year+nyrs[do_site]
-    tmp <- subset(site_ipm, year %in% begin_year:end_year)
-    tmp2 <- dcast(tmp, year~species, value.var = "cover")
-
-    # caclulate observed growth rates
-    # create lagged data frame to only get observed yearly transitions
-    lag_df <- tmp2
-    lag_df$lagyear <- lag_df$year+1
-    colnames(lag_df)[2:(num_spp+1)] <- paste0(colnames(lag_df)[2:(num_spp+1)],"_t0")
-
-    # merge the lag df with observed
-    rm_col <- which(colnames(lag_df)=="year")
-    merged_df <- merge(tmp2, lag_df[,-rm_col], by.x = "year", by.y="lagyear")
-    transitions <- nrow(merged_df)
-    obs_gr <- matrix(nrow=transitions, ncol=num_spp)
-    name_ids1 <- which(colnames(merged_df) %in% species_list)
-    name_ids2 <- which(colnames(merged_df) %in% paste(species_list, "_t0", sep=""))
-    for(ii in 1:transitions){
-      obs_gr[ii,] <- as.numeric(log(merged_df[ii,name_ids1]/merged_df[ii,name_ids2]))
-    }
-    synch[i,do_site] <- as.numeric(community.sync(obs_gr)[1])
-  }
-}
-
-colnames(synch) <- sites
-synch_df <- melt(synch)
-
-obs_pgr_df <- c(as.numeric(out_az$pgr_synchrony[1]),
-                as.numeric(out_id$pgr_synchrony[1]),
-                as.numeric(out_ks$pgr_synchrony[1]),
-                as.numeric(out_mt$pgr_synchrony[1]),
-                as.numeric(out_nm$pgr_synchrony[1]))
-obs_pgr_df <- data.frame(pgr_synch=obs_pgr_df,
-                         Var2=sites)
-
-ggplot()+
-  geom_line(data=synch_df, aes(x=value, color=Var2), stat="density", adjust=2)+
-  geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
-  facet_wrap("Var2", ncol=1)+
-  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
-  guides(color=FALSE)+
-  theme_bw()
-```
-
-\pagebreak{}
-
-```{r figure_2, echo=FALSE, fig.cap="Community synchrony of per capita growth rates (points, solid lines) and percent cover (triangles, dashed lines) from IPM simulations with (INTER) and without (NoINTER) interspecific interactions.
-'F' in x-axis labels refers to a fluctuating environment.
-Errorbars represent the upper and lower 90th quantiles of synchrony from 100 random contiguous 50 year sections from a 2,000 iteration IPM run.
-Points represent the mean synchrony across the 100 sections.", eval=T}
-library(png)
-library(grid)
-img <- readPNG("./components/ipm_pgr_plot.png")
-grid.raster(img)
-```
-
-\pagebreak{}
-
-```{r ibm_sims_figure, echo=FALSE, fig.cap="Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
-Points represent replicate IBM simulations (out of 20 possible) in which no species went extinct.
-Lines show linear fits between plot size (proxy for population size) and species synchrony.
-Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).
-Simulation type codes: C = constant environment, F = fluctuating environment, INTER = interspecific interactions included, NoINTER = interspecific interactions excluded (set to zero).", eval=T}
-library(png)
-library(grid)
-img <- readPNG("./components/ibm_sims_fig.png")
-grid.raster(img)
-```
-
-\pagebreak{}
-
-```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations.", eval=F}
-ggplot(mlist, aes(x=year, y=(cover*100), color=species))+
-  geom_line()+
-  facet_grid(L1~L2, scale="free")
-```
-
-\pagebreak{}
-
-References
-----------

### 2015-11-11 docs/ms_communitySynchrony.Rmd
updates ms intro, methods, and results

@@
    - \usepackage{setspace}
+   - \usepackage{todonotes}
    - \doublespacing
@@
 
+\vspace{5em}
+
+*Authorship*: ATT, CdM, and PBA designed the research, ATT performed the research, and ATT, CdM, ML, and PBA contributed to the analyses and interpretation of results.
+All authors contributed to writing the manuscript.
+
 \renewcommand*{\thefootnote}{\arabic{footnote}}
@@
 ------------
-Biodiversity tends to confer ecosystem stability in the face of environmental fluctuations [@Hector2010; @DeMazancourt2013a; @Hautier2014].
-In the absence of a highly dominant and stable species, asynchronous population dynamics (also called compensatory dynamics) among interacting species is the only way for biodiversity to stabilize aggregate community or ecosystem function [@Loreau2013].
-Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@DeMazancourt2013a; @Gross2014; @Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
-Futhermore, analyses of observational time series tend to show synchronous, rather than asynchrounous, population dynamics among species of similar ecology (e.g., compete for and use similar resources) [@Houlahan2007; @Rocha2011; @Vasseur2014].
-Thus, our understanding of the prevelance and drivers of asynchronous population dynamics is murky at best.
-Only by fully understanding the mechanisms that lead to asynchronous, or synchronous, in natural ecosystems can we predict how future biodiversity loss may impact ecosystem functioning [@Loreau2010].
+Biodiversity tends to stabilize ecosystem function in the face of environmental fluctuations [@Hector2010; @DeMazancourt2013a; @Hautier2014].
+In the absence of a highly dominant and stable species, asynchrony in population dynamics (also called compensatory dynamics) among interacting species is the only way for biodiversity to stabilize aggregate community or ecosystem function [@Ives2002; @Loreau2010; @Loreau2013].
+Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010 ;@DeMazancourt2013a; @Gross2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
+Futhermore, analyses of observational time series tend to show synchronous, rather than asynchrounous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
+Thus, our understanding of the prevalence and drivers of asynchronous population dynamics is limited.
+Only by fully understanding the mechanisms that lead to asynchronous, or synchronous, dynamics in natural ecosystems can we predict how future biodiversity loss may impact ecosystem functioning [@Loreau2010].
 
 All populations fluctuate through time as a result of density-dependence (inter and intraspecific interactions), demographic stochasticity, and fluctuating environmental forces [@Lande2003; @Loreau2008a; @Mutshinda2009; @Gonzalez2009].
+The relative strength of those three forces determines if species fluctuate synchronously or not [@Loreau2008a].
 The seminal study by @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
-The conclusion of @Houlahan2007, that abiotic factors synchronize species' abundances, rests on the implicit assumption that species will fluctuate independently in the absence of temporal niche differences.
+Based on that result, @Houlahan2007 concluded that abiotic factors synchronize species' abundances.
+Such a conclusion rests on the implicit assumption that species will fluctuate independently in the absence of temporal niche differences and that strong interspecific interactions cause compensatory dynamics.
 Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
-However, the null hypothesis of independent fluctuations is the incorrect starting point from which to judge spceies synchrony.
+However, independent fluctations may not be the correct benchmark against which to compare observed dynamics, and interspecific competition actually synchronizes species dynamics [@Loreau2008a].
 
-All organisms exist in variable environments, and those that coexist on similar resources should be expected to respond to the environment in similar ways.
-Based on such a proposition, @Loreau2008a suggest that the true null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
+All organisms exist in variable environments, and those that coexist on similar resources should be expected to respond to the environment in similar ways [@Rocha2011].
+Based on such a proposition, @Loreau2008a suggest that a more appropriate null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
 Then, the judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
 In such a light, Houlahan et al.'s (2007) results can be viewed very differently.
-However, Houlahan et al.'s (2007) meausure of compensatory dynamics was summed species covariances through time.
+However, Houlahan et al.'s (2007) meausure of compensatory dynamics was summed covariances of species abundances through time.
 Summed covariances have no upper or lower bound, making it difficult to interpret their results based on deviations from perfect synchrony.
-In the end, the fact that few communities exhibited negative summed covariance yields little insight as to whether communities tend to be sychronized by environmental forces or not.
+Likewise, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a].
+In the end, the fact that few communities exhibited negative summed covariance yields little insight as to whether communities tend to be sychronized by environmental forces or not [@Loreau2010].
 
@@
 Observational timeseries of species' abundances allows easy quantification of summed covariances [@Houlahan2007] or species synchrony [@Gross2014], but quantifying the relative influence of density-dependence, demographic stochasticity, and environmental stochasticity remains a challenge.
-Indeed, to our knowledge, only one other study has attempted to separate the influences of density-dependence and environmental stochasticity [@Mutshinda2009].
+Indeed, to our knowledge, only one study has attempted to separate the influences of density-dependence and environmental stochasticity directly [@Mutshinda2009].
 @Mutshinda2009 used a statistical technique to decompose synchrony of species' abundances calculate from five time series.
 They found that the effect of environmental forcing on temporal dynamics is much stronger than that of interspecific interactions, across all taxa they analyzed [@Mutshinda2009].
-However, to link environmental forcing to species synchrony, @Mutshinda2009 examined the correlations among species' environmental responses.
+However, the @Mutshinda2009 study did not include data from plant communities that have, so far, received the most evidence for a strong role for interspecific interactions in driving species asynchrony [@Gross2014].
 
 In this paper, we seek to overcome the limitations of previous studies by explicitly testing theory on all the drivers of species synchrony using empirically-based multi-species population models that include the effects of density-dependence, environmental stochasticity, and demographic stochasticity.
-Our theoretical framework is that of @Loreau2008a, which defines expectations for the effects of density-dependence, environmental stochasticity, and demographic stochasticity on species synchrony.
-We confront theoretical expectations with model-based analyses of community time series from five semi-arid grasslands [@Chu2015].
-Our objectives were to (1) reproduce observed species synchrony using multi-species population models, (2) determine the relative influences of environmental stochasticity and interspecific interactions on species synchrony in large populations immune to demographic stochasticity, and (3) determine how dempographic stochasticity augments the influence of the environment and interspecific interactions across a range of populations sizes.
-To achieve our objectives, we analyzed spatially-implicit integral projection models (IPMs) and spatially-explicit individually-based models (IBMs) fit to demographic data from each our study sites.
-We first show that the models are able to reproduce observed synchrony of species' per capita growth rates and then perturb the models to include/exclude the factors to drive synchrony, as identified by theory.
-We used the IPM to model large populations that are immune to demographic stochasticity to isolate the influence of environmental stochasticity; we used the IBM to model populations where demographic stochasticity is present.
-
-Theoretical expectations
-------------------------
-@Loreau2008a derived a statistic to measure community-wide synchrony of per capita population growth rates:
+We follow the theory of @Loreau2008a, who derived a statistic to measure community-wide synchrony of per capita population growth rates:
 
@@
 
+\todo[inline, size=\small]{Claire: I think what we actually want is the non-neutral version here, is there a simple equation for that (something tells me there is not...)?}
+
 where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
@@
 
-Loreau and de Mazancourt show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx 1$ meaning species are expected to fluctuate synchronously.
-At the other limit, where environmental forcing and community regulation are strong relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$ or:
+Loreau and de Mazancourt (2008) show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx \phi_e$, where $phi_e$ is the synchrony of species' responses to the environment.
+At the other limit, where environmental forcing and community regulation are weak relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$.
 
-\begin{equation}
-\phi_{r} = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.
-\end{equation}
+Based on the above theory, we aimed to test the following four hypotheses related the synchrony of per capita growth rates:
 
-Based on the above theory, we expect synchrony in a community where only demographic stochasticity is operating on top of independent fluctuations to be $1/S$.
+1. In large populations ($\sigma^{2}_{d} \approx 0$), removing interspecific interactions should lead to lower synchrony if interspeficic interactions ($\sigma^{2}_{c}$) are strong;
+2.
+In small and medium-sized populations ($\sigma^{2}_{d} \neq 0$), removing interspecific interactions should have the same effect as hypothesized in (1), and removing environmental stochasticity should lead to lower synchrony if environmental forcing ($\sigma^{2}_{e}$) is strong;
+3.
+In small and medium-sized populations ($\sigma^{2}_{d} \neq 0$), synchrony of per capita growth rates ($\phi_r$) should equal the inverse of the number of species in the community (1/*S*) when only demographic stochasticity is present ($\sigma^{2}_{c}$ and $\sigma^{2}_{e} \approx 0$);
+4. The de-synchronizing effect of demographic stochasticity will weaken as population size increases, regardless of presence of interspecific interactions or environmental stochasticity.
 
-Data and Methods
-----------------
-## Study sites and species
+To test these hypotheses we parameterize and simulate dynamic multi-species population models fit long term demographic data from five semi-arid grasslands [@Chu2015].
+We ran simulations directly testing the above hypotheses using a spatially-implicit integral projection model (IPM, hypothesis 1) and a spatially-explicit individual-based model (IBM, hypotheses 2-4).
+Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
+Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models according to each hypothesis should reveal the processes responsible for the degree of species synchrony in our focal grassland communities.
+
+Materials and Methods
+---------------------
+### Data
 In the early 20th century, several range experiment stations in the western U.S. began mapping permanent quadrats, and conducted annual censuses for decades. 
-We use such long-term demographic data sets from five semiarid grasslands in the western United States (described in detail by Chu and Adler 2015).
+We use such long-term demographic data sets from five semiarid grasslands in the western United States (described in detail by @Chu2015).
 Each study site includes a set of 1-$\text{m}^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
 The resulting mapped polygons represent basal cover for grasses and canopy cover for shrubs.
-Data come from the Sonoran desert in Arizona (Anderson et al.
-2012), sagebrush steppe in Idaho (Zachmann et al.
-2010), southern mixed prairie in Kansas (Adler et al.
-2007), northern mixed prairie in Montana (Anderson et al.
-2011), and Chihuahuan desert in New Mexico (Anderson et al.
-in preparation, Chu and Adler 2015).
+Data come from the Sonoran desert in Arizona [@Anderson2012], sagebrush steppe in Idaho [@Zachmann2010], southern mixed prairie in Kansas [@Adler2007], northern mixed prairie in Montana [@Anderson2011], and Chihuahuan desert in New Mexico [Anderson et al.
+in preparation, @Chu2015] (Table 1).
 
-Chu and Adler describe the selection of focal species in each data set, so we offer a brief explanation here.
+Chu and Adler 2015 describe the selection of focal species in each data set, so we offer a brief explanation here.
 Focal species needed to comply with two criteria: (1) be common enough to provide a large sample size for model fitting and (2) co-occur with other focal species often enough to allow us to describe species interactions within the model.
 Thus, Chu and Adler (2015) chose species that occurred in at least 20% of years and then used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
+The location (latitude and longitude), biome, number of years in the time series, and selected species for each site are shown in Table 1.
 
-## Analytical approach
-While the observational data allows us to directly calculate species synchrony within the community, it is impossible to isolate the contributions of demographic and environmental stochasticity on species synchrony based on the data alone.
-So we now turn to a empirically-based modeling approach where we can turn species' interactions and demographic and environmental stochasticity on/off. 
-We use two models: (1) an environmentally stochastic Integral Projection Model (IPM) and (2) a demographically and environmentally stochastic Individual Based Model (IBM).
-Both models are driven by the same vital rate regressions for survival, growth, and recreuitment.
-We use the IPM to explore the drivers of synchrony assuming a large population where demographic stochasticity is weak or absent.
-We use the IBM to isolate the effect of demographic stochasticity and to examine how demographic and environmental stochasticity interact as population size increases (i.e., as the effect of demographic stochasticity weakens).
-The use of two model types (IPM and IBM) is the only way to isolate the effects of environmental stochasticity (an IPM with no demographic stochasticity) and demographic stochasticity (an IBM with no environmental stochasticity).
+### Analytical approach
+Our analysis involved five steps for each study site:
 
-## Statistical models for survival, growth, and recruitment
-### Survival and growth
-We modeled survival probability and growth on individual genets as a function of genet size, the crowding experienced by the focal genet from both heterospecific and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
-Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure).
+1. Calculate observed synchrony of per capita growth rates
+2. Fit statistical models for survival, growth, and recruitment
+3. Build dynamic, multi-species population models (IBMs and IPMs)
+4. Remove interspecific interactions from IPM simulations to isolate the effect of environmental responses
+5.
+Sequentially remove interspecific interactions, environmental fluctuations, and both (isolates effect of demographic stochasticity) from the IBM across a range of population sizes (effects strength of demographic stochasticity)
 
-We follow the approach of Chu and Adler (2015) to estimate crowding, assuming that the crowding experienced by a focal genet depends on distance to each neighbor genet and the neighbor's size, _u_:
+All our code and the data necessary to reproduce the results reported in this paper can be found on GitHub (link) as release v1.0, and we have archived that release on Dryad (link).
 
-\begin{equation}
-w_{ijm,t} = \sum_k e^{-\delta_{jm}d_{ijkm,t}^{2}}u_{km,t}.
-\end{equation}
+#### Calculating observed synchrony
+The data consists of records for individual plant size in quadrats for each year.
+To obtain estimates of percent cover for each focal species in each year, we summed the individual-level data within quadrats and then averaged percent cover, by species, over all quadrats.
+We calculated per capita growth rates as $\text{log}(x_t) - \text{log}(x_{t-1})$, where *x* is species' percent cover in year *t*.
+Using the community time series of per capita growth rates in percent cover, we calculated community synchrony using the metric of @Loreau2008a (Equation 1) in the `synchrony' package [@Gouhier2014] in R [@R2013].
 
-In the above, $w_{ijm,t}$ is the crowding that genet _i_ of species _j_ in year _t_ experiences from neighbors of species _m_.
-The spatial scale over which species _m_ neighbors exert influence on any genet of species _j_ is determined by $\delta_{jm}$.
-The function is applied for all _k_ genets of species _m_ that neighbor the focal genet at time _t_, and $d_{ijkm,t}$ is the distance between genet _i_ in species _j_ and genet _k_ in species _m_.
-When $k=m$, the effect is intraspecific crowding.
-We use regression-specific (survival and growth) $\delta$ values estimated by Chu and Adler (2015).
+#### Fitting statistical models
+Vital rate regressions are the building blocks of our dynamic models (IBM and IPM).
+We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment.
+We modeled survival probability of each genet as function of genet size, temporal variation among years, permanent spatial variation among groups of quadrats, and local neighborhood crowding from conspecific and heterospecific genets.
+In brief, the crowding experienced by a focal genet is a function of the size of and distance to all neighboring genets.
+See @Adler2010 and @Chu2015 for a thorough description of how crowding is estimated.
+Regression coefficients for the effect of crowding by each species can be considered a matrix of interaction coefficients whose diagonals represent intraspecific interations and whose off-diagonals represent interspecific interactions [@Adler2010].
+We modeled growth as the change in size of a genet from one year to the next, which depends on the same factors as the survival model.
+We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
+Genet size and crowding were treated as fixed effects; spatial (quadrat groupings) and temporal (interannual) variation were treated as random effects on the intercept.
 
-### Recruitment
-We model recruitment as a function of plot-level cover, temporal varation among years, and spatial variation among quadrat groups.
+We modeled recruitment at the quadrat scale, rather than the individual scale, because it is impossible to attribute new genets to specific parents [@Chu2015].
+Our recruitment model assumes that the number of recruits produced in each year follows a negative binomial distribution with the mean dependent on the cover of the parent species, permanent spatial variation among groups, temporal variation among years, and inter- and intraspecific interactions as a function of total species' cover in the quadrat.
+We fit the recruitment model using a hierarchical Bayesian approach implemented in JAGS [@Plummer2003] using the `rjags' package [@Plummer2014] in R.
+Again, temporal and spatial variation were treated as random effects.
 
-## Integral projection models (IPM)
-### Model structure
-We built an environmentally and demographically stochastic integral projection model (IPM), where either environmental or demographic stochasticity can be turned off.
-Our IPM follows the specification of Chu and Adler (2015) where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_.
-Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$.
-So, the density function for any size _v_ at time $t+1$ is
+#### Building dynamic multi-species models
+Once we have fit the vital rate stastical models, building the population models is straightforward.
+For the IBM, we initialize the model by randomly assigning plants spatial coordinates, sizes and species identities until each species achieves a density representative of that observed in the data.
+We then project the model forward by using the survival regression to determine whether a genet lives or dies, the growth regression to calculate changes in genet size, and the recruitment regression to add new individuals that are distributed randomly in space.
+Crowding is directly calculated at each time step since each genet is spatially referenced (just like the actual data).
+Demographic stochasticity is an inherent feature of IBMs because individuals with the very same probability of survival, for example, can have different fates.
+The effect of demographic stochasticity on population dynamics depends on population size [@Lande2003], so we can control the strength of demographic stochasticity by simulating the IBM on different sized landscapes.
+Environmental stochasticity is not an inherent feature of IBMs, but is easily included since we fit year-specific temporal random effects for each vital rate regression.
+Thus, to include temporal environmental variation, at each time step we randomly choose a set of estimated survival, growth, and recruitment parameters specific to one observation year.
+We can ignore the random year effects to exclude environmental stochasticity in the IBM.
+For all simulations, we ignore the spatial random effect associated with variation among quadrat groups, so our simulations represent an average quadrat for each site.
 
-\begin{equation}
-n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
-\end{equation}
+The IPM is similar to the IBM in that it is built using the vital rate regressions [@Rees2009; @Rees2014], but it is different in that it is spatially-implicit and does not include demographic stochasticity.
+Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step.
+Temporal variation is included in exactly the same way as for the IBM.
+For full details on the IPMs we use, see @Chu2015.
 
-where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal that describes all possible transitions from size $u$ to $v$ and $\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding experienced from all other species by a genet of size $u_j$ and species $j$.
-The integral is evaluated over all possible sizes between predefined lower (_L_) and upper (_U_) size limits that extend beyond the range of observed genet sizes.
+#### Removing interspecific interactions (IPM)
+To remove interspecific interactions, we simply set the off-diagonals of the interaction matrix for each vital rate regression to zero.
+This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are effectively growing in monoculture.
+We ran IPM simulations with interspecific interactions turned on/off for 2,000 time steps, after an initial 500 iteration burn-in period.
+This allowed species time to reach their stable size distribution.
+We then calculated the synchrony of species' percent cover and per capita growth rates (see *Calculating observed synchrony*) over 100 randomly selected contiguous 50 time step sections.
+These simulations allowed us to isolate the effects of competition and environmental fluctuations in large populations where demographic stochasticity is absent.
+Note, however, what we could not run IPM simulations without environmental fluctuations, because it is only those fluctuations that drive variability in stochastich IPMs.
+In a constant environment, IPM simulations asymptote at stable equilibriums for each species.
 
-The population kernal is defined as the joint contributions of survival (_S_), growth (_G_), and recruitment (_R_):
-
-\begin{equation}
-k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
-\end{equation}
-
-which, said plainly, means we are calculating growth (_G_) for individuals that survive (_S_) from time _t_ to _t+1_ and adding in newly recruited (_R_) individuals of an average sized one-year-old genet for the focal species.
-Our stastical model for recruitment (_R_, described below) returns the number of new recruit produced per quadrat.
-Following previous work, we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM.
-
-### Model simulations
-Our interest is in the (de)synchronizing effects of either demographic or environmental stochasticity.
-To look at these effects, we run three types of model simulations: (1) demographic stochasticity only, (2) environmental stochasticity only, and (3) both.
-Environmental stochasticity is implemented by drawing random year effects for the vital rate regressions.
-We can model a constant environment by simply using vital rate regressions with mean coefficient values.
-To include demographic stochasticity we treat survival as a binomial process and recruitment as a poisson process in the IPM[^stoch].
-See Figure 2 for example model runs.
-
-[^stoch]: This is the crux of the matter here, we need to implement the Poisson approximation at the iteration matrix stage.
+#### Removing interspecific interactions and environmental fluctuations (IBM)
+Since the IBM and IPM are driven by the same vital rate regressions, interspecific interactions were removed as described above for the IPM simulations.
+To remove the effect of a fluctuating environment, we simply removed the temporal (interannual) random effects from the regression equations.
+Thus, if we remove interspecific interactions and the temporal random effects, the only force causing population fluctutations through time is demographic stochasticity.
+We ran a series of simulations for four "experiments": (1) environmental stochasticity on and interspecific interactions present, (2) environmental stochasticity off and interspecific interactions present, (3) environmental stochasticity on and interspecific interactions absent, and (4) environmental stochasticity off and interspecific interactions absent (only demographic stochasticity).
+We ran IBM simulations for 100 time steps, and repeated the simulations 50 times for each "experiment" (interspecific interactions on/off with and without environmental stochasticity).
+From those, we retained only the simulations in which no species went extinct due to demographic stochasticity.
+Synchrony was calculated over the 100 time steps for each no extinction run within a model experiment.
 
@@
 -------
-## Observed and expected synchrony
-Our first task was to calculate the observed and expected community synchrony at each of our study sites.
-Observed synchrony ($\phi_r$) is calculate using equation 1 and expected synchrony ($\text{E}(\phi_r)$) is simply $1/S$ or equation 3.
-Expected synchrony under indpendent fluctuations is lower than the observed synchrony at every site (Table 1).
-This implies that demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$).
-Thus, some combination of density-dependence (species interactions) and environmental forcing are synchronizing population dynamics.
-To isolate those effects, we turn to our simulation results from the IPMs.
+### Observed and simulated community synchrony
+Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 (Table 2).
+Synchrony tended to be higher in communities with few species (Arizona and New Mexico).
+Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub (*Artimisia tripartita*) and three perennial grasses (Table 1).
+All sites exhibited more synchronous dynamics than expected if only demographic stochasticity was operating (E($\phi_r$) = 1/*S* in Table 2).
+The difference between observed $\phi_r$ and E($\phi_r$) represents the effect of the environment and species interactions combined .
+The combined effect of the environment and species interactions is positive in all communities (E($\phi_c + \phi_e$) in Table 2), indicating a synchronizing effect of those forces that counter-balance the desynchronizing effect of demographic stochasticity.
+For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. 1).
 
-## Effect of removing demographic and environmental stochasticity
-At four of our five study areas, the presence of environmental stochasticity via random year effects on vital rate regressions increased community synchrony of per capita population growth rates (Figure 1).
-Only at the Montana site (Northern Mixed Prairie) did incorporating species' responses to environmental conditions decrease community synchrony.
-However, even with similar responses to environmental conditions, community synchrony did not reach 1 at any of the sites.
-Thus, there remains sufficiently strong temporal niche differences to drive the community away from perfect sychrony.
+### Effect of removing interspecific interactions
+The statistical models for vital rates showed that interspecific interactions were weak in all communities (Table X1), as previously found by @Chu2015.
+Consistent with weak interspecific interactions, removing such interactions from IPM simulations resulted in no discernable changes in species synchrony of per capita growth rates or percent cover (Fig.
+2).
+This indicates that community regulation of population dynamics via coupled density-dependence ($\sigma_c$) is relatively weak in these communities.
 
-Theory predicts that demographic stochasticty is the only force that, in isolation, can desynchronize population dynamics.
-However, theory also predicts that demographic stochasticity is likely unable to counteract the synchronizing effect of environmental forces.
-Our results support this prediction in the cases where environmental stochasticity is a synchronizing force (compare ENV and ENVDEMO bars in Figure 1).
+The effect of removing interspecific interactions on synchrony in IBM simulations was also small for per capita growth rates and was not affected by population size (Fig. 3, top panels).
+The only discernable effect of interspecific interactions was in the Arizona and New Mexico communities when environmental stochasticity was also removed.
+In those cases, removing interspecific interactions caused a decrease in species synchrony of per capita growth rates (Fig 3-a,e).
+Removing interspecific interactions had idosyncratic effects of the synchrony of species' percent cover and no clear patterns emerged (Fig. 3, bottom panels)
+
+### Effect of removing environmental stochasticity
+For all five communities, removing environmental stochasticity cause synchrony of per capita growth rates to decrease, whether interspecific interactions were present or not (Fig.
+3, top panels, blue tones vs.
+orange tones).
+Without environmental stochasticity, synchrony of per capita growth rates all approximated the theoretical expectation of 1/S when only demographic stochasticity is operating (Fig. 3, top panels).
+Again, no clear pattern emerges for the effect of removing environmental stochasticity on the synchrony of percent cover (Fig. 3, bottom panels).
+
+### Effect of demographic stochasticity
+The strength of demographic stochasticity decreases with the size of the simulated landscape in our IBMs.
+Thus, in Figure 3 the strength of demographic stochasticity decreases along the x-axis.
+In the absence of environmental stochasticity, the strength of demographic stochasticity has no impact of the synchrony of per capita growth rates (Fig. 3, top panels, blue tones).
+In the presence of environmental stochasticity, synchrony of per capita growth rates is lower when demographic stochasticity is strong (Fig. 3, top panels, orange tones).
+The desynchronizing effect of demographic stochasticity is strongest in small populations (small landscapes) for all communities, but the effect is strongest in Arizona and New Mexico (Fig. 3).
+
+Discussion
+----------
+
+Ackowledgements
+---------------
+This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
+CdM and ML were supported by the TULIP Laboratory of Excellence (ANR-10-LABX-41).
+We thank the original mappers of the permanent quadrats at each site and the digitizers in the Adler lab, without whom this work would not have been possible.
 
@@
 
-```{r get_all_data, echo=FALSE}
+```{r get_all_data, echo=FALSE, eval=T}
 site <- "Kansas"
@@
 
-```{r synchrony_table, echo=FALSE, include=TRUE, results='asis', message=FALSE}
+```{r synchrony_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T}
 get_table_metrics <- function(output){
@@
   obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
-  demo_tmp <- exp_tmp - obs_tmp
+  demo_tmp <- obs_tmp - exp_tmp
   return(c(obs_tmp, exp_tmp, demo_tmp))
@@
                     length(unique(az_data$species)))
+# colnames(table_data) <- c("Site",
+#                           "$\\phi_{r}$", 
+#                           "$\\text{E}(\\phi_{r})$",
+#                           "Species richness")
 colnames(table_data) <- c("Site",
@@
                           "$\\text{E}(\\phi_{r})$", 
-                          "$\\text{E}(\\sigma_{d})$",
+                          "$\\text{E}(\\sigma_{c} + \\sigma_{e})$",
                           "Species richness")

### 2015-11-30 docs/ms_communitySynchrony.Rmd
updates intro text

@@
 
-# Environmental responses synchronize population dynamics in five semi-arid grasslands, but not completely
+# Disentangling the drivers of species synchrony in natural grasslands: environmental forcing, demographic stochasticity, and interspecific interactions
 
@@
 
-Andrew T.
-Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, \newline Stephen Ellner\textsuperscript{3}, and Peter B.
-Adler\textsuperscript{1}
+Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}
 
@@
 \textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}
-
-\textit{\small{\textsuperscript{3}Department of Ecology and Evolutionary Biology, Cornell University, Ithaca, NY 14853, USA}}
 
@@
 Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony through time. 
-We find that the trophically-similar grass species in our dataset have population dynamics that are synchronized by similar responses to environmental fluctuations. 
-Species interactions tend to be weak, so removing interspecific competition did not reduce synchrony. 
+Species' environmental responses were positively, but not perfectly correlated. 
+As a result, environmental variation caused an increase in synchrony relative to an expectation of independent species' responses, but synchrony was not perfect.
+Species interactions tend to be weak, so removing interspecific competition did not affect synchrony. 
 The effect of demographic stochasticity depended on the spatial scale of our modeling; demographic stochasticity was strongest in small populations and was able to modestly counteract the synchronizing effect of the environment.
-Although we found that similar responses to environmental fluctuations synchronized species, dynamics were not completly synchronous. 
 We conclude that niche differentiation among species is strong enough to reduce the synchronizing effect of coupled density dependence via competition, but this is countered by the synchronizing effect of similar environmental responses.
@@
 In the absence of a highly dominant and stable species, asynchrony in population dynamics (also called compensatory dynamics) among interacting species is the only way for biodiversity to stabilize aggregate community or ecosystem function [@Ives2002; @Loreau2010; @Loreau2013].
-Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010 ;@DeMazancourt2013a; @Gross2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
-Futhermore, analyses of observational time series tend to show synchronous, rather than asynchrounous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
+Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010 ;@DeMazancourt2013a; @Gross2014; Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
+Futhermore, analyses of observational time series tend to report synchronous, rather than asynchrounous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
 Thus, our understanding of the prevalence and drivers of asynchronous population dynamics is limited.
 Only by fully understanding the mechanisms that lead to asynchronous, or synchronous, dynamics in natural ecosystems can we predict how future biodiversity loss may impact ecosystem functioning [@Loreau2010].
+However, three sources of confusion have limited our understanding of community synchrony: (1) the metric of synchrony, (2) the baseline expectation of synchrony, and (3) the use of population abundances to calculate synchrony.
+Recent theoretical work has clarified all three sources of confusion [@Loreau2008a; @Loreau2010].
+
+@Loreau2008a derived a statistic to measure community-wide synchrony of per capita population growth rates:
+
+\begin{equation}
+\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
+\end{equation}
+
+where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate. 
+The metric of @Loreau2008 makes clear that synchrony and asynchrony are not binary states, contrary to their common usage in the literature.
+Rather, the synchrony of community dynamics lies along a continuum from perfect asynchrony ($\phi_r=0$) to perfect synchrony ($\phi_r=1$).
+Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony.
+Prior to Loreau and de Mazancourt's (2008) work, the synchrony metric of choice was summed community covariance [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
+However, @Ranta2008 show that the summed community covariance method cannot be used to detect asynchrony or its drivers because values of summed covariance are not exclusive to particular combinations of drivers.
+For example, contrary to prevailing belief, negative community covariance can occur in the absence of competition and positive community covariance can occur in the presence of strong competitive interactions [@Loreau2008a; @Ranta2008].
+Likewise, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a].
+In the end, the fact that few communities exhibited negative summed covariance yields little insight as to whether communities tend to be sychronized by environmental forces or not [@Loreau2010].
+The use of summed community covariance also relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008; @Loreau2010].
+
+Indeed, there is not a single baseline expectation for community synchrony.
+@Loreau2008a suggest there are two expectations: perfect synchrony in neutral communities where species have identical niches, and the inverse of the number of species in the community (1/*S*) when species fluctuate independently.
+Many previous investigations of community synchrony have implicitly assumped that species will fluctuate independently in the absence of temporal niche differences and that strong interspecific interactions cause compensatory dynamics.
+For example, the seminal study by @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
+Based on that result, @Houlahan2007 concluded that abiotic factors synchronize species' abundances.
+However, their null expectation was that species should fluctuate independently.
+Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
+But, if instead we *a priori* assume that species fluctuate in perfect synchrony [@Loreau2008a], judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
+Investigations of community synchrony should therefore focus on both endpoints (1 and 1/*S*) and seek to determine the underlying forces that drive synchrony closer to 1 or 1/*S*.
 
 All populations fluctuate through time as a result of density-dependence (inter and intraspecific interactions), demographic stochasticity, and fluctuating environmental forces [@Lande2003; @Loreau2008a; @Mutshinda2009; @Gonzalez2009].
-The relative strength of those three forces determines if species fluctuate synchronously or not [@Loreau2008a].
-The seminal study by @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
-Based on that result, @Houlahan2007 concluded that abiotic factors synchronize species' abundances.
-Such a conclusion rests on the implicit assumption that species will fluctuate independently in the absence of temporal niche differences and that strong interspecific interactions cause compensatory dynamics.
-Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
-However, independent fluctations may not be the correct benchmark against which to compare observed dynamics, and interspecific competition actually synchronizes species dynamics [@Loreau2008a].
+The relative strengths of those three forces determine the degree of synchrony in population dynamics [@Loreau2008a].
 
@@
 Based on such a proposition, @Loreau2008a suggest that a more appropriate null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
-Then, the judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
+Then, the 
 In such a light, Houlahan et al.'s (2007) results can be viewed very differently.
@@
 Summed covariances have no upper or lower bound, making it difficult to interpret their results based on deviations from perfect synchrony.
-Likewise, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a].
-In the end, the fact that few communities exhibited negative summed covariance yields little insight as to whether communities tend to be sychronized by environmental forces or not [@Loreau2010].
 
@@
 
-Based on the above theory, we aimed to test the following four hypotheses related the synchrony of per capita growth rates:
+Based on the above theory, we aimed to test the following four hypotheses and associated predictions related the synchrony of per capita growth rates:
 
-1. In large populations ($\sigma^{2}_{d} \approx 0$), removing interspecific interactions should lead to lower synchrony if interspeficic interactions ($\sigma^{2}_{c}$) are strong;
+1. Interspecific interactions synchronize population dynamics. Therefore, removing interspecific interactions will decrease synchrony.
 2.
-In small and medium-sized populations ($\sigma^{2}_{d} \neq 0$), removing interspecific interactions should have the same effect as hypothesized in (1), and removing environmental stochasticity should lead to lower synchrony if environmental forcing ($\sigma^{2}_{e}$) is strong;
+Plant species in semiarid ecosystems should respond in similar ways to environmental drivers.
+Therefore, removing environmental variation will decrease synchrony relative to situations were environmental variation is present.
 3.
-In small and medium-sized populations ($\sigma^{2}_{d} \neq 0$), synchrony of per capita growth rates ($\phi_r$) should equal the inverse of the number of species in the community (1/*S*) when only demographic stochasticity is present ($\sigma^{2}_{c}$ and $\sigma^{2}_{e} \approx 0$);
+In populations small enough for demographic stochasticity to influence dynamics (where $\sigma^{2}_{d} \neq 0$), synchrony will equal the inverse of the number of species in the community (1/*S*), in the absence of species interactions and environmental variation ($\sigma^{2}_{c}$ and $\sigma^{2}_{e} \approx 0$).
 4. The de-synchronizing effect of demographic stochasticity will weaken as population size increases, regardless of presence of interspecific interactions or environmental stochasticity.
@@
 ---------------------
+Our current work heavily relies on the data and analysis presented by @Chu2015.
+We refer interested readers to @Chu2015 for full details on the focal plant communities, choice of species, structure of vital rate statistical models, and the ability of fitted population models to reproduce observed dynamics.
+We offer a brief overview here, but note that the only difference between our vital rate models and those fit by @Chu2015 is that we allowed for positive species interactions.
+
 ### Data
@@
 in preparation, @Chu2015] (Table 1).
-
-Chu and Adler 2015 describe the selection of focal species in each data set, so we offer a brief explanation here.
-Focal species needed to comply with two criteria: (1) be common enough to provide a large sample size for model fitting and (2) co-occur with other focal species often enough to allow us to describe species interactions within the model.
-Thus, Chu and Adler (2015) chose species that occurred in at least 20% of years and then used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
-The location (latitude and longitude), biome, number of years in the time series, and selected species for each site are shown in Table 1.
 
@@
 We modeled survival probability of each genet as function of genet size, temporal variation among years, permanent spatial variation among groups of quadrats, and local neighborhood crowding from conspecific and heterospecific genets.
-In brief, the crowding experienced by a focal genet is a function of the size of and distance to all neighboring genets.
-See @Adler2010 and @Chu2015 for a thorough description of how crowding is estimated.
 Regression coefficients for the effect of crowding by each species can be considered a matrix of interaction coefficients whose diagonals represent intraspecific interations and whose off-diagonals represent interspecific interactions [@Adler2010].

### 2015-12-04 docs/ms_communitySynchrony.Rmd
updates ms intro

@@
 
-# Disentangling the drivers of species synchrony in natural grasslands: environmental forcing, demographic stochasticity, and interspecific interactions
+\begin{center}
+\large{Disentangling the drivers of species synchrony in natural grasslands: environmental forcing, demographic stochasticity, and interspecific interactions}
 
@@
 
-Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}
+\normalsize{Andrew T.
+Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B.
+Adler\textsuperscript{1}}
 
@@
 \textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}
+
+\end{center}
 
@@
 ------------
-Biodiversity tends to stabilize ecosystem function in the face of environmental fluctuations [@Hector2010; @DeMazancourt2013a; @Hautier2014].
+Biodiversity tends to stabilize ecosystem function in fluctuating environments [@Hector2010; @DeMazancourt2013a; @Hautier2014].
 In the absence of a highly dominant and stable species, asynchrony in population dynamics (also called compensatory dynamics) among interacting species is the only way for biodiversity to stabilize aggregate community or ecosystem function [@Ives2002; @Loreau2010; @Loreau2013].
-Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010 ;@DeMazancourt2013a; @Gross2014; Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
+Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010; @DeMazancourt2013a; @Gross2014; @Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
 Futhermore, analyses of observational time series tend to report synchronous, rather than asynchrounous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
@@
 Only by fully understanding the mechanisms that lead to asynchronous, or synchronous, dynamics in natural ecosystems can we predict how future biodiversity loss may impact ecosystem functioning [@Loreau2010].
-However, three sources of confusion have limited our understanding of community synchrony: (1) the metric of synchrony, (2) the baseline expectation of synchrony, and (3) the use of population abundances to calculate synchrony.
-Recent theoretical work has clarified all three sources of confusion [@Loreau2008a; @Loreau2010].
 
-@Loreau2008a derived a statistic to measure community-wide synchrony of per capita population growth rates:
+Discovering the mechanisms behind population fluctuations, and whether they are synchronous among species or not, is a long-standing goal of ecology.
+One view is that density-dependent processes like interspecific competition lead to compensatory dyanmics [e.g., @Elton1942; @Tilman1988; @Ives1995].
+An alternative view is that density-independent processes like environmental forcing drive population fluctuations and tend to synchronize dynamics of ecologically-similar species [@Strong1983; @Fischer2001].
+Clearly, these forces are not mutually exclusive, and most research has focused on determining the relative strength of biotic and abiotic forcing on the synchrony of community dynamics [e.g., @Ives1995; @Klug2000; @Houlahan2007; @Mutshinda2009].
+However, recent theory [@Loreau2008a] suggests that previous interpretations of time series have been impaired by a (1) faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects to estimate immediate responses to the environment and competition.
+Moreover, interspecific competition may actually synchronize, rather than de-synchronize, population dynamics [@Loreau2008a; @Loreau2010].
+
+How should synchrony be measured?
+Traditionally, most researchers relied on summed covariances to detect compensatory, and thus asynchronous, dynamics [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
+The idea is that if the summed covariances of species' abundances through time are positive, then abiotic forces structure dynamics since species covar positively through time in response to a shared environmental driver.
+Alternatively, negative summed covariance should indicate that competition structures community fluctuations by driving compensatory dynamics.
+Unfortunately, reliance on such a metric created the search for binary states -- synchronous or not, compensatory dynamics or not -- where a continuum exists [@Loreau2008a].
+Likewise, values of summed covariance are not exclusive to particular combinations of drivers, meaning that negative summed covariances can ocurr in the absence of competition [@Ranta2008].
+Finally, summed covariances have no upper or lower limits and are sensitive to community size, making it difficult to compare different communities [@Loreau2008a].
+In the end, the fact that few communities exhibit negative summed covariance [e.g., @Houlahan2007] yields little insight as to whether communities tend to be sychronized by environmental forces or not [@Loreau2010].
+
+To overcome these limitations, @Loreau2008a derived a new metric of species synchrony based on a multi-species population dynamic model.
+Their proposed metric, best suited to measure community-wide synchrony of per capita population growth rates, is
 
@@
 where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate. 
-The metric of @Loreau2008 makes clear that synchrony and asynchrony are not binary states, contrary to their common usage in the literature.
+The metric of @Loreau2008a makes clear that synchrony and asynchrony are not binary states, contrary to their common usage in the literature.
 Rather, the synchrony of community dynamics lies along a continuum from perfect asynchrony ($\phi_r=0$) to perfect synchrony ($\phi_r=1$).
-Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony.
-Prior to Loreau and de Mazancourt's (2008) work, the synchrony metric of choice was summed community covariance [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
-However, @Ranta2008 show that the summed community covariance method cannot be used to detect asynchrony or its drivers because values of summed covariance are not exclusive to particular combinations of drivers.
-For example, contrary to prevailing belief, negative community covariance can occur in the absence of competition and positive community covariance can occur in the presence of strong competitive interactions [@Loreau2008a; @Ranta2008].
-Likewise, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a].
-In the end, the fact that few communities exhibited negative summed covariance yields little insight as to whether communities tend to be sychronized by environmental forces or not [@Loreau2010].
-The use of summed community covariance also relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008; @Loreau2010].
+Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony within a community and what process increase or decrease sychrony.
 
-Indeed, there is not a single baseline expectation for community synchrony.
+What is the null expectation of species synchrony in competitive communities?
+Unfortunately, there is not a single answer.
+The use of summed community covariance relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008a; @Loreau2010]. 
 @Loreau2008a suggest there are two expectations: perfect synchrony in neutral communities where species have identical niches, and the inverse of the number of species in the community (1/*S*) when species fluctuate independently.
@@
 
-All populations fluctuate through time as a result of density-dependence (inter and intraspecific interactions), demographic stochasticity, and fluctuating environmental forces [@Lande2003; @Loreau2008a; @Mutshinda2009; @Gonzalez2009].
-The relative strengths of those three forces determine the degree of synchrony in population dynamics [@Loreau2008a].
+What temporal variable should we use to measure synchrony?
+In constrast to virtually every study on synchrony that uses population abundance (CITATIONS), @Loreau2008a suggest community-wide synchrony is best viewed through the lense of per capita growth rates.
+On conceptual grounds, per capita growth rates represent the immediate response of species to the environment and competition that we wish to attribute to synchrony.
+On practical grounds, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a].
+Therefore, measuring the synchrony of population abundances through time does not allow inference on the drivers of synchrony.
 
-All organisms exist in variable environments, and those that coexist on similar resources should be expected to respond to the environment in similar ways [@Rocha2011].
-Based on such a proposition, @Loreau2008a suggest that a more appropriate null hypothesis is that coexisting species should fluctuate synchronously, not independently, in the absence of temporal niche differenatiation.
-Then, the 
-In such a light, Houlahan et al.'s (2007) results can be viewed very differently.
-However, Houlahan et al.'s (2007) meausure of compensatory dynamics was summed covariances of species abundances through time.
-Summed covariances have no upper or lower bound, making it difficult to interpret their results based on deviations from perfect synchrony.
-
-Aside from issues relating to the null expectation of species synchrony, disentangling the factors that drive species syncrony is difficult.
+Aside from issues on how to measure synchrony, disentangling the factors that drive species syncrony is difficult.
 Observational timeseries of species' abundances allows easy quantification of summed covariances [@Houlahan2007] or species synchrony [@Gross2014], but quantifying the relative influence of density-dependence, demographic stochasticity, and environmental stochasticity remains a challenge.
@@
 In this paper, we seek to overcome the limitations of previous studies by explicitly testing theory on all the drivers of species synchrony using empirically-based multi-species population models that include the effects of density-dependence, environmental stochasticity, and demographic stochasticity.
-We follow the theory of @Loreau2008a, who derived a statistic to measure community-wide synchrony of per capita population growth rates:
-
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
-\end{equation}
-
-where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
-The statistic is bounded at 0 and 1, where 0 indicates perfect asynchrony and 1 indicates perfect synchrony.
-When applied to a neutral model where species have perfect niche overlap but have independent demographic and environmental fluctuations through time,  expected synchrony ($\phi_r$) is defined as:
+We follow the theory of @Loreau2008a, who show that in a neutral community of *S* species, expected synchrony ($\phi_r$) is defined as:
 
@@
 *S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
+Their analysis isolates three main factors that drive species synchrony in competitive communities: (1) density-dependence via intra- and interspecific competition, (2) environmental fluctuations, and (3) demographic stochasticity.
 
@@
 At the other limit, where environmental forcing and community regulation are weak relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$.
-
-Based on the above theory, we aimed to test the following four hypotheses and associated predictions related the synchrony of per capita growth rates:
+Based on the theory of @Loreau2008a, we aimed to test the following four hypotheses and associated predictions related to the synchrony of per capita growth rates:
 
@@
 
-```{r get_all_synch, echo=FALSE, include=TRUE}
+```{r get_all_synch, echo=FALSE, include=TRUE, eval=T}
 out_ks <- get_comm_synchrony(ts_data = ks_data)

### 2015-12-15 docs/ms_communitySynchrony.Rmd
adds HPC acknowledgment

@@
 \begin{center}
-\large{Disentangling the drivers of species synchrony in natural grasslands: environmental forcing, demographic stochasticity, and interspecific interactions}
+\large{\textbf{Disentangling the drivers of species synchrony in natural grasslands: environmental forcing, demographic stochasticity, and interspecific interactions}}
 
@@
 We thank the original mappers of the permanent quadrats at each site and the digitizers in the Adler lab, without whom this work would not have been possible.
+Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.
 

### 2015-12-16 docs/ms_communitySynchrony.Rmd
updates ms with more-runs IBM results

@@
                           "Species richness")
+table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
+table_data[,4] <- table_data[,2] - table_data[,3] # calculate obs - exp synchrony
 
@@
   guides(color=FALSE)+
+  xlab("Community Synchrony")+
+  ylab("Estimated Probability Density")+
   theme_bw()
@@
 
-```{r figure_2, echo=FALSE, fig.cap="Community synchrony of per capita growth rates (points, solid lines) and percent cover (triangles, dashed lines) from IPM simulations with (INTER) and without (NoINTER) interspecific interactions.
-'F' in x-axis labels refers to a fluctuating environment.
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/ipm_pgr_plot.png}
+  \caption{Community synchrony of per capita growth rates (points, solid lines) and percent cover (triangles, dashed lines) from IPM simulations with (INTER) and without (NoINTER) interspecific interactions.
+`F' in x-axis labels refers to a fluctuating environment.
 Errorbars represent the upper and lower 90th quantiles of synchrony from 100 random contiguous 50 year sections from a 2,000 iteration IPM run.
-Points represent the mean synchrony across the 100 sections.", eval=T}
-library(png)
-library(grid)
-img <- readPNG("./components/ipm_pgr_plot.png")
-grid.raster(img)
-```
+Points represent the mean synchrony across the 100 sections.}
+\end{figure}
 
@@
 
-```{r ibm_sims_figure, echo=FALSE, fig.cap="Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/ibm_sims_fig.png}
+  \caption{Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
 Points represent replicate IBM simulations (out of 20 possible) in which no species went extinct.
@@
 Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).
-Simulation type codes: C = constant environment, F = fluctuating environment, INTER = interspecific interactions included, NoINTER = interspecific interactions excluded (set to zero).", eval=T}
-library(png)
-library(grid)
-img <- readPNG("./components/ibm_sims_fig.png")
-grid.raster(img)
-```
+Simulation type codes: C = constant environment, F = fluctuating environment, INTER = interspecific interactions included, NoINTER = interspecific interactions excluded (set to zero).}
+\end{figure}
 

### 2015-12-16 devel/plot_ibm_sims.R

@@ -7,7 +7,7 @@
 # Clear workspace 
 rm(list=ls(all=TRUE))
 
-totSims <- 20      # number of simulations per site 
+totSims <- 50      # number of simulations per site 
 totT <- 100      # time steps of simulation
 burn.in <- 25    # time steps to discard before calculating cover values
 site_colors <- c("grey45", "steelblue", "slateblue4", "darkorange", "purple")
@@ -28,7 +28,8 @@ library(gridExtra)
 ####
 ####  Read in data and setup up file lists -------------------------------------
 ####
-setwd("../results/ibm_sims/")
+# setwd("../results/ibm_sims/")
+setwd("../../../../../Volumes/A02046115/ibm_synch/results/")
 all_files <- list.files()
 constinterfiles <- all_files[grep("constinter", all_files)]
 constnointerfiles <- all_files[grep("constnointer", all_files)]
@@ -148,8 +149,17 @@ synch_df <- melt(output_df, id.vars = c("experiment", "site", "expansion", "run"
 colnames(synch_df) <- c("experiment", "site", "expansion", "run", 
                         "typesynch", "synch")
 
+# Check for single datapoints
+allcombo <- ddply(synch_df, .(experiment, site, expansion, typesynch), summarise,
+                  nums = length(run))
+singles <- allcombo[which(allcombo$nums<10),]
+singles$remove="yes"
+test <- merge(synch_df, singles, by=c("experiment", "site", "expansion", "typesynch"), all = TRUE)
+ids.out <- which(test$remove=="yes")
+synch_dfplot <- test[-ids.out,]
+
 ### Make plot and save
-ibm_plot <- ggplot(synch_df, aes(x=expansion, y=synch, color=experiment))+
+ibm_plot <- ggplot(synch_dfplot, aes(x=expansion, y=synch, color=experiment))+
   geom_point(alpha=0.5)+
   stat_smooth(se=FALSE, method="lm", size=0.7)+
   scale_color_manual(values = c("steelblue", "slateblue4", "darkorange", "darkred"),
@@ -162,7 +172,8 @@ ibm_plot <- ggplot(synch_df, aes(x=expansion, y=synch, color=experiment))+
   facet_grid(typesynch~site)+
   theme_bw()
 
-png("../../docs/components/ibm_sims_fig.png", height = 5, 
+setwd("~/Repos/community_synchrony/devel")
+png("../docs/components/ibm_sims_fig.png", height = 5, 
     width = 10, units = "in", res=150)
 print(ibm_plot)
 dev.off()

### 2015-12-22 docs/ms_communitySynchrony.Rmd
updates fig 3 caption
adds beginning of discussion

@@
 ----------
+Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations.
+We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.
+Our study produced three main findings that were consistent across all of our focal communities: (1) interspecific interactions had little impact on community synchrony because interspecific interactions were weak; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony since species-specific responses to the environment were positively, but not completely, correlated.
+Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community synchrony approximately equals the synchrony of species' responses to the environment.
+
+### Interspecific interactions had little impact on community synchrony
+We expected community synchrony of per capita growth rates to decrease and synchrony of plant cover to increase when we removed interspecific interactions.
+In the absence of demographic stochasticity, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions for both per capita growth rates and percent cover (Fig.
+2).
+The only notable exception is the Idaho community where, as expected, synchrony of percent cover increased when we removed interspecific interactions.
+Instead, we observed no effect of removing interspecific interactions in the presence of environmental fluctuations, regardless of the presence or absence of demographic stochasticity (Figs. 2,3).
+When we also removed environmental fluctuations, removing interspecific interactions decreased synchrony of per capita growth rates in three of our five communities and increased synchrony of per capita growth rates in two communities (Fig.
+3).
+This result is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics and a recent empirical analysis
+
+### Demographic stochasticity decreases synchrony
+In communities where environmental forcing and community regulation by inter and intraspecific density-dependence are weak, species should fluctuate independently and community synchrony is expected to equal the inverse of the number of species in the community (1/*S*).
+Such independent fluctuations are expected for per capita growth rates regardless of population sizes, whereas synchrony of percent cover should trend toward one (perfect synchrony) as population sizes increase and the effect of demographic stochasticity is diminished.
+We find this to be true in our simulations: for each community, community synchrony of per capita growth rates are approximately equal to 1/*S* when only demographic stochasticity is acting (Fig.
+3; top panels, dark purple lines) and synchrony of percent cover starts at approximately 1/*S* and increases linearly as population sizes increase (Fig.
+3; bottom panels, dark purple lines).
 
@@
   \caption{Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
-Points represent replicate IBM simulations (out of 20 possible) in which no species went extinct.
+Points represent replicate IBM simulations (out of 50 possible) in which no species went extinct.
 Lines show linear fits between plot size (proxy for population size) and species synchrony.

### 2015-12-22 devel/table_spp_interactions.R

@@ -104,3 +104,19 @@ ggplot(df, aes(x=comp))+
   theme_bw()
 
 
+
+####
+####  Plot Each Vital Rate Interaction Coefficients
+####
+df <- melt(unlist(vital_comp_list))
+pieces <- strsplit(rownames(df), split = "[.]")
+df$vitalrate <- sapply(pieces, "[", 1)
+sites <- sapply(pieces, "[", 2)
+df$site <- substr(sites, start = 1, stop = nchar(sites)-1)
+df$interintra <- substr(sites, start = nchar(sites), stop = nchar(sites))
+ggplot(df, aes(x=site, y=value, fill=interintra))+
+  geom_bar(stat="identity", position="dodge")+
+  facet_wrap("vitalrate", scales="free_y")
+
+
+

### 2015-12-23 docs/ms_communitySynchrony.Rmd
more discussion additions

@@
 Such independent fluctuations are expected for per capita growth rates regardless of population sizes, whereas synchrony of percent cover should trend toward one (perfect synchrony) as population sizes increase and the effect of demographic stochasticity is diminished.
-We find this to be true in our simulations: for each community, community synchrony of per capita growth rates are approximately equal to 1/*S* when only demographic stochasticity is acting (Fig.
+Our results conform to expectations: for each community except Arizona, community synchrony of per capita growth rates are approximately equal to 1/*S* when only demographic stochasticity is acting (Fig.
 3; top panels, dark purple lines) and synchrony of percent cover starts at approximately 1/*S* and increases linearly as population sizes increase (Fig.
 3; bottom panels, dark purple lines).
+In the Arizona community, synchrony of per capita growth rates is approximately 1/*S* as expected, but synchrony of precent cover does not increase with population size (Fig. 3f).
+This could indicate that intraspecific density-dependence is abnormally strong in the Arizona community, relative to the other communities we studied.
+Strong intraspecific density-dependence for the two species in the Arizona community would limit population size regardless of landscape size.
+Therefore, populations would not grow large enough to limit the desynchronizing effect of demographic stochasticity.
+In other words, if there is no relationship between landscape size and population size due to strong density-dependence, the effect of demgographic stochasticity will not lessen.
+TODO: CHECK MEAN COVER VERSUS LANDSCAPE SIZE FOR EACH COMMUNITY
+
+Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small.
+Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations.
+These results cofirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
+Only in the Idaho community does synchrony of per capita growth rates approach 1/*S* in a non-constant environment (Fig. 3x).
+This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
+
+### Environmental fluctuations drive community synchrony
+In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations.
+Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment.
+We find that environmental fluctuations drive synchrony of per capita growth rates and plant percent cover in all five communities.
+Removing environmental fluctuations consistently reduced synchrony (Figs. 2,3) and the synchrony of species environmental responses was a strong predictor of community synchrony (Fig. 4).
+This is unsurprising since we know that species interactions are weak in these communities.
+More surprising is that our results appear to be quite general: across five arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
+Can we generalize our results to a vast majority of plant communities?
+
+A recent analysis of several biodiversity-ecosystem function experiments suggests our results are anomolous.
+@Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
+However, their results are based on experimentally-assembled communities that have coexisted for less than 20 years.
+Thus, community assembly processes that can result in competitive exclusion may yet to have been realized.
+In contrast, the communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap (@Chu2015).
 

### 2015-12-23 devel/plot_timeseries.R

@@ -1,0 +1,11 @@
+##
+##  Development Script to Look at Individual Simulated Time Series
+##
+
+rm(list=ls())
+
+par(mfrow=c(1,2))
+ibms <- readRDS("../results/ibm_sims/ibm_Kansas_constnointerexpand1.RDS")
+matplot(ibms[,c(4:5)]*100, type="l")
+ibms <- readRDS("../results/ibm_sims/ibm_Kansas_constnointerexpand5.RDS")
+matplot(ibms[,c(4:5)]*100, type="l")

### 2015-12-26 docs/ms_communitySynchrony.Rmd
updates discussion; adds mono vs polycutlture plot

@@
 Again, no clear pattern emerges for the effect of removing environmental stochasticity on the synchrony of percent cover (Fig. 3, bottom panels).
+Community-wide averages of vital rate random year effects are strong predictors of IPM-simulated synchrony of per capita growth rates ($\text{R}^2$ = 0.67; Fig. 4)
 
@@
 We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.
-Our study produced three main findings that were consistent across all of our focal communities: (1) interspecific interactions had little impact on community synchrony because interspecific interactions were weak; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony since species-specific responses to the environment were positively, but not completely, correlated.
+Our study produced three main findings that were consistent across five plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony since species-specific responses to the environment were positively, but not completely, correlated.
 Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community synchrony approximately equals the synchrony of species' responses to the environment.
@@
 We find that environmental fluctuations drive synchrony of per capita growth rates and plant percent cover in all five communities.
-Removing environmental fluctuations consistently reduced synchrony (Figs. 2,3) and the synchrony of species environmental responses was a strong predictor of community synchrony (Fig. 4).
+Removing environmental fluctuations consistently reduced synchrony (Fig.
+3) and the synchrony of species environmental responses was a strong predictor of community synchrony ($\text{R}^2$ = 0.67; Fig.
+4).
+Likewise, removing interspecific interactions resulted in no discernable change in community synchrony of per capita growth rates (Fig. 2).
+The lack of change indicates a strong correlation between species growing in polyculture (with interspecific interactions) and in monoculture (without interspecific interactions).
+If we consider sychrony in monoculture to represent the synchrony of species' responses to the environment, then our results lead to only one conclusion: environmental fluctuations are the primary driver of community synchrony.
 This is unsurprising since we know that species interactions are weak in these communities.
@@
 @Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
-However, their results are based on experimentally-assembled communities that have coexisted for less than 20 years.
-Thus, community assembly processes that can result in competitive exclusion may yet to have been realized.
-In contrast, the communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap (@Chu2015).
+In fact, they found very weak correlations among species growin in polyculture and species growing in monoculture.
+However, their results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
+Thus, community assembly processes that can result in competitive exclusion may yet to have been realized and transient strong interspecific competition could be masking the synchronizing effects of the environment.
+In contrast, the communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
 
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/envresp_vs_ipmsynch.png}
+  \caption{Average synchrony of IPM-simulated per capita growth rates (from 13 year sections) as a function of the average synchrony of the vital rate random year effects for each site.
+Community sychrony is expected to approximately equal the synchrony of species' environmental responses in the absence of demographic stochasticity and strong interspecific interactions.
+The synchrony of species' random year effects across vital rates is a strong ($\text{R}^2$ = 0.67), but not perfect, predictor of community synchrony.
+Points are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.}
+\end{figure}
+
+\pagebreak{}
+
 ```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations.", eval=F}

### 2015-12-26 devel/plot_yreffect_corrs.R

@@ -12,11 +12,11 @@
 ####  Clean workspace; Load libraries ------------------------------------------
 ####
 rm(list = ls()) # wipe the workspace clean
+library(ggplot2)
 library(reshape2)
 library(plyr)
-library(ggplot2)
 library(synchrony)
-
+library(communitySynchrony)
 
 
 ####
@@ -156,6 +156,81 @@ ggplot(pgr_vr_long, aes(x=pgr_synch, y=value, group=variable))+
   xlab("Synchrony of Yearly Per Capita Growth Rates")+
   ylab("Synchrony of Random Year Effects")+
   theme_bw()
+
+
+
+####
+####  Plot Synchrony of Yearly Per Capita Growth Rates vs. PGR Synchrony
+####
+output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
+mlist <- melt(output_list)
+colnames(mlist)[1:3] <- c("year", "species", "cover")
+sites <- unique(mlist$L1)
+sims <- unique(mlist$L2)
+synch_df <- data.frame(site=NA, experiment=NA, bootnum=NA, 
+                       pgr_synch=NA, abund_synch=NA)
+boots <- 100
+num_iters <- 13
+for(dosite in sites){
+  tmp_data <- subset(mlist, L1==dosite)
+  for(dosim in sims){
+    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
+    for(i in 1:boots){
+      begin_year <- sample(x = 1:(max(tmpsim$year)-num_iters), 1)
+      end_year <- begin_year+num_iters
+      tmp <- subset(tmpsim, year %in% begin_year:end_year)
+      tmpsynch <- get_ipm_synchrony(tmp)
+      tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
+      tmpcast <- dcast(tmp, year~species, value.var = "cover")
+      tmp_abund_synch <- as.numeric(community.sync(tmpcast[2:ncol(tmpcast)])[1])
+      tmp_df <- data.frame(site=dosite, experiment=dosim, 
+                           bootnum=i, pgr_synch=tmp_pgr_synch, 
+                           abund_synch=tmp_abund_synch)
+      synch_df <- rbind(synch_df, tmp_df)
+    }# end boots loop
+  }# end experiment/sim loop
+}# end site loop
+
+synch_dftmp <- synch_df[2:nrow(synch_df),]
[…]

### 2015-12-26 docs/ms_communitySynchrony.Rmd
updates figures with poly vs mono

@@
 geometry: margin=1in
-
+linkcolor: black
+urlcolor: black
 ---
@@
 
+<!--
 \pagebreak{}
@@
 Points are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.}
+\end{figure}
+-->
+\pagebreak{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/poly_vs_mono_synch.png}
+  \caption{Synchrony of communities simulated without interspecific interactions (monoculture) versus synchrony of communities simulated with interspecific interactions (polyculture).
+We used the same sequence of random year effects for both simulations (mono and polyculture) to mimic biodiversity-ecosystem function experiments.
+The dashed line is the line of equality.}
 \end{figure}

### 2015-12-27 docs/ms_communitySynchrony.Rmd
discussion updates

@@
 Many previous investigations of community synchrony have implicitly assumped that species will fluctuate independently in the absence of temporal niche differences and that strong interspecific interactions cause compensatory dynamics.
-For example, the seminal study by @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
+For example, @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
 Based on that result, @Houlahan2007 concluded that abiotic factors synchronize species' abundances.

### 2015-12-28 docs/ms_communitySynchrony.Rmd
updates time series observation figure

@@
 
-```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site.", eval=FALSE}
-# out_ks$percent_cover$site <- "Kansas"
-# out_id$percent_cover$site <- "Idaho"
-# out_mt$percent_cover$site <- "Montana"
-# out_nm$percent_cover$site <- "New Mexico"
-# out_az$percent_cover$site <- "Arizona"
-# cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
-#                     out_mt$percent_cover, out_nm$percent_cover, 
-#                     out_az$percent_cover)
-# 
-# out_ks$growth_rates$site <- "Kansas"
-# out_id$growth_rates$site <- "Idaho"
-# out_mt$growth_rates$site <- "Montana"
-# out_nm$growth_rates$site <- "New Mexico"
-# out_az$growth_rates$site <- "Arizona"
-# pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
-#                   out_mt$growth_rates, out_nm$growth_rates, 
-#                   out_az$growth_rates)
-# all_spp <- unique(pgr_data$species)
-# color_match <- data.frame(species = all_spp,
-#                           color = brewer.pal(11,"Paired"))
-# pgr_data <- merge(pgr_data, color_match, by = "species")
-# cover_data <- merge(cover_data, color_match, by="species")
-# 
-# cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
-#                 geom_line()+
-#                 geom_point()+
-#                 facet_grid(site~., scales = "free_y")+
-#                 guides(color=FALSE)+
-#                 ylab("Mean Percent Cover")+
-#                 xlab("Year")+
-#                 scale_color_manual(values = brewer.pal(11,"Paired"))
-# 
-# pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
-#               geom_line()+
-#               geom_point()+
-#               facet_grid(site~., scales = "free_y")+
-#               guides(color=FALSE)+
-#               ylab("Per Capita Growth Rate")+
-#               xlab("Year")+
-#               scale_color_manual(values = brewer.pal(11,"Paired"))
-# full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
+\pagebreak{}
+
+Figures
+-------
+
+```{r time_series_fig, echo=FALSE, include=TRUE, fig.width=8.5, fig.height=4, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site.
+Panels are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.", eval=T}
+out_ks$percent_cover$site <- "Kansas"
+out_id$percent_cover$site <- "Idaho"
+out_mt$percent_cover$site <- "Montana"
+out_nm$percent_cover$site <- "New Mexico"
+out_az$percent_cover$site <- "Arizona"
+cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
+                    out_mt$percent_cover, out_nm$percent_cover, 
+                    out_az$percent_cover)
+
+out_ks$growth_rates$site <- "Kansas"
+out_id$growth_rates$site <- "Idaho"
+out_mt$growth_rates$site <- "Montana"
+out_nm$growth_rates$site <- "New Mexico"
+out_az$growth_rates$site <- "Arizona"
+pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
+                  out_mt$growth_rates, out_nm$growth_rates, 
+                  out_az$growth_rates)
+all_spp <- unique(pgr_data$species)
+color_match <- data.frame(species = all_spp,
+                          color = brewer.pal(12,"Paired"))
+pgr_data <- merge(pgr_data, color_match, by = "species")
+cover_data <- merge(cover_data, color_match, by="species")
+
+### Create labelling dataframe
+site_list <- unique(cover_data$site)
+textdf <- data.frame(species=NA, year=NA, tot_cover=NA, num_obs=NA, site=NA, color=NA)
+for(dosite in site_list){
+  tmp <- subset(cover_data, site==dosite)
+  if(dosite=="Kansas"){
+    maxyr <- min(unique(tmp$year))
+    textdf <- rbind(textdf,tmp[tmp$year == maxyr, ])
+  }
+  if(dosite!="Kansas"){
+    maxyr <- max(unique(tmp$year))
+    textdf <- rbind(textdf,tmp[tmp$year == maxyr, ])
+    textdf$year <- textdf$year+2
+  }
+}
+textdf <- textdf[2:nrow(textdf),]
+
+### Change site to two-letter codes
+site_list <- sort(unique(cover_data$site))
+sitecode <- c("AZ", "ID", "KS", "MT", "NM")
+for(i in 1:length(site_list)){
+  cover_data[cover_data$site==site_list[i], "site"] <- sitecode[i]
+  pgr_data[pgr_data$site==site_list[i], "site"] <- sitecode[i]
+}
+
+cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
+                geom_line(alpha=0.25)+
+                geom_point(size=2)+
+                facet_grid(site~., scales = "free_y")+
+                guides(color=FALSE)+
+                ylab("Mean Percent Cover")+
+                xlab("Year")+
+                scale_color_manual(values = brewer.pal(12,"Paired"))+
+                theme_bw()
+
+pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
+              geom_line(alpha=0.25)+
+                geom_point(size=2)+
+              facet_grid(site~., scales = "free_y")+
+              guides(color=FALSE)+
+              ylab("Per Capita Growth Rate")+
+              xlab("Year")+
+              scale_color_manual(values = brewer.pal(12,"Paired"))+
+              theme_bw()
+full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
 ```
@@
 
-Figures
--------
 ```{r concept_fig, echo=FALSE, fig.cap="Conceptual figure outlining the effects of interspecific interactions and environmental stochasticity in large populations where demographic stochasticity is effectively absent (A) and in populations of varying size where demographic stochasticity is present (B).", eval=F}

### 2015-12-28 docs/ms_communitySynchrony.Rmd
updates discussion with community assembly and conclusions

@@
    - \usepackage[document]{ragged2e}
+   - \usepackage{color}
 bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
@@
 
+\begin{singlespace}
+
 \begin{center}
@@
 
-\normalsize{Andrew T.
-Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B.
-Adler\textsuperscript{1}}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
 
@@
 
-\vspace{5em}
-
-*Authorship*: ATT, CdM, and PBA designed the research, ATT performed the research, and ATT, CdM, ML, and PBA contributed to the analyses and interpretation of results.
+\textit{Authorship}: ATT, CdM, and PBA designed the research, ATT performed the research, and ATT, CdM, ML, and PBA contributed to the analyses and interpretation of results.
 All authors contributed to writing the manuscript.
 
+\textit{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
+
+\textit{Running Title}: Species synchrony in grasslands
+
+\textit{Article Type}: Letter
+
+\textit{Number of Words}: Abstract = ; Main Text =
+
+\textit{Number of References}:
+
+\textit{Number of Tables and Figures}:  Tables;  Figures
+
+\textit{Corresponding Author}: Andrew Tredennick \\  
+Address: Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA \\
+Phone: +1-970-443-1599 \\
+Fax: +1-435-797-3796 \\
+Email: atredenn@gmail.com
+
+\end{singlespace}
 \renewcommand*{\thefootnote}{\arabic{footnote}}
@@
 
-__Keywords__: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
-
 <!-- \setlength{\parindent}{5ex} -->
+\setlength{\parindent}{5ex}
 
@@
 An alternative view is that density-independent processes like environmental forcing drive population fluctuations and tend to synchronize dynamics of ecologically-similar species [@Strong1983; @Fischer2001].
-Clearly, these forces are not mutually exclusive, and most research has focused on determining the relative strength of biotic and abiotic forcing on the synchrony of community dynamics [e.g., @Ives1995; @Klug2000; @Houlahan2007; @Mutshinda2009].
-However, recent theory [@Loreau2008a] suggests that previous interpretations of time series have been impaired by a (1) faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects to estimate immediate responses to the environment and competition.
+Clearly, these views are not mutually exclusive, and most research has focused on determining the relative strength of biotic and abiotic forcing on the synchrony of community dynamics [e.g., @Ives1995; @Klug2000; @Houlahan2007; @Mutshinda2009].
+However, recent theory [@Loreau2008a] suggests that previous interpretations of time series have been impaired by (1) a faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects and drift to estimate immediate responses to the environment and competition.
 Moreover, interspecific competition may actually synchronize, rather than de-synchronize, population dynamics [@Loreau2008a; @Loreau2010].
@@
 
-where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate. 
+\noindent where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
 The metric of @Loreau2008a makes clear that synchrony and asynchrony are not binary states, contrary to their common usage in the literature.
@@
 
-where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
+\noindent where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
 *S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
@@
 In other words, if there is no relationship between landscape size and population size due to strong density-dependence, the effect of demgographic stochasticity will not lessen.
-TODO: CHECK MEAN COVER VERSUS LANDSCAPE SIZE FOR EACH COMMUNITY
+\todo[inline, size=\small]{CHECK MEAN COVER VERSUS LANDSCAPE SIZE FOR EACH COMMUNITY}
 
@@
 More surprising is that our results appear to be quite general: across five arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
-Can we generalize our results to a vast majority of plant communities?
+Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results are likely very general when applied to natural grasslands.
 
-A recent analysis of several biodiversity-ecosystem function experiments suggests our results are anomolous.
+A challenge to generality of our results is that a recent analysis of several biodiversity-ecosystem function experiments suggests our results may be anomolous.
 @Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
-In fact, they found very weak correlations among species growin in polyculture and species growing in monoculture.
-However, their results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
-Thus, community assembly processes that can result in competitive exclusion may yet to have been realized and transient strong interspecific competition could be masking the synchronizing effects of the environment.
-In contrast, the communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
+This is surprising given the extremely weak species interactions we estimated.
+But the apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
+Gross et al.'s results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
+In contrast, the natural communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
+In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments community assembly processes that can result in competitive exclusion may yet to have been realized.
+Therefore, such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
+This also leads to the prediction that community synchrony is driven by competition in recently assembled communities, while synchrony is driven by environmental fluctuations in communities that have undergone extensive species sorting.
+In other words, in the absence of disturbance and invasion, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
+One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.
+
+The generality of our results only extends to communities composed of a single trophic level. \textcolor{red}{DO WE NEED THIS?}
+
+Our analysis also contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
+Functional traits define how a species will respond to its environment.
+Therefore, greater functional differences among species should lead to lower correlations in species' responses to the environment.
+We found evidence for the negative effect of functional diversity on community synchrony: community synchrony was lowest Idaho (Table 2, Fig.
+2); the only community that contained two life forms (one shrub, three perennial grasses).
+In fact, when we ran simulations with the shrub removed, synchrony of the Idaho community increased substantially and was similar to synchrony in the Kansas community, which also contained three species.
+Functional diversity is especially important in communities with weak interspecific interactions where synchrony depends soley on species' environmental responses.
+Adding species to such communities will always decrease synchrony, and increase temporal stability, so long as the synchrony of species' environmental response is less than 1 [@Loreau2008a; @Loreau2013].
+But functionally diverse suites of species will have the lowest synchrony of environmental responses, and thus lowest community synchrony and highest stability.
+
+Conclusion
+----------
+We used empirically-based, multi-species population models to test predictions arising from theory on the drivers of species synchrony.
+Our results show that community-wide synchrony is predominantly driven by environmental forcing across five perennial plant communities.
+Thus, our results support Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak.
+We have reason to believe that our results are quite general since many studies report strong niche differences in plant communities; meaning that species interactions are weak.
+However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony.
+We conclude that the difference between the Gross et al.
+(2014) study and our own is due to community assembly processes: recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
+Viewing community synchrony through the lens of community assembly processes may provide a way to reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.
 
@@
 
+\setlength{\parindent}{0ex}
 \newpage{}
@@
                 geom_point(size=2)+
-              facet_grid(site~., scales = "free_y")+
+              facet_grid(site~.)+
               guides(color=FALSE)+

### 2015-12-28 docs/ms_communitySynchrony.Rmd
updates ibm figure with burn in removal

@@
 We used empirically-based, multi-species population models to test predictions arising from theory on the drivers of species synchrony.
-Our results show that community-wide synchrony is predominantly driven by environmental forcing across five perennial plant communities.
-Thus, our results support Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak.
-We have reason to believe that our results are quite general since many studies report strong niche differences in plant communities; meaning that species interactions are weak.
+Our results are consistent across five perennial plant communities and show that community-wide synchrony is predominantly driven by environmental forcing.
+Thus, we find support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak.
+We have reason to believe that our results are quite general since many studies report large niche differences, and thus weak interactions, in plant communities.
 However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony.
@@
 ```{r time_series_fig, echo=FALSE, include=TRUE, fig.width=8.5, fig.height=4, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site.
-Panels are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.", eval=T}
+Panels are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.", eval=F}
 out_ks$percent_cover$site <- "Kansas"

### 2015-12-28 devel/plot_abund_v_landscape.R

@@ -1,28 +1,13 @@
-## Script to plot IBM results
+##
+##  Plots mean abundance as a function of landscape size
+##  From IBM simulations
+##
 
-##  Author:       Andrew Tredennick
-##  Email:        atredenn@gmail.com
-##  Date created: 9-22-2015
-
-# Clear workspace 
-rm(list=ls(all=TRUE))
+rm(list=ls())
 
 totSims <- 50      # number of simulations per site 
 totT <- 100      # time steps of simulation
 burn.in <- 25    # time steps to discard before calculating cover values
-site_colors <- c("grey45", "steelblue", "slateblue4", "darkorange", "purple")
-
-####
-####  Load libraries -----------------------------------------------------------
-####
-library(ggplot2)
-library(reshape2)
-library(plyr)
-library(synchrony)
-library(gridExtra)
-
-### Function for inverse of %in%
-"%w/o%" <- function(x, y) x[!x %in% y] # x without y
 
 
 ####
@@ -52,7 +37,7 @@ rm(stringlist)
 
 
 output_df <- data.frame(experiment=NA, site=NA, expansion=NA, 
-                        run=NA, pgr_synch=NA, abund_synch=NA)
+                        run=NA, meanabund=NA)
 
 for(do_exp in unique(files_df$experiment)){
   exp_files <- files_df[which(files_df$experiment==do_exp),"filename"]
@@ -108,26 +93,10 @@ for(do_exp in unique(files_df$experiment)){
           species_names <- species_names[species_names!="Cov"]
           num_spp <- ncol(ts.tmp)
           ts.tmp$timestep <- c(1:nrow(ts.tmp))
-          
-          # Transform to per capita growth rates
-          lagts <- ts.tmp
-          lagts$lagtimestep <- lagts$timestep+1
-          colnames(lagts)[1:(num_spp+1)] <- paste0(colnames(lagts)[1:(num_spp+1)],"_t0") 
-          mergedts <- merge(ts.tmp, lagts, by.x="timestep", by.y="lagtimestep")
-          transitions <- nrow(mergedts)
-          obs_gr <- matrix(nrow=transitions, ncol=num_spp)
-          name_ids1 <- which(colnames(mergedts) %in% paste0("Cov.",species_names))
-          name_ids2 <- which(colnames(mergedts) %in% paste0("Cov.",species_names, "_t0"))
-          for(ii in 1:transitions){
[…]

### 2016-01-18 docs/ms_communitySynchrony.Rmd
small changes to title page

@@
 
-\textit{Authorship}: ATT, CdM, and PBA designed the research, ATT performed the research, and ATT, CdM, ML, and PBA contributed to the analyses and interpretation of results.
-All authors contributed to writing the manuscript.
+\vspace{2em}
 
@@
 
-\textit{Running Title}: Species synchrony in grasslands
+\vspace{2em}
 
-\textit{Article Type}: Letter
+\textbf{Authorship}: ATT, CdM, and PBA designed the research, ATT performed the research, and ATT, CdM, ML, and PBA contributed to the analyses and interpretation of results.
+All authors contributed to writing the manuscript.
 
-\textit{Number of Words}: Abstract = ; Main Text =
+\textbf{Running Title}: Drivers of species synchrony in natural grasslands
 
-\textit{Number of References}:
+\textbf{Article Type}: Letter
 
-\textit{Number of Tables and Figures}:  Tables;  Figures
+\textbf{Number of Words}: Abstract = ; Main Text =
 
-\textit{Corresponding Author}: Andrew Tredennick \\  
+\textbf{Number of References}:
+
+\textbf{Number of Tables and Figures}:  Tables;  Figures
+
+\textbf{Corresponding Author}: Andrew Tredennick \\  
 Address: Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA \\
@@
 What temporal variable should we use to measure synchrony?
-In constrast to virtually every study on synchrony that uses population abundance (CITATIONS), @Loreau2008a suggest community-wide synchrony is best viewed through the lense of per capita growth rates.
+In constrast to virtually every study on synchrony that uses population abundance [e.g., @Houlahan2007; @Mutshinda2009; @Hector2010; @Gross2014; @Vasseur2014], @Loreau2008a suggest community-wide synchrony is best viewed through the lense of per capita growth rates.
 On conceptual grounds, per capita growth rates represent the immediate response of species to the environment and competition that we wish to attribute to synchrony.
-On practical grounds, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a].
+On practical grounds, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a; @Ranta2008].
 Therefore, measuring the synchrony of population abundances through time does not allow inference on the drivers of synchrony.
@@
 They found that the effect of environmental forcing on temporal dynamics is much stronger than that of interspecific interactions, across all taxa they analyzed [@Mutshinda2009].
-However, the @Mutshinda2009 study did not include data from plant communities that have, so far, received the most evidence for a strong role for interspecific interactions in driving species asynchrony [@Gross2014].
+However, the @Mutshinda2009 study did not include data from plant communities that have, so far, received the most evidence for a strong role for interspecific interactions in determining species synchrony [@Gross2014].
 
@@
 
-\pagebreak{}
-
 ```{r concept_fig, echo=FALSE, fig.cap="Conceptual figure outlining the effects of interspecific interactions and environmental stochasticity in large populations where demographic stochasticity is effectively absent (A) and in populations of varying size where demographic stochasticity is present (B).", eval=F}

### 2016-01-20 docs/ms_communitySynchrony.Rmd
updates main functions to return expected sychrony from Claire's new equation

@@
                           "Species richness")
-table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
+# table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
 table_data[,4] <- table_data[,2] - table_data[,3] # calculate obs - exp synchrony

### 2016-01-20 R/community_synchrony.R

@@ -77,13 +77,22 @@ get_comm_synchrony <- function(ts_data){
   
   
   ##  Expected synchrony under independent fluctuations
-  sigma <- numeric(num_spp)
-  sigma_sqr <- numeric(num_spp)
-  for(i in 1:num_spp){
-    sigma[i] <- sd(obs_gr[,i])
-    sigma_sqr[i] <- (sd(obs_gr[,i]))^2
-  }
-  expected_synchrony_ind_flucts <- (sum(sigma_sqr)) / ((sum(sigma))^2)
+  # New formula from Claire
+  ts_freq <- ddply(ts_data, .(year, species), summarise,
+                   avg_totcov = mean(totCover))
+  ts_freq_tmp <- dcast(ts_freq, year~species, value.var = "avg_totcov")
+  ts_freq_tmp$total <- rowSums(ts_freq_tmp[,c(2,3,4)])
+  ts_freq_wide <- colMeans(ts_freq_tmp[,c(2,3,4)] / ts_freq_tmp[,5])
+  expected_cover_synchrony <- 1 / (sum(ts_freq_wide^(1/2)))^2
+  expected_pgr_synchrony <- sum(ts_freq_wide^-1) / (sum(ts_freq_wide^(-1/2)))^2
+  
+#   sigma <- numeric(num_spp)
+#   sigma_sqr <- numeric(num_spp)
+#   for(i in 1:num_spp){
+#     sigma[i] <- sd(obs_gr[,i])
+#     sigma_sqr[i] <- (sd(obs_gr[,i]))^2
+#   }
+#   expected_synchrony_ind_flucts <- (sum(sigma_sqr)) / ((sum(sigma))^2)
   
   
   ##  Output
@@ -92,7 +101,8 @@ get_comm_synchrony <- function(ts_data){
   colnames(obs_gr) <- c("year","species","pgr")
   return(list(stability = stability,
               pgr_synchrony = growth_rate_synchrony,
-              pgr_expected_synch_ind_flucts = expected_synchrony_ind_flucts,
+              pgr_expected_synch_ind_flucts = expected_pgr_synchrony,
+              cover_expected_synch_ind_flucts = expected_cover_synchrony,
               abund_synchrony = synch_abundance,
               growth_rates = obs_gr,
               percent_cover = ts_agg))

### 2016-01-22 docs/ms_communitySynchrony.Rmd
adds percent cover synchrony to table 2

--- 2016-01-18 docs/ms_communitySynchrony.Rmd
@@
 
-```{r get_all_synch, echo=FALSE, include=TRUE, eval=T}
+```{r get_all_synch, echo=FALSE, include=TRUE, eval=T, cache=FALSE}
 out_ks <- get_comm_synchrony(ts_data = ks_data)
@@
 
-```{r synchrony_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T}
+```{r synchrony_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}
 get_table_metrics <- function(output){
@@
   demo_tmp <- obs_tmp - exp_tmp
-  return(c(obs_tmp, exp_tmp, demo_tmp))
+  exp_cover <- round(as.numeric(tmp$cover_expected_synch_ind_flucts), 2)
+  obs_cover <- round(as.numeric(tmp$abund_synchrony[1]), 2)
+  demo_cover <- obs_cover - exp_cover
+  return(c(obs_tmp, exp_tmp, demo_tmp, obs_cover, exp_cover, demo_cover))
 }
@@
 table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
-table_data <- table_data[,c("Site", "V1", "V2", "V3")]
+table_data <- table_data[,c("Site", "V1", "V2", "V3", "V4", "V5", "V6")]
 table_data$spp <- c(length(unique(ks_data$species)),
@@
                           "$\\text{E}(\\phi_{r})$", 
-                          "$\\text{E}(\\sigma_{c} + \\sigma_{e})$",
+                          "$\\text{E}(\\sigma_{r,c} + \\sigma_{r,e})$",
+                          "$\\phi_{C}$", 
+                          "$\\text{E}(\\phi_{C})$",
+                          "$\\text{E}(\\sigma_{C,c} + \\sigma_{C,e})$",
                           "Species richness")
-table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
-table_data[,4] <- table_data[,2] - table_data[,3] # calculate obs - exp synchrony
+# table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
+# table_data[,4] <- table_data[,2] - table_data[,3] # calculate obs - exp synchrony
 

### 2016-01-22 R/community_synchrony.R

@@ -82,7 +82,7 @@ get_comm_synchrony <- function(ts_data){
                    avg_totcov = mean(totCover))
   ts_freq_tmp <- dcast(ts_freq, year~species, value.var = "avg_totcov")
   ts_freq_tmp$total <- rowSums(ts_freq_tmp[,c(2:ncol(ts_freq_tmp))])
-  ts_freq_wide <- colMeans(ts_freq_tmp[,c(2:ncol(ts_freq_tmp)-1)] / ts_freq_tmp[,ncol(ts_freq_tmp)])
+  ts_freq_wide <- colMeans(ts_freq_tmp[,c(2:(ncol(ts_freq_tmp)-1))] / ts_freq_tmp[,ncol(ts_freq_tmp)])
   expected_cover_synchrony <- 1 / (sum(ts_freq_wide^(1/2)))^2
   expected_pgr_synchrony <- sum(ts_freq_wide^-1) / (sum(ts_freq_wide^(-1/2)))^2
   

### 2016-01-22 devel/plot_ipm_sims.R
adds second panel to fig 4 with yearly pgr synchrony

@@ -76,14 +76,47 @@ dev.off()
 library(ggthemes)
 polymono <- agg_synch[,c("site", "typesynch", "experiment", "mean_synch")]
 polymono_wide <- dcast(polymono, site+typesynch~experiment, value.var = "mean_synch")
-ggplot(polymono_wide, aes(x=ENVNOINTER, y=ENVINTER))+
+g1 <- ggplot(polymono_wide, aes(x=ENVNOINTER, y=ENVINTER))+
   geom_abline(aes(intercept=0, slope=1), linetype=3)+
   geom_point(size=3, aes(shape=typesynch))+
   scale_y_continuous(limits=c(0,1))+
   scale_x_continuous(limits=c(0,1))+
   ylab("Species synchrony in polyculture")+
-  xlab("Species synchrony in monoculture")+
+  xlab("Species synchrony \nin monoculture")+
   scale_shape_discrete(name="", labels=c("Per capita growth rate", "Percent cover"))+
   theme_few()+
+  ggtitle("A                                             ")+
   theme(legend.position=c(0.3,0.85))
-ggsave("../docs/components/poly_vs_mono_synch.png", width = 4, height = 3.5, dpi = 150)
+
+pgr_list <- readRDS("../results/ipm_yearly_pgr.RDS")
+site_names <- names(pgr_list)
+out_df <- data.frame(site=NA, pgr_synch=NA)
+for(do_site in site_names){
+  tmp_pgrs <- pgr_list[[do_site]]
+  tmp_synch <- as.numeric(community.sync(tmp_pgrs)[1])
+  tmp_df <- data.frame(site=do_site, pgr_synch=tmp_synch)
+  out_df <- rbind(out_df, tmp_df)
+}
+pgr_synch <- out_df[2:nrow(out_df),]
+polymono_wide$yrpgr <- rep(pgr_synch$pgr_synch, each=2)
+
+g2 <- ggplot(polymono_wide, aes(x=yrpgr, y=ENVINTER))+
+  geom_abline(aes(intercept=0, slope=1), linetype=3)+
+  geom_point(size=3, aes(shape=typesynch))+
+  scale_y_continuous(limits=c(0,1))+
+  scale_x_continuous(limits=c(0,1))+
+  ylab("Species synchrony in polyculture")+
+  xlab("Yearly per capita growth rate synchrony \nin monoculture")+
+  scale_shape_discrete(name="", labels=c("Per capita growth rate", "Percent cover"))+
+  theme_few()+
[…]

### 2016-01-22 runscripts/run_ipm_yearlypgr_simulations.R

@@ -140,7 +140,7 @@ for(do_site in site_list){
   site_output <- list()
   
   constant <- FALSE                 # always run with random year effects
-  spp_interact <- TRUE              # always run with interspp interactions
+  spp_interact <- FALSE              # always run with interspp interactions
   n_spp <- Nspp <- length(spp_list) # all possible forms of Nspp
   A <- 10000                        # area of a 1x1 meter plot, in cm
   maxSize <- max_size               # redo for IPM source script

### 2016-01-25 docs/ms_communitySynchrony.Rmd
adds cover and variability outputs to able 1

@@
    - \usepackage{color}
+   - \usepackage{rotating}
 bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
@@
   demo_cover <- obs_cover - exp_cover
-  return(c(obs_tmp, exp_tmp, demo_tmp, obs_cover, exp_cover, demo_cover))
+  plant_size <- tmp$avg_size
+  cover_var <- tmp$variability
+  return(c(obs_tmp, exp_tmp, demo_tmp, obs_cover, exp_cover, 
+           demo_cover, plant_size, cover_var))
 }
@@
 table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
-table_data <- table_data[,c("Site", "V1", "V2", "V3", "V4", "V5", "V6")]
+table_data <- table_data[,c("Site", "V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8")]
 table_data$spp <- c(length(unique(ks_data$species)),
@@
                           "$\\text{E}(\\sigma_{C,c} + \\sigma_{C,e})$",
+                          "Avg. Plant Size",
+                          "CV of Total Cover",
                           "Species richness")
@@
       sanitize.colnames.function = identity,
-      comment=FALSE)
+      comment=FALSE,
+      size="footnotesize",
+      floating.environment="sidewaystable")
 ```

### 2016-01-25 R/community_synchrony.R

@@ -17,6 +17,9 @@
 #' @return A list of data frames for synchrony and stability metrics for the community.
 
 get_comm_synchrony <- function(ts_data){
+  ##  Get average size of all plants
+  avg_cover <- mean(ts_data$totCover)
+  
   ## aggregate the quadrat-level data for average observed cover
   # divide totCover by 100 to convert from m2 to percent cover in 1m2 plot
   ts_agg <- ddply(ts_data, .(year, species), summarise,
@@ -75,6 +78,10 @@ get_comm_synchrony <- function(ts_data){
   colnames(obs_gr) <- species_list
   obs_gr$year <- merged_df$year
   
+  ##  Total cover variability
+  ts_tot <- apply(ts_mat[2:ncol(ts_mat)], 1, sum)
+  comm_cv <- sd(ts_tot)/mean(ts_tot)
+  
   
   ##  Expected synchrony under independent fluctuations
   # New formula from Claire
@@ -99,7 +106,9 @@ get_comm_synchrony <- function(ts_data){
   obs_gr <- melt(obs_gr, id.vars = "year")
   obs_gr <- obs_gr[with(obs_gr, order(year)), ]
   colnames(obs_gr) <- c("year","species","pgr")
-  return(list(stability = stability,
+  return(list(avg_size = avg_cover,
+              stability = stability,
+              variability = comm_cv,
               pgr_synchrony = growth_rate_synchrony,
               pgr_expected_synch_ind_flucts = expected_pgr_synchrony,
               cover_expected_synch_ind_flucts = expected_cover_synchrony,

### 2016-01-25 docs/ms_communitySynchrony.Rmd
adds constant random year list for IPMs

@@
 -------
-### Observed and simulated community synchrony
-Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 (Table 2).
-Synchrony tended to be higher in communities with few species (Arizona and New Mexico).
+Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
+Synchrony tended to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability (e.g., Montana).
 Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub (*Artimisia tripartita*) and three perennial grasses (Table 1).
-All sites exhibited more synchronous dynamics than expected if only demographic stochasticity was operating (E($\phi_r$) = 1/*S* in Table 2).
-The difference between observed $\phi_r$ and E($\phi_r$) represents the effect of the environment and species interactions combined .
-The combined effect of the environment and species interactions is positive in all communities (E($\phi_c + \phi_e$) in Table 2), indicating a synchronizing effect of those forces that counter-balance the desynchronizing effect of demographic stochasticity.
-For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. 1).
+For per capita growth rates, all sites exhibited more synchronous dynamics than expected if only demographic stochasticity was operating (E($\phi_r$) < $\phi_r$ in Table 2).
+The difference between observed $\phi_r$ and E($\phi_r$) represents the effect of the environmemt, species interactions, and species differences combined.
+The combined effect of the environment, species interactions, and species differences is positive in all communities (E($\phi_c + \phi_e$) in Table 2), indicating a synchronizing effect of those forces that counter-balance the desynchronizing effect of demographic stochasticity.
+In contrast, expected synchrony of percent cover was lower than expected if only demographic stochasticity was operating in two out of the five communities (Kansas and Idaho, Table 2).
 
-### Effect of removing interspecific interactions
-The statistical models for vital rates showed that interspecific interactions were weak in all communities (Table X1), as previously found by @Chu2015.
-Consistent with weak interspecific interactions, removing such interactions from IPM simulations resulted in no discernable changes in species synchrony of per capita growth rates or percent cover (Fig.
-2).
-This indicates that community regulation of population dynamics via coupled density-dependence ($\sigma_c$) is relatively weak in these communities.
-
-The effect of removing interspecific interactions on synchrony in IBM simulations was also small for per capita growth rates and was not affected by population size (Fig. 3, top panels).
+For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig.
+1), meaning that we can use the models to test the importance of different drivers of synchrony in these communities.
+Simulations from the IPM show that removing interspecific interactions has little affect on synchrony of per capita growth rates or percent cover in all five communities (Fig. 2).
+The weak effect of interspecific interactions on synchrony is not surprising because interspecific interactions are weak in our focal communities (Table X1, and see @Chu2015).
+The effect of removing interspecific interactions on synchrony of per capita growth rates in IBM simulations was also small and was not affected by population size (Fig. 3, top panels).
 The only discernable effect of interspecific interactions was in the Arizona and New Mexico communities when environmental stochasticity was also removed.
@@
   return(c(obs_tmp, exp_tmp, demo_tmp, obs_cover, exp_cover, 
-           demo_cover, plant_size, cover_var))
+           demo_cover, cover_var))
 }
@@
 table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
-table_data <- table_data[,c("Site", "V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8")]
+table_data <- table_data[,c("Site", "V1", "V2", "V3", "V4", "V5", "V6", "V7")]
 table_data$spp <- c(length(unique(ks_data$species)),
@@
                           "$\\text{E}(\\sigma_{C,c} + \\sigma_{C,e})$",
-                          "Avg. Plant Size",
                           "CV of Total Cover",

### 2016-01-25 runscripts/generate_randyrs.R

@@ -1,0 +1,16 @@
+##  Get sequence of random year effects for each site
+
+tlimit=2000
+
+# Read in growth parameters to get number of random years
+Gpars_all <- readRDS("../results/growth_params_list.RDS")
+site_list <- names(Gpars_all)
+
+randyr_list <- list()
+for(do_site in site_list){
+  Gpars_tmp <- Gpars_all[[do_site]]
+  Nyrs <- nrow(Gpars_tmp[[1]])
+  randyr_list[[do_site]] <- sample(1:Nyrs, size = tlimit, replace = TRUE)
+}
+
+saveRDS(randyr_list, "../results/randyr_list.RDS")

### 2016-01-25 runscripts/run_ipm_simulations.R

@@ -99,6 +99,8 @@ Gpars_all <- readRDS("../results/growth_params_list.RDS")
 Spars_all <- readRDS("../results/surv_params_list.RDS")
 Rpars_all <- readRDS("../results/recruit_parameters.RDS")
 
+randyrlist <- readRDS("../results/randyr_list.RDS")
+
 site_list <- names(Gpars_all)
 
 output_list <- list()
@@ -145,7 +147,8 @@ for(do_site in site_list){
   }
   
   ##  Set environment time series to be constant across simulations
-  randyrvec <- sample(1:Nyrs, size = tlimit, replace = TRUE)
+  # randyrvec <- sample(1:Nyrs, size = tlimit, replace = TRUE)
+  randyrvec <- randyrlist[[do_site]]
   
   sim_count <- 1
   site_output <- list()

### 2016-01-25 runscripts/ipm_intrinsic_pgr_source.R
updates pgr IPM with same randyear vector as other sims

@@ -8,7 +8,7 @@
 ##  Created: 10-30-2015
 ################################################################################
 
-maxR <- matrix(nrow=Nyrs,ncol=Nspp)
+maxR <- matrix(nrow=tlimit,ncol=Nspp)
 for(jjjj in 1:length(spp_list)){
   # set fixed species
   fixSpp <- spp_list[jjjj] ##need to change if for other species
@@ -79,8 +79,10 @@ for(jjjj in 1:length(spp_list)){
   for (i in 2:(tlimit)){
     
     #draw from observed year effects
-    allYrs=c(1:Nyrs)
-    doYear=allYrs[i-1]
+#     allYrs=c(1:Nyrs)
+#     doYear=allYrs[i-1]
+#     yrSave[i]=doYear
+    doYear <- randyrvec[i]
     yrSave[i]=doYear
     
     #get recruits per area

### 2016-01-25 runscripts/run_ipm_yearlypgr_simulations.R

@@ -91,6 +91,8 @@ Gpars_all <- readRDS("../results/growth_params_list.RDS")
 Spars_all <- readRDS("../results/surv_params_list.RDS")
 Rpars_all <- readRDS("../results/recruit_parameters.RDS")
 
+randyrlist <- readRDS("../results/randyr_list.RDS")
+
 site_list <- names(Gpars_all)
 
 output_list <- list()
@@ -99,6 +101,7 @@ for(do_site in site_list){
   Gpars_tmp <- Gpars_all[[do_site]]
   Spars_tmp <- Spars_all[[do_site]]
   Rpars_tmp <- Rpars_all[[do_site]]
+  randyrvec <- randyrlist[[do_site]]
   
   spp_list <- names(Gpars_tmp)
   Nyrs <- nrow(Gpars_tmp[[1]])
@@ -140,13 +143,13 @@ for(do_site in site_list){
   site_output <- list()
   
   constant <- FALSE                 # always run with random year effects
-  spp_interact <- FALSE              # always run with interspp interactions
+  spp_interact <- FALSE             # always run with interspp interactions OFF
   n_spp <- Nspp <- length(spp_list) # all possible forms of Nspp
   A <- 10000                        # area of a 1x1 meter plot, in cm
   maxSize <- max_size               # redo for IPM source script
   bigM <- iter_matrix_dims          # redo for IPM source script
   NoOverlap.Inter <- FALSE          # heterospecifics allowed to overlap
-  tlimit <- Nyrs+1                  # number of years to simulate
+  tlimit <- 2000                    # number of years to simulate
   size_dir <- "../results/stable_size_dists/" # path for stable size dists
   
   ## Run the IPM from a source script

### 2016-01-25 docs/ms_communitySynchrony.Rmd
updates results with consistent random year effect vectors

@@
 While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
-Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence sychronizes dynamics, demographic stochasticity desynchronizes dynamics, and environmental forcing can either synchronize or desynchronize poplation dynamics depending on the correlation of species-specific environmental responses.
-Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony through time.
+Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence, demographic stochasticity, and environmental forcing.
+However, general trends between synchrony and its drivers are elusive except in special limiting cases.
+Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony.
 Species' environmental responses were positively, but not perfectly correlated. 
@@
 Species interactions tend to be weak, so removing interspecific competition did not affect synchrony. 
-The effect of demographic stochasticity depended on the spatial scale of our modeling; demographic stochasticity was strongest in small populations and was able to modestly counteract the synchronizing effect of the environment.
+The effect of demographic stochasticity depended on the spatial scale of our modeling: demographic stochasticity was strongest in small populations and was able to modestly counteract the synchronizing effect of the environment.
 We conclude that niche differentiation among species is strong enough to reduce the synchronizing effect of coupled density dependence via competition, but this is countered by the synchronizing effect of similar environmental responses.
-The next frontier is to identify the specific environmental drivers of species synchrony to inform predictive efforts.
+Our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments that found species asynchrony to be driven by competition.
+We propose that such differences are due to community assembly processes: recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time.
+Viewing community synchrony through the lens of community assembly processes may provide a way to reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.
 

### 2016-01-25 runscripts/generate_randyrs.R

@@ -1,6 +1,6 @@
 ##  Get sequence of random year effects for each site
 
-tlimit=2000
+tlimit=2500
 
 # Read in growth parameters to get number of random years
 Gpars_all <- readRDS("../results/growth_params_list.RDS")

### 2016-01-25 runscripts/ipm_intrinsic_pgr_source.R

@@ -167,7 +167,7 @@ for(jjjj in 1:length(spp_list)){
   ## Calculate low density growth rate of focal species
   tmp1 <- which(colnames(covSave)==paste(fixSpp, ".t0", sep=""))
   tmp2 <- which(colnames(covSave)==paste(fixSpp, ".t1", sep=""))
-  pgr <- log(covSave[2:tlimit,tmp2]/covSave[2:tlimit,tmp1])
+  pgr <- log(covSave[1:tlimit,tmp2]/covSave[1:tlimit,tmp1])
   maxR[,jjjj] <- pgr
 } # end fixed species loop
 

### 2016-01-25 runscripts/run_ipm_yearlypgr_simulations.R

@@ -149,14 +149,14 @@ for(do_site in site_list){
   maxSize <- max_size               # redo for IPM source script
   bigM <- iter_matrix_dims          # redo for IPM source script
   NoOverlap.Inter <- FALSE          # heterospecifics allowed to overlap
-  tlimit <- 2000                    # number of years to simulate
+  tlimit <- 2500                    # number of years to simulate
   size_dir <- "../results/stable_size_dists/" # path for stable size dists
   
   ## Run the IPM from a source script
   source("ipm_intrinsic_pgr_source.R") 
   
   ##  Save site output to big list
-  output_list[[do_site]] <- maxR
+  output_list[[do_site]] <- maxR[501:tlimit,]
   
 } # end site loop
 

### 2016-01-26 docs/ms_communitySynchrony.Rmd
updates intro text with streamlining attempt

--- 2016-01-25 docs/ms_communitySynchrony.Rmd
@@
 Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time. 
-While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
-Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence sychronizes dynamics, demographic stochasticity desynchronizes dynamics, and environmental forcing can either synchronize or desynchronize poplation dynamics depending on the correlation of species-specific environmental responses.
-Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony through time.
+While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance of species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
+We use stochastic, multi-species population models fit using long term datasets from five semi-arid grasslands to determine the relative importance of the three forces that drive species synchrony: environmental fluctuations, interspecific competition, and demographic stochasticity.
 Species' environmental responses were positively, but not perfectly correlated. 
-As a result, environmental variation caused an increase in synchrony relative to an expectation of independent species' responses, but synchrony was not perfect.
-Species interactions tend to be weak, so removing interspecific competition did not affect synchrony. 
-The effect of demographic stochasticity depended on the spatial scale of our modeling; demographic stochasticity was strongest in small populations and was able to modestly counteract the synchronizing effect of the environment.
-We conclude that niche differentiation among species is strong enough to reduce the synchronizing effect of coupled density dependence via competition, but this is countered by the synchronizing effect of similar environmental responses.
-The next frontier is to identify the specific environmental drivers of species synchrony to inform predictive efforts.
+As a result, community-wide synchrony was higher when environmental variation was present rather than absent, but synchrony was not perfect.
+Removing interspecific competition, which was weak in all communities, did not affect synchrony, and demographic stochasticity only played a role in small populations.
+Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony is approximately equal to the synchrony of species' environmental responses -- a result that may be quite general if natural communities tend to have large niche differences among species.
+However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments, suggesting an interplay between community assembly and the drivers of community synchrony.
 
@@
 ------------
-Biodiversity tends to stabilize ecosystem function in fluctuating environments [@Hector2010; @DeMazancourt2013a; @Hautier2014].
-In the absence of a highly dominant and stable species, asynchrony in population dynamics (also called compensatory dynamics) among interacting species is the only way for biodiversity to stabilize aggregate community or ecosystem function [@Ives2002; @Loreau2010; @Loreau2013].
-Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010; @DeMazancourt2013a; @Gross2014; @Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009].
-Futhermore, analyses of observational time series tend to report synchronous, rather than asynchrounous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
+Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity stabilizes aggregate community or ecosystem function [@Ives2002; @Loreau2010; @Loreau2013].
+Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010; @DeMazancourt2013a; @Gross2014; @Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+Futhermore, analyses of observational time series tend to report synchronous, rather than asynchronous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
 Thus, our understanding of the prevalence and drivers of asynchronous population dynamics is limited.
@@
 Discovering the mechanisms behind population fluctuations, and whether they are synchronous among species or not, is a long-standing goal of ecology.
-One view is that density-dependent processes like interspecific competition lead to compensatory dyanmics [e.g., @Elton1942; @Tilman1988; @Ives1995].
+One view is that density-dependent processes like interspecific competition lead to compensatory dynamics [e.g., @Elton1942; @Tilman1988; @Ives1995].
 An alternative view is that density-independent processes like environmental forcing drive population fluctuations and tend to synchronize dynamics of ecologically-similar species [@Strong1983; @Fischer2001].
 Clearly, these views are not mutually exclusive, and most research has focused on determining the relative strength of biotic and abiotic forcing on the synchrony of community dynamics [e.g., @Ives1995; @Klug2000; @Houlahan2007; @Mutshinda2009].
-However, recent theory [@Loreau2008a] suggests that previous interpretations of time series have been impaired by (1) a faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects and drift to estimate immediate responses to the environment and competition.
-Moreover, interspecific competition may actually synchronize, rather than de-synchronize, population dynamics [@Loreau2008a; @Loreau2010].
+The consensus from analyses of observational time series across a wide range of taxa is that compensatory dynamics are rare and that similar environmental responses among species synchronize community dynamics [@Houlahan2007; @Mutshinda2009].
+In contrast, a recent analysis of several biodiversity-ecosystem function experiments found evidence for compensatory dynamics in grasslands driven by competitive interactions, not species' specific responses to the environment [@Gross2014].
+
+However, recent theory [@Loreau2008a; @Loreau2013] suggests that previous interpretations of time series have been impaired by (1) a faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects and drift to estimate immediate responses to the environment and competition.
 
@@
 Alternatively, negative summed covariance should indicate that competition structures community fluctuations by driving compensatory dynamics.
-Unfortunately, reliance on such a metric created the search for binary states -- synchronous or not, compensatory dynamics or not -- where a continuum exists [@Loreau2008a].
-Likewise, values of summed covariance are not exclusive to particular combinations of drivers, meaning that negative summed covariances can ocurr in the absence of competition [@Ranta2008].
+Unfortunately, reliance on such a metric created the search for binary states -- synchronous or not, compensatory dynamics or not -- where a continuum exists [@Loreau2008a; @Gross2014].
+Likewise, values of summed covariance are not exclusive to particular combinations of drivers, meaning that negative summed covariances can occur in the absence of competition [@Ranta2008].
 Finally, summed covariances have no upper or lower limits and are sensitive to community size, making it difficult to compare different communities [@Loreau2008a].
-In the end, the fact that few communities exhibit negative summed covariance [e.g., @Houlahan2007] yields little insight as to whether communities tend to be sychronized by environmental forces or not [@Loreau2010].
+In the end, the fact that few communities exhibit negative summed covariance [e.g., @Houlahan2007] yields little insight on the relative influence of competition and environmental forcing on community dynamics [@Loreau2010].
 
-To overcome these limitations, @Loreau2008a derived a new metric of species synchrony based on a multi-species population dynamic model.
-Their proposed metric, best suited to measure community-wide synchrony of per capita population growth rates, is
+To overcome many of these limitations, @Loreau2008a derived a new metric of species synchrony based on a multi-species population dynamic model.
+Their proposed metric, which can be applied to per capita growth rates and abundances, is
 
 \begin{equation}
-\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
+\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{x_{i}})^{2}}
 \end{equation}
 
-\noindent where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
+\noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or abundance ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
 The metric of @Loreau2008a makes clear that synchrony and asynchrony are not binary states, contrary to their common usage in the literature.
-Rather, the synchrony of community dynamics lies along a continuum from perfect asynchrony ($\phi_r=0$) to perfect synchrony ($\phi_r=1$).
+Rather, the synchrony of community dynamics lies along a continuum from perfect asynchrony ($\phi=0$) to perfect synchrony ($\phi=1$).
 Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony within a community and what process increase or decrease sychrony.
@@
 Unfortunately, there is not a single answer.
+Theoretical work has isolated three main factors that drive species synchrony in competitive communities: density-dependence via intra- and interspecific competition, environmental fluctuations, and demographic stochasticity [@Loreau2008a; @Loreau2013].
+
+Theory identifies several limiting cases, but none can be considered an expectation in natural communities where the strengths of internal and external forces are unkown.
+For example,
+
 The use of summed community covariance relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008a; @Loreau2010]. 

### 2016-01-26 devel/plot_ipm_sims.R

@@ -78,15 +78,20 @@ polymono <- agg_synch[,c("site", "typesynch", "experiment", "mean_synch")]
 polymono_wide <- dcast(polymono, site+typesynch~experiment, value.var = "mean_synch")
 g1 <- ggplot(polymono_wide, aes(x=ENVNOINTER, y=ENVINTER))+
   geom_abline(aes(intercept=0, slope=1), linetype=3)+
-  geom_point(size=3, aes(shape=typesynch))+
+  geom_point(size=3, aes(shape=typesynch, color=site))+
   scale_y_continuous(limits=c(0,1))+
   scale_x_continuous(limits=c(0,1))+
   ylab("Species synchrony in polyculture")+
   xlab("Species synchrony \nin monoculture")+
-  scale_shape_discrete(name="", labels=c("Per capita growth rate", "Percent cover"))+
+  scale_shape_discrete(name="Temporal Variable", labels=c("Per capita growth rate", "Percent cover"))+
+  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
+                     name = "Site")+
   theme_few()+
+  guides(shape=FALSE)+
   ggtitle("A                                                          ")+
-  theme(legend.position=c(0.3,0.85))
+  theme(legend.position=c(0.2,0.7))+
+  theme(legend.text = element_text(size = 8))+
+  theme(legend.background = element_rect(colour = "grey", fill = NA))
 
 
 
@@ -129,16 +134,17 @@ polymono_wide$yrpgr <- rep(pgr_synch$mean_pgrsynch, each=2)
 
 g2 <- ggplot(polymono_wide, aes(x=yrpgr, y=ENVINTER))+
   geom_abline(aes(intercept=0, slope=1), linetype=3)+
-  geom_point(size=3, aes(shape=typesynch))+
+  geom_point(size=3, aes(shape=typesynch, color=site))+
   scale_y_continuous(limits=c(0,1))+
   scale_x_continuous(limits=c(0,1))+
   ylab("Species synchrony in polyculture")+
   xlab("Yearly per capita growth rate synchrony \nin monoculture")+
-  scale_shape_discrete(name="", labels=c("Per capita growth rate", "Percent cover"))+
+  scale_shape_discrete(name="Temporal Variable", labels=c("Per capita growth rate", "Percent cover"))+
+  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
+                     name = "Site")+
   theme_few()+
[…]

### 2016-01-27 docs/ms_communitySynchrony.Rmd
more intro slimming

@@
 An alternative view is that density-independent processes like environmental forcing drive population fluctuations and tend to synchronize dynamics of ecologically-similar species [@Strong1983; @Fischer2001].
-Clearly, these views are not mutually exclusive, and most research has focused on determining the relative strength of biotic and abiotic forcing on the synchrony of community dynamics [e.g., @Ives1995; @Klug2000; @Houlahan2007; @Mutshinda2009].
-The consensus from analyses of observational time series across a wide range of taxa is that compensatory dynamics are rare and that similar environmental responses among species synchronize community dynamics [@Houlahan2007; @Mutshinda2009].
-In contrast, a recent analysis of several biodiversity-ecosystem function experiments found evidence for compensatory dynamics in grasslands driven by competitive interactions, not species' specific responses to the environment [@Gross2014].
-
+Clearly, these views are not mutually exclusive, and most research has focused on determining the relative strength of biotic and abiotic forcing on the synchrony of community dynamics [e.g., @Ives1995; @Klug2000; @Houlahan2007; @Mutshinda2009; @Gross2014].
 However, recent theory [@Loreau2008a; @Loreau2013] suggests that previous interpretations of time series have been impaired by (1) a faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects and drift to estimate immediate responses to the environment and competition.
@@
 Rather, the synchrony of community dynamics lies along a continuum from perfect asynchrony ($\phi=0$) to perfect synchrony ($\phi=1$).
-Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony within a community and what process increase or decrease sychrony.
+Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony within a community and what process increase or decrease sychrony [e.g., @Gross2014].
 
@@
 Unfortunately, there is not a single answer.
+The use of summed community covariance relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008a; @Loreau2010]. 
+Theory identifies several limiting cases [@Loreau2008a], but none can be considered an expectation in natural communities where the strengths of internal and external forces are unkown.
+For example, independent species fluctuations are expected if only demographic stochasticity if operating and species all have the same abundances and variance [@Loreau2008a].
+Furthermore, asymmetry in species' growth rates, competitive effects, and density-dependence can all interact, making it difficult to identify general trends between synchrony and its drivers.
+For example, competition generally, but not always, has a positive effect on synchrony of per capita growth rates and a negative effect on synchrony of abundances [@Loreau2013].
+Since the mechanisms behind community synchrony are context-dependent, the best way forward is to analyze the drivers of synchrony in natural communities and then relate observed synchrony back to limiting theoretical cases.
+In so doing, we can discover common and rare theoretical cases in nature.
+
+<!-- COMMENT THIS OUT FOR NOW...
+For example, 
 Theoretical work has isolated three main factors that drive species synchrony in competitive communities: density-dependence via intra- and interspecific competition, environmental fluctuations, and demographic stochasticity [@Loreau2008a; @Loreau2013].
 
-Theory identifies several limiting cases, but none can be considered an expectation in natural communities where the strengths of internal and external forces are unkown.
-For example,
-
-The use of summed community covariance relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008a; @Loreau2010]. 
 @Loreau2008a suggest there are two expectations: perfect synchrony in neutral communities where species have identical niches, and the inverse of the number of species in the community (1/*S*) when species fluctuate independently.
@@
 
+-->
+
 What temporal variable should we use to measure synchrony?
@@
 
-Aside from issues on how to measure synchrony, disentangling the factors that drive species syncrony is difficult.
+Aside from issues on how to measure synchrony, disentangling the factors that drive species synchrony is difficult.
 Observational timeseries of species' abundances allows easy quantification of summed covariances [@Houlahan2007] or species synchrony [@Gross2014], but quantifying the relative influence of density-dependence, demographic stochasticity, and environmental stochasticity remains a challenge.
 Indeed, to our knowledge, only one study has attempted to separate the influences of density-dependence and environmental stochasticity directly [@Mutshinda2009].
-@Mutshinda2009 used a statistical technique to decompose synchrony of species' abundances calculate from five time series.
+@Mutshinda2009 used a statistical technique to decompose synchrony of species' abundances calculated from five time series.
 They found that the effect of environmental forcing on temporal dynamics is much stronger than that of interspecific interactions, across all taxa they analyzed [@Mutshinda2009].
@@
 
-In this paper, we seek to overcome the limitations of previous studies by explicitly testing theory on all the drivers of species synchrony using empirically-based multi-species population models that include the effects of density-dependence, environmental stochasticity, and demographic stochasticity.
+In this paper, we seek to overcome the limitations of previous studies by explicitly testing theory on all the drivers of species synchrony using empirically-based, multi-species population models that include the effects of density-dependence, environmental stochasticity, and demographic stochasticity.
 We follow the theory of @Loreau2008a, who show that in a neutral community of *S* species, expected synchrony ($\phi_r$) is defined as:

### 2016-02-03 docs/ms_communitySynchrony.Rmd
trying to get the intro rright; still not

@@
 ------------
-Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity stabilizes aggregate community or ecosystem function [@Ives2002; @Loreau2010; @Loreau2013].
+Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity stabilizes aggregate community or ecosystem function [@Ives2002; @Gonzalez2009; @Loreau2010; @Loreau2013].
 Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010; @DeMazancourt2013a; @Gross2014; @Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009; @Loreau2013].
@@
 
-Discovering the mechanisms behind population fluctuations, and whether they are synchronous among species or not, is a long-standing goal of ecology.
-One view is that density-dependent processes like interspecific competition lead to compensatory dynamics [e.g., @Elton1942; @Tilman1988; @Ives1995].
-An alternative view is that density-independent processes like environmental forcing drive population fluctuations and tend to synchronize dynamics of ecologically-similar species [@Strong1983; @Fischer2001].
-Clearly, these views are not mutually exclusive, and most research has focused on determining the relative strength of biotic and abiotic forcing on the synchrony of community dynamics [e.g., @Ives1995; @Klug2000; @Houlahan2007; @Mutshinda2009; @Gross2014].
-However, recent theory [@Loreau2008a; @Loreau2013] suggests that previous interpretations of time series have been impaired by (1) a faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects and drift to estimate immediate responses to the environment and competition.
+A major roadblock to understanding community synchrony is that many proposed mechanisms lead to similar outcomes.
+Asynchronous fluctuations in species' abundances can result from interspecific competition [@Tilman2000] or species-specific responses to variable environmental conditions [@Loreau2013].
+However, the effects of interspecific competition and environmental fluctuations are not as general as they appear.
+For example, interspecific competition can have either a positive or negative effect on synchrony depending on how strongly species interact, the symmetry of those interactions, and species' growth rates [@Loreau2013].
+Likewise, environmental fluctuations can have either a positive or negative effect on synchrony depending on how similarly species respond to environmental conditions.
+Since predictions arising from the above mechanisms are not mutually exclusive [@Ranta2008], it is difficult to infer mechanism from population time series [e.g., @Houlahan2007].
 
-How should synchrony be measured?
-Traditionally, most researchers relied on summed covariances to detect compensatory, and thus asynchronous, dynamics [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
+Inferring the mechanism driving synchrony in observed population time series has also been hampered by the widespread use of summed community covariance as a metric for detecting compensatory dynamics [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
 The idea is that if the summed covariances of species' abundances through time are positive, then abiotic forces structure dynamics since species covar positively through time in response to a shared environmental driver.
@@
 In the end, the fact that few communities exhibit negative summed covariance [e.g., @Houlahan2007] yields little insight on the relative influence of competition and environmental forcing on community dynamics [@Loreau2010].
+
+However, recent theory [@Loreau2008a; @Loreau2013] suggests that previous interpretations of time series have been impaired by (1) a faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects and drift to estimate immediate responses to the environment and competition.
+
+How should synchrony be measured?
+Traditionally, most researchers relied on summed covariances to detect compensatory, and thus asynchronous, dynamics [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
 

### 2016-02-04 docs/ms_communitySynchrony.Rmd
more intro changes

@@
 ------------
-Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity stabilizes aggregate community or ecosystem function [@Ives2002; @Gonzalez2009; @Loreau2010; @Loreau2013].
-Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010; @DeMazancourt2013a; @Gross2014; @Hautier2014], but how theoretical mechanisms of species asynchrony link to observed dynamics remains unclear [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity can stabilize aggregate community or ecosystem function [@Ives2002; @Gonzalez2009; @Loreau2010; @Loreau2013].
+Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010; @DeMazancourt2013a; @Gross2014; @Hautier2014], but identifying what drives species synchrony in natural communities remains a challenge [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+The challenge lies in partitioning community synchrony among three mechanistic drivers of population fluctuations: inter and intraspecific density-dependence, environmental stochasticity, and demographic stochastcity [@Loreau2008a].
+Ideally, each mechanism would produce a unique signature on community dynamics and we could infer mechanism by comparing observed synchrony to some null expectation [e.g., @Houlahan2007].
+Unfortunately, theoretical work suggests there is no null expectation of community synchrony and the drivers of synchrony interact in complex ways that preclude inference from observations alone [@Loreau2008a; @Loreau2013].
+
+The null expectation for species synchrony has previously been defined as independent species fluctuations.
+Such a null is convenient because it allows the use of summed community covariance to infer whether competition or environmental forces drive community dynamics: if summed covariance is negative, then competition is most important; if summed covariance is positive, then environmental forces are most important [@Houlahan2007].
+However, independent species fluctuations is only expected under very specific conditions where demographic stochasticity alone is operating and all species have similar growth rates and density-dependence [@Loreau2008a; @Loreau2013].
+Likewise, values of summed covariance are not exclusive to particular combinations of drivers, meaning that negative summed covariances can occur in the absence of competition [@Ranta2008].
+Finally, summed covariances have no upper or lower limits and are sensitive to community size, making it difficult to compare different communities [@Loreau2008a].
+In the end, the fact that few communities exhibit negative summed covariance [e.g., @Houlahan2007] yields little insight on the relative influence of competition and environmental forcing on community dynamics [@Loreau2010].
+
+Theory identifies several limiting cases [@Loreau2008a], but none can be considered an expectation in natural communities where the strengths of internal and external forces are unkown.
+For example, independent species fluctuations are expected if only demographic stochasticity if operating and species all have the same abundances and variance [@Loreau2008a].
+Furthermore, asymmetry in species' growth rates, competitive effects, and density-dependence can all interact, making it difficult to identify general trends between synchrony and its drivers.
+For example, competition generally, but not always, has a positive effect on synchrony of per capita growth rates and a negative effect on synchrony of abundances [@Loreau2013].
+Since the mechanisms behind community synchrony are context-dependent, the best way forward is to analyze the drivers of synchrony in natural communities and then relate observed synchrony back to limiting theoretical cases.
+In so doing, we can discover common and rare theoretical cases in nature.
+
+<!-- COMMENT THIS OUT FOR NOW...
 Futhermore, analyses of observational time series tend to report synchronous, rather than asynchronous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
@@
 Likewise, values of summed covariance are not exclusive to particular combinations of drivers, meaning that negative summed covariances can occur in the absence of competition [@Ranta2008].
-Finally, summed covariances have no upper or lower limits and are sensitive to community size, making it difficult to compare different communities [@Loreau2008a].
-In the end, the fact that few communities exhibit negative summed covariance [e.g., @Houlahan2007] yields little insight on the relative influence of competition and environmental forcing on community dynamics [@Loreau2010].
 
@@
 The use of summed community covariance relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008a; @Loreau2010].
-Theory identifies several limiting cases [@Loreau2008a], but none can be considered an expectation in natural communities where the strengths of internal and external forces are unkown.
-For example, independent species fluctuations are expected if only demographic stochasticity if operating and species all have the same abundances and variance [@Loreau2008a].
-Furthermore, asymmetry in species' growth rates, competitive effects, and density-dependence can all interact, making it difficult to identify general trends between synchrony and its drivers.
-For example, competition generally, but not always, has a positive effect on synchrony of per capita growth rates and a negative effect on synchrony of abundances [@Loreau2013].
-Since the mechanisms behind community synchrony are context-dependent, the best way forward is to analyze the drivers of synchrony in natural communities and then relate observed synchrony back to limiting theoretical cases.
-In so doing, we can discover common and rare theoretical cases in nature.
 
-<!-- COMMENT THIS OUT FOR NOW...
 For example, 
@@
 Investigations of community synchrony should therefore focus on both endpoints (1 and 1/*S*) and seek to determine the underlying forces that drive synchrony closer to 1 or 1/*S*.
-
--->
 
@@
 However, the @Mutshinda2009 study did not include data from plant communities that have, so far, received the most evidence for a strong role for interspecific interactions in determining species synchrony [@Gross2014].
+
+-->
 

### 2016-02-04 devel/table_spp_interactions.R

@@ -91,6 +91,14 @@ for(do_vital in vital_rates){
   vital_comp_list[[do_vital]] <- comp_list
 }
 
+df <- melt(unlist(vital_comp_list))
+pieces <- strsplit(rownames(df), split = "[.]")
+df$vitalrate <- sapply(pieces, "[", 1)
+sites <- sapply(pieces, "[", 2)
+df$site <- substr(sites, start = 1, stop = nchar(sites)-1)
+df$interintra <- substr(sites, start = nchar(sites), stop = nchar(sites))
+avg_interintra <- ddply(df, .(site, interintra), summarise,
+                        avg_val = mean(value))
 
 
 ####

### 2016-02-05 docs/ms_communitySynchrony.Rmd
trying new intro

@@
 
-<!-- \setlength{\parindent}{5ex} -->
+\newpage{}
+<!-- THIS HOW TO COMMENT-OUT -->
 \setlength{\parindent}{5ex}
@@
 ------------
+The temporal synchrony of population dynamics among interacting species largely determines ecosystem stability [@Ives2002; @Loreau2013; @Gross2014] and is a function of demographic stochasticity, environmental stochasticity, and competition [@Loreau2008a; @Loreau2013].
+Strong asynchrony (also called compensatory dynamics) can emerge from negative interactions among competitors [@Lehman2000], but can also be the result of species-specific responses to environmental conditions [@Loreau2013].
+However, it may often be the case that communities are composed of species that respond similarly to environmental conditions, which can synchronize community dynamics [@Houlahan2007; @Rocha2011].
+Competition, too, may synchronize dynamics under some conditions [@Loreau2013].
+In fact, demographic stochasticity is the only force with a consistent desynchronizing effect on population dynamics [@Loreau2008a].
+Since environmental forces and competition can both lead to strong synchronous or asynchronous dynamics, deciphering the mechanisms behind species synchrony in natural communities has been a challenge.
+
+One of the largest hurdles to understanding community synchrony has been how we define synchrony.
+Unfortunately, the null expectation of independent species fluctutations and the use of summed community covariance as a metric of synchrony resulted in the perception that synchrony and asynchrony are binary states [@Ives1995; @Klug2000; @Fischer2001; @Houlahan2007], when in fact synchrony is a continuous property [@Loreau2008a].
+Fortunately, two recent theoretical advances clarify how we should view community syhnchrony.
+First, there is no null expectation of community synchrony.
+Theory identifies several limiting cases, but none can be considered null expectations [@Loreau2008a; @Loreau2013].
+For example, independent fluctuations among species is one expected theoretical outcome, but only when demographic stochasticity alone is operating and all species have equal abundances and variances [@Loreau2008a].
+Another expectation is that species fluctuate in perfect synchrony when competition and demographic stochasticity are weak and species' responses to the environment are perfectly correlated [@Loreau2008a].
+It is impossible to *a priori* assign a theoretical expectation to a natural community, suggesting we should abandon the use of any null expectation.
+Second, @Loreau2008a derived a new metric of synchrony ($\phi$) from a multi-species dynamic population model, which makes clear that synchrony and asynchrony are not binary states, contrary to their common usage in the literature.
+Rather, the synchrony of community dynamics lies along a continuum from perfect asynchrony ($\phi=0$) to perfect synchrony ($\phi=1$).
+Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony within a community and what process increase or decrease sychrony [e.g., @Gross2014].
+
+Since there is no null expectation for synchrony, no breakpoint at which dynamics can be considered asynchronous or synchronous, and values of synchrony are not unique to specific processes, it is impossible to infer the relative strength of different mechanisms driving synchrony by calculating summed covariances of population time series [e.g., @Houlahan2007].
+To overcome the limitations of analyzing community covariance alone, @Mutshinda2009 used a statistical approach to decompose fluctuations in abundance into contributions from environmental stochasticity and species' interactions.
+In agreement with Houlahan et al.'s (2007) conclusions, @Mutshinda2009 found that community dynamics were driven by positively correlated environmental responses among species.
+However, the @Mutshinda2009 study did not include data from plant communities that have, so far, received the most evidence for a strong role for interspecific interactions in determining species synchrony [@Gross2014].
+
+In this paper, we explicitly test theory on all the drivers of species synchrony by parameterizing and simulating dynamic multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
+We make no *a priori* assumptions about expected synchrony or the potential for drivers of population dynamics to increase or decrease synchrony.
+Instead, we use the fitted population models to simulate communities with and without interspecific interactions, environmental forcing, and demographic stochasticity.
+Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
+Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
+
+<!-- COMMENT THIS OUT FOR NOW...
+
+We follow the theory of @Loreau2008a, who show that in a neutral community of *S* species, expected synchrony ($\phi_r$) is defined as:
+
+\begin{equation}
+\phi_r = \frac{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(SN)}{[(1/S)(\sum_{i}\sqrt{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(N)})]^2}
+\end{equation}
+
+\todo[inline, size=\small]{Claire: I think what we actually want is the non-neutral version here, is there a simple equation for that (something tells me there is not...)?}
+
+\noindent where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
+*S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
+Their analysis isolates three main factors that drive species synchrony in competitive communities: (1) density-dependence via intra- and interspecific competition, (2) environmental fluctuations, and (3) demographic stochasticity.
+
+Loreau and de Mazancourt (2008) show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx \phi_e$, where $phi_e$ is the synchrony of species' responses to the environment.
+At the other limit, where environmental forcing and community regulation are weak relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$.
+Based on the theory of @Loreau2008a, we aimed to test the following four hypotheses and associated predictions related to the synchrony of per capita growth rates:
+
+1. Interspecific interactions synchronize population dynamics. Therefore, removing interspecific interactions will decrease synchrony.
+2.
+Plant species in semiarid ecosystems should respond in similar ways to environmental drivers.
+Therefore, removing environmental variation will decrease synchrony relative to situations were environmental variation is present.
+3.
+In populations small enough for demographic stochasticity to influence dynamics (where $\sigma^{2}_{d} \neq 0$), synchrony will equal the inverse of the number of species in the community (1/*S*), in the absence of species interactions and environmental variation ($\sigma^{2}_{c}$ and $\sigma^{2}_{e} \approx 0$).
+4. The de-synchronizing effect of demographic stochasticity will weaken as population size increases, regardless of presence of interspecific interactions or environmental stochasticity.
+
+To test these hypotheses we parameterize and simulate dynamic multi-species population models fit long term demographic data from five semi-arid grasslands [@Chu2015].
+We ran simulations directly testing the above hypotheses using a spatially-implicit integral projection model (IPM, hypothesis 1) and a spatially-explicit individual-based model (IBM, hypotheses 2-4).
+Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
+Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models according to each hypothesis should reveal the processes responsible for the degree of species synchrony in our focal grassland communities.
+
 Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity can stabilize aggregate community or ecosystem function [@Ives2002; @Gonzalez2009; @Loreau2010; @Loreau2013].
@@
 
-<!-- COMMENT THIS OUT FOR NOW...
 Futhermore, analyses of observational time series tend to report synchronous, rather than asynchronous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
@@
 -->
-
-In this paper, we seek to overcome the limitations of previous studies by explicitly testing theory on all the drivers of species synchrony using empirically-based, multi-species population models that include the effects of density-dependence, environmental stochasticity, and demographic stochasticity.
-We follow the theory of @Loreau2008a, who show that in a neutral community of *S* species, expected synchrony ($\phi_r$) is defined as:
-
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(SN)}{[(1/S)(\sum_{i}\sqrt{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(N)})]^2}
-\end{equation}
-
-\todo[inline, size=\small]{Claire: I think what we actually want is the non-neutral version here, is there a simple equation for that (something tells me there is not...)?}
-
-\noindent where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
-*S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
-Their analysis isolates three main factors that drive species synchrony in competitive communities: (1) density-dependence via intra- and interspecific competition, (2) environmental fluctuations, and (3) demographic stochasticity.
-
-Loreau and de Mazancourt (2008) show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx \phi_e$, where $phi_e$ is the synchrony of species' responses to the environment.
-At the other limit, where environmental forcing and community regulation are weak relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$.
-Based on the theory of @Loreau2008a, we aimed to test the following four hypotheses and associated predictions related to the synchrony of per capita growth rates:
-
-1. Interspecific interactions synchronize population dynamics. Therefore, removing interspecific interactions will decrease synchrony.
-2.
-Plant species in semiarid ecosystems should respond in similar ways to environmental drivers.
-Therefore, removing environmental variation will decrease synchrony relative to situations were environmental variation is present.
-3.
-In populations small enough for demographic stochasticity to influence dynamics (where $\sigma^{2}_{d} \neq 0$), synchrony will equal the inverse of the number of species in the community (1/*S*), in the absence of species interactions and environmental variation ($\sigma^{2}_{c}$ and $\sigma^{2}_{e} \approx 0$).
-4. The de-synchronizing effect of demographic stochasticity will weaken as population size increases, regardless of presence of interspecific interactions or environmental stochasticity.
-
-To test these hypotheses we parameterize and simulate dynamic multi-species population models fit long term demographic data from five semi-arid grasslands [@Chu2015].
-We ran simulations directly testing the above hypotheses using a spatially-implicit integral projection model (IPM, hypothesis 1) and a spatially-explicit individual-based model (IBM, hypotheses 2-4).
-Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
-Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models according to each hypothesis should reveal the processes responsible for the degree of species synchrony in our focal grassland communities.
 

### 2016-02-05 docs/ms_communitySynchrony.Rmd
updates with new introduction structure

--- 2016-02-04 docs/ms_communitySynchrony.Rmd
@@
 
-\textbf{Number of Tables and Figures}:  Tables;  Figures
+\textbf{Number of Tables and Figures}:  2 Tables;  4 Figures
 
@@
 ------------
+The relationship between species richness and ecosystem stability depends on species synchrony [@Gonzalez2009].
+All else being equal, the stability of ecosystem function is highest when species dynamics are asynchronous [@DeMazancourt2013a].
+Thus, a mechanistic understanding of the diversity-stability relationship requires understanding the processes that increase or decrease species synchrony in nature.
+In the simplest case, when interspecific interactions and demographic stochasticity are weak, species synchrony is entirely determined by the synchrony of species' responses to fluctuating environmental conditions [@Loreau2008a; @Loreau2013].
+Synchrony is highest when species' environmental responses are perfectly correlated, and declines as environmental responses diverge [@Ives2002; @Loreau2008a; @Loreau2013].
+Alternatively, if environmental forcing is weak, interspecific competition can lead to compensatory dynamics that desynchronize species dynamics [@Lehman2000].
+Of course, real communities are subject to some combination of environmental stochasticity, interspecific competition, and demographic stochasticity that complicate predictions from special theoretical cases [@Loreau2013].
+
+The fact that species synchrony emerges from complex interactions among species' environmental responses, species interactions, and demographic stochasticity makes it difficult to disentangle the drivers of synchrony in observational data [e.g., @Houlahan2007].
+Weak synchrony could reflect positive environmental correlations offset by strong competition, or weak environmental correlations and weak competition.
+Likewise, demographic stochasticity can induce ecological drift in small populations that mask signals of competition and the environment [@Loreau2008a; @DeMazancourt2013a].
+Therefore, inference on the processes that drive synchrony from metrics of synchrony, like summed community covariance, alone is prone to error.
+@Mutshinda2009 attempted to overcome some of the limitations of inferring process from pattern by fitting a dynamic population model to several community time series of animal abundance.
+They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
+However, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
+
+Here, we explicitly test theory on the drivers of species synchrony by parameterizing and simulating dynamic, multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
+We use the fitted population models to simulate communities with and without interspecific interactions, environmental forcing, and demographic stochasticity.
+Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
+Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
+Our main focus is on the synchrony of species' per capita growth rates since they represent a species immediate response to environmental conditions and competition.
+However, we also analyze synchrony of plant cover (a metric of abundance) to relate our findings to previous research.
+
+Our approach has several advantages.
+First, we can use the models the directly compare synchrony from communities with and without different processes operating.
+This means we do not have to rely on biased null expectations, such as independent species fluctuations, to determine if competition or environmental responses are more important -- we can test for this directly.
+Second, our models represent natural communities, rather than experimentally-assembled communities.
+This means our results may be more generalizable then results obtained from biodiversity-ecosystem function experiments where strong, but transient, competition is a hallmark [@Gross2014].
+Third, our models are based on individual-level demographic data, which allows us to accurately estimate inter and intraspecific competition.
+Lastly, we can use the models to estimate the synchrony of species' low density growth rates in the absence of interspecific competition.
+Experimental analyses must rely on species' growth in monoculture as an indicator of a species' response to the environemnt, but low density growth rates are a direct measure of a species' immediate response to environmental conditions.
+In combination, our dynamic modeling approach represents the most direct means toward unraveling the process that jointly determine community dynamics, species synchrony, and ultimately ecosystem stability.
+
+<!-- COMMENT THIS OUT FOR NOW...
+
+We follow the theory of @Loreau2008a, who show that in a neutral community of *S* species, expected synchrony ($\phi_r$) is defined as:
+
+\begin{equation}
+\phi_r = \frac{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(SN)}{[(1/S)(\sum_{i}\sqrt{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(N)})]^2}
+\end{equation}
+
+\todo[inline, size=\small]{Claire: I think what we actually want is the non-neutral version here, is there a simple equation for that (something tells me there is not...)?}
+
+\noindent where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
+*S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
+Their analysis isolates three main factors that drive species synchrony in competitive communities: (1) density-dependence via intra- and interspecific competition, (2) environmental fluctuations, and (3) demographic stochasticity.
+
+Loreau and de Mazancourt (2008) show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx \phi_e$, where $phi_e$ is the synchrony of species' responses to the environment.
+At the other limit, where environmental forcing and community regulation are weak relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$.
+Based on the theory of @Loreau2008a, we aimed to test the following four hypotheses and associated predictions related to the synchrony of per capita growth rates:
+
+1. Interspecific interactions synchronize population dynamics. Therefore, removing interspecific interactions will decrease synchrony.
+2.
+Plant species in semiarid ecosystems should respond in similar ways to environmental drivers.
+Therefore, removing environmental variation will decrease synchrony relative to situations were environmental variation is present.
+3.
+In populations small enough for demographic stochasticity to influence dynamics (where $\sigma^{2}_{d} \neq 0$), synchrony will equal the inverse of the number of species in the community (1/*S*), in the absence of species interactions and environmental variation ($\sigma^{2}_{c}$ and $\sigma^{2}_{e} \approx 0$).
+4. The de-synchronizing effect of demographic stochasticity will weaken as population size increases, regardless of presence of interspecific interactions or environmental stochasticity.
+
+To test these hypotheses we parameterize and simulate dynamic multi-species population models fit long term demographic data from five semi-arid grasslands [@Chu2015].
+We ran simulations directly testing the above hypotheses using a spatially-implicit integral projection model (IPM, hypothesis 1) and a spatially-explicit individual-based model (IBM, hypotheses 2-4).
+Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
+Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models according to each hypothesis should reveal the processes responsible for the degree of species synchrony in our focal grassland communities.
+
 Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity can stabilize aggregate community or ecosystem function [@Ives2002; @Gonzalez2009; @Loreau2010; @Loreau2013].

### 2016-02-05 docs/ms_communitySynchrony.Rmd
updates methods with expected sychrony calculations

@@
 However, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
+Thus, despite several efforts to determine the drivers of species synchrony in natural communities, our understanding remains limited.
 
+Much like research on diversity-stability relationships, empirical work has lagged behind theoretical work aimed at deepening our mechanistic understanding of species synchrony.
 Here, we explicitly test theory on the drivers of species synchrony by parameterizing and simulating dynamic, multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
-We use the fitted population models to simulate communities with and without interspecific interactions, environmental forcing, and demographic stochasticity.
+We use the fitted population models to simulate communities with and without the three processes identified by theory as drivers of species synchrony: interspecific interactions, environmental forcing, and demographic stochasticity [@Loreau2008a; @Loreau2013; @DeMazancourt2013a].
 Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
@@
 Experimental analyses must rely on species' growth in monoculture as an indicator of a species' response to the environemnt, but low density growth rates are a direct measure of a species' immediate response to environmental conditions.
-In combination, our dynamic modeling approach represents the most direct means toward unraveling the process that jointly determine community dynamics, species synchrony, and ultimately ecosystem stability.
+In combination, our dynamic modeling approach represents the most direct means toward unraveling the process that jointly determine community dynamics, species synchrony, and, ultimately, ecosystem stability.
 
@@
 ---------------------
-Our current work heavily relies on the data and analysis presented by @Chu2015.
+Our work heavily relies on the data and analysis presented by @Chu2015.
 We refer interested readers to @Chu2015 for full details on the focal plant communities, choice of species, structure of vital rate statistical models, and the ability of fitted population models to reproduce observed dynamics.
@@
 
-1. Calculate observed synchrony of per capita growth rates
+1. Calculate observed and expected synchrony of per capita growth rates and percent cover
 2. Fit statistical models for survival, growth, and recruitment
-3. Build dynamic, multi-species population models (IBMs and IPMs)
+3.
+Build dynamic, multi-species population models: a spatially-implicit integral projection model (IPM; does not include demographic stochasticity) and a spatially-explicit individual based model (IBM; includes demographic stochasticity)
 4. Remove interspecific interactions from IPM simulations to isolate the effect of environmental responses
@@
 
-All our code and the data necessary to reproduce the results reported in this paper can be found on GitHub (link) as release v1.0, and we have archived that release on Dryad (link).
+\noindent All our code and the data necessary to reproduce the results reported in this paper can be found on GitHub (link) as release v1.0, and we have archived that release on Dryad (link).
 
 #### Calculating observed synchrony
-The data consists of records for individual plant size in quadrats for each year.
+The data consist of records for individual plant size in quadrats for each year.
 To obtain estimates of percent cover for each focal species in each year, we summed the individual-level data within quadrats and then averaged percent cover, by species, over all quadrats.
 We calculated per capita growth rates as $\text{log}(x_t) - \text{log}(x_{t-1})$, where *x* is species' percent cover in year *t*.
-Using the community time series of per capita growth rates in percent cover, we calculated community synchrony using the metric of @Loreau2008a (Equation 1) in the `synchrony' package [@Gouhier2014] in R [@R2013].
+Using the community time series of per capita growth rates in percent cover, we calculated community synchrony using the metric of @Loreau2008a in the `synchrony' package [@Gouhier2014] in R [@R2013].
+Specifically, we calculated synchrony as
+
+\begin{equation}
+\phi_x = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{x_{i}})^{2}}
+\end{equation}
+
+\noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or abundance ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
+
+We chose to use the special theoretical case where only demographic stochasticity is operating and all species have similar growth rates and variances to serve as an expectation of synchrony (see Appendix X).
+Specifically, based on a multi-species dynamic population model (Appendix X), we calculate expected synchrony of per capita growth rates ($\text{E}(\phi_r)$) as
+
+\begin{equation}
+\text{E}(\phi_r) = \frac{\sum_{i}p_{i}^{-1}}{\left(\sum_{i}p_{i}^{-1/2}\right)^{2}}
+\end{equation}
+
+\noindent and expected synchrony of percent cover ($\text{E}(\phi_N)$) as
+
+\begin{equation}
+\text{E}(\phi_r) = \frac{1}{\left(\sum_{i}p_{i}^{1/2}\right)^{2}}
+\end{equation}
+
+\noindent where $p_i$ is the average frequency of species *i* ($p_i = N_i/N_Total$).
+
+The difference between observed and expected synchrony represents the combined effect of competition and environmental stochasticity.
+However, it is important to note that comparing observed and expected synchrony does not yield insight on the relative contributions of competition and environmental stochasticity.
+The difference between observed and expected synchrony serves as a metric of how far natural communities are from idealized expectations.
+To understand the influence of competition and environmental stochasticity, we must turn to using population model simulations.
 

### 2016-02-08 docs/ms_communitySynchrony.Rmd
updates abstract; calculates avg synchrony and year effects

@@
 While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance of species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
-We use stochastic, multi-species population models fit using long term datasets from five semi-arid grasslands to determine the relative importance of the three forces that drive species synchrony: environmental fluctuations, interspecific competition, and demographic stochasticity.
-Species' environmental responses were positively, but not perfectly correlated. 
-As a result, community-wide synchrony was higher when environmental variation was present rather than absent, but synchrony was not perfect.
+We used stochastic, multi-species population models fit using long term datasets from five semi-arid grasslands to determine the relative importance of the three forces that drive species synchrony: environmental fluctuations, interspecific competition, and demographic stochasticity.
+Species' environmental responses, quantified as temporal random effects in vital rate regressions, were positively correlated in each community, ranging from $\rho = 0.23$ to $\rho = 0.38$ when averaged over vital rates.
+As a result, community-wide synchrony of per capita growth rates ($\phi_r$), which ranges from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (average $\phi_r = 0.62$) rather than absent (average $\phi_r = 0.43$).
 Removing interspecific competition, which was weak in all communities, did not affect synchrony, and demographic stochasticity only played a role in small populations.
 Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony is approximately equal to the synchrony of species' environmental responses -- a result that may be quite general if natural communities tend to have large niche differences among species.
-However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments, suggesting an interplay between community assembly and the drivers of community synchrony.
+However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments, suggesting that strong transient competition early in community assembly can counteract the often synchronizing effect of environmental fluctuations.
 

### 2016-02-08 devel/plot_ibm_sims.R

@@ -157,6 +157,8 @@ singles$remove="yes"
 test <- merge(synch_df, singles, by=c("experiment", "site", "expansion", "typesynch"), all = TRUE)
 ids.out <- which(test$remove=="yes")
 synch_dfplot <- test[-ids.out,]
+avg_synchs <- ddply(synch_dfplot, .(experiment, expansion, typesynch), summarise,
+                    avg_synch = mean(synch))
 
 ### Make plot and save
 ibm_plot <- ggplot(synch_dfplot, aes(x=expansion, y=synch, color=experiment))+

### 2016-02-08 devel/plot_yreffect_corrs.R

@@ -101,7 +101,7 @@ for(do_vital in vital_rates){
 }
 cor_df <- out_df[2:nrow(out_df),]
 cor_cast <- dcast(cor_df, site~vital_rate)
-
+avg_env_resp <- apply(cor_cast[,2:4], MARGIN = 1, FUN = "mean")
 
 ####
 ####  Calculate synchrony of random year effects -------------------------------

### 2016-02-09 docs/ms_communitySynchrony.Rmd
updates intro and results sections

@@
 In the simplest case, when interspecific interactions and demographic stochasticity are weak, species synchrony is entirely determined by the synchrony of species' responses to fluctuating environmental conditions [@Loreau2008a; @Loreau2013].
-Synchrony is highest when species' environmental responses are perfectly correlated, and declines as environmental responses diverge [@Ives2002; @Loreau2008a; @Loreau2013].
-Alternatively, if environmental forcing is weak, interspecific competition can lead to compensatory dynamics that desynchronize species dynamics [@Lehman2000].
-Of course, real communities are subject to some combination of environmental stochasticity, interspecific competition, and demographic stochasticity that complicate predictions from special theoretical cases [@Loreau2013].
+Synchrony is highest when correlations among species' environmental responses are perfectly postive, intermediate when species' environmental responses are independent, and lowest when the correlations of species' environmental responses are perfectly negative [@Ives2002; @Loreau2008a; @Loreau2013].
+Of course, real communities are subject to many internal and external forces beyond environmental fluctuations that can affect synchrony and complicate predictions from special theoretical cases [@Loreau2013].
 
-The fact that species synchrony emerges from complex interactions among species' environmental responses, species interactions, and demographic stochasticity makes it difficult to disentangle the drivers of synchrony in observational data [e.g., @Houlahan2007].
+Interspecific competition has long been considered a leading driver of species asynchrony (also called compensatory dynamics) [@Elton1942; @Tilman1988; @Ives1995]. 
+If two species compete strongly, then fluctuating densities in one species leads to compensatory fluctuations in the densities of the other species [@Lehman2000].
+Such countervailing fluctuations among competing species means that interspecific competition is typically a desynchronizing force on population abundances [@Loreau2013; @Gross2014].
+However, interspecific competition can actually synchronize abundances when competition is weak and population growth rates are high [@Loreau2013].
+Thus, the prevailing view that competition tends to decrease species synchrony [@Houlahan2007] is only true under certain assumptions [@Loreau2008a].
+
+Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony in small populations.
+In the absence of all other forces, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a]. 
+Interestingly, most previous attempts to decipher the mechanisms behind species synchrony from observational time series relied on a null expectation of independent species fluctuations [e.g., @Houlahan2007].
+A null expectation based on demographic stochasticity operating alone is problematic, however, because its affect is strong only in small populations.
+In natural communities composed of large populations and exposed to environmental fluctuations and competition, independent species fluctuations should not be expected [@Loreau2008a].
+Therefore, inferring the relative importance of competition or the environment based on positive or negative deviations from independent species fluctuations will lead to biased conclusions.
+For example, @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
+Based on that result and the null expecation that species should fluctuate independently, @Houlahan2007 concluded that abiotic factors synchronize species' abundances.
+Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
+But, if instead we *a priori* assume that species fluctuate in perfect synchrony [@Loreau2008a], judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
+
+The fact that species synchrony emerges from complex interactions among species' environmental responses, species interactions, and demographic stochasticity makes it difficult to disentangle the drivers of synchrony in observational data, especially since an unbiased null expectation does not exist [@Loreau2008a].
 Weak synchrony could reflect positive environmental correlations offset by strong competition, or weak environmental correlations and weak competition.
 Likewise, demographic stochasticity can induce ecological drift in small populations that mask signals of competition and the environment [@Loreau2008a; @DeMazancourt2013a].
-Therefore, inference on the processes that drive synchrony from metrics of synchrony, like summed community covariance, alone is prone to error.
-@Mutshinda2009 attempted to overcome some of the limitations of inferring process from pattern by fitting a dynamic population model to several community time series of animal abundance.
+Even the metric used to quantify synchrony can cause confusion.
+For example, @Ranta2008 showed that the most commonly used metric of species synchrony, community covariance, can be positive or negative in the presence or absence of strong competition.
+Thus, it seems that understanding the processes leading to synchrony by analysing patterns alone is impossible.
+
+One potential way forward is to use population models fit to empirical data to understand the influence of different drivers on species synchrony.
+@Mutshinda2009 used such an approach by fitting a dynamic population model to several community time series of insect and bird abundances.
 They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
-However, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Thus, despite several efforts to determine the drivers of species synchrony in natural communities, our understanding remains limited.
+While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
+Thus, despite almost a decade of research aimed toward understanding species synchrony, we still lack direct tests that link environmental fluctuations, interspecific interactions, and demographic stochasticity to species synchrony in natural communities.
 
-Much like research on diversity-stability relationships, empirical work has lagged behind theoretical work aimed at deepening our mechanistic understanding of species synchrony.
 Here, we explicitly test theory on the drivers of species synchrony by parameterizing and simulating dynamic, multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
@@
 However, we also analyze synchrony of plant cover (a metric of abundance) to relate our findings to previous research.
+
+<!-- COMMENT THIS OUT FOR NOW...
 
@@
 In combination, our dynamic modeling approach represents the most direct means toward unraveling the process that jointly determine community dynamics, species synchrony, and, ultimately, ecosystem stability.
-
-<!-- COMMENT THIS OUT FOR NOW...
 
@@
 ---------------------
-Our work heavily relies on the data and analysis presented by @Chu2015.
+Our work relies on the data and analysis presented by @Chu2015.
 We refer interested readers to @Chu2015 for full details on the focal plant communities, choice of species, structure of vital rate statistical models, and the ability of fitted population models to reproduce observed dynamics.
@@
 
-\noindent All our code and the data necessary to reproduce the results reported in this paper can be found on GitHub (link) as release v1.0, and we have archived that release on Dryad (link).
+Our approach has several advantages.
+First, we can use the models the directly compare synchrony from communities with and without different processes operating.
+This means we do not have to rely on biased null expectations, such as independent species fluctuations, to determine if competition or environmental responses are more important -- we can test for this directly.
+Second, our models represent natural communities, rather than experimentally-assembled communities.
+This means our results may be more generalizable then results obtained from biodiversity-ecosystem function experiments where strong, but transient, competition is a hallmark [@Gross2014].
+Third, our models are based on individual-level demographic data, which allows us to accurately estimate inter and intraspecific competition.
+Lastly, we can use the models to estimate the synchrony of species' low density growth rates in the absence of interspecific competition.
+Experimental analyses must rely on species' growth in monoculture as an indicator of a species' response to the environemnt, but low density growth rates are a direct measure of a species' immediate response to environmental conditions.
+
+All our code and the data necessary to reproduce the results reported in this paper can be found on GitHub (link) as release v1.0, and we have archived that release on Dryad (link).
 
@@
 We calculated per capita growth rates as $\text{log}(x_t) - \text{log}(x_{t-1})$, where *x* is species' percent cover in year *t*.
-Using the community time series of per capita growth rates in percent cover, we calculated community synchrony using the metric of @Loreau2008a in the `synchrony' package [@Gouhier2014] in R [@R2013].
+Using the community time series of per capita growth rates or percent cover, we calculated community synchrony using the metric of @Loreau2008a in the `synchrony' package [@Gouhier2014] in R [@R2013].
 Specifically, we calculated synchrony as
@@
 \noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or abundance ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
+$\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [Loreau2008a].
 
@@
 Synchrony tended to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability (e.g., Montana).
-Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub (*Artimisia tripartita*) and three perennial grasses (Table 1).
+Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
 For per capita growth rates, all sites exhibited more synchronous dynamics than expected if only demographic stochasticity was operating (E($\phi_r$) < $\phi_r$ in Table 2).
@@
 1), meaning that we can use the models to test the importance of different drivers of synchrony in these communities.
-Simulations from the IPM show that removing interspecific interactions has little affect on synchrony of per capita growth rates or percent cover in all five communities (Fig. 2).
-The weak effect of interspecific interactions on synchrony is not surprising because interspecific interactions are weak in our focal communities (Table X1, and see @Chu2015).
-The effect of removing interspecific interactions on synchrony of per capita growth rates in IBM simulations was also small and was not affected by population size (Fig. 3, top panels).
-The only discernable effect of interspecific interactions was in the Arizona and New Mexico communities when environmental stochasticity was also removed.
-In those cases, removing interspecific interactions caused a decrease in species synchrony of per capita growth rates (Fig 3-a,e).
-Removing interspecific interactions had idosyncratic effects of the synchrony of species' percent cover and no clear patterns emerged (Fig. 3, bottom panels)
+Removing interspecific interactions from the IPMs did not affect the synchrony of per capita growth rates (Fig.
+2), likely because because interspecific interactions are weak in our focal communities (Table X1, and see @Chu2015).
+For three out five communities, removing interspecific interactions had no effect on the synchrony of percent cover (Fig. 2a,c,d).
+In the two communities where removing interspecific interactions did have an effect, synchrony of percent cover was higher when species interactions were absent (Fig. 2b,e).
+We achieved similar results with the IBM simulations (Fig. 3).
 
-### Effect of removing environmental stochasticity
-For all five communities, removing environmental stochasticity cause synchrony of per capita growth rates to decrease, whether interspecific interactions were present or not (Fig.
+Species' environmental responses, quantified as temporal random effects in vital rate regressions, were positively correlated in each community, ranging from $\rho = 0.23$ to $\rho = 0.38$ when averaged over vital rates (Appendix X).
+Thus, removing environmental stochasticity caused synchrony of per capita growth rates to decrease in all five communities, whether interspecific interactions were present or not (Fig.
 3, top panels, blue tones vs.
 orange tones).
-Without environmental stochasticity, synchrony of per capita growth rates all approximated the theoretical expectation of 1/S when only demographic stochasticity is operating (Fig. 3, top panels).
-Again, no clear pattern emerges for the effect of removing environmental stochasticity on the synchrony of percent cover (Fig. 3, bottom panels).
-Community-wide averages of vital rate random year effects are strong predictors of IPM-simulated synchrony of per capita growth rates ($\text{R}^2$ = 0.67; Fig. 4)
+Without environmental stochasticity or species interactions, synchrony of per capita growth rates all approximated the theoretical expectation of 1/*S*, where *S* is species richness, when only demographic stochasticity is operating (Fig.
+3; top panels, dark blue lines).
+In most cases, removing environmental stochasticity also led to a decrease in the synchrony of percent cover (Fig. 3; bottom panels).
+The Montana community is the clear outlier, where synchrony of percent cover is virtually the same whether environmental fluctuations are present or not (Fig. 3).
 
-### Effect of demographic stochasticity
-The strength of demographic stochasticity decreases with the size of the simulated landscape in our IBMs.
+Demographic stochasticity counteracted the synchronizing effect of the environment in only the smallest populations (Fig. 3).
+The strength of demographic stochasticity decreased with the size of the simulated landscape in our IBMs.
 Thus, in Figure 3 the strength of demographic stochasticity decreases along the x-axis.
-In the absence of environmental stochasticity, the strength of demographic stochasticity has no impact of the synchrony of per capita growth rates (Fig. 3, top panels, blue tones).
+In the absence of environmental stochasticity, the strength of demographic stochasticity has no impact of the synchrony of per capita growth rates or percent cover (Fig. 3, blue tones).
 In the presence of environmental stochasticity, synchrony of per capita growth rates is lower when demographic stochasticity is strong (Fig. 3, top panels, orange tones).
-The desynchronizing effect of demographic stochasticity is strongest in small populations (small landscapes) for all communities, but the effect is strongest in Arizona and New Mexico (Fig. 3).
+The desynchronizing effect of demographic stochasticity on per capita growth rates is strongest in small populations (small landscapes) for all communities (Fig. 3, top panels).
+Demographic stochasticity had a weaker effect on synchrony of percent cover; only in the Arizona community does strong demographic stochasticity counteract the synchronizing effect of the environment (Fig.
+3; bottom panels).
+
+Since species interactions are weak and demographic stochasticity only matters in small populations, the synchrony of species in monoculture (without interspecific interactions) should predict the synchrony of species in polyculture (with interspecific interactions).
+For all five communities and for percent cover and per capita growth rates, the synchrony of species in monoculture is a near-perfect predictor of species synchrony in polyculture (Fig. 4A).
+However, the synchrony of species' low-density growth rates, a more accurate measure of a species' environmental response, was a less accurate predictor of species synchrony in polyculture (Fig. 4B).
 
@@
 ----------
-Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations.
+Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2013a].
 We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.
@@
 3).
-This result is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics and a recent empirical analysis
+This result is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics and a recent empirical analysis [@Gross2014].
 
@@
 In other words, if there is no relationship between landscape size and population size due to strong density-dependence, the effect of demgographic stochasticity will not lessen.
-\todo[inline, size=\small]{CHECK MEAN COVER VERSUS LANDSCAPE SIZE FOR EACH COMMUNITY}
 
@@
 Removing environmental fluctuations consistently reduced synchrony (Fig.
-3) and the synchrony of species environmental responses was a strong predictor of community synchrony ($\text{R}^2$ = 0.67; Fig.
+3) and the synchrony of species environmental responses was a strong, but not perfect, predictor of community synchrony (Fig.
 4).
@@
 
-A challenge to generality of our results is that a recent analysis of several biodiversity-ecosystem function experiments suggests our results may be anomolous.
+A challenge to the generality of our results is that a recent analysis of several biodiversity-ecosystem function experiments came to a completely opposing conclusion: competition drives species synchrony in grasslands [@Gross2014].
 @Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
@@
 One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.
-
-The generality of our results only extends to communities composed of a single trophic level. \textcolor{red}{DO WE NEED THIS?}
 

### 2016-02-09 docs/ms_communitySynchrony.Rmd
fixes typos

@@
 Our approach has several advantages.
-First, we can use the models the directly compare synchrony from communities with and without different processes operating.
+First, we can use the models to directly compare synchrony from communities with and without different processes operating.
 This means we do not have to rely on biased null expectations, such as independent species fluctuations, to determine if competition or environmental responses are more important -- we can test for this directly.
@@
 \noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or abundance ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
-$\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [Loreau2008a].
+$\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [@Loreau2008a].
 
@@
 ----------
-Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2013a].
+Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2013].
 We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.

### 2016-02-09 docs/ms_communitySynchrony.Rmd
removes old commented out text; retrieve here for cached text

@@
 Interestingly, most previous attempts to decipher the mechanisms behind species synchrony from observational time series relied on a null expectation of independent species fluctuations [e.g., @Houlahan2007].
-A null expectation based on demographic stochasticity operating alone is problematic, however, because its affect is strong only in small populations.
+A null expectation based on demographic stochasticity operating alone is problematic, however, because its effect is strong only in small populations.
 In natural communities composed of large populations and exposed to environmental fluctuations and competition, independent species fluctuations should not be expected [@Loreau2008a].
@@
 However, we also analyze synchrony of plant cover (a metric of abundance) to relate our findings to previous research.
-
-<!-- COMMENT THIS OUT FOR NOW...
-
-Our approach has several advantages.
-First, we can use the models the directly compare synchrony from communities with and without different processes operating.
-This means we do not have to rely on biased null expectations, such as independent species fluctuations, to determine if competition or environmental responses are more important -- we can test for this directly.
-Second, our models represent natural communities, rather than experimentally-assembled communities.
-This means our results may be more generalizable then results obtained from biodiversity-ecosystem function experiments where strong, but transient, competition is a hallmark [@Gross2014].
-Third, our models are based on individual-level demographic data, which allows us to accurately estimate inter and intraspecific competition.
-Lastly, we can use the models to estimate the synchrony of species' low density growth rates in the absence of interspecific competition.
-Experimental analyses must rely on species' growth in monoculture as an indicator of a species' response to the environemnt, but low density growth rates are a direct measure of a species' immediate response to environmental conditions.
-In combination, our dynamic modeling approach represents the most direct means toward unraveling the process that jointly determine community dynamics, species synchrony, and, ultimately, ecosystem stability.
-
-We follow the theory of @Loreau2008a, who show that in a neutral community of *S* species, expected synchrony ($\phi_r$) is defined as:
-
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(SN)}{[(1/S)(\sum_{i}\sqrt{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(N)})]^2}
-\end{equation}
-
-\todo[inline, size=\small]{Claire: I think what we actually want is the non-neutral version here, is there a simple equation for that (something tells me there is not...)?}
-
-\noindent where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
-*S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
-Their analysis isolates three main factors that drive species synchrony in competitive communities: (1) density-dependence via intra- and interspecific competition, (2) environmental fluctuations, and (3) demographic stochasticity.
-
-Loreau and de Mazancourt (2008) show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx \phi_e$, where $phi_e$ is the synchrony of species' responses to the environment.
-At the other limit, where environmental forcing and community regulation are weak relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$.
-Based on the theory of @Loreau2008a, we aimed to test the following four hypotheses and associated predictions related to the synchrony of per capita growth rates:
-
-1. Interspecific interactions synchronize population dynamics. Therefore, removing interspecific interactions will decrease synchrony.
-2.
-Plant species in semiarid ecosystems should respond in similar ways to environmental drivers.
-Therefore, removing environmental variation will decrease synchrony relative to situations were environmental variation is present.
-3.
-In populations small enough for demographic stochasticity to influence dynamics (where $\sigma^{2}_{d} \neq 0$), synchrony will equal the inverse of the number of species in the community (1/*S*), in the absence of species interactions and environmental variation ($\sigma^{2}_{c}$ and $\sigma^{2}_{e} \approx 0$).
-4. The de-synchronizing effect of demographic stochasticity will weaken as population size increases, regardless of presence of interspecific interactions or environmental stochasticity.
-
-To test these hypotheses we parameterize and simulate dynamic multi-species population models fit long term demographic data from five semi-arid grasslands [@Chu2015].
-We ran simulations directly testing the above hypotheses using a spatially-implicit integral projection model (IPM, hypothesis 1) and a spatially-explicit individual-based model (IBM, hypotheses 2-4).
-Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
-Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models according to each hypothesis should reveal the processes responsible for the degree of species synchrony in our focal grassland communities.
-
-Asynchrony in population dynamics (also called compensatory dynamics) among interacting species is a key mechanism by which biodiversity can stabilize aggregate community or ecosystem function [@Ives2002; @Gonzalez2009; @Loreau2010; @Loreau2013].
-Experimental work in grassland communities provides strong evidence for the stabilizing effect of asynchronous population dynamics [@Hector2010; @DeMazancourt2013a; @Gross2014; @Hautier2014], but identifying what drives species synchrony in natural communities remains a challenge [@Loreau2008a; @Gonzalez2009; @Loreau2013].
-The challenge lies in partitioning community synchrony among three mechanistic drivers of population fluctuations: inter and intraspecific density-dependence, environmental stochasticity, and demographic stochastcity [@Loreau2008a].
-Ideally, each mechanism would produce a unique signature on community dynamics and we could infer mechanism by comparing observed synchrony to some null expectation [e.g., @Houlahan2007].
-Unfortunately, theoretical work suggests there is no null expectation of community synchrony and the drivers of synchrony interact in complex ways that preclude inference from observations alone [@Loreau2008a; @Loreau2013].
-
-The null expectation for species synchrony has previously been defined as independent species fluctuations.
-Such a null is convenient because it allows the use of summed community covariance to infer whether competition or environmental forces drive community dynamics: if summed covariance is negative, then competition is most important; if summed covariance is positive, then environmental forces are most important [@Houlahan2007].
-However, independent species fluctuations is only expected under very specific conditions where demographic stochasticity alone is operating and all species have similar growth rates and density-dependence [@Loreau2008a; @Loreau2013].
-Likewise, values of summed covariance are not exclusive to particular combinations of drivers, meaning that negative summed covariances can occur in the absence of competition [@Ranta2008].
-Finally, summed covariances have no upper or lower limits and are sensitive to community size, making it difficult to compare different communities [@Loreau2008a].
-In the end, the fact that few communities exhibit negative summed covariance [e.g., @Houlahan2007] yields little insight on the relative influence of competition and environmental forcing on community dynamics [@Loreau2010].
-
-Theory identifies several limiting cases [@Loreau2008a], but none can be considered an expectation in natural communities where the strengths of internal and external forces are unkown.
-For example, independent species fluctuations are expected if only demographic stochasticity if operating and species all have the same abundances and variance [@Loreau2008a].
-Furthermore, asymmetry in species' growth rates, competitive effects, and density-dependence can all interact, making it difficult to identify general trends between synchrony and its drivers.
-For example, competition generally, but not always, has a positive effect on synchrony of per capita growth rates and a negative effect on synchrony of abundances [@Loreau2013].
-Since the mechanisms behind community synchrony are context-dependent, the best way forward is to analyze the drivers of synchrony in natural communities and then relate observed synchrony back to limiting theoretical cases.
-In so doing, we can discover common and rare theoretical cases in nature.
-
-Futhermore, analyses of observational time series tend to report synchronous, rather than asynchronous, population dynamics among species of similar ecology [@Houlahan2007; @Mutshinda2009; @Rocha2011; @Vasseur2014].
-Thus, our understanding of the prevalence and drivers of asynchronous population dynamics is limited.
-Only by fully understanding the mechanisms that lead to asynchronous, or synchronous, dynamics in natural ecosystems can we predict how future biodiversity loss may impact ecosystem functioning [@Loreau2010].
-
-A major roadblock to understanding community synchrony is that many proposed mechanisms lead to similar outcomes.
-Asynchronous fluctuations in species' abundances can result from interspecific competition [@Tilman2000] or species-specific responses to variable environmental conditions [@Loreau2013].
-However, the effects of interspecific competition and environmental fluctuations are not as general as they appear.
-For example, interspecific competition can have either a positive or negative effect on synchrony depending on how strongly species interact, the symmetry of those interactions, and species' growth rates [@Loreau2013].
-Likewise, environmental fluctuations can have either a positive or negative effect on synchrony depending on how similarly species respond to environmental conditions.
-Since predictions arising from the above mechanisms are not mutually exclusive [@Ranta2008], it is difficult to infer mechanism from population time series [e.g., @Houlahan2007].
-
-Inferring the mechanism driving synchrony in observed population time series has also been hampered by the widespread use of summed community covariance as a metric for detecting compensatory dynamics [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
-The idea is that if the summed covariances of species' abundances through time are positive, then abiotic forces structure dynamics since species covar positively through time in response to a shared environmental driver.
-Alternatively, negative summed covariance should indicate that competition structures community fluctuations by driving compensatory dynamics.
-Unfortunately, reliance on such a metric created the search for binary states -- synchronous or not, compensatory dynamics or not -- where a continuum exists [@Loreau2008a; @Gross2014].
-Likewise, values of summed covariance are not exclusive to particular combinations of drivers, meaning that negative summed covariances can occur in the absence of competition [@Ranta2008].
-
-However, recent theory [@Loreau2008a; @Loreau2013] suggests that previous interpretations of time series have been impaired by (1) a faulty metric of synchrony that treated synchrony and asynchrony as binary states, (2) biased null expectations of species synchrony, and (3) the use of population abundances that are strongly influenced by lag effects and drift to estimate immediate responses to the environment and competition.
-
-How should synchrony be measured?
-Traditionally, most researchers relied on summed covariances to detect compensatory, and thus asynchronous, dynamics [e.g., @Ives1995; @Klug2000; @Fischer2001; @Houlahan2007].
-
-To overcome many of these limitations, @Loreau2008a derived a new metric of species synchrony based on a multi-species population dynamic model.
-Their proposed metric, which can be applied to per capita growth rates and abundances, is
-
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{x_{i}})^{2}}
-\end{equation}
-
-\noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or abundance ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
-The metric of @Loreau2008a makes clear that synchrony and asynchrony are not binary states, contrary to their common usage in the literature.
-Rather, the synchrony of community dynamics lies along a continuum from perfect asynchrony ($\phi=0$) to perfect synchrony ($\phi=1$).
-Thus, our focus should not be on whether or not (a)synchronous dynamics occur, but instead on the degree of (a)synchrony within a community and what process increase or decrease sychrony [e.g., @Gross2014].
-
-What is the null expectation of species synchrony in competitive communities?
-Unfortunately, there is not a single answer.
-The use of summed community covariance relied on the null expectation that species fluctuate independently [@Ranta2008], which need not be the case [@Loreau2008a; @Loreau2010].
-
-For example, 
-Theoretical work has isolated three main factors that drive species synchrony in competitive communities: density-dependence via intra- and interspecific competition, environmental fluctuations, and demographic stochasticity [@Loreau2008a; @Loreau2013].
-
-@Loreau2008a suggest there are two expectations: perfect synchrony in neutral communities where species have identical niches, and the inverse of the number of species in the community (1/*S*) when species fluctuate independently.
-Many previous investigations of community synchrony have implicitly assumped that species will fluctuate independently in the absence of temporal niche differences and that strong interspecific interactions cause compensatory dynamics.
-For example, @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
-Based on that result, @Houlahan2007 concluded that abiotic factors synchronize species' abundances.
-However, their null expectation was that species should fluctuate independently.
-Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
-But, if instead we *a priori* assume that species fluctuate in perfect synchrony [@Loreau2008a], judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
-Investigations of community synchrony should therefore focus on both endpoints (1 and 1/*S*) and seek to determine the underlying forces that drive synchrony closer to 1 or 1/*S*.
-
-What temporal variable should we use to measure synchrony?
-In constrast to virtually every study on synchrony that uses population abundance [e.g., @Houlahan2007; @Mutshinda2009; @Hector2010; @Gross2014; @Vasseur2014], @Loreau2008a suggest community-wide synchrony is best viewed through the lense of per capita growth rates.
-On conceptual grounds, per capita growth rates represent the immediate response of species to the environment and competition that we wish to attribute to synchrony.
-On practical grounds, positively covarying abundances and negatively covarying environmental responses are not mutually exclusive [@Loreau2008a; @Ranta2008].
-Therefore, measuring the synchrony of population abundances through time does not allow inference on the drivers of synchrony.
-
-Aside from issues on how to measure synchrony, disentangling the factors that drive species synchrony is difficult.
-Observational timeseries of species' abundances allows easy quantification of summed covariances [@Houlahan2007] or species synchrony [@Gross2014], but quantifying the relative influence of density-dependence, demographic stochasticity, and environmental stochasticity remains a challenge.
-Indeed, to our knowledge, only one study has attempted to separate the influences of density-dependence and environmental stochasticity directly [@Mutshinda2009].
-@Mutshinda2009 used a statistical technique to decompose synchrony of species' abundances calculated from five time series.
-They found that the effect of environmental forcing on temporal dynamics is much stronger than that of interspecific interactions, across all taxa they analyzed [@Mutshinda2009].
-However, the @Mutshinda2009 study did not include data from plant communities that have, so far, received the most evidence for a strong role for interspecific interactions in determining species synchrony [@Gross2014].
-
--->
 
@@
 \begin{equation}
-\text{E}(\phi_r) = \frac{1}{\left(\sum_{i}p_{i}^{1/2}\right)^{2}}
+\text{E}(\phi_N) = \frac{1}{\left(\sum_{i}p_{i}^{1/2}\right)^{2}}
 \end{equation}
 
-\noindent where $p_i$ is the average frequency of species *i* ($p_i = N_i/N_Total$).
+\noindent where $p_i$ is the average frequency of species *i* ($p_i = N_i/N_{Total}$).
 

### 2016-02-09 docs/ms_communitySynchrony.Rmd
adds low density growth rate to methods

@@
 
+\vspace{1em}
+
 \normalsize{Andrew T. Tredennick\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
+
+\vspace{1em}
 
@@
 
-#### Removing interspecific interactions (IPM)
+#### Removing interspecific interactions and quantifying species' environmental responses (IPM)
 To remove interspecific interactions, we simply set the off-diagonals of the interaction matrix for each vital rate regression to zero.
@@
 In a constant environment, IPM simulations asymptote at stable equilibriums for each species.
+
+One way to quantify the importance of environmental stochasticity on synchrony is to compare species synchrony when grown in monoculture versus when grown in polyculture [@Gross2014].
+The assumption is that synchrony in monoculture is representative of the synchrony of species' environmental responses.
+A more direct measure of a species' environmental response is the low density growth rate in a given year in the absence of interspecific competition because then only the environment is influencing the change in abundance [@Adler2010].
+Thus, we compare species synchrony in polyculture (equilibrium densities, with species interactions) to species synchrony in monoculture (equilibrium densities, no species interactions) and to the synchrony of low density growth rates (low densities, no species interactions).
+The difference between synchrony of per capita growth rates in monoculture at equilibrium versus low densities represents the effect of intraspecific density-dependence.
 
@@
 However, the synchrony of species' low-density growth rates, a more accurate measure of a species' environmental response, was a less accurate predictor of species synchrony in polyculture (Fig. 4B).
+Except for synchrony of per capita growth rates in the Idaho and Montana communities, synchrony of low density growth rates in monoculture underpredicted synchrony in polyculture (Fig. 4B).
 

### 2016-02-10 docs/appendix_mathsupport.Rmd
adds new appendix for mathematical support

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
   - \doublespacing
   - \usepackage[document]{ragged2e}
   - \usepackage{color}
   - \usepackage{rotating}
bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in
linkcolor: black
urlcolor: black

---
# Appendix X for TITLE: Mathematical Support

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}

\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

\renewcommand{\theequation}{X\arabic{equation}}

Derivation of Expected Synchrony
--------------------------------
Following @Loreau2013 and @DeMazancourt2013a, we define population growth, ignoring observation error, as

\begin{align}
\tilde{r_i}(t) = r_{mi} \left[ 1- \frac{\tilde{N_i}(t)+\sum_{j \neq i} \alpha_{ij}\tilde{N_j}(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{\tilde{N_i}(t)}} \right]
\end{align}

\noindent where $\tilde{N_i}(t)$ is the biomass of species *i* in year *t*, and $\tilde{r_i}(t)$ is its instantaneous growth rate in year *t* (tildes denote latent quantities that cannot be known due to observation error). 
$r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*. 
Just as in our vital rate statistical model (Appendix X), environmental stochasticity is incorporated as an additive year effect on a species' growth rate.
In the theoretical population model, environmental stochasticity is incorporated as $\sigma_{ei}\mu_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $\mu_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated.
Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first order, normal approximation [@Lande2003; @DeMazancourt2013a].
$\sigma_{di}^2$ is the demographic variance and $\mu_{di}(t)$ are independent normal variables with zero mean and unit variance.

In the general case, **C** is the variance-covariance matrix for population biomasses at steady state.
It solves the Lyapunov equation $\textbf{C}^\infty = \textbf{A} \textbf{C}^\infty \textbf{A}^T + \textbf{B}$, where

\begin{align}
A_{ij} = 
  \begin{cases}
      1-r_{mi} \frac{N_{i}^*}{K_i}, & i=j \\
      -r_{mi} \frac{N_{i}^*}{K_i} \alpha_{ij}, & i \neq j
    \end{cases}
\end{align}

\noindent and

\begin{equation}
B_{ij} = N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})
\end{equation}

\noindent Similarly, **R** is the variance-covariance matrix for population growth rates at steady state

\begin{equation}
R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} \sum_{k,l} \alpha_{ik} \alpha_{jl} C_{kl} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}}.
\end{equation}

\noindent Then, following the synchrony metric of @Loreau2008a, the sychrony of population biomasses is

\begin{equation}
\phi_N = \frac{\sum_{i,j}C_{ij}}{\left( \sum_i \sqrt{C_{ii}} \right)^2}
\end{equation}

\noindent and the synchrony of per capita growth rates is

\begin{equation}
\phi_R = \frac{\sum_{i,j}R_{ij}}{\left( \sum_i \sqrt{R_{ii}} \right)^2}.
\end{equation}

References
----------

### 2016-02-17 docs/ms_communitySynchrony.Rmd
updates results section

--- 2016-02-09 docs/ms_communitySynchrony.Rmd
@@
 We used stochastic, multi-species population models fit using long term datasets from five semi-arid grasslands to determine the relative importance of the three forces that drive species synchrony: environmental fluctuations, interspecific competition, and demographic stochasticity.
-Species' environmental responses, quantified as temporal random effects in vital rate regressions, were positively correlated in each community, ranging from $\rho = 0.23$ to $\rho = 0.38$ when averaged over vital rates.
+Species' environmental responses, quantified as temporal random effects on intercepts in vital rate regressions, were positively correlated in each community, ranging from $\rho = 0.23$ to $\rho = 0.38$ when averaged over vital rates.
 As a result, community-wide synchrony of per capita growth rates ($\phi_r$), which ranges from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (average $\phi_r = 0.62$) rather than absent (average $\phi_r = 0.43$).
@@
 Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony is approximately equal to the synchrony of species' environmental responses -- a result that may be quite general if natural communities tend to have large niche differences among species.
-However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments, suggesting that strong transient competition early in community assembly can counteract the often synchronizing effect of environmental fluctuations.
+However, our results are in contrast to a recent analysis by Gross *et al.* (2014), who found that synchrony was driven by interspecific competition rather than environmental fluctuations in recently assembled experimental communities.
+The difference between our studies suggests that strong transient competition early in community assembly can counteract the, often, synchronizing effect of environmental fluctuations.
 
@@
 If two species compete strongly, then fluctuating densities in one species leads to compensatory fluctuations in the densities of the other species [@Lehman2000].
-Such countervailing fluctuations among competing species means that interspecific competition is typically a desynchronizing force on population abundances [@Loreau2013; @Gross2014].
-However, interspecific competition can actually synchronize abundances when competition is weak and population growth rates are high [@Loreau2013].
-Thus, the prevailing view that competition tends to decrease species synchrony [@Houlahan2007] is only true under certain assumptions [@Loreau2008a].
+All else being equal, interspecific competition generally desynchronizes population abundances but sunchronizes per capita growth rates [@Loreau2008a; @Loreau2013; @Gross2014], except in the special case where competition is weak and population growth rates are high [@Loreau2013].
+When both those conditions are met, the presence of competition synchronizes population abundances [@Loreau2013].
+Nonetheless, theoretical and empirical work agree that competition should generally lead to compensatory dynamics in population abundances.
 
 Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony in small populations.
-In the absence of all other forces, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a]. 
-Interestingly, most previous attempts to decipher the mechanisms behind species synchrony from observational time series relied on a null expectation of independent species fluctuations [e.g., @Houlahan2007].
-A null expectation based on demographic stochasticity operating alone is problematic, however, because its effect is strong only in small populations.
-In natural communities composed of large populations and exposed to environmental fluctuations and competition, independent species fluctuations should not be expected [@Loreau2008a].
-Therefore, inferring the relative importance of competition or the environment based on positive or negative deviations from independent species fluctuations will lead to biased conclusions.
-For example, @Houlahan2007 suggested that asynchronous population fluctuations are rare because they found generally positive correlations in species' abundances through time across a large number of ecosystems.
-Based on that result and the null expecation that species should fluctuate independently, @Houlahan2007 concluded that abiotic factors synchronize species' abundances.
-Thus, if populations are observed to covary positively through time, then the environment must be driving such synchronization.
-But, if instead we *a priori* assume that species fluctuate in perfect synchrony [@Loreau2008a], judgment of whether species fluctate synchronously or not depends on the degree to which observed dynamics deviate from perfect synchrony.
+In the absence of all other forces, and in communities composed of species with similar growth rates, abundances, and demogaphic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
+However, the effect of demographic stochasticity is strong only in small populations, and in theory it cannot oppose, or substantially modulate, the effects of environmental fluctuations and interspecific competition [@Loreau2008a].
 
@@
 
-Here, we explicitly test theory on the drivers of species synchrony by parameterizing and simulating dynamic, multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
-We use the fitted population models to simulate communities with and without the three processes identified by theory as drivers of species synchrony: interspecific interactions, environmental forcing, and demographic stochasticity [@Loreau2008a; @Loreau2013; @DeMazancourt2013a].
+Here, we directly test the importance of proposed drivers of species synchrony by parameterizing and simulating dynamic, multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
+We use the fitted population models to simulate communities with and without the three mechanisms identified by theory as drivers of species synchrony: interspecific interactions, environmental forcing, and demographic stochasticity [@Loreau2008a; @Loreau2013; @DeMazancourt2013a].
 Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
@@
 ---------------------
-Our work relies on the data and analysis presented by @Chu2015.
+Our work relies on the data and population modeling approach presented by @Chu2015.
 We refer interested readers to @Chu2015 for full details on the focal plant communities, choice of species, structure of vital rate statistical models, and the ability of fitted population models to reproduce observed dynamics.
@@
 This means our results may be more generalizable then results obtained from biodiversity-ecosystem function experiments where strong, but transient, competition is a hallmark [@Gross2014].
-Third, our models are based on individual-level demographic data, which allows us to accurately estimate inter and intraspecific competition.
-Lastly, we can use the models to estimate the synchrony of species' low density growth rates in the absence of interspecific competition.
-Experimental analyses must rely on species' growth in monoculture as an indicator of a species' response to the environemnt, but low density growth rates are a direct measure of a species' immediate response to environmental conditions.
+Lastly, our models are based on individual-level demographic data, which allows us to accurately estimate inter and intraspecific competition.
 
@@
 This allowed species time to reach their stable size distribution.
-We then calculated the synchrony of species' percent cover and per capita growth rates (see *Calculating observed synchrony*) over 100 randomly selected contiguous 50 time step sections.
+We then calculated the synchrony of species' percent cover and per capita growth rates (see *Calculating observed synchrony*) over 100 randomly selected contiguous 50 time-step sections.
 These simulations allowed us to isolate the effects of competition and environmental fluctuations in large populations where demographic stochasticity is absent.
@@
 In a constant environment, IPM simulations asymptote at stable equilibriums for each species.
+
+A common way to quantify the importance of environmental stochasticity on synchrony is to compare species synchrony when grown in monoculture versus when grown in polyculture [@Gross2014].
+The assumption is that synchrony in monoculture is representative of the synchrony of species' environmental responses.
+Thus, we compare species synchrony in polyculture (with species interactions) to species synchrony in monoculture (equilibrium densities) based on IPM simulations where demographic stochasticity is absent.
 
@@
 Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
-Synchrony tended to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability (e.g., Montana).
+Synchrony tended to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability in percent cover (e.g., Montana).
 Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
+
+<!-- THIS HOW TO COMMENT-OUT 
 For per capita growth rates, all sites exhibited more synchronous dynamics than expected if only demographic stochasticity was operating (E($\phi_r$) < $\phi_r$ in Table 2).
@@
 In contrast, expected synchrony of percent cover was lower than expected if only demographic stochasticity was operating in two out of the five communities (Kansas and Idaho, Table 2).
+-->
 
@@
       \includegraphics{./components/poly_vs_mono_synch.png}
-  \caption{Synchrony of communities simulated without interspecific interactions (monoculture) versus synchrony of communities simulated with interspecific interactions (polyculture).
-We used the same sequence of random year effects for both simulations (mono and polyculture) to mimic biodiversity-ecosystem function experiments.
+  \caption{Synchrony of species per capita growth rates (filled circles) and percent cover (filled triangles) simulated in monoculture (without species interactions) versus synchrony in polyculture (with species interactions).
+We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem function experiments.
 The dashed line is the line of equality.}

### 2016-02-17 docs/ms_communitySynchrony.Rmd
updates figure 1 with IBM simulations; need to add in year bootstraps

@@
 
-ggplot()+
-  geom_line(data=synch_df, aes(x=value, color=Var2), stat="density", adjust=2)+
+### Get IBM synchs
+synch_df$type <- "IPM"
+synch_df <- synch_df[,2:ncol(synch_df)]
+ibm_poly <- readRDS("../devel/mono_poly_ibm.RDS")
+ibm_poly <- subset(ibm_poly, typesynch=="Per capita growth rate")
+ibm_poly$type <- "IBM"
+ibm_df <- ibm_poly[,c("site","synch","type")]
+names(ibm_df) <- names(synch_df)
+synch_df <- rbind(synch_df, ibm_df)
+
+g1 <- ggplot()+
+  geom_line(data=subset(synch_df, type=="IPM"), aes(x=value, color=Var2), 
+            stat="density", adjust=2)+
   geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
-  facet_wrap("Var2", ncol=1)+
+  facet_wrap("Var2", ncol=1, scales="free_y")+
   scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
@@
   ylab("Estimated Probability Density")+
+  ggtitle("IPM")+
   theme_bw()
+
+g2 <- ggplot()+
+  geom_line(data=subset(synch_df, type=="IBM"), aes(x=value, color=Var2), 
+            stat="density", adjust=2)+
+  geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
+  facet_wrap("Var2", ncol=1, scales="free_y")+
+  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
+  guides(color=FALSE)+
+  xlab("Community Synchrony")+
+  ylab("Estimated Probability Density")+
+  ggtitle("IBM")+
+  theme_bw()
+
+grid.arrange(g1,g2,ncol=2)
 ```

### 2016-02-17 devel/plot_ibm_sims.R

@@ -160,6 +160,10 @@ synch_dfplot <- test[-ids.out,]
 avg_synchs <- ddply(synch_dfplot, .(experiment, expansion, typesynch), summarise,
                     avg_synch = mean(synch))
 
+mono_poly <- subset(synch_dfplot, experiment=="fluctinter" & expansion==5)
+saveRDS(mono_poly, file = "../../../../../Users/atredenn/Repos/community_synchrony/devel/mono_poly_ibm.RDS")
+
+
 ### Make plot and save
 ibm_plot <- ggplot(synch_dfplot, aes(x=expansion, y=synch, color=experiment))+
   geom_point(alpha=0.5)+

### 2016-02-18 docs/ms_communitySynchrony.Rmd
updates figure 1 and captions

@@
 
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.
-Density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.", fig.width=6, fig.height=6, fig.width=3, eval=T}
+```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IPM (left panels) and the IBM (right panels).
+IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.
+IBM density curves come from 100 replicate simulations of 75 iterations each.
+Synchrony values come from simulations where environmental stochasticity and interspecific interactions are present.
+The IBM was run on a 5 by 5 meter landscape to reduce the effect of demographic stochasticity.", fig.width=6, fig.height=6, fig.width=6, eval=T}
 nyrs <- c(max(out_az$stability$num_observations),
@@
   \caption{Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
-Points represent replicate IBM simulations (out of 50 possible) in which no species went extinct.
+Points represent replicate IBM simulations (out of 100 possible) in which no species went extinct.
 Lines show linear fits between plot size (proxy for population size) and species synchrony.

### 2016-02-18 devel/plot_ibm_sims.R

@@ -7,7 +7,7 @@
 # Clear workspace 
 rm(list=ls(all=TRUE))
 
-totSims <- 50      # number of simulations per site 
+totSims <- 100   # number of simulations per site 
 totT <- 100      # time steps of simulation
 burn.in <- 25    # time steps to discard before calculating cover values
 site_colors <- c("grey45", "steelblue", "slateblue4", "darkorange", "purple")

### 2016-02-18 docs/appendix_temporalRandomEffects.Rmd
adds temporal random effects table appendix

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
   - \doublespacing
   - \usepackage[document]{ragged2e}
   - \usepackage{color}
   - \usepackage{rotating}
bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in
linkcolor: black
urlcolor: black

---
# Appendix X for TITLE: Correlations of Species' Random Year Effects

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}

\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

```{r get_randeffs, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE}
####
####  Clean workspace; Load libraries ------------------------------------------
####
rm(list = ls()) # wipe the workspace clean
library(ggplot2)
library(reshape2)
library(plyr)
library(synchrony)
library(communitySynchrony)
library(xtable)

####
####  Read in growth rate matrices ---------------------------------------------
####
pgr_list <- readRDS("../results/ipm_yearly_pgr.RDS")
site_names <- names(pgr_list)
out_df <- data.frame(site=NA, pgr_synch=NA)
for(do_site in site_names){
  tmp_pgrs <- pgr_list[[do_site]]
  tmp_synch <- as.numeric(community.sync(tmp_pgrs)[1])
  tmp_df <- data.frame(site=do_site, pgr_synch=tmp_synch)
  out_df <- rbind(out_df, tmp_df)
}
pgr_synch <- out_df[2:nrow(out_df),]

####
####  Read in regression fits --------------------------------------------------
####
all_files <- list.files("../results/")
param_files <- all_files[grep("param*", all_files)]
num_vitals <- length(param_files)

list_yr_effects <- list()
for(j in 1:num_vitals){
  param_list <- readRDS(paste0("../results/",param_files[j]))
  vital_name <- unlist(strsplit(param_files[j], "_"))[1]

  if(vital_name != "recruit"){

    site_names <- names(param_list)
    site_year_list <- list()
    for(do_site in site_names){
      tmp_site <- param_list[[do_site]]
      num_spp <- length(tmp_site) 
      tmp_yr_mat <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
      for(i in 1:num_spp){
        spp_df <- tmp_site[[i]]
        tmp_yr_mat[,i] <- spp_df$Intercept.yr
      } # end species within site loop
      site_year_list[[do_site]] <- tmp_yr_mat
    } # end site within vital rate loop

  } # end NOT recruit loops

  if(vital_name == "recruit"){

    site_names <- names(param_list)
    site_year_list <- list()
    for(do_site in site_names){
      tmp_site <- param_list[[do_site]]
      tmp_yrs <- tmp_site[grep("intcpt.yr", rownames(tmp_site)),"Mean"]
      stringlist <- strsplit(names(tmp_yrs), ",")
      num_spp <- length(unique(sapply(stringlist, "[[", 2)))
      mat_yrs <- t(matrix(tmp_yrs, nrow=num_spp, byrow = TRUE))
      site_year_list[[do_site]] <- mat_yrs
    } # end site within vital rate loop

  } # end YES recruit loops

  list_yr_effects[[vital_name]] <- site_year_list
} # end vital rate loop

####
####  Calculate average correlation of year effects ----------------------------
####
vital_rates <- names(list_yr_effects)
out_df <- data.frame(site=NA, vital_rate=NA, corr=NA)
for(do_vital in vital_rates){
  tmp_vital <- list_yr_effects[[do_vital]]
  site_names <- names(tmp_vital)
  for(do_site in site_names){
    tmp_site <- tmp_vital[[do_site]]
    tmp_cor <- cor(tmp_site)
    avg_cor <- mean(tmp_cor[upper.tri(tmp_cor)])
    tmp_out <- data.frame(site=do_site, vital_rate=do_vital, corr=avg_cor)
    out_df <- rbind(out_df, tmp_out)
  }
}
cor_df <- out_df[2:nrow(out_df),]
cor_cast <- dcast(cor_df, site~vital_rate)
avg_env_resp <- apply(cor_cast[,2:4], MARGIN = 1, FUN = "mean")

####
####  Calculate synchrony of random year effects -------------------------------
####
vital_rates <- names(list_yr_effects)
out_df <- data.frame(site=NA, vital_rate=NA, synch=NA)
for(do_vital in vital_rates){
  tmp_vital <- list_yr_effects[[do_vital]]
  site_names <- names(tmp_vital)
  for(do_site in site_names){
    tmp_site <- tmp_vital[[do_site]]
    site_synch <- as.numeric(community.sync(tmp_site)[1])
    tmp_out <- data.frame(site=do_site, vital_rate=do_vital, synch=site_synch)
    out_df <- rbind(out_df, tmp_out)
  }
}

out_df <- out_df[2:nrow(out_df),]
out_cast <- dcast(out_df, site~vital_rate)
rm(list=setdiff(ls(), c("out_cast", "out_df", "cor_df", 
                        "cor_cast", "pgr_synch")))

####
####  Save output for tables
####
outlist <- list(synch_long=out_df, synch_wide=out_cast,
                corr_long=cor_df, corr_wide=cor_cast,
                pgr_synch=pgr_synch)
```

```{r make_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
synch_cap <- "Pearson's correlations among species' temporal random effects by site and vital rate."
print(xtable(cor_cast, caption = synch_cap),
      caption.placement="top",
      include.rownames = F, 
      sanitize.colnames.function = identity,
      comment=FALSE)
```

### 2016-02-18 docs/ms_communitySynchrony.Rmd
adds temporal random effects table appendix

@@
 In the absence of all other forces, and in communities composed of species with similar growth rates, abundances, and demogaphic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
-However, the effect of demographic stochasticity is strong only in small populations, and in theory it cannot oppose, or substantially modulate, the effects of environmental fluctuations and interspecific competition [@Loreau2008a].
+However, the effect of demographic stochasticity is strong only in small populations, and in theory it cannot oppose or amplify the effects of environmental fluctuations and interspecific competition [@Loreau2008a].
 
@@
 Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations.
-These results cofirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
+These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
 Only in the Idaho community does synchrony of per capita growth rates approach 1/*S* in a non-constant environment (Fig. 3x).

### 2016-02-18 docs/appendix_speciesInteractions.Rmd
adds species interaction tables appendix

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
   - \doublespacing
   - \usepackage[document]{ragged2e}
   - \usepackage{color}
   - \usepackage{rotating}
bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in
linkcolor: black
urlcolor: black

---
# Appendix X for TITLE: Correlations of Species' Random Year Effects

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}

\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

\renewcommand{\thetable}{X\arabic{table}}

```{r get_interactions, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE}
####
####  Clean Workspace an Load Libraries ----------------------------------------
####
rm(list = ls()) # wipe the workspace clean
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)

####
####  Get Vital Rate Satistical Results Files ----------------------------------
####
vital_rates <- c("growth", "surv", "recruit")
result_files <- list.files("../results/")
vr_files <-result_files[grep(paste(vital_rates,collapse="|"), 
                              list.files("../results/"))]

####
####  Loop Over Files and Extract Crowding Effects -----------------------------
####
# Growth and survival first
vital_mat_list <- list()
for(do_vital in vital_rates[c(1,2)]){
  vr_do <- vr_files[grep(do_vital,vr_files)]
  tmp_vital <- readRDS(paste0("../results/",vr_do))
  sites <- names(tmp_vital)
  mat_list <- list()
  for(do_site in sites){
    tmp <- tmp_vital[[do_site]]
    spp <- names(tmp)
    tmpmat <- matrix(0,nrow = length(spp), ncol = length(spp))
    for(i in 1:length(spp)){
      do_spp <- spp[i]
      tmpspp <- tmp[[do_spp]]
      tmpw <- tmpspp[1,grep("crowd*", colnames(tmpspp))]
      tmpmat[i,] <- as.numeric(tmpw)
    }
    rownames(tmpmat) <- spp
    colnames(tmpmat) <- spp
    mat_list[[do_site]] <- tmpmat
  }
  vital_mat_list[[do_vital]] <- mat_list 
}

# Recruitment
vr_do <- vr_files[grep("recruit",vr_files)]
tmp_vital <- readRDS(paste0("../results/",vr_do))
names_vital <- readRDS(paste0("../results/",vr_files[1]))
sites <- names(tmp_vital)
mat_list <- list()
for(do_site in sites){
  tmp <- tmp_vital[[do_site]]
  spp <- names(names_vital[[do_site]])
  tmpdd <- tmp[grep("dd", rownames(tmp)),"Mean"]
  tmpmat <- matrix(tmpdd,length(spp),length(spp), byrow = FALSE)
  rownames(tmpmat) <- spp
  colnames(tmpmat) <- spp
  mat_list[[do_site]] <- tmpmat
}
vital_mat_list[["recruit"]] <- mat_list 
```

```{r make_tables, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
names(vital_mat_list) <- c("growth", "survival", "recruitment")
for(do_vital in names(vital_mat_list)){
  tmp_list <- vital_mat_list[[do_vital]]
  for(do_site in names(tmp_list)){
    site_list <- tmp_list[[do_site]]
    synch_cap <- paste0("Interaction coefficients for ", do_vital, " regressions in ", do_site, ".")
    print(xtable(site_list, caption = synch_cap),
      caption.placement="top",
      include.rownames = TRUE, 
      sanitize.colnames.function = identity,
      comment=FALSE)
  }
}
```

### 2016-02-18 docs/ms_communitySynchrony.Rmd
adds rank correlations to results section

@@
 
-\noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or abundance ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
+\noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or percent cover ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
 $\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [@Loreau2008a].
@@
 We ran a series of simulations for four "experiments": (1) environmental stochasticity on and interspecific interactions present, (2) environmental stochasticity off and interspecific interactions present, (3) environmental stochasticity on and interspecific interactions absent, and (4) environmental stochasticity off and interspecific interactions absent (only demographic stochasticity).
-We ran IBM simulations for 100 time steps, and repeated the simulations 50 times for each "experiment" (interspecific interactions on/off with and without environmental stochasticity).
+We ran IBM simulations for 100 time steps, and repeated the simulations 100 times for each "experiment" (interspecific interactions on/off with and without environmental stochasticity).
 From those, we retained only the simulations in which no species went extinct due to demographic stochasticity.
@@
 1), meaning that we can use the models to test the importance of different drivers of synchrony in these communities.
-Removing interspecific interactions from the IPMs did not affect the synchrony of per capita growth rates (Fig.
-2), likely because because interspecific interactions are weak in our focal communities (Table X1, and see @Chu2015).
-For three out five communities, removing interspecific interactions had no effect on the synchrony of percent cover (Fig. 2a,c,d).
-In the two communities where removing interspecific interactions did have an effect, synchrony of percent cover was higher when species interactions were absent (Fig. 2b,e).
-We achieved similar results with the IBM simulations (Fig. 3).
+However, IBM-simulated synchrony is consistently lower than IPM-simulated synchrony (Fig. 1), likely due to the desynchronizing effect of demographic stochasticity.
+Removing interspecific interactions from the IPMs had little effect on the synchrony of per capita growth rates (Fig.
+2) because interspecific interactions are weak in our focal communities (Table X1, and see @Chu2015).
+Only in Arizona does removing species interactions result in a small decline in synchrony of per capita growth rates (Fig. 2a).
+For two out five communities, removing interspecific interactions had little effect on the synchrony of percent cover (Fig. 2c,d; Kansas and Montana).
+In the three communities where removing interspecific interactions did have an effect (Arizona, Idaho and New Mexico), synchrony of percent cover was higher when species interactions were absent (Fig.
+2a,b,e).
+We achieved similar results with the IBM simulations: removing interspecific interactions had little effect on synchrony of per capita growth rates, but increased synchrony of percent cover in Idaho and New Mexico (Fig.
+3).
+In Idaho, however, synchrony of per capita growth rates is similar among all simulations with species interactions present, regardless of the presence of environmental fluctuations (Fig. 3b).
 
 Species' environmental responses, quantified as temporal random effects in vital rate regressions, were positively correlated in each community, ranging from $\rho = 0.23$ to $\rho = 0.38$ when averaged over vital rates (Appendix X).
-Thus, removing environmental stochasticity caused synchrony of per capita growth rates to decrease in all five communities, whether interspecific interactions were present or not (Fig.
-3, top panels, blue tones vs.
-orange tones).
+Thus, removing environmental stochasticity caused synchrony of per capita growth rates to decrease in all five communities (Fig. 3a-e).
+In our simulations synchrony of per capita growth rates was higher when environmental variation was present (average $\phi_r = 0.62$) rather than absent (average $\phi_r = 0.43$).
+This is true whether species interactions are present or not, except for Idaho (Fig. 3b).
+Removing environmental fluctuations also decreased synchrony of percent cover, except in Idaho and Montana (Fig. 3f-j).
+In Idaho, synchrony of percent cover is similar among simulations with and without environmental fluctuations when species interactions are not present, but synchrony is lowest in simulation where species interactions are present and environmental fluctuations are absent (Fig.
+3g).
+In Montana, all simulations result in the same values of synchrony of percent cover (Fig. 3h).
+
 Without environmental stochasticity or species interactions, synchrony of per capita growth rates all approximated the theoretical expectation of 1/*S*, where *S* is species richness, when only demographic stochasticity is operating (Fig.
 3; top panels, dark blue lines).
-In most cases, removing environmental stochasticity also led to a decrease in the synchrony of percent cover (Fig. 3; bottom panels).
-The Montana community is the clear outlier, where synchrony of percent cover is virtually the same whether environmental fluctuations are present or not (Fig. 3).
-
 Demographic stochasticity counteracted the synchronizing effect of the environment in only the smallest populations (Fig. 3).
@@
 Thus, in Figure 3 the strength of demographic stochasticity decreases along the x-axis.
-In the absence of environmental stochasticity, the strength of demographic stochasticity has no impact of the synchrony of per capita growth rates or percent cover (Fig. 3, blue tones).
-In the presence of environmental stochasticity, synchrony of per capita growth rates is lower when demographic stochasticity is strong (Fig. 3, top panels, orange tones).
-The desynchronizing effect of demographic stochasticity on per capita growth rates is strongest in small populations (small landscapes) for all communities (Fig. 3, top panels).
+In the absence of environmental stochasticity and species interactions, the strength of demographic stochasticity has no impact of the synchrony of per capita growth rates or percent cover (Fig.
+3; dark blue lines).
+In the presence of environmental stochasticity, synchrony of per capita growth rates is lower when demographic stochasticity is strong (Fig. 3a-e; orange tones).
+The desynchronizing effect of demographic stochasticity on per capita growth rates is strongest in small populations (small landscapes) for all communities (Fig. 3a-e).
 Demographic stochasticity had a weaker effect on synchrony of percent cover; only in the Arizona community does strong demographic stochasticity counteract the synchronizing effect of the environment (Fig.
@@
 Since species interactions are weak and demographic stochasticity only matters in small populations, the synchrony of species in monoculture (without interspecific interactions) should predict the synchrony of species in polyculture (with interspecific interactions).
-For all five communities and for percent cover and per capita growth rates, the synchrony of species in monoculture is a near-perfect predictor of species synchrony in polyculture (Fig. 4A).
+For all five communities and for percent cover (rank correlation = 0.8) and per capita growth rates (rank correlation = 1), the synchrony of species in monoculture is strongly associated with species synchrony in polyculture (Fig.
+4).
+
+<!--
 However, the synchrony of species' low-density growth rates, a more accurate measure of a species' environmental response, was a less accurate predictor of species synchrony in polyculture (Fig. 4B).
 Except for synchrony of per capita growth rates in the Idaho and Montana communities, synchrony of low density growth rates in monoculture underpredicted synchrony in polyculture (Fig. 4B).
+-->
 

### 2016-02-18 devel/plot_ipm_sims.R

@@ -69,6 +69,14 @@ print(ipm_pgr_plot)
 dev.off()
 
 
+####
+####  Calculate some correlations
+####
+cor(subset(polymono_wide, typesynch=="pgr_synch")[,"ENVINTER"],
+    subset(polymono_wide, typesynch=="pgr_synch")[,"ENVNOINTER"], method = "spearman")
+cor(subset(polymono_wide, typesynch=="abund_synch")[,"ENVINTER"],
+    subset(polymono_wide, typesynch=="abund_synch")[,"ENVNOINTER"], method = "spearman")
+
 
 ####
 ####  Plot synchrony in polyculture vs synchrony in monoculture

### 2016-02-19 docs/ms_communitySynchrony.Rmd
updates citations in intro

@@
 ------------
-The relationship between species richness and ecosystem stability depends on species synchrony [@Gonzalez2009].
+The relationship between species richness and ecosystem stability in grasslands depends on species synchrony [@Tilman2006; @Isbell2009; @Hector2010; @Hautier2014].
 All else being equal, the stability of ecosystem function is highest when species dynamics are asynchronous [@DeMazancourt2013a].
@@
   ggtitle("IPM")+
+  scale_x_continuous(limits=c(0,1))+
   theme_bw()
@@
   ggtitle("IBM")+
+  scale_x_continuous(limits=c(0,1))+
   theme_bw()

### 2016-02-19 docs/ms_communitySynchrony.Rmd
updates conclusions

@@
 ### Demographic stochasticity decreases synchrony
-In communities where environmental forcing and community regulation by inter and intraspecific density-dependence are weak, species should fluctuate independently and community synchrony is expected to equal the inverse of the number of species in the community (1/*S*).
-Such independent fluctuations are expected for per capita growth rates regardless of population sizes, whereas synchrony of percent cover should trend toward one (perfect synchrony) as population sizes increase and the effect of demographic stochasticity is diminished.
-Our results conform to expectations: for each community except Arizona, community synchrony of per capita growth rates are approximately equal to 1/*S* when only demographic stochasticity is acting (Fig.
-3; top panels, dark purple lines) and synchrony of percent cover starts at approximately 1/*S* and increases linearly as population sizes increase (Fig.
-3; bottom panels, dark purple lines).
-In the Arizona community, synchrony of per capita growth rates is approximately 1/*S* as expected, but synchrony of precent cover does not increase with population size (Fig. 3f).
-This could indicate that intraspecific density-dependence is abnormally strong in the Arizona community, relative to the other communities we studied.
-Strong intraspecific density-dependence for the two species in the Arizona community would limit population size regardless of landscape size.
-Therefore, populations would not grow large enough to limit the desynchronizing effect of demographic stochasticity.
-In other words, if there is no relationship between landscape size and population size due to strong density-dependence, the effect of demgographic stochasticity will not lessen.
-
-Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small.
+Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small, in agreement with theory [@Loreau2008a].
 Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations.
@@
 ### Environmental fluctuations drive community synchrony
-In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations.
+In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations [@Loreau2013].
 Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment.
 We find that environmental fluctuations drive synchrony of per capita growth rates and plant percent cover in all five communities.
-Removing environmental fluctuations consistently reduced synchrony (Fig.
-3) and the synchrony of species environmental responses was a strong, but not perfect, predictor of community synchrony (Fig.
-4).
-Likewise, removing interspecific interactions resulted in no discernable change in community synchrony of per capita growth rates (Fig. 2).
-The lack of change indicates a strong correlation between species growing in polyculture (with interspecific interactions) and in monoculture (without interspecific interactions).
-If we consider sychrony in monoculture to represent the synchrony of species' responses to the environment, then our results lead to only one conclusion: environmental fluctuations are the primary driver of community synchrony.
+Removing environmental fluctuations consistently reduced synchrony (Fig. 3) and the synchrony of species in monoculture is a strong predictor of synchrony of species in polyculture (Fig. 4).
+Likewise, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
+If we consider sychrony in monoculture to represent the synchrony of species' responses to the environment, then our results lead to only one conclusion: environmental fluctuations are the primary driver of community-wide species synchrony.
 This is unsurprising since we know that species interactions are weak in these communities.
 More surprising is that our results appear to be quite general: across five arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
-Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results are likely very general when applied to natural grasslands.
+Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural grasslands.
 
@@
 ----------
-We used empirically-based, multi-species population models to test predictions arising from theory on the drivers of species synchrony.
-Our results are consistent across five perennial plant communities and show that community-wide synchrony is predominantly driven by environmental forcing.
-Thus, we find support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak.
+Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
+When we removed environmental variation, we found that synchrony decreased (Fig. 3).
+A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
+Such a conclusion is only half correct.
+The perennial grasses we studied do have similar responses to the environment (Table X1-X10), which will tend to synchronize dynamics.
+However, if community-wide species synchrony is less than 1, some degree of compensatory dynamics must be present [@Loreau2008a].
+Since observed and simulated synchrony is less than 1 in all our focal communities (Fig. 1), we conclude that compensatory dynamics are common in natural grasslands.
+But, in agreement with other studies, we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
+
+We found strong support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak, since removing environmental fluctuations was the only simulate that affected synchrony.
 We have reason to believe that our results are quite general since many studies report large niche differences, and thus weak interactions, in plant communities.
-However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony.
-We conclude that the difference between the Gross et al.
-(2014) study and our own is due to community assembly processes: recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
+However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
+The difference between the Gross et al.
+(2014) study and our own may be due to community assembly processes: recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
 Viewing community synchrony through the lens of community assembly processes may provide a way to reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.

### 2016-02-19 docs/ms_communitySynchrony.Rmd
adds letters to figures; updates conclusion section and discussion with rare species caveats

@@
 
-\textbf{Corresponding Author}: Andrew Tredennick \\  
-Address: Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA \\
-Phone: +1-970-443-1599 \\
-Fax: +1-435-797-3796 \\
-Email: atredenn@gmail.com
+\textbf{Corresponding Author}: \\
+\hspace{2em}Andrew Tredennick \\  
+\hspace{2em}Department of Wildland Resources and the Ecology Center\\
+\hspace{2em}Utah State University \\
+\hspace{2em}5230 Old Main Hill \\
+\hspace{2em}Logan, Utah 84322 USA \\
+\hspace{2em}Phone: +1-970-443-1599 \\
+\hspace{2em}Fax: +1-435-797-3796 \\
+\hspace{2em}Email: atredenn@gmail.com
 
@@
 
-Interspecific competition has long been considered a leading driver of species asynchrony (also called compensatory dynamics) [@Elton1942; @Tilman1988; @Ives1995]. 
+Interspecific competition has long been considered a leading driver of species asynchrony (also called compensatory dynamics) [@Tilman1988; @Ives1995]. 
 If two species compete strongly, then fluctuating densities in one species leads to compensatory fluctuations in the densities of the other species [@Lehman2000].
@@
 Nonetheless, theoretical and empirical work agree that competition should generally lead to compensatory dynamics in population abundances.
+However, competition-induced compensatory dynamics may or may not increase ecosystem stability, depending on how much competition destabilizes population dynamics [@Loreau2013].
 
-Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony in small populations.
+Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony.
 In the absence of all other forces, and in communities composed of species with similar growth rates, abundances, and demogaphic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
@@
 While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Thus, despite almost a decade of research aimed toward understanding species synchrony, we still lack direct tests that link environmental fluctuations, interspecific interactions, and demographic stochasticity to species synchrony in natural communities.
+Despite almost a decade of research aimed toward understanding species synchrony, we still lack direct tests that link environmental fluctuations, interspecific interactions, and demographic stochasticity to species synchrony in natural communities.
 
@@
 ----------
-Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2013].
+Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2008a].
 We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.
-Our study produced three main findings that were consistent across five plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony since species-specific responses to the environment were positively, but not completely, correlated.
+Our study produced three main findings that were consistent across five plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
 Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community synchrony approximately equals the synchrony of species' responses to the environment.
@@
 
-A challenge to the generality of our results is that a recent analysis of several biodiversity-ecosystem function experiments came to a completely opposing conclusion: competition drives species synchrony in grasslands [@Gross2014].
+One challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
+Rare species, which our analysis does not include, could be kept at low densities by competitive suppression due to strong interspecific interactions.
+If we had been able to model rare species, we may have found a larger effect of competition on community synchrony.
+However, rare species may occupy rare niches that do not overlap with other species, in which case synchrony would remain predominantly driven by environmental responses.
+Rare species could also be limited by species-specific natural enemies whose effects are density-independent [@Mordecai2011].
+Under such conditions, synchrony will depend on the degree of host-specificity and the pathogen's response to environmental conditions.
+Niether our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].
+
+Another challenge to the generality of our results is a recent analysis of several biodiversity-ecosystem function experiments that came to a completely opposing conclusion: competition drives species synchrony in grasslands [@Gross2014].
 @Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
@@
 
-Our analysis also contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
+Our analysis contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
 Functional traits define how a species will respond to its environment.
@@
 
-Conclusion
-----------
+Conclusions
+-----------
 Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
@@
 Since observed and simulated synchrony is less than 1 in all our focal communities (Fig. 1), we conclude that compensatory dynamics are common in natural grasslands.
-But, in agreement with other studies, we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
+But, in agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
 
-We found strong support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak, since removing environmental fluctuations was the only simulate that affected synchrony.
+We found strong support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak, since removing environmental fluctuations was the only simulation that affected synchrony.
 We have reason to believe that our results are quite general since many studies report large niche differences, and thus weak interactions, in plant communities.

### 2016-02-22 docs/ms_communitySynchrony.Rmd
adds to competition discussion section

@@
 Our study produced three main findings that were consistent across five plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
-Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community synchrony approximately equals the synchrony of species' responses to the environment.
+Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community-wide species synchrony is determined by the synchrony of species' responses to environmental conditions.
 
 ### Interspecific interactions had little impact on community synchrony
-We expected community synchrony of per capita growth rates to decrease and synchrony of plant cover to increase when we removed interspecific interactions.
+We expected community synchrony of per capita growth rates to decrease and synchrony of plant cover to increase when we removed interspecific interactions [@Loreau2008a and Appendix X].
 In the absence of demographic stochasticity, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions for both per capita growth rates and percent cover (Fig.
@@
 The only notable exception is the Idaho community where, as expected, synchrony of percent cover increased when we removed interspecific interactions.
-Instead, we observed no effect of removing interspecific interactions in the presence of environmental fluctuations, regardless of the presence or absence of demographic stochasticity (Figs. 2,3).
 When we also removed environmental fluctuations, removing interspecific interactions decreased synchrony of per capita growth rates in three of our five communities and increased synchrony of per capita growth rates in two communities (Fig.
 3).
-This result is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics and a recent empirical analysis [@Gross2014].
+The lack of an effect of interspecific competition on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
+
+Our results do not contradict the idea that competition can lead to compensatory dynamics, but they do highlight the fact that interspecific competition must be relatively strong to influence species synchrony.
+The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables X1-X10).
+Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap increases.
+In that sense, our results are not surprising: interspecific interactions are weak, so of course removing them does not affect synchrony.
+More surprising is that our focal communities have little niche overlap. 
+But, as discussed by @Chu2015, little niche overlap becomes less surprising when we consider the virtually limitless niche axes upon which species can partition themselves [@Clark2010].
 

### 2016-02-26 docs/ms_communitySynchrony.Rmd
adds observed vs predicted synchrony plot

@@
 \begin{center}
-\large{\textbf{Disentangling the drivers of species synchrony in natural grasslands: environmental forcing, demographic stochasticity, and interspecific interactions}}
+\large{\textbf{Disentangling the drivers of species synchrony in natural plant communities: environmental forcing, demographic stochasticity, and interspecific interactions}}
 
@@
 
-\textbf{Running Title}: Drivers of species synchrony in natural grasslands
+\textbf{Running Title}: Drivers of species synchrony in natural plant communities
 
@@
 
-Introduction
+INTRODUCTION
 ------------
@@
 Thus, it seems that understanding the processes leading to synchrony by analysing patterns alone is impossible.
+So far, no study has rigorously integrated theoretical and empirical approaches to definitively elucidate the mechanisms underlying community-wide species synchrony in natural communities.
 
+In this paper, we refine previous theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013] and species synchrony [@Loreau2008a].
+We use the theory to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
+We then confront our theoretical predictions with data from five semi-arid plant communities by using empirically-based population models.
+The fitted population models allow us to estimate species dynamics in monoculture and, thus, test our predictions.
+We also use the population models to directly manipulate the three factors that jointly determine species synchrony: interspecific interactions, environmental forcing, and demographic stochasticity.
+Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
+
+<!--
 One potential way forward is to use population models fit to empirical data to understand the influence of different drivers on species synchrony.
@@
 However, we also analyze synchrony of plant cover (a metric of abundance) to relate our findings to previous research.
+-->
 
-Materials and Methods
+MATERIALS AND METHODS
 ---------------------
+### Theoretical model and predictions
+Following @Loreau2013 and @DeMazancourt2013a, we define population growth, ignoring observation error, as
+
+\begin{align}
+r_i(t) = r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{N_i(t)}} \right]
+\end{align}
+
+\noindent where $N_i(t)$ is the biomass of species *i* in year *t*, and $r_i(t)$ is its instantaneous growth rate in year *t*. 
+$r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*.
+Environmental stochasticity is incorporated as $\sigma_{ei}\mu_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $\mu_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated.
+Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first order, normal approximation [@Lande2003; @DeMazancourt2013a].
+$\sigma_{di}^2$ is the demographic variance and $\mu_{di}(t)$ are independent normal variables with zero mean and unit variance.
+
+CLAIRE'S TEXT ON DERIVING THE PREDICTION...
+
+We derived predictions for synchrony of growth rates and synchrony of abundance for two limiting cases that are analytically tractable.
+First, we assume no interspecific interactions, no environmental stochasticity, identical intrinsic growth rates, and that demographic stochasticity is operating but all species have identical demographic variances.
+This limiting case, $\mathcal{M}_{D}$, represents a community where dynamics are driven by demographic stochasticity alone.
+Our predictions are
+
+\begin{equation}
+\phi_{N,\mathcal{M}_{D}} = \frac{1}{\left(\sum_i p_i^{1/2} \right)^2}
+\end{equation}
+
+\noindent{} for synchrony of population abundances and
+
+\begin{equation}
+\phi_{R,\mathcal{M}_{D}} = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
+\end{equation}
+
+\noindent for synchrony of per capita growth rates.
+In both equations, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
+When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a].
+
+Our second limiting case assumes only environmental stochasticity is operating.
+Thus, there are no interspecific interactions, demographic stochasticity is absent, intrinsic growth rates are identical, and environmental variance is identical for all species.
+Our predictions for this limiting case, $\mathcal{M}_{E}$, are
+
+\begin{equation}
+\phi_{N,\mathcal{M}_{E}} = \sum_{i,j} p_i p_j \text{cov}(\mu_{ei},\mu_{ej})
+\end{equation}
+
+\noindent{} for synchrony of population abundances and
+
+\begin{equation}
+\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(\mu_{ei},\mu_{ej})}{S^2}
+\end{equation}
+
+\noindent{} for synchrony of per capita growth rates.
+In both equations, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$, and $\text{cov}(\mu_{ei},\mu_{ej})$ is the standardize covariance (correlation) of environmental responses between species *i* and species *j*.
+Confronting our theoretical predictions with data requires estimates of species dynamics in monoculture to calculate the covariance of species' environmental responses.
+To estimate environmental responses in natural communities, we turn to our population models built using long term demographic data.
+
+<!--
 Our work relies on the data and population modeling approach presented by @Chu2015.
@@
 We offer a brief overview here, but note that the only difference between our vital rate models and those fit by @Chu2015 is that we allowed for positive species interactions.
+-->
 
-### Data
+### Empirically-based population models
+#### Data
 In the early 20th century, several range experiment stations in the western U.S. began mapping permanent quadrats, and conducted annual censuses for decades. 
@@
 
+<!--
 ### Analytical approach
@@
 
+-->
+
 #### Calculating observed synchrony
@@
 
+<!--
 We chose to use the special theoretical case where only demographic stochasticity is operating and all species have similar growth rates and variances to serve as an expectation of synchrony (see Appendix X).
@@
 To understand the influence of competition and environmental stochasticity, we must turn to using population model simulations.
+-->
 

### 2016-02-26 devel/prediction_obs_sims_plot.R

@@ -1,0 +1,225 @@
+##  R script to plot comparisons between observed/simulated synchrony and
+##  expectations under two limiting conditions
+
+rm(list=ls())
+
+library(ggplot2)
+library(ggthemes)
+library(plyr)
+library(reshape2)
+library(communitySynchrony)
+library(synchrony)
+
+####
+#### Observed synchrony (cover and growth rates)
+####
+site <- "Kansas"
+spp_list <- c("BOCU","BOHI","SCSC")
+num_spp <- length(spp_list)
+ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  ks_data <- rbind(ks_data, spp_data)
+} #end species looping for raw data
+ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
+
+tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
+tmp2<-which(ks_data$quad_data=="q27")
+tmp3<-which(ks_data$quad=="q28")
+tmp4<-which(ks_data$quad=="q30")
+tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
+tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
+tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
+ks_data<-ks_data[-tmp,]
+
+# exclude the records later than 1968, to keep the same random year effect...
+ks_data<-subset(ks_data,year<68)
+
+
+site <- "Idaho"
+spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
+num_spp <- length(spp_list)
+id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  id_data <- rbind(id_data, spp_data)
[…]

### 2016-02-26 runscripts/envstoch_predictions.R

@@ -90,6 +90,7 @@ for(do_site in site_names){
   synchrony_predictions <- rbind(synchrony_predictions, tmpdf)
 }
 
-
+synchrony_predictions[2:nrow(synchrony_predictions),]
+saveRDS(synchrony_predictions,"../results/envstoch_predictions.RDS")
 
 

### 2016-02-26 docs/ms_communitySynchrony.Rmd
updates predictions figure and adds text for theory section

@@
 
-Abstract
--------------
-Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time. 
-While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance of species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
-We used stochastic, multi-species population models fit using long term datasets from five semi-arid grasslands to determine the relative importance of the three forces that drive species synchrony: environmental fluctuations, interspecific competition, and demographic stochasticity.
-Species' environmental responses, quantified as temporal random effects on intercepts in vital rate regressions, were positively correlated in each community, ranging from $\rho = 0.23$ to $\rho = 0.38$ when averaged over vital rates.
-As a result, community-wide synchrony of per capita growth rates ($\phi_r$), which ranges from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (average $\phi_r = 0.62$) rather than absent (average $\phi_r = 0.43$).
-Removing interspecific competition, which was weak in all communities, did not affect synchrony, and demographic stochasticity only played a role in small populations.
-Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony is approximately equal to the synchrony of species' environmental responses -- a result that may be quite general if natural communities tend to have large niche differences among species.
-However, our results are in contrast to a recent analysis by Gross *et al.* (2014), who found that synchrony was driven by interspecific competition rather than environmental fluctuations in recently assembled experimental communities.
-The difference between our studies suggests that strong transient competition early in community assembly can counteract the, often, synchronizing effect of environmental fluctuations.
+\begin{abstract}
+Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem function, but elucidating the mechanisms underlying species synchrony remains a challenge. 
+Here, we extend theory to predict species synchrony under limiting cases that we couple with an empirical analysis from five plant communities.
+When parameterized with empirical estimates, our prediction under the limiting case of only environmental stochasticity operating best explains observed synchrony of per capita growth rates.
+In model simulations, average synchrony of per capita growth rates, which ranges from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Removing interspecific competition and demographic stochasticity had small effects on synchrony.
+Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony is approximately equal to the synchrony of species' environmental responses.
+\end{abstract}
 
@@
 
-In this paper, we refine previous theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013] and species synchrony [@Loreau2008a].
+In this paper, we refine previous theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a].
 We use the theory to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
@@
 
-MATERIALS AND METHODS
+THEORETICAL MODEL
 ---------------------
-### Theoretical model and predictions
-Following @Loreau2013 and @DeMazancourt2013a, we define population growth, ignoring observation error, as
+### Materials and methods
+Following @Loreau2008a and @DeMazancourt2013a, we define population growth, ignoring observation error, as
 
 \begin{align}
-r_i(t) = r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{N_i(t)}} \right]
+r_i(t) &= \text{ln}N_i(t+1) - \text{ln}N_i(t) \\
+&= r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{N_i(t)}} \right]
 \end{align}
@@
 
+### Results
 We derived predictions for synchrony of growth rates and synchrony of abundance for two limiting cases that are analytically tractable.
@@
 
-### Empirically-based population models
+EMPIRICAL ANALYSIS
+------------------
+### Materials and methods
 #### Data
@@
 
-Results
--------
+### Results
+
 Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
@@
 For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig.
-1), meaning that we can use the models to test the importance of different drivers of synchrony in these communities.
-However, IBM-simulated synchrony is consistently lower than IPM-simulated synchrony (Fig. 1), likely due to the desynchronizing effect of demographic stochasticity.
+X1), meaning that we can use the models to test the importance of different drivers of synchrony in these communities.
+However, IBM-simulated synchrony is consistently lower than IPM-simulated synchrony (Fig. X1), likely due to the desynchronizing effect of demographic stochasticity.
+
+Across the five communities, our limiting case prediction of community-wide synchrony when only environmental stochasticity is operating best matched observed synchrony for per capita growth rates (Fig.
+1).
+The regression line between observed growth rate synchrony and predicted growth rate synchrony under prediction $\mathcal{M}_E$ was the only one whose slope and intercept were not significantly different from zero (*P* = 0.88) and one (*P* = 0.87), respectively (Supplementary Information X).
+Our prediction under $\mathcal{M}_E$ for percent cover did not fair as well, with a slope and intercept that were significantly different from zero or one, respectively.
+The regression lines between observed and predicted growth rates under prediction $\mathcal{M}_D$ had intercepts or slopes significantly different from zero or one, respectively, for both percent cover and per capita growth rates.
+
 Removing interspecific interactions from the IPMs had little effect on the synchrony of per capita growth rates (Fig.
@@
 
-Discussion
+DISCUSSION
 ----------
@@
 
-Conclusions
+CONCLUSIONS
 -----------
@@
 
-Ackowledgements
+ACKNOWLEDGMENTS
 ---------------
@@
 table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
-table_data <- table_data[,c("Site", "V1", "V2", "V3", "V4", "V5", "V6", "V7")]
+table_data <- table_data[,c("Site", "V1", "V4", "V7")]
 table_data$spp <- c(length(unique(ks_data$species)),
@@
                     length(unique(az_data$species)))
+colnames(table_data) <- c("Site",
+                          "$\\phi_{r}$", 
+                          "$\\phi_{C}$",
+                          "CV of Total Cover",
+                          "Species richness")
 # colnames(table_data) <- c("Site",
@@
 #                           "$\\text{E}(\\phi_{r})$", 
+#                           "$\\text{E}(\\sigma_{r,c} + \\sigma_{r,e})$",
+#                           "$\\phi_{C}$", 
+#                           "$\\text{E}(\\phi_{C})$",
+#                           "$\\text{E}(\\sigma_{C,c} + \\sigma_{C,e})$",
+#                           "CV of Total Cover",
 #                           "Species richness")
-colnames(table_data) <- c("Site",
-                          "$\\phi_{r}$", 
-                          "$\\text{E}(\\phi_{r})$", 
-                          "$\\text{E}(\\sigma_{r,c} + \\sigma_{r,e})$",
-                          "$\\phi_{C}$", 
-                          "$\\text{E}(\\phi_{C})$",
-                          "$\\text{E}(\\sigma_{C,c} + \\sigma_{C,e})$",
-                          "CV of Total Cover",
-                          "Species richness")
 # table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
@@
 
-synch_cap <- "Observed and expected synchrony among species' per capita growth rates and expected effect of demographic stochasticity on observed synchrony.
+synch_cap <- "Observed synchrony among species' per capita growth rates ($\\phi_{r}$), observed synchrony among species' percent cover ($\\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community.
 Species richness values reflect the number of species analyzed from the community, not the actual richness."
@@
       comment=FALSE,
-      size="footnotesize",
-      floating.environment="sidewaystable")
+      size="normalsize")
 ```
@@
 Synchrony values come from simulations where environmental stochasticity and interspecific interactions are present.
-The IBM was run on a 5 by 5 meter landscape to reduce the effect of demographic stochasticity.", fig.width=6, fig.height=6, fig.width=6, eval=T}
+The IBM was run on a 5 by 5 meter landscape to reduce the effect of demographic stochasticity.", fig.width=6, fig.height=6, fig.width=6, eval=F}
 nyrs <- c(max(out_az$stability$num_observations),
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=3in]{./components/prediction_observed.png}
+  \caption{Observed versus predicted synchrony for percent cover and growth rates under our two limiting case predictions.
+The observed synchrony of per capita growth rates was best predicted by our theoretical prediction in the limiting case of only environmental stochasticity operating ($\phi_R$; green line and points in left panel): it was the only regression line whose slope and intercept did not significantly differ from zero or one, respectively.}
+\end{figure}
+
 \pagebreak{}
@@
   \centering
-      \includegraphics{./components/ipm_pgr_plot.png}
+      \includegraphics[width=6in]{./components/ipm_pgr_plot.png}
   \caption{Community synchrony of per capita growth rates (points, solid lines) and percent cover (triangles, dashed lines) from IPM simulations with (INTER) and without (NoINTER) interspecific interactions.
@@
   \centering
-      \includegraphics{./components/ibm_sims_fig.png}
+      \includegraphics[width=6in]{./components/ibm_sims_fig.png}
   \caption{Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
@@
   \centering
-      \includegraphics{./components/poly_vs_mono_synch.png}
+      \includegraphics[width=4in]{./components/poly_vs_mono_synch.png}
   \caption{Synchrony of species per capita growth rates (filled circles) and percent cover (filled triangles) simulated in monoculture (without species interactions) versus synchrony in polyculture (with species interactions).

### 2016-02-26 devel/prediction_obs_sims_plot.R

@@ -195,31 +195,45 @@ plot_df <- data.frame(site=c(rep(site_names,4)),
                                        exp_synchrony_env$growthrate_prediction,
                                        exp_synchrony_demo$cover,
                                        exp_synchrony_demo$cover),
-                      typesynch=c(rep("cover", 5),
-                                  rep("growth rate",5),
-                                  rep("cover", 5),
-                                  rep("growth rate",5)),
+                      typesynch=c(rep("Percent Cover", 5),
+                                  rep("Growth Rate",5),
+                                  rep("Percent Cover", 5),
+                                  rep("Growth Rate",5)),
                       predtype=c(rep("env",5*2),
                                  rep("demo",5*2)))
 
 
 mycolors <- c("#9D6188","#97A861")
-ggplot(plot_df, aes(x=pred_synchrony, y=obs_synchrony, color=predtype))+
+gout <- ggplot(plot_df, aes(x=pred_synchrony, y=obs_synchrony, color=predtype))+
   geom_point(size=3)+
   stat_smooth(method="lm", se=FALSE, size=1)+
   geom_abline(aes(intercept=0, slope=1), linetype=2, color="grey")+
   scale_y_continuous(limits=c(0,1))+
   scale_x_continuous(limits=c(0,1))+
-  facet_wrap("typesynch", ncol=2)+
-  scale_color_manual(values=mycolors)+
+  facet_wrap("typesynch", ncol=1)+
+  scale_color_manual(values=mycolors, labels=c(expression(M[D]), expression(M[E])), name="")+
   xlab("Predicted Synchrony")+
   ylab("Observed Synchrony")+
   # guides(shape=FALSE)+
-  theme_few()
+  theme_few()+
+  theme(legend.position=c(0.2,0.95),
+        legend.background = element_rect(colour = NA, fill = NA))
 
+png("../docs/components/prediction_observed.png", width = 3, height=5, units = "in", res=100)
+print(gout)
+dev.off()
 
 corrs_all <- ddply(plot_df, .(typesynch, predtype), summarise,
                    value = cor(pred_synchrony, obs_synchrony))
 
-predict(lm(obs_synchrony~pred_synchrony, data=subset(plot_df, typesynch=="growth rate" & predtype=="env")))
[…]

### 2016-02-29 docs/ms_communitySynchrony.Rmd
updates discussion with theoretical prediction mentions

@@
 Here, we extend theory to predict species synchrony under limiting cases that we couple with an empirical analysis from five plant communities.
-When parameterized with empirical estimates, our prediction under the limiting case of only environmental stochasticity operating best explains observed synchrony of per capita growth rates.
+When parameterized with empirical estimates, our prediction of species synchrony under only environmental stochasticity best predicts observed synchrony of per capita growth rates.
 In model simulations, average synchrony of per capita growth rates, which ranges from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
 Removing interspecific competition and demographic stochasticity had small effects on synchrony.
-Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony is approximately equal to the synchrony of species' environmental responses.
+Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony can be predicted by the synchrony of species' environmental responses.
 \end{abstract}
@@
 ------------
-The relationship between species richness and ecosystem stability in grasslands depends on species synchrony [@Tilman2006; @Isbell2009; @Hector2010; @Hautier2014].
+The relationship between species richness and ecosystem stability in plant communities depends on species synchrony [@Tilman2006; @Isbell2009; @Hector2010; @Hautier2014].
 All else being equal, the stability of ecosystem function is highest when species dynamics are asynchronous [@DeMazancourt2013a].
@@
 If two species compete strongly, then fluctuating densities in one species leads to compensatory fluctuations in the densities of the other species [@Lehman2000].
-All else being equal, interspecific competition generally desynchronizes population abundances but sunchronizes per capita growth rates [@Loreau2008a; @Loreau2013; @Gross2014], except in the special case where competition is weak and population growth rates are high [@Loreau2013].
+In the absence of other forces, interspecific competition generally desynchronizes population abundances but synchronizes per capita growth rates [@Loreau2008a; @Loreau2013; @Gross2014], except in the special case where competition is weak and population growth rates are high [@Loreau2013].
 When both those conditions are met, the presence of competition synchronizes population abundances [@Loreau2013].
@@
 Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony.
-In the absence of all other forces, and in communities composed of species with similar growth rates, abundances, and demogaphic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
+In the absence of environmental fluctuations and interspecific competition, and in communities composed of species with similar growth rates, abundances, and demographic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
 However, the effect of demographic stochasticity is strong only in small populations, and in theory it cannot oppose or amplify the effects of environmental fluctuations and interspecific competition [@Loreau2008a].
@@
 Thus, we compare species synchrony in polyculture (with species interactions) to species synchrony in monoculture (equilibrium densities) based on IPM simulations where demographic stochasticity is absent.
+We also use simulation results from species growing in monoculture to estimate the covariance of species' responses to the environment ($cov(u_{ie}, u_{je})$) for parameterizing Equations 5-6.
 
@@
 
-Across the five communities, our limiting case prediction of community-wide synchrony when only environmental stochasticity is operating best matched observed synchrony for per capita growth rates (Fig.
+Across the five communities, our limiting case prediction of community-wide synchrony when only environmental stochasticity is operating ($\mathcal{M}_E$) best matched observed synchrony for per capita growth rates (Fig.
 1).
@@
 Our prediction under $\mathcal{M}_E$ for percent cover did not fair as well, with a slope and intercept that were significantly different from zero or one, respectively.
-The regression lines between observed and predicted growth rates under prediction $\mathcal{M}_D$ had intercepts or slopes significantly different from zero or one, respectively, for both percent cover and per capita growth rates.
+The regression lines between observed and predicted growth rates under prediction $\mathcal{M}_D$ (demographic stochasticity only) had intercepts and/or slopes significantly different from zero or one, respectively, for both percent cover and per capita growth rates.
 
@@
 ----------
-Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2008a].
-We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.
-Our study produced three main findings that were consistent across five plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
-Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community-wide species synchrony is determined by the synchrony of species' responses to environmental conditions.
+Our study produced three main findings that were consistent across five natural plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
+These results are in agreement with our theoretical prediction for species synchrony under the limiting case of community dynamics being driven by environmental fluctuations alone.
+For all of our results, from empirical tests of theoretical predictions to our model simulations, synchrony of per capita growth rates were more consitent and predictive of community dynamics that percent cover.
+Thus, we primarily focus on those results in our discussion, and echo the call of @Loreau2008a to view species synchrony through the lens of per capita growth rates.
 
@@
 ### Demographic stochasticity decreases synchrony
-Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small, in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations.
+Our theoretical prediction for species synchrony in a neutral community with no species interactions or environmental fluctuations ($\mathcal{M}_D$) received no support when parameterized with emprical estimates (Fig.
+1).
+Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small (Fig.
+3), in agreement with theory [@Loreau2008a].
+Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 3).
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
@@
 ### Environmental fluctuations drive community synchrony
-In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations [@Loreau2013].
-Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment.
-We find that environmental fluctuations drive synchrony of per capita growth rates and plant percent cover in all five communities.
-Removing environmental fluctuations consistently reduced synchrony (Fig. 3) and the synchrony of species in monoculture is a strong predictor of synchrony of species in polyculture (Fig. 4).
-Likewise, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
-If we consider sychrony in monoculture to represent the synchrony of species' responses to the environment, then our results lead to only one conclusion: environmental fluctuations are the primary driver of community-wide species synchrony.
+In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations (Equations 5-6).
+Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment [@Loreau2008a].
+Several lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
+First, of our two limiting-case theoretical predictions for species synchrony (Equations 3-6), only our prediction for communities driven by environmental fluctuations alone ($\mathcal{M}_E$) received empirical support (Fig.
+1).
+Second, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
+Lastly, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
+3), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig.
+4).
+If we consider synchrony in monoculture to represent the synchrony of species' responses to the environment, then our results lead to only one conclusion: environmental fluctuations are the primary driver of community-wide species synchrony.
 This is unsurprising since we know that species interactions are weak in these communities.
 More surprising is that our results appear to be quite general: across five arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
-Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural grasslands.
+Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural plant communities.
 
@@
 
-Our analysis contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
+### The importance of functional diversity
+Our analysis also contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
 Functional traits define how a species will respond to its environment.
@@
 2); the only community that contained two life forms (one shrub, three perennial grasses).
-In fact, when we ran simulations with the shrub removed, synchrony of the Idaho community increased substantially and was similar to synchrony in the Kansas community, which also contained three species.
+In fact, when we ran simulations with the shrub removed, synchrony of the Idaho community increased substantially and was similar to synchrony in the Kansas community (Fig.
+SX), which also contained three species.
 Functional diversity is especially important in communities with weak interspecific interactions where synchrony depends soley on species' environmental responses.
@@
 
-We found strong support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak, since removing environmental fluctuations was the only simulation that affected synchrony.
+We found strong support for our theoretical prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak: removing environmental fluctuations was the only simulation that affected synchrony and our limiting-case prediction where only environmental stochasticity is operating faired well when confronted with emprical estiamtes.
 We have reason to believe that our results are quite general since many studies report large niche differences, and thus weak interactions, in plant communities.

### 2016-03-04 docs/ms_communitySynchrony.Rmd
adds collated IBM sims results

@@
 Thus, we compare species synchrony in polyculture (with species interactions) to species synchrony in monoculture (equilibrium densities) based on IPM simulations where demographic stochasticity is absent.
-We also use simulation results from species growing in monoculture to estimate the covariance of species' responses to the environment ($cov(u_{ie}, u_{je})$) for parameterizing Equations 5-6.
+We also use simulation results from species growing in monoculture to estimate the covariance of species' responses to the environment ($\text{cov}(\mu_{ie}, \mu_{je})$) for parameterizing Equations 5-6.
 

### 2016-03-07 docs/ms_communitySynchrony.Rmd
updates to new version of ms with new figures and no percent cover synchrony

--- 2016-02-29 docs/ms_communitySynchrony.Rmd
@@
 ------------
-The relationship between species richness and ecosystem stability in plant communities depends on species synchrony [@Tilman2006; @Isbell2009; @Hector2010; @Hautier2014].
-All else being equal, the stability of ecosystem function is highest when species dynamics are asynchronous [@DeMazancourt2013a].
-Thus, a mechanistic understanding of the diversity-stability relationship requires understanding the processes that increase or decrease species synchrony in nature.
-In the simplest case, when interspecific interactions and demographic stochasticity are weak, species synchrony is entirely determined by the synchrony of species' responses to fluctuating environmental conditions [@Loreau2008a; @Loreau2013].
-Synchrony is highest when correlations among species' environmental responses are perfectly postive, intermediate when species' environmental responses are independent, and lowest when the correlations of species' environmental responses are perfectly negative [@Ives2002; @Loreau2008a; @Loreau2013].
-Of course, real communities are subject to many internal and external forces beyond environmental fluctuations that can affect synchrony and complicate predictions from special theoretical cases [@Loreau2013].
+The earth's ecosystems are in a period of upheaval due to species extinctions [@Cardinale2012], rapid changes in community composition [@Dornelas2014a; @Vellend2013], and human-driven changes in the abiotic environment [@Vitousek1997].
+Such changes can have profound impacts on the provisioning and stability of ecosystem services [@Loreau2001a; @Hooper2005; @Rockstrom2009].
+In particular, studies of experimental grassland communities show that greater species richness leads to greater productivity [@Tilman2001] and temporal stability of productivity [@Tilman2006; @Hector2010].
+The stabilizing effect of species richness stems from individual species fluctuating asynchronously through time [@Hector2010; @Isbell2009; @DeMazancourt2013a].
+Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], which implies that species losses will reduce ecosystem stability.
+Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem function [@Hautier2014].
+Thus, it is becoming clear that predicting the impacts of global change requires a mechanistic understanding of species synchrony in natural settings.
 
-Interspecific competition has long been considered a leading driver of species asynchrony (also called compensatory dynamics) [@Tilman1988; @Ives1995]. 
-If two species compete strongly, then fluctuating densities in one species leads to compensatory fluctuations in the densities of the other species [@Lehman2000].
-In the absence of other forces, interspecific competition generally desynchronizes population abundances but synchronizes per capita growth rates [@Loreau2008a; @Loreau2013; @Gross2014], except in the special case where competition is weak and population growth rates are high [@Loreau2013].
-When both those conditions are met, the presence of competition synchronizes population abundances [@Loreau2013].
-Nonetheless, theoretical and empirical work agree that competition should generally lead to compensatory dynamics in population abundances.
-However, competition-induced compensatory dynamics may or may not increase ecosystem stability, depending on how much competition destabilizes population dynamics [@Loreau2013].
+The mechanisms that underpin species synchrony are complex [@Loreau2008a], and whether synchrony in grassland communities is primarily an outcome of species-specific responses to environmental conditions [@Hautier2014] or competition [@Gross2014] remains controversial [reviewed in @Gonzalez2009].
+Fortunately, theoretical studies have identified the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and interspecific competition [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+Unfortunately, quantifying the relative strengths of each driver on the degree of synchrony from community time series [e.g., @Houlahan2007] is impossible because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
+For example, weak synchrony could reflect positive environmental correlations (synchronizing effect) offset by strong competition (desynchronizing effect), or negative environmental correlations and weak competition.
 
-Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony.
-In the absence of environmental fluctuations and interspecific competition, and in communities composed of species with similar growth rates, abundances, and demographic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
-However, the effect of demographic stochasticity is strong only in small populations, and in theory it cannot oppose or amplify the effects of environmental fluctuations and interspecific competition [@Loreau2008a].
+The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific competition is to remove them one-by-one, and in combination.
+This could be done experimentally, but empirically-based models of interacting populations offer a way to exploit existing data sets from natural communities.
+For example, @Mutshinda2009 fit a dynamic population model to several community time series of insect and bird abundances.
+They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
+While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
+Furthermore, their analysis did not include data from grassland communities, which have received the most experimental [@Tilman2001] and theoretical [@DeMazancourt2013a] attention.
+Thus, there remains a gap in our ability to draw clear lines of inference from theoretical and experimental work to natural plant communities.
 
-The fact that species synchrony emerges from complex interactions among species' environmental responses, species interactions, and demographic stochasticity makes it difficult to disentangle the drivers of synchrony in observational data, especially since an unbiased null expectation does not exist [@Loreau2008a].
-Weak synchrony could reflect positive environmental correlations offset by strong competition, or weak environmental correlations and weak competition.
-Likewise, demographic stochasticity can induce ecological drift in small populations that mask signals of competition and the environment [@Loreau2008a; @DeMazancourt2013a].
-Even the metric used to quantify synchrony can cause confusion.
-For example, @Ranta2008 showed that the most commonly used metric of species synchrony, community covariance, can be positive or negative in the presence or absence of strong competition.
-Thus, it seems that understanding the processes leading to synchrony by analysing patterns alone is impossible.
-So far, no study has rigorously integrated theoretical and empirical approaches to definitively elucidate the mechanisms underlying community-wide species synchrony in natural communities.
-
-In this paper, we refine previous theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a].
+Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony.
+First, we refine previous theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a].
 We use the theory to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
@@
 The fitted population models allow us to estimate species dynamics in monoculture and, thus, test our predictions.
-We also use the population models to directly manipulate the three factors that jointly determine species synchrony: interspecific interactions, environmental forcing, and demographic stochasticity.
+Second, we use the multi-species population models to perform simulation experiments designed to isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific competition on community-wide species synchrony.
 Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
@@
 THEORETICAL MODEL
----------------------
+-----------------
 ### Materials and methods
@@
 
-CLAIRE'S TEXT ON DERIVING THE PREDICTION...
+Our objective is to derive analytical predictions for species synchrony under special limiting cases.
+We constrain our focus to synchrony of per capita growth rates, rather than abundance, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008].
+To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION...
 
 ### Results
-We derived predictions for synchrony of growth rates and synchrony of abundance for two limiting cases that are analytically tractable.
+We derived predictions for synchrony of species' growth rates for two limiting cases that are analytically tractable.
 First, we assume no interspecific interactions, no environmental stochasticity, identical intrinsic growth rates, and that demographic stochasticity is operating but all species have identical demographic variances.
 This limiting case, $\mathcal{M}_{D}$, represents a community where dynamics are driven by demographic stochasticity alone.
-Our predictions are
-
-\begin{equation}
-\phi_{N,\mathcal{M}_{D}} = \frac{1}{\left(\sum_i p_i^{1/2} \right)^2}
-\end{equation}
-
-\noindent{} for synchrony of population abundances and
+Our prediction is
 
@@
 \noindent for synchrony of per capita growth rates.
-In both equations, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
-When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a].
+$p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
+When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, synchrony equal 1/S [@Loreau2008a].
 
@@
 Thus, there are no interspecific interactions, demographic stochasticity is absent, intrinsic growth rates are identical, and environmental variance is identical for all species.
-Our predictions for this limiting case, $\mathcal{M}_{E}$, are
-
-\begin{equation}
-\phi_{N,\mathcal{M}_{E}} = \sum_{i,j} p_i p_j \text{cov}(\mu_{ei},\mu_{ej})
-\end{equation}
-
-\noindent{} for synchrony of population abundances and
+Our prediction for this limiting case, $\mathcal{M}_{E}$, is
 
@@
 \noindent{} for synchrony of per capita growth rates.
-In both equations, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$, and $\text{cov}(\mu_{ei},\mu_{ej})$ is the standardize covariance (correlation) of environmental responses between species *i* and species *j*.
+Again, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$, and $\text{cov}(\mu_{ei},\mu_{ej})$ is the standardize covariance (correlation) of environmental responses between species *i* and species *j*.
 Confronting our theoretical predictions with data requires estimates of species dynamics in monoculture to calculate the covariance of species' environmental responses.
@@
 \begin{equation}
-\phi_x = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{x_{i}})^{2}}
+\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
 \end{equation}
 
-\noindent where $\sigma_{x_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) or percent cover ($N_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate or abundance.
+\noindent where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
 $\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [@Loreau2008a].
@@
 
-#### Removing interspecific interactions and quantifying species' environmental responses (IPM)
+#### Removing interspecific interactions (IPM)
 To remove interspecific interactions, we simply set the off-diagonals of the interaction matrix for each vital rate regression to zero.
@@
 This allowed species time to reach their stable size distribution.
-We then calculated the synchrony of species' percent cover and per capita growth rates (see *Calculating observed synchrony*) over 100 randomly selected contiguous 50 time-step sections.
-These simulations allowed us to isolate the effects of competition and environmental fluctuations in large populations where demographic stochasticity is absent.
-Note, however, what we could not run IPM simulations without environmental fluctuations, because it is only those fluctuations that drive variability in stochastich IPMs.
-In a constant environment, IPM simulations asymptote at stable equilibriums for each species.
-
-A common way to quantify the importance of environmental stochasticity on synchrony is to compare species synchrony when grown in monoculture versus when grown in polyculture [@Gross2014].
-The assumption is that synchrony in monoculture is representative of the synchrony of species' environmental responses.
-Thus, we compare species synchrony in polyculture (with species interactions) to species synchrony in monoculture (equilibrium densities) based on IPM simulations where demographic stochasticity is absent.
-We also use simulation results from species growing in monoculture to estimate the covariance of species' responses to the environment ($cov(u_{ie}, u_{je})$) for parameterizing Equations 5-6.
+We then calculated the synchrony of species' per capita growth rates over 100 randomly selected contiguous 50 time-step sections.
+These simulations allowed us to isolate the effect of removing competition in large populations where demographic stochasticity is absent.
+We also used simulation results from species growing in monoculture to estimate the covariance of species' responses to the environment ($\text{cov}(\mu_{ie}, \mu_{je})$) for parameterizing Equation 3.
 
@@
 ### Results
-
 Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
@@
 Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
-
-<!-- THIS HOW TO COMMENT-OUT 
-For per capita growth rates, all sites exhibited more synchronous dynamics than expected if only demographic stochasticity was operating (E($\phi_r$) < $\phi_r$ in Table 2).
-The difference between observed $\phi_r$ and E($\phi_r$) represents the effect of the environmemt, species interactions, and species differences combined.
-The combined effect of the environment, species interactions, and species differences is positive in all communities (E($\phi_c + \phi_e$) in Table 2), indicating a synchronizing effect of those forces that counter-balance the desynchronizing effect of demographic stochasticity.
-In contrast, expected synchrony of percent cover was lower than expected if only demographic stochasticity was operating in two out of the five communities (Kansas and Idaho, Table 2).
--->
-
 For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig.
 X1), meaning that we can use the models to test the importance of different drivers of synchrony in these communities.
-However, IBM-simulated synchrony is consistently lower than IPM-simulated synchrony (Fig. X1), likely due to the desynchronizing effect of demographic stochasticity.
+However, IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. X1), likely due to the desynchronizing effect of demographic stochasticity.
 
@@
 The regression line between observed growth rate synchrony and predicted growth rate synchrony under prediction $\mathcal{M}_E$ was the only one whose slope and intercept were not significantly different from zero (*P* = 0.88) and one (*P* = 0.87), respectively (Supplementary Information X).
-Our prediction under $\mathcal{M}_E$ for percent cover did not fair as well, with a slope and intercept that were significantly different from zero or one, respectively.
-The regression lines between observed and predicted growth rates under prediction $\mathcal{M}_D$ (demographic stochasticity only) had intercepts and/or slopes significantly different from zero or one, respectively, for both percent cover and per capita growth rates.
+The regression lines between observed and predicted growth rates under prediction $\mathcal{M}_D$ (demographic stochasticity only) had intercepts and/or slopes significantly different from zero or one, respectively (Fig.
+1).
 
-Removing interspecific interactions from the IPMs had little effect on the synchrony of per capita growth rates (Fig.
-2) because interspecific interactions are weak in our focal communities (Table X1, and see @Chu2015).
-Only in Arizona does removing species interactions result in a small decline in synchrony of per capita growth rates (Fig. 2a).
-For two out five communities, removing interspecific interactions had little effect on the synchrony of percent cover (Fig. 2c,d; Kansas and Montana).
-In the three communities where removing interspecific interactions did have an effect (Arizona, Idaho and New Mexico), synchrony of percent cover was higher when species interactions were absent (Fig.
-2a,b,e).
-We achieved similar results with the IBM simulations: removing interspecific interactions had little effect on synchrony of per capita growth rates, but increased synchrony of percent cover in Idaho and New Mexico (Fig.
-3).
-In Idaho, however, synchrony of per capita growth rates is similar among all simulations with species interactions present, regardless of the presence of environmental fluctuations (Fig. 3b).
-
-Species' environmental responses, quantified as temporal random effects in vital rate regressions, were positively correlated in each community, ranging from $\rho = 0.23$ to $\rho = 0.38$ when averaged over vital rates (Appendix X).
-Thus, removing environmental stochasticity caused synchrony of per capita growth rates to decrease in all five communities (Fig. 3a-e).
-In our simulations synchrony of per capita growth rates was higher when environmental variation was present (average $\phi_r = 0.62$) rather than absent (average $\phi_r = 0.43$).
-This is true whether species interactions are present or not, except for Idaho (Fig. 3b).
-Removing environmental fluctuations also decreased synchrony of percent cover, except in Idaho and Montana (Fig. 3f-j).
-In Idaho, synchrony of percent cover is similar among simulations with and without environmental fluctuations when species interactions are not present, but synchrony is lowest in simulation where species interactions are present and environmental fluctuations are absent (Fig.
-3g).
-In Montana, all simulations result in the same values of synchrony of percent cover (Fig. 3h).
-
-Without environmental stochasticity or species interactions, synchrony of per capita growth rates all approximated the theoretical expectation of 1/*S*, where *S* is species richness, when only demographic stochasticity is operating (Fig.
-3; top panels, dark blue lines).
-Demographic stochasticity counteracted the synchronizing effect of the environment in only the smallest populations (Fig. 3).
-The strength of demographic stochasticity decreased with the size of the simulated landscape in our IBMs.
-Thus, in Figure 3 the strength of demographic stochasticity decreases along the x-axis.
-In the absence of environmental stochasticity and species interactions, the strength of demographic stochasticity has no impact of the synchrony of per capita growth rates or percent cover (Fig.
-3; dark blue lines).
-In the presence of environmental stochasticity, synchrony of per capita growth rates is lower when demographic stochasticity is strong (Fig. 3a-e; orange tones).
-The desynchronizing effect of demographic stochasticity on per capita growth rates is strongest in small populations (small landscapes) for all communities (Fig. 3a-e).
-Demographic stochasticity had a weaker effect on synchrony of percent cover; only in the Arizona community does strong demographic stochasticity counteract the synchronizing effect of the environment (Fig.
-3; bottom panels).
-
-Since species interactions are weak and demographic stochasticity only matters in small populations, the synchrony of species in monoculture (without interspecific interactions) should predict the synchrony of species in polyculture (with interspecific interactions).
-For all five communities and for percent cover (rank correlation = 0.8) and per capita growth rates (rank correlation = 1), the synchrony of species in monoculture is strongly associated with species synchrony in polyculture (Fig.
-4).
+Simulation experiments reveal that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in snychrony of species growth rates in four out five communities (Fig.
+2a).
+Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by X% in Arizona, X% in Kansas, X% in Montana, and X% in New Mexico.
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2a), but the effect was small (X% increase).
+Species interactions are weak in this communitie (Table SX), so removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
+Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2a).
+The desynchronizing effect of demographic stochasticity was able to modestly counteract the synchronizing force of the environment, but not enough to match the synchrony of species when only demographic stochasticity is operating (Fig.
+2b).
+For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
+Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more noisy, as expected by theory [@Loreau2008].
 
@@
 
+<!--
 ### The importance of functional diversity
@@
 But functionally diverse suites of species will have the lowest synchrony of environmental responses, and thus lowest community synchrony and highest stability.
+-->
 
@@
   \centering
-      \includegraphics[width=3in]{./components/prediction_observed.png}
-  \caption{Observed versus predicted synchrony for percent cover and growth rates under our two limiting case predictions.
+      \includegraphics[width=3in]{./components/formatted_fig1.png}
+  \caption{Observed versus predicted synchrony of growth rates under our two limiting case predictions.
 The observed synchrony of per capita growth rates was best predicted by our theoretical prediction in the limiting case of only environmental stochasticity operating ($\phi_R$; green line and points in left panel): it was the only regression line whose slope and intercept did not significantly differ from zero or one, respectively.}
@@
   \centering
-      \includegraphics[width=6in]{./components/ipm_pgr_plot.png}
-  \caption{Community synchrony of per capita growth rates (points, solid lines) and percent cover (triangles, dashed lines) from IPM simulations with (INTER) and without (NoINTER) interspecific interactions.
-`F' in x-axis labels refers to a fluctuating environment.
-Errorbars represent the upper and lower 90th quantiles of synchrony from 100 random contiguous 50 year sections from a 2,000 iteration IPM run.
-Points represent the mean synchrony across the 100 sections.}
+      \includegraphics[width=6in]{./components/formatted_fig2.png}
+  \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.}
 \end{figure}
@@
 
+<!--
 \begin{figure}[!ht]
@@
 \end{figure}
+-->
 

### 2016-03-07 devel/plot_ipm_sims.R

@@ -84,14 +84,32 @@ cor(subset(polymono_wide, typesynch=="abund_synch")[,"ENVINTER"],
 library(ggthemes)
 polymono <- agg_synch[,c("site", "typesynch", "experiment", "mean_synch")]
 polymono_wide <- dcast(polymono, site+typesynch~experiment, value.var = "mean_synch")
-g1 <- ggplot(polymono_wide, aes(x=ENVNOINTER, y=ENVINTER))+
+# g1 <- ggplot(polymono_wide, aes(x=ENVNOINTER, y=ENVINTER))+
+#   geom_abline(aes(intercept=0, slope=1), linetype=3)+
+#   geom_point(size=3, aes(shape=typesynch, color=site))+
+#   scale_y_continuous(limits=c(0,1))+
+#   scale_x_continuous(limits=c(0,1))+
+#   ylab("Species synchrony in polyculture")+
+#   xlab("Species synchrony in monoculture")+
+#   scale_shape_discrete(name="Temporal Variable", labels=c("Per capita growth rate", "Percent cover"))+
+#   scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
+#                      name = "",
+#                      labels = c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico"))+
+#   theme_few()+
+#   guides(shape=FALSE)+
+#   # ggtitle("A                                                          ")+
+#   theme(legend.position=c(0.2,0.8))+
+#   theme(legend.text = element_text(size = 8))+
+#   theme(legend.background = element_rect(colour = NA, fill = NA))
+
+g1 <- ggplot(subset(polymono_wide,typesynch=="pgr_synch"), aes(x=ENVNOINTER, y=ENVINTER))+
   geom_abline(aes(intercept=0, slope=1), linetype=3)+
-  geom_point(size=3, aes(shape=typesynch, color=site))+
+  geom_point(size=3, aes(color=site))+
   scale_y_continuous(limits=c(0,1))+
   scale_x_continuous(limits=c(0,1))+
   ylab("Species synchrony in polyculture")+
   xlab("Species synchrony in monoculture")+
-  scale_shape_discrete(name="Temporal Variable", labels=c("Per capita growth rate", "Percent cover"))+
[…]

### 2016-03-07 devel/prediction_obs_sims_plot.R

@@ -204,22 +204,22 @@ plot_df <- data.frame(site=c(rep(site_names,4)),
 
 
 mycolors <- c("#9D6188","#97A861")
-gout <- ggplot(plot_df, aes(x=pred_synchrony, y=obs_synchrony, color=predtype))+
+gout <- ggplot(subset(plot_df, typesynch=="Growth Rate"), aes(x=pred_synchrony, y=obs_synchrony, color=predtype))+
   geom_point(size=3)+
   stat_smooth(method="lm", se=FALSE, size=1)+
   geom_abline(aes(intercept=0, slope=1), linetype=2, color="grey")+
   scale_y_continuous(limits=c(0,1))+
   scale_x_continuous(limits=c(0,1))+
-  facet_wrap("typesynch", ncol=1)+
+  # facet_wrap("typesynch", ncol=1)+
   scale_color_manual(values=mycolors, labels=c(expression(M[D]), expression(M[E])), name="")+
   xlab("Predicted Synchrony")+
   ylab("Observed Synchrony")+
   # guides(shape=FALSE)+
   theme_few()+
-  theme(legend.position=c(0.2,0.95),
+  theme(legend.position=c(0.1,0.85),
         legend.background = element_rect(colour = NA, fill = NA))
 
-png("../docs/components/prediction_observed.png", width = 3, height=5, units = "in", res=100)
+png("../docs/components/prediction_observed.png", width = 4, height=4, units = "in", res=100)
 print(gout)
 dev.off()
 

### 2016-03-08 docs/ms_communitySynchrony.Rmd
adds synchrony percent diffs results and to results text

@@
 Our objective is to derive analytical predictions for species synchrony under special limiting cases.
-We constrain our focus to synchrony of per capita growth rates, rather than abundance, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008].
+We constrain our focus to synchrony of per capita growth rates, rather than abundance, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
 To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION...
@@
 2a).
-Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by X% in Arizona, X% in Kansas, X% in Montana, and X% in New Mexico.
-Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2a), but the effect was small (X% increase).
-Species interactions are weak in this communitie (Table SX), so removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
-Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2a).
+Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2a), but the effect was small (9% increase).
+Species interactions are weak in this communitie (Table SX), so removing interspecific competition had little effect on synchrony (Fig. 2a; "No Comp." simulations).
+Removing competition caused, at most, a 5% change in synchrony (Table SX).
+Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2a), with an average 6% increase.
 The desynchronizing effect of demographic stochasticity was able to modestly counteract the synchronizing force of the environment, but not enough to match the synchrony of species when only demographic stochasticity is operating (Fig.
@@
 For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
-Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more noisy, as expected by theory [@Loreau2008].
+Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more noisy, as expected by theory [@Loreau2008a].
 
@@
 These results are in agreement with our theoretical prediction for species synchrony under the limiting case of community dynamics being driven by environmental fluctuations alone.
-For all of our results, from empirical tests of theoretical predictions to our model simulations, synchrony of per capita growth rates were more consitent and predictive of community dynamics that percent cover.
-Thus, we primarily focus on those results in our discussion, and echo the call of @Loreau2008a to view species synchrony through the lens of per capita growth rates.
 
 ### Interspecific interactions had little impact on community synchrony
-We expected community synchrony of per capita growth rates to decrease and synchrony of plant cover to increase when we removed interspecific interactions [@Loreau2008a and Appendix X].
-In the absence of demographic stochasticity, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions for both per capita growth rates and percent cover (Fig.
-2).
+We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a and Supplementary Information].
+In the absence of demographic stochasticity, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2a).
 The only notable exception is the Idaho community where, as expected, synchrony of percent cover increased when we removed interspecific interactions.
@@
 However, rare species may occupy rare niches that do not overlap with other species, in which case synchrony would remain predominantly driven by environmental responses.
-Rare species could also be limited by species-specific natural enemies whose effects are density-independent [@Mordecai2011].
-Under such conditions, synchrony will depend on the degree of host-specificity and the pathogen's response to environmental conditions.
+Rare species could also be limited by generalist natural enemies whose effects are density-independent but alter competitive hierarchies [@Mordecai2011].
+Under such conditions, synchrony will depend on the degree of pathogen-induced fitness differences and the pathogen's response to environmental conditions.
 Niether our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].
@@
       \includegraphics[width=6in]{./components/formatted_fig2.png}
-  \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.}
+  \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
+(a) Synchrony of species' growth rates for each study area from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
++ No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
++ No E.S.'').
+Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation.
+(b) Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S.
++ No Comp.'').
+The strength of demographic stochasticity decreases as landscape size increases.}
 \end{figure}

### 2016-03-08 devel/make_ms_plots.R

@@ -97,6 +97,10 @@ ggplot(plot_df, aes(x=simulation, y=synchrony, fill=site))+
 
 ggsave("../docs/components/all_sims_results.png", width = 10, height = 4, units="in", dpi=75)
   
+##  Calculate percent differences for Results
+plot_df$control <- rep(plot_df[which(plot_df$simulation=="1All Drivers"),"synchrony"], times=nrow(plot_df)/nsites)
+plot_df$percent_diff <- with(plot_df, abs(synchrony-control)/((synchrony+control)/2)*100)
+write.csv(plot_df, "../results/synchsims_percent_diffs.csv")
 
 ##  Make demographic stochasiticty plot for all landscape sizes -----
 ibm_demo_rms <- subset(ibm_synch_agg, typesynch=="Per capita growth rate")

### 2016-03-08 docs/ms_communitySynchrony.Rmd
fixes typos; changes title

@@
 \begin{center}
-\large{\textbf{Disentangling the drivers of species synchrony in natural plant communities: environmental forcing, demographic stochasticity, and interspecific interactions}}
+\large{\textbf{Environmental responses, not species interactions, determine species synchrony in five natural plant communities}}
 
@@
 While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Furthermore, their analysis did not include data from grassland communities, which have received the most experimental [@Tilman2001] and theoretical [@DeMazancourt2013a] attention.
-Thus, there remains a gap in our ability to draw clear lines of inference from theoretical and experimental work to natural plant communities.
+Furthermore, although @Mutshinda2009 quantify the relative importance of environmental stochasticity and competition to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony changes when each factor is removed.
+Lastly, their analysis did not include data from grassland communities, which have received the most experimental [@Tilman2001; @Hector2010] and theoretical [@DeMazancourt2013a] attention.
+Thus, we lack the ability to draw clear lines of inference from theoretical and experimental work to natural plant communities.
 
@@
 Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2a), but the effect was small (9% increase).
-Species interactions are weak in this communitie (Table SX), so removing interspecific competition had little effect on synchrony (Fig. 2a; "No Comp." simulations).
+Species interactions are weak in these communities [Table SX and @Chu2015], so removing interspecific competition had little effect on synchrony (Fig. 2a; "No Comp." simulations).
 Removing competition caused, at most, a 5% change in synchrony (Table SX).
@@
 ----------
-Our study produced three main findings that were consistent across five natural plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
+Our study produced three main findings that were generally consistent across five natural plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
 These results are in agreement with our theoretical prediction for species synchrony under the limiting case of community dynamics being driven by environmental fluctuations alone.

### 2016-03-08 docs/ms_communitySynchrony.Rmd
updates discussion with Idaho outlier explanation

@@
 The regression line between observed growth rate synchrony and predicted growth rate synchrony under prediction $\mathcal{M}_E$ was the only one whose slope and intercept were not significantly different from zero (*P* = 0.88) and one (*P* = 0.87), respectively (Supplementary Information X).
-The regression lines between observed and predicted growth rates under prediction $\mathcal{M}_D$ (demographic stochasticity only) had intercepts and/or slopes significantly different from zero or one, respectively (Fig.
+The regression line between observed and predicted growth rates under prediction $\mathcal{M}_D$ (demographic stochasticity only) had an intercept and slope significantly different from zero and one, respectively (Fig.
 1).
 
-Simulation experiments reveal that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in snychrony of species growth rates in four out five communities (Fig.
+Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out five communities (Fig.
 2a).
@@
 We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a and Supplementary Information].
-In the absence of demographic stochasticity, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2a).
-The only notable exception is the Idaho community where, as expected, synchrony of percent cover increased when we removed interspecific interactions.
-When we also removed environmental fluctuations, removing interspecific interactions decreased synchrony of per capita growth rates in three of our five communities and increased synchrony of per capita growth rates in two communities (Fig.
-3).
+Contrary to expectations, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2a).
 The lack of an effect of interspecific competition on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
@@
 Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small (Fig.
-3), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 3).
+2b), in agreement with theory [@Loreau2008a].
+Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 2b).
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
-Only in the Idaho community does synchrony of per capita growth rates approach 1/*S* in a non-constant environment (Fig. 3x).
+Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of specuies) in a non-constant environment (Fig. 2b).
 This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
@@
 ### Environmental fluctuations drive community synchrony
-In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations (Equations 5-6).
+In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations (Equation 4).
 Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment [@Loreau2008a].
 Several lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
-First, of our two limiting-case theoretical predictions for species synchrony (Equations 3-6), only our prediction for communities driven by environmental fluctuations alone ($\mathcal{M}_E$) received empirical support (Fig.
+First, of our two limiting-case theoretical predictions for species synchrony (Equations 3-4), only our prediction for communities driven by environmental fluctuations alone ($\mathcal{M}_E$) received empirical support (Fig.
 1).
-Second, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
+Second, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2a).
 Lastly, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
-3), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig.
-4).
-If we consider synchrony in monoculture to represent the synchrony of species' responses to the environment, then our results lead to only one conclusion: environmental fluctuations are the primary driver of community-wide species synchrony.
-This is unsurprising since we know that species interactions are weak in these communities.
-More surprising is that our results appear to be quite general: across five arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
+2a), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig.
+3).
+Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony.
+Likewise, our results appear to be quite general: across four arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
+Only in the Idaho community did not removing environmental fluctuations not cause a large decrease in synchrony, but removing competition had no effect either.
 Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural plant communities.
 
-One challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
+The insensitivity of species synchrony to model perturbations in the Idaho community challenges the consistency of our results, but its outlier status may have more to do with environmental variability than intrinsic species synchrony.
+Regardless of how synchronously or asynchronously species in a community respond to environmental fluctuations, if the environment does not fluctuate substantially then removing such fluctuations will not impact species synchrony.
+Indeed, the Idaho community had the lowest variability in percent cover over the observation period (Table 2).
+At the same time, correlations among species' temporal random effects for each vital rate in the Idaho community were similar to, or greater than, those in the other communities we studied (Table SX).
+These lines of evidence lead us to conclude that either environmental fluctuations are weak in the Idaho community or our model was unable to detect temporal fluctuations in the data.
+In other words, species' responses to environmental conditions are likely the main determinant of synchrony in the Idaho community, but we were unable to detect the effect of the environment due to model misspecification or a lack of a strong environmental signal in the data.
+
+A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
 Rare species, which our analysis does not include, could be kept at low densities by competitive suppression due to strong interspecific interactions.
@@
 Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
-When we removed environmental variation, we found that synchrony decreased (Fig. 3).
+When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2a).
 A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
@@
 However, if community-wide species synchrony is less than 1, some degree of compensatory dynamics must be present [@Loreau2008a].
-Since observed and simulated synchrony is less than 1 in all our focal communities (Fig. 1), we conclude that compensatory dynamics are common in natural grasslands.
+Since observed and simulated synchrony is less than 1 in all our focal communities (Table 2 and Fig. 2), we conclude that compensatory dynamics are common in natural grasslands.
 But, in agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
-
-We found strong support for our theoretical prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak: removing environmental fluctuations was the only simulation that affected synchrony and our limiting-case prediction where only environmental stochasticity is operating faired well when confronted with emprical estiamtes.
-We have reason to believe that our results are quite general since many studies report large niche differences, and thus weak interactions, in plant communities.
-However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
-The difference between the Gross et al.
-(2014) study and our own may be due to community assembly processes: recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
+Such a conclusion is in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
+It may be that recently assembled communities, such as experimental plots, exhibit strong species interactions that will weaken over time [@Kokkoris1999].
 Viewing community synchrony through the lens of community assembly processes may provide a way to reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.

### 2016-03-08 devel/plot_yreffect_corrs.R

@@ -83,6 +83,22 @@ for(j in 1:num_vitals){
 } # end vital rate loop
 
 
+## Plot histograms of year effects by vital rate
+vital_rates <- names(list_yr_effects)
+par(mfrow=c(3,5))
+for(do_vital in vital_rates){
+  tmp_vital <- list_yr_effects[[do_vital]]
+  site_names <- names(tmp_vital)
+  for(do_site in site_names){
+    tmp_site <- tmp_vital[[do_site]]
+    hist(tmp_site, main=paste(do_site,do_vital), xlab="Year Effect", xlim=c(-3,8))
+  }
+}
+
+##  Get standard deviation of year effect posterior
+
+
+
 ####
 ####  Calculate average correlation of year effects ----------------------------
 ####

### 2016-03-09 devel/make_ms_plots.R
adds errorbars to figs; closes #19

@@ -53,7 +53,9 @@ ipm_synch <- subset(ipm_synch, typesynch=="pgr_synch")
 ##  Read in IBM results -----
 ibm_synch_all <- readRDS("../results/ibm_sims_collated.RDS") 
 ibm_synch_agg <- ddply(ibm_synch_all, .(experiment, site, expansion, typesynch), summarise,
-                   avg_synch = mean(synch))
+                   avg_synch = mean(synch),
+                   up_synch = quantile(synch, 0.95),
+                   lo_synch = quantile(synch, 0.05))
 ibm_synch <- subset(ibm_synch_agg, expansion==5 & typesynch=="Per capita growth rate")
 
 ##  Combine results -----
@@ -77,23 +79,40 @@ all_experiments <- c(control_synch$avg_synch,
                      nocomp_synch$avg_synch,
                      nodsnocomp_synch$mean_synch, 
                      noesnocomp_synch$avg_synch)
+all_ups <- c(control_synch$up_synch,
+             nods_synch$up_synch,
+             noes_synch$up_synch,
+             nocomp_synch$up_synch,
+             nodsnocomp_synch$up_synch, 
+             noesnocomp_synch$up_synch)
+all_downs <- c(control_synch$lo_synch,
+               nods_synch$lo_synch,
+               noes_synch$lo_synch,
+               nocomp_synch$lo_synch,
+               nodsnocomp_synch$lo_synch, 
+               noesnocomp_synch$lo_synch)
 
 plot_df <- data.frame(site = rep(site_labels, times=length(sim_names)),
                       simulation = rep(sim_names_order, each=nsites),
-                      synchrony = all_experiments)
+                      synchrony = all_experiments,
+                      upper_synch = all_ups,
+                      lower_synch = all_downs)
 
 ##  Make the main (all sims) plot -----
-ggplot(plot_df, aes(x=simulation, y=synchrony, fill=site))+
+ggplot(plot_df, aes(x=simulation, y=synchrony, fill=site, color=site))+
   geom_bar(stat="identity")+
+  geom_errorbar(aes(ymin=lower_synch, ymax=upper_synch), color="white", size=1, width=0.25)+
+  geom_errorbar(aes(ymin=lower_synch, ymax=upper_synch), width=0.25)+
   facet_wrap("site", nrow=1)+
[…]

### 2016-03-09 docs/ms_communitySynchrony.Rmd
adds IPM no-ARTR scripts and sims for #18

@@
   \centering
-      \includegraphics[width=6in]{./components/formatted_fig2.png}
+      \includegraphics[width=6in]{./components/all_sims_results.png}
   \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
-(a) Synchrony of species' growth rates for each study area from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
+Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
 + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
@@
 (b) Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S.
++ No Comp.'').
+The strength of demographic stochasticity decreases as landscape size increases.}
+\end{figure}
+
+\pagebreak{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/ibm_sims_across_landscape.png}
+  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S.
 + No Comp.'').

### 2016-03-09 runscripts/Idaho_NoARTR_sims/ibm_largepop_wrapper.R

@@ -1,0 +1,168 @@
+##  Wrapper for IBM runs to compare to IPM runs
+##
+##  The IBM is run on a different landscape sizes (plot size) to
+##  to test the effect demographic stochasticity.
+
+##  Author:       Andrew Tredennick, Peter Adler
+##  Email:        atredenn@gmail.com
+##  Date created: 10-23-2015
+
+
+## Clear the workspace
+rm(list=ls())
+
+
+####
+####  Load libraries -----------------------------------------------------------
+####
+library(communitySynchrony) # inhouse package
+library(synchrony)          # calculates Claire's synchrony metric
+library(boot)               # for the IBM
+library(mvtnorm)            # for the IBM
+library(msm)                # for the IBM
+
+
+
+####
+####  Set up global variables and simulation config ----------------------------
+####
+totSims <- 20      # number of simulations per site (1 here since using large landscape)
+totT <- 100      # time steps of simulation
+burn.in <- 25    # time steps to discard before calculating cover values
+L <- 100          # dimension of square quadrat (cm)
+doGroup <- NA     # NA for spatial avg., values for a specific group
+constant.vec <- c(FALSE, FALSE, TRUE, TRUE) # TRUE for constant env.; FALSE for random year effects
+sppinter.vec <- c(FALSE, TRUE, FALSE, TRUE) # TRUE for interspp interactions; FALSE for no interspp interactions
+filename.flag <- c("fluctnointer", "fluctinter", "constnointer", "constinter")
+# constant.vec <- c(FALSE, TRUE, TRUE) # TRUE for constant env.; FALSE for random year effects
+# sppinter.vec <- c(TRUE, FALSE, TRUE) # TRUE for interspp interactions; FALSE for no interspp interactions
+# filename.flag <- c("fluctinter", "constnointer", "constinter")
+
+## Looping over different landscape sizes
+expand_vec <- c(1,2,3,4,5) # 1 = 1x1 m^2, 2 = 2x2m^2, etc
+
+
+####
+####  Read in regression parameters; list for all sites and species ------------
+####
+Gpars_all <- readRDS("../results/growth_params_list.RDS")
+Spars_all <- readRDS("../results/surv_params_list.RDS")
+Rpars_all <- readRDS("../results/recruit_parameters.RDS")
+
+site_names <- names(Gpars_all)
[…]

### 2016-03-09 runscripts/Idaho_NoARTR_sims/ibm_skeleton.R

@@ -1,0 +1,245 @@
+##############################################################
+##  Individually-based model for a multiple species,
+##  with density-dependence and explicit space (random
+##  spatial pattern)
+##
+##  Gaussian distance function
+##
+##  This version uses the negative binomial recruitment function
+##
+## Initialized from observed size distributions
+##############################################################
+
+#############################################################
+##  This is a skeleton version to be called from a wrapper ##
+#############################################################
+
+
+####
+####  Vital rate function subroutines ------------------------------------------
+####
+## Growth
+grow <- function(Gpars,doSpp,doYear,sizes,crowding){
+  # crowding and nb are vectors of dim Nspp
+  logsizes=log(sizes)
+  
+  mu <- Gpars$intcpt[doSpp]+Gpars$intcpt.yr[doYear,doSpp]+ # intercept
+        (Gpars$slope[doSpp]+Gpars$slope.yr[doYear,doSpp])*logsizes+ # size effect
+        Gpars$nb[doSpp,]%*%crowding # competition
+  
+  tmp <- which(mu<log(minSize)*1.5)         # we will kill vanishingly small plants...below
+  mu[tmp] <- log(minSize)                   # truncate tiny sizes (they cause problems in sigma2)
+  sigma2 <- Gpars$sigma2.a[doSpp]*exp(Gpars$sigma2.b[doSpp]*mu)
+  out <- exp(rnorm(length(sizes),mu,sqrt(sigma2)))
+  if(sum(is.na(out))>0) browser()
+  out[tmp] <- 0                             # here's the killing
+  out[out>maxSize[doSpp]] <- maxSize[doSpp] # truncate big plants
+  return(out)
+}
+
+## Survival
+survive <- function(Spars,doSpp,doYear,sizes,crowding){
+  logsizes <- log(sizes)
+  
+  mu <- Spars$intcpt[doSpp]+Spars$intcpt.yr[doYear,doSpp]+ # intercept
+        (Spars$slope[doSpp]+Spars$slope.yr[doYear,doSpp])*logsizes+ # size effect
[…]

### 2016-03-09 runscripts/Idaho_NoARTR_sims/run_ibm.R

@@ -1,0 +1,389 @@
+##  Individual-based model with no environmental variance
+##  Used to test against the demographic approximation of the IPM
+
+######  FROM PETER #########################################
+# Individually-based model for a multiple species,
+# with density-dependence and explicit space (random
+# spatial pattern)
+
+# Gaussian distance function
+
+# This version uses the negative binomial recruitment function
+
+# Initialized from observed size distributions
+############################################################
+
+##  Clear the workspace and load libraries
+rm(list=ls())
+library(communitySynchrony)
+library(synchrony)
+library(boot)
+library(mvtnorm)
+library(msm)
+
+
+make.P.values <- function(v,u,muWG,muWS, #state variables
+                          Gpars,Spars,doYear,doSpp){  #growth arguments
+  S(u,muWS,Spars,doYear,doSpp)*G(v,u,muWG,Gpars,doYear,doSpp) 
+}
+
+make.P.matrix <- function(v,muWG,muWS,Gpars,Spars,doYear,doSpp) {
+  muWG=expandW(v,v,muWG)
+  muWS=expandW(v,v,muWS)
+  
+  P.matrix=outer(v,v,make.P.values,muWG,muWS,Gpars,Spars,doYear,doSpp)
+  return(h[dospps]*P.matrix)
+} 
+
+
+####
+####  Set some global variables for the simulation
+####
+totSims=100
+totT=150     # time steps of simulation
+burn.in=50  # time steps to discard before calculating cover values
+L=100       # dimension of square quadrat (cm)
+expand=2    # 1 = 1x1 m^2, 2 = 2x2m^2, etc
+init.cover=c(0,1,1,1)   # in % cover
+# maxSize=c(8000,500,500,500)
+maxSize=c(3000,202,260,225) 
+minSize=0.2
+iter_matrix_dims <- c(50,75,50,75)
+do_site="Idaho"
+Nyrs=22
+myCol=c("black","darkgreen","blue","red")
+doGroup=NA  # NA for spatial avg., values 1-6 for a specific group
+constant=T        
+
+out.nt <- list()
+out.nt[[1]] <- array(0,c(totT,iter_matrix_dims[2], totSims))
[…]

### 2016-03-09 runscripts/Idaho_NoARTR_sims/run_ipm_simulations.R

@@ -1,0 +1,223 @@
+############################################################
+##  Script to run IPM simulations for each site.
+##  
+##  Authors: Andrew Tredennick, Peter Adler, Chengjin Chu
+##  Email:   atredenn@gmail.com
+##  Created: 9-11-2015
+############################################################
+
+##  Clear the workspace
+rm(list=ls(all=TRUE))
+
+##  Set the IPM simulation run time and burn in
+tlimit <- 2500
+burn_in <- 500
+
+##  Set up looping vectors for spp interactions
+inter_comp <- c(TRUE, FALSE)
+sim_names <- c("ENVINTER", "ENVNOINTER")
+
+
+####
+####  Load libraries
+####
+library(communitySynchrony)
+library(boot)
+library(mvtnorm)
+library(msm)
+library(statmod)  
+
+
+####
+####  IPM subroutines
+####
+## Make combined kernel
+make.K.values=function(v,u,muWG,muWS, #state variables
+                       Rpars,rpa,Gpars,Spars,doYear,doSpp){  #growth arguments
+  f(v,u,Rpars,rpa,doSpp)+S(u,muWS,Spars,doYear,doSpp)*G(v,u,muWG,Gpars,doYear,doSpp) 
+}
+
+## Function to make iteration matrix based only on mean crowding
+make.K.matrix=function(v,muWG,muWS,Rpars,rpa,Gpars,Spars,doYear,doSpp) {
+  muWG=expandW(v,v,muWG)
+  muWS=expandW(v,v,muWS)
+  K.matrix=outer(v,v,make.K.values,muWG,muWS,Rpars,rpa,Gpars,Spars,doYear,doSpp)
+  return(h[doSpp]*K.matrix)
+}
+
+## Function to format the W matrix for the outer product
+expandW=function(v,u,W){
+  if(dim(W)[1]!=length(u)) stop("Check size of W")
+  Nspp=dim(W)[2]
+  W=as.vector(W)
+  W=matrix(W,length(W),ncol=length(v))
+  W=as.vector(t(W))
+  W=matrix(W,nrow=length(u)*length(v),ncol=Nspp)
+  return(W)
+}
+
+## Function to calculate size-dependent crowding, assuming no overlap
+# Growth
+wrijG=function(r,i,j){
+  return(2*pi*integrate(function(z) z*exp(-alphaG[i,j]*(z^2))*Cr[[j]](z-r),r,r+r.U[j])$value+
[…]

### 2016-03-09 runscripts/Idaho_NoARTR_sims/run_ipm_source.R

@@ -1,0 +1,142 @@
+############################################################
+##  IPM skeleton to be sourced after set up.
+##  
+##  Authors: Andrew Tredennick, Peter Adler, Chengjin Chu
+##  Email:   atredenn@gmail.com
+##  Created: 9-11-2015
+############################################################
+
+
+####
+####  Simulation length, Matrix size and initial vectors
+####
+v=v.r=b.r=expv=Cr=WmatG=WmatS=list(Nspp)
+h=r.L=r.U=Ctot=numeric(Nspp)
+for(i in 1:Nspp){
+  # minimum (0.9*minimum size from data) and maximum sizes (1.1*maximum size from data)
+  L=log(0.2)
+  U=log(maxSize[i])*1.1     
+  
+  # boundary points b and mesh points y. Note: b chops up the size interval (L-U) into bigM-equal-sized portions.
+  b = L+c(0:bigM[i])*(U-L)/bigM[i] 
+  
+  # v calculates the middle of each n-equal-sized portion.
+  v[[i]] = 0.5*(b[1:bigM[i]]+b[2:(bigM[i]+1)])
+  
+  # step size for midpoint rule. (see equations 4 and 5 in Ellner and Rees (2006) Am Nat.)
+  h[i] = v[[i]][2]-v[[i]][1]  
+  
+  # variables for Wr approximation
+  b.r[[i]]=sqrt(exp(b)/pi)
+  v.r[[i]]=sqrt(exp(v[[i]])/pi)
+  expv[[i]]=exp(v[[i]])
+  r.L[i] = sqrt(exp(L)/pi)
+  r.U[i] = sqrt(exp(U)/pi)
+  WmatG[[i]]=matrix(NA,length(v.r[[i]]),Nspp)  # storage of size-specific W values for each focal species
+  WmatS[[i]]=matrix(NA,length(v.r[[i]]),Nspp)
+} # next species
+
+tmp=range(v.r)
+size.range=seq(tmp[1],tmp[2],length=50) # range across all possible sizes
+
+
+## initial population density vector
+nt=v
+for(i in 1:Nspp) nt[[i]][]=0.001
+if(do_site=="Idaho") nt[[1]] <- 0
+new.nt=nt
+
[…]

### 2016-03-09 docs/ms_communitySynchrony.Rmd
first attempt at manuscript reording for #20

@@
 ------------
-The earth's ecosystems are in a period of upheaval due to species extinctions [@Cardinale2012], rapid changes in community composition [@Dornelas2014a; @Vellend2013], and human-driven changes in the abiotic environment [@Vitousek1997].
+The earth's ecosystems are in a period of rapid change due to species extinctions [@Cardinale2012], changes in community composition [@Dornelas2014a; @Vellend2013], and human-driven changes in the abiotic environment [@Vitousek1997].
 Such changes can have profound impacts on the provisioning and stability of ecosystem services [@Loreau2001a; @Hooper2005; @Rockstrom2009].
@@
 
-The mechanisms that underpin species synchrony are complex [@Loreau2008a], and whether synchrony in grassland communities is primarily an outcome of species-specific responses to environmental conditions [@Hautier2014] or competition [@Gross2014] remains controversial [reviewed in @Gonzalez2009].
-Fortunately, theoretical studies have identified the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and interspecific competition [@Loreau2008a; @Gonzalez2009; @Loreau2013].
-Unfortunately, quantifying the relative strengths of each driver on the degree of synchrony from community time series [e.g., @Houlahan2007] is impossible because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
+Theory identifies the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and interspecific competition [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+For example, in a community composed of large (no demographic stochasticity) and weakly interacting (no interspecific competition) populations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
+However, such a prediction relies on a relatively simple population dynamical model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) that all species have similar growth rates.
+Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains unkown.
+Indeed, no study has explicitly tested theory on the drivers of species synchrony in natural communities.
+
+In grasslands, most empirical studies have focused on whether species synchrony is primarily an outcome of species-specific responses to environmental conditions [@Hautier2014] or competition [@Gross2014].
+Even beyond grassland studies, whether competition or environmental responses drive compensatory dynamics remains controversial [reviewed in @Gonzalez2009].
+In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from community time series [e.g., @Houlahan2007] is impossible.
+This is because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
 For example, weak synchrony could reflect positive environmental correlations (synchronizing effect) offset by strong competition (desynchronizing effect), or negative environmental correlations and weak competition.
@@
 Furthermore, although @Mutshinda2009 quantify the relative importance of environmental stochasticity and competition to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony changes when each factor is removed.
-Lastly, their analysis did not include data from grassland communities, which have received the most experimental [@Tilman2001; @Hector2010] and theoretical [@DeMazancourt2013a] attention.
-Thus, we lack the ability to draw clear lines of inference from theoretical and experimental work to natural plant communities.
 
 Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony.
-First, we refine previous theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a].
-We use the theory to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
+Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific competition on species synchrony in natural plant communities.
+To achieve these objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
 We then confront our theoretical predictions with data from five semi-arid plant communities by using empirically-based population models.
-The fitted population models allow us to estimate species dynamics in monoculture and, thus, test our predictions.
 Second, we use the multi-species population models to perform simulation experiments designed to isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific competition on community-wide species synchrony.
@@
 -----------------
-### Materials and methods
+### The model
+Our objective is to derive analytical predictions for species synchrony under special limiting cases that may or may not be realized in natural communities.
+We constrain our focus to synchrony of per capita growth rates, rather than abundance, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
 Following @Loreau2008a and @DeMazancourt2013a, we define population growth, ignoring observation error, as
@@
 
-Our objective is to derive analytical predictions for species synchrony under special limiting cases.
-We constrain our focus to synchrony of per capita growth rates, rather than abundance, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
 To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION...
 
-### Results
-We derived predictions for synchrony of species' growth rates for two limiting cases that are analytically tractable.
-First, we assume no interspecific interactions, no environmental stochasticity, identical intrinsic growth rates, and that demographic stochasticity is operating but all species have identical demographic variances.
+### Predictions
+Our first prediction assumes no interspecific interactions, no environmental stochasticity, identical intrinsic growth rates, and that demographic stochasticity is operating but all species have identical demographic variances.
 This limiting case, $\mathcal{M}_{D}$, represents a community where dynamics are driven by demographic stochasticity alone.
@@
 
-<!--
-Our work relies on the data and population modeling approach presented by @Chu2015.
-We refer interested readers to @Chu2015 for full details on the focal plant communities, choice of species, structure of vital rate statistical models, and the ability of fitted population models to reproduce observed dynamics.
-We offer a brief overview here, but note that the only difference between our vital rate models and those fit by @Chu2015 is that we allowed for positive species interactions.
--->
-
 EMPIRICAL ANALYSIS
@@
 in preparation, @Chu2015] (Table 1).
-
-<!--
-### Analytical approach
-Our analysis involved five steps for each study site:
-
-1. Calculate observed and expected synchrony of per capita growth rates and percent cover
-2. Fit statistical models for survival, growth, and recruitment
-3.
-Build dynamic, multi-species population models: a spatially-implicit integral projection model (IPM; does not include demographic stochasticity) and a spatially-explicit individual based model (IBM; includes demographic stochasticity)
-4. Remove interspecific interactions from IPM simulations to isolate the effect of environmental responses
-5.
-Sequentially remove interspecific interactions, environmental fluctuations, and both (isolates effect of demographic stochasticity) from the IBM across a range of population sizes (effects strength of demographic stochasticity)
-
-Our approach has several advantages.
-First, we can use the models to directly compare synchrony from communities with and without different processes operating.
-This means we do not have to rely on biased null expectations, such as independent species fluctuations, to determine if competition or environmental responses are more important -- we can test for this directly.
-Second, our models represent natural communities, rather than experimentally-assembled communities.
-This means our results may be more generalizable then results obtained from biodiversity-ecosystem function experiments where strong, but transient, competition is a hallmark [@Gross2014].
-Lastly, our models are based on individual-level demographic data, which allows us to accurately estimate inter and intraspecific competition.
-
-All our code and the data necessary to reproduce the results reported in this paper can be found on GitHub (link) as release v1.0, and we have archived that release on Dryad (link).
-
--->
 
@@
 
-<!--
-We chose to use the special theoretical case where only demographic stochasticity is operating and all species have similar growth rates and variances to serve as an expectation of synchrony (see Appendix X).
-Specifically, based on a multi-species dynamic population model (Appendix X), we calculate expected synchrony of per capita growth rates ($\text{E}(\phi_r)$) as
-
-\begin{equation}
-\text{E}(\phi_r) = \frac{\sum_{i}p_{i}^{-1}}{\left(\sum_{i}p_{i}^{-1/2}\right)^{2}}
-\end{equation}
-
-\noindent and expected synchrony of percent cover ($\text{E}(\phi_N)$) as
-
-\begin{equation}
-\text{E}(\phi_N) = \frac{1}{\left(\sum_{i}p_{i}^{1/2}\right)^{2}}
-\end{equation}
-
-\noindent where $p_i$ is the average frequency of species *i* ($p_i = N_i/N_{Total}$).
-
-The difference between observed and expected synchrony represents the combined effect of competition and environmental stochasticity.
-However, it is important to note that comparing observed and expected synchrony does not yield insight on the relative contributions of competition and environmental stochasticity.
-The difference between observed and expected synchrony serves as a metric of how far natural communities are from idealized expectations.
-To understand the influence of competition and environmental stochasticity, we must turn to using population model simulations.
--->
-
 #### Fitting statistical models
-Vital rate regressions are the building blocks of our dynamic models (IBM and IPM).
+Vital rate regressions are the building blocks of our dynamic models: an integral projection model (IPM) and an individual-based model (IBM).
 We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment.
@@
 Crowding is directly calculated at each time step since each genet is spatially referenced (just like the actual data).
-Demographic stochasticity is an inherent feature of IBMs because individuals with the very same probability of survival, for example, can have different fates.
-The effect of demographic stochasticity on population dynamics depends on population size [@Lande2003], so we can control the strength of demographic stochasticity by simulating the IBM on different sized landscapes.
 Environmental stochasticity is not an inherent feature of IBMs, but is easily included since we fit year-specific temporal random effects for each vital rate regression.
 Thus, to include temporal environmental variation, at each time step we randomly choose a set of estimated survival, growth, and recruitment parameters specific to one observation year.
-We can ignore the random year effects to exclude environmental stochasticity in the IBM.
 For all simulations, we ignore the spatial random effect associated with variation among quadrat groups, so our simulations represent an average quadrat for each site.
@@
 
-#### Removing interspecific interactions (IPM)
-To remove interspecific interactions, we simply set the off-diagonals of the interaction matrix for each vital rate regression to zero.
+#### Simulation experiments
+We performed removal simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) were removed one-by-one and combination.
+To remove interspecific interactions, we set the off-diagonals of the interaction matrix for each vital rate regression to zero.
 This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are effectively growing in monoculture.
-We ran IPM simulations with interspecific interactions turned on/off for 2,000 time steps, after an initial 500 iteration burn-in period.
+To remove the effect of a fluctuating environment, we removed the temporal (interannual) random effects from the regression equations.
+To remove the effect of demographic stochasticity, we use the IPM rather than the IBM because the IPM does not include demographic stochasticity and demographic stochasticity cannot be removed from IBM simulations.
+Since the effect of demographic stochasticity on population dynamics depends on population size [@Lande2003], we can control the strength of demographic stochasticity by simulating the IBM on different area sizes (e.g., plot sizes).
+Note that the IPM must include environmental stochasticity because otherwise all populations go to a stable equilibrium, and the IBM must include demographic stochasticity because it explicitly models individual plants.
+
+Using the above methods, we ran six simulation experiments:
+(1) All drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) present using the IBM;
+(2) Demographic stochasticity removed using the IPM;
+(3) Environmental stochastcity removed suing the IBM;
+(4) Interspecific competition removed using the IBM;
+(5) Interspecific competition and demographic stochasticity removed using the IPM;
+(6) Interspecific competition and environmental stochasticity removed using the IBM.
+We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
 This allowed species time to reach their stable size distribution.
 We then calculated the synchrony of species' per capita growth rates over 100 randomly selected contiguous 50 time-step sections.
-These simulations allowed us to isolate the effect of removing competition in large populations where demographic stochasticity is absent.
-We also used simulation results from species growing in monoculture to estimate the covariance of species' responses to the environment ($\text{cov}(\mu_{ie}, \mu_{je})$) for parameterizing Equation 3.
-
-#### Removing interspecific interactions and environmental fluctuations (IBM)
-Since the IBM and IPM are driven by the same vital rate regressions, interspecific interactions were removed as described above for the IPM simulations.
-To remove the effect of a fluctuating environment, we simply removed the temporal (interannual) random effects from the regression equations.
-Thus, if we remove interspecific interactions and the temporal random effects, the only force causing population fluctutations through time is demographic stochasticity.
-We ran a series of simulations for four "experiments": (1) environmental stochasticity on and interspecific interactions present, (2) environmental stochasticity off and interspecific interactions present, (3) environmental stochasticity on and interspecific interactions absent, and (4) environmental stochasticity off and interspecific interactions absent (only demographic stochasticity).
-We ran IBM simulations for 100 time steps, and repeated the simulations 100 times for each "experiment" (interspecific interactions on/off with and without environmental stochasticity).
+We ran IBM simulations for 100 time steps, and repeated the simulations 100 times for each simulation experiment.
 From those, we retained only the simulations in which no species went extinct due to demographic stochasticity.
 Synchrony was calculated over the 100 time steps for each no extinction run within a model experiment.
+To explore the effect of demographic stochasticity in different sized populations, we ran simulations (1) and (6) on plot sizes of 1, 2, 3, 4, and 5 square meters.
+
+Results from our simulation experiments also allow us to test our theoretical predictions.
+First, in the absence of interspecific competition and demographic stochasticity, species are effectively growing in large monocultures.
+Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(\mu_{ie}, \mu_{je})$) and parameterize Equation 3.
+Parameterizing Equation 2 does not require simulation output because the only parameters are the species' relative abundances.
+Second, simulations (5) and (6) represent the simulated version of our limiting case theoretical predictions.
+Thus, we directly test the theoretical predictions by comparing them to observed synchrony and simulated synchrony.
 
@@
 Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more noisy, as expected by theory [@Loreau2008a].
-
-<!--
-However, the synchrony of species' low-density growth rates, a more accurate measure of a species' environmental response, was a less accurate predictor of species synchrony in polyculture (Fig. 4B).
-Except for synchrony of per capita growth rates in the Idaho and Montana communities, synchrony of low density growth rates in monoculture underpredicted synchrony in polyculture (Fig. 4B).
--->
 

### 2016-03-09 runscripts/Idaho_NoARTR_sims/ibm_largepop_wrapper.R

@@ -5,49 +5,56 @@
 
 ##  Author:       Andrew Tredennick, Peter Adler
 ##  Email:        atredenn@gmail.com
-##  Date created: 10-23-2015
+##  Date created: 12-15-2015
 
 
-## Clear the workspace
-rm(list=ls())
 
+####
+####  Set Simulation ID
+args <- commandArgs(trailingOnly = F)
+myargument <- args[length(args)]
+myargument <- sub("-","",myargument)
+sim_id <- as.numeric(myargument)
+# sim_id=1
 
 ####
 ####  Load libraries -----------------------------------------------------------
 ####
-library(communitySynchrony) # inhouse package
-library(synchrony)          # calculates Claire's synchrony metric
 library(boot)               # for the IBM
 library(mvtnorm)            # for the IBM
 library(msm)                # for the IBM
+source("import_regression_params_fxns")
 
 
 
 ####
 ####  Set up global variables and simulation config ----------------------------
 ####
-totSims <- 20      # number of simulations per site (1 here since using large landscape)
+totSims <- 100      # number of simulations per site (1 here since using large landscape)
 totT <- 100      # time steps of simulation
 burn.in <- 25    # time steps to discard before calculating cover values
 L <- 100          # dimension of square quadrat (cm)
 doGroup <- NA     # NA for spatial avg., values for a specific group
-constant.vec <- c(FALSE, FALSE, TRUE, TRUE) # TRUE for constant env.; FALSE for random year effects
-sppinter.vec <- c(FALSE, TRUE, FALSE, TRUE) # TRUE for interspp interactions; FALSE for no interspp interactions
-filename.flag <- c("fluctnointer", "fluctinter", "constnointer", "constinter")
-# constant.vec <- c(FALSE, TRUE, TRUE) # TRUE for constant env.; FALSE for random year effects
-# sppinter.vec <- c(TRUE, FALSE, TRUE) # TRUE for interspp interactions; FALSE for no interspp interactions
-# filename.flag <- c("fluctinter", "constnointer", "constinter")
-
-## Looping over different landscape sizes
+constant.vec <- c(FALSE,TRUE) # TRUE for constant env.; FALSE for random year effects
[…]

### 2016-03-09 runscripts/Idaho_NoARTR_sims/ibm_skeleton.R

@@ -177,7 +177,7 @@ for(iSim in 1:totSims){
   tmp=c(iSim,1,0,A,N)
   output=rbind(output,tmp)
   
-  pb <- txtProgressBar(min=2, max=totT, char="+", style=3, width=65)
+  # pb <- txtProgressBar(min=2, max=totT, char="+", style=3, width=65)
   for(tt in 2:(totT)){
     
     # draw year effects
@@ -240,6 +240,6 @@ for(iSim in 1:totSims){
       }
     }
     
-    setTxtProgressBar(pb, tt)
+    # setTxtProgressBar(pb, tt)
   } # next tt
 } # next iSim

### 2016-03-09 runscripts/Idaho_NoARTR_sims/import_regression_params_fxns.R

@@ -1,0 +1,130 @@
+#' Import and format growth parameters for IPM
+#' 
+#' @author Andrew Tredennick
+#' @param do_site Focal site (character scalar).
+#' @param species_list Character vector of four letter species' codes.
+#' @param Nyrs Number of random effects years.
+#' @param Gdata_species Growth parameters matrix from regression output.
+
+format_growth_params <- function(do_site, species_list, Nyrs, Gdata_species){
+  Nspp <- length(species_list)
+  Ngrp <- length(which(!is.na(Gdata_species[[1]]$Group)))
+  Gpars <- list(intcpt=rep(NA,Nspp),intcpt.yr=matrix(0,Nyrs,Nspp), 
+                intcpt.gr=matrix(0,Ngrp,Nspp),
+                slope=rep(NA,Nspp),slope.yr=matrix(0,Nyrs,Nspp),
+                nb=matrix(0,Nspp,Nspp),alpha=matrix(NA,Nspp,Nspp),
+                sigma2.a=rep(NA,Nspp),sigma2.b=rep(NA,Nspp))
+  
+  for(i in 1:Nspp){
+    Gdata <- Gdata_species[[i]]
+    Gpars$intcpt[i]=Gdata$Intercept[1]
+    tmp=which(names(Gdata)=="Group")
+    if(length(tmp)>0) Gpars$intcpt.gr[,i]=Gdata$Group[!is.na(Gdata$Group)] 
+    Gpars$intcpt.yr[,i]=Gdata$Intercept.yr
+    Gpars$slope[i]=Gdata$logarea.t0[1]
+    # random effects on slope
+    tmp=which(names(Gdata)=="logarea.t0.yr")
+    if(length(tmp)>0) Gpars$slope.yr[,i]=Gdata[,tmp]
+    # get competition coefficients
+    tmp=paste("crowdV",1:length(species_list),sep="")
+    tmp=which(is.element(names(Gdata),tmp))
+    if(length(tmp)>0) Gpars$nb[i,]=as.numeric(Gdata[1,tmp])
+    
+    Gpars$alpha[i,]=Gdata$alpha[1:length(species_list)]
+    Gpars$sigma2.a[i]=Gdata$sigma.a[1]
+    Gpars$sigma2.b[i]=Gdata$sigma.b[1]
+  } # next i
+  return(Gpars)
+} # end function
+
+
+
+
+#' Import and format survival parameters for IPM
+#' 
+#' @author Andrew Tredennick
+#' @param do_site Focal site (character scalar).
[…]

### 2016-03-09 runscripts/Idaho_NoARTR_sims/run_ibm.R

@@ -1,389 +1,0 @@
-##  Individual-based model with no environmental variance
-##  Used to test against the demographic approximation of the IPM
-
-######  FROM PETER #########################################
-# Individually-based model for a multiple species,
-# with density-dependence and explicit space (random
-# spatial pattern)
-
-# Gaussian distance function
-
-# This version uses the negative binomial recruitment function
-
-# Initialized from observed size distributions
-############################################################
-
-##  Clear the workspace and load libraries
-rm(list=ls())
-library(communitySynchrony)
-library(synchrony)
-library(boot)
-library(mvtnorm)
-library(msm)
-
-
-make.P.values <- function(v,u,muWG,muWS, #state variables
-                          Gpars,Spars,doYear,doSpp){  #growth arguments
-  S(u,muWS,Spars,doYear,doSpp)*G(v,u,muWG,Gpars,doYear,doSpp) 
-}
-
-make.P.matrix <- function(v,muWG,muWS,Gpars,Spars,doYear,doSpp) {
-  muWG=expandW(v,v,muWG)
-  muWS=expandW(v,v,muWS)
-  
-  P.matrix=outer(v,v,make.P.values,muWG,muWS,Gpars,Spars,doYear,doSpp)
-  return(h[dospps]*P.matrix)
-} 
-
-
-####
-####  Set some global variables for the simulation
-####
-totSims=100
-totT=150     # time steps of simulation
-burn.in=50  # time steps to discard before calculating cover values
-L=100       # dimension of square quadrat (cm)
-expand=2    # 1 = 1x1 m^2, 2 = 2x2m^2, etc
-init.cover=c(0,1,1,1)   # in % cover
-# maxSize=c(8000,500,500,500)
-maxSize=c(3000,202,260,225) 
-minSize=0.2
-iter_matrix_dims <- c(50,75,50,75)
-do_site="Idaho"
-Nyrs=22
-myCol=c("black","darkgreen","blue","red")
-doGroup=NA  # NA for spatial avg., values 1-6 for a specific group
-constant=T        
-
-out.nt <- list()
-out.nt[[1]] <- array(0,c(totT,iter_matrix_dims[2], totSims))
[…]

### 2016-03-09 docs/ms_communitySynchrony.Rmd
updates figures and figure referencing

@@
 Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out five communities (Fig.
-2a).
+2).
 Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
-Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2a), but the effect was small (9% increase).
-Species interactions are weak in these communities [Table SX and @Chu2015], so removing interspecific competition had little effect on synchrony (Fig. 2a; "No Comp." simulations).
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase).
+Species interactions are weak in these communities [Table SX and @Chu2015], so removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
 Removing competition caused, at most, a 5% change in synchrony (Table SX).
-Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2a), with an average 6% increase.
+Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2), with an average 6% increase.
 The desynchronizing effect of demographic stochasticity was able to modestly counteract the synchronizing force of the environment, but not enough to match the synchrony of species when only demographic stochasticity is operating (Fig.
-2b).
+3).
 For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
@@
 We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a and Supplementary Information].
-Contrary to expectations, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2a).
+Contrary to expectations, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
 The lack of an effect of interspecific competition on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
@@
 Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small (Fig.
-2b), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 2b).
+3), in agreement with theory [@Loreau2008a].
+Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 3).
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
-Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of specuies) in a non-constant environment (Fig. 2b).
+Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of specuies) in a non-constant environment (Fig. 3).
 This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
@@
 1).
-Second, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2a).
+Second, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
 Lastly, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
-2a), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig.
+2), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig.
 3).
@@
 Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
-When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2a).
+When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2).
 A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
@@
   \centering
-      \includegraphics[width=6in]{./components/all_sims_results.png}
+      \includegraphics[width=6in]{./components/formatted_fig2.png}
   \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
@@
 Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation.
-(b) Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S.
-+ No Comp.'').
-The strength of demographic stochasticity decreases as landscape size increases.}
+Error bars represent the upper and lower 95\% quantiles from model simulations.}
 \end{figure}
@@
 + No Comp.'').
-The strength of demographic stochasticity decreases as landscape size increases.}
+The strength of demographic stochasticity decreases as landscape size increases.
+Error bars represent the upper and lower 95\% quantiles from model simulations.}
 \end{figure}

### 2016-03-10 docs/ms_communitySynchrony.Rmd
adds new description of Idaho outlier status

@@
 
-The insensitivity of species synchrony to model perturbations in the Idaho community challenges the consistency of our results, but its outlier status may have more to do with environmental variability than intrinsic species synchrony.
-Regardless of how synchronously or asynchronously species in a community respond to environmental fluctuations, if the environment does not fluctuate substantially then removing such fluctuations will not impact species synchrony.
-Indeed, the Idaho community had the lowest variability in percent cover over the observation period (Table 2).
-At the same time, correlations among species' temporal random effects for each vital rate in the Idaho community were similar to, or greater than, those in the other communities we studied (Table SX).
-These lines of evidence lead us to conclude that either environmental fluctuations are weak in the Idaho community or our model was unable to detect temporal fluctuations in the data.
-In other words, species' responses to environmental conditions are likely the main determinant of synchrony in the Idaho community, but we were unable to detect the effect of the environment due to model misspecification or a lack of a strong environmental signal in the data.
+The insensitivity of species synchrony to model perturbations in the Idaho community challenges the consistency of our results.
+Idaho's outlier status is due to its functional diversity.
+In fact, when we removed the shrub (*Artemisia tripartita*) from our analyses, results from the Idaho data set conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
+SX).
+Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.
+*A. tripartita* has a facilitative effect on each grass species across all vital rates, except for a small competitive effect on *H. comata*'s survival probability (Tables SX, SX, SX).
+At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables SX, SX, SX).
+Thus, interspecific interactions play a relatively large role in determining synchrony in Idaho, but only when *A. tripartita* is at equilibrium abundance.
+The removal of environmental variation causes a decrease in the abundance of *A. tripartita* and, in turn, a weakening of the overall effect of competition on species synchrony.
+This would explain why synchrony actually increases when we remove environmental fluctuations in Idaho (Fig. 2).
 

### 2016-03-10 docs/ms_communitySynchrony.Rmd
restructures ms to close #20

@@
 
-\textbf{Number of Tables and Figures}:  2 Tables;  4 Figures
+\textbf{Number of Tables and Figures}:  3 Tables;  3 Figures
 
@@
 Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
-
-<!--
-One potential way forward is to use population models fit to empirical data to understand the influence of different drivers on species synchrony.
-@Mutshinda2009 used such an approach by fitting a dynamic population model to several community time series of insect and bird abundances.
-They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
-While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Despite almost a decade of research aimed toward understanding species synchrony, we still lack direct tests that link environmental fluctuations, interspecific interactions, and demographic stochasticity to species synchrony in natural communities.
-
-Here, we directly test the importance of proposed drivers of species synchrony by parameterizing and simulating dynamic, multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
-We use the fitted population models to simulate communities with and without the three mechanisms identified by theory as drivers of species synchrony: interspecific interactions, environmental forcing, and demographic stochasticity [@Loreau2008a; @Loreau2013; @DeMazancourt2013a].
-Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
-Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
-Our main focus is on the synchrony of species' per capita growth rates since they represent a species immediate response to environmental conditions and competition.
-However, we also analyze synchrony of plant cover (a metric of abundance) to relate our findings to previous research.
--->
 
@@
 (4) Interspecific competition removed using the IBM;
-(5) Interspecific competition and demographic stochasticity removed using the IPM;
+(5) Interspecific competition and demographic stochasticity removed using the IPM; and
 (6) Interspecific competition and environmental stochasticity removed using the IBM.
@@
 ----------
-Our study produced three main findings that were generally consistent across five natural plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
-These results are in agreement with our theoretical prediction for species synchrony under the limiting case of community dynamics being driven by environmental fluctuations alone.
+Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by analagous simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.
 
-### Interspecific interactions had little impact on community synchrony
-We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a and Supplementary Information].
-Contrary to expectations, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
-The lack of an effect of interspecific competition on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
-
-Our results do not contradict the idea that competition can lead to compensatory dynamics, but they do highlight the fact that interspecific competition must be relatively strong to influence species synchrony.
-The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables X1-X10).
-Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap increases.
-In that sense, our results are not surprising: interspecific interactions are weak, so of course removing them does not affect synchrony.
-More surprising is that our focal communities have little niche overlap. 
-But, as discussed by @Chu2015, little niche overlap becomes less surprising when we consider the virtually limitless niche axes upon which species can partition themselves [@Clark2010].
+### Simulations support theoretical predictions
 
 ### Demographic stochasticity decreases synchrony
-Our theoretical prediction for species synchrony in a neutral community with no species interactions or environmental fluctuations ($\mathcal{M}_D$) received no support when parameterized with emprical estimates (Fig.
-1).
 Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small (Fig.
@@
 Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment [@Loreau2008a].
-Several lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
-First, of our two limiting-case theoretical predictions for species synchrony (Equations 3-4), only our prediction for communities driven by environmental fluctuations alone ($\mathcal{M}_E$) received empirical support (Fig.
-1).
-Second, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
-Lastly, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
+Two lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
+First, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
+Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
 2), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig.
@@
 Likewise, our results appear to be quite general: across four arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
-Only in the Idaho community did not removing environmental fluctuations not cause a large decrease in synchrony, but removing competition had no effect either.
+Only in the Idaho community did removing environmental fluctuations not cause a large decrease in synchrony, but removing competition had no effect either.
 Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural plant communities.
@@
 
-Another challenge to the generality of our results is a recent analysis of several biodiversity-ecosystem function experiments that came to a completely opposing conclusion: competition drives species synchrony in grasslands [@Gross2014].
-@Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
+### Interspecific interactions had little impact on community synchrony
+We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a and Supplementary Information].
+Contrary to expectations, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
+The lack of an effect of interspecific competition on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
+
+Our results do not contradict the idea that competition can lead to compensatory dynamics, but they do highlight the fact that interspecific competition must be relatively strong to influence species synchrony.
+The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables X1-X10).
+Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap increases.
+In that sense, our results are not surprising: interspecific interactions are weak, so of course removing them does not affect synchrony.
+More surprising is that our focal communities have little niche overlap. 
+But, as discussed by @Chu2015, little niche overlap becomes less surprising when we consider the virtually limitless niche axes upon which species can partition themselves [@Clark2010].
+
+Our results are in direct contrast to a recent analysis of several biodiversity-ecosystem function experiments that found competition drives species synchrony in grasslands [@Gross2014].
 This is surprising given the extremely weak species interactions we estimated.
@@
 One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.
-
-<!--
-### The importance of functional diversity
-Our analysis also contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
-Functional traits define how a species will respond to its environment.
-Therefore, greater functional differences among species should lead to lower correlations in species' responses to the environment.
-We found evidence for the negative effect of functional diversity on community synchrony: community synchrony was lowest Idaho (Table 2, Fig.
-2); the only community that contained two life forms (one shrub, three perennial grasses).
-In fact, when we ran simulations with the shrub removed, synchrony of the Idaho community increased substantially and was similar to synchrony in the Kansas community (Fig.
-SX), which also contained three species.
-Functional diversity is especially important in communities with weak interspecific interactions where synchrony depends soley on species' environmental responses.
-Adding species to such communities will always decrease synchrony, and increase temporal stability, so long as the synchrony of species' environmental response is less than 1 [@Loreau2008a; @Loreau2013].
-But functionally diverse suites of species will have the lowest synchrony of environmental responses, and thus lowest community synchrony and highest stability.
--->
 
@@
 
-table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
+table_data <- as.data.frame(rbind(get_table_metrics(out_az),
                                   get_table_metrics(out_id),
+                                  get_table_metrics(out_ks),
                                   get_table_metrics(out_mt),
-                                  get_table_metrics(out_nm),
-                                  get_table_metrics(out_az)))
-table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
+                                  get_table_metrics(out_nm)))
+alltable_data <- table_data
+table_data$Site <- c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico")
 table_data <- table_data[,c("Site", "V1", "V4", "V7")]
-table_data$spp <- c(length(unique(ks_data$species)),
+table_data$spp <- c(length(unique(az_data$species)),
                     length(unique(id_data$species)),
+                    length(unique(ks_data$species)),
                     length(unique(mt_data$species)),
-                    length(unique(nm_data$species)),
-                    length(unique(az_data$species)))
+                    length(unique(nm_data$species)))
 colnames(table_data) <- c("Site",
@@
 print(xtable(table_data, caption = synch_cap),
+      caption.placement="top",
+      include.rownames = F, 
+      sanitize.colnames.function = identity,
+      comment=FALSE,
+      size="normalsize")
+```
+
+\pagebreak{}
+
+```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}
+exp_synchrony_env <- readRDS("../results/envstoch_predictions.RDS")
+exp_synchrony_env <- exp_synchrony_env[complete.cases(exp_synchrony_env),]
+exp_synchrony_demo <- alltable_data[,c("V2","V5")]
+exp_synchrony_demo$site <- table_data$Site
+colnames(exp_synchrony_demo) <- c("pgr", "cover", "site")
+sims_output <- readRDS("../results/allsims_plot_data.RDS")
+
+pred_table <- data.frame(site = table_data$Site,
+                         exp_demo = exp_synchrony_demo$pgr,
+                         sim_demo = round(subset(sims_output, simulation=="6No Comp. + No E.S.")["synchrony"],2),
+                         exp_env = round(exp_synchrony_env$growthrate_prediction,2),
+                         sim_env = round(subset(sims_output, simulation=="5No Comp. + No D.S.")["synchrony"],2))
+
+colnames(pred_table) <- c("Site", 
+                          "Predicted $\\phi_{R,\\mathcal{M}_D}$",
+                          "Simulated $\\phi_{R,\\mathcal{M}_D}$",
+                          "Predicted $\\phi_{R,\\mathcal{M}_E}$",
+                          "Simulated $\\phi_{R,\\mathcal{M}_E}$")
+
+pred_cap <- "Predicted and simulated synchrony of per capita growth rates under the two limiting cases."
+print(xtable(pred_table, caption = pred_cap),
       caption.placement="top",
@@
 \pagebreak{}
+-->
 
@@
 
-<!--
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{./components/ibm_sims_fig.png}
-  \caption{Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations.
-Points represent replicate IBM simulations (out of 100 possible) in which no species went extinct.
-Lines show linear fits between plot size (proxy for population size) and species synchrony.
-Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes).
-Simulation type codes: C = constant environment, F = fluctuating environment, INTER = interspecific interactions included, NoINTER = interspecific interactions excluded (set to zero).}
-\end{figure}
--->
-
-<!--
-\pagebreak{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/envresp_vs_ipmsynch.png}
-  \caption{Average synchrony of IPM-simulated per capita growth rates (from 13 year sections) as a function of the average synchrony of the vital rate random year effects for each site.
-Community sychrony is expected to approximately equal the synchrony of species' environmental responses in the absence of demographic stochasticity and strong interspecific interactions.
-The synchrony of species' random year effects across vital rates is a strong ($\text{R}^2$ = 0.67), but not perfect, predictor of community synchrony.
-Points are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.}
-\end{figure}
--->
-\pagebreak{}
-
 \begin{figure}[!ht]
@@
 
-\pagebreak{}
-
 References

### 2016-03-10 devel/make_ms_plots.R

@@ -121,6 +121,9 @@ plot_df$control <- rep(plot_df[which(plot_df$simulation=="1All Drivers"),"synchr
 plot_df$percent_diff <- with(plot_df, abs(synchrony-control)/((synchrony+control)/2)*100)
 write.csv(plot_df, "../results/synchsims_percent_diffs.csv")
 
+##  Save the plot data
+saveRDS(plot_df, "../results/allsims_plot_data.RDS")
+
 ##  Make demographic stochasiticty plot for all landscape sizes -----
 ibm_demo_rms <- subset(ibm_synch_agg, typesynch=="Per capita growth rate")
 ibm_demo_rms <- ibm_demo_rms[which(ibm_demo_rms$experiment %in% c("fluctinter", "constnointer")),]

### 2016-03-10 docs/ms_communitySynchrony.Rmd
updates theory discussion and prediction table caption

@@
 This is because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
-For example, weak synchrony could reflect positive environmental correlations (synchronizing effect) offset by strong competition (desynchronizing effect), or negative environmental correlations and weak competition.
+For example, weak synchrony of population abundances could reflect positive environmental correlations (synchronizing effect) offset by strong competition (desynchronizing effect), or negative environmental correlations and weak competition.
 
@@
 To estimate environmental responses in natural communities, we turn to our population models built using long term demographic data.
+
+Qualitative predictions here...?
 
@@
 
-Across the five communities, our limiting case prediction of community-wide synchrony when only environmental stochasticity is operating ($\mathcal{M}_E$) best matched observed synchrony for per capita growth rates (Fig.
-1).
-The regression line between observed growth rate synchrony and predicted growth rate synchrony under prediction $\mathcal{M}_E$ was the only one whose slope and intercept were not significantly different from zero (*P* = 0.88) and one (*P* = 0.87), respectively (Supplementary Information X).
-The regression line between observed and predicted growth rates under prediction $\mathcal{M}_D$ (demographic stochasticity only) had an intercept and slope significantly different from zero and one, respectively (Fig.
-1).
+Across the five communities, our limiting case predictions closely matched synchrony from simulations designed to test each prediction (Table 3).
+The correlation between our analytical predictions and simulated synchrony was 0.97 for $\phi_{R,\mathcal{M}_D}$ and 0.997 for $\phi_{R,\mathcal{M}_E}$.
+The largest difference between predicted and simulated synchrony was 0.05 in New Mexico for $\phi_{R,\mathcal{M}_D}$ (Table 3).
 
@@
 ### Simulations support theoretical predictions
+Our theoretical predictions were derived from a relatively simple model of population dynamics and required several simplifying assumptions.
+Yet, our theoretical predictions were strongly supported by our empirically-based model simulations of natural plant communities (Table 3).
+Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony.
+The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem function experiments [@DeMazancourt2013a].
+Thus, there is accumulating evidence, now from experimental and observational studies, that the population dynamical approach is an extremely useful modeling framework for linking population dynamics to community and ecosystem level properties [@Loreau2010].
 
@@
 3), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 3).
+Even in small populations, however, demographic stochasticity is not a strong enough to completely counteract the synchronizing effects of environmental fluctuations and induce independent species fluctuations (Fig.
+3).
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
@@
 The insensitivity of species synchrony to model perturbations in the Idaho community challenges the consistency of our results.
-Idaho's outlier status is due to its functional diversity.
-In fact, when we removed the shrub (*Artemisia tripartita*) from our analyses, results from the Idaho data set conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
+However, Idaho's outlier status is soley due to its functional diversity.
+When we removed the shrub (*Artemisia tripartita*) from our analyses, results from the Idaho data set conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
 SX).
@@
 At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables SX, SX, SX).
-Thus, interspecific interactions play a relatively large role in determining synchrony in Idaho, but only when *A. tripartita* is at equilibrium abundance.
-The removal of environmental variation causes a decrease in the abundance of *A. tripartita* and, in turn, a weakening of the overall effect of competition on species synchrony.
-This would explain why synchrony actually increases when we remove environmental fluctuations in Idaho (Fig. 2).
+We know synchrony is affected by interspecific competition [@Loreau2008a], but how facilitative effects manifest themselves is unknown.
+The interaction of facilitation and competition is clearly capable of having a large effect on species synchrony, and future theoretical efforts should aim to include a wider range of species interactions.
 
@@
 
-pred_cap <- "Predicted and simulated synchrony of per capita growth rates under the two limiting cases."
+pred_cap <- "Comparisons between our analytical predictions and simulation results for synchrony of species' per capita growth rates.
+Analytical predictions represent two limiting cases where only demographic stochasticity is operating ($\\phi_{R,\\mathcal{M}_D}$) and where only environmental stochasticity is operating ($\\phi_{R,\\mathcal{M}_E}$).
+Simulated synchrony values come from our empirically-based, multi-species population models when simulated under conditions that match the limiting case conditions (e.g., environmental stochasticity and competition removed for $\\mathcal{M}_D$)."
 print(xtable(pred_table, caption = pred_cap),

### 2016-03-11 docs/ms_communitySynchrony.Rmd
adds pba edits to ms

@@
 \begin{abstract}
-Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem function, but elucidating the mechanisms underlying species synchrony remains a challenge.
-Here, we extend theory to predict species synchrony under limiting cases that we couple with an empirical analysis from five plant communities.
-When parameterized with empirical estimates, our prediction of species synchrony under only environmental stochasticity best predicts observed synchrony of per capita growth rates.
-In model simulations, average synchrony of per capita growth rates, which ranges from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem function, but identifying the mechanisms that determine synchrony remains a challenge.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific competition, and demographic stochasticity.
+We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
+In all five communities, theoretical predictions closely matched simulated synchrony for the limiting cases where only demographic stochasticity or environmental variation are present, meaning that theory developed with simple models also applies to our empirical models.
+Simulations experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
 Removing interspecific competition and demographic stochasticity had small effects on synchrony.
-Our results suggest that all five communities represent the theoretical limiting case where demographic stochasticity and competition are weak, and community-wide synchrony can be predicted by the synchrony of species' environmental responses.
+In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
 \end{abstract}
@@
 ------------
-The earth's ecosystems are in a period of rapid change due to species extinctions [@Cardinale2012], changes in community composition [@Dornelas2014a; @Vellend2013], and human-driven changes in the abiotic environment [@Vitousek1997].
-Such changes can have profound impacts on the provisioning and stability of ecosystem services [@Loreau2001a; @Hooper2005; @Rockstrom2009].
-In particular, studies of experimental grassland communities show that greater species richness leads to greater productivity [@Tilman2001] and temporal stability of productivity [@Tilman2006; @Hector2010].
+Ecosystems are being transformed by species extinctions [@Cardinale2012], changes in community composition [@Dornelas2014a; @Vellend2013], and anthropogenic environmental change [@Vitousek1997], impacting the provisioning and stability of ecosystem services [@Loreau2001a; @Hooper2005; @Rockstrom2009].
+Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
 The stabilizing effect of species richness stems from individual species fluctuating asynchronously through time [@Hector2010; @Isbell2009; @DeMazancourt2013a].
-Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], which implies that species losses will reduce ecosystem stability.
+Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
 Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem function [@Hautier2014].
-Thus, it is becoming clear that predicting the impacts of global change requires a mechanistic understanding of species synchrony in natural settings.
+Predicting the impacts of global change requires a mechanistic understanding of species synchrony in natural settings.
 
@@
 For example, in a community composed of large (no demographic stochasticity) and weakly interacting (no interspecific competition) populations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
-However, such a prediction relies on a relatively simple population dynamical model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) that all species have similar growth rates.
-Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains unkown.
-Indeed, no study has explicitly tested theory on the drivers of species synchrony in natural communities.
+However, such a prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
+Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains unkown: no study has explicitly tested theory on the drivers of species synchrony in natural communities.
 
@@
 The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific competition is to remove them one-by-one, and in combination.
-This could be done experimentally, but empirically-based models of interacting populations offer a way to exploit existing data sets from natural communities.
+In principle, this could be done experimentally, but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
 For example, @Mutshinda2009 fit a dynamic population model to several community time series of insect and bird abundances.
@@
 While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Furthermore, although @Mutshinda2009 quantify the relative importance of environmental stochasticity and competition to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony changes when each factor is removed.
+Furthermore, although @Mutshinda2009 quantified the relative importance of environmental stochasticity and competition to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
 
@@
 To achieve these objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
-We then confront our theoretical predictions with data from five semi-arid plant communities by using empirically-based population models.
-Second, we use the multi-species population models to perform simulation experiments designed to isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific competition on community-wide species synchrony.
-Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
+We then confront our theoretical predictions with simulations from the empirically-based population models.
+Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific competition on community-wide species synchrony.
+Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, and demographic and environmental stochasticity), and that these models successfully reproduce observed community dynamics [@Chu2015], perturbing the models can reveal the processes that determine species synchrony in our focal grassland communities.
 
@@
 ### The model
-Our objective is to derive analytical predictions for species synchrony under special limiting cases that may or may not be realized in natural communities.
-We constrain our focus to synchrony of per capita growth rates, rather than abundance, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
+While existing theory has identified the factors driving synchrony, we do not have a simple expression to predict synchrony in a particular community with all factors operating simultaneously.
+However, we can derive analytical predictions for species synchrony under special limiting cases.
+We constrain our focus to synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
 Following @Loreau2008a and @DeMazancourt2013a, we define population growth, ignoring observation error, as
@@
 Regression coefficients for the effect of crowding by each species can be considered a matrix of interaction coefficients whose diagonals represent intraspecific interations and whose off-diagonals represent interspecific interactions [@Adler2010].
+These interaction coefficients can take positive (facilitative) or negative (competitive) values.
 We modeled growth as the change in size of a genet from one year to the next, which depends on the same factors as the survival model.
@@
 We then project the model forward by using the survival regression to determine whether a genet lives or dies, the growth regression to calculate changes in genet size, and the recruitment regression to add new individuals that are distributed randomly in space.
-Crowding is directly calculated at each time step since each genet is spatially referenced (just like the actual data).
+Crowding is directly calculated at each time step since each genet is spatially referenced (as in the observed data).
 Environmental stochasticity is not an inherent feature of IBMs, but is easily included since we fit year-specific temporal random effects for each vital rate regression.
-Thus, to include temporal environmental variation, at each time step we randomly choose a set of estimated survival, growth, and recruitment parameters specific to one observation year.
+To include temporal environmental variation, at each time step we randomly choose a set of estimated survival, growth, and recruitment parameters specific to one observation year.
 For all simulations, we ignore the spatial random effect associated with variation among quadrat groups, so our simulations represent an average quadrat for each site.
 
-The IPM is similar to the IBM in that it is built using the vital rate regressions [@Rees2009; @Rees2014], but it is different in that it is spatially-implicit and does not include demographic stochasticity.
+The IPM uses the same vital rate regressions as the IBM [@Rees2009; @Rees2014], but it is spatially-implicit and does not include demographic stochasticity.
 Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step.
@@
 #### Simulation experiments
-We performed removal simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) were removed one-by-one and combination.
+We performed removal simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) were removed one-by-one and in combination.
 To remove interspecific interactions, we set the off-diagonals of the interaction matrix for each vital rate regression to zero.
@@
 To remove the effect of a fluctuating environment, we removed the temporal (interannual) random effects from the regression equations.
-To remove the effect of demographic stochasticity, we use the IPM rather than the IBM because the IPM does not include demographic stochasticity and demographic stochasticity cannot be removed from IBM simulations.
-Since the effect of demographic stochasticity on population dynamics depends on population size [@Lande2003], we can control the strength of demographic stochasticity by simulating the IBM on different area sizes (e.g., plot sizes).
-Note that the IPM must include environmental stochasticity because otherwise all populations go to a stable equilibrium, and the IBM must include demographic stochasticity because it explicitly models individual plants.
+To remove the effect of demographic stochasticity, we use the IPM rather than the IBM because the IPM does not include demographic stochasticity (demographic stochasticity cannot be removed from the IBM).
+Since the effect of demographic stochasticity on population dynamics depends on population size [@Lande2003], we can control the strength of demographic stochasticity by simulating the IBM on areas (e.g.
+plots) of different size.
 
-Using the above methods, we ran six simulation experiments:
-(1) All drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) present using the IBM;
-(2) Demographic stochasticity removed using the IPM;
-(3) Environmental stochastcity removed suing the IBM;
-(4) Interspecific competition removed using the IBM;
-(5) Interspecific competition and demographic stochasticity removed using the IPM; and
-(6) Interspecific competition and environmental stochasticity removed using the IBM.
+We conducted the following six simulation experiments:
+(1) IBM: All drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) present;
+(2) IPM: Demographic stochasticity removed;
+(3) IBM: Environmental stochastcity removed;
+(4) IBM: Interspecific competition removed;
+(5) IPM: Interspecific competition and demographic stochasticity removed;
+(6) IBM: Interspecific competition and environmental stochasticity removed.
 We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
@@
 Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
-For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig.
-X1), meaning that we can use the models to test the importance of different drivers of synchrony in these communities.
-However, IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. X1), likely due to the desynchronizing effect of demographic stochasticity.
+For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. X1).
+IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. X1), likely due to the desynchronizing effect of demographic stochasticity.
 
-Across the five communities, our limiting case predictions closely matched synchrony from simulations designed to test each prediction (Table 3).
+Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Table 3).
 The correlation between our analytical predictions and simulated synchrony was 0.97 for $\phi_{R,\mathcal{M}_D}$ and 0.997 for $\phi_{R,\mathcal{M}_E}$.
@@
 Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase).
-Species interactions are weak in these communities [Table SX and @Chu2015], so removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
+Species interactions are weak in these communities [Table SX and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
 Removing competition caused, at most, a 5% change in synchrony (Table SX).
-Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2), with an average 6% increase.
-The desynchronizing effect of demographic stochasticity was able to modestly counteract the synchronizing force of the environment, but not enough to match the synchrony of species when only demographic stochasticity is operating (Fig.
+Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
+2), with an average 6% increase [IS THIS FOR THE SMALLEST SIMULATED PLOT SIZE?].
+The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
 3).
 For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
-Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more noisy, as expected by theory [@Loreau2008a].
+Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more variable, as expected by theory [@Loreau2008a].
 
@@
 ----------
-Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by analagous simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.
+Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, though not perfectly, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.
 
 ### Simulations support theoretical predictions
-Our theoretical predictions were derived from a relatively simple model of population dynamics and required several simplifying assumptions.
-Yet, our theoretical predictions were strongly supported by our empirically-based model simulations of natural plant communities (Table 3).
+Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
+However, the theoretical predictions were strongly supported by simulations of population models fit to long-term data from natural plant communities (Table 3).
 Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony.
 The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem function experiments [@DeMazancourt2013a].
-Thus, there is accumulating evidence, now from experimental and observational studies, that the population dynamical approach is an extremely useful modeling framework for linking population dynamics to community and ecosystem level properties [@Loreau2010].
+The accumulating evidence from both experimental and observational studies demonstrates the utility of linking population dynamics to community and ecosystem level properties [@Loreau2010].
+[THIS LAST SENTENCE SUCKS]
 
 ### Demographic stochasticity decreases synchrony
-Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small (Fig.
+Demographic stochasticity partially counteracted the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations were small (Fig.
 3), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity is not a strong enough to completely counteract the synchronizing effects of environmental fluctuations and induce independent species fluctuations (Fig.
+Even in small populations, however, demographic stochasticity was not strong enough to completely counteract the synchronizing effects of environmental fluctuations and induce independent species fluctuations (Fig.
 3).
@@
 Two lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
-First, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
+First, in our simulation experiments, removing interspecific interactions resulted in no discernible change in community-wide species synchrony of per capita growth rates (Fig. 2).
 Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
-2), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig.
+2), and the synchrony of species in monoculture was a very strong predictor of synchrony of species in polyculture (Fig.
 3).
-Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony.
-Likewise, our results appear to be quite general: across four arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
-Only in the Idaho community did removing environmental fluctuations not cause a large decrease in synchrony, but removing competition had no effect either.
+Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony in the communities we studied.
 Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural plant communities.
 
-The insensitivity of species synchrony to model perturbations in the Idaho community challenges the consistency of our results.
-However, Idaho's outlier status is soley due to its functional diversity.
-When we removed the shrub (*Artemisia tripartita*) from our analyses, results from the Idaho data set conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
+In the Idaho community, removing environmental fluctuations did not cause a large decrease in synchrony.
+However, that result appears to be an artifact.
+Removing environmental variation results in a negative invasion growth rate for *A.
+tripartita*.
+Although we only analyzed IBM runs in which *A.
+tripartita* had not yet gone extinct, it was at much lower abundance than in the other simulation runs.
+When we removed the shrub (*Artemisia tripartita*) from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
 SX).
@@
 Under such conditions, synchrony will depend on the degree of pathogen-induced fitness differences and the pathogen's response to environmental conditions.
-Niether our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].
+Neither our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].
 
@@
 We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a and Supplementary Information].
-Contrary to expectations, community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
+We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
 The lack of an effect of interspecific competition on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
@@
 In that sense, our results are not surprising: interspecific interactions are weak, so of course removing them does not affect synchrony.
-More surprising is that our focal communities have little niche overlap. 
-But, as discussed by @Chu2015, little niche overlap becomes less surprising when we consider the virtually limitless niche axes upon which species can partition themselves [@Clark2010].
 
-Our results are in direct contrast to a recent analysis of several biodiversity-ecosystem function experiments that found competition drives species synchrony in grasslands [@Gross2014].
-This is surprising given the extremely weak species interactions we estimated.
-But the apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
-Gross et al.'s results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
-In contrast, the natural communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
-In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments community assembly processes that can result in competitive exclusion may yet to have been realized.
-Therefore, such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
-This also leads to the prediction that community synchrony is driven by competition in recently assembled communities, while synchrony is driven by environmental fluctuations in communities that have undergone extensive species sorting.
-In other words, in the absence of disturbance and invasion, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
-One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.
+However, our results do contrast with a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives species synchrony in grasslands [@Gross2014].
+The apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
+Gross et al.'s results are based on rapidly assembling experimental communities that are, at most, 11 years old.
+The natural, relatively undisturbed communities we studied consist of species that have co-occurred for decades [@Chu2015] and represent a much later stage of community assembly.
+In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments the processes driving competitive exclusion are still operating.
+Such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
+In contrast, synchrony may be driven by environmental fluctuations in older communities.
+In other words, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
+One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct repeat the analysis of @Gross2014 every few years.
 
@@
 The perennial grasses we studied do have similar responses to the environment (Table X1-X10), which will tend to synchronize dynamics.
-However, if community-wide species synchrony is less than 1, some degree of compensatory dynamics must be present [@Loreau2008a].
-Since observed and simulated synchrony is less than 1 in all our focal communities (Table 2 and Fig. 2), we conclude that compensatory dynamics are common in natural grasslands.
-But, in agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
+However, if community-wide species synchrony is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
+In agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
 Such a conclusion is in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
-It may be that recently assembled communities, such as experimental plots, exhibit strong species interactions that will weaken over time [@Kokkoris1999].
-Viewing community synchrony through the lens of community assembly processes may provide a way to reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.
+Recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
+Future research on the influence of community assembly on synchrony could reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.
 

### 2016-03-11 docs/cache/appendix_X1.Rmd
adds online supporting information document

--- 2015-10-23 docs/prelim_analysis.Rmd
@@
 ---
-# Environmental responses synchronize population \newline dynamics in five semi-arid grasslands
+# Appendix X for Environmental responses synchronize population \newline dynamics in five semi-arid grasslands: Extra Methods
 
@@
 
-Abstract
--------------
-Temporal asynchrony among interacting species is an important mechanism through which species richness can stabilize biomass production through time. 
-While asynchrony is a key mechanism linking biodiversity and ecosystem function, empirical support for the prevalance species asynchrony is mixed, and no studies have rigorously disentangled the forces driving species (a)synchrony in nature.
-Theory identifies three drivers of species synchrony: inter- and intraspecific density dependence sychronizes dynamics, demographic stochasticity desynchronizes dynamics, and environmental forcing can either synchronize or desynchronize poplation dynamics depending on the correlation of species-specific environmental responses.
-Here we use stochastic population models fit using long term datasets from five semi-arid grasslands to identify the forces that increase or decrease species synchrony through time. 
-We find that the trophically-similar grass species in our dataset have population dynamics that are synchronized by similar responses to environmental fluctuations. 
-Species interactions tend to be weak, so removing interspecific competition did not reduce synchrony. 
-The effect of demographic stochasticity depended on the spatial scale of our modeling; demographic stochasticity was strongest in small populations and was able to modestly counteract the synchronizing effect of the environment.
-Although we found that similar responses to environmental fluctuations synchronized species, dynamics were not completly synchronous. 
-We conclude that niche differentiation among species is strong enough to reduce the synchronizing effect of coupled density dependence via competition, but this is countered by the synchronizing effect of similar environmental responses.
-The next frontier is to identify the specific environmental drivers of species synchrony to inform predictive efforts.
+Detailed Study Area Descriptions
+--------------------------------
+We use the exact same data as presented in Chu and Adler (2015). From their paper:
 
-__Keywords__: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
+>Many range experiment stations in the western U.S.
+began mapping permanent quadrats in the early 20th century, and continued annual censuses for decades.
+Here we focus on five long-term data sets (Fig.
+1), four of which have been distributed publicly (Adler et al.
+2007b, Zachmann et al.
+2010, Anderson et al.
+2011, 2012).
+At each study site, all individual plants within each 1-m2 quadrat were identified and mapped annually using a pantograph (Hill 1920; Fig.
+2).
+Mapped polygons represent basal cover for grasses and canopy cover for shrubs (Fig.
+2).
 
-Introduction
-------------
-Asynchrony in population dynamics among co-occuring plant species stabilizes ecosystem-level properties like annual biomass production (Hector et al.
-2010, Loreau and de Mazancourt 2013, de Mazancourt et al.
-2013, Hautier et al.
-2014).
-Species-specific responses to environmental conditions, in particular, can stabilize ecosystem properties even as environmental forcings fluctuate through time. 
-Such asynchrony of species dynamics in response to the environment has the potential to buffer ecosystems subjected to an increasingly variable envirnmental conditions.
+>To fit our population models, we needed to study species common enough to provide a large sample size (i.e.
+the most common species at each site).
+In order to describe species interactions, we needed to select species that co-occurred together across many years.
+Thus, our analysis focused on the niche differences and average fitness differences of the most common, co-occurring species at each site, which implies that the average fitness differences among species could be relatively small.
+To select these species, we first identified species that occurred in at least 20% of years.
+Then we used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
+Based on the degree of aggregation of quadrats on the NMDS plot, we selected quadrats and species for our full analysis.
+We also confirmed these selections with local personnel from each study site.
+Below we further describe each study site and our criteria for quadrat and species selections.
 
-Theory suggests that community synchrony is driven by a combination of internal and external forces. 
-Internal to the community, demographic stochasticity can induce asynchronous dynamics when species tend to fluctuate independently in a constant environment (Loreau and de Mazancourt 2008). 
-However, species live in non-constant environments that act as strong external drivers of community dynamics. 
-Environmental variability can induce asynchronous dynamics when species' responses to environmental conditions are temporally uncorrelated. 
-When species respond similarly to environmental conditions, community dynamics become more synchronized (Gonzalez and Loreau 2009, Loreau and de Mazancourt 2008).
+>_Sonoran desert, Arizona_
 
-It is impossible to draw inference on the relative contribution of demographic and environmental stochasticity on synchrony based on empirical estimates alone, so our analysis involves two steps.  
-First, we will use longterm datasets from five North American grasslands (Fig. 1; Chu and Adler 2015) to calculate community synchrony. 
-Second, we will fit multispecies population models to the data for dominant species to evaluate the relative contribution of species' environmental responses, interspecific interactions, and demographic stochasticity to community synchrony.
-Sparse data on sub-dominant species constrained model fitting to dominant species that account for ##% - ##% of local plant abundance at their respective sites. 
-Our focus is on temporal trends of per capita growth rates becuase these best reflect the short-term responses of populations to environmental conditions that fluctuate on annual time scales (Loreau and de Mazancourt 2008).
+>The Sonoran desert data set comes from 178 permanent quadrats located on semi-desert grasslands at the Santa Rita Experimental Range (31º50’ N, 110º53’ W; elevation 1150 meters), Arizona (Anderson et al.
+2012).
+Mean annual precipitation was 450 mm with the monthly variation from 5 mm (May) to 107 mm (July), and mean annual temperature was 16 °C with the monthly variation from 8 °C (January) to 26 °C (July) during the sampling period.
+These quadrats were mapped annually from 1915 to 1933 in most cases.
+The dominant flora varies across the broad range of soils and topography that occur within the semi-desert grasslands.
+For our present analysis, we chose 32 quadrats dominated by black grama Bouteloua eriopoda (BOER) and Bouteloua rothrockii (BORO).
 
-What is expected by theory?[^devel]
----------------------------
-[^devel]: Note that this is still in development...so some of it may be wrong.
+>_Sagebrush steppe, Idaho_
 
-Loreau and de Mazancourt (2008) derived a statistic to measure community-wide synchrony of per capita population growth rates:
+>The sagebrush data set comes from the U.S.
+Sheep Experiment Station, currently a US Department of Agriculture Agricultural Research Service field station, located 9.6 km north of Dubois, Idaho (44.2°N, 112.1°W; elevation 1500 meters).
+26 quadrats were established between 1926 and 1932 (Zachmann et al.
+2010).
+During the period of sampling (1926 – 1957), mean annual precipitation was 270 mm with the monthly variation from 16 mm (March) to 42 mm (June), and mean annual temperature was 6 °C with the monthly variation from -8 °C (January) to 21 °C (July).
+The vegetation is dominated by the shrub, Artemisia tripartita (ARTR), and the C3 perennial bunchgrasses Pseudoroegneria spicata (PSSP), Hesperostipa comata (HECO), and Poa secunda (POSE).
 
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
-\end{equation}
+>_Southern mixed prairie, Kansas_
 
-where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
-The statistic is bounded at 0 and 1, where 0 indicates perfect asynchrony and 1 indicates perfect synchrony.
-When applied to a neutral model where species have perfect niche overlap but have independent demographic and environmental fluctuations through time,  expected synchrony ($\phi_r$) is defined as:
+>The southern mixed prairie data set represents the pioneering work of Albertson and colleagues (Albertson and Tomanek 1965), who established 51 permanent quadrats inside and outside livestock exclosures near Hays, Kansas (38.8°N, 99.3°W; elevation 650 meters).
+The mean annual precipitation was 580 mm with the monthly variation from 10 mm (January) to 103 mm (June), and mean annual temperature was 12 °C with the monthly variation from -2.0 °C (January) to 27 °C (July).
+These quadrats were mapped from 1932 to 1972 at the end of each growing season (Adler et al.
+2007b).
+Variation in soil depth and texture creates distinct plant communities.
+Many of the quadrats were located on shortgrass communities dominated by Bouteloua gracilis and Buchloë dactyloides.
+We had difficulty working with these species, which can be hard to distinguish in the field in the absence of inflorescences, and were often mapped inconsistently.
+Therefore, we focused on 7 quadrats dominated by Bouteloua curtipendula (BOCU), Bouteloua hirsuta (BOHI), and Schizachyrium scoparium (SCSC).
 
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(SN)}{[(1/S)(\sum_{i}\sqrt{\sigma^{2}_{c} + \sigma^{2}_{e} + \sigma^{2}_{d}/(N)})]^2}
-\end{equation}
+>_Northern mixed prairie, Montana_
 
-where $\sigma^{2}_{c}$ is the temporal variance of per capita population growth rates due to regulation of community size (density-dependence), $\sigma^{2}_{e}$ is environmental variance, and $\sigma^{2}_{d}$ is demographic variance.
-*S* is the species richness of the focal community and *N* is the harmonic temporal mean of either the entire community (no subscript in numerator) or species *i* (with subscript in denominator).
+>The northern mixed prairie data set comes from the Fort Keogh Livestock and Range Research Laboratory, another USDA Agricultural Research Service field station.
+The study site is located on alluvial plains near the Tongue River (46o19’N, 105o48’W; elevation 720 meters).
+Mean annual precipitation was 343 mm with the monthly variation from 7 mm (February) to 74 mm (June), and mean annual temperature was 8 °C with the monthly variation from -7 °C (January) to 25 °C (July).
+44 quadrats, distributed across six pastures assigned to one of three grazing intensities, were mapped annually (with some exceptions) from 1932 through 1945 (Anderson et al.
+2011).
+Our analysis includes 19 quadrats in which Bouteloua gracilis (BOGR), Hesperostipa comata (HECO), Pascopyrum smithii (PASM), and Poa secunda (POSE) were prevalent.
 
-Loreau and de Mazancourt show that when demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$), $\phi_r \approx 1$ meaning species are expected to fluctuate synchronously.
-At the other limit, where environmental forcing and community regulation are strong relative to demographic stochasticity ($\sigma^{2}_{c}+\sigma^{2}_{e} << \sigma^{2}_{d}/N_i$), species are expected to fluctuate independently, and $\phi_r \approx 1/S$ or:
+>_Chihuahuan desert, New Mexico_
 
-\begin{equation}
-\phi_{r} = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.
-\end{equation}
+>The last data set consists of 76 permanent quadrats located in Chihuahuan desert plant communities at the Jornada Experimental Range (32.62°N, 106.67°W; elevation 1260 meters), New Mexico.
+Mean annual precipitation was 264 mm with the monthly variation from 6 mm (April) to 48 mm (August), and mean annual temperature was 14 °C with the monthly variation from 4 °C (January) to 26 °C (July).
+Quadrats were established from 1915-1920 and mapping was conducted annually until the 1960’s (it continues on roughly 5 yr intervals to the present).
+Due to the changes in mapping frequency in the 1960’s, and the effects of a severe drought in the 1950’s that triggered a conversion of many quadrats from grassland to shrubland, we analyzed data for the period from 1915 to 1950.
+We selected 48 quadrats in which both Bouteloua eriopoda (BOER) and Sporobolus flexuosus (SPFL) were common (Anderson et al.
+in preparation).
 
-Based on the above theory, we expect synchrony in a community where only demographic stochasticity is operating on top of independent fluctuations to be $1/S$.
-
-Data and Methods
-----------------
-## Study sites and species
-In the early 20th century, several range experiment stations in the western U.S. began mapping permanent quadrats, and conducted annual censuses for decades. 
-We use such long-term demographic data sets from five semiarid grasslands in the western United States (described in detail by Chu and Adler 2015).
-Each study site includes a set of 1-$\text{m}^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
-The resulting mapped polygons represent basal cover for grasses and canopy cover for shrubs.
-Data come from the Sonoran desert in Arizona (Anderson et al.
-2012), sagebrush steppe in Idaho (Zachmann et al.
-2010), southern mixed prairie in Kansas (Adler et al.
-2007), northern mixed prairie in Montana (Anderson et al.
-2011), and Chihuahuan desert in New Mexico (Anderson et al.
-in preparation, Chu and Adler 2015).
-
-Chu and Adler describe the selection of focal species in each data set, so we offer a brief explanation here.
-Focal species needed to comply with two criteria: (1) be common enough to provide a large sample size for model fitting and (2) co-occur with other focal species often enough to allow us to describe species interactions within the model.
-Thus, Chu and Adler (2015) chose species that occurred in at least 20% of years and then used Nonmetric Multidimensional Scaling (NMDS) to identify quadrats with a similar composition of these candidate species.
-
-## Analytical approach
-While the observational data allows us to directly calculate species synchrony within the community, it is impossible to isolate the contributions of demographic and environmental stochasticity on species synchrony based on the data alone.
-So we now turn to a empirically-based modeling approach where we can turn species' interactions and demographic and environmental stochasticity on/off. 
-We use two models: (1) an environmentally stochastic Integral Projection Model (IPM) and (2) a demographically and environmentally stochastic Individual Based Model (IBM).
-Both models are driven by the same vital rate regressions for survival, growth, and recreuitment.
-We use the IPM to explore the drivers of synchrony assuming a large population where demographic stochasticity is weak or absent.
-We use the IBM to isolate the effect of demographic stochasticity and to examine how demographic and environmental stochasticity interact as population size increases (i.e., as the effect of demographic stochasticity weakens).
-The use of two model types (IPM and IBM) is the only way to isolate the effects of environmental stochasticity (an IPM with no demographic stochasticity) and demographic stochasticity (an IBM with no environmental stochasticity).
-
-## Statistical models for survival, growth, and recruitment
-### Survival and growth
+Statistical models for survival, growth, and recruitment
+--------------------------------------------------------
+## Survival and growth
 We modeled survival probability and growth on individual genets as a function of genet size, the crowding experienced by the focal genet from both heterospecific and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
@@
 
-### Recruitment
+## Recruitment
 We model recruitment as a function of plot-level cover, temporal varation among years, and spatial variation among quadrat groups.
@@
 [^stoch]: This is the crux of the matter here, we need to implement the Poisson approximation at the iteration matrix stage.
-
-Results
--------
-## Observed and expected synchrony
-Our first task was to calculate the observed and expected community synchrony at each of our study sites.
-Observed synchrony ($\phi_r$) is calculate using equation 1 and expected synchrony ($\text{E}(\phi_r)$) is simply $1/S$ or equation 3.
-Expected synchrony under indpendent fluctuations is lower than the observed synchrony at every site (Table 1).
-This implies that demographic stochasticity is weak relative to community regulation and environmental forcing ($\sigma^{2}_{c}+\sigma^{2}_{e} >> \sigma^{2}_{d}/N_i$).
-Thus, some combination of density-dependence (species interactions) and environmental forcing are synchronizing population dynamics.
-To isolate those effects, we turn to our simulation results from the IPMs.
-
-## Effect of removing demographic and environmental stochasticity
-At four of our five study areas, the presence of environmental stochasticity via random year effects on vital rate regressions increased community synchrony of per capita population growth rates (Figure 1).
-Only at the Montana site (Northern Mixed Prairie) did incorporating species' responses to environmental conditions decrease community synchrony.
-However, even with similar responses to environmental conditions, community synchrony did not reach 1 at any of the sites.
-Thus, there remains sufficiently strong temporal niche differences to drive the community away from perfect sychrony.
-
-Theory predicts that demographic stochasticty is the only force that, in isolation, can desynchronize population dynamics.
-However, theory also predicts that demographic stochasticity is likely unable to counteract the synchronizing effect of environmental forces.
-Our results support this prediction in the cases where environmental stochasticity is a synchronizing force (compare ENV and ENVDEMO bars in Figure 1).
-
-\newpage{}
-
-Tables
-------
-
-\singlespacing
-
-Table 1: Site descriptions and focal species.
-\footnotesize
-
-| Site Name | Biome | Location (Lat, Lon) | Obs. Years | Species |
-| --------- | ----- | ------------------- | ---------- | ------- |
-| Arizona | Sonoran Desert | 31&deg;50' N, 110&deg;53' W | 1915-1933 | *Bouteloua eriopoda* |
-| | | | | *Bouteloua rothrockii* |
-| Idaho | Sagebrush steppe | 44.2&deg; N, 112.1&deg; W | 1926-1957 | *Artemisia tripartita* |
-| | | | | *Pseudoroegneria spicata* |
-| | | | | *Hesperostipa comata* |
-| | | | | *Poa secunda* |
-| Kansas | Southern mixed prairie | 38.8&deg; N, 99.3&deg; W | 1932-1972 | *Bouteloua curtipendula* |
-| | | | | *Bouteloua hirsuta* |
-| | | | | *Schizachyrium scoparium* |
-| Montana | Northern mixed prairie | 46&deg;19' N, 105&deg;48' W | 1926-1957 | *Bouteloua gracilis* |
-| | | | | *Hesperostipa comata* |
-| | | | | *Pascopyrum smithii* |
-| | | | | *Poa secunda* |
-| New Mexico | Chihuahuan Desert | 32.62&deg; N, 106.67&deg; W | 1915-1950 | *Bouteloua eriopoda* |
-| | | | | *Sporobolus flexuosus* |
-
-```{r load_libs, echo=FALSE, message=FALSE, warning=FALSE}
-library(communitySynchrony)
-library(plyr)
-library(reshape2)
-library(synchrony)
-library(ggplot2)
-library(xtable)
-library(gridExtra)
-library(RColorBrewer)
-```
-
-```{r get_all_data, echo=FALSE}
-site <- "Kansas"
-spp_list <- c("BOCU","BOHI","SCSC")
-num_spp <- length(spp_list)
-ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  ks_data <- rbind(ks_data, spp_data)
-} #end species looping for raw data
-ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
-
-tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
-tmp2<-which(ks_data$quad_data=="q27")
-tmp3<-which(ks_data$quad=="q28")
-tmp4<-which(ks_data$quad=="q30")
-tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
-tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
-tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
-ks_data<-ks_data[-tmp,]
-
-# exclude the records later than 1968, to keep the same random year effect...
-ks_data<-subset(ks_data,year<68)
-
-site <- "Idaho"
-spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
-num_spp <- length(spp_list)
-id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  id_data <- rbind(id_data, spp_data)
-} #end species looping for raw data
-id_data <- id_data[2:nrow(id_data),] #remove first NA row
-id_data <- subset(id_data, species!="ARTR") #take out the shrub
-
-site <- "Montana"
-spp_list <- sort(c("BOGR","HECO","PASM","POSE"))
-num_spp <- length(spp_list)
-mt_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  mt_data <- rbind(mt_data, spp_data)
-} #end species looping for raw data
-mt_data <- mt_data[2:nrow(mt_data),] #remove first NA row
-
-site <- "NewMexico"
-spp_list <- sort(c("BOER","SPFL"))
-num_spp <- length(spp_list)
-nm_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  nm_data <- rbind(nm_data, spp_data)
-} #end species looping for raw data
-nm_data <- nm_data[2:nrow(nm_data),] #remove first NA row
-
-site <- "Arizona"
-spp_list <- sort(c("BOER","BORO"))
-num_spp <- length(spp_list)
-az_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
-for(dospp in 1:num_spp){ #loop through species to read in data
-  spp_now <- spp_list[dospp]
-  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
-  spp_data <- read.csv(quad_file)
-  spp_data$species <- spp_now
-  az_data <- rbind(az_data, spp_data)
-} #end species looping for raw data
-az_data <- az_data[2:nrow(az_data),] #remove first NA row
-```
-
-```{r get_all_synch, echo=FALSE, include=TRUE}
-out_ks <- get_comm_synchrony(ts_data = ks_data)
-out_id <- get_comm_synchrony(ts_data = id_data)
-out_mt <- get_comm_synchrony(ts_data = mt_data)
-out_nm <- get_comm_synchrony(ts_data = nm_data)
-out_az <- get_comm_synchrony(ts_data = az_data)
-```
-
-```{r synchrony_table, echo=FALSE, include=TRUE, results='asis', message=FALSE}
-get_table_metrics <- function(output){
-  tmp <- output
-  exp_tmp <- round(tmp$pgr_expected_synch_ind_flucts, 2)
-  obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
-  demo_tmp <- exp_tmp - obs_tmp
-  return(c(obs_tmp, exp_tmp, demo_tmp))
-}
-
-table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
-                                  get_table_metrics(out_id),
-                                  get_table_metrics(out_mt),
-                                  get_table_metrics(out_nm),
-                                  get_table_metrics(out_az)))
-table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
-table_data <- table_data[,c("Site", "V1", "V2", "V3")]
-table_data$spp <- c(length(unique(ks_data$species)),
-                    length(unique(id_data$species)),
-                    length(unique(mt_data$species)),
-                    length(unique(nm_data$species)),
-                    length(unique(az_data$species)))
-colnames(table_data) <- c("Site",
-                          "$\\phi_{r}$", 
-                          "$\\text{E}(\\phi_{r})$", 
-                          "$\\text{E}(\\sigma_{d})$",
-                          "Species richness")
-
-synch_cap <- "Observed and expected synchrony among species' per capita growth rates and expected effect of demographic stochasticity on observed synchrony.
-Species richness values reflect the number of species analyzed from the community, not the actual richness."
-print(xtable(table_data, caption = synch_cap),
-      caption.placement="top",
-      include.rownames = F, 
-      sanitize.colnames.function = identity,
-      comment=FALSE)
-```
-
-```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site."}
-# out_ks$percent_cover$site <- "Kansas"
-# out_id$percent_cover$site <- "Idaho"
-# out_mt$percent_cover$site <- "Montana"
-# out_nm$percent_cover$site <- "New Mexico"
-# out_az$percent_cover$site <- "Arizona"
-# cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
-#                     out_mt$percent_cover, out_nm$percent_cover, 
-#                     out_az$percent_cover)
-# 
-# out_ks$growth_rates$site <- "Kansas"
-# out_id$growth_rates$site <- "Idaho"
-# out_mt$growth_rates$site <- "Montana"
-# out_nm$growth_rates$site <- "New Mexico"
-# out_az$growth_rates$site <- "Arizona"
-# pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
-#                   out_mt$growth_rates, out_nm$growth_rates, 
-#                   out_az$growth_rates)
-# all_spp <- unique(pgr_data$species)
-# color_match <- data.frame(species = all_spp,
-#                           color = brewer.pal(11,"Paired"))
-# pgr_data <- merge(pgr_data, color_match, by = "species")
-# cover_data <- merge(cover_data, color_match, by="species")
-# 
-# cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
-#                 geom_line()+
-#                 geom_point()+
-#                 facet_grid(site~., scales = "free_y")+
-#                 guides(color=FALSE)+
-#                 ylab("Mean Percent Cover")+
-#                 xlab("Year")+
-#                 scale_color_manual(values = brewer.pal(11,"Paired"))
-# 
-# pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
-#               geom_line()+
-#               geom_point()+
-#               facet_grid(site~., scales = "free_y")+
-#               guides(color=FALSE)+
-#               ylab("Per Capita Growth Rate")+
-#               xlab("Year")+
-#               scale_color_manual(values = brewer.pal(11,"Paired"))
-# full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
-```
-
-\newpage{}
-
-Figures
--------
-```{r read_data, echo=FALSE, eval=F}
-output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
-mlist <- melt(output_list)
-```
-
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and IPM simulated (solid density curves) synchrony of species per capita growth rates based on changes in percent cover.", fig.width=6, fig.height=3, eval=F}
-nyrs <- c(max(out_az$stability$num_observations),
-          max(out_id$stability$num_observations),
-          max(out_ks$stability$num_observations),
-          max(out_mt$stability$num_observations),
-          max(out_nm$stability$num_observations))
-
-boots <- 100
-sites <- sort(table_data$Site)
-sites[5] <- "NewMexico"
-synch <- matrix(NA, boots, length(sites))
-
-for(do_site in 1:length(sites)){
-  full_ipm_outs <- subset(mlist, L2=="ENVINTER")
-  site_ipm <- subset(full_ipm_outs, L1==sites[do_site])
-  if(sites[do_site]=="Idaho") site_ipm <- subset(site_ipm, Var2!="ARTR")
-  species_list <- unique(site_ipm$Var2)
-  for(i in 1:boots){
-    begin_year <- sample(x = 1:(max(site_ipm$Var1)-nyrs[do_site]), 1)
-    end_year <- begin_year+nyrs[1]
-    tmp <- subset(site_ipm, Var1 %in% begin_year:end_year)
-    tmp2 <- dcast(tmp, Var1~Var2, value.var = "value")
-
-    # caclulate observed growth rates
-    # create lagged data frame to only get observed yearly transitions
-    lag_df <- tmp2
-    lag_df$lagyear <- lag_df$Var1+1
-    colnames(lag_df)[2:(num_spp+1)] <- paste0(colnames(lag_df)[2:(num_spp+1)],"_t0")
-
-    # merge the lag df with observed
-    rm_col <- which(colnames(lag_df)=="Var1")
-    merged_df <- merge(tmp2, lag_df[,-rm_col], by.x = "Var1", by.y="lagyear")
-    transitions <- nrow(merged_df)
-    obs_gr <- matrix(nrow=transitions, ncol=num_spp)
-    name_ids1 <- which(colnames(merged_df) %in% species_list)
-    name_ids2 <- which(colnames(merged_df) %in% paste(species_list, "_t0", sep=""))
-    for(ii in 1:transitions){
-      obs_gr[ii,] <- as.numeric(log(merged_df[ii,name_ids1]/merged_df[ii,name_ids2]))
-    }
-    synch[i,do_site] <- as.numeric(community.sync(obs_gr)[1])
-  }
-}
-
-colnames(synch) <- sites
-synch_df <- melt(synch)
-
-obs_pgr_df <- c(as.numeric(out_az$pgr_synchrony[1]),
-                as.numeric(out_id$pgr_synchrony[1]),
-                as.numeric(out_ks$pgr_synchrony[1]),
-                as.numeric(out_mt$pgr_synchrony[1]),
-                as.numeric(out_nm$pgr_synchrony[1]))
-obs_pgr_df <- data.frame(pgr_synch=obs_pgr_df,
-                         Var2=sites)
-
-ggplot()+
-  geom_line(data=synch_df, aes(x=value, color=Var2), stat="density", adjust=1)+
-  geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
-  facet_wrap("Var2")+
-  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
-  guides(color=FALSE)+
-  theme_bw()
-```
-
-```{r figure_2, echo=FALSE, fig.cap="Results from IPM simulations with difference sources of stochasticity present/absent at each site.
-DEMO = demographic stochasticity only; ENV = environmental stochasticity only; ENVDEMO = both demograhpic and environmental stochasticity.", fig.width=6, eval=F}
-colnames(mlist)[1:3] <- c("year", "species", "cover")
-sites <- unique(mlist$L1)
-sims <- unique(mlist$L2)
-pgr_list <- list()
-for(dosite in sites){
-  tmp_data <- subset(mlist, L1==dosite)
-  if(dosite=="Idaho"){
-    tmp_data <- subset(tmp_data, species!="ARTR")
-  }
-  within_site_list <- list()
-  for(dosim in sims){
-    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
-    tmpsynch <- get_ipm_synchrony(tmpsim)
-    tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
-    within_site_list[[dosim]] <- tmp_pgr_synch
-  }
-  pgr_list[[dosite]] <- within_site_list
-}
-pgr_synch <- melt(pgr_list)
-expected_synch <- table_data[,c(1,3)]
-expected_synch$L2 <- "E(SYNCH)"
-colnames(expected_synch) <- c("L1", "value", "L2")
-expected_synch <- expected_synch[,c("value", "L2", "L1")]
-expected_synch[which(expected_synch[,"L1"]=="New Mexico"),"L1"] <- "NewMexico"
-
-pgr_synch <- rbind(pgr_synch, expected_synch)
-
-# ggplot(pgr_synch, aes(x=L2, y=value, fill=L2))+
-#   geom_bar(stat="identity", position=position_dodge(0.9), color="grey")+
-#   facet_wrap("L1", scales="free_x")+
-#   scale_fill_manual(values = c("grey45", "steelblue", "slateblue4"))+
-#   guides(fill=FALSE)+
-#   scale_y_continuous(limits=c(0,1))+
-#   xlab("Source of stochasticity")+
-#   ylab("Species synchrony")
-
-pgrsynch_plot <- ggplot(pgr_synch, aes(x=L2, y=value, color=L1, group=L1))+
-  geom_line()+
-  geom_point(size=5)+
-  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
-                     name = "Site")+
-  scale_y_continuous(limits=c(0,1))+
-  xlab("Simulation")+
-  ylab("Species synchrony")+
-  theme_bw()
-
-print(pgrsynch_plot)
-
-####
-####  Compare observed and IPM simulated synchrony
-####
-nsites <- length(sites)
-
-```
-
-\newpage
-
-Additional Figures
-------------------
-```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations.", eval=F}
-ggplot(mlist, aes(x=year, y=(cover*100), color=species))+
-  geom_line()+
-  facet_grid(L1~L2, scale="free")
-```

### 2016-03-11 docs/cache/appendix_mathsupport.Rmd
adds online supporting information document

--- 2016-02-10 docs/appendix_mathsupport.Rmd
@@
 
+Equations A# and A# show that synchrony population biomasses and growth rates emerge from complex interactions among species' intrinsic growth rates, interspecific interactions, environmental stochasticity, and demographic stochasticity.
+Given these complexities, it is impossible to determine expected effects of different parameters in a multi-species case.
+Thus, we analyze a simplified case where interspecific interactions are zero.
+The variance-covariance matrix of population biomasses at steady state then becomes
+
+\begin{equation}
+c_{ij} = \frac{B_{ij}}{1-a_i a_j} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
+\end{equation}
+
+and the variance-covariance matrix of population growth rates becomes
+
+\begin{equation}
+R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} C_{ij} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}},
+\end{equation}
+
+which simplifies to
+
+\begin{equation}
+R_{ij} = \frac{1}{1- \frac{r_{mi}r_{mj}}{r_{mi}+r_{mj}}} \left( \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}} \right)
+\end{equation}
+
+\noindent This is still not a general result, however, because synchrony in population sizes and synchrony in growth rates use weighted factors of the environmental variances and covariances with species-specific parameters.
+So, we further assume that, along with interspecific interactions being zero, all species have identical growth rates, environmental stochasticity is absent, and all species have identical demographic variance.
+This represents a theoretical limiting case where the community consists of identical species coexisting in a constant environment where only demographic stochasticity causes temporal fluctuations.
+Under such conditions, synchrony of population biomasses is
+
+\begin{equation}
+\phi_N = \frac{1}{\left(\sum_i p_i^{1/2} \right)^2}
+\end{equation}
+
+\noindent and synchrony of growth rates is
+
+\begin{equation}
+\phi_N = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
+\end{equation}
+
+\noindent where $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
+When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a].
+
 References

### 2016-03-11 docs/ms_communitySynchrony.Rmd
adds online supporting information document

@@
 
-To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION...
+To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION... (see Online Supporting Information).
 
@@
 To estimate environmental responses in natural communities, we turn to our population models built using long term demographic data.
-
-Qualitative predictions here...?
 
@@
 Vital rate regressions are the building blocks of our dynamic models: an integral projection model (IPM) and an individual-based model (IBM).
-We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment.
+We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment (see Online Supporting Information for full details).
 We modeled survival probability of each genet as function of genet size, temporal variation among years, permanent spatial variation among groups of quadrats, and local neighborhood crowding from conspecific and heterospecific genets.
@@
 The IPM uses the same vital rate regressions as the IBM [@Rees2009; @Rees2014], but it is spatially-implicit and does not include demographic stochasticity.
-Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step.
+Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step (Supporting Online Information).
 Temporal variation is included in exactly the same way as for the IBM.
@@
 Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
-For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. X1).
-IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. X1), likely due to the desynchronizing effect of demographic stochasticity.
+For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
+IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to the desynchronizing effect of demographic stochasticity.
 
@@
 Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
-Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase).
-Species interactions are weak in these communities [Table SX and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
-Removing competition caused, at most, a 5% change in synchrony (Table SX).
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S2).
+Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
+Removing competition caused, at most, a 5% change in synchrony (Table S2).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
-2), with an average 6% increase [IS THIS FOR THE SMALLEST SIMULATED PLOT SIZE?].
+2), with an average 6% increase over synchrony from IBM simulations on a five meter area.
 The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
@@
 For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
-Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more variable, as expected by theory [@Loreau2008a].
+Results for synchrony of percent cover are qualitatively similar (Figs. S2-S3) but more variable, as expected by theory [@Loreau2008a].
 
@@
 The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem function experiments [@DeMazancourt2013a].
-The accumulating evidence from both experimental and observational studies demonstrates the utility of linking population dynamics to community and ecosystem level properties [@Loreau2010].
-[THIS LAST SENTENCE SUCKS]
 
@@
 tripartita* had not yet gone extinct, it was at much lower abundance than in the other simulation runs.
-When we removed the shrub (*Artemisia tripartita*) from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
-SX).
+When we removed *A.
+tripartita* from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
+S4).
 Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.
-*A. tripartita* has a facilitative effect on each grass species across all vital rates, except for a small competitive effect on *H. comata*'s survival probability (Tables SX, SX, SX).
-At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables SX, SX, SX).
+*A. tripartita* has a facilitative effect on each grass species across all vital rates, except for a small competitive effect on *H. comata*'s survival probability (Tables S4, S9, S14).
+At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables S4, S9, S14).
 We know synchrony is affected by interspecific competition [@Loreau2008a], but how facilitative effects manifest themselves is unknown.
@@
 Our results do not contradict the idea that competition can lead to compensatory dynamics, but they do highlight the fact that interspecific competition must be relatively strong to influence species synchrony.
-The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables X1-X10).
+The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables S1, S3-S17).
 Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap increases.

### 2016-03-11 docs/online_supporting_information.Rmd
adds online supporting information document

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
   - \doublespacing
   - \usepackage[document]{ragged2e}
   - \usepackage{color}
   - \usepackage{rotating}
bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "ms_communitySynchrony"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

\textbf{ONLINE SUPPORTING INFORMATION}

\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\tableofcontents
\listoffigures
\listoftables
\newpage{}

Derivation of limiting case predictions for synchrony of per capita growth rates (Equations 3-4)
--------------------------------
Following @Loreau2013 and @DeMazancourt2013a, we define population growth, ignoring observation error, as

\begin{align}
\tilde{r_i}(t) = r_{mi} \left[ 1- \frac{\tilde{N_i}(t)+\sum_{j \neq i} \alpha_{ij}\tilde{N_j}(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{\tilde{N_i}(t)}} \right]
\end{align}

\noindent where $N_i(t)$ is the biomass of species *i* in year *t*, and $r_i(t)$ is its instantaneous growth rate in year *t*. 
$r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*. 
Environmental stochasticity is incorporated as $\sigma_{ei}\mu_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $\mu_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated.
Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first order, normal approximation [@Lande2003; @DeMazancourt2013a].
$\sigma_{di}^2$ is the demographic variance and $\mu_{di}(t)$ are independent normal variables with zero mean and unit variance.

In the general case, **C** is the variance-covariance matrix for population biomasses at steady state.
It solves the Lyapunov equation $\textbf{C}^\infty = \textbf{A} \textbf{C}^\infty \textbf{A}^T + \textbf{B}$, where

\begin{align}
A_{ij} = 
  \begin{cases}
      1-r_{mi} \frac{N_{i}^*}{K_i}, & i=j \\
      -r_{mi} \frac{N_{i}^*}{K_i} \alpha_{ij}, & i \neq j
    \end{cases}
\end{align}

\noindent and

\begin{equation}
B_{ij} = N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})
\end{equation}

\noindent Similarly, **R** is the variance-covariance matrix for population growth rates at steady state

\begin{equation}
R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} \sum_{k,l} \alpha_{ik} \alpha_{jl} C_{kl} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}}.
\end{equation}

\noindent Then, following the synchrony metric of @Loreau2008a, the sychrony of population biomasses is

\begin{equation}
\phi_N = \frac{\sum_{i,j}C_{ij}}{\left( \sum_i \sqrt{C_{ii}} \right)^2}
\end{equation}

\noindent and the synchrony of per capita growth rates is

\begin{equation}
\phi_R = \frac{\sum_{i,j}R_{ij}}{\left( \sum_i \sqrt{R_{ii}} \right)^2}.
\end{equation}

Synchrony of population biomasses and growth rates emerge from complex interactions among species' intrinsic growth rates, interspecific interactions, environmental stochasticity, and demographic stochasticity.
Given these complexities, it is impossible to determine expected effects of different parameters in a multi-species case.
Thus, we analyze a simplified case where interspecific interactions are zero.
The variance-covariance matrix of population biomasses at steady state then becomes

\begin{equation}
c_{ij} = \frac{B_{ij}}{1-a_i a_j} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
\end{equation}

and the variance-covariance matrix of population growth rates becomes

\begin{equation}
R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} C_{ij} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}},
\end{equation}

which simplifies to

\begin{equation}
R_{ij} = \frac{1}{1- \frac{r_{mi}r_{mj}}{r_{mi}+r_{mj}}} \left( \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}} \right)
\end{equation}

\noindent This is still not a general result, however, because synchrony in population sizes and synchrony in growth rates use weighted factors of the environmental variances and covariances with species-specific parameters.
So, we further assume that, along with interspecific interactions being zero, all species have identical growth rates, environmental stochasticity is absent, and all species have identical demographic variance.
This represents a theoretical limiting case where the community consists of identical species coexisting in a constant environment where only demographic stochasticity causes temporal fluctuations.
Under such conditions, synchrony of population biomasses is

\begin{equation}
\phi_N = \frac{1}{\left(\sum_i p_i^{1/2} \right)^2}
\end{equation}

\noindent and synchrony of growth rates is

\begin{equation}
\phi_R = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
\end{equation}

\noindent where $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a]. The prediction represented in Equation S11 is Equation 3 in the main text, which we refer to as $\phi_{R,\mathcal{M_D}}$.

Another limiting case is where only environmental stochasticity is operating: no interspecific interactions, no demographic stochasticity, identical intrinsic growth rates, and environmental stochasticity with same standard deviation for all species.
Under such constraints, the synchrony of of species' biomasses is

\begin{equation}
\phi_N = \sum_{i,j} p_i p_j \text{cov}(u_{ei}, u_{ej})
\end{equation}

\noindent{} and the synchrony of species' growth rates is

\begin{equation}
\phi_R = \frac{\sum_{i,j} p_i p_j \text{cov}(u_{ei}, u_{ej})}{S^2}
\end{equation}

\noindent The prediction represented in Equation S13 is Equation 4 in the main text, which we refer to as $\phi_{R,\mathcal{M_E}}$. If we know the covariance matrix of the species environmental responses, we can calculate the above expectations directly. Note that the environmental responses are normalized in the above equations, therefore the covariances are correlations in the above equations.

\newpage{}

Materials and methods details
-----------------------------
### Vital rate statistical models
We modeled survival probability and growth on individual genets as a function of genet size, the crowding experienced by the focal genet from both heterospecific and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups. Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure).

We follow the approach of @Chu2015 to estimate crowding, assuming that the crowding experienced by a focal genet depends on distance to each neighbor genet and the neighbor's size, _u_:

\begin{equation}
w_{ijm,t} = \sum_k e^{-\delta_{jm}d_{ijkm,t}^{2}}u_{km,t}.
\end{equation}

In the above, $w_{ijm,t}$ is the crowding that genet _i_ of species _j_ in year _t_ experiences from neighbors of species _m_. The spatial scale over which species _m_ neighbors exert influence on any genet of species _j_ is determined by $\delta_{jm}$. The function is applied for all _k_ genets of species _m_ that neighbor the focal genet at time _t_, and $d_{ijkm,t}$ is the distance between genet _i_ in species _j_ and genet _k_ in species _m_. When $k=m$, the effect is intraspecific crowding. We use regression-specific (survival and growth) $\delta$ values estimated by Chu and Adler (2015).

We used logistic regression to model survival probability ($S$) of genet $i$ from species $j$ in quadrat group $g$ from time $t$ to $t+1$:

\begin{align}
\text{logit}(S_{ijg,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jg} + \beta^{S}_{j,t}x_{ij,t} + \boldsymbol{\omega}^{S}_{j} \textbf{w}_{ij,t}
\end{align}

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope parameter for size, $\phi^{S}_{jg}$ is the random effect of quadrat group location, and $\omega$ is a vector of per capita interaction coefficients which determine the impact of crowding, \textbf{w}, by each species on the focal species.

We modeled genet growth, conditional on survival, in a similar manner:

\begin{align}
\mu_{ijg,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jg} + \beta^{G}_{j,t}\mu_{ij,t} + \boldsymbol{\omega}^{G}_{j} \textbf{w}_{ij,t}
\end{align}

where $\mu$ is log genet size and all other parameters are as described for the survival regression. We capture non-constant error variance in growth by modeling the variance around the growth regression ($\varepsilon$) as a nonlinear function of predicted genet size:

\begin{align}
\varepsilon_{ij,t} = a e^{b \mu_{ijg,t+1}}
\end{align}

Our data allows us to track new recruits, but we cannot assign a specific parent to new genets. THerefore, we model recruitment at the quadrat level: the number of new individuals of species $j$ in quadrat $q$ recruiting at time $t+1$ as a function of quadrat "effective cover" ($A'$) in the previous year ($t$). Effective cover is a mixture of observed cover ($A$) in the focal quadrat ($q$) and the mean cover across the entire group ($\bar{A}$) of $Q$ quadrats in which $q$ is located:

\begin{equation}
A'_{jq,t} = p_{j}A_{jq,t} + (1-p_{j})\bar{A}_{jQ,t}
\end{equation}

where $p$ is a mixing fraction between 0 and 1 that is estimated within the model.

We assume the number of individuals, $y^{R}$, recruiting at time $t+1$ follows a negative binomial distribution:

\begin{equation}
y^{R}_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

where $\lambda$ is the mean intensity and $\zeta$ is the size parameter. We define $\lambda$ as:

\begin{equation}
\lambda_{jq,t+1} = A'_{jq,t}e^{(\gamma^{R}_{j,t} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} + \omega^{R}\sqrt{A'_{q,t}})}
\end{equation}

where $A'$ is effective cover ($\text{cm}^2$) of species $j$ in quadrat $q$ and all other terms are as in the survival and growth regressions.

### Multi-species populations models
The individually based model (IBM) is straighforward and described in the main text, so here focus on the structure of the integral projeciton model (IPM).
We built an environmentally stochastic IPM. Our IPM follows the specification of @Chu2015 where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_. Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. So, the density function for any size _v_ at time $t+1$ is

\begin{equation}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
\end{equation}

where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal that describes all possible transitions from size $u$ to $v$ and $\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding experienced from all other species by a genet of size $u_j$ and species $j$. The integral is evaluated over all possible sizes between predefined lower (_L_) and upper (_U_) size limits that extend beyond the range of observed genet sizes.

The population kernal is defined as the joint contributions of survival (_S_), growth (_G_), and recruitment (_R_):

\begin{equation}
k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
\end{equation}

\noindent{} which means we are calculating growth (_G_) for individuals that survive (_S_) from time _t_ to _t+1_ and adding in newly recruited (_R_) individuals of an average sized one-year-old genet for the focal species. Our stastical model for recruitment (_R_, described below) returns the number of new recruit produced per quadrat. Following previous work, we assume that fecundity increases linearly with size ($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$) to incorporate the recruitment function in the spatially-implicit IPM.

\newpage{}

Figures
-------
\begin{figure}[!ht]
  \centering
      \includegraphics[width=4in]{./components/figureS1.pdf}
  \caption{Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IPM (left panels) and the IBM (right panels). IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set. IBM density curves come from 100 replicate simulations of 75 iterations each. Synchrony values come from simulations where environmental stochasticity and interspecific interactions are present. The IBM was run on a 5 by 5 meter landscape to reduce the effect of demographic stochasticity.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/figureS2_simscover.png}
  \caption{Community-wide species percent cover from model simulation experiments. Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp. + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp. + No E.S.''). Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation. Error bars represent the upper and lower 95\% quantiles from model simulations.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/figureS3_ibmcover.png}
  \caption{Synchrony of species' percent cover for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S. + No Comp.''). The strength of demographic stochasticity decreases as landscape size increases. Error bars represent the upper and lower 95\% quantiles from model simulations.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics{./components/all_sims_supp_noARTR.png}
  \caption{Community-wide species percent cover from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed. Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp. + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp. + No E.S.''). Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation. Error bars represent the upper and lower 95\% quantiles from model simulations.}
\end{figure}

\newpage{}

Tables
------
```{r avg_spp_interactions, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)
####
####  Get Vital Rate Satistical Results Files ----------------------------------
####
vital_rates <- c("growth", "surv", "recruit")
result_files <- list.files("../results/")
vr_files <-result_files[grep(paste(vital_rates,collapse="|"), 
                              list.files("../results/"))]

####
####  Loop Over Files and Extract Crowding Effects -----------------------------
####
# Growth and survival first
vital_mat_list <- list()
for(do_vital in vital_rates[c(1,2)]){
  vr_do <- vr_files[grep(do_vital,vr_files)]
  tmp_vital <- readRDS(paste0("../results/",vr_do))
  sites <- names(tmp_vital)
  mat_list <- list()
  for(do_site in sites){
    tmp <- tmp_vital[[do_site]]
    spp <- names(tmp)
    tmpmat <- matrix(0,nrow = length(spp), ncol = length(spp))
    for(i in 1:length(spp)){
      do_spp <- spp[i]
      tmpspp <- tmp[[do_spp]]
      tmpw <- tmpspp[1,grep("crowd*", colnames(tmpspp))]
      tmpmat[i,] <- as.numeric(tmpw)
    }
    rownames(tmpmat) <- spp
    colnames(tmpmat) <- spp
    mat_list[[do_site]] <- tmpmat
  }
  vital_mat_list[[do_vital]] <- mat_list 
}

# Recruitment
vr_do <- vr_files[grep("recruit",vr_files)]
tmp_vital <- readRDS(paste0("../results/",vr_do))
names_vital <- readRDS(paste0("../results/",vr_files[1]))
sites <- names(tmp_vital)
mat_list <- list()
for(do_site in sites){
  tmp <- tmp_vital[[do_site]]
  spp <- names(names_vital[[do_site]])
  tmpdd <- tmp[grep("dd", rownames(tmp)),"Mean"]
  tmpmat <- matrix(tmpdd,length(spp),length(spp), byrow = FALSE)
  rownames(tmpmat) <- spp
  colnames(tmpmat) <- spp
  mat_list[[do_site]] <- tmpmat
}
vital_mat_list[["recruit"]] <- mat_list

####
####  Calculate Average Inter and Intraspecific Interactions -------------------
####
vital_comp_list <- list()
for(do_vital in vital_rates){
  tmp <- vital_mat_list[[do_vital]]
  sites <- names(tmp)
  comp_list <- list()
  for(do_site in sites){
    tmpsite <- tmp[[do_site]]
    avgintra <- mean(diag(tmpsite))
    avginter <- mean(c(tmpsite[upper.tri(tmpsite)], tmpsite[lower.tri(tmpsite)]))
    avgcomp <- matrix(c(avginter, avgintra),1,2)
    colnames(avgcomp) <- c("inter", "intra")
    comp_list[[do_site]] <- avgcomp
  }
  vital_comp_list[[do_vital]] <- comp_list
}

df <- melt(unlist(vital_comp_list))
pieces <- strsplit(rownames(df), split = "[.]")
df$vitalrate <- sapply(pieces, "[", 1)
sites <- sapply(pieces, "[", 2)
df$site <- substr(sites, start = 1, stop = nchar(sites)-1)
df$interintra <- substr(sites, start = nchar(sites), stop = nchar(sites))
avg_interintra <- ddply(df, .(site, interintra), summarise,
                        avg_val = mean(value))
avg_interintra[which(avg_interintra$interintra==1),"interintra"] <- "Interspecific Competition"
avg_interintra[which(avg_interintra$interintra==2),"interintra"] <- "Intraspecific Competition"
interintra_table <- dcast(avg_interintra, site~interintra, value.var = "avg_val")

comp_cap <- "Average interaction coefficients across vital rates for each community."
print(xtable(interintra_table, caption = comp_cap),
  caption.placement="top",
  include.rownames = FALSE, 
  sanitize.colnames.function = identity,
  comment=FALSE)
```

```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)

perc_changes <- read.csv("../results/synchsims_percent_diffs.csv")
perc_changes <- perc_changes[which(perc_changes$simulation!="1All Drivers"), c("site", "simulation", "percent_diff")]
table_diff <- dcast(perc_changes, simulation~site, value.var = "percent_diff")

diff_cap <- "Percent differences of synchrony of per capita growth rates between each removal simulation experiment and the 'All Drivers' simulation."
print(xtable(table_diff, caption = diff_cap),
  caption.placement="top",
  include.rownames = FALSE, 
  sanitize.colnames.function = identity,
  comment=FALSE)
```

```{r get_interactions, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE}
####
####  Get Vital Rate Satistical Results Files ----------------------------------
####
vital_rates <- c("growth", "surv", "recruit")
result_files <- list.files("../results/")
vr_files <-result_files[grep(paste(vital_rates,collapse="|"), 
                              list.files("../results/"))]

####
####  Loop Over Files and Extract Crowding Effects -----------------------------
####
# Growth and survival first
vital_mat_list <- list()
for(do_vital in vital_rates[c(1,2)]){
  vr_do <- vr_files[grep(do_vital,vr_files)]
  tmp_vital <- readRDS(paste0("../results/",vr_do))
  sites <- names(tmp_vital)
  mat_list <- list()
  for(do_site in sites){
    tmp <- tmp_vital[[do_site]]
    spp <- names(tmp)
    tmpmat <- matrix(0,nrow = length(spp), ncol = length(spp))
    for(i in 1:length(spp)){
      do_spp <- spp[i]
      tmpspp <- tmp[[do_spp]]
      tmpw <- tmpspp[1,grep("crowd*", colnames(tmpspp))]
      tmpmat[i,] <- as.numeric(tmpw)
    }
    rownames(tmpmat) <- spp
    colnames(tmpmat) <- spp
    mat_list[[do_site]] <- tmpmat
  }
  vital_mat_list[[do_vital]] <- mat_list 
}

# Recruitment
vr_do <- vr_files[grep("recruit",vr_files)]
tmp_vital <- readRDS(paste0("../results/",vr_do))
names_vital <- readRDS(paste0("../results/",vr_files[1]))
sites <- names(tmp_vital)
mat_list <- list()
for(do_site in sites){
  tmp <- tmp_vital[[do_site]]
  spp <- names(names_vital[[do_site]])
  tmpdd <- tmp[grep("dd", rownames(tmp)),"Mean"]
  tmpmat <- matrix(tmpdd,length(spp),length(spp), byrow = FALSE)
  rownames(tmpmat) <- spp
  colnames(tmpmat) <- spp
  mat_list[[do_site]] <- tmpmat
}
vital_mat_list[["recruit"]] <- mat_list 
```

```{r make_tables, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
names(vital_mat_list) <- c("growth", "survival", "recruitment")
for(do_vital in names(vital_mat_list)){
  tmp_list <- vital_mat_list[[do_vital]]
  for(do_site in names(tmp_list)){
    site_list <- tmp_list[[do_site]]
    synch_cap <- paste0("Interaction coefficients for ", do_vital, " regressions in ", do_site, ".")
    print(xtable(site_list, caption = synch_cap),
      caption.placement="top",
      include.rownames = TRUE, 
      sanitize.colnames.function = identity,
      comment=FALSE)
  }
}
```

\newpage{}

References
----------

### 2016-03-11 devel/make_ms_plots.R

@@ -44,11 +44,11 @@ synch_dftmp <- synch_df[2:nrow(synch_df),]
 synch_df <- melt(synch_dftmp, id.vars = c("site", "experiment", "bootnum"))
 colnames(synch_df) <- c("site", "experiment", "bootnum", "typesynch", "synch")
 
-ipm_synch <- ddply(synch_df, .(site, experiment, typesynch), summarise,
+ipm_synch_all <- ddply(synch_df, .(site, experiment, typesynch), summarise,
                    mean_synch = mean(synch),
                    up_synch = quantile(synch, 0.95),
                    lo_synch = quantile(synch, 0.05))
-ipm_synch <- subset(ipm_synch, typesynch=="pgr_synch")
+ipm_synch <- subset(ipm_synch_all, typesynch=="pgr_synch")
 
 ##  Read in IBM results -----
 ibm_synch_all <- readRDS("../results/ibm_sims_collated.RDS") 
@@ -148,3 +148,98 @@ ggsave("../docs/components/ibm_sims_across_landscape.png", width = 10, height = 
 
 
 
+##  Make supplement plots for percent cover ------------
+ipm_synch_all <- ddply(synch_df, .(site, experiment, typesynch), summarise,
+                       mean_synch = mean(synch),
+                       up_synch = quantile(synch, 0.95),
+                       lo_synch = quantile(synch, 0.05))
+ipm_synch <- subset(ipm_synch_all, typesynch=="abund_synch")
+
+##  Read in IBM results -----
+ibm_synch_all <- readRDS("../results/ibm_sims_collated.RDS") 
+ibm_synch_agg <- ddply(ibm_synch_all, .(experiment, site, expansion, typesynch), summarise,
+                       avg_synch = mean(synch),
+                       up_synch = quantile(synch, 0.95),
+                       lo_synch = quantile(synch, 0.05))
+ibm_synch <- subset(ibm_synch_agg, expansion==5 & typesynch=="Cover (%)")
+
+##  Combine results -----
+sim_names <- c("All Drivers", "No D.S.", "No E.S.", "No Comp.", "No Comp. + No D.S.", "No Comp. + No E.S.")
+sim_names_order <- paste0(1:length(sim_names),sim_names)
+site_names <- unique(ipm_synch$site)
+nsites <- length(site_names)
+site_labels <- site_names
+site_labels[which(site_labels=="NewMexico")] <- "New Mexico"
+
[…]

### 2016-03-11 docs/ms_communitySynchrony.Rmd
updates supp table order

@@
 Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
-Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S2).
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S1).
+Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S2).
 Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
@@
 In other words, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
-One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct repeat the analysis of @Gross2014 every few years.
+One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.
 
@@
 Such a conclusion is only half correct.
-The perennial grasses we studied do have similar responses to the environment (Table X1-X10), which will tend to synchronize dynamics.
+The perennial grasses we studied do have similar responses to the environment (Table S2), which will tend to synchronize dynamics.
 However, if community-wide species synchrony is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].

### 2016-03-11 docs/online_supporting_information.Rmd
updates supp table order

@@
 ------
+```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
+library(reshape2)
+library(plyr)
+library(ggplot2)
+library(xtable)
+
+perc_changes <- read.csv("../results/synchsims_percent_diffs.csv")
+perc_changes <- perc_changes[which(perc_changes$simulation!="1All Drivers"), c("site", "simulation", "percent_diff")]
+table_diff <- dcast(perc_changes, simulation~site, value.var = "percent_diff")
+
+diff_cap <- "Percent differences of synchrony of per capita growth rates between each removal simulation experiment and the 'All Drivers' simulation."
+print(xtable(table_diff, caption = diff_cap),
+  caption.placement="top",
+  include.rownames = FALSE, 
+  sanitize.colnames.function = identity,
+  comment=FALSE)
+```
+
+```{r corr_yeareffs, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
+library(reshape2)
+library(plyr)
+library(ggplot2)
+library(xtable)
+####
+####  Read in regression fits --------------------------------------------------
+####
+all_files <- list.files("../results/")
+param_files <- all_files[grep("param*", all_files)]
+num_vitals <- length(param_files)
+
+list_yr_effects <- list()
+for(j in 1:num_vitals){
+  param_list <- readRDS(paste0("../results/",param_files[j]))
+  vital_name <- unlist(strsplit(param_files[j], "_"))[1]
+
+  if(vital_name != "recruit"){
+
+    site_names <- names(param_list)
+    site_year_list <- list()
+    for(do_site in site_names){
+      tmp_site <- param_list[[do_site]]
+      num_spp <- length(tmp_site) 
+      tmp_yr_int <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
+      tmp_yr_slope <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
+      for(i in 1:num_spp){
+        spp_df <- tmp_site[[i]]
+        tmp_yr_int[,i] <- spp_df$Intercept.yr
+        col2use <- grep("logarea", colnames(spp_df))
+        col2use2 <- grep(".yr", colnames(spp_df[col2use]))
+        tmp_yr_slope[,i] <- spp_df[,col2use[col2use2]]
+      } # end species within site loop
+      site_year_list[[do_site]][["intercept"]] <- tmp_yr_int
+      site_year_list[[do_site]][["slope"]] <- tmp_yr_slope
+    } # end site within vital rate loop
+
+  } # end NOT recruit loops
+
+  if(vital_name == "recruit"){
+
+    site_names <- names(param_list)
+    site_year_list <- list()
+    for(do_site in site_names){
+      tmp_site <- param_list[[do_site]]
+      tmp_yrs <- tmp_site[grep("intcpt.yr", rownames(tmp_site)),"Mean"]
+      stringlist <- strsplit(names(tmp_yrs), ",")
+      num_spp <- length(unique(sapply(stringlist, "[[", 2)))
+      mat_yrs <- t(matrix(tmp_yrs, nrow=num_spp, byrow = TRUE))
+      site_year_list[[do_site]] <- mat_yrs
+    } # end site within vital rate loop
+
+  } # end YES recruit loops
+
+  list_yr_effects[[vital_name]] <- site_year_list
+} # end vital rate loop
+
+# ## Plot histograms of year effects by vital rate
+# vital_rates <- names(list_yr_effects)
+# par(mfrow=c(3,5))
+# for(do_vital in vital_rates){
+#   tmp_vital <- list_yr_effects[[do_vital]]
+#   site_names <- names(tmp_vital)
+#   for(do_site in site_names){
+#     tmp_site <- tmp_vital[[do_site]]
+#     hist(tmp_site, main=paste(do_site,do_vital), xlab="Year Effect", xlim=c(-3,8))
+#   }
+# }
+
+##  Get standard deviation of year effect posterior
+
+####
+####  Calculate average correlation of year effects ----------------------------
+####
+vital_rates <- names(list_yr_effects)
+out_df <- data.frame(site=NA, vital_rate=NA, term=NA, corr=NA)
+for(do_vital in vital_rates){
+  tmp_vital <- list_yr_effects[[do_vital]]
+  site_names <- names(tmp_vital)
+  for(do_site in site_names){
+    tmp_site <- tmp_vital[[do_site]]
+
+    if(do_vital != "recruit"){
+      for(i in 1:length(tmp_site)){
+        tmp_cor <- cor(tmp_site[[i]])
+        avg_cor <- mean(tmp_cor[upper.tri(tmp_cor)])
+        tmp_out <- data.frame(site=do_site, vital_rate=do_vital, term=names(tmp_site)[i], corr=avg_cor)
+        out_df <- rbind(out_df, tmp_out)
+      }
+    }
+
+    if(do_vital == "recruit"){
+      tmp_cor <- cor(tmp_site)
+      avg_cor <- mean(tmp_cor[upper.tri(tmp_cor)])
+      tmp_out <- data.frame(site=do_site, vital_rate=do_vital, term="intercept", corr=avg_cor)
+      out_df <- rbind(out_df, tmp_out)
+    }
+
+  }
+}
+cor_df <- out_df[2:nrow(out_df),]
+cor_cast <- dcast(cor_df, site+term~vital_rate)
+
+cor.cap <- "Correlations of species' year random effects for each site by term, where term refers to the random effect on the slope or the intercept."
+print(xtable(cor_cast, caption = cor.cap),
+  caption.placement="top",
+  include.rownames = FALSE, 
+  sanitize.colnames.function = identity,
+  comment=FALSE)
+```
+
 ```{r avg_spp_interactions, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
@@
 
-```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
-library(reshape2)
-library(plyr)
-library(ggplot2)
-library(xtable)
-
-perc_changes <- read.csv("../results/synchsims_percent_diffs.csv")
-perc_changes <- perc_changes[which(perc_changes$simulation!="1All Drivers"), c("site", "simulation", "percent_diff")]
-table_diff <- dcast(perc_changes, simulation~site, value.var = "percent_diff")
-
-diff_cap <- "Percent differences of synchrony of per capita growth rates between each removal simulation experiment and the 'All Drivers' simulation."
-print(xtable(table_diff, caption = diff_cap),
-  caption.placement="top",
-  include.rownames = FALSE, 
-  sanitize.colnames.function = identity,
-  comment=FALSE)
-```
-
 ```{r get_interactions, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE}

### 2016-03-11 devel/plot_yreffect_corrs.R

@@ -54,12 +54,17 @@ for(j in 1:num_vitals){
     for(do_site in site_names){
       tmp_site <- param_list[[do_site]]
       num_spp <- length(tmp_site) 
-      tmp_yr_mat <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
+      tmp_yr_int <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
+      tmp_yr_slope <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
       for(i in 1:num_spp){
         spp_df <- tmp_site[[i]]
-        tmp_yr_mat[,i] <- spp_df$Intercept.yr
+        tmp_yr_int[,i] <- spp_df$Intercept.yr
+        col2use <- grep("logarea", colnames(spp_df))
+        col2use2 <- grep(".yr", colnames(spp_df[col2use]))
+        tmp_yr_slope[,i] <- spp_df[,col2use[col2use2]]
       } # end species within site loop
-      site_year_list[[do_site]] <- tmp_yr_mat
+      site_year_list[[do_site]][["intercept"]] <- tmp_yr_int
+      site_year_list[[do_site]][["slope"]] <- tmp_yr_slope
     } # end site within vital rate loop
     
   } # end NOT recruit loops
@@ -83,17 +88,17 @@ for(j in 1:num_vitals){
 } # end vital rate loop
 
 
-## Plot histograms of year effects by vital rate
-vital_rates <- names(list_yr_effects)
-par(mfrow=c(3,5))
-for(do_vital in vital_rates){
-  tmp_vital <- list_yr_effects[[do_vital]]
-  site_names <- names(tmp_vital)
-  for(do_site in site_names){
-    tmp_site <- tmp_vital[[do_site]]
-    hist(tmp_site, main=paste(do_site,do_vital), xlab="Year Effect", xlim=c(-3,8))
-  }
-}
+# ## Plot histograms of year effects by vital rate
+# vital_rates <- names(list_yr_effects)
+# par(mfrow=c(3,5))
+# for(do_vital in vital_rates){
+#   tmp_vital <- list_yr_effects[[do_vital]]
+#   site_names <- names(tmp_vital)
+#   for(do_site in site_names){
+#     tmp_site <- tmp_vital[[do_site]]
[…]

### 2016-03-17 docs/ms_communitySynchrony.Rmd
updates figures re CDM comments

@@
 ---
+
+<!--    - \usepackage{times}   -->
+
 ```{r caching, include=FALSE}
@@
                                   get_table_metrics(out_nm)))
+table_data$Site <- c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico")
 alltable_data <- table_data
-table_data$Site <- c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico")
-table_data <- table_data[,c("Site", "V1", "V4", "V7")]
+colnames(table_data) <- c("obs_gr_synch", "demo_only_exp_grsynch",
+                          "diff_noneed", "obs_cover_synch",
+                          "demo_only_exp_coversynch", "diffcover_noneed",
+                          "cover_variance", "site")
+saveRDS(table_data, "../results/calculate_site_metrics.RDS")
+
+table_data <- alltable_data[,c("Site", "V1", "V4", "V7")]
 table_data$spp <- c(length(unique(az_data$species)),
@@
                          exp_env = round(exp_synchrony_env$growthrate_prediction,2),
-                         sim_env = round(subset(sims_output, simulation=="5No Comp. + No D.S.")["synchrony"],2))
+                         sim_env = round(subset(sims_output, simulation=="2No Comp. + No D.S.")["synchrony"],2))
 
@@
   \centering
-      \includegraphics[width=6in]{./components/formatted_fig2.png}
+      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure1.png}
   \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
@@
   \centering
-      \includegraphics[width=6in]{./components/ibm_sims_across_landscape.png}
+      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
   \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S.
@@
   \centering
-      \includegraphics[width=4in]{./components/poly_vs_mono_synch.png}
-  \caption{Synchrony of species per capita growth rates (filled circles) and percent cover (filled triangles) simulated in monoculture (without species interactions) versus synchrony in polyculture (with species interactions).
+      \includegraphics[width=4in]{./components/formatted_figures/formatted_figure3.png}
+  \caption{Synchrony of species per capita growth rates simulated in monoculture (without species interactions) versus synchrony in polyculture (with species interactions).
 We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem function experiments.

### 2016-03-17 devel/make_ms_plots.R

@@ -60,11 +60,12 @@ ibm_synch <- subset(ibm_synch_agg, expansion==5 & typesynch=="Per capita growth 
 
 ##  Combine results -----
 sim_names <- c("All Drivers", "No D.S.", "No E.S.", "No Comp.", "No Comp. + No D.S.", "No Comp. + No E.S.")
-sim_names_order <- paste0(1:length(sim_names),sim_names)
+sim_names_order <- paste0(c(3,1,5,4,2,6),sim_names)
 site_names <- unique(ipm_synch$site)
 nsites <- length(site_names)
 site_labels <- site_names
 site_labels[which(site_labels=="NewMexico")] <- "New Mexico"
+site_labels_order <- paste0(c(2,5,3,4,1), site_labels)
 
 # Get vectors of synchrony for each "experiment"
 control_synch <- subset(ibm_synch, experiment=="fluctinter")
@@ -92,23 +93,48 @@ all_downs <- c(control_synch$lo_synch,
                nodsnocomp_synch$lo_synch, 
                noesnocomp_synch$lo_synch)
 
-plot_df <- data.frame(site = rep(site_labels, times=length(sim_names)),
+plot_df <- data.frame(site = rep(site_labels_order, times=length(sim_names)),
                       simulation = rep(sim_names_order, each=nsites),
                       synchrony = all_experiments,
                       upper_synch = all_ups,
                       lower_synch = all_downs)
 
+
+site_labels <- site_labels[order(site_labels_order)]
+sim_labels <- sim_names[order(sim_names_order)]
+
+# Bring in theoretical predictions and observed values
+theory_env <- readRDS("../results/envstoch_predictions.RDS")
+theory_env <- theory_env[2:nrow(theory_env), c("site", "growthrate_prediction")]
+metrics <- readRDS("../results/calculate_site_metrics.RDS")
+theory_demo <- metrics[,c("demo_only_exp_grsynch", "obs_gr_synch", "site")]
+theory_demo[which(theory_demo$site=="New Mexico"),"site"] <- "NewMexico"
+theoretical_preds_and_obs <- merge(theory_env, theory_demo)
+colnames(theoretical_preds_and_obs) <- c("site", "envonly_pred", "demonly_pred", "obs")
[…]

### 2016-03-17 devel/plot_ipm_sims.R

@@ -102,7 +102,9 @@ polymono_wide <- dcast(polymono, site+typesynch~experiment, value.var = "mean_sy
 #   theme(legend.text = element_text(size = 8))+
 #   theme(legend.background = element_rect(colour = NA, fill = NA))
 
-g1 <- ggplot(subset(polymono_wide,typesynch=="pgr_synch"), aes(x=ENVNOINTER, y=ENVINTER))+
+plot_df <- subset(polymono_wide,typesynch=="pgr_synch")
+plot_df$site <- c("2Arizona", "5Idaho", "3Kansas", "4Montana", "1New Mexico")
+g1 <- ggplot(plot_df, aes(x=ENVNOINTER, y=ENVINTER))+
   geom_abline(aes(intercept=0, slope=1), linetype=3)+
   geom_point(size=3, aes(color=site))+
   scale_y_continuous(limits=c(0,1))+
@@ -112,10 +114,9 @@ g1 <- ggplot(subset(polymono_wide,typesynch=="pgr_synch"), aes(x=ENVNOINTER, y=E
   # scale_shape_discrete(name="Temporal Variable", labels=c("Per capita growth rate", "Percent cover"))+
   scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
                      name = "",
-                     labels = c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico"))+
+                     labels = c("New Mexico", "Arizona", "Kansas", "Montana", "Idaho"))+
   theme_few()+
   guides(shape=FALSE)+
-  # ggtitle("A                                                          ")+
   theme(legend.position=c(0.2,0.8))+
   theme(legend.text = element_text(size = 8))+
   theme(legend.background = element_rect(colour = NA, fill = NA))

### 2016-03-18 docs/ms_communitySynchrony.Rmd
incorporates ML edits

@@
    - \doublespacing
-   - \usepackage[document]{ragged2e}
    - \usepackage{color}
@@
 
-<!--    - \usepackage{times}   -->
+<!--    - \usepackage{times} 
+   - \usepackage[document]{ragged2e} -->
 
@@
 
-\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}
+\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Theoretical and Experimental Ecology Station, CNRS and Paul Sabatier University, Moulis, 09200, France}}
 
@@
 
-\textit{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, stability
+\textit{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
 
@@
 
-\textbf{Authorship}: ATT, CdM, and PBA designed the research, ATT performed the research, and ATT, CdM, ML, and PBA contributed to the analyses and interpretation of results.
-All authors contributed to writing the manuscript.
+\textbf{Authorship}: ATT, CdM, and PBA designed the research.
+CdM and ML designed and analyzed the theoretical model.
+ATT and PBA performed the empirical analysis.
+All authors contributed to the interpretation and presentation of results.
+ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
 
@@
 \begin{abstract}
-Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem function, but identifying the mechanisms that determine synchrony remains a challenge.
+Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
 Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific competition, and demographic stochasticity.
@@
 In all five communities, theoretical predictions closely matched simulated synchrony for the limiting cases where only demographic stochasticity or environmental variation are present, meaning that theory developed with simple models also applies to our empirical models.
-Simulations experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Simulation experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
 Removing interspecific competition and demographic stochasticity had small effects on synchrony.
@@
 Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
-Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem function [@Hautier2014].
+Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
 Predicting the impacts of global change requires a mechanistic understanding of species synchrony in natural settings.
@@
 We constrain our focus to synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
+However, we provide analytical predictions for synchrony of abundances in the Online Supporting Information.
 Following @Loreau2008a and @DeMazancourt2013a, we define population growth, ignoring observation error, as
@@
 r_i(t) &= \text{ln}N_i(t+1) - \text{ln}N_i(t) \\
-&= r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{N_i(t)}} \right]
+&= r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}u_{ei}(t) + \frac{\sigma_{di}u_{di}(t)}{\sqrt{N_i(t)}} \right]
 \end{align}
 
-\noindent where $N_i(t)$ is the biomass of species *i* in year *t*, and $r_i(t)$ is its instantaneous growth rate in year *t*. 
+\noindent where $N_i(t)$ is the biomass of species *i* in year *t*, and $r_i(t)$ is its population growth rate in year *t*. 
 $r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*.
-Environmental stochasticity is incorporated as $\sigma_{ei}\mu_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $\mu_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated.
-Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first order, normal approximation [@Lande2003; @DeMazancourt2013a].
-$\sigma_{di}^2$ is the demographic variance and $\mu_{di}(t)$ are independent normal variables with zero mean and unit variance.
+Environmental stochasticity is incorporated as $\sigma_{ei}u_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $u_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated between species.
+Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first-order, normal approximation [@Lande2003; @DeMazancourt2013a].
+$\sigma_{di}^2$ is the demographic variance and $u_{di}(t)$ are independent normal variables with zero mean and unit variance.
 
@@
 \begin{equation}
-\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(\mu_{ei},\mu_{ej})}{S^2}
+\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}p_ip_j\text{cov}(u_{ei},u_{ej})}{S^2}
 \end{equation}
@@
 \noindent{} for synchrony of per capita growth rates.
-Again, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$, and $\text{cov}(\mu_{ei},\mu_{ej})$ is the standardize covariance (correlation) of environmental responses between species *i* and species *j*.
+Again, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$, and $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
 Confronting our theoretical predictions with data requires estimates of species dynamics in monoculture to calculate the covariance of species' environmental responses.
-To estimate environmental responses in natural communities, we turn to our population models built using long term demographic data.
+To estimate environmental responses in natural communities, we turn to our population models built using long-term demographic data.
 
@@
 In the early 20th century, several range experiment stations in the western U.S. began mapping permanent quadrats, and conducted annual censuses for decades. 
-We use such long-term demographic data sets from five semiarid grasslands in the western United States (described in detail by @Chu2015).
+We use such long-term demographic data sets from five semiarid grasslands in the western United States [described in detail by @Chu2015].
 Each study site includes a set of 1-$\text{m}^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
@@
 $\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [@Loreau2008a].
+We use the same equation to calculate observed synchrony of species' percent cover, which we present to relate our results to previous findings, even though we focus on synchrony of growth rates in our model simulations (see below).
 
@@
 We modeled survival probability of each genet as function of genet size, temporal variation among years, permanent spatial variation among groups of quadrats, and local neighborhood crowding from conspecific and heterospecific genets.
-Regression coefficients for the effect of crowding by each species can be considered a matrix of interaction coefficients whose diagonals represent intraspecific interations and whose off-diagonals represent interspecific interactions [@Adler2010].
+Regression coefficients for the effect of crowding by each species can be considered a matrix of interaction coefficients whose diagonals represent intraspecific interactions and whose off-diagonals represent interspecific interactions [@Adler2010].
 These interaction coefficients can take positive (facilitative) or negative (competitive) values.
@@
 #### Building dynamic multi-species models
-Once we have fit the vital rate stastical models, building the population models is straightforward.
+Once we have fit the vital rate statistical models, building the population models is straightforward.
 For the IBM, we initialize the model by randomly assigning plants spatial coordinates, sizes and species identities until each species achieves a density representative of that observed in the data.
@@
 (2) IPM: Demographic stochasticity removed;
-(3) IBM: Environmental stochastcity removed;
+(3) IBM: Environmental stochasticity removed;
 (4) IBM: Interspecific competition removed;
@@
 First, in the absence of interspecific competition and demographic stochasticity, species are effectively growing in large monocultures.
-Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(\mu_{ie}, \mu_{je})$) and parameterize Equation 3.
+Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(u_{ie}, u_{je})$) and parameterize Equation 3.
 Parameterizing Equation 2 does not require simulation output because the only parameters are the species' relative abundances.
@@
 
-Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out five communities (Fig.
-2).
+Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out of five communities (Fig.
+1).
 Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
-Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S1).
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 1), but the effect was small (9% increase; Table S1).
 Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S2).
-Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
+Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 1; "No Comp." simulations).
 Removing competition caused, at most, a 5% change in synchrony (Table S2).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
-2), with an average 6% increase over synchrony from IBM simulations on a five meter area.
+1), with an average 6% increase over synchrony from IBM simulations on a five meter area.
 The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
-3).
+2).
 For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
@@
 Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony.
-The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem function experiments [@DeMazancourt2013a].
+The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem functioning experiments [@DeMazancourt2013a].
 
@@
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
-Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of specuies) in a non-constant environment (Fig. 3).
+Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of species) in a non-constant environment (Fig. 2).
 This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
@@
 Rare species, which our analysis does not include, could be kept at low densities by competitive suppression due to strong interspecific interactions.
-If we had been able to model rare species, we may have found a larger effect of competition on community synchrony.
+If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
 However, rare species may occupy rare niches that do not overlap with other species, in which case synchrony would remain predominantly driven by environmental responses.
@@
 The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables S1, S3-S17).
-Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap increases.
+Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap decreases.
 In that sense, our results are not surprising: interspecific interactions are weak, so of course removing them does not affect synchrony.
 
-However, our results do contrast with a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives species synchrony in grasslands [@Gross2014].
+However, our results do contrast with a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives species synchrony in grasslands [@Gross2014].
 The apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
@@
 The natural, relatively undisturbed communities we studied consist of species that have co-occurred for decades [@Chu2015] and represent a much later stage of community assembly.
-In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments the processes driving competitive exclusion are still operating.
+In theory, species interactions should weaken through time as community assembly processes [@Kokkoris1999], meaning that in biodiversity-ecosystem functioning experiments the processes driving competitive exclusion are still operating.
 Such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
@@
 In other words, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
-One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.
+One way to test this prediction is to continue collecting data from biodiversity-ecoystem functioning experiments and conduct the analysis of @Gross2014 every few years.
 
@@
 In agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
-Such a conclusion is in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
+Such a conclusion is in contrast to a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives community synchrony [@Gross2014].
 Recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
@@
 colnames(table_data) <- c("Site",
-                          "$\\phi_{r}$", 
+                          "$\\phi_{R}$", 
                           "$\\phi_{C}$",
@@
   \caption{Synchrony of species per capita growth rates simulated in monoculture (without species interactions) versus synchrony in polyculture (with species interactions).
-We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem function experiments.
+We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
 The dashed line is the line of equality.}

### 2016-03-18 docs/ms_communitySynchrony.Rmd
adds CdM edits and figure updates to main text

@@
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-In all five communities, theoretical predictions closely matched simulated synchrony for the limiting cases where only demographic stochasticity or environmental variation are present, meaning that theory developed with simple models also applies to our empirical models.
 Simulation experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
@@
 Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
-The stabilizing effect of species richness stems from individual species fluctuating asynchronously through time [@Hector2010; @Isbell2009; @DeMazancourt2013a].
+The stabilizing effect of species richness stems from individual species responding to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity).
+[@Hector2010; @Isbell2009; @DeMazancourt2013a].
 Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
@@
 The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific competition is to remove them one-by-one, and in combination.
-In principle, this could be done experimentally, but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
+In principle, this could be done in an extremely controlled laboratory setting, but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
 For example, @Mutshinda2009 fit a dynamic population model to several community time series of insect and bird abundances.
@@
 However, we can derive analytical predictions for species synchrony under special limiting cases.
-We constrain our focus to synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
-However, we provide analytical predictions for synchrony of abundances in the Online Supporting Information.
+We focus on synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
+We present equivalent results for synchrony of species abundances in the Online Supporting Information, and show that they lead to the same overall conclusions as synchrony of per capita growth rates.
 Following @Loreau2008a and @DeMazancourt2013a, we define population growth, ignoring observation error, as
@@
 
-To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION... (see Online Supporting Information).
+To derive analytical predictions we solved a first-order approximation of Equation 2 (Online Supporting Information).
 
@@
 
-\noindent{} for synchrony of per capita growth rates.
-Again, $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$, and $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
-Confronting our theoretical predictions with data requires estimates of species dynamics in monoculture to calculate the covariance of species' environmental responses.
+\noindent{} for synchrony of per capita growth rates, where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
+Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in monoculture (no interspecific interactions) to calculate the covariance of species' environmental responses.
 To estimate environmental responses in natural communities, we turn to our population models built using long-term demographic data.
@@
 We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
-Genet size and crowding were treated as fixed effects; spatial (quadrat groupings) and temporal (interannual) variation were treated as random effects on the intercept.
+Genet size and crowding were treated as fixed effects; spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
 
@@
 Once we have fit the vital rate statistical models, building the population models is straightforward.
-For the IBM, we initialize the model by randomly assigning plants spatial coordinates, sizes and species identities until each species achieves a density representative of that observed in the data.
+For the IBM, we initialize the model by randomly assigning plants spatial coordinates, sizes, and species identities until each species achieves a density representative of that observed in the data.
 We then project the model forward by using the survival regression to determine whether a genet lives or dies, the growth regression to calculate changes in genet size, and the recruitment regression to add new individuals that are distributed randomly in space.
@@
 
-The IPM uses the same vital rate regressions as the IBM [@Rees2009; @Rees2014], but it is spatially-implicit and does not include demographic stochasticity.
+The IPM uses the same vital rate regressions as the IBM [@Rees2009; @Rees2014], but it is spatially implicit and does not include demographic stochasticity.
 Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step (Supporting Online Information).
@@
 plots) of different size.
+Indeed, results from an IBM with infinite population size would equal results from the IPM.
+Given computational constraints, the largest landscape we simulate is a 5 square meter plot.
 
@@
 To explore the effect of demographic stochasticity in different sized populations, we ran simulations (1) and (6) on plot sizes of 1, 2, 3, 4, and 5 square meters.
+All other IBM simulations were run on a 5 square meter landscape.
 
@@
 2).
-For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
+For all five communities, the synchrony of species' growth rates in monoculture almost perfectly matched species synchrony in polyculture (Fig. 3).
 Results for synchrony of percent cover are qualitatively similar (Figs. S2-S3) but more variable, as expected by theory [@Loreau2008a].
@@
 Demographic stochasticity partially counteracted the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations were small (Fig.
-3), in agreement with theory [@Loreau2008a].
+2), in agreement with theory [@Loreau2008a].
 Even in small populations, however, demographic stochasticity was not strong enough to completely counteract the synchronizing effects of environmental fluctuations and induce independent species fluctuations (Fig.
-3).
+2).
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
@@
 This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
+
+Our analysis of how demographic stochasticity affects synchrony across population sizes also demonstrates that synchrony depends on the observation area.
+As the observation area increases, population size increases and the desynchronizing effect of demographic stochasticity lessens (Fig. 2).
+Thus, our results suggest that community-wide species synchrony will increase as the observation area increases.
+Such a conclusion assumes, however, that species richness remains constant as observation area increases, which is unlikely [@Taylor1961].
+Recent theoretical work has begun to explore the linkage between ecosystem stability and spatial scale [@Wang2014; @Wang2016], and our results suggest that including demographic stochasticity in theoretical models of metacommunity dynamics may be important for understanding the role of species synchrony in determining ecosystem stability across spatial scales.
 
@@
 Two lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
-First, in our simulation experiments, removing interspecific interactions resulted in no discernible change in community-wide species synchrony of per capita growth rates (Fig. 2).
+First, in our simulation experiments, removing interspecific interactions resulted in no discernible change in community-wide species synchrony of per capita growth rates (Fig. 1).
 Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
-2), and the synchrony of species in monoculture was a very strong predictor of synchrony of species in polyculture (Fig.
+1), and the synchrony of species in monoculture was a very strong predictor of synchrony of species in polyculture (Fig.
 3).
@@
 
-```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}
+```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=F, cache=FALSE}
 exp_synchrony_env <- readRDS("../results/envstoch_predictions.RDS")
@@
 Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation.
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+Error bars represent the upper and lower 95\% quantiles from model simulations.
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
 \end{figure}
@@
       \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
-  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S.
-+ No Comp.'').
-The strength of demographic stochasticity decreases as landscape size increases.
+  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
+Only'') and when environmental stochasticity is also present removed (``D.S.
++ E.S.'').
+The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
+The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
+Theoretically, ``D.S.
+Only'' simulations should remain constant across landscape size, whereas ``D.S.
++ E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
 Error bars represent the upper and lower 95\% quantiles from model simulations.}
@@
       \includegraphics[width=4in]{./components/formatted_figures/formatted_figure3.png}
-  \caption{Synchrony of species per capita growth rates simulated in monoculture (without species interactions) versus synchrony in polyculture (with species interactions).
+  \caption{Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions).
 We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
-The dashed line is the line of equality.}
+The dashed line is the line of equality.
+Simulation results in this figure are analagous to ``No Comp.
++ No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.
+1, but here we control the time series of random year effects.}
 \end{figure}

### 2016-03-18 devel/make_ms_plots.R

@@ -152,7 +152,7 @@ saveRDS(plot_df, "../results/allsims_plot_data.RDS")
 
 ##  Make demographic stochasiticty plot for all landscape sizes -----
 ibm_demo_rms <- subset(ibm_synch_agg, typesynch=="Per capita growth rate")
-ibm_demo_rms <- ibm_demo_rms[which(ibm_demo_rms$experiment %in% c("fluctinter", "constinter")),]
+ibm_demo_rms <- ibm_demo_rms[which(ibm_demo_rms$experiment %in% c("fluctnointer", "constnointer")),]
 ibm_demo_rms[which(ibm_demo_rms$site == "NewMexico"),"site"] <- "1New Mexico"
 ibm_demo_rms[which(ibm_demo_rms$site == "Arizona"),"site"] <- "2Arizona"
 ibm_demo_rms[which(ibm_demo_rms$site == "Kansas"),"site"] <- "3Kansas"
@@ -170,7 +170,7 @@ ggplot(ibm_demo_rms, aes(x=expansion, y=avg_synch, color=site))+
   xlab(expression(paste("Simulated Area (", m^2,")")))+
   ylab("Synchrony of Species' Growth Rates")+
   scale_y_continuous(limits=c(0,1))+
-  scale_shape_discrete(name="",labels=c("No E.S", "All Drivers"))+
+  scale_shape_discrete(name="",labels=c("D.S. Only", "D.S + E.S."))+
   # scale_linetype_discrete(name="",labels=c("No E.S", "All Drivers"))+
   guides(color=FALSE)+
   theme_few()+

### 2016-03-18 docs/ms_communitySynchrony.Rmd
last changes to demoggraphic discussion section

@@
 2), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity was not strong enough to completely counteract the synchronizing effects of environmental fluctuations and induce independent species fluctuations (Fig.
+Even in small populations, however, demographic stochasticity was not strong enough to completely counteract the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
 2).
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
-Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of species) in a non-constant environment (Fig. 2).
+Only in the Idaho community does synchrony of per capita growth rates approach $\mathcal{M}_D$ in a non-constant environment (Fig. 2).
 This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
@@
 As the observation area increases, population size increases and the desynchronizing effect of demographic stochasticity lessens (Fig. 2).
-Thus, our results suggest that community-wide species synchrony will increase as the observation area increases.
+Thus, our results suggest that community-wide species synchrony will increase as the observation area increases, rising from $\mathcal{M}_D$ to $\mathcal{M}_E$.
 Such a conclusion assumes, however, that species richness remains constant as observation area increases, which is unlikely [@Taylor1961].

### 2016-03-18 docs/online_supporting_information.Rmd
adds CdM's text for deriving predictions

@@
 \begin{align}
-\tilde{r_i}(t) = r_{mi} \left[ 1- \frac{\tilde{N_i}(t)+\sum_{j \neq i} \alpha_{ij}\tilde{N_j}(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{\tilde{N_i}(t)}} \right]
+r_i(t) = r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}u_{ei}(t) + \frac{\sigma_{di}u_{di}(t)}{\sqrt{N_i(t)}} \right]
 \end{align}
 
-\noindent where $N_i(t)$ is the biomass of species *i* in year *t*, and $r_i(t)$ is its instantaneous growth rate in year *t*. 
+\noindent where $N_i(t)$ is the biomass of species *i* in year *t*, and $r_i(t)$ is its growth rate in year *t*. 
 $r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*.
-Environmental stochasticity is incorporated as $\sigma_{ei}\mu_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $\mu_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated.
+Environmental stochasticity is incorporated as $\sigma_{ei}u_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $u_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated between species.
 Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first order, normal approximation [@Lande2003; @DeMazancourt2013a].
-$\sigma_{di}^2$ is the demographic variance and $\mu_{di}(t)$ are independent normal variables with zero mean and unit variance.
+$\sigma_{di}^2$ is the demographic variance and $u_{di}(t)$ are independent normal variables with zero mean and unit variance.
 
-In the general case, **C** is the variance-covariance matrix for population biomasses at steady state.
-It solves the Lyapunov equation $\textbf{C}^\infty = \textbf{A} \textbf{C}^\infty \textbf{A}^T + \textbf{B}$, where
+First-order approximations of the temporal variance of total community biomass are obtained as follows [@Ives1995; @Hughes2000; @Ives2002; @Loreau2008a; @DeMazancourt2013a].
+Let $\delta N_i(t) = N_i(t) - N^*_i$ denote the deviation of observed species *i*’s biomass from its equilibrium value in the community, $N^*_i$, in the absence of stochasticity.
+Equation (S1) can be Taylor expanded around $\delta N_i(t) = u_{ei}(t) = u_{di}(t) = u^S_{oi}(t) = 0$ to yield, after dropping terms of order two and higher,
+
+\begin{equation}
+\boldsymbol{\delta}\textbf{N}(t+1) = \textbf{A}\boldsymbol{\delta}\textbf{N}(t) + \textbf{z}(t),
+\end{equation}
+
+\noindent where $\boldsymbol{\delta}\textbf{N}(t)$ is the vector of deviations of species biomasses from their deterministic equilibrium value, **A** is the community matrix, also known as the Jacobian matrix around the equilibrium, with elements $(a_{ij})_{1<i,j<S}$:
 
@@
 
-\noindent and
+\noindent and $\textbf{z}(t)$ is a vector that encapsulates the effects of environmental and demographic stochasticity whose elements are
 
 \begin{equation}
-B_{ij} = N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})
+z_i(t) = N^*_i \sigma_{ei} u_{ei}(t) + \sqrt{N^*_i} \sigma_{di} u_{di}(t)
+\end{equation}
+
+When the system reaches a stationary distribution, the variances and covariances between species biomass time series are:
+
+\begin{equation}
+\big\langle \boldsymbol{\delta}\textbf{N}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle = \textbf{C}^\infty = (\text{cov}(N_i,N_j))_{1<i,j<S} = (\text{cov}(\delta N_i, \delta N_j))_{1<i,j<S}
+\end{equation}
+
+\noindent where $\boldsymbol{\delta}\textbf{N}^T$ is the transpose of vector $\boldsymbol{\delta}\textbf{N}$, i.e. a row vector.
+
+Our assumptions listed above lead to the following correlation structure of **z**:
+
+\begin{equation}
+\big\langle \textbf{z}(t) \textbf{z}(t)^T \big\rangle = \textbf{B}
+\end{equation}
+
+\begin{equation}
+\big\langle \textbf{z}(t-s) \textbf{z}(t)^T \big\rangle = 0 \qquad \text{for} \: s>0
+\end{equation}
+
+We use Equation S2 to write a dynamical equation for the covariance **C**:
+
+\begin{footnotesize}
+\begin{align}
+\textbf{C}(t+1) =& \big\langle \boldsymbol{\delta}\textbf{N}(t+1) \boldsymbol{\delta}\textbf{N}(t+1)^T \big\rangle \\
+=& \textbf{A}\big\langle \boldsymbol{\delta}\textbf{N}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle \textbf{A}^T + \textbf{A}\big\langle \boldsymbol{\delta}\textbf{N}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle + \big\langle \textbf{z}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle \textbf{A}^T + \big\langle \textbf{z}(t) \textbf{z}(t)^T \big\rangle \\
+=& \textbf{A} \textbf{C}(t) \textbf{A}^T + \textbf{Z}_0
+\end{align}
+\end{footnotesize}
+
+\noindent Taking the limit $t \rightarrow \infty$ on both sides, we get:
+
+\begin{equation}
+\textbf{C}^\infty = \textbf{A} \textbf{C}^\infty \textbf{A}^T + \textbf{B}
+\end{equation}
+
+\noindent where **A** is as in Equation S3 and **B** is:
+
+\begin{equation}
+B_{ij} = N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})
 \end{equation}
@@
 \begin{equation}
-R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} \sum_{k,l} \alpha_{ik} \alpha_{jl} C_{kl} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}}.
+R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} \sum_{k,l} \alpha_{ik} \alpha_{jl} C_{kl} + \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{\sqrt{N_{i}^* N_{j}^*}}.
 \end{equation}
@@
 \begin{equation}
-c_{ij} = \frac{B_{ij}}{1-a_i a_j} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
+c_{ij} = \frac{B_{ij}}{1-a_i a_j} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
 \end{equation}
@@
 \begin{equation}
-R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} C_{ij} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}},
+R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} C_{ij} + \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{\sqrt{N_{i}^* N_{j}^*}},
 \end{equation}
@@
 \begin{equation}
-R_{ij} = \frac{1}{1- \frac{r_{mi}r_{mj}}{r_{mi}+r_{mj}}} \left( \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}} \right)
+R_{ij} = \frac{1}{1- \frac{r_{mi}r_{mj}}{r_{mi}+r_{mj}}} \left( \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{\sqrt{N_{i}^* N_{j}^*}} \right)
 \end{equation}
 
-\noindent This is still not a general result, however, because synchrony in population sizes and synchrony in growth rates use weighted factors of the environmental variances and covariances with species-specific parameters.
+\noindent Note that synchrony in population sizes and synchrony in growth rates use weighted factors of the environmental variances and covariances with species-specific parameters.
 So, we further assume that, along with interspecific interactions being zero, all species have identical growth rates, environmental stochasticity is absent, and all species have identical demographic variance.
@@
 When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a].
-The prediction represented in Equation S11 is Equation 3 in the main text, which we refer to as $\phi_{R,\mathcal{M_D}}$.
+The prediction represented in Equation S20 is Equation 3 in the main text, which we refer to as $\phi_{R,\mathcal{M_D}}$.
 
@@
 
-\noindent The prediction represented in Equation S13 is Equation 4 in the main text, which we refer to as $\phi_{R,\mathcal{M_E}}$.
+\noindent The prediction represented in Equation S22 is Equation 4 in the main text, which we refer to as $\phi_{R,\mathcal{M_E}}$.
 If we know the covariance matrix of the species environmental responses, we can calculate the above expectations directly.
@@
 ------
-```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
+```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=F, cache=FALSE, warning=FALSE}
 library(reshape2)
@@
 perc_changes <- read.csv("../results/synchsims_percent_diffs.csv")
-perc_changes <- perc_changes[which(perc_changes$simulation!="1All Drivers"), c("site", "simulation", "percent_diff")]
+perc_changes <- perc_changes[which(perc_changes$simulation!="3All Drivers"), c("site", "simulation", "percent_diff")]
 table_diff <- dcast(perc_changes, simulation~site, value.var = "percent_diff")

### 2016-03-18 docs/online_supporting_information.Rmd
changes line spacing to single

@@
    - \usepackage{todonotes}
-   - \doublespacing
-   - \usepackage[document]{ragged2e}
    - \usepackage{color}
@@
 
+\begin{center} 
 \textbf{ONLINE SUPPORTING INFORMATION}
+\end{center}
 
@@
 When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a].
-The prediction represented in Equation S20 is Equation 3 in the main text, which we refer to as $\phi_{R,\mathcal{M_D}}$.
+The prediction represented in Equation S20 is Equation 3 in the main text, which we refer to as $\phi_{R,\mathcal{M}_D}$.
 
@@
 
-\noindent The prediction represented in Equation S22 is Equation 4 in the main text, which we refer to as $\phi_{R,\mathcal{M_E}}$.
+\noindent The prediction represented in Equation S22 is Equation 4 in the main text, which we refer to as $\phi_{R,\mathcal{M}_E}$.
 If we know the covariance matrix of the species environmental responses, we can calculate the above expectations directly.
@@
 
-Figures
--------
+Supporting Figures
+------------------
 \begin{figure}[!ht]
@@
 
-Tables
-------
+Supporting Tables
+-----------------
 ```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=F, cache=FALSE, warning=FALSE}

### 2016-03-18 docs/ms_communitySynchrony.Rmd
changes site orders in table to go by species richness

@@
 $\sigma_{di}^2$ is the demographic variance and $u_{di}(t)$ are independent normal variables with zero mean and unit variance.
-
-To derive analytical predictions we solved a first-order approximation of Equation 2 (Online Supporting Information).
+To derive analytical predictions we solved a first-order approximation of Equation 2 [@DeMazancourt2013a and Online Supporting Information].
 
@@
 | --------- | ----- | ------------------- | ---------- | ------- |
+| New Mexico | Chihuahuan Desert | 32.62&deg; N, 106.67&deg; W | 1915-1950 | *Bouteloua eriopoda* |
+| | | | | *Sporobolus flexuosus* |
 | Arizona | Sonoran Desert | 31&deg;50' N, 110&deg;53' W | 1915-1933 | *Bouteloua eriopoda* |
 | | | | | *Bouteloua rothrockii* |
-| Idaho | Sagebrush steppe | 44.2&deg; N, 112.1&deg; W | 1926-1957 | *Artemisia tripartita* |
-| | | | | *Pseudoroegneria spicata* |
-| | | | | *Hesperostipa comata* |
-| | | | | *Poa secunda* |
 | Kansas | Southern mixed prairie | 38.8&deg; N, 99.3&deg; W | 1932-1972 | *Bouteloua curtipendula* |
@@
 | | | | | *Poa secunda* |
-| New Mexico | Chihuahuan Desert | 32.62&deg; N, 106.67&deg; W | 1915-1950 | *Bouteloua eriopoda* |
-| | | | | *Sporobolus flexuosus* |
+| Idaho | Sagebrush steppe | 44.2&deg; N, 112.1&deg; W | 1926-1957 | *Artemisia tripartita* |
+| | | | | *Pseudoroegneria spicata* |
+| | | | | *Hesperostipa comata* |
+| | | | | *Poa secunda* |
 
@@
                           "Species richness")
+myorder <- c(5,1,3,4,2)
+xtable_data <- table_data[myorder,]
 # colnames(table_data) <- c("Site",
@@
 Species richness values reflect the number of species analyzed from the community, not the actual richness."
-print(xtable(table_data, caption = synch_cap),
+print(xtable(xtable_data, caption = synch_cap),
       caption.placement="top",
@@
 
-```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=F, cache=FALSE}
+```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}
 exp_synchrony_env <- readRDS("../results/envstoch_predictions.RDS")
@@
 
+myorder <- c(5,1,3,4,2)
+xpred_table <- pred_table[myorder,]
 pred_cap <- "Comparisons between our analytical predictions and simulation results for synchrony of species' per capita growth rates.
@@
 Simulated synchrony values come from our empirically-based, multi-species population models when simulated under conditions that match the limiting case conditions (e.g., environmental stochasticity and competition removed for $\\mathcal{M}_D$)."
-print(xtable(pred_table, caption = pred_cap),
+print(xtable(xpred_table, caption = pred_cap),
       caption.placement="top",

### 2016-03-18 devel/make_ms_plots.R
updates SI plots to match main text ordering of sites and colors

@@ -143,7 +143,7 @@ ggplot(data=plot_df)+
 ggsave("../docs/components/all_sims_results.png", width = 10, height = 4, units="in", dpi=75)
   
 ##  Calculate percent differences for Results
-plot_df$control <- rep(plot_df[which(plot_df$simulation=="1All Drivers"),"synchrony"], times=nrow(plot_df)/nsites)
+plot_df$control <- rep(plot_df[which(plot_df$simulation=="3All Drivers"),"synchrony"], times=nrow(plot_df)/nsites)
 plot_df$percent_diff <- with(plot_df, abs(synchrony-control)/((synchrony+control)/2)*100)
 write.csv(plot_df, "../results/synchsims_percent_diffs.csv")
 
@@ -198,11 +198,12 @@ ibm_synch <- subset(ibm_synch_agg, expansion==5 & typesynch=="Cover (%)")
 
 ##  Combine results -----
 sim_names <- c("All Drivers", "No D.S.", "No E.S.", "No Comp.", "No Comp. + No D.S.", "No Comp. + No E.S.")
-sim_names_order <- paste0(1:length(sim_names),sim_names)
+sim_names_order <- paste0(c(3,1,5,4,2,6),sim_names)
 site_names <- unique(ipm_synch$site)
 nsites <- length(site_names)
 site_labels <- site_names
 site_labels[which(site_labels=="NewMexico")] <- "New Mexico"
+site_labels_order <- paste0(c(2,5,3,4,1), site_labels)
 
 # Get vectors of synchrony for each "experiment"
 control_synch <- subset(ibm_synch, experiment=="fluctinter")
@@ -230,23 +231,32 @@ all_downs <- c(control_synch$lo_synch,
                nodsnocomp_synch$lo_synch, 
                noesnocomp_synch$lo_synch)
 
-plot_df <- data.frame(site = rep(site_labels, times=length(sim_names)),
+plot_df <- data.frame(site = rep(site_labels_order, times=length(sim_names)),
                       simulation = rep(sim_names_order, each=nsites),
                       synchrony = all_experiments,
                       upper_synch = all_ups,
                       lower_synch = all_downs)
 
+
+site_labels <- site_labels[order(site_labels_order)]
+sim_labels <- sim_names[order(sim_names_order)]
+
 ##  Make the main (all sims) plot -----
[…]

### 2016-03-18 docs/ms_communitySynchrony.Rmd
updates SI figures to match main text figure ordering and colors

@@
 Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 1; "No Comp." simulations).
-Removing competition caused, at most, a 5% change in synchrony (Table S2).
+Removing competition caused, at most, a 5% change in synchrony (Fig. 1).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
@@
 
-synch_cap <- "Observed synchrony among species' per capita growth rates ($\\phi_{r}$), observed synchrony among species' percent cover ($\\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community.
+synch_cap <- "Observed synchrony among species' per capita growth rates ($\\phi_{R}$), observed synchrony among species' percent cover ($\\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community.
 Species richness values reflect the number of species analyzed from the community, not the actual richness."

### 2016-03-18 docs/online_supporting_information.Rmd
updates SI figures to match main text figure ordering and colors

@@
       \includegraphics[width=6in]{./components/figureS3_ibmcover.png}
-  \caption{Synchrony of species' percent cover for each study area from IBM simulations across different landscape sizes when all drivers are present (``All Drivers'') and where only demographic stochasticity is present (``No E.S.
-+ No Comp.'').
-The strength of demographic stochasticity decreases as landscape size increases.
+  \caption{Synchrony of species' percent cover for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
+Only'') and when environmental stochasticity is also present removed (``D.S.
++ E.S.'').
+The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
 Error bars represent the upper and lower 95\% quantiles from model simulations.}

### 2016-03-18 runscripts/Idaho_NoARTR_sims/make_ms_plots_noARTR.R

@@ -7,7 +7,7 @@ library(plyr)
 library(synchrony)
 library(communitySynchrony)
 
-site_colors <- c("steelblue")
+site_colors <- c("purple")
 
 
 ##  Read in IPM results ---------
@@ -62,7 +62,7 @@ ibm_synch <- subset(ibm_synch_agg, expansion==5 & typesynch=="Per capita growth 
 
 ##  Combine results -----
 sim_names <- c("All Drivers", "No D.S.", "No E.S.", "No Comp.", "No Comp. + No D.S.", "No Comp. + No E.S.")
-sim_names_order <- paste0(1:length(sim_names),sim_names)
+sim_names_order <- paste0(c(3,1,5,4,2,6),sim_names)
 site_names <- unique(ipm_synch$site)
 nsites <- length(site_names)
 site_labels <- site_names
@@ -123,25 +123,25 @@ ggsave("../../docs/components/all_sims_supp_noARTR.png", width = 4, height = 4, 
 # write.csv(plot_df, "../results/synchsims_percent_diffs.csv")
 
 ##  Make demographic stochasiticty plot for all landscape sizes -----
-ibm_demo_rms <- subset(ibm_synch_agg, typesynch=="Per capita growth rate")
-ibm_demo_rms <- ibm_demo_rms[which(ibm_demo_rms$experiment %in% c("fluctinter", "constnointer")),]
-
-ggplot(ibm_demo_rms, aes(x=expansion, y=avg_synch, color=site))+
-  geom_line(aes(linetype=experiment))+
-  geom_point(aes(shape=experiment), size=3)+
-  geom_errorbar(aes(ymin=lo_synch, ymax=up_synch), width=0.25)+
-  scale_color_manual(values=site_colors, labels=site_labels, name="")+
-  xlab(expression(paste("Simulated Area (", m^2,")")))+
-  ylab("Synchrony of Species' Growth Rates")+
-  scale_y_continuous(limits=c(0,1))+
-  scale_shape_discrete(name="",labels=c("No E.S + No Comp.", "All Drivers"))+
-  scale_linetype_discrete(name="",labels=c("No E.S + No Comp.", "All Drivers"))+
-  guides(color=FALSE)+
-  theme_few()+
-  theme(legend.position=c(0.3,0.8),
-        legend.background = element_rect(fill = NA))
-
[…]

### 2016-03-19 docs/ms_communitySynchrony.Rmd
changes results to all caps

@@
 
-### Results
+RESULTS
+-------
 Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).

### 2016-03-19 docs/ms_communitySynchrony.Rmd
udpates fig S1 site order

@@
 
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IPM (left panels) and the IBM (right panels).
+```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IBM (top panels) and the IPM (bottom panels).
 IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.

### 2016-03-19 docs/online_supporting_information.Rmd
udpates fig S1 site order

@@
   \centering
-      \includegraphics[width=4in]{./components/figureS1.pdf}
-  \caption{Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IPM (left panels) and the IBM (right panels).
+      \includegraphics[width=6in]{./components/figureS1.png}
+  \caption{Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IBM (top panels) and the IPM (bottom panels).
 IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.

### 2016-03-19 devel/make_figS1.R

@@ -1,0 +1,244 @@
+##  Main plot for synchrony results
+
+library(ggplot2)
+library(ggthemes)
+library(reshape2)
+library(plyr)
+library(synchrony)
+library(communitySynchrony)
+library(gridExtra)
+
+site <- "Kansas"
+spp_list <- c("BOCU","BOHI","SCSC")
+num_spp <- length(spp_list)
+ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  ks_data <- rbind(ks_data, spp_data)
+} #end species looping for raw data
+ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
+
+tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
+tmp2<-which(ks_data$quad_data=="q27")
+tmp3<-which(ks_data$quad=="q28")
+tmp4<-which(ks_data$quad=="q30")
+tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
+tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
+tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
+ks_data<-ks_data[-tmp,]
+
+# exclude the records later than 1968, to keep the same random year effect...
+ks_data<-subset(ks_data,year<68)
+
+
+site <- "Idaho"
+spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
+num_spp <- length(spp_list)
+id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  id_data <- rbind(id_data, spp_data)
+} #end species looping for raw data
+id_data <- id_data[2:nrow(id_data),] #remove first NA row
[…]

### 2016-03-21 docs/ms_communitySynchrony.Rmd
adds number of refs to title page
adds color formatting for coauthor review of sections

@@
 
-\textit{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
+\textbf{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
 
@@
 
-\textbf{Number of References}:
+\textbf{Number of References}: 46
 
@@
 plots) of different size.
-Indeed, results from an IBM with infinite population size would equal results from the IPM.
-Given computational constraints, the largest landscape we simulate is a 5 square meter plot.
+\textcolor{red}{Indeed, results from an IBM with infinite population size would equal results from the IPM.
+Given computational constraints, the largest landscape we simulate is a 5 square meter plot.}
 
@@
 
-RESULTS
--------
+### Results
 Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
@@
 1), with an average 6% increase over synchrony from IBM simulations on a five meter area.
-The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
+\textcolor{red}{The desynchronizing effect of demographic stochasticity, which increases as population size decreases, modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
 2).
+Synchrony from simulations with demographic and environmental stochasticity decreased as the simulated area decreased, representing a shift from synchrony being driven by environmental responses (e.g., $\mathcal{M}_E$) to being driven by a combination of demographic and environmental stochasticity (e.g., between $\mathcal{M}_E$ and $\mathcal{M}_D$).}
 For all five communities, the synchrony of species' growth rates in monoculture almost perfectly matched species synchrony in polyculture (Fig. 3).
@@
 
-Our analysis of how demographic stochasticity affects synchrony across population sizes also demonstrates that synchrony depends on the observation area.
+\textcolor{red}{Our analysis of how demographic stochasticity affects synchrony across population sizes also demonstrates that synchrony depends on the observation area.
 As the observation area increases, population size increases and the desynchronizing effect of demographic stochasticity lessens (Fig. 2).
@@
 Such a conclusion assumes, however, that species richness remains constant as observation area increases, which is unlikely [@Taylor1961].
-Recent theoretical work has begun to explore the linkage between ecosystem stability and spatial scale [@Wang2014; @Wang2016], and our results suggest that including demographic stochasticity in theoretical models of metacommunity dynamics may be important for understanding the role of species synchrony in determining ecosystem stability across spatial scales.
+Recent theoretical work has begun to explore the linkage between ecosystem stability and spatial scale [@Wang2014; @Wang2016], and our results suggest that including demographic stochasticity in theoretical models of metacommunity dynamics may be important for understanding the role of species synchrony in determining ecosystem stability across spatial scales.}
 
@@
 ### Interspecific interactions had little impact on community synchrony
-We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a and Supplementary Information].
+We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a].
 We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
@@
 When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2).
-A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
+\textcolor{blue}{A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
 Such a conclusion is only half correct.
 The perennial grasses we studied do have similar responses to the environment (Table S2), which will tend to synchronize dynamics.
-However, if community-wide species synchrony is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
+However, if community-wide species synchrony is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].}
 In agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.

### 2016-03-22 docs/ms_communitySynchrony.Rmd
adds ML final edits

--- 2016-03-19 docs/ms_communitySynchrony.Rmd
@@
 \begin{center}
-\large{\textbf{Environmental responses, not species interactions, determine species synchrony in five natural plant communities}}
+\large{\textbf{Environmental responses, not species interactions, determine species synchrony in natural plant communities}}
 
@@
 Indeed, results from an IBM with infinite population size would equal results from the IPM.
-Given computational constraints, the largest landscape we simulate is a 5 square meter plot.
+Given computational constraints, the largest landscape we simulate is a 25 $\text{m}^2$ plot.
 
@@
 1), with an average 6% increase over synchrony from IBM simulations on a five meter area.
-The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
+The desynchronizing effect of demographic stochasticity, which increases as population size decreases, modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
 2).
+Synchrony from simulations with demographic and environmental stochasticity decreased as the simulated area decreased, representing a shift from synchrony being driven by environmental responses (e.g., $\mathcal{M}_E$) to being driven by a combination of demographic and environmental stochasticity (e.g., between $\mathcal{M}_E$ and $\mathcal{M}_D$).
 For all five communities, the synchrony of species' growth rates in monoculture almost perfectly matched species synchrony in polyculture (Fig. 3).
@@
 We thank the original mappers of the permanent quadrats at each site and the digitizers in the Adler lab, without whom this work would not have been possible.
-Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.
+Compute, storage, and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.
 

### 2016-03-22 docs/ms_communitySynchrony.Rmd
adds CdM final edits

@@
 
-Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Table 3).
+Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Table S1).
 The correlation between our analytical predictions and simulated synchrony was 0.97 for $\phi_{R,\mathcal{M}_D}$ and 0.997 for $\phi_{R,\mathcal{M}_E}$.
@@
 Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
-Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 1), but the effect was small (9% increase; Table S1).
-Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S2).
-Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 1; "No Comp." simulations).
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 1), but the effect was small (9% increase; Table S2).
+Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S3).
+Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 1; "No Comp." simulations).
 Removing competition caused, at most, a 5% change in synchrony (Fig. 1).
@@
 
-```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}
-exp_synchrony_env <- readRDS("../results/envstoch_predictions.RDS")
-exp_synchrony_env <- exp_synchrony_env[complete.cases(exp_synchrony_env),]
-exp_synchrony_demo <- alltable_data[,c("V2","V5")]
-exp_synchrony_demo$site <- table_data$Site
-colnames(exp_synchrony_demo) <- c("pgr", "cover", "site")
-sims_output <- readRDS("../results/allsims_plot_data.RDS")
-
-pred_table <- data.frame(site = table_data$Site,
-                         exp_demo = exp_synchrony_demo$pgr,
-                         sim_demo = round(subset(sims_output, simulation=="6No Comp. + No E.S.")["synchrony"],2),
-                         exp_env = round(exp_synchrony_env$growthrate_prediction,2),
-                         sim_env = round(subset(sims_output, simulation=="2No Comp. + No D.S.")["synchrony"],2))
-
-colnames(pred_table) <- c("Site", 
-                          "Predicted $\\phi_{R,\\mathcal{M}_D}$",
-                          "Simulated $\\phi_{R,\\mathcal{M}_D}$",
-                          "Predicted $\\phi_{R,\\mathcal{M}_E}$",
-                          "Simulated $\\phi_{R,\\mathcal{M}_E}$")
-
-myorder <- c(5,1,3,4,2)
-xpred_table <- pred_table[myorder,]
-pred_cap <- "Comparisons between our analytical predictions and simulation results for synchrony of species' per capita growth rates.
-Analytical predictions represent two limiting cases where only demographic stochasticity is operating ($\\phi_{R,\\mathcal{M}_D}$) and where only environmental stochasticity is operating ($\\phi_{R,\\mathcal{M}_E}$).
-Simulated synchrony values come from our empirically-based, multi-species population models when simulated under conditions that match the limiting case conditions (e.g., environmental stochasticity and competition removed for $\\mathcal{M}_D$)."
-print(xtable(xpred_table, caption = pred_cap),
-      caption.placement="top",
-      include.rownames = F, 
-      sanitize.colnames.function = identity,
-      comment=FALSE,
-      size="normalsize")
-```
-
-\pagebreak{}
-
 Figures
@@
 + No E.S.'').
-Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation.
+Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
 Error bars represent the upper and lower 95\% quantiles from model simulations.

### 2016-03-22 docs/online_supporting_information.Rmd
adds CdM final edits

@@
    - \usepackage{rotating}
+   - \usepackage{morefloats}
 bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
@@
 \begin{equation}
-c_{ij} = \frac{B_{ij}}{1-a_i a_j} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
+c_{ij} = \frac{B_{ij}}{1-A_ii A_jj} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
 \end{equation}
 
-and the variance-covariance matrix of population growth rates becomes
-
-\begin{equation}
-R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} C_{ij} + \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{\sqrt{N_{i}^* N_{j}^*}},
-\end{equation}
-
-which simplifies to
+and the variance-covariance matrix of population growth rates simplifies to
 
@@
   \centering
-      \includegraphics[width=6in]{./components/figureS2_simscover.png}
-  \caption{Community-wide species percent cover from model simulation experiments.
+      \includegraphics[width=6in]{./components/formatted_figures/formatted_figureS2.png}
+  \caption{Community-wide synchrony of species percent cover from model simulation experiments.
 Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
@@
 + No E.S.'').
-Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation.
+Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
 Error bars represent the upper and lower 95\% quantiles from model simulations.}
@@
   \centering
-      \includegraphics[width=6in]{./components/figureS3_ibmcover.png}
+      \includegraphics[width=6in]{./components/formatted_figures/formatted_figureS3.png}
   \caption{Synchrony of species' percent cover for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
@@
   \centering
-      \includegraphics{./components/all_sims_supp_noARTR.png}
-  \caption{Community-wide species percent cover from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
+      \includegraphics[width=3in]{./components/formatted_figures/formatted_figureS4.png}
+  \caption{Community-wide synchrony of species percent cover from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
 Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
@@
 + No E.S.'').
-Abbreviations within the bars for the Arizona site indicate whether the IBM or IPM was used for a particular simulation.
 Error bars represent the upper and lower 95\% quantiles from model simulations.}
@@
 -----------------
-```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=F, cache=FALSE, warning=FALSE}
+```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}
+library(communitySynchrony)
+library(plyr)
+library(reshape2)
+library(synchrony)
+library(ggplot2)
+library(xtable)
+library(gridExtra)
+library(RColorBrewer)
+site <- "Kansas"
+spp_list <- c("BOCU","BOHI","SCSC")
+num_spp <- length(spp_list)
+ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  ks_data <- rbind(ks_data, spp_data)
+} #end species looping for raw data
+ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
+
+tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
+tmp2<-which(ks_data$quad_data=="q27")
+tmp3<-which(ks_data$quad=="q28")
+tmp4<-which(ks_data$quad=="q30")
+tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
+tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
+tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
+ks_data<-ks_data[-tmp,]
+
+# exclude the records later than 1968, to keep the same random year effect...
+ks_data<-subset(ks_data,year<68)
+
+site <- "Idaho"
+spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
+num_spp <- length(spp_list)
+id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  id_data <- rbind(id_data, spp_data)
+} #end species looping for raw data
+id_data <- id_data[2:nrow(id_data),] #remove first NA row
+# id_data <- subset(id_data, species!="ARTR") #take out the shrub
+
+site <- "Montana"
+spp_list <- sort(c("BOGR","HECO","PASM","POSE"))
+num_spp <- length(spp_list)
+mt_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  mt_data <- rbind(mt_data, spp_data)
+} #end species looping for raw data
+mt_data <- mt_data[2:nrow(mt_data),] #remove first NA row
+
+site <- "NewMexico"
+spp_list <- sort(c("BOER","SPFL"))
+num_spp <- length(spp_list)
+nm_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  nm_data <- rbind(nm_data, spp_data)
+} #end species looping for raw data
+nm_data <- nm_data[2:nrow(nm_data),] #remove first NA row
+
+site <- "Arizona"
+spp_list <- sort(c("BOER","BORO"))
+num_spp <- length(spp_list)
+az_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
+for(dospp in 1:num_spp){ #loop through species to read in data
+  spp_now <- spp_list[dospp]
+  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
+  spp_data <- read.csv(quad_file)
+  spp_data$species <- spp_now
+  az_data <- rbind(az_data, spp_data)
+} #end species looping for raw data
+az_data <- az_data[2:nrow(az_data),] #remove first NA row
+out_ks <- get_comm_synchrony(ts_data = ks_data)
+out_id <- get_comm_synchrony(ts_data = id_data)
+out_mt <- get_comm_synchrony(ts_data = mt_data)
+out_nm <- get_comm_synchrony(ts_data = nm_data)
+out_az <- get_comm_synchrony(ts_data = az_data)
+get_table_metrics <- function(output){
+  tmp <- output
+  exp_tmp <- round(tmp$pgr_expected_synch_ind_flucts, 2)
+  obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
+  demo_tmp <- obs_tmp - exp_tmp
+  exp_cover <- round(as.numeric(tmp$cover_expected_synch_ind_flucts), 2)
+  obs_cover <- round(as.numeric(tmp$abund_synchrony[1]), 2)
+  demo_cover <- obs_cover - exp_cover
+  plant_size <- tmp$avg_size
+  cover_var <- tmp$variability
+  return(c(obs_tmp, exp_tmp, demo_tmp, obs_cover, exp_cover, 
+           demo_cover, cover_var))
+}
+
+table_data <- as.data.frame(rbind(get_table_metrics(out_az),
+                                  get_table_metrics(out_id),
+                                  get_table_metrics(out_ks),
+                                  get_table_metrics(out_mt),
+                                  get_table_metrics(out_nm)))
+table_data$Site <- c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico")
+alltable_data <- table_data
+exp_synchrony_env <- readRDS("../results/envstoch_predictions.RDS")
+exp_synchrony_env <- exp_synchrony_env[complete.cases(exp_synchrony_env),]
+exp_synchrony_demo <- alltable_data[,c("V2","V5")]
+exp_synchrony_demo$site <- table_data$Site
+colnames(exp_synchrony_demo) <- c("pgr", "cover", "site")
+sims_output <- readRDS("../results/allsims_plot_data.RDS")
+
+pred_table <- data.frame(site = table_data$Site,
+                         exp_demo = exp_synchrony_demo$pgr,
+                         sim_demo = round(subset(sims_output, simulation=="6No Comp. + No E.S.")["synchrony"],2),
+                         exp_env = round(exp_synchrony_env$growthrate_prediction,2),
+                         sim_env = round(subset(sims_output, simulation=="2No Comp. + No D.S.")["synchrony"],2))
+
+colnames(pred_table) <- c("Site", 
+                          "Predicted $\\phi_{R,\\mathcal{M}_D}$",
+                          "Simulated $\\phi_{R,\\mathcal{M}_D}$",
+                          "Predicted $\\phi_{R,\\mathcal{M}_E}$",
+                          "Simulated $\\phi_{R,\\mathcal{M}_E}$")
+
+myorder <- c(5,1,3,4,2)
+xpred_table <- pred_table[myorder,]
+pred_cap <- "Comparisons between our analytical predictions and simulation results for synchrony of species' per capita growth rates.
+Analytical predictions represent two limiting cases where only demographic stochasticity is operating ($\\phi_{R,\\mathcal{M}_D}$) and where only environmental stochasticity is operating ($\\phi_{R,\\mathcal{M}_E}$).
+Simulated synchrony values come from our empirically-based, multi-species population models when simulated under conditions that match the limiting case conditions (e.g., environmental stochasticity and competition removed for $\\mathcal{M}_D$)."
+print(xtable(xpred_table, caption = pred_cap),
+      caption.placement="top",
+      include.rownames = F, 
+      sanitize.colnames.function = identity,
+      comment=FALSE,
+      size="normalsize")
+```
+
+```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
 library(reshape2)
@@
 
-```{r get_interactions, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE}
+```{r get_interactions, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE, eval=T}
 ####

### 2016-03-22 devel/make_ms_plots.R

@@ -10,6 +10,10 @@ library(communitySynchrony)
 site_colors <- c("grey45", "steelblue", "slateblue4", "darkorange", "purple")
 
 
+
+####
+####  MAIN TEXT FIGURES 1-2
+####
 ##  Read in IPM results ---------
 output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
 mlist <- melt(output_list)
@@ -159,12 +163,12 @@ ibm_demo_rms[which(ibm_demo_rms$site == "Kansas"),"site"] <- "3Kansas"
 ibm_demo_rms[which(ibm_demo_rms$site == "Montana"),"site"] <- "4Montana"
 ibm_demo_rms[which(ibm_demo_rms$site == "Idaho"),"site"] <- "5Idaho"
 
-ggplot(ibm_demo_rms, aes(x=expansion, y=avg_synch, color=site))+
+ggplot(ibm_demo_rms, aes(x=(expansion^2), y=avg_synch, color=site))+
   geom_hline(data=theoretical_preds_and_obs, aes(yintercept=envonly_pred), color="grey15", linetype=3)+
   geom_hline(data=theoretical_preds_and_obs, aes(yintercept=demonly_pred), color="grey15", linetype=2)+
   geom_line(aes(group=experiment))+
   geom_point(aes(shape=experiment), size=3)+
-  geom_errorbar(aes(ymin=lo_synch, ymax=up_synch), width=0.25)+
+  geom_errorbar(aes(ymin=lo_synch, ymax=up_synch), width=1.5)+
   facet_wrap("site", nrow=1)+
   scale_color_manual(values=site_colors, labels=site_labels, name="")+
   xlab(expression(paste("Simulated Area (", m^2,")")))+
@@ -177,10 +181,14 @@ ggplot(ibm_demo_rms, aes(x=expansion, y=avg_synch, color=site))+
   theme(legend.position=c(0.1,0.2),
         legend.background = element_rect(fill = NA))
 
-ggsave("../docs/components/ibm_sims_across_landscape.png", width = 10, height = 3, units="in", dpi=75)
+ggsave("../docs/components/ibm_sims_across_landscape.png", width = 12, height = 3, units="in", dpi=75)
 
 
 
+
+####
+####  SUPPLEMENTAL FIGURES 2-3
+####
 ##  Make supplement plots for percent cover ------------
 ipm_synch_all <- ddply(synch_df, .(site, experiment, typesynch), summarise,
                        mean_synch = mean(synch),
[…]

### 2016-03-22 runscripts/Idaho_NoARTR_sims/make_ms_plots_noARTR.R

@@ -100,6 +100,7 @@ plot_df <- data.frame(site = rep(site_labels, times=length(sim_names)),
                       upper_synch = all_ups,
                       lower_synch = all_downs)
 
+sim_labels <- sim_names[order(sim_names_order)]
 ##  Make the main (all sims) plot -----
 ggplot(plot_df, aes(x=simulation, y=synchrony, fill=site, color=site))+
   geom_bar(stat="identity")+
@@ -109,14 +110,16 @@ ggplot(plot_df, aes(x=simulation, y=synchrony, fill=site, color=site))+
   scale_color_manual(values=site_colors, labels=site_labels, name="")+
   xlab("Simulation Experiment")+
   ylab("Synchrony of Species' Growth Rates")+
-  scale_x_discrete(labels=sim_names)+
+  scale_x_discrete(labels=sim_labels)+
   scale_y_continuous(limits=c(0,1))+
   theme_few()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   guides(fill=FALSE,color=FALSE)
 
 ggsave("../../docs/components/all_sims_supp_noARTR.png", width = 4, height = 4, units="in", dpi=75)
-  
+ggsave("../../docs/components/formatted_figures/formatted_figureS4.png", width = 4, height = 4, units="in", dpi=75)  
+
+
 ##  Calculate percent differences for Results
 # plot_df$control <- rep(plot_df[which(plot_df$simulation=="1All Drivers"),"synchrony"], times=nrow(plot_df)/nsites)
 # plot_df$percent_diff <- with(plot_df, abs(synchrony-control)/((synchrony+control)/2)*100)

### 2016-03-22 docs/INTERACTION_TABLES_MASTER.tex
updates table structure with master tex file

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\usepackage{color}
\usepackage{rotating}
\usepackage{morefloats}

\begin{document}

\maketitle

\begin{center} 
\textbf{ONLINE SUPPORTING INFORMATION}
\end{center}

\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\tableofcontents
\listoffigures
\listoftables
\newpage{}

\subsection{Derivation of limiting case predictions for synchrony of per
capita growth rates (Equations
3-4)}\label{derivation-of-limiting-case-predictions-for-synchrony-of-per-capita-growth-rates-equations-3-4}

Following Loreau and {{de Mazancourt}} (2013) and {{de Mazancourt}} et
al. (2013), we define population growth, ignoring observation error, as

\begin{align}
r_i(t) = r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}u_{ei}(t) + \frac{\sigma_{di}u_{di}(t)}{\sqrt{N_i(t)}} \right]
\end{align}

\noindent where \(N_i(t)\) is the biomass of species \emph{i} in year
\emph{t}, and \(r_i(t)\) is its growth rate in year \emph{t}. \(r_{mi}\)
is species \emph{i}'s intrinsic rate of increase, \(K_i\) is its
carrying capacity, and \(\alpha_{ij}\) is the interspecific competition
coefficient representing the effect of species \emph{j} on species
\emph{i}. Environmental stochasticity is incorporated as
\(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the environmental
variance and \(u_{ei}\) are normal random variables with zero mean and
unit variance that are independent through time but may be correlated
between species. Demographic stochasticity arises from variations in
births and deaths among individuals (e.g., same states, different
fates), and is included in the model as a first order, normal
approximation (Lande et al. 2003, {{de Mazancourt}} et al. 2013).
\(\sigma_{di}^2\) is the demographic variance and \(u_{di}(t)\) are
independent normal variables with zero mean and unit variance.

First-order approximations of the temporal variance of total community
biomass are obtained as follows (Ives 1995, Hughes and Roughgarden 2000,
Ives and Hughes 2002, Loreau and {{de Mazancourt}} 2008, {{de
Mazancourt}} et al. 2013). Let \(\delta N_i(t) = N_i(t) - N^*_i\) denote
the deviation of observed species \emph{i}'s biomass from its
equilibrium value in the community, \(N^*_i\), in the absence of
stochasticity. Equation (S1) can be Taylor expanded around
\(\delta N_i(t) = u_{ei}(t) = u_{di}(t) = u^S_{oi}(t) = 0\) to yield,
after dropping terms of order two and higher,

\begin{equation}
\boldsymbol{\delta}\textbf{N}(t+1) = \textbf{A}\boldsymbol{\delta}\textbf{N}(t) + \textbf{z}(t),
\end{equation}

\noindent where \(\boldsymbol{\delta}\textbf{N}(t)\) is the vector of
deviations of species biomasses from their deterministic equilibrium
value, \textbf{A} is the community matrix, also known as the Jacobian
matrix around the equilibrium, with elements \((a_{ij})_{1<i,j<S}\):

\begin{align}
A_{ij} = 
  \begin{cases}
      1-r_{mi} \frac{N_{i}^*}{K_i}, & i=j \\
      -r_{mi} \frac{N_{i}^*}{K_i} \alpha_{ij}, & i \neq j
    \end{cases}
\end{align}

\noindent and \(\textbf{z}(t)\) is a vector that encapsulates the
effects of environmental and demographic stochasticity whose elements
are

\begin{equation}
z_i(t) = N^*_i \sigma_{ei} u_{ei}(t) + \sqrt{N^*_i} \sigma_{di} u_{di}(t)
\end{equation}

When the system reaches a stationary distribution, the variances and
covariances between species biomass time series are:

\begin{equation}
\big\langle \boldsymbol{\delta}\textbf{N}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle = \textbf{C}^\infty = (\text{cov}(N_i,N_j))_{1<i,j<S} = (\text{cov}(\delta N_i, \delta N_j))_{1<i,j<S}
\end{equation}

\noindent where \(\boldsymbol{\delta}\textbf{N}^T\) is the transpose of
vector \(\boldsymbol{\delta}\textbf{N}\), i.e.~a row vector.

Our assumptions listed above lead to the following correlation structure
of \textbf{z}:

\begin{equation}
\big\langle \textbf{z}(t) \textbf{z}(t)^T \big\rangle = \textbf{B}
\end{equation}

\begin{equation}
\big\langle \textbf{z}(t-s) \textbf{z}(t)^T \big\rangle = 0 \qquad \text{for} \: s>0
\end{equation}

We use Equation S2 to write a dynamical equation for the covariance
\textbf{C}:

\begin{footnotesize}
\begin{align}
\textbf{C}(t+1) =& \big\langle \boldsymbol{\delta}\textbf{N}(t+1) \boldsymbol{\delta}\textbf{N}(t+1)^T \big\rangle \\
=& \textbf{A}\big\langle \boldsymbol{\delta}\textbf{N}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle \textbf{A}^T + \textbf{A}\big\langle \boldsymbol{\delta}\textbf{N}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle + \big\langle \textbf{z}(t) \boldsymbol{\delta}\textbf{N}(t)^T \big\rangle \textbf{A}^T + \big\langle \textbf{z}(t) \textbf{z}(t)^T \big\rangle \\
=& \textbf{A} \textbf{C}(t) \textbf{A}^T + \textbf{Z}_0
\end{align}
\end{footnotesize}

\noindent Taking the limit \(t \rightarrow \infty\) on both sides, we
get:

\begin{equation}
\textbf{C}^\infty = \textbf{A} \textbf{C}^\infty \textbf{A}^T + \textbf{B}
\end{equation}

\noindent where \textbf{A} is as in Equation S3 and \textbf{B} is:

\begin{equation}
B_{ij} = N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})
\end{equation}

\noindent Similarly, \textbf{R} is the variance-covariance matrix for
population growth rates at steady state

\begin{equation}
R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} \sum_{k,l} \alpha_{ik} \alpha_{jl} C_{kl} + \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{\sqrt{N_{i}^* N_{j}^*}}.
\end{equation}

\noindent Then, following the synchrony metric of Loreau and {{de
Mazancourt}} (2008), the sychrony of population biomasses is

\begin{equation}
\phi_N = \frac{\sum_{i,j}C_{ij}}{\left( \sum_i \sqrt{C_{ii}} \right)^2}
\end{equation}

\noindent and the synchrony of per capita growth rates is

\begin{equation}
\phi_R = \frac{\sum_{i,j}R_{ij}}{\left( \sum_i \sqrt{R_{ii}} \right)^2}.
\end{equation}

Synchrony of population biomasses and growth rates emerge from complex
interactions among species' intrinsic growth rates, interspecific
interactions, environmental stochasticity, and demographic
stochasticity. Given these complexities, it is impossible to determine
expected effects of different parameters in a multi-species case. Thus,
we analyze a simplified case where interspecific interactions are zero.
The variance-covariance matrix of population biomasses at steady state
then becomes

\begin{equation}
c_{ij} = \frac{B_{ij}}{1-A_ii A_jj} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
\end{equation}

and the variance-covariance matrix of population growth rates simplifies
to

\begin{equation}
R_{ij} = \frac{1}{1- \frac{r_{mi}r_{mj}}{r_{mi}+r_{mj}}} \left( \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{\sqrt{N_{i}^* N_{j}^*}} \right)
\end{equation}

\noindent Note that synchrony in population sizes and synchrony in
growth rates use weighted factors of the environmental variances and
covariances with species-specific parameters. So, we further assume
that, along with interspecific interactions being zero, all species have
identical growth rates, environmental stochasticity is absent, and all
species have identical demographic variance. This represents a
theoretical limiting case where the community consists of identical
species coexisting in a constant environment where only demographic
stochasticity causes temporal fluctuations. Under such conditions,
synchrony of population biomasses is

\begin{equation}
\phi_N = \frac{1}{\left(\sum_i p_i^{1/2} \right)^2}
\end{equation}

\noindent and synchrony of growth rates is

\begin{equation}
\phi_R = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
\end{equation}

\noindent where \(p_i\) is the average frequency of species \emph{i},
\(p_i = N_i/N_T\). When all species have identical abundances and
\(p_i = 1/S\), where \emph{S} is species richness, the both synchrony
values equal 1/S (Loreau and {{de Mazancourt}} 2008). The prediction
represented in Equation S20 is Equation 3 in the main text, which we
refer to as \(\phi_{R,\mathcal{M}_D}\).

Another limiting case is where only environmental stochasticity is
operating: no interspecific interactions, no demographic stochasticity,
identical intrinsic growth rates, and environmental stochasticity with
same standard deviation for all species. Under such constraints, the
synchrony of of species' biomasses is

\begin{equation}
\phi_N = \sum_{i,j} p_i p_j \text{cov}(u_{ei}, u_{ej})
\end{equation}

\noindent{} and the synchrony of species' growth rates is

\begin{equation}
\phi_R = \frac{\sum_{i,j} p_i p_j \text{cov}(u_{ei}, u_{ej})}{S^2}
\end{equation}

\noindent The prediction represented in Equation S22 is Equation 4 in
the main text, which we refer to as \(\phi_{R,\mathcal{M}_E}\). If we
know the covariance matrix of the species environmental responses, we
can calculate the above expectations directly. Note that the
environmental responses are normalized in the above equations, therefore
the covariances are correlations in the above equations.

\newpage{}

\subsection{Materials and methods
details}\label{materials-and-methods-details}

\subsubsection{Vital rate statistical
models}\label{vital-rate-statistical-models}

We modeled survival probability and growth on individual genets as a
function of genet size, the crowding experienced by the focal genet from
both heterospecific and conspecific genets in its neighborhood
(described below), temporal varation among years, and spatial variation
among quadrat groups. Groups are sets of quadrats located in close
proximity within a pasture or grazing exclosure.

We follow the approach of Chu and Adler (2015) to estimate crowding,
assuming that the crowding experienced by a focal genet depends on
distance to each neighbor genet and the neighbor's size, \emph{u}:

\begin{equation}
w_{ijm,t} = \sum_k e^{-\delta_{jm}d_{ijkm,t}^{2}}u_{km,t}.
\end{equation}

In the above, \(w_{ijm,t}\) is the crowding that genet \emph{i} of
species \emph{j} in year \emph{t} experiences from neighbors of species
\emph{m}. The spatial scale over which species \emph{m} neighbors exert
influence on any genet of species \emph{j} is determined by
\(\delta_{jm}\). The function is applied for all \emph{k} genets of
species \emph{m} that neighbor the focal genet at time \emph{t}, and
\(d_{ijkm,t}\) is the distance between genet \emph{i} in species
\emph{j} and genet \emph{k} in species \emph{m}. When \(k=m\), the
effect is intraspecific crowding. We use regression-specific (survival
and growth) \(\delta\) values estimated by Chu and Adler (2015).

We used logistic regression to model survival probability (\(S\)) of
genet \(i\) from species \(j\) in quadrat group \(g\) from time \(t\) to
\(t+1\):

\begin{align}
\text{logit}(S_{ijg,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jg} + \beta^{S}_{j,t}x_{ij,t} + \boldsymbol{\omega}^{S}_{j} \textbf{w}_{ij,t}
\end{align}

where \(x_{ij,t}\) is the log of genet size, \(\gamma^{S}_{j,t}\) is a
year-specific intercept, \(\beta^{S}_{j,t}\) is the year-specific slope
parameter for size, \(\phi^{S}_{jg}\) is the random effect of quadrat
group location, and \(\omega\) is a vector of per capita interaction
coefficients which determine the impact of crowding, \textbf{w}, by each
species on the focal species.

We modeled genet growth, conditional on survival, in a similar manner:

\begin{align}
\mu_{ijg,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jg} + \beta^{G}_{j,t}\mu_{ij,t} + \boldsymbol{\omega}^{G}_{j} \textbf{w}_{ij,t}
\end{align}

where \(\mu\) is log genet size and all other parameters are as
described for the survival regression. We capture non-constant error
variance in growth by modeling the variance around the growth regression
(\(\varepsilon\)) as a nonlinear function of predicted genet size:

\begin{align}
\varepsilon_{ij,t} = a e^{b \mu_{ijg,t+1}}
\end{align}

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. THerefore, we model recruitment at the
quadrat level: the number of new individuals of species \(j\) in quadrat
\(q\) recruiting at time \(t+1\) as a function of quadrat ``effective
cover'' (\(A'\)) in the previous year (\(t\)). Effective cover is a
mixture of observed cover (\(A\)) in the focal quadrat (\(q\)) and the
mean cover across the entire group (\(\bar{A}\)) of \(Q\) quadrats in
which \(q\) is located:

\begin{equation}
A'_{jq,t} = p_{j}A_{jq,t} + (1-p_{j})\bar{A}_{jQ,t}
\end{equation}

where \(p\) is a mixing fraction between 0 and 1 that is estimated
within the model.

We assume the number of individuals, \(y^{R}\), recruiting at time
\(t+1\) follows a negative binomial distribution:

\begin{equation}
y^{R}_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

where \(\lambda\) is the mean intensity and \(\zeta\) is the size
parameter. We define \(\lambda\) as:

\begin{equation}
\lambda_{jq,t+1} = A'_{jq,t}e^{(\gamma^{R}_{j,t} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} + \omega^{R}\sqrt{A'_{q,t}})}
\end{equation}

where \(A'\) is effective cover (\(\text{cm}^2\)) of species \(j\) in
quadrat \(q\) and all other terms are as in the survival and growth
regressions.

\subsubsection{Multi-species populations
models}\label{multi-species-populations-models}

The individually based model (IBM) is straighforward and described in
the main text, so here focus on the structure of the integral projeciton
model (IPM). We built an environmentally stochastic IPM. Our IPM follows
the specification of Chu and Adler (2015) where the population of
species \emph{j} is a density function \(n(u_{j},t)\) giving the density
of sized-\emph{u} genets at time \emph{t}. Genet size is on the natural
log scale, so that \(n(u_{j},t)du\) is the number of genets whose area
(on the arithmetic scale) is between \(e^{u_{j}}\) and \(e^{u_{j}+du}\).
So, the density function for any size \emph{v} at time \(t+1\) is

\begin{equation}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
\end{equation}

where \(k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})\) is the population kernal
that describes all possible transitions from size \(u\) to \(v\) and
\(\bar{\bold{w_{j}}}\) is a vector of estimates of average crowding
experienced from all other species by a genet of size \(u_j\) and
species \(j\). The integral is evaluated over all possible sizes between
predefined lower (\emph{L}) and upper (\emph{U}) size limits that extend
beyond the range of observed genet sizes.

The population kernal is defined as the joint contributions of survival
(\emph{S}), growth (\emph{G}), and recruitment (\emph{R}):

\begin{equation}
k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
\end{equation}

\noindent{} which means we are calculating growth (\emph{G}) for
individuals that survive (\emph{S}) from time \emph{t} to \emph{t+1} and
adding in newly recruited (\emph{R}) individuals of an average sized
one-year-old genet for the focal species. Our stastical model for
recruitment (\emph{R}, described below) returns the number of new
recruit produced per quadrat. Following previous work, we assume that
fecundity increases linearly with size
(\(R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})\))
to incorporate the recruitment function in the spatially-implicit IPM.

\newpage{}

\subsection{Supporting Figures}\label{supporting-figures}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/figureS1.png}
  \caption{Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IBM (top panels) and the IPM (bottom panels). IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set. IBM density curves come from 100 replicate simulations of 75 iterations each. Synchrony values come from simulations where environmental stochasticity and interspecific interactions are present. The IBM was run on a 5 by 5 meter landscape to reduce the effect of demographic stochasticity.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/formatted_figures/formatted_figureS2.png}
  \caption{Community-wide synchrony of species percent cover from model simulation experiments. Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp. + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp. + No E.S.''). Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation. Error bars represent the upper and lower 95\% quantiles from model simulations.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/formatted_figures/formatted_figureS3.png}
  \caption{Synchrony of species' percent cover for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S. Only'') and when environmental stochasticity is also present removed (``D.S. + E.S.''). The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase. Error bars represent the upper and lower 95\% quantiles from model simulations.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/formatted_figures/formatted_figureS4.png}
  \caption{Community-wide synchrony of species percent cover from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed. Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp. + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp. + No E.S.''). Error bars represent the upper and lower 95\% quantiles from model simulations.}
\end{figure}

\newpage{}

\subsection{Supporting Tables}\label{supporting-tables}

\begin{table}[ht]
\centering
\caption{Comparisons between our analytical predictions and simulation results for synchrony of species' per capita growth rates. Analytical predictions represent two limiting cases where only demographic stochasticity is operating ($\phi_{R,\mathcal{M}_D}$) and where only environmental stochasticity is operating ($\phi_{R,\mathcal{M}_E}$). Simulated synchrony values come from our empirically-based, multi-species population models when simulated under conditions that match the limiting case conditions (e.g., environmental stochasticity and competition removed for $\mathcal{M}_D$).} 
{\normalsize
\begin{tabular}{lrrrr}
  \hline
Site & Predicted $\phi_{R,\mathcal{M}_D}$ & Simulated $\phi_{R,\mathcal{M}_D}$ & Predicted $\phi_{R,\mathcal{M}_E}$ & Simulated $\phi_{R,\mathcal{M}_E}$ \\ 
  \hline
New Mexico & 0.56 & 0.51 & 0.86 & 0.85 \\ 
  Arizona & 0.51 & 0.52 & 0.82 & 0.81 \\ 
  Kansas & 0.35 & 0.39 & 0.62 & 0.63 \\ 
  Montana & 0.31 & 0.33 & 0.47 & 0.49 \\ 
  Idaho & 0.29 & 0.31 & 0.43 & 0.41 \\ 
   \hline
\end{tabular}
}
\end{table}

\newpage{}

\begin{table}[ht]
\centering
\caption{Percent differences of synchrony of per capita growth rates between each removal simulation experiment and the 'All Drivers' simulation.} 
\begin{tabular}{lrrrrr}
  \hline
simulation & 1New Mexico & 2Arizona & 3Kansas & 4Montana & 5Idaho \\ 
  \hline
1No D.S. & 5.00 & 4.94 & 7.13 & 4.42 & 10.10 \\ 
  2No Comp. + No D.S. & 4.25 & 2.08 & 7.61 & 4.39 & 8.10 \\ 
  4No Comp. & 0.98 & 0.06 & 4.84 & 1.01 & 0.11 \\ 
  5No E.S. & 40.14 & 33.20 & 47.84 & 38.52 & 9.28 \\ 
  6No Comp. + No E.S. & 45.95 & 41.48 & 38.21 & 33.07 & 20.33 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\begin{table}[ht]
\centering
\caption{Correlations of species' year random effects for each site by term, where term refers to the random effect on the slope or the intercept.} 
\begin{tabular}{llrrr}
  \hline
site & term & growth & recruit & surv \\ 
  \hline
Arizona & intercept & 0.49 & -0.05 & 0.45 \\ 
  Arizona & slope & 0.68 &  & 0.15 \\ 
  Idaho & intercept & 0.46 & 0.38 & 0.29 \\ 
  Idaho & slope & 0.50 &  & 0.47 \\ 
  Kansas & intercept & 0.47 & 0.29 & 0.24 \\ 
  Kansas & slope & 0.18 &  & 0.16 \\ 
  Montana & intercept & 0.26 & 0.04 & 0.39 \\ 
  Montana & slope & 0.10 &  & 0.01 \\ 
  NewMexico & intercept & 0.54 & -0.15 & 0.64 \\ 
  NewMexico & slope & 0.32 &  & 0.15 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\begin{table}[ht]
\centering
\caption{Average interaction coefficients across vital rates for each community.} 
\begin{tabular}{lrr}
  \hline
site & Interspecific Competition & Intraspecific Competition \\ 
  \hline
Arizona & -0.17 & -0.64 \\ 
  Idaho & -0.04 & -0.71 \\ 
  Kansas & 0.02 & -0.35 \\ 
  Montana & 0.04 & -0.98 \\ 
  NewMexico & -0.12 & -0.62 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\subsubsection{Interaction Coefficients for Vital Rates in Arizona}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for growth regressions in Arizona.} 
\begin{tabular}{rrr}
  \hline
 & BOER & BORO \\ 
  \hline
BOER & -0.0033 & -0.0171 \\ 
  BORO & -0.0050 & -0.0772 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for survival regressions in Arizona.} 
\begin{tabular}{rrr}
  \hline
 & BOER & BORO \\ 
  \hline
BOER & -0.2791 & -0.0816 \\ 
  BORO & -0.0200 & -0.6101 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for recruitment regressions in Arizona.} 
\begin{tabular}{rrr}
  \hline
 & BOER & BORO \\ 
  \hline
BOER & -0.5304 & -0.3063 \\ 
  BORO & -0.5798 & -2.3237 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\subsubsection{Interaction Coefficients for Vital Rates in Idaho}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for growth regressions in Idaho.} 
\begin{tabular}{rrrrr}
  \hline
 & ARTR & HECO & POSE & PSSP \\ 
  \hline
ARTR & -0.0300 & 0.0129 & 0.0117 & 0.0003 \\ 
  HECO & -0.0000 & -0.0285 & 0.0084 & -0.0028 \\ 
  POSE & 0.0001 & 0.0032 & -0.0845 & -0.0043 \\ 
  PSSP & -0.0003 & -0.0027 & -0.0004 & -0.0448 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for survival regressions in Idaho.} 
\begin{tabular}{rrrrr}
  \hline
 & ARTR & HECO & POSE & PSSP \\ 
  \hline
ARTR & -0.0236 & -0.0202 & 0.0314 & 0.0158 \\ 
  HECO & -0.0005 & -1.0206 & -0.0052 & 0.0125 \\ 
  POSE & 0.0001 & 0.0006 & -0.9091 & 0.0121 \\ 
  PSSP & 0.0009 & -0.0013 & 0.0246 & -0.1543 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for recruitment regressions in Idaho.} 
\begin{tabular}{rrrrr}
  \hline
 & ARTR & HECO & POSE & PSSP \\ 
  \hline
ARTR & -0.4054 & 0.2132 & 0.0120 & 0.2389 \\ 
  HECO & -0.5674 & -1.7722 & -0.1476 & -0.3184 \\ 
  POSE & -0.2499 & -0.0328 & -2.2873 & -0.1448 \\ 
  PSSP & -0.2580 & -0.1687 & 0.0365 & -1.7675 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\subsubsection{Interaction Coefficients for Vital Rates in Kansas}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for growth regressions in Kansas.} 
\begin{tabular}{rrrr}
  \hline
 & BOCU & BOHI & SCSC \\ 
  \hline
BOCU & -0.0017 & 0.0011 & 0.0010 \\ 
  BOHI & 0.0002 & -0.0046 & -0.0078 \\ 
  SCSC & -0.0003 & 0.0007 & -0.0124 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for survival regressions in Kansas.} 
\begin{tabular}{rrrr}
  \hline
 & BOCU & BOHI & SCSC \\ 
  \hline
BOCU & -0.1087 & 0.0006 & 0.0010 \\ 
  BOHI & 0.0021 & -0.0403 & -0.0153 \\ 
  SCSC & -0.0064 & -0.0066 & -0.0270 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for recruitment regressions in Kansas.} 
\begin{tabular}{rrrr}
  \hline
 & BOCU & BOHI & SCSC \\ 
  \hline
BOCU & -1.1584 & -0.0242 & -0.0524 \\ 
  BOHI & 0.3387 & -0.8620 & 0.0123 \\ 
  SCSC & 0.0615 & -0.0308 & -0.9543 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\subsubsection{Interaction Coefficients for Vital Rates in Montana}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for growth regressions in Montana.} 
\begin{tabular}{rrrrr}
  \hline
 & BOGR & HECO & PASM & POSE \\ 
  \hline
BOGR & -0.0121 & -0.0002 & -0.0229 & -0.0014 \\ 
  HECO & -0.0023 & -0.0778 & 0.0142 & 0.0061 \\ 
  PASM & -0.0005 & -0.0039 & -0.0601 & -0.0033 \\ 
  POSE & 0.0005 & -0.0101 & -0.0404 & -0.1431 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for survival regressions in Montana.} 
\begin{tabular}{rrrrr}
  \hline
 & BOGR & HECO & PASM & POSE \\ 
  \hline
BOGR & -0.1280 & 0.0211 & 0.1243 & 0.0173 \\ 
  HECO & -0.0043 & -0.5661 & 0.1386 & 0.0013 \\ 
  PASM & 0.0001 & -0.0044 & -1.3046 & 0.0012 \\ 
  POSE & 0.0010 & 0.0085 & 0.1198 & -1.2434 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for recruitment regressions in Montana.} 
\begin{tabular}{rrrrr}
  \hline
 & BOGR & HECO & PASM & POSE \\ 
  \hline
BOGR & -0.8916 & -0.2522 & -0.1334 & 0.0610 \\ 
  HECO & 0.0852 & -1.4191 & -0.0685 & -0.0422 \\ 
  PASM & 0.1252 & 0.3093 & -4.1823 & 0.3032 \\ 
  POSE & 0.1577 & 0.4906 & 0.0709 & -1.7065 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\subsubsection{Interaction Coefficients for Vital Rates in New Mexico}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for growth regressions in New Mexico.} 
\begin{tabular}{rrr}
  \hline
 & BOER & SPFL \\ 
  \hline
BOER & -0.0045 & -0.0013 \\ 
  SPFL & -0.0041 & -0.0176 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for survival regressions in New Mexico.} 
\begin{tabular}{rrr}
  \hline
 & BOER & SPFL \\ 
  \hline
BOER & -0.1575 & -0.0040 \\ 
  SPFL & -0.0120 & -0.2807 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Interaction coefficients for recruitment regressions in New Mexico.} 
\begin{tabular}{rrr}
  \hline
 & BOER & SPFL \\ 
  \hline
BOER & -0.7631 & -0.6198 \\ 
  SPFL & -0.0684 & -2.5220 \\ 
   \hline
\end{tabular}
\end{table}

\newpage{}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
recruitment stage to stabilize grassland coexistence. Ecological
Monographs 85:373--392.

{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
Loreau. 2013. Predicting ecosystem stability from community composition
and biodiversity. Ecology Letters 16:617--625.

Hughes, J. B., and J. Roughgarden. 2000. Species Diversity and Biomass
Stability. The American Naturalist 155:618--627.

Ives, A. 1995. Predicting the Response of Populations to Environmental
Change. Ecology 76:926--941.

Ives, A. R., and J. B. Hughes. 2002. General Relationships Between
Species Diversity and Stability in Competitive Systems. The American
Naturalist 159:388--395.

Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
dynamics in ecology and conservation.

Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
drivers: neutral and nonneutral community dynamics in fluctuating
environments. The American Naturalist 172:E48--E66.

Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
stability: A synthesis of underlying mechanisms. Ecology Letters
16:106--115.

\end{document}

### 2016-03-22 docs/online_supporting_information.Rmd
updates table structure with master tex file

@@
     fig_caption: true
-    keep_tex: false
+    keep_tex: true
 fontsize: 12pt
@@
 We modeled survival probability and growth on individual genets as a function of genet size, the crowding experienced by the focal genet from both heterospecific and conspecific genets in its neighborhood (described below), temporal varation among years, and spatial variation among quadrat groups.
-Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure).
+Groups are sets of quadrats located in close proximity within a pasture or grazing exclosure.
 
@@
     synch_cap <- paste0("Interaction coefficients for ", do_vital, " regressions in ", do_site, ".")
-    print(xtable(site_list, caption = synch_cap),
+    # digitvec <- rep(3,ncol(site_list))
+    print(xtable(site_list, caption = synch_cap, digits = 4),
       caption.placement="top",

### 2016-03-22 inst/examples/demog_stoch_test.tex
updates table structure with master tex file

\documentclass[12pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{todonotes}

\begin{document}

\normalsize

\section{Comparing IPM with approximate demographic stochasticity to
IBM}\label{comparing-ipm-with-approximate-demographic-stochasticity-to-ibm}

Our approach to incorporating demographic stochasticity in integral
projection models (IPMs) is new, and thus requires testing.
Individually-based models (IBMs) inherently include demographic
stochasticity (e.g., ``coin flips'' for survival of a genet from a
Bernoulli trial). IBMs, then, provide the baseline against which to
compare an IPM with our approximation to demogrpahic stochasticity. If
the two models produce similar equilibrium cover and species synchrony,
then we can conclude our approximation is valid. Here we use the Idaho
dataset as a test case, and simulate communities from the IPM with
demographic stochastisticy and the IBM. We ignore temporal variation due
to random year effects to isolate variation from demographic
stochasticity alone.

Figure 1 compares the distribution of percent cover over 2,000
iterations (after a 500 iteration burn in) for each species from the IBM
and the IPM with approximate demographic stochasticity. There are some
small differences, but overall the approximation leads to similar long
term population dynamics when compared to the IBM. The IPM and IBM also
show similar interannual dynamics, as signified by the very similar
synchrony metrics (Table 1). Moreover, species synchrony from models
where only demographic stochasticity is acting matches theoretical
expectations: theory predicts synchrony = 1/\emph{S} (where \emph{S} is
species richness) when demographic stochasticity is much stronger than
the effects of environmental stochasticity and species interactions.
Indeed, the IPM produces synchrony = 0.3 (1/3 with 3 species) when we
exclude random year effects (species interactions are present, but not
influential due to small competition coefficients).

\begin{figure}[htbp]
\centering
\includegraphics{./demog_stoch_test_files/figure-latex/figs.pdf}
\caption{Density estimates of equilibrium cover from 2,000 iterations
from each model.}
\end{figure}

\begin{table}[ht]
\centering
\caption{Species synchrony of percent cover and per capita growth rates.} 
\begin{tabular}{lrr}
  \hline
Model & Cover Sychrony & Growth Rate Synchrony \\ 
  \hline
IPM & 0.33 & 0.34 \\ 
  IBM & 0.34 & 0.33 \\ 
   \hline
\end{tabular}
\end{table}

\end{document}

### 2016-03-22 docs/ms_communitySynchrony.Rmd
adds PBA line edits to main ms

@@
 plots) of different size.
-Indeed, results from an IBM with infinite population size would equal results from the IPM.
+Indeed, results from an IBM with infinite population size would converge on results from the IPM.
 Given computational constraints, the largest landscape we simulate is a 25 $\text{m}^2$ plot.
@@
 Synchrony was calculated over the 100 time steps for each no extinction run within a model experiment.
-To explore the effect of demographic stochasticity in different sized populations, we ran simulations (1) and (6) on plot sizes of 1, 2, 3, 4, and 5 square meters.
-All other IBM simulations were run on a 5 square meter landscape.
+To explore the effect of demographic stochasticity in different sized populations, we ran simulations (1) and (6) on plot sizes of 1, 4, 9, 16, and 25 $\text{m}^2$.
+All other IBM simulations were run on a 25 $\text{m}^2$ landscape.
 
@@
 2).
-Synchrony from simulations with demographic and environmental stochasticity decreased as the simulated area decreased, representing a shift from synchrony being driven by environmental responses (e.g., $\mathcal{M}_E$) to being driven by a combination of demographic and environmental stochasticity (e.g., between $\mathcal{M}_E$ and $\mathcal{M}_D$).
+In the largest, 25 $\text{m}^2$ plots, synchrony was driven by environmental stochasticity (e.g., $\mathcal{M}_E$).
+At 1 $\text{m}^2$, synchrony reflected demographic stochasticity and environmental stochasticity (e.g., between $\mathcal{M}_E$ and $\mathcal{M}_D$).
+
 For all five communities, the synchrony of species' growth rates in monoculture almost perfectly matched species synchrony in polyculture (Fig. 3).
@@
 
-Our analysis of how demographic stochasticity affects synchrony across population sizes also demonstrates that synchrony depends on the observation area.
+Our analysis of how demographic stochasticity affects synchrony demonstrates that synchrony depends on the observation area.
 As the observation area increases, population size increases and the desynchronizing effect of demographic stochasticity lessens (Fig. 2).
@@
 However, our results do contrast with a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives species synchrony in grasslands [@Gross2014].
-The apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
+The apparent inconsistency between our results and those of @Gross2014 may be explained by the differences between our studies.
 Gross et al.'s results are based on rapidly assembling experimental communities that are, at most, 11 years old.
 The natural, relatively undisturbed communities we studied consist of species that have co-occurred for decades [@Chu2015] and represent a much later stage of community assembly.
-In theory, species interactions should weaken through time as community assembly processes [@Kokkoris1999], meaning that in biodiversity-ecosystem functioning experiments the processes driving competitive exclusion are still operating.
+In theory, species interactions should weaken through time as community assembly proceeds [@Kokkoris1999], meaning that in biodiversity-ecosystem functioning experiments the processes driving competitive exclusion are still operating.
 Such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
@@
 A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
-Such a conclusion is only half correct.
+Such a conclusion misses an important subtlety.
 The perennial grasses we studied do have similar responses to the environment (Table S2), which will tend to synchronize dynamics.
@@
 In agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
-Such a conclusion is in contrast to a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives community synchrony [@Gross2014].
+This result contrasts with a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives community synchrony [@Gross2014].
 Recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
@@
 Error bars represent the upper and lower 95\% quantiles from model simulations.
+All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
 Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}

### 2016-03-22 docs/online_supporting_information.Rmd
breaks apart interaction SI table by vital rate

@@
 df$interintra <- substr(sites, start = nchar(sites), stop = nchar(sites))
-avg_interintra <- ddply(df, .(site, interintra), summarise,
-                        avg_val = mean(value))
-avg_interintra[which(avg_interintra$interintra==1),"interintra"] <- "Interspecific Competition"
-avg_interintra[which(avg_interintra$interintra==2),"interintra"] <- "Intraspecific Competition"
-interintra_table <- dcast(avg_interintra, site~interintra, value.var = "avg_val")
 
-comp_cap <- "Average interaction coefficients across vital rates for each community."
-print(xtable(interintra_table, caption = comp_cap),
+dftable <- dcast(df, site+interintra~vitalrate, value.var = "value")
+colnames(dftable) <- c("Site", "Interaction Type", "Growth", "Recruitment", "Survival")
+dftable[which(dftable$`Interaction Type`==1),"Interaction Type"] <- "Interspecific"
+dftable[which(dftable$`Interaction Type`==2),"Interaction Type"] <- "Intraspecific"
+# avg_interintra <- ddply(df, .(site, interintra), summarise,
+#                         avg_val = mean(value))
+# avg_interintra[which(avg_interintra$interintra==1),"interintra"] <- "Interspecific Competition"
+# avg_interintra[which(avg_interintra$interintra==2),"interintra"] <- "Intraspecific Competition"
+# interintra_table <- dcast(avg_interintra, site~interintra, value.var = "avg_val")
+
+comp_cap <- "Average interaction coefficients for each vital rate for each community."
+print(xtable(dftable, caption = comp_cap, digits=4),
   caption.placement="top",

### 2016-03-23 docs/ms_communitySynchrony.Rmd
updates ms with text on percent cover synchrony and abundances at landscape sizes

@@
 1), with an average 6% increase over synchrony from IBM simulations on a five meter area.
+
 The desynchronizing effect of demographic stochasticity, which increases as population size decreases, modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
@@
 At 1 $\text{m}^2$, synchrony reflected demographic stochasticity and environmental stochasticity (e.g., between $\mathcal{M}_E$ and $\mathcal{M}_D$).
+For context, population sizes increased from an average of 17 individuals per community in 1 $\text{m}^2$ IBM simulations to an average of 357 individuals per community in 25 $\text{m}^2$ IBM simulations.
 
 For all five communities, the synchrony of species' growth rates in monoculture almost perfectly matched species synchrony in polyculture (Fig. 3).
-Results for synchrony of percent cover are qualitatively similar (Figs. S2-S3) but more variable, as expected by theory [@Loreau2008a].
+Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Figs. S2-S3).
 
@@
 Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, though not perfectly, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.
+We also found that analyses based on synchrony of species' percent cover were uninformative (Online Supporting Information), implying that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
 

### 2016-03-23 devel/plot_abund_v_landscape.R

@@ -3,6 +3,9 @@
 ##  From IBM simulations
 ##
 
+library(plyr)
+library(reshape2)
+
 rm(list=ls())
 
 totSims <- 50      # number of simulations per site 
@@ -37,7 +40,7 @@ rm(stringlist)
 
 
 output_df <- data.frame(experiment=NA, site=NA, expansion=NA, 
-                        run=NA, meanabund=NA)
+                        run=NA, meannum=NA)
 
 for(do_exp in unique(files_df$experiment)){
   exp_files <- files_df[which(files_df$experiment==do_exp),"filename"]
@@ -49,7 +52,7 @@ for(do_exp in unique(files_df$experiment)){
     for(i in 1:nsims){ # loop over plot size simulations
       tmp <- readRDS(as.character(tmp_files[i]))
       tmp <- as.data.frame(tmp, row.names = c(1:nrow(tmp)))
-      cover.columns <- grep("Cov.", colnames(tmp))
+      plantnum.columns <- grep("N.", colnames(tmp))
       
       # Get rid of ARTR column for Idaho, for now
       # if(do_site=="Idaho"){cover.columns <- cover.columns[2:length(cover.columns)]}
@@ -60,14 +63,14 @@ for(do_exp in unique(files_df$experiment)){
       
       for(jj in 1:nrow(end.tmp)){ # loop over simulations within plot size
         tmpjj <- subset(end.tmp, run==jj)
-        endcover <- tmpjj[,cover.columns]
+        endplantnum <- tmpjj[,plantnum.columns]
         
         # Set extinct to "yes" if no runs coexist
-        if(nrow(endcover)==0){extinct[jj]<-"yes"}
+        if(nrow(endplantnum)==0){extinct[jj]<-"yes"}
         
         # Nested if/then for coexistence runs
-        if(nrow(endcover)>0){
-          if(length(which(endcover==0))>0){extinct[jj]<-"yes"} 
+        if(nrow(endplantnum)>0){
+          if(length(which(endplantnum==0))>0){extinct[jj]<-"yes"} 
         } # end if/then for extinction flagging
         
       } # end loop for sims within plot size
@@ -88,15 +91,15 @@ for(do_exp in unique(files_df$experiment)){
         
         for(k in runs){ # loop over coexistence runs
           tmpsynch <- subset(tmp4synch, run==k)
-          ts.tmp <- tmpsynch[burn.in:totT,cover.columns]
+          ts.tmp <- tmpsynch[burn.in:totT,plantnum.columns]
[…]

### 2016-03-23 docs/ms_communitySynchrony_preprintVersion.Rmd
adds preprint version

--- 2016-03-23 docs/ms_communitySynchrony.Rmd
@@
 ---
-layout: 12pt
+layout: 10pt
 header-includes:
    - \usepackage{lineno}
-   - \linenumbers
    - \usepackage{setspace}
-   - \usepackage{todonotes}
-   - \doublespacing
-   - \usepackage{color}
    - \usepackage{rotating}
@@
     keep_tex: false
-fontsize: 12pt
+fontsize: 10pt
 geometry: margin=1in
@@
 
-\vspace{2em}
-
-\textbf{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
-
-\vspace{2em}
-
-\textbf{Authorship}: ATT, CdM, and PBA designed the research.
-CdM and ML designed and analyzed the theoretical model.
-ATT and PBA performed the empirical analysis.
-All authors contributed to the interpretation and presentation of results.
-ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
-
-\textbf{Running Title}: Drivers of species synchrony in natural plant communities
-
-\textbf{Article Type}: Letter
-
-\textbf{Number of Words}: Abstract = ; Main Text =
-
-\textbf{Number of References}: 46
-
-\textbf{Number of Tables and Figures}:  3 Tables;  3 Figures
-
-\textbf{Corresponding Author}: \\
-\hspace{2em}Andrew Tredennick \\  
-\hspace{2em}Department of Wildland Resources and the Ecology Center\\
-\hspace{2em}Utah State University \\
-\hspace{2em}5230 Old Main Hill \\
-\hspace{2em}Logan, Utah 84322 USA \\
-\hspace{2em}Phone: +1-970-443-1599 \\
-\hspace{2em}Fax: +1-435-797-3796 \\
-\hspace{2em}Email: atredenn@gmail.com
-
 \end{singlespace}
@@
 \setcounter{footnote}{0}
-
-\newpage{}
 
@@
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+\\
+\\
+\textit{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
 \end{abstract}
 
-\newpage{}
 <!-- THIS HOW TO COMMENT-OUT -->
 \setlength{\parindent}{5ex}
+\twocolumn{}
 
@@
 
+\begin{tiny}
 \begin{align}
@@
 \end{align}
+\end{tiny}
 
@@
 
+\textbf{Authorship}: ATT, CdM, and PBA designed the research.
+CdM and ML designed and analyzed the theoretical model.
+ATT and PBA performed the empirical analysis.
+All authors contributed to the interpretation and presentation of results.
+ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
+
 \setlength{\parindent}{0ex}
 \newpage{}
+
+\onecolumn{}
 

### 2016-03-23 docs/ms_communitySynchrony.Rmd
SI tex with good tables; updates SI table ref in main text

@@
 
-\textbf{Number of Tables and Figures}:  3 Tables;  3 Figures
+\textbf{Number of Tables and Figures}:  2 Tables;  3 Figures
 
@@
 
-Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Table S1).
+Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Fig. 1 and Table S1).
 The correlation between our analytical predictions and simulated synchrony was 0.97 for $\phi_{R,\mathcal{M}_D}$ and 0.997 for $\phi_{R,\mathcal{M}_E}$.
-The largest difference between predicted and simulated synchrony was 0.05 in New Mexico for $\phi_{R,\mathcal{M}_D}$ (Table 3).
+The largest difference between predicted and simulated synchrony was 0.05 in New Mexico for $\phi_{R,\mathcal{M}_D}$ (Table S1).
 
@@
 Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, though not perfectly, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.
-We also found that analyses based on synchrony of species' percent cover were uninformative (Online Supporting Information), implying that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
+We also found that analyses based on synchrony of species' percent cover were uninformative (Figs.
+S2-S3), implying that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
 
@@
 However, that result appears to be an artifact. 
-Removing environmental variation results in a negative invasion growth rate for *A.
-tripartita*.
-Although we only analyzed IBM runs in which *A.
-tripartita* had not yet gone extinct, it was at much lower abundance than in the other simulation runs.
+Removing environmental variation results in a negative invasion growth rate for *A. tripartita*. 
+Although we only analyzed IBM runs in which *A. tripartita* had not yet gone extinct, it was at much lower abundance than in the other simulation runs.
 When we removed *A.
@@
 Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.
-*A. tripartita* has a facilitative effect on each grass species across all vital rates, except for a small competitive effect on *H. comata*'s survival probability (Tables S4, S9, S14).
-At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables S4, S9, S14).
+*A. tripartita* has a facilitative effect on each grass species across all vital rates, except for a small competitive effect on *H. comata*'s survival probability (Tables S8-S10).
+At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables S8-S10).
 We know synchrony is affected by interspecific competition [@Loreau2008a], but how facilitative effects manifest themselves is unknown.

### 2016-03-23 docs/online_supporting_information.tex
SI tex with good tables; updates SI table ref in main text

--- 2016-03-22 docs/INTERACTION_TABLES_MASTER.tex
@@
 \centering
-\caption{Average interaction coefficients across vital rates for each community.} 
-\begin{tabular}{lrr}
+\caption{Average interaction coefficients for each vital rate for each community.} 
+\begin{tabular}{llrrr}
   \hline
-site & Interspecific Competition & Intraspecific Competition \\ 
+Site & Interaction Type & Growth & Recruitment & Survival \\ 
   \hline
-Arizona & -0.17 & -0.64 \\ 
-  Idaho & -0.04 & -0.71 \\ 
-  Kansas & 0.02 & -0.35 \\ 
-  Montana & 0.04 & -0.98 \\ 
-  NewMexico & -0.12 & -0.62 \\ 
+Arizona & Interspecific & -0.0111 & -0.4430 & -0.0508 \\ 
+  Arizona & Intraspecific & -0.0402 & -1.4271 & -0.4446 \\ 
+  Idaho & Interspecific & 0.0022 & -0.1156 & 0.0059 \\ 
+  Idaho & Intraspecific & -0.0469 & -1.5581 & -0.5269 \\ 
+  Kansas & Interspecific & -0.0008 & 0.0509 & -0.0041 \\ 
+  Kansas & Intraspecific & -0.0062 & -0.9916 & -0.0587 \\ 
+  Montana & Interspecific & -0.0053 & 0.0922 & 0.0354 \\ 
+  Montana & Intraspecific & -0.0733 & -2.0499 & -0.8105 \\ 
+  NewMexico & Interspecific & -0.0027 & -0.3441 & -0.0080 \\ 
+  NewMexico & Intraspecific & -0.0111 & -1.6425 & -0.2191 \\ 
    \hline

### 2016-03-23 devel/formula.tex
adds main text word count and tex file

% Created by tikzDevice version 0.10.1 on 2016-02-26 15:02:09
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (108.08,212.04) circle (  3.20);

\path[fill=fillColor] ( 84.15, 75.92) circle (  3.20);

\path[fill=fillColor] ( 93.19, 82.20) circle (  3.20);

\path[fill=fillColor] ( 94.94,165.97) circle (  3.20);

\path[fill=fillColor] (114.56,237.18) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] ( 85.24,212.04) circle (  3.20);

\path[fill=fillColor] ( 72.87, 75.92) circle (  3.20);

\path[fill=fillColor] ( 69.23, 82.20) circle (  3.20);

\path[fill=fillColor] ( 72.87,165.97) circle (  3.20);

\path[fill=fillColor] ( 88.87,237.18) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 69.23, 89.02) --
	( 69.48, 90.92) --
	( 69.72, 92.82) --
	( 69.97, 94.72) --
	( 70.22, 96.62) --
	( 70.47, 98.53) --
	( 70.72,100.43) --
	( 70.97,102.33) --
	( 71.22,104.23) --
	( 71.46,106.13) --
	( 71.71,108.03) --
	( 71.96,109.93) --
	( 72.21,111.83) --
	( 72.46,113.74) --
	( 72.71,115.64) --
	( 72.96,117.54) --
	( 73.21,119.44) --
	( 73.45,121.34) --
	( 73.70,123.24) --
	( 73.95,125.14) --
	( 74.20,127.05) --
	( 74.45,128.95) --
	( 74.70,130.85) --
	( 74.95,132.75) --
	( 75.20,134.65) --
	( 75.44,136.55) --
	( 75.69,138.45) --
	( 75.94,140.35) --
	( 76.19,142.26) --
	( 76.44,144.16) --
	( 76.69,146.06) --
	( 76.94,147.96) --
	( 77.19,149.86) --
	( 77.43,151.76) --
	( 77.68,153.66) --
	( 77.93,155.56) --
	( 78.18,157.47) --
	( 78.43,159.37) --
	( 78.68,161.27) --
	( 78.93,163.17) --
	( 79.17,165.07) --
	( 79.42,166.97) --
	( 79.67,168.87) --
	( 79.92,170.77) --
	( 80.17,172.68) --
	( 80.42,174.58) --
	( 80.67,176.48) --
	( 80.92,178.38) --
	( 81.16,180.28) --
	( 81.41,182.18) --
	( 81.66,184.08) --
	( 81.91,185.98) --
	( 82.16,187.89) --
	( 82.41,189.79) --
	( 82.66,191.69) --
	( 82.91,193.59) --
	( 83.15,195.49) --
	( 83.40,197.39) --
	( 83.65,199.29) --
	( 83.90,201.19) --
	( 84.15,203.10) --
	( 84.40,205.00) --
	( 84.65,206.90) --
	( 84.89,208.80) --
	( 85.14,210.70) --
	( 85.39,212.60) --
	( 85.64,214.50) --
	( 85.89,216.40) --
	( 86.14,218.31) --
	( 86.39,220.21) --
	( 86.64,222.11) --
	( 86.88,224.01) --
	( 87.13,225.91) --
	( 87.38,227.81) --
	( 87.63,229.71) --
	( 87.88,231.61) --
	( 88.13,233.52) --
	( 88.38,235.42) --
	( 88.63,237.32) --
	( 88.87,239.22);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 84.15, 71.19) --
	( 84.53, 73.36) --
	( 84.92, 75.52) --
	( 85.30, 77.69) --
	( 85.69, 79.86) --
	( 86.07, 82.02) --
	( 86.46, 84.19) --
	( 86.84, 86.35) --
	( 87.23, 88.52) --
	( 87.61, 90.68) --
	( 88.00, 92.85) --
	( 88.38, 95.02) --
	( 88.77, 97.18) --
	( 89.15, 99.35) --
	( 89.54,101.51) --
	( 89.92,103.68) --
	( 90.31,105.84) --
	( 90.69,108.01) --
	( 91.08,110.18) --
	( 91.46,112.34) --
	( 91.85,114.51) --
	( 92.23,116.67) --
	( 92.62,118.84) --
	( 93.00,121.00) --
	( 93.39,123.17) --
	( 93.77,125.34) --
	( 94.16,127.50) --
	( 94.54,129.67) --
	( 94.93,131.83) --
	( 95.31,134.00) --
	( 95.70,136.16) --
	( 96.08,138.33) --
	( 96.46,140.50) --
	( 96.85,142.66) --
	( 97.23,144.83) --
	( 97.62,146.99) --
	( 98.00,149.16) --
	( 98.39,151.33) --
	( 98.77,153.49) --
	( 99.16,155.66) --
	( 99.54,157.82) --
	( 99.93,159.99) --
	(100.31,162.15) --
	(100.70,164.32) --
	(101.08,166.49) --
	(101.47,168.65) --
	(101.85,170.82) --
	(102.24,172.98) --
	(102.62,175.15) --
	(103.01,177.31) --
	(103.39,179.48) --
	(103.78,181.65) --
	(104.16,183.81) --
	(104.55,185.98) --
	(104.93,188.14) --
	(105.32,190.31) --
	(105.70,192.47) --
	(106.09,194.64) --
	(106.47,196.81) --
	(106.86,198.97) --
	(107.24,201.14) --
	(107.63,203.30) --
	(108.01,205.47) --
	(108.40,207.63) --
	(108.78,209.80) --
	(109.17,211.97) --
	(109.55,214.13) --
	(109.94,216.30) --
	(110.32,218.46) --
	(110.71,220.63) --
	(111.09,222.80) --
	(111.48,224.96) --
	(111.86,227.13) --
	(112.25,229.29) --
	(112.63,231.46) --
	(113.02,233.62) --
	(113.40,235.79) --
	(113.79,237.96) --
	(114.17,240.12) --
	(114.56,242.29);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.49, 34.03) -- (124.53,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.49, 34.03) rectangle (124.53,264.40);
\end{scope}
\begin{scope}
\path[clip] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (190.57,230.89) circle (  3.20);

\path[fill=fillColor] (162.76,157.59) circle (  3.20);

\path[fill=fillColor] (176.06,119.90) circle (  3.20);

\path[fill=fillColor] (165.59,155.50) circle (  3.20);

\path[fill=fillColor] (193.62,224.61) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (168.29,230.89) circle (  3.20);

\path[fill=fillColor] (155.92,157.59) circle (  3.20);

\path[fill=fillColor] (152.28,119.90) circle (  3.20);

\path[fill=fillColor] (155.92,155.50) circle (  3.20);

\path[fill=fillColor] (171.93,224.61) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (152.28,131.37) --
	(152.53,132.71) --
	(152.78,134.05) --
	(153.03,135.40) --
	(153.28,136.74) --
	(153.53,138.08) --
	(153.78,139.42) --
	(154.02,140.76) --
	(154.27,142.11) --
	(154.52,143.45) --
	(154.77,144.79) --
	(155.02,146.13) --
	(155.27,147.47) --
	(155.52,148.81) --
	(155.77,150.16) --
	(156.01,151.50) --
	(156.26,152.84) --
	(156.51,154.18) --
	(156.76,155.52) --
	(157.01,156.87) --
	(157.26,158.21) --
	(157.51,159.55) --
	(157.75,160.89) --
	(158.00,162.23) --
	(158.25,163.58) --
	(158.50,164.92) --
	(158.75,166.26) --
	(159.00,167.60) --
	(159.25,168.94) --
	(159.50,170.28) --
	(159.74,171.63) --
	(159.99,172.97) --
	(160.24,174.31) --
	(160.49,175.65) --
	(160.74,176.99) --
	(160.99,178.34) --
	(161.24,179.68) --
	(161.49,181.02) --
	(161.73,182.36) --
	(161.98,183.70) --
	(162.23,185.04) --
	(162.48,186.39) --
	(162.73,187.73) --
	(162.98,189.07) --
	(163.23,190.41) --
	(163.47,191.75) --
	(163.72,193.10) --
	(163.97,194.44) --
	(164.22,195.78) --
	(164.47,197.12) --
	(164.72,198.46) --
	(164.97,199.80) --
	(165.22,201.15) --
	(165.46,202.49) --
	(165.71,203.83) --
	(165.96,205.17) --
	(166.21,206.51) --
	(166.46,207.86) --
	(166.71,209.20) --
	(166.96,210.54) --
	(167.21,211.88) --
	(167.45,213.22) --
	(167.70,214.57) --
	(167.95,215.91) --
	(168.20,217.25) --
	(168.45,218.59) --
	(168.70,219.93) --
	(168.95,221.27) --
	(169.19,222.62) --
	(169.44,223.96) --
	(169.69,225.30) --
	(169.94,226.64) --
	(170.19,227.98) --
	(170.44,229.33) --
	(170.69,230.67) --
	(170.94,232.01) --
	(171.18,233.35) --
	(171.43,234.69) --
	(171.68,236.03) --
	(171.93,237.38);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (162.76,138.10) --
	(163.15,139.13) --
	(163.54,140.17) --
	(163.94,141.20) --
	(164.33,142.23) --
	(164.72,143.27) --
	(165.11,144.30) --
	(165.50,145.34) --
	(165.89,146.37) --
	(166.28,147.40) --
	(166.67,148.44) --
	(167.06,149.47) --
	(167.45,150.51) --
	(167.84,151.54) --
	(168.23,152.57) --
	(168.62,153.61) --
	(169.01,154.64) --
	(169.40,155.68) --
	(169.79,156.71) --
	(170.18,157.74) --
	(170.57,158.78) --
	(170.96,159.81) --
	(171.36,160.85) --
	(171.75,161.88) --
	(172.14,162.91) --
	(172.53,163.95) --
	(172.92,164.98) --
	(173.31,166.02) --
	(173.70,167.05) --
	(174.09,168.08) --
	(174.48,169.12) --
	(174.87,170.15) --
	(175.26,171.19) --
	(175.65,172.22) --
	(176.04,173.25) --
	(176.43,174.29) --
	(176.82,175.32) --
	(177.21,176.36) --
	(177.60,177.39) --
	(177.99,178.42) --
	(178.38,179.46) --
	(178.78,180.49) --
	(179.17,181.53) --
	(179.56,182.56) --
	(179.95,183.59) --
	(180.34,184.63) --
	(180.73,185.66) --
	(181.12,186.70) --
	(181.51,187.73) --
	(181.90,188.76) --
	(182.29,189.80) --
	(182.68,190.83) --
	(183.07,191.87) --
	(183.46,192.90) --
	(183.85,193.93) --
	(184.24,194.97) --
	(184.63,196.00) --
	(185.02,197.04) --
	(185.41,198.07) --
	(185.80,199.10) --
	(186.20,200.14) --
	(186.59,201.17) --
	(186.98,202.21) --
	(187.37,203.24) --
	(187.76,204.27) --
	(188.15,205.31) --
	(188.54,206.34) --
	(188.93,207.38) --
	(189.32,208.41) --
	(189.71,209.44) --
	(190.10,210.48) --
	(190.49,211.51) --
	(190.88,212.55) --
	(191.27,213.58) --
	(191.66,214.61) --
	(192.05,215.65) --
	(192.44,216.68) --
	(192.83,217.72) --
	(193.23,218.75) --
	(193.62,219.78);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (127.54, 34.03) -- (207.59,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.54, 34.03) rectangle (207.59,264.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49,264.40) rectangle (124.53,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51,267.41) {cover};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54,264.40) rectangle (207.59,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56,267.41) {growth rate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 41.20) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 93.56) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,145.91) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,198.27) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,250.62) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 44.51) --
	( 44.49, 44.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 96.86) --
	( 44.49, 96.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,149.22) --
	( 44.49,149.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,201.57) --
	( 44.49,201.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,253.93) --
	( 44.49,253.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.12, 29.77) --
	( 48.12, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.32, 29.77) --
	( 66.32, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.51, 29.77) --
	( 84.51, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.70, 29.77) --
	(102.70, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.89, 29.77) --
	(120.89, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 48.12, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 66.32, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (102.70, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (120.89, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] (131.18, 29.77) --
	(131.18, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.37, 29.77) --
	(149.37, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.56, 29.77) --
	(167.56, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.76, 29.77) --
	(185.76, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.95, 29.77) --
	(203.95, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (131.18, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (149.37, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (185.76, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (203.95, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (126.04,  9.03) {Predicted Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 17.30,149.22) {Observed Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (216.45,125.38) rectangle (268.17,173.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (220.72,162.17) {\bfseries predtype};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,144.10) rectangle (235.18,158.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (227.95,151.33) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,151.33) -- (233.73,151.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,129.64) rectangle (235.18,144.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (227.95,136.87) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,136.87) -- (233.73,136.87);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (236.98,148.02) {$\mathcal{M}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (236.98,133.57) {NA};
\end{scope}
\end{tikzpicture}
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (108.08,212.04) circle (  3.20);

\path[fill=fillColor] ( 84.15, 75.92) circle (  3.20);

\path[fill=fillColor] ( 93.19, 82.20) circle (  3.20);

\path[fill=fillColor] ( 94.94,165.97) circle (  3.20);

\path[fill=fillColor] (114.56,237.18) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] ( 85.24,212.04) circle (  3.20);

\path[fill=fillColor] ( 72.87, 75.92) circle (  3.20);

\path[fill=fillColor] ( 69.23, 82.20) circle (  3.20);

\path[fill=fillColor] ( 72.87,165.97) circle (  3.20);

\path[fill=fillColor] ( 88.87,237.18) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 69.23, 89.02) --
	( 69.48, 90.92) --
	( 69.72, 92.82) --
	( 69.97, 94.72) --
	( 70.22, 96.62) --
	( 70.47, 98.53) --
	( 70.72,100.43) --
	( 70.97,102.33) --
	( 71.22,104.23) --
	( 71.46,106.13) --
	( 71.71,108.03) --
	( 71.96,109.93) --
	( 72.21,111.83) --
	( 72.46,113.74) --
	( 72.71,115.64) --
	( 72.96,117.54) --
	( 73.21,119.44) --
	( 73.45,121.34) --
	( 73.70,123.24) --
	( 73.95,125.14) --
	( 74.20,127.05) --
	( 74.45,128.95) --
	( 74.70,130.85) --
	( 74.95,132.75) --
	( 75.20,134.65) --
	( 75.44,136.55) --
	( 75.69,138.45) --
	( 75.94,140.35) --
	( 76.19,142.26) --
	( 76.44,144.16) --
	( 76.69,146.06) --
	( 76.94,147.96) --
	( 77.19,149.86) --
	( 77.43,151.76) --
	( 77.68,153.66) --
	( 77.93,155.56) --
	( 78.18,157.47) --
	( 78.43,159.37) --
	( 78.68,161.27) --
	( 78.93,163.17) --
	( 79.17,165.07) --
	( 79.42,166.97) --
	( 79.67,168.87) --
	( 79.92,170.77) --
	( 80.17,172.68) --
	( 80.42,174.58) --
	( 80.67,176.48) --
	( 80.92,178.38) --
	( 81.16,180.28) --
	( 81.41,182.18) --
	( 81.66,184.08) --
	( 81.91,185.98) --
	( 82.16,187.89) --
	( 82.41,189.79) --
	( 82.66,191.69) --
	( 82.91,193.59) --
	( 83.15,195.49) --
	( 83.40,197.39) --
	( 83.65,199.29) --
	( 83.90,201.19) --
	( 84.15,203.10) --
	( 84.40,205.00) --
	( 84.65,206.90) --
	( 84.89,208.80) --
	( 85.14,210.70) --
	( 85.39,212.60) --
	( 85.64,214.50) --
	( 85.89,216.40) --
	( 86.14,218.31) --
	( 86.39,220.21) --
	( 86.64,222.11) --
	( 86.88,224.01) --
	( 87.13,225.91) --
	( 87.38,227.81) --
	( 87.63,229.71) --
	( 87.88,231.61) --
	( 88.13,233.52) --
	( 88.38,235.42) --
	( 88.63,237.32) --
	( 88.87,239.22);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 84.15, 71.19) --
	( 84.53, 73.36) --
	( 84.92, 75.52) --
	( 85.30, 77.69) --
	( 85.69, 79.86) --
	( 86.07, 82.02) --
	( 86.46, 84.19) --
	( 86.84, 86.35) --
	( 87.23, 88.52) --
	( 87.61, 90.68) --
	( 88.00, 92.85) --
	( 88.38, 95.02) --
	( 88.77, 97.18) --
	( 89.15, 99.35) --
	( 89.54,101.51) --
	( 89.92,103.68) --
	( 90.31,105.84) --
	( 90.69,108.01) --
	( 91.08,110.18) --
	( 91.46,112.34) --
	( 91.85,114.51) --
	( 92.23,116.67) --
	( 92.62,118.84) --
	( 93.00,121.00) --
	( 93.39,123.17) --
	( 93.77,125.34) --
	( 94.16,127.50) --
	( 94.54,129.67) --
	( 94.93,131.83) --
	( 95.31,134.00) --
	( 95.70,136.16) --
	( 96.08,138.33) --
	( 96.46,140.50) --
	( 96.85,142.66) --
	( 97.23,144.83) --
	( 97.62,146.99) --
	( 98.00,149.16) --
	( 98.39,151.33) --
	( 98.77,153.49) --
	( 99.16,155.66) --
	( 99.54,157.82) --
	( 99.93,159.99) --
	(100.31,162.15) --
	(100.70,164.32) --
	(101.08,166.49) --
	(101.47,168.65) --
	(101.85,170.82) --
	(102.24,172.98) --
	(102.62,175.15) --
	(103.01,177.31) --
	(103.39,179.48) --
	(103.78,181.65) --
	(104.16,183.81) --
	(104.55,185.98) --
	(104.93,188.14) --
	(105.32,190.31) --
	(105.70,192.47) --
	(106.09,194.64) --
	(106.47,196.81) --
	(106.86,198.97) --
	(107.24,201.14) --
	(107.63,203.30) --
	(108.01,205.47) --
	(108.40,207.63) --
	(108.78,209.80) --
	(109.17,211.97) --
	(109.55,214.13) --
	(109.94,216.30) --
	(110.32,218.46) --
	(110.71,220.63) --
	(111.09,222.80) --
	(111.48,224.96) --
	(111.86,227.13) --
	(112.25,229.29) --
	(112.63,231.46) --
	(113.02,233.62) --
	(113.40,235.79) --
	(113.79,237.96) --
	(114.17,240.12) --
	(114.56,242.29);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.49, 34.03) -- (124.53,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.49, 34.03) rectangle (124.53,264.40);
\end{scope}
\begin{scope}
\path[clip] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (190.57,230.89) circle (  3.20);

\path[fill=fillColor] (162.76,157.59) circle (  3.20);

\path[fill=fillColor] (176.06,119.90) circle (  3.20);

\path[fill=fillColor] (165.59,155.50) circle (  3.20);

\path[fill=fillColor] (193.62,224.61) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (168.29,230.89) circle (  3.20);

\path[fill=fillColor] (155.92,157.59) circle (  3.20);

\path[fill=fillColor] (152.28,119.90) circle (  3.20);

\path[fill=fillColor] (155.92,155.50) circle (  3.20);

\path[fill=fillColor] (171.93,224.61) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (152.28,131.37) --
	(152.53,132.71) --
	(152.78,134.05) --
	(153.03,135.40) --
	(153.28,136.74) --
	(153.53,138.08) --
	(153.78,139.42) --
	(154.02,140.76) --
	(154.27,142.11) --
	(154.52,143.45) --
	(154.77,144.79) --
	(155.02,146.13) --
	(155.27,147.47) --
	(155.52,148.81) --
	(155.77,150.16) --
	(156.01,151.50) --
	(156.26,152.84) --
	(156.51,154.18) --
	(156.76,155.52) --
	(157.01,156.87) --
	(157.26,158.21) --
	(157.51,159.55) --
	(157.75,160.89) --
	(158.00,162.23) --
	(158.25,163.58) --
	(158.50,164.92) --
	(158.75,166.26) --
	(159.00,167.60) --
	(159.25,168.94) --
	(159.50,170.28) --
	(159.74,171.63) --
	(159.99,172.97) --
	(160.24,174.31) --
	(160.49,175.65) --
	(160.74,176.99) --
	(160.99,178.34) --
	(161.24,179.68) --
	(161.49,181.02) --
	(161.73,182.36) --
	(161.98,183.70) --
	(162.23,185.04) --
	(162.48,186.39) --
	(162.73,187.73) --
	(162.98,189.07) --
	(163.23,190.41) --
	(163.47,191.75) --
	(163.72,193.10) --
	(163.97,194.44) --
	(164.22,195.78) --
	(164.47,197.12) --
	(164.72,198.46) --
	(164.97,199.80) --
	(165.22,201.15) --
	(165.46,202.49) --
	(165.71,203.83) --
	(165.96,205.17) --
	(166.21,206.51) --
	(166.46,207.86) --
	(166.71,209.20) --
	(166.96,210.54) --
	(167.21,211.88) --
	(167.45,213.22) --
	(167.70,214.57) --
	(167.95,215.91) --
	(168.20,217.25) --
	(168.45,218.59) --
	(168.70,219.93) --
	(168.95,221.27) --
	(169.19,222.62) --
	(169.44,223.96) --
	(169.69,225.30) --
	(169.94,226.64) --
	(170.19,227.98) --
	(170.44,229.33) --
	(170.69,230.67) --
	(170.94,232.01) --
	(171.18,233.35) --
	(171.43,234.69) --
	(171.68,236.03) --
	(171.93,237.38);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (162.76,138.10) --
	(163.15,139.13) --
	(163.54,140.17) --
	(163.94,141.20) --
	(164.33,142.23) --
	(164.72,143.27) --
	(165.11,144.30) --
	(165.50,145.34) --
	(165.89,146.37) --
	(166.28,147.40) --
	(166.67,148.44) --
	(167.06,149.47) --
	(167.45,150.51) --
	(167.84,151.54) --
	(168.23,152.57) --
	(168.62,153.61) --
	(169.01,154.64) --
	(169.40,155.68) --
	(169.79,156.71) --
	(170.18,157.74) --
	(170.57,158.78) --
	(170.96,159.81) --
	(171.36,160.85) --
	(171.75,161.88) --
	(172.14,162.91) --
	(172.53,163.95) --
	(172.92,164.98) --
	(173.31,166.02) --
	(173.70,167.05) --
	(174.09,168.08) --
	(174.48,169.12) --
	(174.87,170.15) --
	(175.26,171.19) --
	(175.65,172.22) --
	(176.04,173.25) --
	(176.43,174.29) --
	(176.82,175.32) --
	(177.21,176.36) --
	(177.60,177.39) --
	(177.99,178.42) --
	(178.38,179.46) --
	(178.78,180.49) --
	(179.17,181.53) --
	(179.56,182.56) --
	(179.95,183.59) --
	(180.34,184.63) --
	(180.73,185.66) --
	(181.12,186.70) --
	(181.51,187.73) --
	(181.90,188.76) --
	(182.29,189.80) --
	(182.68,190.83) --
	(183.07,191.87) --
	(183.46,192.90) --
	(183.85,193.93) --
	(184.24,194.97) --
	(184.63,196.00) --
	(185.02,197.04) --
	(185.41,198.07) --
	(185.80,199.10) --
	(186.20,200.14) --
	(186.59,201.17) --
	(186.98,202.21) --
	(187.37,203.24) --
	(187.76,204.27) --
	(188.15,205.31) --
	(188.54,206.34) --
	(188.93,207.38) --
	(189.32,208.41) --
	(189.71,209.44) --
	(190.10,210.48) --
	(190.49,211.51) --
	(190.88,212.55) --
	(191.27,213.58) --
	(191.66,214.61) --
	(192.05,215.65) --
	(192.44,216.68) --
	(192.83,217.72) --
	(193.23,218.75) --
	(193.62,219.78);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (127.54, 34.03) -- (207.59,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.54, 34.03) rectangle (207.59,264.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49,264.40) rectangle (124.53,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51,267.41) {cover};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54,264.40) rectangle (207.59,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56,267.41) {growth rate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 41.20) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 93.56) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,145.91) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,198.27) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,250.62) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 44.51) --
	( 44.49, 44.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 96.86) --
	( 44.49, 96.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,149.22) --
	( 44.49,149.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,201.57) --
	( 44.49,201.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,253.93) --
	( 44.49,253.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.12, 29.77) --
	( 48.12, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.32, 29.77) --
	( 66.32, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.51, 29.77) --
	( 84.51, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.70, 29.77) --
	(102.70, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.89, 29.77) --
	(120.89, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 48.12, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 66.32, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (102.70, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (120.89, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] (131.18, 29.77) --
	(131.18, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.37, 29.77) --
	(149.37, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.56, 29.77) --
	(167.56, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.76, 29.77) --
	(185.76, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.95, 29.77) --
	(203.95, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (131.18, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (149.37, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (185.76, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (203.95, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (126.04,  9.03) {Predicted Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 17.30,149.22) {Observed Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (216.45,125.38) rectangle (268.17,173.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (220.72,162.17) {\bfseries predtype};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,144.10) rectangle (235.18,158.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (227.95,151.33) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,151.33) -- (233.73,151.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,129.64) rectangle (235.18,144.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (227.95,136.87) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,136.87) -- (233.73,136.87);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (236.98,149.14) {M};
\end{scope}
\end{tikzpicture}

\end{document}

### 2016-03-23 docs/ms_communitySynchrony.Rmd
adds main text word count and tex file

@@
 ---
-layout: 12pt
+bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
+csl: components/ecology.csl
+fontsize: 12pt
+geometry: margin=1in
 header-includes:
@@
 - \usepackage{rotating}
-bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
-csl: components/ecology.csl
-
-## rmarkdown render options
+layout: 12pt
+linkcolor: black
 output:
   pdf_document:
-    fig_caption: true
-    keep_tex: false
-fontsize: 12pt
-geometry: margin=1in
-linkcolor: black
+    fig_caption: yes
+    keep_tex: yes
+  word_document: default
 urlcolor: black
@@
 
-\textbf{Number of Words}: Abstract = ; Main Text =
+\textbf{Number of Words}: Abstract = 138; Main Text = 4929 (including section titles and in-line citations)
 

### 2016-03-23 docs/ms_communitySynchrony.tex
adds main text word count and tex file

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\doublespacing
\usepackage{color}
\usepackage{rotating}

\begin{document}

\maketitle

\begin{singlespace}

\begin{center}
\large{\textbf{Environmental responses, not species interactions, determine species synchrony in natural plant communities}}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\vspace{1em}

\normalsize{Andrew T. Tredennick\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}

\vspace{1em}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Theoretical and Experimental Ecology Station, CNRS and Paul Sabatier University, Moulis, 09200, France}}

\end{center}

\vspace{2em}

\textbf{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability

\vspace{2em}

\textbf{Authorship}: ATT, CdM, and PBA designed the research. CdM and ML designed and analyzed the theoretical model. ATT and PBA performed the empirical analysis. All authors contributed to the interpretation and presentation of results. ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.

\textbf{Running Title}: Drivers of species synchrony in natural plant communities

\textbf{Article Type}: Letter

\textbf{Number of Words}: Abstract = 138; Main Text = 4929 (including section titles and in-line citations)

\textbf{Number of References}: 46

\textbf{Number of Tables and Figures}:  2 Tables;  3 Figures

\textbf{Corresponding Author}: \\
\hspace{2em}Andrew Tredennick \\  
\hspace{2em}Department of Wildland Resources and the Ecology Center\\
\hspace{2em}Utah State University \\
\hspace{2em}5230 Old Main Hill \\
\hspace{2em}Logan, Utah 84322 USA \\
\hspace{2em}Phone: +1-970-443-1599 \\
\hspace{2em}Fax: +1-435-797-3796 \\
\hspace{2em}Email: atredenn@gmail.com

\end{singlespace}\renewcommand*{\thefootnote}{\arabic{footnote}}

\setcounter{footnote}{0}

\newpage{}

\begin{abstract}
Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific competition, and demographic stochasticity. We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
Simulation experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
Removing interspecific competition and demographic stochasticity had small effects on synchrony. In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
\end{abstract}

\newpage{} \setlength{\parindent}{5ex}

\subsection{INTRODUCTION}\label{introduction}

Ecosystems are being transformed by species extinctions (Cardinale et
al. 2012), changes in community composition (Vellend et al. 2013,
Dornelas et al. 2014), and anthropogenic environmental change (Vitousek
et al. 1997), impacting the provisioning and stability of ecosystem
services (Loreau et al. 2001, Hooper et al. 2005, Rockstrom et al.
2009). Experiments have provided compelling evidence that decreases in
species richness will decrease productivity (Tilman et al. 2001) and the
temporal stability of productivity (Tilman et al. 2006, Hector et al.
2010). The stabilizing effect of species richness stems from individual
species responding to environmental fluctuations (environmental
stochasticity), or fluctuating asynchronously because of random chance
events (demographic stochasticity). (Isbell et al. 2009, Hector et al.
2010, {{de Mazancourt}} et al. 2013). Species richness affects synchrony
because larger species pools are more likely to contain species that
respond disimilarly to environmental conditions (Yachi and Loreau 1999),
implying that species losses will reduce ecosystem stability. Even
without species losses, abiotic homogenization can weaken compensatory
dynamics and, in turn, decrease temporal stability of ecosystem
functioning (Hautier et al. 2014). Predicting the impacts of global
change requires a mechanistic understanding of species synchrony in
natural settings.

Theory identifies the three main determinants of species synchrony as
environmental stochasticity, demographic stochasticity, and
interspecific competition (Loreau and {{de Mazancourt}} 2008, 2013,
Gonzalez and Loreau 2009). For example, in a community composed of large
(no demographic stochasticity) and weakly interacting (no interspecific
competition) populations, community-wide species synchrony should be
determined by the covariance of species' responses to the environment
(Loreau and {{de Mazancourt}} 2008). However, such a prediction relies
on a relatively simple population model and only holds under two
assumptions: (i) species' responses to the environment are similar in
magnitude and (ii) all species have similar growth rates. Whether such
theoretical predictions hold in natural communities where species
differences are unlikely to be symmetrical remains unkown: no study has
explicitly tested theory on the drivers of species synchrony in natural
communities.

In grasslands, most empirical studies have focused on whether species
synchrony is primarily an outcome of species-specific responses to
environmental conditions (Hautier et al. 2014) or competition (Gross et
al. 2014). Even beyond grassland studies, whether competition or
environmental responses drive compensatory dynamics remains
controversial (reviewed in Gonzalez and Loreau 2009). In part,
controversy remains because quantifying the relative strengths of each
driver on the degree of synchrony from community time series (e.g.,
Houlahan et al. 2007) is impossible. This is because an unbiased null
expectation for synchrony does not exist (Loreau and {{de Mazancourt}}
2008) and observed synchrony can arise from non-unique combinations of
factors (Ranta et al. 2008). For example, weak synchrony of population
abundances could reflect positive environmental correlations
(synchronizing effect) offset by strong competition (desynchronizing
effect), or negative environmental correlations and weak competition.

The best way to quantify the effects of environmental stochasticity,
demographic stochasticity, and interspecific competition is to remove
them one-by-one, and in combination. In principle, this could be done in
an extremely controlled laboratory setting, but empirically-based models
of interacting populations, fit with data sets from natural communities,
offer a practical alternative. For example, Mutshinda et al. (2009) fit
a dynamic population model to several community time series of insect
and bird abundances. They used a statistical technique to decompose
temporal variation into competition and environmental components, and
found that postively correlated environmental responses among species
determined community dynamics. While a major step forward, Mutshinda et
al.'s (2009) modeling technique relied on abundance data that may or may
not reliably capture competitive interactions that occur at the
individual level. Furthermore, although Mutshinda et al. (2009)
quantified the relative importance of environmental stochasticity and
competition to explain the observed variation of species synchrony, they
did not use the model to quantify how much synchrony would change when
each factor is removed.

Here, we use multi-species population models fit to long-term
demographic data from five semi-arid plant communities to test theory on
the drivers of species synchrony. Our objectives are to (1) derive and
test theoretical predictions of species synchrony and (2) determine the
relative influence of environmental stochasticity, demographic
stochasticity, and interspecific competition on species synchrony in
natural plant communities. To achieve these objectives, we first refine
theory that has been used to predict the effects of species richness on
ecosystem stability ({{de Mazancourt}} et al. 2013) and species
synchrony (Loreau and {{de Mazancourt}} 2008) to generate predictions of
community-wide species synchrony under two limiting cases derived from
the dynamics of individual species in monoculture. We then confront our
theoretical predictions with simulations from the empirically-based
population models. Second, we use the multi-species population models to
perform simulation experiments that isolate the effects of environmental
stochasticity, demographic stochasticity, and interspecific competition
on community-wide species synchrony. Given that our population models
capture the essential features of community dynamics important to
synchrony (density-dependence, and demographic and environmental
stochasticity), and that these models successfully reproduce observed
community dynamics (Chu and Adler 2015), perturbing the models can
reveal the processes that determine species synchrony in our focal
grassland communities.

\subsection{THEORETICAL MODEL}\label{theoretical-model}

\subsubsection{The model}\label{the-model}

While existing theory has identified the factors driving synchrony, we
do not have a simple expression to predict synchrony in a particular
community with all factors operating simultaneously. However, we can
derive analytical predictions for species synchrony under special
limiting cases. We focus on synchrony of per capita growth rates, rather
than abundances, because growth rates represent the instantaneous
response of species to the environment and competition, and are less
susceptible to the legacy effects of drift and disturbance (Loreau and
{{de Mazancourt}} 2008). We present equivalent results for synchrony of
species abundances in the Online Supporting Information, and show that
they lead to the same overall conclusions as synchrony of per capita
growth rates. Following Loreau and {{de Mazancourt}} (2008) and {{de
Mazancourt}} et al. (2013), we define population growth, ignoring
observation error, as

\begin{align}
r_i(t) &= \text{ln}N_i(t+1) - \text{ln}N_i(t) \\
&= r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}u_{ei}(t) + \frac{\sigma_{di}u_{di}(t)}{\sqrt{N_i(t)}} \right]
\end{align}

\noindent where \(N_i(t)\) is the biomass of species \emph{i} in year
\emph{t}, and \(r_i(t)\) is its population growth rate in year \emph{t}.
\(r_{mi}\) is species \emph{i}'s intrinsic rate of increase, \(K_i\) is
its carrying capacity, and \(\alpha_{ij}\) is the interspecific
competition coefficient representing the effect of species \emph{j} on
species \emph{i}. Environmental stochasticity is incorporated as
\(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the environmental
variance and \(u_{ei}\) are normal random variables with zero mean and
unit variance that are independent through time but may be correlated
between species. Demographic stochasticity arises from variations in
births and deaths among individuals (e.g., same states, different
fates), and is included in the model as a first-order, normal
approximation (Lande et al. 2003, {{de Mazancourt}} et al. 2013).
\(\sigma_{di}^2\) is the demographic variance and \(u_{di}(t)\) are
independent normal variables with zero mean and unit variance. To derive
analytical predictions we solved a first-order approximation of Equation
2 ({{de Mazancourt}} et al. 2013 and Online Supporting Information).

\subsubsection{Predictions}\label{predictions}

Our first prediction assumes no interspecific interactions, no
environmental stochasticity, identical intrinsic growth rates, and that
demographic stochasticity is operating but all species have identical
demographic variances. This limiting case, \(\mathcal{M}_{D}\),
represents a community where dynamics are driven by demographic
stochasticity alone. Our prediction is

\begin{equation}
\phi_{R,\mathcal{M}_{D}} = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
\end{equation}

\noindent for synchrony of per capita growth rates. \(p_i\) is the
average frequency of species \emph{i}, \(p_i = N_i/N_T\). When all
species have identical abundances and \(p_i = 1/S\), where \emph{S} is
species richness, synchrony equal 1/S (Loreau and {{de Mazancourt}}
2008).

Our second limiting case assumes only environmental stochasticity is
operating. Thus, there are no interspecific interactions, demographic
stochasticity is absent, intrinsic growth rates are identical, and
environmental variance is identical for all species. Our prediction for
this limiting case, \(\mathcal{M}_{E}\), is

\begin{equation}
\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}p_ip_j\text{cov}(u_{ei},u_{ej})}{S^2}
\end{equation}

\noindent{} for synchrony of per capita growth rates, where
\(\text{cov}(u_{ei},u_{ej})\) is the standardized covariance of
environmental responses between species \emph{i} and species \emph{j}.
Confronting our theoretical predictions with data requires estimates of
species dynamics of large populations (no demographic stochasticity) in
monoculture (no interspecific interactions) to calculate the covariance
of species' environmental responses. To estimate environmental responses
in natural communities, we turn to our population models built using
long-term demographic data.

\subsection{EMPIRICAL ANALYSIS}\label{empirical-analysis}

\subsubsection{Materials and methods}\label{materials-and-methods}

\paragraph{Data}\label{data}

In the early 20th century, several range experiment stations in the
western U.S. began mapping permanent quadrats, and conducted annual
censuses for decades. We use such long-term demographic data sets from
five semiarid grasslands in the western United States (described in
detail by Chu and Adler 2015). Each study site includes a set of
1-\(\text{m}^2\) permanent quadrats within which all individual plants
were identified and mapped annually using a pantograph (Hill 1920). The
resulting mapped polygons represent basal cover for grasses and canopy
cover for shrubs. Data come from the Sonoran desert in Arizona (Anderson
et al. 2012), sagebrush steppe in Idaho (Zachmann et al. 2010), southern
mixed prairie in Kansas (Adler et al. 2007), northern mixed prairie in
Montana (Anderson et al. 2011), and Chihuahuan desert in New Mexico
(Anderson et al. in preparation, Chu and Adler 2015) (Table 1).

\paragraph{Calculating observed
synchrony}\label{calculating-observed-synchrony}

The data consist of records for individual plant size in quadrats for
each year. To obtain estimates of percent cover for each focal species
in each year, we summed the individual-level data within quadrats and
then averaged percent cover, by species, over all quadrats. We
calculated per capita growth rates as
\(\text{log}(x_t) - \text{log}(x_{t-1})\), where \emph{x} is species'
percent cover in year \emph{t}. Using the community time series of per
capita growth rates or percent cover, we calculated community synchrony
using the metric of Loreau and {{de Mazancourt}} (2008) in the
`synchrony' package (Gouhier and Guichard 2014) in R (R Core Team 2013).
Specifically, we calculated synchrony as

\begin{equation}
\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
\end{equation}

\noindent where \(\sigma_{r_{i}}\) is the temporal variance of species
\emph{i}`s per capita population growth rate (\(r_i\)) and
\(\sigma^{2}_{T}\) is the temporal variance of the aggregate
community-level growth rate. \(\phi\) ranges from 0 at perfect
asynchrony to 1 at perfect synchrony (Loreau and {{de Mazancourt}}
2008). We use the same equation to calculate observed synchrony of
species' percent cover, which we present to relate our results to
previous findings, even though we focus on synchrony of growth rates in
our model simulations (see below).

\paragraph{Fitting statistical models}\label{fitting-statistical-models}

Vital rate regressions are the building blocks of our dynamic models: an
integral projection model (IPM) and an individual-based model (IBM). We
followed the approach of Chu and Adler (2015) to fit statistical models
for survival, growth, and recruitment (see Online Supporting Information
for full details). We modeled survival probability of each genet as
function of genet size, temporal variation among years, permanent
spatial variation among groups of quadrats, and local neighborhood
crowding from conspecific and heterospecific genets. Regression
coefficients for the effect of crowding by each species can be
considered a matrix of interaction coefficients whose diagonals
represent intraspecific interactions and whose off-diagonals represent
interspecific interactions (Adler et al. 2010). These interaction
coefficients can take positive (facilitative) or negative (competitive)
values. We modeled growth as the change in size of a genet from one year
to the next, which depends on the same factors as the survival model. We
fit the survival and growth regressions using INLA (Rue et al. 2014), a
statistical package for fitting generalized linear mixed effects models
via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
2013). Genet size and crowding were treated as fixed effects; spatial
(quadrat groupings) variation was treated as a random effect on the
intercept and temporal (interannual) variation was treated as random
effects on the intercept and the effect of genet size in the previous
year (Online Supporting Information).

We modeled recruitment at the quadrat scale, rather than the individual
scale, because it is impossible to attribute new genets to specific
parents (Chu and Adler 2015). Our recruitment model assumes that the
number of recruits produced in each year follows a negative binomial
distribution with the mean dependent on the cover of the parent species,
permanent spatial variation among groups, temporal variation among
years, and inter- and intraspecific interactions as a function of total
species' cover in the quadrat. We fit the recruitment model using a
hierarchical Bayesian approach implemented in JAGS (Plummer 2003) using
the `rjags' package (Plummer 2014) in R. Again, temporal and spatial
variation were treated as random effects.

\paragraph{Building dynamic multi-species
models}\label{building-dynamic-multi-species-models}

Once we have fit the vital rate statistical models, building the
population models is straightforward. For the IBM, we initialize the
model by randomly assigning plants spatial coordinates, sizes, and
species identities until each species achieves a density representative
of that observed in the data. We then project the model forward by using
the survival regression to determine whether a genet lives or dies, the
growth regression to calculate changes in genet size, and the
recruitment regression to add new individuals that are distributed
randomly in space. Crowding is directly calculated at each time step
since each genet is spatially referenced (as in the observed data).
Environmental stochasticity is not an inherent feature of IBMs, but is
easily included since we fit year-specific temporal random effects for
each vital rate regression. To include temporal environmental variation,
at each time step we randomly choose a set of estimated survival,
growth, and recruitment parameters specific to one observation year. For
all simulations, we ignore the spatial random effect associated with
variation among quadrat groups, so our simulations represent an average
quadrat for each site.

The IPM uses the same vital rate regressions as the IBM (Rees and Ellner
2009, Rees et al. 2014), but it is spatially implicit and does not
include demographic stochasticity. Following Chu and Adler (2015), we
use a mean field approximation that captures the essential features of
spatial patterning to define the crowding index at each time step
(Supporting Online Information). Temporal variation is included in
exactly the same way as for the IBM. For full details on the IPMs we
use, see Chu and Adler (2015).

\paragraph{Simulation experiments}\label{simulation-experiments}

We performed removal simulation experiments where drivers (environmental
stochasticity, demographic stochasticity, or interspecific competition)
were removed one-by-one and in combination. To remove interspecific
interactions, we set the off-diagonals of the interaction matrix for
each vital rate regression to zero. This retains intraspecific
interactions, and thus density-dependence, and results in simulations
where species are effectively growing in monoculture. To remove the
effect of a fluctuating environment, we removed the temporal
(interannual) random effects from the regression equations. To remove
the effect of demographic stochasticity, we use the IPM rather than the
IBM because the IPM does not include demographic stochasticity
(demographic stochasticity cannot be removed from the IBM). Since the
effect of demographic stochasticity on population dynamics depends on
population size (Lande et al. 2003), we can control the strength of
demographic stochasticity by simulating the IBM on areas (e.g.~plots) of
different size. Indeed, results from an IBM with infinite population
size would converge on results from the IPM. Given computational
constraints, the largest landscape we simulate is a 25 \(\text{m}^2\)
plot.

We conducted the following six simulation experiments: (1) IBM: All
drivers (environmental stochasticity, demographic stochasticity, or
interspecific competition) present; (2) IPM: Demographic stochasticity
removed; (3) IBM: Environmental stochasticity removed; (4) IBM:
Interspecific competition removed; (5) IPM: Interspecific competition
and demographic stochasticity removed; (6) IBM: Interspecific
competition and environmental stochasticity removed. We ran IPM
simulations for 2,000 time steps, after an initial 500 iteration burn-in
period. This allowed species time to reach their stable size
distribution. We then calculated the synchrony of species' per capita
growth rates over 100 randomly selected contiguous 50 time-step
sections. We ran IBM simulations for 100 time steps, and repeated the
simulations 100 times for each simulation experiment. From those, we
retained only the simulations in which no species went extinct due to
demographic stochasticity. Synchrony was calculated over the 100 time
steps for each no extinction run within a model experiment. To explore
the effect of demographic stochasticity in different sized populations,
we ran simulations (1) and (6) on plot sizes of 1, 4, 9, 16, and 25
\(\text{m}^2\). All other IBM simulations were run on a 25
\(\text{m}^2\) landscape.

Results from our simulation experiments also allow us to test our
theoretical predictions. First, in the absence of interspecific
competition and demographic stochasticity, species are effectively
growing in large monocultures. Therefore, we can use results from
simulation (5) to estimate the covariance of species' responses to the
environment (\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation
3. Parameterizing Equation 2 does not require simulation output because
the only parameters are the species' relative abundances. Second,
simulations (5) and (6) represent the simulated version of our limiting
case theoretical predictions. Thus, we directly test the theoretical
predictions by comparing them to observed synchrony and simulated
synchrony.

\subsubsection{Results}\label{results}

Synchrony of species' per capita growth rates at our study sites ranged
from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to
0.92 (Table 2). Synchrony tended to be higher in communities with few
species (Arizona and New Mexico) and/or with relatively high temporal
variability in percent cover (e.g., Montana). Synchrony was lowest in
Idaho, the only data set that includes two life forms: a shrub and three
perennial grasses (Table 1). For all five communities, species synchrony
from IPM and IBM simulations closely approximated observed synchrony
(Fig. S1). IBM-simulated synchrony is consistently, but only slightly,
lower than IPM-simulated synchrony (Fig. S1), likely due to the
desynchronizing effect of demographic stochasticity.

Across the five communities, our limiting case predictions closely
matched synchrony from the corresponding simulation experiment (Fig. 1
and Table S1). The correlation between our analytical predictions and
simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) and 0.997
for \(\phi_{R,\mathcal{M}_E}\). The largest difference between predicted
and simulated synchrony was 0.05 in New Mexico for
\(\phi_{R,\mathcal{M}_D}\) (Table S1).

Simulation experiments revealed that removing environmental fluctuations
has the largest impact on synchrony, leading to a reduction in synchrony
of species growth rates in four out of five communities (Fig. 1).
Removing environmental fluctuations (``No E.S'' simulations) decreased
synchrony by 33\% in Arizona, 48\% in Kansas, 39\% in Montana, and 40\%
in New Mexico. Only in Idaho did removing environmental fluctuations
cause an increase in synchrony (Fig. 1), but the effect was small (9\%
increase; Table S2). Overall, species' temporal random effects in the
statistical vital rate models are positively correlated (Table S3).
Species interactions are weak in these communities (Table S4 and Chu and
Adler 2015), and removing interspecific competition had little effect on
synchrony (Fig. 1; ``No Comp.'' simulations). Removing competition
caused, at most, a 5\% change in synchrony (Fig. 1). Removing
demographic stochasticity (``No D.S.'' simulations) caused synchrony to
increase slightly in all communities (Fig. 1), with an average 6\%
increase over synchrony from IBM simulations on a five meter area.

The desynchronizing effect of demographic stochasticity, which increases
as population size decreases, modestly counteracted the synchronizing
force of the environment, but not enough to lower synchrony to the level
observed when only demographic stochasticity is operating (Fig. 2). In
the largest, 25 \(\text{m}^2\) plots, synchrony was driven by
environmental stochasticity (e.g., \(\mathcal{M}_E\)). At 1
\(\text{m}^2\), synchrony reflected demographic stochasticity and
environmental stochasticity (e.g., between \(\mathcal{M}_E\) and
\(\mathcal{M}_D\)). For context, population sizes increased from an
average of 17 individuals per community in 1 \(\text{m}^2\) IBM
simulations to an average of 357 individuals per community in 25
\(\text{m}^2\) IBM simulations.

For all five communities, the synchrony of species' growth rates in
monoculture almost perfectly matched species synchrony in polyculture
(Fig. 3). Results for synchrony of percent cover are qualitatively
similar, but simulation results were more variable and less consistent
with analytical predictions and observed synchrony (Figs. S2-S3).

\subsection{DISCUSSION}\label{discussion}

Our study produced four main findings that were generally consistent
across five natural plant communities: (1) limiting-case predictions
from the theoretical model were well-supported by simulations from the
empirical models; (2) demographic stochasticity decreased community
synchrony, as expected by theory, and its effect was largest in small
populations; (3) environmental fluctuations increased community
synchrony relative to simulations in constant environments because
species-specific responses to the environment were positively, though
not perfectly, correlated; and (4) interspecific interactions were weak
and therefore had little impact on community synchrony. We also found
that analyses based on synchrony of species' percent cover were
uninformative (Figs. S2-S3), implying that it is difficult to decipher
mechanisms of species synchrony from abundance time series, as expected
by theory (Loreau and {{de Mazancourt}} 2008).

\subsubsection{Simulations support theoretical
predictions}\label{simulations-support-theoretical-predictions}

Our theoretical predictions were derived from a simple model of
population dynamics and required several simplifying assumptions,
raising questions about their relevance to natural communities. However,
the theoretical predictions were strongly supported by simulations of
population models fit to long-term data from natural plant communities
(Table 3). Such a close match between our analytical predictions and the
simulation results should inspire confidence in the ability of simple
models to inform our understanding of species synchrony. The theoretical
population dynamical model we used here also successfully predicted
ecosystem stability in several biodiversity-ecosystem functioning
experiments ({{de Mazancourt}} et al. 2013).

\subsubsection{Demographic stochasticity decreases
synchrony}\label{demographic-stochasticity-decreases-synchrony}

Demographic stochasticity partially counteracted the synchronizing
effects of environmental fluctuations and interspecific interactions on
per capita growth rates, but only when populations were small (Fig. 2),
in agreement with theory (Loreau and {{de Mazancourt}} 2008). Even in
small populations, however, demographic stochasticity was not strong
enough to completely counteract the synchronizing effects of
environmental fluctuations and match the analytical prediction where
only demographic stochasticity is operating (\(\mathcal{M}_D\) in Fig.
2). These results confirm the theoretical argument by Loreau and {{de
Mazancourt}} (2008) that independent fluctuations among interacting
species in a non-constant environment should be rare. Only in the Idaho
community does synchrony of per capita growth rates approach
\(\mathcal{M}_D\) in a non-constant environment (Fig. 2). This is most
likely due to the strong effect of demographic stochasticity on the
shrub \emph{Artemisia tripartita} since even a 5 \(\text{m}^2\) quadrat
would only contain a few individuals of that species.

Our analysis of how demographic stochasticity affects synchrony
demonstrates that synchrony depends on the observation area. As the
observation area increases, population size increases and the
desynchronizing effect of demographic stochasticity lessens (Fig. 2).
Thus, our results suggest that community-wide species synchrony will
increase as the observation area increases, rising from
\(\mathcal{M}_D\) to \(\mathcal{M}_E\). Such a conclusion assumes,
however, that species richness remains constant as observation area
increases, which is unlikely (Taylor 1961). Recent theoretical work has
begun to explore the linkage between ecosystem stability and spatial
scale (Wang and Loreau 2014, 2016), and our results suggest that
including demographic stochasticity in theoretical models of
metacommunity dynamics may be important for understanding the role of
species synchrony in determining ecosystem stability across spatial
scales.

\subsubsection{Environmental fluctuations drive community
synchrony}\label{environmental-fluctuations-drive-community-synchrony}

In large populations where interspecific interactions are weak,
synchrony is expected to be driven exclusively by environmental
fluctuations (Equation 4). Under such conditions community synchrony
should approximately equal the synchrony of species' responses to the
environment (Loreau and {{de Mazancourt}} 2008). Two lines of evidence
lead us to conclude that environmental fluctuations drive species
synchrony in our focal plant communities. First, in our simulation
experiments, removing interspecific interactions resulted in no
discernible change in community-wide species synchrony of per capita
growth rates (Fig. 1). Second, removing environmental fluctuations from
simulations consistently reduced synchrony (Fig. 1), and the synchrony
of species in monoculture was a very strong predictor of synchrony of
species in polyculture (Fig. 3). Our results lead us to conclude that
environmental fluctuations, not species interactions, are the primary
driver of community-wide species synchrony in the communities we
studied. Given accumulating evidence that niche differences in natural
communities are large (reviewed in Chu and Adler 2015), and thus species
interactions are likely to be weak, our results may be general in
natural plant communities.

In the Idaho community, removing environmental fluctuations did not
cause a large decrease in synchrony. However, that result appears to be
an artifact. Removing environmental variation results in a negative
invasion growth rate for \emph{A. tripartita}. Although we only analyzed
IBM runs in which \emph{A. tripartita} had not yet gone extinct, it was
at much lower abundance than in the other simulation runs. When we
removed \emph{A. tripartita} from all simulations, the Idaho results
conformed with results from all other sites: removing environmental
stochasticity caused a significant reduction in species synchrony (Fig.
S4). Our main results for Idaho (Fig. 2), with \emph{A. tripartita}
included, demonstrate how the processes that determine species synchrony
interact in complex ways. \emph{A. tripartita} has a facilitative effect
on each grass species across all vital rates, except for a small
competitive effect on \emph{H. comata}'s survival probability (Tables
S8-S10). At the same time, all the perennial grasses have negative
effects on each other for each vital rate (Tables S8-S10). We know
synchrony is affected by interspecific competition (Loreau and {{de
Mazancourt}} 2008), but how facilitative effects manifest themselves is
unknown. The interaction of facilitation and competition is clearly
capable of having a large effect on species synchrony, and future
theoretical efforts should aim to include a wider range of species
interactions.

A challenge to the generality of our results is that we were only able
to model common, co-occuring species (see Chu and Adler 2015). Rare
species, which our analysis does not include, could be kept at low
densities by competitive suppression due to strong interspecific
interactions. If we had been able to model rare species, we might have
found a larger effect of competition on community synchrony. However,
rare species may occupy rare niches that do not overlap with other
species, in which case synchrony would remain predominantly driven by
environmental responses. Rare species could also be limited by
generalist natural enemies whose effects are density-independent but
alter competitive hierarchies (Mordecai 2011). Under such conditions,
synchrony will depend on the degree of pathogen-induced fitness
differences and the pathogen's response to environmental conditions.
Neither our model nor current theory explicitly considers the effects of
pathogens on species synchrony, and this highlights the need for
theoretical work on the interaction between mechanisms of species
coexistence and mechanisms of species synchrony (Loreau 2010).

\subsubsection{Interspecific interactions had little impact on community
synchrony}\label{interspecific-interactions-had-little-impact-on-community-synchrony}

We expected community synchrony of per capita growth rates to decrease
when we removed interspecific interactions (Loreau and {{de Mazancourt}}
2008). We found that community synchrony was virtually indistinguishable
between simulations with and without interspecific interactions (Fig.
2). The lack of an effect of interspecific competition on synchrony is
in contrast to a large body of theoretical work that predicts a strong
role for competition in creating compensatory dynamics (Tilman 1988) and
a recent empirical analysis (Gross et al. 2014).

Our results do not contradict the idea that competition can lead to
compensatory dynamics, but they do highlight the fact that interspecific
competition must be relatively strong to influence species synchrony.
The communities we analyzed are composed of species with very little
niche overlap (Chu and Adler 2015) and weak interspecific interactions
(Tables S1, S3-S17). Mechanistic consumer-resource models (Lehman and
Tilman 2000) and phenomenological Lotka-Volterra models (Lehman and
Tilman 2000, Loreau and {{de Mazancourt}} 2013) both confirm that the
effect of competition on species synchrony diminishes as niche overlap
decreases. In that sense, our results are not surprising: interspecific
interactions are weak, so of course removing them does not affect
synchrony.

However, our results do contrast with a recent analysis of several
biodiversity-ecosystem functioning experiments showing that competition
drives species synchrony in grasslands (Gross et al. 2014). The apparent
inconsistency between our results and those of Gross et al. (2014) may
be explained by the differences between our studies. Gross et al.'s
results are based on rapidly assembling experimental communities that
are, at most, 11 years old. The natural, relatively undisturbed
communities we studied consist of species that have co-occurred for
decades (Chu and Adler 2015) and represent a much later stage of
community assembly. In theory, species interactions should weaken
through time as community assembly proceeds (Kokkoris et al. 1999),
meaning that in biodiversity-ecosystem functioning experiments the
processes driving competitive exclusion are still operating. Such
communities may be highly influenced by transient, but strong,
interspecific competition that can mask the synchronizing effects of the
environment. In contrast, synchrony may be driven by environmental
fluctuations in older communities. In other words, the dominant driver
of community synchrony should shift from competition to environmental
fluctuations through time. One way to test this prediction is to
continue collecting data from biodiversity-ecoystem functioning
experiments and conduct the analysis of Gross et al. (2014) every few
years.

\subsection{CONCLUSIONS}\label{conclusions}

Species-specific responses to temporally fluctuating environmental
conditions is an important mechanism underlying asynchronous population
dynamics and, in turn, ecosystem stability (Loreau and {{de Mazancourt}}
2013). When we removed environmental variation, we found that synchrony
decreased in four out of the five grassland communities we studied (Fig.
2). A tempting conclusion is that our study confirms that compensatory
dynamics are rare in natural communities, and that ecologically-similar
species will exhibit synchronous dynamics (e.g., Houlahan et al. 2007).
Such a conclusion misses an important subtlety. The perennial grasses we
studied do have similar responses to the environment (Table S2), which
will tend to synchronize dynamics. However, if community-wide species
synchrony is less than 1, as it is in all our focal communities, some
degree of compensatory dynamics must be present (Loreau and {{de
Mazancourt}} 2008). In agreement with other studies (Rocha et al. 2011,
Vasseur et al. 2014), we find that environmental responses are primarily
responsible for the degree of synchrony among ecologically-similar
species. This result contrasts with a recent analysis of several
biodiversity-ecosystem functioning experiments showing that competition
drives community synchrony (Gross et al. 2014). Recently assembled
communities, such as experimental plots, may exhibit strong species
interactions that will weaken over time (Kokkoris et al. 1999). Future
research on the influence of community assembly on synchrony could
reconcile inconsistent observations on the biotic and abiotic forces
that jointly drive community dynamics.

\subsection{ACKNOWLEDGMENTS}\label{acknowledgments}

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040). CdM and ML were supported by the
TULIP Laboratory of Excellence (ANR-10-LABX-41). We thank the original
mappers of the permanent quadrats at each site and the digitizers in the
Adler lab, without whom this work would not have been possible. Compute,
storage, and other resources from the Division of Research Computing in
the Office of Research and Graduate Studies at Utah State University are
gratefully acknowledged.

\setlength{\parindent}{0ex} \newpage{}

\subsection{Tables}\label{tables}

\singlespacing

Table 1: Site descriptions and focal species. \footnotesize

\begin{longtable}[c]{@{}lllll@{}}
\toprule
Site Name & Biome & Location (Lat, Lon) & Obs. Years &
Species\tabularnewline
\midrule
\endhead
New Mexico & Chihuahuan Desert & 32.62° N, 106.67° W & 1915-1950 &
\emph{Bouteloua eriopoda}\tabularnewline
& & & & \emph{Sporobolus flexuosus}\tabularnewline
Arizona & Sonoran Desert & 31°50' N, 110°53' W & 1915-1933 &
\emph{Bouteloua eriopoda}\tabularnewline
& & & & \emph{Bouteloua rothrockii}\tabularnewline
Kansas & Southern mixed prairie & 38.8° N, 99.3° W & 1932-1972 &
\emph{Bouteloua curtipendula}\tabularnewline
& & & & \emph{Bouteloua hirsuta}\tabularnewline
& & & & \emph{Schizachyrium scoparium}\tabularnewline
Montana & Northern mixed prairie & 46°19' N, 105°48' W & 1926-1957 &
\emph{Bouteloua gracilis}\tabularnewline
& & & & \emph{Hesperostipa comata}\tabularnewline
& & & & \emph{Pascopyrum smithii}\tabularnewline
& & & & \emph{Poa secunda}\tabularnewline
Idaho & Sagebrush steppe & 44.2° N, 112.1° W & 1926-1957 &
\emph{Artemisia tripartita}\tabularnewline
& & & & \emph{Pseudoroegneria spicata}\tabularnewline
& & & & \emph{Hesperostipa comata}\tabularnewline
& & & & \emph{Poa secunda}\tabularnewline
\bottomrule
\end{longtable}

\normalsize

\pagebreak{}

\begin{table}[ht]
\centering
\caption{Observed synchrony among species' per capita growth rates ($\phi_{R}$), observed synchrony among species' percent cover ($\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community. Species richness values reflect the number of species analyzed from the community, not the actual richness.} 
{\normalsize
\begin{tabular}{lrrrr}
  \hline
Site & $\phi_{R}$ & $\phi_{C}$ & CV of Total Cover & Species richness \\ 
  \hline
New Mexico & 0.86 & 0.92 & 0.51 &   2 \\ 
  Arizona & 0.89 & 0.80 & 0.47 &   2 \\ 
  Kansas & 0.54 & 0.15 & 0.30 &   3 \\ 
  Montana & 0.53 & 0.58 & 0.54 &   4 \\ 
  Idaho & 0.36 & 0.18 & 0.19 &   4 \\ 
   \hline
\end{tabular}
}
\end{table}

\pagebreak{}

\subsection{Figures}\label{figures}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure1.png}
  \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments. Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp. + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp. + No E.S.''). Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation. Error bars represent the upper and lower 95\% quantiles from model simulations. All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape. Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S. Only'') and when environmental stochasticity is also present removed (``D.S. + E.S.''). The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line). The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase. Theoretically, ``D.S. Only'' simulations should remain constant across landscape size, whereas ``D.S. + E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing. Error bars represent the upper and lower 95\% quantiles from model simulations.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=4in]{./components/formatted_figures/formatted_figure3.png}
  \caption{Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions). We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments. The dashed line is the line of equality. Simulation results in this figure are analagous to ``No Comp. + No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig. 1, but here we control the time series of random year effects.}
\end{figure}

\pagebreak{}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
perennial plants: An embarrassment of niches. Ecology Letters
13:1019--1029.

Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
mapped quadrats from Kansas prairie: demographic information for
herbaceous plants. Ecology 88:2673.

Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
of semi-desert grassland plants in permanent quadrats mapped from 1915
to 1947. Ecology 93:1492--1492.

Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
92:1703.

Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
and its impact on humanity. Nature 489:59--67.

Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
recruitment stage to stabilize grassland coexistence. Ecological
Monographs 85:373--392.

{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
Loreau. 2013. Predicting ecosystem stability from community composition
and biodiversity. Ecology Letters 16:617--625.

Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
biodiversity change but not systematic loss. Science 344:296--9.

Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
Compensatory Dynamics in Ecological Communities. Annual Review of
Ecology, Evolution, and Systematics 40:393--414.

Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
variability in space and time. Methods in Ecology and Evolution
5:524--533.

Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
temporal stability of biomass production: a new analysis of recent
biodiversity experiments. The American naturalist 183:1--12.

Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
of diversity in natural grasslands. Nature 508:521--5.

Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
diversity on grassland productivity through population asynchrony and
overyielding. Ecology 91:2213--2220.

Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
on ecosystem functioning: a consensus of current knowledge. Ecological
Monographs 75:3--35.

Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
Compensatory dynamics are rare in natural ecological communities.
Proceedings of the National Academy of Sciences of the United States of
America 104:3273--3277.

Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
productivity and the temporal stability of productivity: Patterns and
processes. Ecology Letters 12:443--451.

Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
species interaction strength in assembled theoretical competition
communities. Ecology Letters 2:70--74.

Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
dynamics in ecology and conservation.

Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
Productivity in Competitive Communities. The American Naturalist
156:534--552.

Loreau, M. 2010. From Polutations to Ecosystems: Theoretical Fondations
for a New Ecological Synthesis.

Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
drivers: neutral and nonneutral community dynamics in fluctuating
environments. The American Naturalist 172:E48--E66.

Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
stability: A synthesis of underlying mechanisms. Ecology Letters
16:106--115.

Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
knowledge and future challenges. Science 294:804--808.

Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
theory, concepts, and empirical work. Ecological Monographs 81:429--441.

Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
community dynamics? Proceedings of the Royal Society B: Biological
Sciences 276:2923--2929.

Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
3rd international workshop on distributed statistical computing (dSC
2003). march.

Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.

R Core Team. 2013. R: A language and environment for statistical
computing.

Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
O'Hara. 2008. Detecting compensatory dynamics in competitive communities
under environmental forcing. Oikos 117:1907--1911.

Rees, M., and S. P. Ellner. 2009. Integral projection models for
populations in temporally varying environments. Ecological Monographs
79:575--594.

Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
projection models: A user's guide.

Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
species have similar dynamics. Journal of Ecology 99:1453--1459.

Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
operating space for humanity. Nature 461:472--475.

Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
for Latent Gaussian Models Using Integrated Nested Laplace
Approximations. Journal of the Royal Statistical Society, Series B
71:319--392.

Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
analysis of latent Gaussian models using Integrated Nested Laplace
Approximaxion.

Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
189:732--735.

Tilman, D. 1988. Plant strategies and the dynamics and structure of
plant communities. Plant strategies and the dynamics and structure of
plant communities.:360 pp.

Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
ecosystem stability in a decade-long grassland experiment. Nature
441:629--632.

Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
2001. Diversity and productivity in a long-term grassland experiment.
Science 294:843--845.

Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
2014. Synchronous dynamics of zooplankton competitors prevail in
temperate lake ecosystems. Proceedings. Biological sciences / The Royal
Society 281:20140633.

Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
Beauséjour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf. 2013.
Global meta-analysis reveals no net change in local-scale plant
biodiversity over time. Proceedings of the National Academy of Sciences
of the United States of America 110:19456--9.

Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
Human Domination of Earth' s Ecosystems. Science 277:494--499.

Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.

Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
across scales in metacommunities. Ecology Letters.

Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
in a fluctuating environment: the insurance hypothesis. Proceedings of
the National Academy of Sciences of the United States of America
96:1463--1468.

Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
sagebrush steppe: long-term data for analyzing demographic rates and
plant--plant interactions. Ecology 91:3427.

\end{document}

### 2016-03-24 docs/ms_communitySynchrony.tex
updates figure formatting

@@
 
+\subsection{Table Titles}
+
+TABLE 1: Site descriptions and focal species.
+
+TABLE 2: Observed synchrony among species' per capita growth rates ($\phi_{R}$), observed synchrony among species' percent cover ($\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community.
+Species richness values reflect the number of species analyzed from the community, not the actual richness.
+
+\newpage{}
+
 \subsection{Tables}\label{tables}
@@
 
+\doublespacing
+\subsection{Figure Legends}
+
+FIGURE 1: Community-wide species synchrony of per capita growth rates from model simulation experiments.
+Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
++ No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
++ No E.S.'').
+Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
+Error bars represent the upper and lower 95\% quantiles from model simulations.
+All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.
+
+FIGURE 2: Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
+Only'') and when environmental stochasticity is also present removed (``D.S.
++ E.S.'').
+The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
+The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
+Theoretically, ``D.S.
+Only'' simulations should remain constant across landscape size, whereas ``D.S.
++ E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
+Error bars represent the upper and lower 95\% quantiles from model simulations.
+
+FIGURE 3: Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions).
+We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
+The dashed line is the line of equality.
+Simulation results in this figure are analogous to ``No Comp.
++ No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.
+1, but here we control the time series of random year effects.
+
+\newpage{}
+\singlespacing
+
 \subsection{Figures}\label{figures}
@@
 All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
-Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.}
 \end{figure}
@@
 The dashed line is the line of equality.
-Simulation results in this figure are analagous to ``No Comp.
+Simulation results in this figure are analogous to ``No Comp.
 + No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.

### 2016-03-24 docs/online_supporting_information.tex
updates figure formatting

@@
 \begin{equation}
-c_{ij} = \frac{B_{ij}}{1-A_ii A_jj} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
+c_{ij} = \frac{B_{ij}}{1-A_{ii} A_{jj}} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
 \end{equation}

### 2016-03-24 docs/online_supporting_information.tex
fixes fig s3 legend typo

@@
       \includegraphics[width=3in]{./components/formatted_figures/formatted_figureS4.png}
-  \caption{Community-wide synchrony of species percent cover from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
+  \caption{Community-wide synchrony of species' per capita growth rates from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
 Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.

### 2016-03-24 docs/online_supporting_information.Rmd
fixers typo in growth rate prediction equations

@@
 When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a].
-The prediction represented in Equation S20 is Equation 3 in the main text, which we refer to as $\phi_{R,\mathcal{M}_D}$.
+The prediction represented in Equation S19 is Equation 3 in the main text, which we refer to as $\phi_{R,\mathcal{M}_D}$.
 
@@
 \begin{equation}
-\phi_R = \frac{\sum_{i,j} p_i p_j \text{cov}(u_{ei}, u_{ej})}{S^2}
+\phi_R = \frac{\sum_{i,j} \text{cov}(u_{ei}, u_{ej})}{S^2}
 \end{equation}
 
-\noindent The prediction represented in Equation S22 is Equation 4 in the main text, which we refer to as $\phi_{R,\mathcal{M}_E}$.
+\noindent The prediction represented in Equation S21 is Equation 4 in the main text, which we refer to as $\phi_{R,\mathcal{M}_E}$.
 If we know the covariance matrix of the species environmental responses, we can calculate the above expectations directly.

### 2016-03-24 docs/online_supporting_information.tex
fixers typo in growth rate prediction equations

@@
 values equal 1/S (Loreau and {{de Mazancourt}} 2008). The prediction
-represented in Equation S20 is Equation 3 in the main text, which we
+represented in Equation S19 is Equation 3 in the main text, which we
 refer to as \(\phi_{R,\mathcal{M}_D}\).
@@
 \begin{equation}
-\phi_R = \frac{\sum_{i,j} p_i p_j \text{cov}(u_{ei}, u_{ej})}{S^2}
+\phi_R = \frac{\sum_{i,j} \text{cov}(u_{ei}, u_{ej})}{S^2}
 \end{equation}
 
-\noindent The prediction represented in Equation S22 is Equation 4 in
+\noindent The prediction represented in Equation S21 is Equation 4 in
 the main text, which we refer to as \(\phi_{R,\mathcal{M}_E}\). If we

### 2016-04-01 docs/ms_communitySynchrony.Rmd
fixers typo in growth rate prediction equations
adds text to ms on percent cover results

@@
 \begin{equation}
-\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}p_ip_j\text{cov}(u_{ei},u_{ej})}{S^2}
+\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(u_{ei},u_{ej})}{S^2}
 \end{equation}
@@
 For all five communities, the synchrony of species' growth rates in monoculture almost perfectly matched species synchrony in polyculture (Fig. 3).
-Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Figs. S2-S3).
+Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Online Supporting Information, Figs.
+S2-S3).
 
@@
 Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, though not perfectly, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.
-We also found that analyses based on synchrony of species' percent cover were uninformative (Figs.
-S2-S3), implying that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
+We also found that analyses based on synchrony of species' percent cover, rather than growth rates, were uninformative (Figs.
+S2-S3) since the linear approximation required for analytical predictions is a stronger assumption for abundance than growth rates, especially given relatively short time-series (Online Supporting Information).
+Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
 

### 2016-04-01 docs/ms_communitySynchrony.tex
fixers typo in growth rate prediction equations
adds text to ms on percent cover results

--- 2016-03-23 docs/ms_communitySynchrony.tex
@@
 similar, but simulation results were more variable and less consistent
-with analytical predictions and observed synchrony (Figs. S2-S3).
+with analytical predictions and observed synchrony (Online Supporting
+Information, Figs. S2-S3).
 
@@
 and therefore had little impact on community synchrony. We also found
-that analyses based on synchrony of species' percent cover were
-uninformative (Figs. S2-S3), implying that it is difficult to decipher
-mechanisms of species synchrony from abundance time series, as expected
-by theory (Loreau and {{de Mazancourt}} 2008).
+that analyses based on synchrony of species' percent cover, rather than
+growth rates, were uninformative (Figs. S2-S3) since the linear
+approximation required for analytical predictions is a stronger
+assumption for abundance than growth rates, especially given relatively
+short time-series (Online Supporting Information). Thus, our results
+provide further evidence that it is difficult to decipher mechanisms of
+species synchrony from abundance time series, as expected by theory
+(Loreau and {{de Mazancourt}} 2008).
 
@@
 
-Loreau, M. 2010. From Polutations to Ecosystems: Theoretical Fondations
+Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
 for a New Ecological Synthesis.
@@
 Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beauséjour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf. 2013.
-Global meta-analysis reveals no net change in local-scale plant
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
 biodiversity over time. Proceedings of the National Academy of Sciences

### 2016-04-01 docs/online_supporting_information.Rmd
adds results from percent cover section

@@
 
+Results for synchrony of percent cover
+--------------------------------------
+Synchrony of percent cover from population model simulations did not compare well with observed synchrony of cover or analytical predictions for synchrony cover (Fig. S2, S3).
+We did not expect synchrony of percent cover from simulations to compare well with observed synchrony of cover because our models represent equilibrium dynamics rather than a specific set of observed years.
+Also, unlike growth rates, synchrony in cover is highly influenced by drift and legacy effects [@Loreau2008a].
+
+More interesting is that our simulation results did not compare well with analytical predictions.
+One issue is that some species are not well-regulated around their equilibrium cover, so that our linear approximation required for anayltical predictions is likely to fail.
+Analytical predictions for synchrony of growth rates are more informative than synchrony in population sizes because the linear approximation is more proximate.
+
+Another issue is the influence of demographic stochasticity in some communities, for example, in Idaho where our simulations with only demographic stochasticity yielded results far different from analytical predicitons (Fig.
+S2).
+Our analytical predictions require the unrealistic assumption that demographic variances among species are equal.
+However, in Idaho, *A. tripartita* has much higher demographic stochasticity than the other species, so that variation in its abundance dominates.
+Thus, the assumption that all species have similar stochasticity fails.
+In combination, our results from analyzing synchrony of percent cover indicate that the best way to decipher the mechanisms contributing to community synchrony is to analyze growth rates, in agreement with the theoretical arguments of @Loreau2008a.
+
+\newpage{}
+
 Supporting Figures
@@
       \includegraphics[width=3in]{./components/formatted_figures/formatted_figureS4.png}
-  \caption{Community-wide synchrony of species percent cover from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
-Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
+  \caption{Community-wide synchrony of species' growth rates from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
+Synchrony of species' growth rates are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
 + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.

### 2016-04-01 docs/online_supporting_information.tex
adds results from percent cover section

@@
 \begin{equation}
-c_{ij} = \frac{B_{ij}}{1-A_{ii} A_{jj}} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
+c_{ij} = \frac{B_{ij}}{1-A_ii A_jj} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(u_{ei},u_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(u_{di},u_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
 \end{equation}
@@
 
+\subsection{Results for synchrony of percent
+cover}\label{results-for-synchrony-of-percent-cover}
+
+Synchrony of percent cover from population model simulations did not
+compare well with observed synchrony of cover or analytical predictions
+for synchrony cover (Fig. S2, S3). We did not expect synchrony of
+percent cover from simulations to compare well with observed synchrony
+of cover because our models represent equilibrium dynamics rather than a
+specific set of observed years. Also, unlike growth rates, synchrony in
+cover is highly influenced by drift and legacy effects (Loreau and {{de
+Mazancourt}} 2008).
+
+More interesting is that our simulation results did not compare well
+with analytical predictions. One issue is that some species are not
+well-regulated around their equilibrium cover, so that our linear
+approximation required for anayltical predictions is likely to fail.
+Analytical predictions for synchrony of growth rates are more
+informative than synchrony in population sizes because the linear
+approximation is more proximate.
+
+Another issue is the influence of demographic stochasticity in some
+communities, for example, in Idaho where our simulations with only
+demographic stochasticity yielded results far different from analytical
+predicitons (Fig. S2). Our analytical predictions require the
+unrealistic assumption that demographic variances among species are
+equal. However, in Idaho, \emph{A. tripartita} has much higher
+demographic stochasticity than the other species, so that variation in
+its abundance dominates. Thus, the assumption that all species have
+similar stochasticity fails. In combination, our results from analyzing
+synchrony of percent cover indicate that the best way to decipher the
+mechanisms contributing to community synchrony is to analyze growth
+rates, in agreement with the theoretical arguments of Loreau and {{de
+Mazancourt}} (2008).
+
+\newpage{}
+
 \subsection{Supporting Figures}\label{supporting-figures}
@@
       \includegraphics[width=3in]{./components/formatted_figures/formatted_figureS4.png}
-  \caption{Community-wide synchrony of species' per capita growth rates from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
-Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
+  \caption{Community-wide synchrony of species' growth rates from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
+Synchrony of species' growth rates are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
 + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
@@
   \hline
-New Mexico & 0.56 & 0.51 & 0.86 & 0.85 \\ 
-  Arizona & 0.51 & 0.52 & 0.82 & 0.81 \\ 
-  Kansas & 0.35 & 0.39 & 0.62 & 0.63 \\ 
+New Mexico & 0.56 & 0.51 & 0.86 & 0.86 \\ 
+  Arizona & 0.51 & 0.52 & 0.82 & 0.82 \\ 
+  Kansas & 0.35 & 0.39 & 0.62 & 0.62 \\ 
   Montana & 0.31 & 0.33 & 0.47 & 0.49 \\ 
@@
 \end{table}
-
-\newpage{}
 
@@
   \hline
-1No D.S. & 5.00 & 4.94 & 7.13 & 4.42 & 10.10 \\ 
-  2No Comp. + No D.S. & 4.25 & 2.08 & 7.61 & 4.39 & 8.10 \\ 
+1No D.S. & 4.82 & 3.99 & 7.19 & 5.37 & 9.53 \\ 
+  2No Comp. + No D.S. & 4.74 & 2.70 & 6.81 & 4.53 & 7.69 \\ 
   4No Comp. & 0.98 & 0.06 & 4.84 & 1.01 & 0.11 \\ 
@@
 
-\newpage{}
-
 \begin{table}[ht]
@@
 
-\newpage{}
-
-\subsubsection{Interaction Coefficients for Vital Rates in Arizona}
-
 \begin{table}[ht]
@@
 \end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in Arizona.} 
-\begin{tabular}{rrr}
-  \hline
- & BOER & BORO \\ 
-  \hline
-BOER & -0.2791 & -0.0816 \\ 
-  BORO & -0.0200 & -0.6101 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for recruitment regressions in Arizona.} 
-\begin{tabular}{rrr}
-  \hline
- & BOER & BORO \\ 
-  \hline
-BOER & -0.5304 & -0.3063 \\ 
-  BORO & -0.5798 & -2.3237 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\newpage{}
-
-\subsubsection{Interaction Coefficients for Vital Rates in Idaho}
-
-\begin{table}[ht]
+\end{table}\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}
-
-\begin{table}[ht]
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for growth regressions in Kansas.} 
+\begin{tabular}{rrrr}
+  \hline
+ & BOCU & BOHI & SCSC \\ 
+  \hline
+BOCU & -0.0017 & 0.0011 & 0.0010 \\ 
+  BOHI & 0.0002 & -0.0046 & -0.0078 \\ 
+  SCSC & -0.0003 & 0.0007 & -0.0124 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for growth regressions in Montana.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & BOGR & HECO & PASM & POSE \\ 
+  \hline
+BOGR & -0.0121 & -0.0002 & -0.0229 & -0.0014 \\ 
+  HECO & -0.0023 & -0.0778 & 0.0142 & 0.0061 \\ 
+  PASM & -0.0005 & -0.0039 & -0.0601 & -0.0033 \\ 
+  POSE & 0.0005 & -0.0101 & -0.0404 & -0.1431 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for growth regressions in NewMexico.} 
+\begin{tabular}{rrr}
+  \hline
+ & BOER & SPFL \\ 
+  \hline
+BOER & -0.0045 & -0.0013 \\ 
+  SPFL & -0.0041 & -0.0176 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in Arizona.} 
+\begin{tabular}{rrr}
+  \hline
+ & BOER & BORO \\ 
+  \hline
+BOER & -0.2791 & -0.0816 \\ 
+  BORO & -0.0200 & -0.6101 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}
-
-\begin{table}[ht]
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in Kansas.} 
+\begin{tabular}{rrrr}
+  \hline
+ & BOCU & BOHI & SCSC \\ 
+  \hline
+BOCU & -0.1087 & 0.0006 & 0.0010 \\ 
+  BOHI & 0.0021 & -0.0403 & -0.0153 \\ 
+  SCSC & -0.0064 & -0.0066 & -0.0270 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in Montana.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & BOGR & HECO & PASM & POSE \\ 
+  \hline
+BOGR & -0.1280 & 0.0211 & 0.1243 & 0.0173 \\ 
+  HECO & -0.0043 & -0.5661 & 0.1386 & 0.0013 \\ 
+  PASM & 0.0001 & -0.0044 & -1.3046 & 0.0012 \\ 
+  POSE & 0.0010 & 0.0085 & 0.1198 & -1.2434 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in NewMexico.} 
+\begin{tabular}{rrr}
+  \hline
+ & BOER & SPFL \\ 
+  \hline
+BOER & -0.1575 & -0.0040 \\ 
+  SPFL & -0.0120 & -0.2807 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for recruitment regressions in Arizona.} 
+\begin{tabular}{rrr}
+  \hline
+ & BOER & BORO \\ 
+  \hline
+BOER & -0.5304 & -0.3063 \\ 
+  BORO & -0.5798 & -2.3237 \\ 
+   \hline
+\end{tabular}
+\end{table}\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}
-
-\newpage{}
-
-\subsubsection{Interaction Coefficients for Vital Rates in Kansas}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for growth regressions in Kansas.} 
-\begin{tabular}{rrrr}
-  \hline
- & BOCU & BOHI & SCSC \\ 
-  \hline
-BOCU & -0.0017 & 0.0011 & 0.0010 \\ 
-  BOHI & 0.0002 & -0.0046 & -0.0078 \\ 
-  SCSC & -0.0003 & 0.0007 & -0.0124 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in Kansas.} 
-\begin{tabular}{rrrr}
-  \hline
- & BOCU & BOHI & SCSC \\ 
-  \hline
-BOCU & -0.1087 & 0.0006 & 0.0010 \\ 
-  BOHI & 0.0021 & -0.0403 & -0.0153 \\ 
-  SCSC & -0.0064 & -0.0066 & -0.0270 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
+\end{table}\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}
-
-\newpage{}
-
-\subsubsection{Interaction Coefficients for Vital Rates in Montana}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for growth regressions in Montana.} 
-\begin{tabular}{rrrrr}
-  \hline
- & BOGR & HECO & PASM & POSE \\ 
-  \hline
-BOGR & -0.0121 & -0.0002 & -0.0229 & -0.0014 \\ 
-  HECO & -0.0023 & -0.0778 & 0.0142 & 0.0061 \\ 
-  PASM & -0.0005 & -0.0039 & -0.0601 & -0.0033 \\ 
-  POSE & 0.0005 & -0.0101 & -0.0404 & -0.1431 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in Montana.} 
-\begin{tabular}{rrrrr}
-  \hline
- & BOGR & HECO & PASM & POSE \\ 
-  \hline
-BOGR & -0.1280 & 0.0211 & 0.1243 & 0.0173 \\ 
-  HECO & -0.0043 & -0.5661 & 0.1386 & 0.0013 \\ 
-  PASM & 0.0001 & -0.0044 & -1.3046 & 0.0012 \\ 
-  POSE & 0.0010 & 0.0085 & 0.1198 & -1.2434 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
+\end{table}\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}
-
-\newpage{}
-
-\subsubsection{Interaction Coefficients for Vital Rates in New Mexico}
-
-\begin{table}[ht]
+\end{table}\begin{table}[ht]
 \centering
-\caption{Interaction coefficients for growth regressions in New Mexico.} 
-\begin{tabular}{rrr}
-  \hline
- & BOER & SPFL \\ 
-  \hline
-BOER & -0.0045 & -0.0013 \\ 
-  SPFL & -0.0041 & -0.0176 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in New Mexico.} 
-\begin{tabular}{rrr}
-  \hline
- & BOER & SPFL \\ 
-  \hline
-BOER & -0.1575 & -0.0040 \\ 
-  SPFL & -0.0120 & -0.2807 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for recruitment regressions in New Mexico.} 
+\caption{Interaction coefficients for recruitment regressions in NewMexico.} 
 \begin{tabular}{rrr}

### 2016-04-01 docs/ms_communitySynchrony.Rmd
SUBMITTED VERSION TO ELE

@@
 output:
+  word_document: default
   pdf_document:
@@
     keep_tex: yes
-  word_document: default
 urlcolor: black
@@
 -------
-
 ```{r time_series_fig, echo=FALSE, include=TRUE, fig.width=8.5, fig.height=4, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site.

### 2016-04-01 docs/ms_communitySynchrony.tex
SUBMITTED VERSION TO ELE

@@
 
-\textbf{Number of Words}: Abstract = 138; Main Text = 4929 (including section titles and in-line citations)
+\textbf{Number of Words}: Abstract = 138; Main Text = 4965 (including section titles and in-line citations)
 
@@
 
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
+perennial plants: An embarrassment of niches. Ecology Letters
+13:1019--1029.
+
+Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
+mapped quadrats from Kansas prairie: demographic information for
+herbaceous plants. Ecology 88:2673.
+
+Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
+of semi-desert grassland plants in permanent quadrats mapped from 1915
+to 1947. Ecology 93:1492--1492.
+
+Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
+mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
+92:1703.
+
+Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
+P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
+Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
+and its impact on humanity. Nature 489:59--67.
+
+Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
+recruitment stage to stabilize grassland coexistence. Ecological
+Monographs 85:373--392.
+
+{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
+Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
+Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
+Loreau. 2013. Predicting ecosystem stability from community composition
+and biodiversity. Ecology Letters 16:617--625.
+
+Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
+Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
+biodiversity change but not systematic loss. Science 344:296--9.
+
+Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. Annual Review of
+Ecology, Evolution, and Systematics 40:393--414.
+
+Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
+variability in space and time. Methods in Ecology and Evolution
+5:524--533.
+
+Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
+Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
+temporal stability of biomass production: a new analysis of recent
+biodiversity experiments. The American naturalist 183:1--12.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
+on ecosystem functioning: a consensus of current knowledge. Ecological
+Monographs 75:3--35.
+
+Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
+Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
+Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
+Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
+Compensatory dynamics are rare in natural ecological communities.
+Proceedings of the National Academy of Sciences of the United States of
+America 104:3273--3277.
+
+Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
+productivity and the temporal stability of productivity: Patterns and
+processes. Ecology Letters 12:443--451.
+
+Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
+species interaction strength in assembled theoretical competition
+communities. Ecology Letters 2:70--74.
+
+Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
+dynamics in ecology and conservation.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
+for a New Ecological Synthesis.
+
+Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
+drivers: neutral and nonneutral community dynamics in fluctuating
+environments. The American Naturalist 172:E48--E66.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
+Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
+and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
+knowledge and future challenges. Science 294:804--808.
+
+Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
+theory, concepts, and empirical work. Ecological Monographs 81:429--441.
+
+Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
+community dynamics? Proceedings of the Royal Society B: Biological
+Sciences 276:2923--2929.
+
+Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
+Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
+3rd international workshop on distributed statistical computing (dSC
+2003). march.
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+
+R Core Team. 2013. R: A language and environment for statistical
+computing.
+
+Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
+O'Hara. 2008. Detecting compensatory dynamics in competitive communities
+under environmental forcing. Oikos 117:1907--1911.
+
+Rees, M., and S. P. Ellner. 2009. Integral projection models for
+populations in temporally varying environments. Ecological Monographs
+79:575--594.
+
+Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
+projection models: A user's guide.
+
+Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
+species have similar dynamics. Journal of Ecology 99:1453--1459.
+
+Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
+A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
+Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
+C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
+Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
+Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
+Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
+Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
+Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
+Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
+Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
+operating space for humanity. Nature 461:472--475.
+
+Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
+for Latent Gaussian Models Using Integrated Nested Laplace
+Approximations. Journal of the Royal Statistical Society, Series B
+71:319--392.
+
+Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
+Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
+analysis of latent Gaussian models using Integrated Nested Laplace
+Approximaxion.
+
+Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
+189:732--735.
+
+Tilman, D. 1988. Plant strategies and the dynamics and structure of
+plant communities. Plant strategies and the dynamics and structure of
+plant communities.:360 pp.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
+2001. Diversity and productivity in a long-term grassland experiment.
+Science 294:843--845.
+
+Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
+Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
+W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
+2014. Synchronous dynamics of zooplankton competitors prevail in
+temperate lake ecosystems. Proceedings. Biological sciences / The Royal
+Society 281:20140633.
+
+Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
+Beauséjour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf. 2013.
+Global meta-analysis reveals no net change in local-scale plant
+biodiversity over time. Proceedings of the National Academy of Sciences
+of the United States of America 110:19456--9.
+
+Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
+Human Domination of Earth' s Ecosystems. Science 277:494--499.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
+across scales in metacommunities. Ecology Letters.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
+sagebrush steppe: long-term data for analyzing demographic rates and
+plant--plant interactions. Ecology 91:3427.
+
+\newpage{}
+
+\subsection{Table Titles}\label{tables}
+TABLE 1: Site descriptions and focal species.
+
+TABLE 2: Observed synchrony among species' per capita growth rates ($\phi_{R}$), observed synchrony among species' percent cover ($\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community.
+Species richness values reflect the number of species analyzed from the community, not the actual richness.
+
+\newpage{}
+
 \subsection{Tables}\label{tables}
-
 \singlespacing
@@
 
+\subsection{Figures Captions}
+\doublespacing
+
+FIGURE 1: Community-wide species synchrony of per capita growth rates from model simulation experiments.
+Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
++ No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
++ No E.S.'').
+Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
+Error bars represent the upper and lower 95\% quantiles from model simulations.
+All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.
+
+FIGURE 2: Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
+Only'') and when environmental stochasticity is also present removed (``D.S.
++ E.S.'').
+The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
+The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
+Theoretically, ``D.S.
+Only'' simulations should remain constant across landscape size, whereas ``D.S.
++ E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
+Error bars represent the upper and lower 95\% quantiles from model simulations.
+
+FIGURE 3: Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions).
+We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
+The dashed line is the line of equality.
+Simulation results in this figure are analogous to ``No Comp.
++ No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.
+1, but here we control the time series of random year effects.
+
+\newpage{}
+
+\singlespacing
+
 \subsection{Figures}\label{figures}
@@
 
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
-Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
-perennial plants: An embarrassment of niches. Ecology Letters
-13:1019--1029.
-
-Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
-mapped quadrats from Kansas prairie: demographic information for
-herbaceous plants. Ecology 88:2673.
-
-Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
-of semi-desert grassland plants in permanent quadrats mapped from 1915
-to 1947. Ecology 93:1492--1492.
-
-Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
-mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
-92:1703.
-
-Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
-P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
-Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
-and its impact on humanity. Nature 489:59--67.
-
-Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
-recruitment stage to stabilize grassland coexistence. Ecological
-Monographs 85:373--392.
-
-{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
-Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
-Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
-Loreau. 2013. Predicting ecosystem stability from community composition
-and biodiversity. Ecology Letters 16:617--625.
-
-Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
-Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
-biodiversity change but not systematic loss. Science 344:296--9.
-
-Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
-Compensatory Dynamics in Ecological Communities. Annual Review of
-Ecology, Evolution, and Systematics 40:393--414.
-
-Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
-variability in space and time. Methods in Ecology and Evolution
-5:524--533.
-
-Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
-Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
-temporal stability of biomass production: a new analysis of recent
-biodiversity experiments. The American naturalist 183:1--12.
-
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
-
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
-
-Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
-on ecosystem functioning: a consensus of current knowledge. Ecological
-Monographs 75:3--35.
-
-Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
-Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
-Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
-Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
-Compensatory dynamics are rare in natural ecological communities.
-Proceedings of the National Academy of Sciences of the United States of
-America 104:3273--3277.
-
-Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
-productivity and the temporal stability of productivity: Patterns and
-processes. Ecology Letters 12:443--451.
-
-Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
-species interaction strength in assembled theoretical competition
-communities. Ecology Letters 2:70--74.
-
-Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
-dynamics in ecology and conservation.
-
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
-
-Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
-for a New Ecological Synthesis.
-
-Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
-drivers: neutral and nonneutral community dynamics in fluctuating
-environments. The American Naturalist 172:E48--E66.
-
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
-
-Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
-Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
-and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
-knowledge and future challenges. Science 294:804--808.
-
-Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
-theory, concepts, and empirical work. Ecological Monographs 81:429--441.
-
-Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
-community dynamics? Proceedings of the Royal Society B: Biological
-Sciences 276:2923--2929.
-
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
-
-R Core Team. 2013. R: A language and environment for statistical
-computing.
-
-Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
-O'Hara. 2008. Detecting compensatory dynamics in competitive communities
-under environmental forcing. Oikos 117:1907--1911.
-
-Rees, M., and S. P. Ellner. 2009. Integral projection models for
-populations in temporally varying environments. Ecological Monographs
-79:575--594.
-
-Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
-projection models: A user's guide.
-
-Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
-species have similar dynamics. Journal of Ecology 99:1453--1459.
-
-Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
-A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
-Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
-C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
-Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
-Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
-Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
-Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
-Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
-Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
-Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
-operating space for humanity. Nature 461:472--475.
-
-Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
-for Latent Gaussian Models Using Integrated Nested Laplace
-Approximations. Journal of the Royal Statistical Society, Series B
-71:319--392.
-
-Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
-Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
-analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
-
-Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
-189:732--735.
-
-Tilman, D. 1988. Plant strategies and the dynamics and structure of
-plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
-
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
-
-Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
-2001. Diversity and productivity in a long-term grassland experiment.
-Science 294:843--845.
-
-Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
-Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
-W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
-2014. Synchronous dynamics of zooplankton competitors prevail in
-temperate lake ecosystems. Proceedings. Biological sciences / The Royal
-Society 281:20140633.
-
-Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
-2013. Global meta-analysis reveals no net change in local-scale plant
-biodiversity over time. Proceedings of the National Academy of Sciences
-of the United States of America 110:19456--9.
-
-Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
-Human Domination of Earth' s Ecosystems. Science 277:494--499.
-
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
-
-Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
-
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
-
-Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
-sagebrush steppe: long-term data for analyzing demographic rates and
-plant--plant interactions. Ecology 91:3427.
-
 \end{document}

### 2016-04-04 docs/ms_communitySynchrony_PREPRINT.tex
adds preprint version of ms

--- 2016-04-01 docs/ms_communitySynchrony.tex
@@
 
-\usepackage{lineno}
-\linenumbers
 \usepackage{setspace}
 \usepackage{todonotes}
-\doublespacing
 \usepackage{color}
@@
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
+\normalsize{Andrew T.
+Tredennick\textsuperscript{1}\footnote{Correspondence: Email: atredenn@gmail.com}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B.
+Adler\textsuperscript{1}}
 
@@
 
-\vspace{2em}
-
-\textbf{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
-
-\vspace{2em}
-
-\textbf{Authorship}: ATT, CdM, and PBA designed the research.
-CdM and ML designed and analyzed the theoretical model.
-ATT and PBA performed the empirical analysis.
-All authors contributed to the interpretation and presentation of results.
-ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
-
-\textbf{Running Title}: Drivers of species synchrony in natural plant communities
-
-\textbf{Article Type}: Letter
-
-\textbf{Number of Words}: Abstract = 138; Main Text = 4965 (including section titles and in-line citations)
-
-\textbf{Number of References}: 46
-
-\textbf{Number of Tables and Figures}:  2 Tables;  3 Figures
-
-\textbf{Corresponding Author}: \\
-\hspace{2em}Andrew Tredennick \\  
-\hspace{2em}Department of Wildland Resources and the Ecology Center\\
-\hspace{2em}Utah State University \\
-\hspace{2em}5230 Old Main Hill \\
-\hspace{2em}Logan, Utah 84322 USA \\
-\hspace{2em}Phone: +1-970-443-1599 \\
-\hspace{2em}Fax: +1-435-797-3796 \\
-\hspace{2em}Email: atredenn@gmail.com
-
 \end{singlespace}\renewcommand*{\thefootnote}{\arabic{footnote}}
@@
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+
+\vspace{2em}
+
+\noindent{}\textbf{Keywords}\\
+\noindent{}synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
 \end{abstract}
 
-\newpage{} \setlength{\parindent}{5ex}
+\setlength{\parindent}{5ex}
 
@@
 gratefully acknowledged.
+
+\subsection{STATEMENT OF AUTHORSHIP}
+ATT, CdM, and PBA designed the research.
+CdM and ML designed and analyzed the theoretical model.
+ATT and PBA performed the empirical analysis.
+All authors contributed to the interpretation and presentation of results.
+ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
 

### 2016-04-04 docs/ms_communitySynchrony_PREPRINT.tex
adds peer review disclosure to preprint

@@
 \usepackage{rotating}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\chead{\color{red}THIS MANUSCRIPT HAS NOT UNDERGONE PEER REVIEW}
+\lhead{}
+\rhead{}
 
@@
 
+\thispagestyle{fancy}
 \begin{singlespace}
@@
 stochasticity), or fluctuating asynchronously because of random chance
-events (demographic stochasticity). (Isbell et al. 2009, Hector et al.
+events (demographic stochasticity) (Isbell et al. 2009, Hector et al.
 2010, {{de Mazancourt}} et al. 2013). Species richness affects synchrony

### 2016-05-20 R/recruitment_function.R
updates vital rate functions to report parameter correlations

@@ -10,7 +10,8 @@
 #' @return Matrix of statistical results by fitted parameter.
 
 recruit_mcmc <- function(dataframe, n_adapt=5000, n_update=10000, 
-                         n_samples=20000, n_thin=50, sppList){
+                         n_samples=20000, n_thin=50, sppList,
+                         fig_outfile){
   D <- dataframe
   # Calculate mean cover by group and year
   tmpD <- D[,c("quad","year","Group",paste("cov.",sppList,sep=""))]
@@ -91,6 +92,40 @@ recruit_mcmc <- function(dataframe, n_adapt=5000, n_update=10000,
                    inits = inits, n.adapt = n.Adapt)
   update(jm, n.iter=n.Up)
   out <- coda.samples(jm, variable.names=params, n.iter=n.Samp, n.thin=n.Thin)
+  out1 <- rbind(out[[1]],out[[2]],out[[3]])
+  toget1 <- grep("dd", colnames(out1))
+  toget2 <- grep("intcpt.mu", colnames(out1))
+  out.corrs <- out1[,c(toget1,toget2)]
+  library(IDPmisc)
+  
+  panel.hist <- function(x, ...)
+  {
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(usr[1:2], 0, 1.5) )
+    h <- hist(x, plot = FALSE)
+    breaks <- h$breaks; nB <- length(breaks)
+    y <- h$counts; y <- y/max(y)
+    rect(breaks[-nB], 0, breaks[-1], y, col="blue4", ...)
+  }
+  
+  panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
+  {
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(0, 1, 0, 1))
+    r <- abs(cor(x, y, method = "spearman"))
+    txt <- format(c(r, 0.123456789), digits=digits)[1]
+    txt <- paste(prefix, txt, sep="")
+    if(missing(cex.cor)) cex <- 0.8/strwidth(txt)
+    text(0.5, 0.5, txt, cex = cex * r)
+  }
+  
+  betterPairs <- function(YourData){
+    return(pairs(YourData, lower.panel=function(...) {par(new=TRUE);ipanel.smooth(...)}, diag.panel=panel.hist, upper.panel=panel.cor))
+  }
+  
+  png(filename = fig_outfile)
+  betterPairs(data.frame(out.corrs))
[…]

### 2016-05-20 R/vital_rate_regression_functions.R

@@ -34,7 +34,10 @@ get_growth_params <- function(dataframe, crowd_mat, alpha){
                   family=c("gaussian"), verbose=FALSE,
                   control.predictor = list(link = 1),
                   control.compute=list(dic=T,mlik=T),
-                  control.inla = list(h = 1e-10))
+                  control.inla = list(h = 1e-10),
+                  control.fixed = list(correlation.matrix=TRUE))
+  
+  tmp_fixed_covariance <- outINLA$misc$lincomb.derived.covariance.matrix
   
   # Fit variance
   x <- outINLA$summary.fitted.values$mean #fitted values from INLA
@@ -64,7 +67,7 @@ get_growth_params <- function(dataframe, crowd_mat, alpha){
   params$sigma.a=NA; params$sigma.a[1]=coef(outVar)[1] 
   params$sigma.b=NA; params$sigma.b[1]=coef(outVar)[2]
   
-  return(params)
+  return(list(params=params, covariance=tmp_fixed_covariance))
 }
 
 
@@ -106,7 +109,10 @@ get_survival_params <- function(dataframe, crowd_mat, alpha){
                   control.compute=list(dic=T,mlik=T),
                   control.predictor = list(link = 1),
                   control.inla = list(h = 1e-10),
-                  Ntrials=rep(1,nrow(D)))
+                  Ntrials=rep(1,nrow(D)),
+                  control.fixed = list(correlation.matrix=TRUE))
+  
+  tmp_fixed_covariance <- outINLA$misc$lincomb.derived.covariance.matrix
   
   #Collect parameters
   #random year and group effects
@@ -125,7 +131,7 @@ get_survival_params <- function(dataframe, crowd_mat, alpha){
   params$alpha=NA; params$alpha[1:length(alpha)]=alpha
   colnames(params)[5] <- "Intercept"
   
-  return(params)
+  return(list(params=params, covariance=tmp_fixed_covariance))
 }
 
 

### 2016-05-20 runscripts/call_recruitment_fxn.R

@@ -73,7 +73,8 @@ for(do_site in site_list){
   } # end species loop
   D[is.na(D)] <- 0  # replace missing values
   
-  recruit_params <- recruit_mcmc(dataframe = D, sppList = species_list)
+  outfile_now <- paste0("../results/param_covariance/recruitment_",do_site,"_",do_species,"_paramcorrs.png")
+  recruit_params <- recruit_mcmc(dataframe = D, sppList = species_list, fig_outfile = outfile_now)
   recruit_params_site_list[[do_site]] <- recruit_params
 } # end site loop
 

### 2016-05-20 runscripts/call_vitalrate_fxns.R

@@ -93,6 +93,13 @@ for(do_site in site_list){
                              crowd_mat = crowd_growth_now,
                              alpha = alpha_now)
     
+    # Save the parameter covariance matrix
+    tmp_fixed_covariance <- tmp[[2]]
+    tmp_file <- paste0(do_site,"_",do_species,"_fixedcovariance")
+    saveRDS(tmp_fixed_covariance, paste0("../results/param_covariance/growth_", tmp_file, ".RDS"))
+    
+    tmp <- tmp[[1]] # get just the parameter estimates
+    
     # Save in temporary list
     growth_params[[do_species]] <- tmp
   } #end species loop
@@ -174,6 +181,13 @@ for(do_site in site_list){
     tmp <- get_survival_params(dataframe = D,
                                crowd_mat = crowd_surv_now,
                                alpha = alpha_now)
+    
+    # Save the parameter covariance matrix
+    tmp_fixed_covariance <- tmp[[2]]
+    tmp_file <- paste0(do_site,"_",do_species,"_fixedcovariance")
+    saveRDS(tmp_fixed_covariance, paste0("../results/param_covariance/surv_", tmp_file, ".RDS"))
+    
+    tmp <- tmp[[1]] # get just the parameter estimates
     
     # Save in temporary list
     surv_params[[do_species]] <- tmp

### 2016-05-20 docs/revision_docs/response_to_reviews_ele.Rmd
adds response to reviewers doc

---
fontsize: 12pt
geometry: margin=1in
header-includes:
- \usepackage{setspace}
- \usepackage{color}
- \usepackage{rotating}
layout: 12pt
linkcolor: black
output:
  pdf_document:
    fig_caption: yes
    keep_tex: no
urlcolor: black
---

Dr. Thrall,

Thank you for the opportunity to submit a revised version of our manuscript "Environmental responses, not species interactions, determine species synchrony in natural plant communities" for consideration for publication in *Ecology Letters* as a Letter.

## Point-by-Point Responses to Reviewers
### Reviewer 1: Major Comments
1. \emph{Low species richness: The results focus on species-poor communities (N <= 4). However, the introduction and the discussion are framed in terms of the biodiversity loss due to global change, which generally refers to relatively species-rich communities. Some of these same authors have shown that species richness typically has a strong negative effect on the synchrony of species’ response to the environment and a weak positive effect on demographic stochasticity (e.g., Fig. 5 of de Mazancourt et al. 2013). Overall, these results suggest that in more species-rich communities, the role of species’ synchronous response to the environment might be greatly reduced and that of species interactions and/or demographic stochasticity increased, thus upending the authors’ conclusions. Hence, I think the authors need to mention this important caveat in the discussion section and potentially change their title to something more specific (e.g., “Environmental responses, not species interactions, determine species synchrony in **depauperate** plant communities”).}

We agree with the reviewer that our study is limited by the fact that we can only model common, co-occuring species due to sample size constraints.

2. \emph{Limited novelty: The main novelty in this manuscript is the derivation of two limiting cases using an existing model (Loreau and de Mazancourt 2008). The models fit to the data (IPMs and IBMs) have already been published elsewhere and are thus not novel. Hence, I'm left wondering if the novelty of the limiting cases is sufficiently high to warrant publication in Ecology Letters. One could argue that identifying the relative importance of demographic stochasticity, environmental variation and competition on species synchrony in natural communities is novel. However, given the issue raised in point 1 (potential biases due to low species richness), the generality of these results is likely to be limited to species-poor systems.}

While we are pleased the reviewer appreciates the novelty of the derivation of the limiting case predictions, we disagree that the empirical analysis (``models fit to the data'') is not novel.
The reviewer is of course correct that the data and models have been used elsewhere (Chu and Adler 2015, Adler et al. 2010), we contend that the novelty lies in how the models are applied, rather than the models in and of themselves. In particular, this is the first use of the Chu and Adler (2015) models and data to investigate the processes that determine ecosystem functioning (Chu and Adler focus on the maintenance of coexistence). More generally, to our knowledge, our paper represents the first instance of applying the tools of demographic population modeling to answer questions about ecosystem functioning.

The reviewer makes a good point about the generality of our results given the relatively low number of species we were able to model. But, please see our response above on the generality of our results despite the low species richness represented in our dynamic multi-species population models.

3. \emph{Independence of drivers: Do species interactions, demographic stochasticity and environmental variation have purely additive effects on species synchrony? Figure 1 suggests not. For instance, in the first panel (New Mexico), the height of the second bar (No comp + No DS = ES) plus the height of the last bar (No comp + No ES = DS) do not add up to the height of the fourth bar (No comp = DS + ES). Hence, there is a subadditive or antagonistic effect of these drivers (ES and DS) on species synchrony. In the presence of such interactions, identifying the relative contribution of each driver based on their independent effect when operating in isolation is likely to yield poor results. The only way to infer the relative importance of the drivers is to conduct an in silico factorial experiment on field-based IBM/IPM models (e.g., Figure 1). Hence, although the analytical approximations can be used to generate some interesting baseline results, they do not really provide insights about the drivers of synchrony in natural communities. If in the end we have to rely on digital experiments applied to field-parameterized models, what do the approximations really provide? The authors should feel free to push back here as I may very well be wrong…}

4. \emph{Linear approximation: The species appear to fluctuate significantly over time (0.19 < CV < 0.54 in Table 1) yet a linear approximation was used via Taylor expansion to estimate species synchrony under one of the limiting cases (pure environmental stochasticity; equation 4 and S21). Although linearization around the equilibrium is only expected to work for infinitesimally small deviations, the authors show remarkably good agreement between the synchrony values obtained from their field-based IPMs/IBMs and those predicted by their linear approximation. I think it’s important to discuss (1) the robustness of this approach for systems that experience mild perturbations from the equilibrium [I’m surprised it worked so well in this case] and (2) the potential limitations of this approximation for systems exhibiting stronger perturbations away from the equilibrium (e.g., CV > 1?). This point is briefly addressed when describing the applicability of the approximations for predicting the synchrony of species abundance (lines 318-321), but I think expanding the discussion would be worthwhile.}

### Reviewer 1: Minor Comments
1. \emph{Line 371-372: I would qualify this statement by mentioning that the results are unlikely to apply to species-rich plant communities. They may be broadly applicable to species-poor communities though.}

2. \emph{Figure 1 caption: Awkward phrasing for the sentence beginning with ``Error bars represent the upper and lower 95\% quantiles from the model simulations''. It should read something like ``Error bars represent the 2.5\% and the 97.5\% quantiles from the model simulations'', as there is no such as thing as ‘lower’ and ‘upper’ 95\% quantiles.}

Good catch. We have edited the phrasing.

### Reviewer 2: Major Comments
1. \emph{How does the oversimplification of these plant systems into 2, 3 and 4 species communities affect the accuracy and the relevance of the models? The authors admit this constraint and discuss some potential biases this may have created in their results in lns 389-401. The authors could expand this discussion even more more by speculating how synchrony and the importance of environmental stochasticity might change in communities that have more species, and also in communities where additional the additional species richness is not just due to rare species. For example, if communities have 8 species that are all relatively equally abundant, is interspecific competition likely to become more important, and will environmental stochasticity have a similar effect or different effect in more diverse and complex systems?}

2. \emph{The results are rather obvious that environmental stochasticity will be the major driver of synchrony in these systems because the common species in these communities appear to have very weak interspecific interactions (and the authors admit to this, e.g. lns 417-418). The somewhat obvious result may make the complex modeling almost seem unnecessary. However, the approach the authors take is one that can be applied to other systems and other long-term data sets. And the models help clarify and quantify the relative importance of the different factors in an elegant way. Finally, the congruence between the IPM and IBM simulations and the observed synchrony in real systems is very compelling and suggests that the simulation experiments are an effective way to assess the relative importance of these various factors for synchrony in these systems (Fig 1, Table S1).}

3. \emph{What do these results mean for communities that do not resemble these plant communities? The authors are careful to not overgeneralize their results and they consistently discuss how their results are applicable to plant communities that resemble these systems (e.g. lns 368-372). However, the authors also explain how this study is applicable and of broader interest for understanding synchrony in communities that differ from these plant communities, e.g. consumer-resource models (lns 415-417) and rapidly assembled experimental plant communities, (lns 419-430) where interspecific interactions are much stronger. But more might be discussed here (see comment \#4 below).}

4. \emph{In the plant systems studied here, niche differentiation is strong (Chu and Adler 2015), resulting in weak interspecific competition (e.g. Table S4). Thus, species growth rates reflect environmental stochasticity because growth is influenced more by abiotic forces such as precipitation and temperature as opposed to competition for resources, or predation. How common is this scenario in other communities? In many other systems, biotic interactions are likely to be much more important in driving population dynamics (e.g. predator-prey, consumer-resource oscillations). In communities where biotic interactions are strong, abiotic forces such as environmental stochasticity may have a different effect on synchrony. Can the authors comment on this?}

5. \emph{Most discussions of synchrony are presented in the context of how synchrony contributes to ecological stability. In this study, synchrony was expressed in terms of growth rates rather than abundance or biomass, making a clear connection between synchrony of different species to overall stability of plant communities more difficult. Based on the role of synchronous dynamics for ecological stability, one would predict that as plant synchrony goes up, the stability of these systems should decrease. In this study, the only measure of ecological stability that was presented was the CV of above ground plant biomass (Table 2), and the authors do not discuss this data in any detail. From the data, it appears there is no connection between synchrony and plant above ground biomass stability, but the authors do not raise this issue anywhere in the manuscript, other than to say that using abundance data to detect synchrony is difficult and uninformative. (lns 317-323). Can the authors provide any additional commentary or explanation of the apparent lack of relationship between synchrony of plant growth rates and CV of above ground plant biomass?}

### Reviewer 2: Minor Comments
1. \emph{Ln 142-132, Equn 3.  It would be helpful to define $phi_R$ as synchrony in the text.
Lns 263-264. It should be Equations 4 and Equations 3 (not Equn 3 and Equn 2).}

2. \emph{Extra period (.) on line 52.}

Period is removed.

3. \emph{Last sentence paragraph 1, lns 58-59. Perhaps you want to focus this sentence a little ‘Predicting the impacts of global changes for the stability of ecosystems requires a mechanistic ….’}

4. \emph{In Online Supporting Information: Lns 22, ‘number’ is misspelled.}

Thanks for taking the time to review the supporting information! Spelling fixed.

5. \emph{Figure S3. It would be helpful to have a fuller figure legend here that includes reference to ‘the horizontal dashed and dotted lines …’ like you have in Figure 2 in the main text.}

We have added a fuller figure legend.

6. \emph{Table S5-S19. You do not define the shorthand species names anywhere, e.g. BOER, BORO, etc. It would be helpful to have a legend somewhere, or refer the reader to Table 1 for the full species names.}

Good catch. We have added a guide for readers before the tables and refer the reader to Table 1.

### Reviewer 3: Major Comments
1. \emph{First, I question the validity of virtual experiments where, unrealistically, species interactions and other sources of stochasticity are deactivated by setting some parameters to zero while leaving the others unchanged. Because the survival, growth and recruitment models are GLMs fitted to empirical data, one can expect the parameters to exhibit some correlation: changing the magnitude of some parameters would affect that of other parameters. A more credible analysis could be conducted by refitting the models, for example, without species interactions. This however would not solve the problem, because the mere presence of a species can influence the parameters of another species even if they do not interact directly. For example, if a bush overshadows a grass, it could have a detrimental effect on the growth of the grass but mitigate the effect of dry years on the grass, therefore changing the way it responds to environmental fluctuations. Indeed, there are many examples in the literature where species interactions are competitive in some way and facilitative in other ways. Unfortunately the only way to assess the response of a species to environmental fluctuations when grown in isolation is to grow it in isolation. I think this is a major flaw of the study that can be addressed only by providing some empirical evidence of the validity of the approach.}

This is a valid concern, and we thank the reviewer for bringing it to our attention. As the reviewer notes, the only way to disentangle the determinants of species synchrony in natural communities is with in silico simulations from empirically-based models. Thus, to gain any understanding of synchrony in natural communities means we must accept certain limitations of the approach. That said, we are in effect performing a sensitivity analysis: how much does synchrony change when we alter a given parameter? In our case, instead of slightly perturbing parameters, as is the case in many sensitivity analyses, we simply "turn them off". The reviewer's concern did prompt us to investigate the correlation between key parameters in our model. Fortunately, most parameters are uncorrelated, meaning that the magnitude of one coefficient does not affect the magnitude of another coefficient. Thus, our in silico experiments are valid and robust. We have added text about this to the Results section.

2. \emph{Second, the role of demographic stochasticity, although weak, is very likely overestimated in these analyses. This is because the intensity of demographic stochasticity is not estimated from empirical data, but from simulations on an arbitrary quadrat size of 25 m2. Demographic stochasticity is stronger in small populations and the population in a quadrat of 25m2 is very likely small compared to the actual population size. There is a simple way to check for this: using the parameters for survival and recruitment estimated from the data, in conjunction with the stable size distribution estimated by the IPM, one could calculate the intensity of demographic stochasticity as a function of quadrat size. This analysis should be conducted, and I expect it to show that the influence of demographic stochasticity on community stability is negligible in realistically sized grassland populations.}

We are a little confused by this comment. In fact, we show that demographic stochasticity is a function of quadrat size in Figure 2 (as suggested by the reviewer). Likewise, given the close correspondence between synchrony from IPM simulations (which approximate an infinitely large population with no demographic stochasticity) and synchrony from IBM simulations in a 25m2 quadrat (Fig. S1), we conclude that demographic is not overestimated.

3. \emph{Third, the nearly perfect match between the simplified synchrony metrics and the simulated community synchrony might be an artefact due to the way the data has been simulated and the metrics calculated. Specifically, eq. 4 defines the “environmental stochasticity only” synchrony index. To calculate this index the correlations of responses to environmental fluctuations, cov(uei, uej) between species pairs are needed. However this quantities are unknown, and cannot be estimated easily from empirical data, as acknowledged by the authors. Consequently, they used the correlations of growth rates from the IPM simulation with environmental stochasticity only to
parametrize the equations. In a second step, they use the same simulated growth rates correlations to calculate the community synchrony index (eq. 5). A quick look at eq 4. and eq 5. shows that these two values are inevitably very similar when calculated on the same data, unless there is extreme unevenness in environmental variances among species. Indeed when all the species have the same environmental variance, the two indices are equal. The problem, of course, is that eq. 4 needs to be parametrized with the covariance of the responses of environmental fluctuations instead and that these are unknown.}

4. \emph{A similar reasoning can be applied to eq. 3, which can be derived directly from eq. 5 by assuming that all species have the same demographic variance, except for a factor proportional to the inverse of species abundance. This normal assumption, derived by Lande, will necessarily be met in IBM simulations, as a consequence of the central limit theorem: demographic variance results from the sum of the variances of a large number of individuals. Therefore, in a similar way that the environmental synchrony index, the demographic synchrony index and community synchrony index will be very similar unless strong unevenness in demographic variance among species is present. In a nutshell, the match of the limiting case indices and synchrony indices on simulated data is a statistical inevitability.}

### Reviewer 3: Minor Comments
1. \emph{P3, line61: interspecific interactions are not limited to competition, and both predator-prey and facilitation have been shown to potentially play a role in community synchrony (Ives 2001)}

2. \emph{P4, line 70: I know at least of two studies that have addressed this exact question: (Mutshinda et al. 2009, Thibaut et al. 2012).}

3. \emph{P 4. L 76-77: this is unnecessarily controversial. Estimating the relative strength of each driver from time-series is indeed possible by fitting a community dynamics model to the data. What is impossible, as argued by Ranta, is to do so from the covariance matrix of species abundances as Houlahan did.}

Fair point. We have replace this sentence.

4. \emph{P 15 L 338-339: demographic stochasticity can never counteract the effects of environmental fluctuations, because demographic fluctuations are independent among species. Adding a positive value to the diagonal of a covariance matrix can reduce correlations but not eliminate them.}

Good catch.

### 2016-05-26 docs/online_supporting_information.Rmd
updates parameter corr table generator

@@
 all_files <- list.files("../results/")
-param_files <- all_files[grep("param*", all_files)]
+param_files1 <- all_files[grep("param*", all_files)]
+param_files <- param_files1[grep("RDS", param_files1)]
 num_vitals <- length(param_files)
@@
 result_files <- list.files("../results/")
-vr_files <-result_files[grep(paste(vital_rates,collapse="|"), 
+vr_files1 <-result_files[grep(paste(vital_rates,collapse="|"), 
                               list.files("../results/"))]
+vr_files <- vr_files1[grep("RDS", vr_files1)]
 

### 2016-05-26 docs/online_supporting_information.tex
updates parameter corr table generator

--- 2016-04-01 docs/online_supporting_information.tex
@@
   \hline
-Arizona & Interspecific & -0.0111 & -0.4430 & -0.0508 \\ 
-  Arizona & Intraspecific & -0.0402 & -1.4271 & -0.4446 \\ 
-  Idaho & Interspecific & 0.0022 & -0.1156 & 0.0059 \\ 
-  Idaho & Intraspecific & -0.0469 & -1.5581 & -0.5269 \\ 
-  Kansas & Interspecific & -0.0008 & 0.0509 & -0.0041 \\ 
-  Kansas & Intraspecific & -0.0062 & -0.9916 & -0.0587 \\ 
-  Montana & Interspecific & -0.0053 & 0.0922 & 0.0354 \\ 
-  Montana & Intraspecific & -0.0733 & -2.0499 & -0.8105 \\ 
-  NewMexico & Interspecific & -0.0027 & -0.3441 & -0.0080 \\ 
-  NewMexico & Intraspecific & -0.0111 & -1.6425 & -0.2191 \\ 
+Arizona & Interspecific & -0.0111 & -0.4534 & -0.0507 \\ 
+  Arizona & Intraspecific & -0.0402 & -1.4209 & -0.4446 \\ 
+  Idaho & Interspecific & 0.0022 & -0.1091 & 0.0059 \\ 
+  Idaho & Intraspecific & -0.0469 & -1.5569 & -0.5269 \\ 
+  Kansas & Interspecific & -0.0008 & 0.0463 & -0.0041 \\ 
+  Kansas & Intraspecific & -0.0062 & -0.9945 & -0.0586 \\ 
+  Montana & Interspecific & -0.0053 & 0.0920 & 0.0354 \\ 
+  Montana & Intraspecific & -0.0733 & -2.0492 & -0.8103 \\ 
+  NewMexico & Interspecific & -0.0027 & -0.3461 & -0.0080 \\ 
+  NewMexico & Intraspecific & -0.0111 & -1.6473 & -0.2191 \\ 
    \hline
@@
   SPFL & -0.0684 & -2.5220 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Arizona, vital rate = growth, and species = BOER.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.34E-01 & -1.15E-01 & -9.09E-02 \\ 
+  logarea.t0 & -1.34E-01 & 1.00E+00 & -2.09E-02 & 3.67E-02 \\ 
+  crowdV1 & -1.15E-01 & -2.09E-02 & 1.00E+00 & 1.02E-01 \\ 
+  crowdV2 & -9.09E-02 & 3.67E-02 & 1.02E-01 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Arizona, vital rate = growth, and species = BORO.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -5.64E-02 & -7.69E-02 & -7.72E-02 \\ 
+  logarea.t0 & -5.64E-02 & 1.00E+00 & 3.00E-02 & 3.60E-04 \\ 
+  crowdV1 & -7.69E-02 & 3.00E-02 & 1.00E+00 & 6.23E-02 \\ 
+  crowdV2 & -7.72E-02 & 3.60E-04 & 6.23E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = growth, and species = ARTR.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -8.98E-02 & -4.69E-02 & -4.01E-02 & -2.91E-02 & -7.15E-02 \\ 
+  logarea.t0 & -8.98E-02 & 1.00E+00 & 3.69E-02 & 5.78E-03 & 5.03E-05 & -3.27E-03 \\ 
+  crowdV1 & -4.69E-02 & 3.69E-02 & 1.00E+00 & -1.67E-02 & -3.06E-02 & -1.05E-03 \\ 
+  crowdV2 & -4.01E-02 & 5.78E-03 & -1.67E-02 & 1.00E+00 & -2.54E-02 & 8.19E-02 \\ 
+  crowdV3 & -2.91E-02 & 5.03E-05 & -3.06E-02 & -2.54E-02 & 1.00E+00 & 3.12E-02 \\ 
+  crowdV4 & -7.15E-02 & -3.27E-03 & -1.05E-03 & 8.19E-02 & 3.12E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = growth, and species = HECO.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.25E-01 & -4.30E-02 & -1.10E-01 & -6.36E-02 & -1.03E-01 \\ 
+  logarea.t0 & -1.25E-01 & 1.00E+00 & 4.95E-03 & 5.15E-02 & 2.80E-03 & 6.47E-02 \\ 
+  crowdV1 & -4.30E-02 & 4.95E-03 & 1.00E+00 & -1.11E-02 & 2.73E-02 & 6.86E-02 \\ 
+  crowdV2 & -1.10E-01 & 5.15E-02 & -1.11E-02 & 1.00E+00 & -5.17E-02 & 5.57E-02 \\ 
+  crowdV3 & -6.36E-02 & 2.80E-03 & 2.73E-02 & -5.17E-02 & 1.00E+00 & 1.17E-02 \\ 
+  crowdV4 & -1.03E-01 & 6.47E-02 & 6.86E-02 & 5.57E-02 & 1.17E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = growth, and species = POSE.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -9.56E-02 & -4.68E-02 & -8.67E-02 & -7.52E-02 & -1.32E-01 \\ 
+  logarea.t0 & -9.56E-02 & 1.00E+00 & 1.29E-02 & 1.77E-03 & 9.43E-03 & 4.44E-02 \\ 
+  crowdV1 & -4.68E-02 & 1.29E-02 & 1.00E+00 & 4.50E-02 & 7.39E-03 & 5.64E-02 \\ 
+  crowdV2 & -8.67E-02 & 1.77E-03 & 4.50E-02 & 1.00E+00 & -5.26E-03 & 7.84E-02 \\ 
+  crowdV3 & -7.52E-02 & 9.43E-03 & 7.39E-03 & -5.26E-03 & 1.00E+00 & 1.48E-02 \\ 
+  crowdV4 & -1.32E-01 & 4.44E-02 & 5.64E-02 & 7.84E-02 & 1.48E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = growth, and species = PSSP.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -8.15E-02 & -4.77E-02 & -5.04E-02 & -6.31E-02 & -8.25E-02 \\ 
+  logarea.t0 & -8.15E-02 & 1.00E+00 & 3.28E-02 & 3.53E-02 & 1.95E-02 & 3.99E-02 \\ 
+  crowdV1 & -4.77E-02 & 3.28E-02 & 1.00E+00 & 3.60E-02 & 1.27E-02 & -9.47E-03 \\ 
+  crowdV2 & -5.04E-02 & 3.53E-02 & 3.60E-02 & 1.00E+00 & -1.01E-02 & 1.52E-02 \\ 
+  crowdV3 & -6.31E-02 & 1.95E-02 & 1.27E-02 & -1.01E-02 & 1.00E+00 & 2.27E-02 \\ 
+  crowdV4 & -8.25E-02 & 3.99E-02 & -9.47E-03 & 1.52E-02 & 2.27E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Kansas, vital rate = growth, and species = BOCU.} 
+\begin{tabular}{rrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.27E-01 & -7.88E-02 & -6.53E-02 & -2.72E-02 \\ 
+  logarea.t0 & -1.27E-01 & 1.00E+00 & 7.37E-02 & 8.64E-02 & 4.48E-02 \\ 
+  crowdV1 & -7.88E-02 & 7.37E-02 & 1.00E+00 & -1.12E-02 & 2.39E-02 \\ 
+  crowdV2 & -6.53E-02 & 8.64E-02 & -1.12E-02 & 1.00E+00 & 6.34E-02 \\ 
+  crowdV3 & -2.72E-02 & 4.48E-02 & 2.39E-02 & 6.34E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Kansas, vital rate = growth, and species = BOHI.} 
+\begin{tabular}{rrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.27E-01 & -3.75E-02 & -3.16E-02 & -3.57E-02 \\ 
+  logarea.t0 & -1.27E-01 & 1.00E+00 & 5.16E-02 & 7.55E-02 & 1.74E-01 \\ 
+  crowdV1 & -3.75E-02 & 5.16E-02 & 1.00E+00 & -6.11E-02 & 3.83E-02 \\ 
+  crowdV2 & -3.16E-02 & 7.55E-02 & -6.11E-02 & 1.00E+00 & 6.58E-02 \\ 
+  crowdV3 & -3.57E-02 & 1.74E-01 & 3.83E-02 & 6.58E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Kansas, vital rate = growth, and species = SCSC.} 
+\begin{tabular}{rrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -2.59E-01 & -1.43E-01 & -7.78E-02 & -9.80E-02 \\ 
+  logarea.t0 & -2.59E-01 & 1.00E+00 & 2.02E-01 & 1.85E-01 & 2.40E-01 \\ 
+  crowdV1 & -1.43E-01 & 2.02E-01 & 1.00E+00 & 4.84E-02 & 9.45E-02 \\ 
+  crowdV2 & -7.78E-02 & 1.85E-01 & 4.84E-02 & 1.00E+00 & 2.75E-02 \\ 
+  crowdV3 & -9.80E-02 & 2.40E-01 & 9.45E-02 & 2.75E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = growth, and species = BOGR.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -4.52E-02 & -7.52E-02 & -3.67E-02 & -4.32E-02 & -2.79E-02 \\ 
+  logarea.t0 & -4.52E-02 & 1.00E+00 & 2.29E-02 & 1.73E-02 & 1.72E-02 & 2.35E-03 \\ 
+  crowdV1 & -7.52E-02 & 2.29E-02 & 1.00E+00 & 3.98E-02 & -6.01E-02 & -6.43E-03 \\ 
+  crowdV2 & -3.67E-02 & 1.73E-02 & 3.98E-02 & 1.00E+00 & 3.07E-02 & 2.20E-02 \\ 
+  crowdV3 & -4.32E-02 & 1.72E-02 & -6.01E-02 & 3.07E-02 & 1.00E+00 & 2.40E-02 \\ 
+  crowdV4 & -2.79E-02 & 2.35E-03 & -6.43E-03 & 2.20E-02 & 2.40E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = growth, and species = HECO.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -2.31E-02 & -8.93E-02 & -5.50E-02 & -9.75E-02 & -4.23E-02 \\ 
+  logarea.t0 & -2.31E-02 & 1.00E+00 & 1.75E-02 & -2.02E-02 & 1.17E-03 & 1.38E-02 \\ 
+  crowdV1 & -8.93E-02 & 1.75E-02 & 1.00E+00 & 9.21E-02 & 5.16E-02 & -1.44E-02 \\ 
+  crowdV2 & -5.50E-02 & -2.02E-02 & 9.21E-02 & 1.00E+00 & 3.90E-02 & -3.61E-03 \\ 
+  crowdV3 & -9.75E-02 & 1.17E-03 & 5.16E-02 & 3.90E-02 & 1.00E+00 & 3.92E-02 \\ 
+  crowdV4 & -4.23E-02 & 1.38E-02 & -1.44E-02 & -3.61E-03 & 3.92E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = growth, and species = PASM.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & 1.65E-02 & -6.08E-02 & -4.11E-02 & -1.10E-01 & -2.98E-02 \\ 
+  logarea.t0 & 1.65E-02 & 1.00E+00 & 4.15E-03 & 4.42E-03 & -7.75E-03 & 1.04E-02 \\ 
+  crowdV1 & -6.08E-02 & 4.15E-03 & 1.00E+00 & 9.00E-02 & 8.59E-02 & -3.36E-02 \\ 
+  crowdV2 & -4.11E-02 & 4.42E-03 & 9.00E-02 & 1.00E+00 & 2.85E-02 & 1.31E-02 \\ 
+  crowdV3 & -1.10E-01 & -7.75E-03 & 8.59E-02 & 2.85E-02 & 1.00E+00 & 3.59E-03 \\ 
+  crowdV4 & -2.98E-02 & 1.04E-02 & -3.36E-02 & 1.31E-02 & 3.59E-03 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = growth, and species = POSE.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -4.09E-02 & -6.28E-02 & -3.56E-02 & -8.91E-02 & -4.55E-02 \\ 
+  logarea.t0 & -4.09E-02 & 1.00E+00 & -8.02E-03 & 2.17E-02 & 4.47E-02 & 1.00E-02 \\ 
+  crowdV1 & -6.28E-02 & -8.02E-03 & 1.00E+00 & 1.03E-01 & 6.54E-02 & 9.60E-03 \\ 
+  crowdV2 & -3.56E-02 & 2.17E-02 & 1.03E-01 & 1.00E+00 & 5.30E-02 & 4.18E-02 \\ 
+  crowdV3 & -8.91E-02 & 4.47E-02 & 6.54E-02 & 5.30E-02 & 1.00E+00 & 2.95E-02 \\ 
+  crowdV4 & -4.55E-02 & 1.00E-02 & 9.60E-03 & 4.18E-02 & 2.95E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = NewMexico, vital rate = growth, and species = BOER.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.33E-01 & -1.25E-01 & -3.79E-02 \\ 
+  logarea.t0 & -1.33E-01 & 1.00E+00 & 3.30E-02 & 2.14E-02 \\ 
+  crowdV1 & -1.25E-01 & 3.30E-02 & 1.00E+00 & 5.46E-02 \\ 
+  crowdV2 & -3.79E-02 & 2.14E-02 & 5.46E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = NewMexico, vital rate = growth, and species = SPFL.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea.t0 & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -3.24E-01 & -1.29E-01 & -9.84E-02 \\ 
+  logarea.t0 & -3.24E-01 & 1.00E+00 & 3.89E-02 & 4.28E-02 \\ 
+  crowdV1 & -1.29E-01 & 3.89E-02 & 1.00E+00 & 3.82E-02 \\ 
+  crowdV2 & -9.84E-02 & 4.28E-02 & 3.82E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Arizona, vital rate = surv, and species = BOER.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.36E-01 & -9.25E-02 & -9.08E-02 \\ 
+  logarea & -1.36E-01 & 1.00E+00 & -3.15E-02 & 1.56E-02 \\ 
+  crowdV1 & -9.25E-02 & -3.15E-02 & 1.00E+00 & 3.54E-02 \\ 
+  crowdV2 & -9.08E-02 & 1.56E-02 & 3.54E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Arizona, vital rate = surv, and species = BORO.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -8.81E-03 & -2.77E-02 & -5.40E-02 \\ 
+  logarea & -8.81E-03 & 1.00E+00 & 1.02E-02 & -1.25E-02 \\ 
+  crowdV1 & -2.77E-02 & 1.02E-02 & 1.00E+00 & 5.37E-02 \\ 
+  crowdV2 & -5.40E-02 & -1.25E-02 & 5.37E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = surv, and species = ARTR.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.73E-01 & -2.05E-01 & -6.82E-02 & -7.73E-02 & -2.43E-01 \\ 
+  logarea & -1.73E-01 & 1.00E+00 & 1.66E-04 & -2.14E-02 & 1.31E-02 & 6.83E-02 \\ 
+  crowdV1 & -2.05E-01 & 1.66E-04 & 1.00E+00 & -2.11E-02 & -2.09E-02 & 2.15E-02 \\ 
+  crowdV2 & -6.82E-02 & -2.14E-02 & -2.11E-02 & 1.00E+00 & -3.73E-04 & 3.67E-02 \\ 
+  crowdV3 & -7.73E-02 & 1.31E-02 & -2.09E-02 & -3.73E-04 & 1.00E+00 & 1.82E-02 \\ 
+  crowdV4 & -2.43E-01 & 6.83E-02 & 2.15E-02 & 3.67E-02 & 1.82E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = surv, and species = HECO.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & 6.74E-02 & -8.51E-02 & -1.31E-01 & -8.00E-02 & -1.08E-01 \\ 
+  logarea & 6.74E-02 & 1.00E+00 & 1.65E-02 & -2.28E-02 & -1.04E-02 & 1.87E-02 \\ 
+  crowdV1 & -8.51E-02 & 1.65E-02 & 1.00E+00 & -8.42E-03 & 1.36E-02 & 1.27E-01 \\ 
+  crowdV2 & -1.31E-01 & -2.28E-02 & -8.42E-03 & 1.00E+00 & 2.71E-03 & 3.61E-02 \\ 
+  crowdV3 & -8.00E-02 & -1.04E-02 & 1.36E-02 & 2.71E-03 & 1.00E+00 & -8.60E-03 \\ 
+  crowdV4 & -1.08E-01 & 1.87E-02 & 1.27E-01 & 3.61E-02 & -8.60E-03 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = surv, and species = POSE.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & 5.11E-02 & -6.64E-02 & -5.27E-02 & -7.80E-02 & -1.01E-01 \\ 
+  logarea & 5.11E-02 & 1.00E+00 & 1.46E-02 & 6.55E-03 & -2.42E-02 & 1.35E-03 \\ 
+  crowdV1 & -6.64E-02 & 1.46E-02 & 1.00E+00 & 6.15E-02 & -2.11E-02 & 6.79E-02 \\ 
+  crowdV2 & -5.27E-02 & 6.55E-03 & 6.15E-02 & 1.00E+00 & 5.89E-03 & 6.17E-02 \\ 
+  crowdV3 & -7.80E-02 & -2.42E-02 & -2.11E-02 & 5.89E-03 & 1.00E+00 & -1.04E-01 \\ 
+  crowdV4 & -1.01E-01 & 1.35E-03 & 6.79E-02 & 6.17E-02 & -1.04E-01 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Idaho, vital rate = surv, and species = PSSP.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & 6.03E-02 & -6.59E-02 & -2.53E-02 & -5.16E-02 & -9.37E-02 \\ 
+  logarea & 6.03E-02 & 1.00E+00 & 1.46E-02 & 7.86E-03 & -3.97E-03 & -4.80E-02 \\ 
+  crowdV1 & -6.59E-02 & 1.46E-02 & 1.00E+00 & 1.28E-02 & 1.80E-02 & 3.74E-02 \\ 
+  crowdV2 & -2.53E-02 & 7.86E-03 & 1.28E-02 & 1.00E+00 & -4.17E-03 & 2.34E-02 \\ 
+  crowdV3 & -5.16E-02 & -3.97E-03 & 1.80E-02 & -4.17E-03 & 1.00E+00 & 2.05E-02 \\ 
+  crowdV4 & -9.37E-02 & -4.80E-02 & 3.74E-02 & 2.34E-02 & 2.05E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Kansas, vital rate = surv, and species = BOCU.} 
+\begin{tabular}{rrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -2.04E-01 & -5.64E-02 & -7.87E-02 & -5.55E-02 \\ 
+  logarea & -2.04E-01 & 1.00E+00 & -2.86E-02 & 3.25E-02 & 1.81E-02 \\ 
+  crowdV1 & -5.64E-02 & -2.86E-02 & 1.00E+00 & -3.70E-02 & 4.25E-02 \\ 
+  crowdV2 & -7.87E-02 & 3.25E-02 & -3.70E-02 & 1.00E+00 & 1.10E-01 \\ 
+  crowdV3 & -5.55E-02 & 1.81E-02 & 4.25E-02 & 1.10E-01 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Kansas, vital rate = surv, and species = BOHI.} 
+\begin{tabular}{rrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.09E-01 & -2.29E-02 & -2.75E-02 & -1.80E-02 \\ 
+  logarea & -1.09E-01 & 1.00E+00 & -8.16E-03 & -6.52E-02 & -2.59E-02 \\ 
+  crowdV1 & -2.29E-02 & -8.16E-03 & 1.00E+00 & 5.98E-02 & 3.19E-02 \\ 
+  crowdV2 & -2.75E-02 & -6.52E-02 & 5.98E-02 & 1.00E+00 & 1.07E-01 \\ 
+  crowdV3 & -1.80E-02 & -2.59E-02 & 3.19E-02 & 1.07E-01 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Kansas, vital rate = surv, and species = SCSC.} 
+\begin{tabular}{rrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.84E-01 & -3.85E-02 & -3.41E-02 & -4.71E-02 \\ 
+  logarea & -1.84E-01 & 1.00E+00 & 1.41E-02 & -1.72E-02 & -6.63E-04 \\ 
+  crowdV1 & -3.85E-02 & 1.41E-02 & 1.00E+00 & 5.31E-02 & 5.63E-02 \\ 
+  crowdV2 & -3.41E-02 & -1.72E-02 & 5.31E-02 & 1.00E+00 & 7.50E-02 \\ 
+  crowdV3 & -4.71E-02 & -6.63E-04 & 5.63E-02 & 7.50E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = surv, and species = BOGR.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -2.42E-02 & -6.26E-02 & -3.93E-02 & -4.13E-02 & -3.32E-02 \\ 
+  logarea & -2.42E-02 & 1.00E+00 & -1.46E-02 & 1.17E-02 & 1.22E-02 & 4.58E-03 \\ 
+  crowdV1 & -6.26E-02 & -1.46E-02 & 1.00E+00 & 2.74E-02 & -6.18E-02 & -2.90E-02 \\ 
+  crowdV2 & -3.93E-02 & 1.17E-02 & 2.74E-02 & 1.00E+00 & 1.82E-02 & 2.28E-02 \\ 
+  crowdV3 & -4.13E-02 & 1.22E-02 & -6.18E-02 & 1.82E-02 & 1.00E+00 & 2.29E-02 \\ 
+  crowdV4 & -3.32E-02 & 4.58E-03 & -2.90E-02 & 2.28E-02 & 2.29E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = surv, and species = HECO.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & 5.49E-02 & -6.27E-02 & -7.44E-02 & -8.86E-02 & -4.69E-02 \\ 
+  logarea & 5.49E-02 & 1.00E+00 & -4.33E-03 & -4.33E-02 & 1.92E-02 & 1.60E-02 \\ 
+  crowdV1 & -6.27E-02 & -4.33E-03 & 1.00E+00 & 5.86E-02 & 2.63E-02 & -1.98E-02 \\ 
+  crowdV2 & -7.44E-02 & -4.33E-02 & 5.86E-02 & 1.00E+00 & 4.25E-02 & 2.18E-02 \\ 
+  crowdV3 & -8.86E-02 & 1.92E-02 & 2.63E-02 & 4.25E-02 & 1.00E+00 & 4.26E-02 \\ 
+  crowdV4 & -4.69E-02 & 1.60E-02 & -1.98E-02 & 2.18E-02 & 4.26E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = surv, and species = PASM.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & 1.01E-01 & -2.99E-02 & -2.06E-02 & -7.10E-02 & -2.57E-02 \\ 
+  logarea & 1.01E-01 & 1.00E+00 & -1.55E-03 & 1.35E-03 & -1.67E-02 & 7.47E-03 \\ 
+  crowdV1 & -2.99E-02 & -1.55E-03 & 1.00E+00 & 3.63E-02 & 7.62E-02 & -1.90E-02 \\ 
+  crowdV2 & -2.06E-02 & 1.35E-03 & 3.63E-02 & 1.00E+00 & 2.19E-02 & 6.00E-03 \\ 
+  crowdV3 & -7.10E-02 & -1.67E-02 & 7.62E-02 & 2.19E-02 & 1.00E+00 & 1.35E-02 \\ 
+  crowdV4 & -2.57E-02 & 7.47E-03 & -1.90E-02 & 6.00E-03 & 1.35E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = Montana, vital rate = surv, and species = POSE.} 
+\begin{tabular}{rrrrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 & crowdV3 & crowdV4 \\ 
+  \hline
+(Intercept) & 1.00E+00 & 1.08E-02 & -6.78E-02 & -3.62E-02 & -1.01E-01 & -1.11E-01 \\ 
+  logarea & 1.08E-02 & 1.00E+00 & -8.40E-03 & 1.51E-02 & 3.08E-02 & -5.58E-02 \\ 
+  crowdV1 & -6.78E-02 & -8.40E-03 & 1.00E+00 & 5.16E-02 & 4.40E-02 & 5.77E-03 \\ 
+  crowdV2 & -3.62E-02 & 1.51E-02 & 5.16E-02 & 1.00E+00 & 3.43E-02 & 4.24E-02 \\ 
+  crowdV3 & -1.01E-01 & 3.08E-02 & 4.40E-02 & 3.43E-02 & 1.00E+00 & 3.12E-02 \\ 
+  crowdV4 & -1.11E-01 & -5.58E-02 & 5.77E-03 & 4.24E-02 & 3.12E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = NewMexico, vital rate = surv, and species = BOER.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -1.55E-01 & -8.07E-02 & -3.74E-02 \\ 
+  logarea & -1.55E-01 & 1.00E+00 & -3.38E-02 & 1.40E-02 \\ 
+  crowdV1 & -8.07E-02 & -3.38E-02 & 1.00E+00 & 4.56E-02 \\ 
+  crowdV2 & -3.74E-02 & 1.40E-02 & 4.56E-02 & 1.00E+00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Parameter correlations for site = NewMexico, vital rate = surv, and species = SPFL.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & (Intercept) & logarea & crowdV1 & crowdV2 \\ 
+  \hline
+(Intercept) & 1.00E+00 & -2.30E-01 & -6.01E-02 & -8.89E-02 \\ 
+  logarea & -2.30E-01 & 1.00E+00 & -9.22E-03 & -1.57E-03 \\ 
+  crowdV1 & -6.01E-02 & -9.22E-03 & 1.00E+00 & 4.85E-02 \\ 
+  crowdV2 & -8.89E-02 & -1.57E-03 & 4.85E-02 & 1.00E+00 \\ 
    \hline

### 2016-05-26 devel/make_paramcorr_tables.R

@@ -12,14 +12,16 @@ all_files <- tmp_files[grep(".RDS", tmp_files)]
 if(grep("surv_grow_paramcorrs", list.files("../results/")) != 0){
   file.remove("../results/surv_grow_paramcorrs.txt")
 }
-
+newpages <- seq(3,length(all_files), by=3)
 for(i in 1:length(all_files)){
   tmp_corrs <- readRDS(paste0(in_dir,all_files[i]))
   vital <- strsplit(all_files[i], "_")[[1]][1]
   site <- strsplit(all_files[i], "_")[[1]][2]
   species <- strsplit(all_files[i], "_")[[1]][3]
-  cap <- paste(site,vital,species)
-  cat(print(xtable(tmp_corrs, digits=-2, caption = cap), caption.placement = "top"), file = "../results/surv_grow_paramcorrs.txt", append = TRUE, sep=" ")
+  cap <- paste0("Parameter correlations for site = ", site, ", vital rate = ",vital, ", and species = ",species,".")
+  cat(print(xtable(tmp_corrs, digits=-2, caption = cap), caption.placement = "top", comment=FALSE), file = "../results/surv_grow_paramcorrs.txt", append = TRUE, sep="\n")
+  cat("\n\n", file = "../results/surv_grow_paramcorrs.txt", append = TRUE)
+  if(i %in% newpages){cat("\\newpage{} \n\n", file = "../results/surv_grow_paramcorrs.txt", append = TRUE)}
 }
 
 

### 2016-06-01 docs/online_supporting_information.Rmd
adds new figure for stability

@@
 Error bars represent the upper and lower 95\% quantiles from model simulations.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/cv_vs_synchrony.png}
+  \caption{Relationship between species synchrony (percent cover, top panels; growth rates, bottom panels) and the coefficient of variation of summed community cover.
+Each point represents a single time series from simulations and is color coded according to simulation experiment.
+Results from IBM simulations are from 25$\text{m}^2$ landscapes.
+Within experiments, the CV of total cover increases with synchrony of percent cover.
+The relationship between CV of total cover and synchrony of per capita growth rates within experiments is weaker.
+Among experiments, CV of total cover is lowest when the synchrony of per capita growth rates is lowest, confirming the link between species synchrony and ecosystem stability.
+Because our main focus is on species synchrony, we simply removed environmental variation rather than breaking the covariance structure of species environmental responses.
+Therefore, CV of total cover is drastically reduced when we remove environmental variation (``No E.S.'' points).
+However, regardless of environmental variance, CV of total cover is always reduced when synhrony is reduced.}
 \end{figure}

### 2016-06-01 docs/online_supporting_information.tex
adds new figure for stability

--- 2016-04-01 docs/online_supporting_information.tex
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/cv_vs_synchrony.png}
+  \caption{Relationship between species synchrony (percent cover, top panels; growth rates, bottom panels) and the coefficient of variation of summed community cover.
+Each point represents a single time series from simulations and is color coded according to simulation experiment.
+Results from IBM simulations are from 25$\text{m}^2$ landscapes.
+Within experiments, the CV of total cover increases with synchrony of percent cover.
+The relationship between CV of total cover and synchrony of per capita growth rates within experiments is weaker.
+Among experiments, CV of total cover is lowest when the synchrony of per capita growth rates is lowest, confirming the link between species synchrony and ecosystem stability.
+Because our main focus is on species synchrony, we simply removed environmental variation rather than breaking the covariance structure of species environmental responses.
+Therefore, CV of total cover is drastically reduced when we remove environmental variation (``No E.S.'' points).
+However, regardless of environmental variance, CV of total cover is always reduced when synhrony is reduced.}
+\end{figure}
+
+\newpage{}
+
 \subsection{Supporting Tables}\label{supporting-tables}

### 2016-06-02 docs/ms_communitySynchrony.Rmd
addresses minor comments from reviewers

@@
 Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
-The stabilizing effect of species richness stems from individual species responding to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity).
-[@Hector2010; @Isbell2009; @DeMazancourt2013a].
+The stabilizing effect of species richness stems from individual species responding to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
 Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
 Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
-Predicting the impacts of global change requires a mechanistic understanding of species synchrony in natural settings.
+Predicting the impacts of global change requires a mechanistic understanding of species synchrony in natural settings because of the inherent link between synchrony and ecosystem stability.
 
-Theory identifies the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and interspecific competition [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+Theory identifies the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and interspecific interactions [@Loreau2008a; @Gonzalez2009; @Loreau2013].
 For example, in a community composed of large (no demographic stochasticity) and weakly interacting (no interspecific competition) populations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
 However, such a prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
-Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains unkown: no study has explicitly tested theory on the drivers of species synchrony in natural communities.
+Whether such theoretical predictions hold in natural plant communities where species differences are unlikely to be symmetrical remains unkown: no study has explicitly tested theory on the drivers of species synchrony in natural plant communities.
 
@@
 Even beyond grassland studies, whether competition or environmental responses drive compensatory dynamics remains controversial [reviewed in @Gonzalez2009].
-In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from community time series [e.g., @Houlahan2007] is impossible.
+In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from the covariance matrix of species abundances [e.g., @Houlahan2007] is impossible.
 This is because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
@@
 This limiting case, $\mathcal{M}_{D}$, represents a community where dynamics are driven by demographic stochasticity alone.
-Our prediction is
+Our prediction for the synchrony of per capita growth rates for $\mathcal{M}_{D}$, $\phi_{R,\mathcal{M}_{D}}$, is
 
 \begin{equation}
-\phi_{R,\mathcal{M}_{D}} = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
+\phi_{R,\mathcal{M}_{D}} = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2},
 \end{equation}
 
-\noindent for synchrony of per capita growth rates.
-$p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
+\noindent where $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
 When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, synchrony equal 1/S [@Loreau2008a].
 
-Our second limiting case assumes only environmental stochasticity is operating.
-Thus, there are no interspecific interactions, demographic stochasticity is absent, intrinsic growth rates are identical, and environmental variance is identical for all species.
-Our prediction for this limiting case, $\mathcal{M}_{E}$, is
+Our second limiting case assumes only environmental stochasticity is operating ($\mathcal{M}_{E}$).
+Thus, we assume there are no interspecific interactions, demographic stochasticity is absent, intrinsic growth rates are identical, and environmental variance is identical for all species.
+Our prediction for the synchrony of per capita growth rates for $\mathcal{M}_{E}$, $\phi_{R,\mathcal{M}_{E}}$, is
 
 \begin{equation}
-\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(u_{ei},u_{ej})}{S^2}
+\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(u_{ei},u_{ej})}{S^2},
 \end{equation}
 
-\noindent{} for synchrony of per capita growth rates, where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
+\noindent{} where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
 Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in monoculture (no interspecific interactions) to calculate the covariance of species' environmental responses.
@@
 First, in the absence of interspecific competition and demographic stochasticity, species are effectively growing in large monocultures.
-Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(u_{ie}, u_{je})$) and parameterize Equation 3.
-Parameterizing Equation 2 does not require simulation output because the only parameters are the species' relative abundances.
+Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(u_{ie}, u_{je})$) and parameterize Equation 4.
+Parameterizing Equation 3 does not require simulation output because the only parameters are the species' relative abundances.
 Second, simulations (5) and (6) represent the simulated version of our limiting case theoretical predictions.
@@
 Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
-Error bars represent the upper and lower 95\% quantiles from model simulations.
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.
 All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
@@
 + E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
 \end{figure}

### 2016-06-02 docs/online_supporting_information.Rmd
addresses minor comments from reviewers

@@
   \caption{Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IBM (top panels) and the IPM (bottom panels).
-IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.
+IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the number of observation years for each data set.
 IBM density curves come from 100 replicate simulations of 75 iterations each.
@@
 Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
 \end{figure}
@@
 + E.S.'').
+The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
 The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
 \end{figure}
@@
 + No E.S.'').
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
 \end{figure}
@@
 -----------------
+In all tables, species codes are as follows (also see Table 1 in the main text):
+
+* *Bouteloua eriopoda* = BOER
+* *Sporobolus flexuosus* = SPFL
+* *Bouteloua rothrockii* = BORO
+* *Bouteloua curtipendula* = BOCU
+* *Bouteloua hirsuta* = BOHI
+* *Schizachyrium scoparium* = SCSC
+* *Bouteloua gracilis* = BOGR
+* *Hesperostipa comata* = HECO
+* *Pascopyrum smithii* = PASM
+* *Poa secunda* = POSE
+* *Artemisia tripartita* = ARTR
+* *Pseudoroegneria spicata* = PSSP
+
 ```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}

### 2016-06-02 docs/online_supporting_information.tex
addresses minor comments from reviewers

@@
 \subsection{Supporting Tables}\label{supporting-tables}
+
+In all tables, species codes are as follows (also see Table 1 in the
+main text):
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \emph{Bouteloua eriopoda} = BOER
+\item
+  \emph{Sporobolus flexuosus} = SPFL
+\item
+  \emph{Bouteloua rothrockii} = BORO
+\item
+  \emph{Bouteloua curtipendula} = BOCU
+\item
+  \emph{Bouteloua hirsuta} = BOHI
+\item
+  \emph{Schizachyrium scoparium} = SCSC
+\item
+  \emph{Bouteloua gracilis} = BOGR
+\item
+  \emph{Hesperostipa comata} = HECO
+\item
+  \emph{Pascopyrum smithii} = PASM
+\item
+  \emph{Poa secunda} = POSE
+\item
+  \emph{Artemisia tripartita} = ARTR
+\item
+  \emph{Pseudoroegneria spicata} = PSSP
+\end{itemize}
 

### 2016-06-21 docs/ms_communitySynchrony.Rmd
adds correlation of cv and pgr synch to ms

@@
 xtable_data <- table_data[myorder,]
+sink(file = "../results/correlation_Rsynch_cv.txt")
+cat("correlation between synchrony of growth rates and cover CV is:")
+cor(table_data$`CV of Total Cover`, table_data[,2])
+Sys.Date()
+sink()
 # colnames(table_data) <- c("Site",

### 2016-06-22 docs/ms_communitySynchrony.Rmd
adds CI figs for survival and growth competiiton random effects

@@
 output:
-  word_document: default
   pdf_document:
     fig_caption: yes
-    keep_tex: yes
+    keep_tex: no
 urlcolor: black

### 2016-06-22 devel/plot_comp_randeff_cis.R

@@ -1,0 +1,112 @@
+################################################################################
+##  plot_comp_randeff_cis.R: plots the 95% CIs for the yearly random offsets  ##
+##  on competition for survival, growth, and recruitment.                     ##
+################################################################################
+
+rm(list=ls(all.names = TRUE)) # clear the workspace
+library(ggplot2)
+library(gridExtra)
+
+
+####
+####  SURVIVAL
+####
+surv_param_summaries <- readRDS("../results/surv_summaries_yrlycomp_list.RDS")
+sites <- names(surv_param_summaries)
+plot_list <- list()
+for(do_site in sites){
+  site_summary <- surv_param_summaries[[do_site]]
+  species <- names(site_summary)
+  for(do_species in species){
+    species_summary <- site_summary[[do_species]]
+    comp_ids <- grep("W", names(species_summary))
+    for(i in comp_ids){
+      comp_summary <- species_summary[[i]]
+      colnames(comp_summary)[c(4,6)] <- c("low", "high")
+      comp_summary$year <- 1:nrow(comp_summary)
+      new_plot <- ggplot(data=comp_summary, aes(x=year, y=mean))+
+        geom_hline(aes(yintercept=0), color="red")+
+        geom_errorbar(aes(ymin=low, ymax=high), width=0.25)+
+        # geom_point(size=4)+
+        ylab("yearly competition effect")+
+        ggtitle(paste(do_site,do_species,names(species_summary)[i]))+
+        theme(plot.title = element_text(size = 12),
+              axis.title = element_text(size=8))
+      plot_list <- c(plot_list, list(new_plot))
+    } # end competition coefficient loop
+  } # end species loop
+} # end site loop
+
+png(filename = "randyear_compCIs_survival.png", width = 1280, height = 960, units = "px")
+do.call(grid.arrange, c(plot_list, list(ncol = 7)))
+dev.off()
+
+
+
+####
+####  GROWTH
+####
+grow_param_summaries <- readRDS("../results/growth_summaries_yrlycomp_list.RDS")
+sites <- names(grow_param_summaries)
+plot_list <- list()
+for(do_site in sites){
+  site_summary <- grow_param_summaries[[do_site]]
+  species <- names(site_summary)
+  for(do_species in species){
+    species_summary <- site_summary[[do_species]]
+    comp_ids <- grep("W", names(species_summary))
+    for(i in comp_ids){
[…]

### 2016-06-22 docs/ms_communitySynchrony.Rmd
ms revisions

@@
 - \usepackage{rotating}
+- \usepackage{color, soul}
 layout: 12pt
@@
 Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
-Predicting the impacts of global change requires a mechanistic understanding of species synchrony in natural settings because of the inherent link between synchrony and ecosystem stability.
+Predicting the impacts of global change on ecosystem stability requires a mechanistic understanding of species synchrony in natural settings because of the inherent link between synchrony and stability.
 
@@
 However, such a prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
-Whether such theoretical predictions hold in natural plant communities where species differences are unlikely to be symmetrical remains unkown: no study has explicitly tested theory on the drivers of species synchrony in natural plant communities.
+\hl{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains relatively unkown because only a few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \hl{and they did not consider demographic stochasiticity.}
 
@@
 Even beyond grassland studies, whether competition or environmental responses drive compensatory dynamics remains controversial [reviewed in @Gonzalez2009].
-In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from the covariance matrix of species abundances [e.g., @Houlahan2007] is impossible.
+In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from \hl{the covariance matrix of species abundances} [e.g., @Houlahan2007] is impossible.
 This is because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
@@
 They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
-While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Furthermore, although @Mutshinda2009 quantified the relative importance of environmental stochasticity and competition to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
+@Thibaut2012 \hl{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
+While a major step forward, Mutshinda et al.'s (2009) and Thibaut et al.'s (2012) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
+Furthermore, although both studies quantified the relative importance of environmental stochasticity and competition to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
 
@@
 However, we can derive analytical predictions for species synchrony under special limiting cases.
+\hl{The limiting case predictions we derive serve as baselines against which we can compare, and thus understand, results from empirically-based simulations (described below).}
 We focus on synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
@@
 To derive analytical predictions we solved a first-order approximation of Equation 2 [@DeMazancourt2013a and Online Supporting Information].
+\hl{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos} [@Loreau2008a].
+\hl{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are typically more well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small} (e.g., $r < 2$).
 
@@
 We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
-Genet size and crowding were treated as fixed effects; spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
+\hl{Crowding was treated as a fixed effect without a temporal component because the 95% credible intervals for random year effects on crowding broadly overlapped zero.}
+Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
 
@@
 To remove interspecific interactions, we set the off-diagonals of the interaction matrix for each vital rate regression to zero.
-This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are effectively growing in monoculture.
+This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are \hl{growing in isolation}.
+\hl{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a species in a no-interaction model.}
 To remove the effect of a fluctuating environment, we removed the temporal (interannual) random effects from the regression equations.
@@
 ### Results
-Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
-Synchrony tended to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability in percent cover (e.g., Montana).
-Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
+Synchrony of species' per capita growth rates at our study sites range from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
+Synchrony tends to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability in percent cover (e.g., Montana).
+Synchrony is lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
+\hl{Synchrony of per capita growth rates and CV of percent cover are positively correlated} (Pearson’s $\rho$ = 0.72).
 For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
@@
 Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
+\hl{Observed synchrony of per capita growth rates were positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
 
@@
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
+\hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S#-S#), as required by our predictions.}
 However, the theoretical predictions were strongly supported by simulations of population models fit to long-term data from natural plant communities (Table 3).
-Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony.
-The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem functioning experiments [@DeMazancourt2013a].
+Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony, \hl{and allows us to place our simulation results within the context of contemporary theory}.
 
@@
 2), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity was not strong enough to completely counteract the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
+Even in small populations, however, demographic stochasticity was not strong enough to \hl{compensate} the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
 2).
@@
 A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
-Rare species, which our analysis does not include, could be kept at low densities by competitive suppression due to strong interspecific interactions.
+\hl{Most communities are dominated by few common species and many rare species}, \hl{meaning that the low number of common species in our focal commiunities is not unusual.} 
+\hl{Rather, the generality of our results hinges upon the influence of rare species.
+Rare species could be kept at low densities by competitive suppression due to strong interspecific interactions.
 If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
-However, rare species may occupy rare niches that do not overlap with other species, in which case synchrony would remain predominantly driven by environmental responses.
+However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species} [@Yenni2012], \hl{in which case synchrony would remain predominantly driven by environmental responses.}
 Rare species could also be limited by generalist natural enemies whose effects are density-independent but alter competitive hierarchies [@Mordecai2011].
@@
 One way to test this prediction is to continue collecting data from biodiversity-ecoystem functioning experiments and conduct the analysis of @Gross2014 every few years.
+
+\hl{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
+Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap} [@Vasseur2007], \hl{dispersal} [@Gouhier2010], \hl{and the strength of top-down control} [@Bauer2014] \hl{are all likely to affect community synchrony.}
+\hl{However, our results do agree with findings from fish, bird, and insect studies showing that species interactions had little effect on synchrony} [@Mutshinda2009; @Thibaut2012].
 

### 2016-06-22 R/recruitment_function_yrlycomp.R

@@ -10,8 +10,7 @@
 #' @return Matrix of statistical results by fitted parameter.
 
 recruit_mcmc_yrlr <- function(dataframe, n_adapt=5000, n_update=10000, 
-                         n_samples=20000, n_thin=50, sppList,
-                         fig_outfile){
+                         n_samples=20000, n_thin=50, sppList){
   D <- dataframe
   # Calculate mean cover by group and year
   tmpD <- D[,c("quad","year","Group",paste("cov.",sppList,sep=""))]
@@ -91,7 +90,8 @@ recruit_mcmc_yrlr <- function(dataframe, n_adapt=5000, n_update=10000,
   out <- coda.samples(jm, variable.names=params, n.iter=n.Samp, n.thin=n.Thin)
 
   zmStat <- summary(out)$stat
-  return(zmStat)
+  zmQuants <- summary(out)$quantile
+  return(list(stats=zmStat,quants=zmQuants))
 } # end function
   
 

### 2016-06-22 runscripts/call_recruitment_fxn_yrlycomp.R

@@ -11,7 +11,8 @@ removes <- c(grep("*.csv", site_list),
              grep("Crowding", site_list))
 site_list <- site_list[-removes]
 
-recruit_params_site_list <- list()
+recruit_params_stat_list <- list()
+recruit_params_quant_list <- list()
 for(do_site in site_list){
   species_list <- list.files(paste(path_to_data, do_site, "/", sep=""))
   if(length(grep("_", species_list)) > 0){
@@ -87,10 +88,11 @@ for(do_site in site_list){
   } # end species loop
   D[is.na(D)] <- 0  # replace missing values
   
-  outfile_now <- paste0("../results/param_covariance/recruitment_",do_site,"_",do_species,"_paramcorrs.png")
-  recruit_params <- recruit_mcmc_yrlr(dataframe = D, sppList = species_list, fig_outfile = outfile_now)
-  recruit_params_site_list[[do_site]] <- recruit_params
+  recruit_params <- recruit_mcmc_yrlr(dataframe = D, sppList = species_list)
+  recruit_params_stat_list[[do_site]] <- recruit_params[[1]]
+  recruit_params_quant_list[[do_site]] <- recruit_params[[2]]
 } # end site loop
 
 # Save the output
-saveRDS(recruit_params_site_list, "../results/recruit_parameters_yrlycomp.RDS")
+saveRDS(recruit_params_stat_list, "../results/recruit_parameters_yrlycomp.RDS")
+saveRDS(recruit_params_quant_list, "../results/recruit_quants_yrlycomp.RDS")

### 2016-06-22 docs/ms_communitySynchrony.Rmd
more revisions

@@
 - \doublespacing
-- \usepackage{color}
 - \usepackage{rotating}
@@
 Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
-Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific competition, and demographic stochasticity.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \hl{interactions}, and demographic stochasticity.
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
 Simulation experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
-Removing interspecific competition and demographic stochasticity had small effects on synchrony.
+Removing interspecific \hl{interactions} and demographic stochasticity had small effects on synchrony.
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
@@
 Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
-Predicting the impacts of global change on ecosystem stability requires a mechanistic understanding of species synchrony in natural settings because of the inherent link between synchrony and stability.
+\hl{Predicting the impacts of global change on ecosystem stability requires a mechanistic understanding of species synchrony in natural settings because of the inherent link between synchrony and stability.}
 
-Theory identifies the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and interspecific interactions [@Loreau2008a; @Gonzalez2009; @Loreau2013].
-For example, in a community composed of large (no demographic stochasticity) and weakly interacting (no interspecific competition) populations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
+Theory identifies the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and \hl{interspecific interactions} [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+For example, in a community composed of large (no demographic stochasticity) and weakly interacting (no interspecific \hl{interactions}) populations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
 However, such a prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
@@
 
-The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific competition is to remove them one-by-one, and in combination.
+The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} is to remove them one-by-one, and in combination.
 In principle, this could be done in an extremely controlled laboratory setting, but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
@@
 While a major step forward, Mutshinda et al.'s (2009) and Thibaut et al.'s (2012) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Furthermore, although both studies quantified the relative importance of environmental stochasticity and competition to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
+Furthermore, although both studies quantified the relative importance of environmental stochasticity and \hl{interspecific interactions} to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
 
 Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony.
-Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific competition on species synchrony in natural plant communities.
+Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} on species synchrony in natural plant communities.
 To achieve these objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
 We then confront our theoretical predictions with simulations from the empirically-based population models.
-Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific competition on community-wide species synchrony.
+Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} on community-wide species synchrony.
 Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, and demographic and environmental stochasticity), and that these models successfully reproduce observed community dynamics [@Chu2015], perturbing the models can reveal the processes that determine species synchrony in our focal grassland communities.
@@
 #### Data
-In the early 20th century, several range experiment stations in the western U.S. began mapping permanent quadrats, and conducted annual censuses for decades. 
-We use such long-term demographic data sets from five semiarid grasslands in the western United States [described in detail by @Chu2015].
-Each study site includes a set of 1-$\text{m}^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
+We use long-term demographic data from five semiarid grasslands in the western United States [described in detail by @Chu2015].
+Each site includes a set of 1-$\text{m}^2$ permanent quadrats within which all individual plants were identified and mapped annually using a pantograph (Hill 1920).
 The resulting mapped polygons represent basal cover for grasses and canopy cover for shrubs.
@@
 We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
-\hl{Crowding was treated as a fixed effect without a temporal component because the 95% credible intervals for random year effects on crowding broadly overlapped zero.}
+\hl{Crowding was treated as a fixed effect without a temporal component because the 95\% credible intervals for random year effects on crowding broadly overlapped zero.}
 Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
@@
 Our recruitment model assumes that the number of recruits produced in each year follows a negative binomial distribution with the mean dependent on the cover of the parent species, permanent spatial variation among groups, temporal variation among years, and inter- and intraspecific interactions as a function of total species' cover in the quadrat.
-We fit the recruitment model using a hierarchical Bayesian approach implemented in JAGS [@Plummer2003] using the `rjags' package [@Plummer2014] in R.
+We fit the recruitment model using a hierarchical Bayesian approach implemented in JAGS [@Plummer2003] using the `rjags' package [@Plummer2014] in R [@R2013].
 Again, temporal and spatial variation were treated as random effects.
@@
 #### Simulation experiments
-We performed removal simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) were removed one-by-one and in combination.
+We performed removal simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific \hl{interactions}) were removed one-by-one and in combination.
 To remove interspecific interactions, we set the off-diagonals of the interaction matrix for each vital rate regression to zero.
@@
 We conducted the following six simulation experiments:
-(1) IBM: All drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) present;
+(1) IBM: All drivers (environmental stochasticity, demographic stochasticity, or \hl{interspecific interactions}) present;
 (2) IPM: Demographic stochasticity removed;
 (3) IBM: Environmental stochasticity removed;
-(4) IBM: Interspecific competition removed;
-(5) IPM: Interspecific competition and demographic stochasticity removed;
-(6) IBM: Interspecific competition and environmental stochasticity removed.
+(4) IBM: \hl{Interspecific interactions} removed;
+(5) IPM: \hl{Interspecific interactions} and demographic stochasticity removed;
+(6) IBM: \hl{Interspecific interactions} and environmental stochasticity removed.
 We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
@@
 Results from our simulation experiments also allow us to test our theoretical predictions.
-First, in the absence of interspecific competition and demographic stochasticity, species are effectively growing in large monocultures.
+First, in the absence of \hl{interspecific interactions} and demographic stochasticity, \hl{populations fluctuate in response to the environment alone}.
 Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(u_{ie}, u_{je})$) and parameterize Equation 4.
@@
 Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S3).
-Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 1; "No Comp." simulations).
+Species interactions are weak in these communities [Table S4 and @Chu2015], and removing \hl{interspecific interactions} had little effect on synchrony (Fig. 1; "No Comp." simulations).
 Removing competition caused, at most, a 5% change in synchrony (Fig. 1).
@@
 
-For all five communities, the synchrony of species' growth rates in monoculture almost perfectly matched species synchrony in polyculture (Fig. 3).
+For all five communities, the synchrony of species' growth rates when growing in isolation almost perfectly matched species synchrony in polyculture (Fig. 3).
 Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Online Supporting Information, Figs.
@@
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
-\hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S#-S#), as required by our predictions.}
+\hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. Sx-Sx), as required by our predictions.}
 However, the theoretical predictions were strongly supported by simulations of population models fit to long-term data from natural plant communities (Table 3).
@@
 Only in the Idaho community does synchrony of per capita growth rates approach $\mathcal{M}_D$ in a non-constant environment (Fig. 2).
-This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
+This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a \hl{2}5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
 
@@
 Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
-1), and the synchrony of species in monoculture was a very strong predictor of synchrony of species in polyculture (Fig.
+1), and the synchrony of species in isolation was a very strong predictor of synchrony of species in polyculture (Fig.
 3).
@@
 We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
-The lack of an effect of interspecific competition on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
+The lack of an effect of \hl{interspecific interactions} on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
 
@@
 Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap} [@Vasseur2007], \hl{dispersal} [@Gouhier2010], \hl{and the strength of top-down control} [@Bauer2014] \hl{are all likely to affect community synchrony.}
-\hl{However, our results do agree with findings from fish, bird, and insect studies showing that species interactions had little effect on synchrony} [@Mutshinda2009; @Thibaut2012].
 
@@
 xtable_data <- table_data[myorder,]
-sink(file = "../results/correlation_Rsynch_cv.txt")
-cat("correlation between synchrony of growth rates and cover CV is:")
-cor(table_data$`CV of Total Cover`, table_data[,2])
-Sys.Date()
-sink()
+
 # colnames(table_data) <- c("Site",
@@
       size="normalsize")
+```
+
+```{r print_sync_cv_cor, echo=FALSE, eval=FALSE, results=FALSE}
+sink(file = "../results/correlation_Rsynch_cv.txt")
+cat("correlation between synchrony of growth rates and cover CV is:")
+cor(table_data$`CV of Total Cover`, table_data[,2])
+Sys.Date()
+sink()
 ```
@@
   \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
-Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
-+ No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
+Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific interactions present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific interactions removed (``No Comp.''), interspecific interactions and demographic stochasticity removed (``No Comp.
++ No D.S.''), and interspecific interactions and environmental stochasticity removed (``No Comp.
 + No E.S.'').
@@
       \includegraphics[width=4in]{./components/formatted_figures/formatted_figure3.png}
-  \caption{Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions).
-We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
+  \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).
+We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.
 The dashed line is the line of equality.
 Simulation results in this figure are analagous to ``No Comp.
-+ No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.
++ No D.S.'' (species in isolation) and ``No D.S.'' (species in competition) in Fig.
 1, but here we control the time series of random year effects.}

### 2016-06-23 docs/ms_communitySynchrony.Rmd
R1 version; sent to coauthors

@@
 Species interactions are weak in these communities [Table S4 and @Chu2015], and removing \hl{interspecific interactions} had little effect on synchrony (Fig. 1; "No Comp." simulations).
-Removing competition caused, at most, a 5% change in synchrony (Fig. 1).
+Removing \hl{interspecific interactions} caused, at most, a 5% change in synchrony (Fig. 1).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
@@
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
-\hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. Sx-Sx), as required by our predictions.}
+\hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
 However, the theoretical predictions were strongly supported by simulations of population models fit to long-term data from natural plant communities (Table 3).
@@
 tripartita* from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
-S4).
+S8).
 Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.
@@
   \centering
-      \includegraphics[width=4in]{./components/formatted_figures/formatted_figure3.png}
+      \includegraphics[width=3in]{./components/formatted_figures/formatted_figure3.png}
   \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).

### 2016-06-23 docs/online_supporting_information.Rmd
R1 version; sent to coauthors

@@
   \centering
+      \includegraphics[width=6in]{./components/formatted_figures/environmental_variances_cover.png}
+  \caption{Variance of percent cover for each species (along x-axes) in each site through time from simulations with only environmental stochasticity operating (IPM with no species interactions).}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/environmental_variances_pgr.png}
+  \caption{Variance of per capita growth rates for each species (along x-axes) in each site through time from simulations with only environmental stochasticity operating (IPM with no species interactions).}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/demographic_variances_cover.png}
+  \caption{Variance of percent cover for each species (along x-axes) in each site through time from simulations with only demographic stochasticity operating (IBM simulated on 5 by 5 meter landscape).}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/demographic_variances_pgr.png}
+  \caption{Variance of per capita growth rates for each species (along x-axes) in each site through time from simulations with only demographic stochasticity operating (IBM simulated on 5 by 5 meter landscape).}
+\end{figure}
+
+\pagebreak{}
+
+\begin{figure}[!ht]
+  \centering
       \includegraphics[width=3in]{./components/formatted_figures/formatted_figureS4.png}
@@
 
+<!---
 \begin{figure}[!ht]
@@
 \newpage{}
+--->
 
@@
   vr_do <- vr_files[grep(do_vital,vr_files)]
+  torms <- grep("yr", vr_do)
+  vr_do <- vr_do[-torms]
   tmp_vital <- readRDS(paste0("../results/",vr_do))
@@
 vr_do <- vr_files[grep("recruit",vr_files)]
+torms <- c(grep("yr", vr_do), grep("MONT", vr_do))
+vr_do <- vr_do[-torms]
 tmp_vital <- readRDS(paste0("../results/",vr_do))

### 2016-06-23 docs/online_supporting_information.tex
R1 version; sent to coauthors

@@
 Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
-\end{figure}
-
-\newpage{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{./components/cv_vs_synchrony.png}
-  \caption{Relationship between species synchrony (percent cover, top panels; growth rates, bottom panels) and the coefficient of variation of summed community cover.
-Each point represents a single time series from simulations and is color coded according to simulation experiment.
-Results from IBM simulations are from 25$\text{m}^2$ landscapes.
-Within experiments, the CV of total cover increases with synchrony of percent cover.
-The relationship between CV of total cover and synchrony of per capita growth rates within experiments is weaker.
-Among experiments, CV of total cover is lowest when the synchrony of per capita growth rates is lowest, confirming the link between species synchrony and ecosystem stability.
-Because our main focus is on species synchrony, we simply removed environmental variation rather than breaking the covariance structure of species environmental responses.
-Therefore, CV of total cover is drastically reduced when we remove environmental variation (``No E.S.'' points).
-However, regardless of environmental variance, CV of total cover is always reduced when synhrony is reduced.}
 \end{figure}
@@
 New Mexico & 0.56 & 0.51 & 0.86 & 0.86 \\ 
-  Arizona & 0.51 & 0.52 & 0.82 & 0.82 \\ 
+  Arizona & 0.51 & 0.52 & 0.82 & 0.81 \\ 
   Kansas & 0.35 & 0.39 & 0.62 & 0.62 \\ 
-  Montana & 0.31 & 0.33 & 0.47 & 0.49 \\ 
-  Idaho & 0.29 & 0.31 & 0.43 & 0.41 \\ 
+  Montana & 0.31 & 0.31 & 0.47 & 0.50 \\ 
+  Idaho & 0.29 & 0.31 & 0.43 & 0.42 \\ 
    \hline
@@
   \hline
-1No D.S. & 4.82 & 3.99 & 7.19 & 5.37 & 9.53 \\ 
-  2No Comp. + No D.S. & 4.74 & 2.70 & 6.81 & 4.53 & 7.69 \\ 
-  4No Comp. & 0.98 & 0.06 & 4.84 & 1.01 & 0.11 \\ 
-  5No E.S. & 40.14 & 33.20 & 47.84 & 38.52 & 9.28 \\ 
-  6No Comp. + No E.S. & 45.95 & 41.48 & 38.21 & 33.07 & 20.33 \\ 
+1No D.S. & 5.53 & 4.28 & 6.01 & 5.27 & 11.95 \\ 
+  2No Comp. + No D.S. & 5.56 & 2.06 & 7.03 & 6.21 & 9.10 \\ 
+  4No Comp. & 0.98 & 0.06 & 4.84 & 1.43 & 0.11 \\ 
+  5No E.S. & 40.14 & 33.20 & 47.84 & 39.06 & 9.28 \\ 
+  6No Comp. + No E.S. & 45.95 & 41.48 & 38.21 & 39.44 & 20.33 \\ 
    \hline
@@
   \hline
-Arizona & intercept & 0.49 & -0.05 & 0.45 \\ 
-  Arizona & slope & 0.68 &  & 0.15 \\ 
-  Idaho & intercept & 0.46 & 0.38 & 0.29 \\ 
+Arizona & intercept & 0.51 & -0.05 & 0.51 \\ 
+  Arizona & slope & 0.68 &  & 0.16 \\ 
+  Idaho & intercept & 0.45 & 0.38 & 0.25 \\ 
   Idaho & slope & 0.50 &  & 0.47 \\ 
-  Kansas & intercept & 0.47 & 0.29 & 0.24 \\ 
-  Kansas & slope & 0.18 &  & 0.16 \\ 
-  Montana & intercept & 0.26 & 0.04 & 0.39 \\ 
-  Montana & slope & 0.10 &  & 0.01 \\ 
-  NewMexico & intercept & 0.54 & -0.15 & 0.64 \\ 
-  NewMexico & slope & 0.32 &  & 0.15 \\ 
+  Kansas & intercept & 0.44 & 0.29 & 0.26 \\ 
+  Kansas & slope & 0.19 &  & 0.16 \\ 
+  Montana & intercept & 0.29 & 0.07 & 0.41 \\ 
+  Montana & slope & 0.11 &  & 0.02 \\ 
+  NewMexico & intercept & 0.51 & -0.15 & 0.65 \\ 
+  NewMexico & slope & 0.37 &  & 0.20 \\ 
    \hline
@@
   Kansas & Intraspecific & -0.0062 & -0.9945 & -0.0586 \\ 
-  Montana & Interspecific & -0.0053 & 0.0920 & 0.0354 \\ 
-  Montana & Intraspecific & -0.0733 & -2.0492 & -0.8103 \\ 
+  Montana & Interspecific & -0.0059 & 0.0691 & 0.0342 \\ 
+  Montana & Intraspecific & -0.0801 & -2.0900 & -0.8652 \\ 
   NewMexico & Interspecific & -0.0027 & -0.3461 & -0.0080 \\ 

### 2016-06-23 devel/plot_comp_randeff_cis.R
adds recruitment temporal competition CI check figure

@@ -81,32 +81,37 @@ dev.off()
 ####
 ####  RECRUITMENT
 ####
-rec_param_summaries <- readRDS("../results/recruit_parameters_yrlycomp.RDS")
+# density-dependence (dd) is a 3-D array: i focal species, j competing species, k year
+rec_param_summaries <- readRDS("../results/recruit_quants_yrlycomp.RDS")
 sites <- names(rec_param_summaries)
 plot_list <- list()
 for(do_site in sites){
   site_summary <- rec_param_summaries[[do_site]]
-  species <- names(site_summary)
-  for(do_species in species){
-    species_summary <- site_summary[[do_species]]
-    comp_ids <- grep("W", names(species_summary))
-    for(i in comp_ids){
-      comp_summary <- species_summary[[i]]
-      colnames(comp_summary)[c(4,6)] <- c("low", "high")
-      comp_summary$year <- 1:nrow(comp_summary)
-      new_plot <- ggplot(data=comp_summary, aes(x=year, y=mean))+
-        geom_hline(aes(yintercept=0), color="red")+
-        geom_errorbar(aes(ymin=low, ymax=high), width=0.25)+
-        # geom_point(size=4)+
-        ylab("yearly competition effect")+
-        ggtitle(paste(do_site,do_species,names(species_summary)[i]))+
-        theme(plot.title = element_text(size = 12),
-              axis.title = element_text(size=8))
-      plot_list <- c(plot_list, list(new_plot))
-    } # end competition coefficient loop
-  } # end species loop
+  comp_summary <- site_summary[grep("dd\\[",rownames(site_summary)),]
+  species_numid <- unique(substr(rownames(comp_summary), 4, 4))
+  species_names <- names(surv_param_summaries[[do_site]])
+  
+  for(i in 1:length(species_numid)){
+    focal_species_rows <- grep(paste0("dd\\[", species_numid[i]), rownames(comp_summary))
+    tmp_densdep <- as.data.frame(comp_summary[focal_species_rows,])
+    colnames(tmp_densdep)[c(1,5)] <- c("low", "high")
+    tmp_densdep$year <- c(1:nrow(tmp_densdep))
+    tmp_densdep$focal <- substr(rownames(tmp_densdep), 4, 4)
+    tmp_densdep$competitor <- substr(rownames(tmp_densdep), 6, 6)
+    tmp_densdep$intra <- "yes"
[…]

### 2016-06-24 docs/ms_communitySynchrony.Rmd
fixes small typo in ms

@@
 However, such a prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
-\hl{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains relatively unkown because only a few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \hl{and they did not consider demographic stochasiticity.}
+\hl{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains relatively unkown because only a few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \hl{and they did not consider demographic stochasticity.}
 

### 2016-06-24 docs/ms_communitySynchrony.Rmd
adds PBA edits

@@
 \begin{abstract}
-Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+Temporal asynchrony among species \hl{helps} diversity \hl{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
 Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \hl{interactions}, and demographic stochasticity.
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-Simulation experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+\hl{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
 Removing interspecific \hl{interactions} and demographic stochasticity had small effects on synchrony.
@@
 Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
-The stabilizing effect of species richness stems from individual species responding to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
+The stabilizing effect of species richness stems from individual species responding \hl{in different ways} to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
 Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
 Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
-\hl{Predicting the impacts of global change on ecosystem stability requires a mechanistic understanding of species synchrony in natural settings because of the inherent link between synchrony and stability.}
+\hl{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}
 
-Theory identifies the three main determinants of species synchrony as environmental stochasticity, demographic stochasticity, and \hl{interspecific interactions} [@Loreau2008a; @Gonzalez2009; @Loreau2013].
-For example, in a community composed of large (no demographic stochasticity) and weakly interacting (no interspecific \hl{interactions}) populations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
-However, such a prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
-\hl{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical remains relatively unkown because only a few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \hl{and they did not consider demographic stochasticity.}
+Theory identifies three main determinants of species synchrony: environmental stochasticity, demographic stochasticity, and \hl{interspecific interactions} [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+For example, in a community composed of large \hl{populations} (no demographic stochasticity) \hl{with weak interspecific interations}, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
+However, \hl{this} prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
+\hl{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \hl{and they did not consider demographic stochasiticity.}
 
@@
 For example, @Mutshinda2009 fit a dynamic population model to several community time series of insect and bird abundances.
-They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
+They used a statistical technique to decompose temporal variation into competition and environmental components, and found that positively correlated environmental responses among species determined community dynamics.
 @Thibaut2012 \hl{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
@@
 However, we can derive analytical predictions for species synchrony under special limiting cases.
-\hl{The limiting case predictions we derive serve as baselines against which we can compare, and thus understand, results from empirically-based simulations (described below).}
+\hl{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
 We focus on synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
@@
 \hl{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos} [@Loreau2008a].
-\hl{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are typically more well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small} (e.g., $r < 2$).
+\hl{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small} (e.g., $r < 2$).
 
@@
 \noindent{} where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
-Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in monoculture (no interspecific interactions) to calculate the covariance of species' environmental responses.
+Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in \hl{growing in isolation} (no interspecific interactions) to calculate the covariance of species' environmental responses.
 To estimate environmental responses in natural communities, we turn to our population models built using long-term demographic data.
@@
 
-\noindent where $\sigma_{r_{i}}$ is the temporal variance of species *i*'s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
+\noindent where $\sigma_{r_{i}}$ is the temporal variance of species *i'*s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
 $\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [@Loreau2008a].
@@
 
-We modeled recruitment at the quadrat scale, rather than the individual scale, because it is impossible to attribute new genets to specific parents [@Chu2015].
+We modeled recruitment at the quadrat scale, rather than the individual scale, because the original data do not attribute new genets to specific parents [@Chu2015].
 Our recruitment model assumes that the number of recruits produced in each year follows a negative binomial distribution with the mean dependent on the cover of the parent species, permanent spatial variation among groups, temporal variation among years, and inter- and intraspecific interactions as a function of total species' cover in the quadrat.
@@
 #### Simulation experiments
-We performed removal simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific \hl{interactions}) were removed one-by-one and in combination.
+We performed simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific \hl{interactions}) were removed one-by-one and in combination.
 To remove interspecific interactions, we set the off-diagonals of the interaction matrix for each vital rate regression to zero.
 This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are \hl{growing in isolation}.
-\hl{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a species in a no-interaction model.}
+\hl{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
 To remove the effect of a fluctuating environment, we removed the temporal (interannual) random effects from the regression equations.
@@
 Results from our simulation experiments also allow us to test our theoretical predictions.
-First, in the absence of \hl{interspecific interactions} and demographic stochasticity, \hl{populations fluctuate in response to the environment alone}.
+First, in the absence of \hl{interspecific interactions} and demographic stochasticity, \hl{populations can only fluctuate in response to the environment}.
 Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(u_{ie}, u_{je})$) and parameterize Equation 4.
@@
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
-1), with an average 6% increase over synchrony from IBM simulations on a five meter area.
+1), with an average 6% increase over synchrony from IBM simulations on a 25$\text{m}^2$ area.
 
@@
 Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
-\hl{Observed synchrony of per capita growth rates were positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
+\hl{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
 
@@
 \hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
-However, the theoretical predictions were strongly supported by simulations of population models fit to long-term data from natural plant communities (Table 3).
-Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony, \hl{and allows us to place our simulation results within the context of contemporary theory}.
+However, the theoretical predictions \hl{closely matched results from} simulations of population models fit to long-term data from natural plant communities (Table 3).
+Such \hl{strong agreement} between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony \hl{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.
 

### 2016-07-01 R/import_regression_params_fxns_yrlycomp.R
adds yearly comp sims for Montana

@@ -35,8 +35,8 @@ format_growth_params_yrly <- function(do_site, species_list, Nyrs, Gdata_species
     if(length(tmp2)>0) Gpars$nb.yr[,,i]=as.matrix(Gdata[,tmp2])
     
     Gpars$alpha[i,]=Gdata$alpha[1:length(species_list)]
-    Gpars$sigma2.a[i]=0#Gdata$sigma.a[1]
-    Gpars$sigma2.b[i]=0#Gdata$sigma.b[1]
+    Gpars$sigma2.a[i]=Gdata$sigma.a[1]
+    Gpars$sigma2.b[i]=Gdata$sigma.b[1]
   } # next i
   return(Gpars)
 } # end function
@@ -83,11 +83,11 @@ format_survival_params_yrly <- function(do_site, species_list, Nyrs, Sdata_speci
     # get competition coefficients
     # get competition coefficients
     tmp=paste("W",1:length(species_list),sep="")
-    tmp=which(is.element(names(Gdata),tmp))
-    if(length(tmp)>0) Gpars$nb[i,]=as.numeric(Gdata[1,tmp])
+    tmp=which(is.element(names(Sdata),tmp))
+    if(length(tmp)>0) Spars$nb[i,]=as.numeric(Sdata[1,tmp])
     tmp2=paste("W",1:length(species_list), ".yr",sep="")
-    tmp2=which(is.element(names(Gdata),tmp2))
-    if(length(tmp2)>0) Gpars$nb.yr[,,i]=as.matrix(Gdata[,tmp2])
+    tmp2=which(is.element(names(Sdata),tmp2))
+    if(length(tmp2)>0) Spars$nb.yr[,,i]=as.matrix(Sdata[,tmp2])
     
     Spars$alpha[i,]=Sdata$alpha[1:length(species_list)]
   } # next i
@@ -113,17 +113,23 @@ format_recruitment_params_yrly <- function(do_site, species_list, Nyrs,
   Rpars <- list(intcpt.mu=rep(0,Nspp),intcpt.yr=matrix(0,Nyrs,Nspp),
                 intcpt.tau=rep(100,Nspp),
                 intcpt.gr=matrix(NA,Ngrp,Nspp),g.tau=rep(NA,Nspp),
-                dd=matrix(NA,Nspp,Nspp),theta=rep(NA,Nspp),
+                dd=matrix(NA,Nspp,Nspp),dd.yr=array(NA, dim=c(Nyrs,Nspp,Nspp)),theta=rep(NA,Nspp),
[…]

### 2016-07-01 R/ipm_kernel_vitalrate_fxns_yrly.R

@@ -8,10 +8,10 @@
 #' @param doYear Current year for getting year level random effects.
 #' @param doSpp Numeric scalar for species.
 
-G <- function(v,u,W,Gpars,doYear,doSpp){
+G_yr <- function(v,u,W,Gpars,doYear,doSpp){
   mu <- Gpars$intcpt[doSpp]+Gpars$intcpt.yr[doYear,doSpp]+
     (Gpars$slope[doSpp]+Gpars$slope.yr[doYear,doSpp])*u+
-    W%*%(Gpars$nb[doSpp,])
+    W%*%(Gpars$nb[doSpp,]+Gpars$nb.yr[doYear,,doSpp])
   sigma2 <- Gpars$sigma2.a[doSpp]*exp(Gpars$sigma2.b[doSpp]*mu)
   out <- dnorm(v,mu,sqrt(sigma2))
   return(out)
@@ -27,10 +27,10 @@ G <- function(v,u,W,Gpars,doYear,doSpp){
 #' @param doYear Current year for getting year level random effects.
 #' @param doSpp Numeric scalar for species.
 
-S <- function(u,W,Spars,doYear,doSpp){
+S_yr <- function(u,W,Spars,doYear,doSpp){
   mu <- Spars$intcpt[doSpp]+Spars$intcpt.yr[doYear,doSpp]+
     (Spars$slope[doSpp]+Spars$slope.yr[doYear,doSpp])*u+
-    W%*%(Spars$nb[doSpp,])
+    W%*%(Spars$nb[doSpp,]+Spars$nb.yr[doYear,,doSpp])
   return(inv.logit(mu))
 }
 
@@ -43,14 +43,14 @@ S <- function(u,W,Spars,doYear,doSpp){
 #' @param cover Current time step cover in m^2. Length is num_species.
 #' @param doYear Numeric scalar for current simulation year.
 
-get_rpa=function(Rpars,cover,doYear,A){
+get_rpa_yr=function(Rpars,cover,doYear,A){
   # cover is in m^2 per m^2; convert to % scale:
   cover2=cover*100
   # calculate recruits
   Nspp=length(cover)
   mu=rep(NA,Nspp)
   for(i in 1:Nspp){
-    mu[i]=cover2[i]*exp(Rpars$intcpt.yr[doYear,i]+sqrt(cover2)%*%Rpars$dd[i,]) 
[…]

### 2016-07-01 runscripts/run_ipm_simulations_yrlcomp.R

@@ -34,7 +34,7 @@ library(statmod)
 ## Make combined kernel
 make.K.values=function(v,u,muWG,muWS, #state variables
                        Rpars,rpa,Gpars,Spars,doYear,doSpp){  #growth arguments
-  f(v,u,Rpars,rpa,doSpp)+S(u,muWS,Spars,doYear,doSpp)*G(v,u,muWG,Gpars,doYear,doSpp) 
+  f_yr(v,u,Rpars,rpa,doSpp)+S_yr(u,muWS,Spars,doYear,doSpp)*G_yr(v,u,muWG,Gpars,doYear,doSpp) 
 }
 
 ## Function to make iteration matrix based only on mean crowding
@@ -115,11 +115,11 @@ for(do_site in "Montana"){
   
   #Import and format parameters
   site_path <- paste("../data/", do_site, sep="")
-  Gpars <- format_growth_params(do_site = do_site, species_list = spp_list, 
+  Gpars <- format_growth_params_yrly(do_site = do_site, species_list = spp_list, 
                                 Nyrs = Nyrs, Gdata_species = Gpars_tmp)
-  Spars <- format_survival_params(do_site = do_site, species_list = spp_list, 
+  Spars <- format_survival_params_yrly(do_site = do_site, species_list = spp_list, 
                                   Nyrs = Nyrs, Sdata_species = Spars_tmp)
-  Rpars <- format_recruitment_params(do_site = do_site, species_list = spp_list, 
+  Rpars <- format_recruitment_params_yrly(do_site = do_site, species_list = spp_list, 
                                      Nyrs = Nyrs, Rdata_species = Rpars_tmp,
                                      path_to_site_data = site_path)
   
@@ -170,24 +170,41 @@ for(do_site in "Montana"){
     
     # Turn off competition if spp_interact==FALSE
     if(spp_interact==FALSE){
-      rnbtmp <- Rpars$dd
+      rnbtmp <- Rpars$dd.yr
       rnbtmp[] <- 0
-      diag(rnbtmp) <- diag(Rpars$dd)
-      Rpars$dd <- rnbtmp
+      for(ispp in 1:Nspp){
+        rtmp_intra <- Rpars$dd.yr[,ispp,ispp]
+        rnbtmp[,ispp,ispp] <- rtmp_intra
+      }
+      Rpars$dd.yr <- rnbtmp
       
       gnbtmp <- Gpars$nb
       gnbtmp[] <- 0
[…]

### 2016-07-01 devel/plot_montana_ipm_yearlycomp.R
adds montana figure with yearly comp removed

@@ -1,0 +1,71 @@
+######################################################################################
+##  plot_montana_ipm_yearlycomp.R: script to calculate and plot the synchrony       ##
+##  of species in the Montana community with and without yearly competition effects ##
+######################################################################################
+
+rm(list=ls(all.names = TRUE)) # clear the workspace
+
+library(reshape2)
+library(plyr)
+library(ggplot2)
+library(communitySynchrony)
+library(synchrony)
+library(ggthemes)
+
+
+
+##  Read in IPM results ---------
+output_list <- readRDS("../results/ipm_comp_nocomp_sims_yearlycomp_Montana.RDS")
+mlist <- melt(output_list)
+colnames(mlist)[1:3] <- c("year", "species", "cover")
+sites <- unique(mlist$L1)
+sims <- unique(mlist$L2)
+synch_df <- data.frame(site=NA, experiment=NA, bootnum=NA, 
+                       pgr_synch=NA, abund_synch=NA)
+boots <- 100
+num_iters <- 50
+for(dosite in "Montana"){
+  tmp_data <- subset(mlist, L1==dosite)
+  for(dosim in sims){
+    tmpsim <- subset(tmp_data, L2==dosim)[c("year", "species", "cover")]
+    for(i in 1:boots){
+      begin_year <- sample(x = 1:(max(tmpsim$year)-num_iters), 1)
+      end_year <- begin_year+num_iters
+      tmp <- subset(tmpsim, year %in% begin_year:end_year)
+      tmpsynch <- get_ipm_synchrony(tmp)
+      tmp_pgr_synch <- as.numeric(tmpsynch$pgr_synchrony["obs"])
+      tmpcast <- dcast(tmp, year~species, value.var = "cover")
+      tmp_abund_synch <- as.numeric(community.sync(tmpcast[2:ncol(tmpcast)])[1])
+      tmp_df <- data.frame(site=dosite, experiment=dosim, 
+                           bootnum=i, pgr_synch=tmp_pgr_synch, 
+                           abund_synch=tmp_abund_synch)
+      synch_df <- rbind(synch_df, tmp_df)
+    }# end boots loop
+  }# end experiment/sim loop
+}# end site loop
+
+synch_dftmp <- synch_df[2:nrow(synch_df),]
+synch_df <- melt(synch_dftmp, id.vars = c("site", "experiment", "bootnum"))
[…]

### 2016-07-01 docs/ms_communitySynchrony.Rmd
adding ms back in

@@
 
-\begin{abstract}
-Temporal asynchrony among species \hl{helps} diversity \hl{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
-Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \hl{interactions}, and demographic stochasticity.
+Abstract
+--------
+Temporal asynchrony among species \textcolot{blue{helps} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \textcolor{blue}{interactions}, and demographic stochasticity.
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-\hl{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
-Removing interspecific \hl{interactions} and demographic stochasticity had small effects on synchrony.
+\textcolor{blue}{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Removing interspecific \textcolor{blue}{interactions} and demographic stochasticity had small effects on synchrony.
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
-\end{abstract}
 
@@
 Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
-The stabilizing effect of species richness stems from individual species responding \hl{in different ways} to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
+The stabilizing effect of species richness stems from individual species responding \textcolor{blue}{in different ways} to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
 Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
 Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
-\hl{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}
+\textcolor{blue}{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}
 
-Theory identifies three main determinants of species synchrony: environmental stochasticity, demographic stochasticity, and \hl{interspecific interactions} [@Loreau2008a; @Gonzalez2009; @Loreau2013].
-For example, in a community composed of large \hl{populations} (no demographic stochasticity) \hl{with weak interspecific interations}, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
-However, \hl{this} prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
-\hl{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \hl{and they did not consider demographic stochasiticity.}
+Theory identifies three main determinants of species synchrony: environmental stochasticity, demographic stochasticity, and \textcolor{blue}{interspecific interactions} [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+For example, in a community composed of large \textcolor{blue}{populations} (no demographic stochasticity) \textcolor{blue}{with weak interspecific interations}, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
+However, \textcolor{blue}{this} prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
+\textcolor{blue}{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \textcolor{blue}{and they did not consider demographic stochasiticity.}
 
@@
 Even beyond grassland studies, whether competition or environmental responses drive compensatory dynamics remains controversial [reviewed in @Gonzalez2009].
-In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from \hl{the covariance matrix of species abundances} [e.g., @Houlahan2007] is impossible.
+In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from \textcolor{blue}{the covariance matrix of species abundances} [e.g., @Houlahan2007] is impossible.
 This is because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
@@
 
-The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} is to remove them one-by-one, and in combination.
+The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific \textcolor{blue}{interactions} is to remove them one-by-one, and in combination.
 In principle, this could be done in an extremely controlled laboratory setting, but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
@@
 They used a statistical technique to decompose temporal variation into competition and environmental components, and found that positively correlated environmental responses among species determined community dynamics.
-@Thibaut2012 \hl{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
+@Thibaut2012 \textcolor{blue}{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
 While a major step forward, Mutshinda et al.'s (2009) and Thibaut et al.'s (2012) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Furthermore, although both studies quantified the relative importance of environmental stochasticity and \hl{interspecific interactions} to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
+Furthermore, although both studies quantified the relative importance of environmental stochasticity and \textcolor{blue}{interspecific interactions} to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
 
 Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony.
-Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} on species synchrony in natural plant communities.
+Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific \textcolor{blue}{interactions} on species synchrony in natural plant communities.
 To achieve these objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
 We then confront our theoretical predictions with simulations from the empirically-based population models.
-Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} on community-wide species synchrony.
+Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific \textcolor{blue}{interactions} on community-wide species synchrony.
 Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, and demographic and environmental stochasticity), and that these models successfully reproduce observed community dynamics [@Chu2015], perturbing the models can reveal the processes that determine species synchrony in our focal grassland communities.
@@
 However, we can derive analytical predictions for species synchrony under special limiting cases.
-\hl{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
+\textcolor{blue}{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
 We focus on synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
@@
 To derive analytical predictions we solved a first-order approximation of Equation 2 [@DeMazancourt2013a and Online Supporting Information].
-\hl{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos} [@Loreau2008a].
-\hl{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small} (e.g., $r < 2$).
+\textcolor{blue}{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos} [@Loreau2008a].
+\textcolor{blue}{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small} (e.g., $r < 2$).
 
@@
 This limiting case, $\mathcal{M}_{D}$, represents a community where dynamics are driven by demographic stochasticity alone.
-Our prediction for the synchrony of per capita growth rates for $\mathcal{M}_{D}$, $\phi_{R,\mathcal{M}_{D}}$, is
+Our prediction \textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{D}$, $\phi_{R,\mathcal{M}_{D}}$, is}
 
@@
 Thus, we assume there are no interspecific interactions, demographic stochasticity is absent, intrinsic growth rates are identical, and environmental variance is identical for all species.
-Our prediction for the synchrony of per capita growth rates for $\mathcal{M}_{E}$, $\phi_{R,\mathcal{M}_{E}}$, is
+Our prediction \textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{E}$, $\phi_{R,\mathcal{M}_{E}}$, is }
 
@@
 \noindent{} where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
-Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in \hl{growing in isolation} (no interspecific interactions) to calculate the covariance of species' environmental responses.
+Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in \textcolor{blue}{growing in isolation} (no interspecific interactions) to calculate the covariance of species' environmental responses.
 To estimate environmental responses in natural communities, we turn to our population models built using long-term demographic data.
@@
 We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
-\hl{Crowding was treated as a fixed effect without a temporal component because the 95\% credible intervals for random year effects on crowding broadly overlapped zero.}
+\textcolor{blue}{Crowding was treated as a fixed effect without a temporal component because the 95\% credible intervals for random year effects on crowding broadly overlapped zero.}
 Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
@@
 #### Simulation experiments
-We performed simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific \hl{interactions}) were removed one-by-one and in combination.
+We performed simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific \textcolor{blue}{interactions}) were removed one-by-one and in combination.
 To remove interspecific interactions, we set the off-diagonals of the interaction matrix for each vital rate regression to zero.
-This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are \hl{growing in isolation}.
-\hl{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
+This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are \textcolor{blue}{growing in isolation}.
+\textcolor{blue}{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
 To remove the effect of a fluctuating environment, we removed the temporal (interannual) random effects from the regression equations.
@@
 We conducted the following six simulation experiments:
-(1) IBM: All drivers (environmental stochasticity, demographic stochasticity, or \hl{interspecific interactions}) present;
+(1) IBM: All drivers (environmental stochasticity, demographic stochasticity, or \textcolor{blue}{interspecific interactions}) present;
 (2) IPM: Demographic stochasticity removed;
 (3) IBM: Environmental stochasticity removed;
-(4) IBM: \hl{Interspecific interactions} removed;
-(5) IPM: \hl{Interspecific interactions} and demographic stochasticity removed;
-(6) IBM: \hl{Interspecific interactions} and environmental stochasticity removed.
+(4) IBM: \textcolor{blue}{Interspecific interactions} removed;
+(5) IPM: \textcolor{blue}{Interspecific interactions} and demographic stochasticity removed;
+(6) IBM: \textcolor{blue}{Interspecific interactions} and environmental stochasticity removed.
 We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
@@
 Results from our simulation experiments also allow us to test our theoretical predictions.
-First, in the absence of \hl{interspecific interactions} and demographic stochasticity, \hl{populations can only fluctuate in response to the environment}.
+First, in the absence of \textcolor{blue}{interspecific interactions} and demographic stochasticity, \textcolor{blue}{populations can only fluctuate in response to the environment}.
 Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(u_{ie}, u_{je})$) and parameterize Equation 4.
@@
 Synchrony is lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
-\hl{Synchrony of per capita growth rates and CV of percent cover are positively correlated} (Pearson’s $\rho$ = 0.72).
+\textcolor{blue}{Synchrony of per capita growth rates and CV of percent cover are positively correlated} (Pearson’s $\rho$ = 0.72).
 For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
@@
 Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S3).
-Species interactions are weak in these communities [Table S4 and @Chu2015], and removing \hl{interspecific interactions} had little effect on synchrony (Fig. 1; "No Comp." simulations).
-Removing \hl{interspecific interactions} caused, at most, a 5% change in synchrony (Fig. 1).
+Species interactions are weak in these communities [Table S4 and @Chu2015], and removing \textcolor{blue}{interspecific interactions} had little effect on synchrony (Fig. 1; "No Comp." simulations).
+Removing \textcolor{blue}{interspecific interactions} caused, at most, a 5% change in synchrony (Fig. 1).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
@@
 Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
-\hl{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
+\textcolor{blue}{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
 
@@
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
-\hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
-However, the theoretical predictions \hl{closely matched results from} simulations of population models fit to long-term data from natural plant communities (Table 3).
-Such \hl{strong agreement} between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony \hl{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.
+\textcolor{blue}{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
+However, the theoretical predictions \textcolor{blue}{closely matched results from} simulations of population models fit to long-term data from natural plant communities (Table 3).
+Such \textcolor{blue}{strong agreement} between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony \textcolor{blue}{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.
 
@@
 2), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity was not strong enough to \hl{compensate} the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
+Even in small populations, however, demographic stochasticity was not strong enough to \textcolor{blue}{compensate} the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
 2).
@@
 Only in the Idaho community does synchrony of per capita growth rates approach $\mathcal{M}_D$ in a non-constant environment (Fig. 2).
-This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a \hl{2}5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
+This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a \textcolor{blue}{2}5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
 
@@
 A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
-\hl{Most communities are dominated by few common species and many rare species}, \hl{meaning that the low number of common species in our focal commiunities is not unusual.} 
-\hl{Rather, the generality of our results hinges upon the influence of rare species.
+\textcolor{blue}{Most communities are dominated by few common species and many rare species}, \textcolor{blue}{meaning that the low number of common species in our focal commiunities is not unusual.} 
+\textcolor{blue}{Rather, the generality of our results hinges upon the influence of rare species.
 Rare species could be kept at low densities by competitive suppression due to strong interspecific interactions.
 If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
-However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species} [@Yenni2012], \hl{in which case synchrony would remain predominantly driven by environmental responses.}
+However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species} [@Yenni2012], \textcolor{blue}{in which case synchrony would remain predominantly driven by environmental responses.}
 Rare species could also be limited by generalist natural enemies whose effects are density-independent but alter competitive hierarchies [@Mordecai2011].
@@
 We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
-The lack of an effect of \hl{interspecific interactions} on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
+The lack of an effect of \textcolor{blue}{interspecific interactions} on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
 
@@
 
-\hl{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
-Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap} [@Vasseur2007], \hl{dispersal} [@Gouhier2010], \hl{and the strength of top-down control} [@Bauer2014] \hl{are all likely to affect community synchrony.}
+\textcolor{blue}{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
+Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap} [@Vasseur2007], \textcolor{blue}{dispersal} [@Gouhier2010], \textcolor{blue}{and the strength of top-down control} [@Bauer2014] \textcolor{blue}{are all likely to affect community synchrony.}
 
@@
 Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
-Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.
-All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
+\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.} All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
 Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
@@
 + E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
+\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}}
 \end{figure}

### 2016-07-02 docs/ms_communitySynchrony.Rmd
final edits for ELE resubmission

@@
     fig_caption: yes
-    keep_tex: no
+    keep_tex: yes
 urlcolor: black
@@
 
-\textbf{Number of Words}: Abstract = 138; Main Text = 4929 (including section titles and in-line citations)
+\textbf{Number of Words}: Abstract = 136; Main Text = 4992
 
@@
 
-Abstract
---------
-Temporal asynchrony among species \textcolot{blue{helps} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+\begin{abstract}
+Temporal asynchrony among species \textcolor{blue}{helps} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
 Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \textcolor{blue}{interactions}, and demographic stochasticity.
@@
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+\end{abstract}
 
@@
 \noindent{} where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
-Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in \textcolor{blue}{growing in isolation} (no interspecific interactions) to calculate the covariance of species' environmental responses.
+Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) \textcolor{blue}{growing in isolation} (no interspecific interactions) to calculate the covariance of species' environmental responses.
 To estimate environmental responses in natural communities, we turn to our population models built using long-term demographic data.
@@
 We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
-\textcolor{blue}{Crowding was treated as a fixed effect without a temporal component because the 95\% credible intervals for random year effects on crowding broadly overlapped zero.}
+\textcolor{blue}{Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.}
 Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
@@
 A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
-\textcolor{blue}{Most communities are dominated by few common species and many rare species}, \textcolor{blue}{meaning that the low number of common species in our focal commiunities is not unusual.} 
+\textcolor{blue}{Most communities are dominated by few common species and many rare species} [@McGill2007], \textcolor{blue}{meaning that the low number of common species in our focal commiunities is not unusual.}
 \textcolor{blue}{Rather, the generality of our results hinges upon the influence of rare species.
@@
 Compute, storage, and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.
+We also thank three anonymous reviewers who provided critical feedback that improved the manuscript.
 

### 2016-07-02 docs/ms_communitySynchrony.tex
final edits for ELE resubmission

--- 2016-04-01 docs/ms_communitySynchrony.tex
@@
 \doublespacing
-\usepackage{color}
 \usepackage{rotating}
+\usepackage{color, soul}
 
@@
 \begin{abstract}
-Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
-Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific competition, and demographic stochasticity.
+Temporal asynchrony among species \textcolor{blue}{helps} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \textcolor{blue}{interactions}, and demographic stochasticity.
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-Simulation experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
-Removing interspecific competition and demographic stochasticity had small effects on synchrony.
+\textcolor{blue}{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Removing interspecific \textcolor{blue}{interactions} and demographic stochasticity had small effects on synchrony.
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
@@
 2010). The stabilizing effect of species richness stems from individual
-species responding to environmental fluctuations (environmental
-stochasticity), or fluctuating asynchronously because of random chance
-events (demographic stochasticity). (Isbell et al. 2009, Hector et al.
-2010, {{de Mazancourt}} et al. 2013). Species richness affects synchrony
-because larger species pools are more likely to contain species that
-respond disimilarly to environmental conditions (Yachi and Loreau 1999),
+species responding \textcolor{blue}{in different ways} to environmental
+fluctuations (environmental stochasticity), or fluctuating
+asynchronously because of random chance events (demographic
+stochasticity) (Isbell et al. 2009, Hector et al. 2010, {{de
+Mazancourt}} et al. 2013). Species richness affects synchrony because
+larger species pools are more likely to contain species that respond
+disimilarly to environmental conditions (Yachi and Loreau 1999),
 implying that species losses will reduce ecosystem stability. Even
@@
 dynamics and, in turn, decrease temporal stability of ecosystem
-functioning (Hautier et al. 2014). Predicting the impacts of global
-change requires a mechanistic understanding of species synchrony in
-natural settings.
-
-Theory identifies the three main determinants of species synchrony as
+functioning (Hautier et al. 2014).
+\textcolor{blue}{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}
+
+Theory identifies three main determinants of species synchrony:
 environmental stochasticity, demographic stochasticity, and
-interspecific competition (Loreau and {{de Mazancourt}} 2008, 2013,
-Gonzalez and Loreau 2009). For example, in a community composed of large
-(no demographic stochasticity) and weakly interacting (no interspecific
-competition) populations, community-wide species synchrony should be
-determined by the covariance of species' responses to the environment
-(Loreau and {{de Mazancourt}} 2008). However, such a prediction relies
-on a relatively simple population model and only holds under two
-assumptions: (i) species' responses to the environment are similar in
-magnitude and (ii) all species have similar growth rates. Whether such
-theoretical predictions hold in natural communities where species
-differences are unlikely to be symmetrical remains unkown: no study has
-explicitly tested theory on the drivers of species synchrony in natural
-communities.
+\textcolor{blue}{interspecific interactions} (Loreau and {{de
+Mazancourt}} 2008, 2013, Gonzalez and Loreau 2009). For example, in a
+community composed of large \textcolor{blue}{populations} (no
+demographic stochasticity)
+\textcolor{blue}{with weak interspecific interations}, community-wide
+species synchrony should be determined by the covariance of species'
+responses to the environment (Loreau and {{de Mazancourt}} 2008).
+However, \textcolor{blue}{this} prediction relies on a relatively simple
+population model and only holds under two assumptions: (i) species'
+responses to the environment are similar in magnitude and (ii) all
+species have similar growth rates.
+\textcolor{blue}{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities}
+(Mutshinda et al. 2009, Thibaut et al. 2012),
+\textcolor{blue}{and they did not consider demographic stochasiticity.}
 
@@
 controversy remains because quantifying the relative strengths of each
-driver on the degree of synchrony from community time series (e.g.,
+driver on the degree of synchrony from
+\textcolor{blue}{the covariance matrix of species abundances} (e.g.,
 Houlahan et al. 2007) is impossible. This is because an unbiased null
@@
 The best way to quantify the effects of environmental stochasticity,
-demographic stochasticity, and interspecific competition is to remove
-them one-by-one, and in combination. In principle, this could be done in
-an extremely controlled laboratory setting, but empirically-based models
-of interacting populations, fit with data sets from natural communities,
-offer a practical alternative. For example, Mutshinda et al. (2009) fit
-a dynamic population model to several community time series of insect
-and bird abundances. They used a statistical technique to decompose
-temporal variation into competition and environmental components, and
-found that postively correlated environmental responses among species
-determined community dynamics. While a major step forward, Mutshinda et
-al.'s (2009) modeling technique relied on abundance data that may or may
+demographic stochasticity, and interspecific
+\textcolor{blue}{interactions} is to remove them one-by-one, and in
+combination. In principle, this could be done in an extremely controlled
+laboratory setting, but empirically-based models of interacting
+populations, fit with data sets from natural communities, offer a
+practical alternative. For example, Mutshinda et al. (2009) fit a
+dynamic population model to several community time series of insect and
+bird abundances. They used a statistical technique to decompose temporal
+variation into competition and environmental components, and found that
+positively correlated environmental responses among species determined
+community dynamics. Thibaut et al. (2012)
+\textcolor{blue}{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
+While a major step forward, Mutshinda et al.'s (2009) and Thibaut et
+al.'s (2012) modeling technique relied on abundance data that may or may
 not reliably capture competitive interactions that occur at the
-individual level. Furthermore, although Mutshinda et al. (2009)
-quantified the relative importance of environmental stochasticity and
-competition to explain the observed variation of species synchrony, they
-did not use the model to quantify how much synchrony would change when
-each factor is removed.
+individual level. Furthermore, although both studies quantified the
+relative importance of environmental stochasticity and
+\textcolor{blue}{interspecific interactions} to explain the observed
+variation of species synchrony, they did not use the model to quantify
+how much synchrony would change when each factor is removed.
 
@@
 relative influence of environmental stochasticity, demographic
-stochasticity, and interspecific competition on species synchrony in
-natural plant communities. To achieve these objectives, we first refine
-theory that has been used to predict the effects of species richness on
-ecosystem stability ({{de Mazancourt}} et al. 2013) and species
-synchrony (Loreau and {{de Mazancourt}} 2008) to generate predictions of
-community-wide species synchrony under two limiting cases derived from
-the dynamics of individual species in monoculture. We then confront our
-theoretical predictions with simulations from the empirically-based
-population models. Second, we use the multi-species population models to
-perform simulation experiments that isolate the effects of environmental
-stochasticity, demographic stochasticity, and interspecific competition
-on community-wide species synchrony. Given that our population models
+stochasticity, and interspecific \textcolor{blue}{interactions} on
+species synchrony in natural plant communities. To achieve these
+objectives, we first refine theory that has been used to predict the
+effects of species richness on ecosystem stability ({{de Mazancourt}} et
+al. 2013) and species synchrony (Loreau and {{de Mazancourt}} 2008) to
+generate predictions of community-wide species synchrony under two
+limiting cases derived from the dynamics of individual species in
+monoculture. We then confront our theoretical predictions with
+simulations from the empirically-based population models. Second, we use
+the multi-species population models to perform simulation experiments
+that isolate the effects of environmental stochasticity, demographic
+stochasticity, and interspecific \textcolor{blue}{interactions} on
+community-wide species synchrony. Given that our population models
 capture the essential features of community dynamics important to
@@
 derive analytical predictions for species synchrony under special
-limiting cases. We focus on synchrony of per capita growth rates, rather
-than abundances, because growth rates represent the instantaneous
-response of species to the environment and competition, and are less
-susceptible to the legacy effects of drift and disturbance (Loreau and
-{{de Mazancourt}} 2008). We present equivalent results for synchrony of
+limiting cases.
+\textcolor{blue}{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
+We focus on synchrony of per capita growth rates, rather than
+abundances, because growth rates represent the instantaneous response of
+species to the environment and competition, and are less susceptible to
+the legacy effects of drift and disturbance (Loreau and {{de
+Mazancourt}} 2008). We present equivalent results for synchrony of
 species abundances in the Online Supporting Information, and show that
@@
 2 ({{de Mazancourt}} et al. 2013 and Online Supporting Information).
+\textcolor{blue}{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos}
+(Loreau and {{de Mazancourt}} 2008).
+\textcolor{blue}{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small}
+(e.g., \(r < 2\)).
 
@@
 represents a community where dynamics are driven by demographic
-stochasticity alone. Our prediction is
+stochasticity alone. Our prediction
+\textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{D}$, $\phi_{R,\mathcal{M}_{D}}$, is}
 
 \begin{equation}
-\phi_{R,\mathcal{M}_{D}} = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
+\phi_{R,\mathcal{M}_{D}} = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2},
 \end{equation}
 
-\noindent for synchrony of per capita growth rates. \(p_i\) is the
-average frequency of species \emph{i}, \(p_i = N_i/N_T\). When all
-species have identical abundances and \(p_i = 1/S\), where \emph{S} is
-species richness, synchrony equal 1/S (Loreau and {{de Mazancourt}}
-2008).
+\noindent where \(p_i\) is the average frequency of species \emph{i},
+\(p_i = N_i/N_T\). When all species have identical abundances and
+\(p_i = 1/S\), where \emph{S} is species richness, synchrony equal 1/S
+(Loreau and {{de Mazancourt}} 2008).
 
 Our second limiting case assumes only environmental stochasticity is
-operating. Thus, there are no interspecific interactions, demographic
-stochasticity is absent, intrinsic growth rates are identical, and
-environmental variance is identical for all species. Our prediction for
-this limiting case, \(\mathcal{M}_{E}\), is
+operating (\(\mathcal{M}_{E}\)). Thus, we assume there are no
+interspecific interactions, demographic stochasticity is absent,
+intrinsic growth rates are identical, and environmental variance is
+identical for all species. Our prediction
+\textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{E}$, $\phi_{R,\mathcal{M}_{E}}$, is }
 
 \begin{equation}
-\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(u_{ei},u_{ej})}{S^2}
+\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(u_{ei},u_{ej})}{S^2},
 \end{equation}
 
-\noindent{} for synchrony of per capita growth rates, where
-\(\text{cov}(u_{ei},u_{ej})\) is the standardized covariance of
-environmental responses between species \emph{i} and species \emph{j}.
-Confronting our theoretical predictions with data requires estimates of
-species dynamics of large populations (no demographic stochasticity) in
-monoculture (no interspecific interactions) to calculate the covariance
-of species' environmental responses. To estimate environmental responses
-in natural communities, we turn to our population models built using
-long-term demographic data.
+\noindent{} where \(\text{cov}(u_{ei},u_{ej})\) is the standardized
+covariance of environmental responses between species \emph{i} and
+species \emph{j}. Confronting our theoretical predictions with data
+requires estimates of species dynamics of large populations (no
+demographic stochasticity) \textcolor{blue}{growing in isolation} (no
+interspecific interactions) to calculate the covariance of species'
+environmental responses. To estimate environmental responses in natural
+communities, we turn to our population models built using long-term
+demographic data.
 
@@
 
-In the early 20th century, several range experiment stations in the
-western U.S. began mapping permanent quadrats, and conducted annual
-censuses for decades. We use such long-term demographic data sets from
-five semiarid grasslands in the western United States (described in
-detail by Chu and Adler 2015). Each study site includes a set of
-1-\(\text{m}^2\) permanent quadrats within which all individual plants
-were identified and mapped annually using a pantograph (Hill 1920). The
-resulting mapped polygons represent basal cover for grasses and canopy
-cover for shrubs. Data come from the Sonoran desert in Arizona (Anderson
-et al. 2012), sagebrush steppe in Idaho (Zachmann et al. 2010), southern
-mixed prairie in Kansas (Adler et al. 2007), northern mixed prairie in
-Montana (Anderson et al. 2011), and Chihuahuan desert in New Mexico
-(Anderson et al. in preparation, Chu and Adler 2015) (Table 1).
+We use long-term demographic data from five semiarid grasslands in the
+western United States (described in detail by Chu and Adler 2015). Each
+site includes a set of 1-\(\text{m}^2\) permanent quadrats within which
+all individual plants were identified and mapped annually using a
+pantograph (Hill 1920). The resulting mapped polygons represent basal
+cover for grasses and canopy cover for shrubs. Data come from the
+Sonoran desert in Arizona (Anderson et al. 2012), sagebrush steppe in
+Idaho (Zachmann et al. 2010), southern mixed prairie in Kansas (Adler et
+al. 2007), northern mixed prairie in Montana (Anderson et al. 2011), and
+Chihuahuan desert in New Mexico (Anderson et al. in preparation, Chu and
+Adler 2015) (Table 1).
 
@@
 \noindent where \(\sigma_{r_{i}}\) is the temporal variance of species
-\emph{i}`s per capita population growth rate (\(r_i\)) and
+\emph{i'}s per capita population growth rate (\(r_i\)) and
 \(\sigma^{2}_{T}\) is the temporal variance of the aggregate
@@
 via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
-2013). Genet size and crowding were treated as fixed effects; spatial
-(quadrat groupings) variation was treated as a random effect on the
-intercept and temporal (interannual) variation was treated as random
+2013).
+\textcolor{blue}{Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.}
+Spatial (quadrat groupings) variation was treated as a random effect on
+the intercept and temporal (interannual) variation was treated as random
 effects on the intercept and the effect of genet size in the previous
@@
 We modeled recruitment at the quadrat scale, rather than the individual
-scale, because it is impossible to attribute new genets to specific
+scale, because the original data do not attribute new genets to specific
 parents (Chu and Adler 2015). Our recruitment model assumes that the
@@
 hierarchical Bayesian approach implemented in JAGS (Plummer 2003) using
-the `rjags' package (Plummer 2014) in R. Again, temporal and spatial
-variation were treated as random effects.
+the `rjags' package (Plummer 2014) in R (R Core Team 2013). Again,
+temporal and spatial variation were treated as random effects.
 
@@
 
-We performed removal simulation experiments where drivers (environmental
-stochasticity, demographic stochasticity, or interspecific competition)
-were removed one-by-one and in combination. To remove interspecific
-interactions, we set the off-diagonals of the interaction matrix for
-each vital rate regression to zero. This retains intraspecific
-interactions, and thus density-dependence, and results in simulations
-where species are effectively growing in monoculture. To remove the
-effect of a fluctuating environment, we removed the temporal
-(interannual) random effects from the regression equations. To remove
-the effect of demographic stochasticity, we use the IPM rather than the
-IBM because the IPM does not include demographic stochasticity
+We performed simulation experiments where drivers (environmental
+stochasticity, demographic stochasticity, or interspecific
+\textcolor{blue}{interactions}) were removed one-by-one and in
+combination. To remove interspecific interactions, we set the
+off-diagonals of the interaction matrix for each vital rate regression
+to zero. This retains intraspecific interactions, and thus
+density-dependence, and results in simulations where species are
+\textcolor{blue}{growing in isolation}.
+\textcolor{blue}{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
+To remove the effect of a fluctuating environment, we removed the
+temporal (interannual) random effects from the regression equations. To
+remove the effect of demographic stochasticity, we use the IPM rather
+than the IBM because the IPM does not include demographic stochasticity
 (demographic stochasticity cannot be removed from the IBM). Since the
@@
 drivers (environmental stochasticity, demographic stochasticity, or
-interspecific competition) present; (2) IPM: Demographic stochasticity
-removed; (3) IBM: Environmental stochasticity removed; (4) IBM:
-Interspecific competition removed; (5) IPM: Interspecific competition
-and demographic stochasticity removed; (6) IBM: Interspecific
-competition and environmental stochasticity removed. We ran IPM
-simulations for 2,000 time steps, after an initial 500 iteration burn-in
-period. This allowed species time to reach their stable size
-distribution. We then calculated the synchrony of species' per capita
-growth rates over 100 randomly selected contiguous 50 time-step
-sections. We ran IBM simulations for 100 time steps, and repeated the
-simulations 100 times for each simulation experiment. From those, we
-retained only the simulations in which no species went extinct due to
-demographic stochasticity. Synchrony was calculated over the 100 time
-steps for each no extinction run within a model experiment. To explore
-the effect of demographic stochasticity in different sized populations,
-we ran simulations (1) and (6) on plot sizes of 1, 4, 9, 16, and 25
-\(\text{m}^2\). All other IBM simulations were run on a 25
-\(\text{m}^2\) landscape.
+\textcolor{blue}{interspecific interactions}) present; (2) IPM:
+Demographic stochasticity removed; (3) IBM: Environmental stochasticity
+removed; (4) IBM: \textcolor{blue}{Interspecific interactions} removed;
+(5) IPM: \textcolor{blue}{Interspecific interactions} and demographic
+stochasticity removed; (6) IBM:
+\textcolor{blue}{Interspecific interactions} and environmental
+stochasticity removed. We ran IPM simulations for 2,000 time steps,
+after an initial 500 iteration burn-in period. This allowed species time
+to reach their stable size distribution. We then calculated the
+synchrony of species' per capita growth rates over 100 randomly selected
+contiguous 50 time-step sections. We ran IBM simulations for 100 time
+steps, and repeated the simulations 100 times for each simulation
+experiment. From those, we retained only the simulations in which no
+species went extinct due to demographic stochasticity. Synchrony was
+calculated over the 100 time steps for each no extinction run within a
+model experiment. To explore the effect of demographic stochasticity in
+different sized populations, we ran simulations (1) and (6) on plot
+sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM simulations
+were run on a 25 \(\text{m}^2\) landscape.
 
 Results from our simulation experiments also allow us to test our
-theoretical predictions. First, in the absence of interspecific
-competition and demographic stochasticity, species are effectively
-growing in large monocultures. Therefore, we can use results from
-simulation (5) to estimate the covariance of species' responses to the
-environment (\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation
-3. Parameterizing Equation 2 does not require simulation output because
-the only parameters are the species' relative abundances. Second,
+theoretical predictions. First, in the absence of
+\textcolor{blue}{interspecific interactions} and demographic
+stochasticity,
+\textcolor{blue}{populations can only fluctuate in response to the environment}.
+Therefore, we can use results from simulation (5) to estimate the
+covariance of species' responses to the environment
+(\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation 4.
+Parameterizing Equation 3 does not require simulation output because the
+only parameters are the species' relative abundances. Second,
 simulations (5) and (6) represent the simulated version of our limiting
@@
 
-Synchrony of species' per capita growth rates at our study sites ranged
+Synchrony of species' per capita growth rates at our study sites range
 from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to
-0.92 (Table 2). Synchrony tended to be higher in communities with few
+0.92 (Table 2). Synchrony tends to be higher in communities with few
 species (Arizona and New Mexico) and/or with relatively high temporal
-variability in percent cover (e.g., Montana). Synchrony was lowest in
+variability in percent cover (e.g., Montana). Synchrony is lowest in
 Idaho, the only data set that includes two life forms: a shrub and three
-perennial grasses (Table 1). For all five communities, species synchrony
+perennial grasses (Table 1).
+\textcolor{blue}{Synchrony of per capita growth rates and CV of percent cover are positively correlated}
+(Pearson's \(\rho\) = 0.72). For all five communities, species synchrony
 from IPM and IBM simulations closely approximated observed synchrony
@@
 Species interactions are weak in these communities (Table S4 and Chu and
-Adler 2015), and removing interspecific competition had little effect on
-synchrony (Fig. 1; ``No Comp.'' simulations). Removing competition
-caused, at most, a 5\% change in synchrony (Fig. 1). Removing
-demographic stochasticity (``No D.S.'' simulations) caused synchrony to
-increase slightly in all communities (Fig. 1), with an average 6\%
-increase over synchrony from IBM simulations on a five meter area.
+Adler 2015), and removing \textcolor{blue}{interspecific interactions}
+had little effect on synchrony (Fig. 1; ``No Comp.'' simulations).
+Removing \textcolor{blue}{interspecific interactions} caused, at most, a
+5\% change in synchrony (Fig. 1). Removing demographic stochasticity
+(``No D.S.'' simulations) caused synchrony to increase slightly in all
+communities (Fig. 1), with an average 6\% increase over synchrony from
+IBM simulations on a 25\(\text{m}^2\) area.
 
@@
 
-For all five communities, the synchrony of species' growth rates in
-monoculture almost perfectly matched species synchrony in polyculture
-(Fig. 3). Results for synchrony of percent cover are qualitatively
-similar, but simulation results were more variable and less consistent
-with analytical predictions and observed synchrony (Online Supporting
-Information, Figs. S2-S3).
+For all five communities, the synchrony of species' growth rates when
+growing in isolation almost perfectly matched species synchrony in
+polyculture (Fig. 3). Results for synchrony of percent cover are
+qualitatively similar, but simulation results were more variable and
+less consistent with analytical predictions and observed synchrony
+(Online Supporting Information, Figs. S2-S3).
 
@@
 (Loreau and {{de Mazancourt}} 2008).
+\textcolor{blue}{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
 
@@
 population dynamics and required several simplifying assumptions,
-raising questions about their relevance to natural communities. However,
-the theoretical predictions were strongly supported by simulations of
-population models fit to long-term data from natural plant communities
-(Table 3). Such a close match between our analytical predictions and the
-simulation results should inspire confidence in the ability of simple
-models to inform our understanding of species synchrony. The theoretical
-population dynamical model we used here also successfully predicted
-ecosystem stability in several biodiversity-ecosystem functioning
-experiments ({{de Mazancourt}} et al. 2013).
+raising questions about their relevance to natural communities.
+\textcolor{blue}{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
+However, the theoretical predictions
+\textcolor{blue}{closely matched results from} simulations of population
+models fit to long-term data from natural plant communities (Table 3).
+Such \textcolor{blue}{strong agreement} between our analytical
+predictions and the simulation results should inspire confidence in the
+ability of simple models to inform our understanding of species
+synchrony
+\textcolor{blue}{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.
 
@@
 small populations, however, demographic stochasticity was not strong
-enough to completely counteract the synchronizing effects of
+enough to \textcolor{blue}{compensate} the synchronizing effects of
 environmental fluctuations and match the analytical prediction where
@@
 likely due to the strong effect of demographic stochasticity on the
-shrub \emph{Artemisia tripartita} since even a 5 \(\text{m}^2\) quadrat
-would only contain a few individuals of that species.
+shrub \emph{Artemisia tripartita} since even a \textcolor{blue}{2}5
+\(\text{m}^2\) quadrat would only contain a few individuals of that
+species.
 
@@
 simulations consistently reduced synchrony (Fig. 1), and the synchrony
-of species in monoculture was a very strong predictor of synchrony of
+of species in isolation was a very strong predictor of synchrony of
 species in polyculture (Fig. 3). Our results lead us to conclude that
@@
 stochasticity caused a significant reduction in species synchrony (Fig.
-S4). Our main results for Idaho (Fig. 2), with \emph{A. tripartita}
+S8). Our main results for Idaho (Fig. 2), with \emph{A. tripartita}
 included, demonstrate how the processes that determine species synchrony
@@
 A challenge to the generality of our results is that we were only able
-to model common, co-occuring species (see Chu and Adler 2015). Rare
-species, which our analysis does not include, could be kept at low
-densities by competitive suppression due to strong interspecific
-interactions. If we had been able to model rare species, we might have
-found a larger effect of competition on community synchrony. However,
-rare species may occupy rare niches that do not overlap with other
-species, in which case synchrony would remain predominantly driven by
-environmental responses. Rare species could also be limited by
-generalist natural enemies whose effects are density-independent but
-alter competitive hierarchies (Mordecai 2011). Under such conditions,
-synchrony will depend on the degree of pathogen-induced fitness
-differences and the pathogen's response to environmental conditions.
-Neither our model nor current theory explicitly considers the effects of
-pathogens on species synchrony, and this highlights the need for
-theoretical work on the interaction between mechanisms of species
-coexistence and mechanisms of species synchrony (Loreau 2010).
+to model common, co-occuring species (see Chu and Adler 2015).
+\textcolor{blue}{Most communities are dominated by few common species and many rare species}
+(McGill et al. 2007),
+\textcolor{blue}{meaning that the low number of common species in our focal commiunities is not unusual.}
+\textcolor{blue}{Rather, the generality of our results hinges upon the influence of rare species.
+Rare species could be kept at low densities by competitive suppression due to strong interspecific interactions.
+If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
+However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species}
+(Yenni et al. 2012),
+\textcolor{blue}{in which case synchrony would remain predominantly driven by environmental responses.}
+Rare species could also be limited by generalist natural enemies whose
+effects are density-independent but alter competitive hierarchies
+(Mordecai 2011). Under such conditions, synchrony will depend on the
+degree of pathogen-induced fitness differences and the pathogen's
+response to environmental conditions. Neither our model nor current
+theory explicitly considers the effects of pathogens on species
+synchrony, and this highlights the need for theoretical work on the
+interaction between mechanisms of species coexistence and mechanisms of
+species synchrony (Loreau 2010).
 
@@
 between simulations with and without interspecific interactions (Fig.
-2). The lack of an effect of interspecific competition on synchrony is
-in contrast to a large body of theoretical work that predicts a strong
-role for competition in creating compensatory dynamics (Tilman 1988) and
-a recent empirical analysis (Gross et al. 2014).
+2). The lack of an effect of
+\textcolor{blue}{interspecific interactions} on synchrony is in contrast
+to a large body of theoretical work that predicts a strong role for
+competition in creating compensatory dynamics (Tilman 1988) and a recent
+empirical analysis (Gross et al. 2014).
 
@@
 years.
+
+\textcolor{blue}{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
+Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap}
+(Vasseur and Fox 2007), \textcolor{blue}{dispersal} (Gouhier et al.
+2010), \textcolor{blue}{and the strength of top-down control} (Bauer et
+al. 2014)
+\textcolor{blue}{are all likely to affect community synchrony.}
 
@@
 the Office of Research and Graduate Studies at Utah State University are
-gratefully acknowledged.
-
-\setlength{\parindent}{0ex} \newpage{}
+gratefully acknowledged. We also thank three anonymous reviewers who
+provided critical feedback that improved the manuscript.
+
+\setlength{\parindent}{0ex}
+
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
+perennial plants: An embarrassment of niches. Ecology Letters
+13:1019--1029.
+
+Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
+mapped quadrats from Kansas prairie: demographic information for
+herbaceous plants. Ecology 88:2673.
+
+Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
+of semi-desert grassland plants in permanent quadrats mapped from 1915
+to 1947. Ecology 93:1492--1492.
+
+Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
+mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
+92:1703.
+
+Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
+Functional Similarity, and Top-Down Control Drive Synchronization and
+the Reliability of Ecosystem Function. The American Naturalist
+183:394--409.
+
+Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
+P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
+Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
+and its impact on humanity. Nature 489:59--67.
+
+Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
+recruitment stage to stabilize grassland coexistence. Ecological
+Monographs 85:373--392.
+
+{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
+Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
+Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
+Loreau. 2013. Predicting ecosystem stability from community composition
+and biodiversity. Ecology Letters 16:617--625.
+
+Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
+Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
+biodiversity change but not systematic loss. Science 344:296--9.
+
+Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. Annual Review of
+Ecology, Evolution, and Systematics 40:393--414.
+
+Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
+variability in space and time. Methods in Ecology and Evolution
+5:524--533.
+
+Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
+Stability of Food Webs in Metacommunities. The American Naturalist
+175:E16--E34.
+
+Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
+Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
+temporal stability of biomass production: a new analysis of recent
+biodiversity experiments. The American naturalist 183:1--12.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
+on ecosystem functioning: a consensus of current knowledge. Ecological
+Monographs 75:3--35.
+
+Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
+Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
+Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
+Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
+Compensatory dynamics are rare in natural ecological communities.
+Proceedings of the National Academy of Sciences of the United States of
+America 104:3273--3277.
+
+Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
+productivity and the temporal stability of productivity: Patterns and
+processes. Ecology Letters 12:443--451.
+
+Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
+species interaction strength in assembled theoretical competition
+communities. Ecology Letters 2:70--74.
+
+Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
+dynamics in ecology and conservation.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
+for a New Ecological Synthesis.
+
+Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
+drivers: neutral and nonneutral community dynamics in fluctuating
+environments. The American Naturalist 172:E48--E66.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
+Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
+and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
+knowledge and future challenges. Science 294:804--808.
+
+McGill, B. J., R. S. Etienne, J. S. Gray, D. Alonso, M. J. Anderson, H.
+K. Benecha, M. Dornelas, B. J. Enquist, J. L. Green, F. He, A. H.
+Hurlbert, A. E. Magurran, P. A. Marquet, B. A. Maurer, A. Ostling, C. U.
+Soykan, K. I. Ugland, and E. P. White. 2007. Species abundance
+distributions: Moving beyond single prediction theories to integration
+within an ecological framework. Ecology Letters 10:995--1015.
+
+Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
+theory, concepts, and empirical work. Ecological Monographs 81:429--441.
+
+Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
+community dynamics? Proceedings of the Royal Society B: Biological
+Sciences 276:2923--2929.
+
+Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
+Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
+3rd international workshop on distributed statistical computing (dSC
+2003). march.
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+
+R Core Team. 2013. R: A language and environment for statistical
+computing.
+
+Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
+O'Hara. 2008. Detecting compensatory dynamics in competitive communities
+under environmental forcing. Oikos 117:1907--1911.
+
+Rees, M., and S. P. Ellner. 2009. Integral projection models for
+populations in temporally varying environments. Ecological Monographs
+79:575--594.
+
+Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
+projection models: A user's guide.
+
+Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
+species have similar dynamics. Journal of Ecology 99:1453--1459.
+
+Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
+A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
+Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
+C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
+Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
+Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
+Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
+Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
+Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
+Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
+Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
+operating space for humanity. Nature 461:472--475.
+
+Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
+for Latent Gaussian Models Using Integrated Nested Laplace
+Approximations. Journal of the Royal Statistical Society, Series B
+71:319--392.
+
+Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
+Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
+analysis of latent Gaussian models using Integrated Nested Laplace
+Approximaxion.
+
+Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
+189:732--735.
+
+Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
+and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
+
+Tilman, D. 1988. Plant strategies and the dynamics and structure of
+plant communities. Plant strategies and the dynamics and structure of
+plant communities.:360 pp.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
+2001. Diversity and productivity in a long-term grassland experiment.
+Science 294:843--845.
+
+Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
+Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
+W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
+2014. Synchronous dynamics of zooplankton competitors prevail in
+temperate lake ecosystems. Proceedings. Biological sciences / The Royal
+Society 281:20140633.
+
+Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
+stabilize food web dynamics by increasing synchrony. Ecology Letters
+10:1066--1074.
+
+Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
+biodiversity over time. Proceedings of the National Academy of Sciences
+of the United States of America 110:19456--9.
+
+Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
+Human Domination of Earth' s Ecosystems. Science 277:494--499.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
+across scales in metacommunities. Ecology Letters.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+Yenni, G., P. B. Adler, and S. K. {Morgan Ernest}. 2012. Strong
+self-limitation promotes the persistence of rare species. Ecology
+93:456--461.
+
+Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
+sagebrush steppe: long-term data for analyzing demographic rates and
+plant--plant interactions. Ecology 91:3427.
+
+\newpage{}
 
@@
   Kansas & 0.54 & 0.15 & 0.30 &   3 \\ 
-  Montana & 0.53 & 0.58 & 0.54 &   4 \\ 
+  Montana & 0.53 & 0.54 & 0.52 &   4 \\ 
   Idaho & 0.36 & 0.18 & 0.19 &   4 \\ 
@@
   \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
-Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
-+ No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
+Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific interactions present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific interactions removed (``No Comp.''), interspecific interactions and demographic stochasticity removed (``No Comp.
++ No D.S.''), and interspecific interactions and environmental stochasticity removed (``No Comp.
 + No E.S.'').
 Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
-Error bars represent the upper and lower 95\% quantiles from model simulations.
-All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
+\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.} All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
 Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
@@
 + E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}}
 \end{figure}
@@
   \centering
-      \includegraphics[width=4in]{./components/formatted_figures/formatted_figure3.png}
-  \caption{Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions).
-We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
+      \includegraphics[width=3in]{./components/formatted_figures/formatted_figure3.png}
+  \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).
+We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.
 The dashed line is the line of equality.
 Simulation results in this figure are analagous to ``No Comp.
-+ No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.
++ No D.S.'' (species in isolation) and ``No D.S.'' (species in competition) in Fig.
 1, but here we control the time series of random year effects.}
@@
 
-\pagebreak{}
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
-Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
-perennial plants: An embarrassment of niches. Ecology Letters
-13:1019--1029.
-
-Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
-mapped quadrats from Kansas prairie: demographic information for
-herbaceous plants. Ecology 88:2673.
-
-Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
-of semi-desert grassland plants in permanent quadrats mapped from 1915
-to 1947. Ecology 93:1492--1492.
-
-Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
-mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
-92:1703.
-
-Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
-P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
-Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
-and its impact on humanity. Nature 489:59--67.
-
-Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
-recruitment stage to stabilize grassland coexistence. Ecological
-Monographs 85:373--392.
-
-{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
-Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
-Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
-Loreau. 2013. Predicting ecosystem stability from community composition
-and biodiversity. Ecology Letters 16:617--625.
-
-Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
-Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
-biodiversity change but not systematic loss. Science 344:296--9.
-
-Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
-Compensatory Dynamics in Ecological Communities. Annual Review of
-Ecology, Evolution, and Systematics 40:393--414.
-
-Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
-variability in space and time. Methods in Ecology and Evolution
-5:524--533.
-
-Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
-Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
-temporal stability of biomass production: a new analysis of recent
-biodiversity experiments. The American naturalist 183:1--12.
-
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
-
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
-
-Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
-on ecosystem functioning: a consensus of current knowledge. Ecological
-Monographs 75:3--35.
-
-Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
-Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
-Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
-Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
-Compensatory dynamics are rare in natural ecological communities.
-Proceedings of the National Academy of Sciences of the United States of
-America 104:3273--3277.
-
-Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
-productivity and the temporal stability of productivity: Patterns and
-processes. Ecology Letters 12:443--451.
-
-Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
-species interaction strength in assembled theoretical competition
-communities. Ecology Letters 2:70--74.
-
-Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
-dynamics in ecology and conservation.
-
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
-
-Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
-for a New Ecological Synthesis.
-
-Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
-drivers: neutral and nonneutral community dynamics in fluctuating
-environments. The American Naturalist 172:E48--E66.
-
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
-
-Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
-Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
-and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
-knowledge and future challenges. Science 294:804--808.
-
-Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
-theory, concepts, and empirical work. Ecological Monographs 81:429--441.
-
-Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
-community dynamics? Proceedings of the Royal Society B: Biological
-Sciences 276:2923--2929.
-
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
-
-R Core Team. 2013. R: A language and environment for statistical
-computing.
-
-Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
-O'Hara. 2008. Detecting compensatory dynamics in competitive communities
-under environmental forcing. Oikos 117:1907--1911.
-
-Rees, M., and S. P. Ellner. 2009. Integral projection models for
-populations in temporally varying environments. Ecological Monographs
-79:575--594.
-
-Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
-projection models: A user's guide.
-
-Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
-species have similar dynamics. Journal of Ecology 99:1453--1459.
-
-Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
-A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
-Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
-C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
-Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
-Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
-Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
-Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
-Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
-Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
-Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
-operating space for humanity. Nature 461:472--475.
-
-Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
-for Latent Gaussian Models Using Integrated Nested Laplace
-Approximations. Journal of the Royal Statistical Society, Series B
-71:319--392.
-
-Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
-Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
-analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
-
-Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
-189:732--735.
-
-Tilman, D. 1988. Plant strategies and the dynamics and structure of
-plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
-
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
-
-Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
-2001. Diversity and productivity in a long-term grassland experiment.
-Science 294:843--845.
-
-Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
-Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
-W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
-2014. Synchronous dynamics of zooplankton competitors prevail in
-temperate lake ecosystems. Proceedings. Biological sciences / The Royal
-Society 281:20140633.
-
-Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
-2013. Global meta-analysis reveals no net change in local-scale plant
-biodiversity over time. Proceedings of the National Academy of Sciences
-of the United States of America 110:19456--9.
-
-Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
-Human Domination of Earth' s Ecosystems. Science 277:494--499.
-
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
-
-Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
-
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
-
-Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
-sagebrush steppe: long-term data for analyzing demographic rates and
-plant--plant interactions. Ecology 91:3427.
-
 \end{document}

### 2016-07-02 docs/online_supporting_information.Rmd
final edits for ELE resubmission

@@
 vr_do <- vr_files[grep("recruit",vr_files)]
+vr_do <- vr_do[which(vr_do=="recruit_parameters.RDS")]
 tmp_vital <- readRDS(paste0("../results/",vr_do))

### 2016-07-02 docs/INTERACTION_TABLES_MASTER.tex
R1 to ELE version

@@
 + E.S.'').
+The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
 The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
+\end{figure}
+
+\pagebreak{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/environmental_variances_cover.png}
+  \caption{Variance of percent cover for each species (along x-axes) in each site through time from simulations with only environmental stochasticity operating (IPM with no species interactions).}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/environmental_variances_pgr.png}
+  \caption{Variance of per capita growth rates for each species (along x-axes) in each site through time from simulations with only environmental stochasticity operating (IPM with no species interactions).}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/demographic_variances_cover.png}
+  \caption{Variance of percent cover for each species (along x-axes) in each site through time from simulations with only demographic stochasticity operating (IBM simulated on 5 by 5 meter landscape).}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/demographic_variances_pgr.png}
+  \caption{Variance of per capita growth rates for each species (along x-axes) in each site through time from simulations with only demographic stochasticity operating (IBM simulated on 5 by 5 meter landscape).}
 \end{figure}
@@
       \includegraphics[width=3in]{./components/formatted_figures/formatted_figureS4.png}
-  \caption{Community-wide synchrony of species percent cover from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
-Synchrony of species' percent cover for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
+  \caption{Community-wide synchrony of species' growth rates from model simulation experiments for the Idaho community with \textit{Artemisia tripartita} removed.
+Synchrony of species' growth rates are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
 + No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
 + No E.S.'').
-Error bars represent the upper and lower 95\% quantiles from model simulations.}
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
 \end{figure}
@@
   \hline
-New Mexico & 0.56 & 0.51 & 0.86 & 0.85 \\ 
+New Mexico & 0.56 & 0.51 & 0.86 & 0.86 \\ 
   Arizona & 0.51 & 0.52 & 0.82 & 0.81 \\ 
-  Kansas & 0.35 & 0.39 & 0.62 & 0.63 \\ 
-  Montana & 0.31 & 0.33 & 0.47 & 0.49 \\ 
-  Idaho & 0.29 & 0.31 & 0.43 & 0.41 \\ 
+  Kansas & 0.35 & 0.39 & 0.62 & 0.62 \\ 
+  Montana & 0.31 & 0.31 & 0.47 & 0.50 \\ 
+  Idaho & 0.29 & 0.31 & 0.43 & 0.42 \\ 
    \hline
@@
   \hline
-1No D.S. & 5.00 & 4.94 & 7.13 & 4.42 & 10.10 \\ 
-  2No Comp. + No D.S. & 4.25 & 2.08 & 7.61 & 4.39 & 8.10 \\ 
-  4No Comp. & 0.98 & 0.06 & 4.84 & 1.01 & 0.11 \\ 
-  5No E.S. & 40.14 & 33.20 & 47.84 & 38.52 & 9.28 \\ 
-  6No Comp. + No E.S. & 45.95 & 41.48 & 38.21 & 33.07 & 20.33 \\ 
+1No D.S. & 5.53 & 4.28 & 6.01 & 5.27 & 11.95 \\ 
+  2No Comp. + No D.S. & 5.56 & 2.06 & 7.03 & 6.21 & 9.10 \\ 
+  4No Comp. & 0.98 & 0.06 & 4.84 & 1.43 & 0.11 \\ 
+  5No E.S. & 40.14 & 33.20 & 47.84 & 39.06 & 9.28 \\ 
+  6No Comp. + No E.S. & 45.95 & 41.48 & 38.21 & 39.44 & 20.33 \\ 
    \hline
@@
 
-\newpage{}
-
 \begin{table}[ht]
 \centering
-\caption{Average interaction coefficients across vital rates for each community.} 
-\begin{tabular}{lrr}
+\caption{Average interaction coefficients for each vital rate for each community.} 
+\begin{tabular}{llrrr}
   \hline
-site & Interspecific Competition & Intraspecific Competition \\ 
+Site & Interaction Type & Growth & Recruitment & Survival \\ 
   \hline
-Arizona & -0.17 & -0.64 \\ 
-  Idaho & -0.04 & -0.71 \\ 
-  Kansas & 0.02 & -0.35 \\ 
-  Montana & 0.04 & -0.98 \\ 
-  NewMexico & -0.12 & -0.62 \\ 
+Arizona & Interspecific & -0.0111 & -0.4534 & -0.0507 \\ 
+  Arizona & Intraspecific & -0.0402 & -1.4209 & -0.4446 \\ 
+  Idaho & Interspecific & 0.0022 & -0.1091 & 0.0059 \\ 
+  Idaho & Intraspecific & -0.0469 & -1.5569 & -0.5269 \\ 
+  Kansas & Interspecific & -0.0008 & 0.0463 & -0.0041 \\ 
+  Kansas & Intraspecific & -0.0062 & -0.9945 & -0.0586 \\ 
+  Montana & Interspecific & -0.0059 & 0.0691 & 0.0342 \\ 
+  Montana & Intraspecific & -0.0801 & -2.0900 & -0.8652 \\ 
+  NewMexico & Interspecific & -0.0027 & -0.3461 & -0.0080 \\ 
+  NewMexico & Intraspecific & -0.0111 & -1.6473 & -0.2191 \\ 
    \hline
@@
   \hline
-BOER & -0.2791 & -0.0816 \\ 
-  BORO & -0.0200 & -0.6101 \\ 
+BOER & -0.2791 & -0.0814 \\ 
+  BORO & -0.0200 & -0.6102 \\ 
    \hline
@@
   \hline
-BOER & -0.5304 & -0.3063 \\ 
-  BORO & -0.5798 & -2.3237 \\ 
+BOER & -0.5184 & -0.3290 \\ 
+  BORO & -0.5779 & -2.3235 \\ 
    \hline
@@
   \hline
-ARTR & -0.0236 & -0.0202 & 0.0314 & 0.0158 \\ 
-  HECO & -0.0005 & -1.0206 & -0.0052 & 0.0125 \\ 
-  POSE & 0.0001 & 0.0006 & -0.9091 & 0.0121 \\ 
-  PSSP & 0.0009 & -0.0013 & 0.0246 & -0.1543 \\ 
+ARTR & -0.0236 & -0.0199 & 0.0314 & 0.0158 \\ 
+  HECO & -0.0005 & -1.0206 & -0.0051 & 0.0125 \\ 
+  POSE & 0.0001 & 0.0006 & -0.9088 & 0.0121 \\ 
+  PSSP & 0.0009 & -0.0013 & 0.0248 & -0.1547 \\ 
    \hline
@@
   \hline
-ARTR & -0.4054 & 0.2132 & 0.0120 & 0.2389 \\ 
-  HECO & -0.5674 & -1.7722 & -0.1476 & -0.3184 \\ 
-  POSE & -0.2499 & -0.0328 & -2.2873 & -0.1448 \\ 
-  PSSP & -0.2580 & -0.1687 & 0.0365 & -1.7675 \\ 
+ARTR & -0.4003 & 0.2243 & 0.0117 & 0.2360 \\ 
+  HECO & -0.5681 & -1.7594 & -0.1437 & -0.3195 \\ 
+  POSE & -0.2505 & -0.0131 & -2.2928 & -0.1420 \\ 
+  PSSP & -0.2405 & -0.1389 & 0.0347 & -1.7751 \\ 
    \hline
@@
   \hline
-BOCU & -0.1087 & 0.0006 & 0.0010 \\ 
+BOCU & -0.1084 & 0.0006 & 0.0010 \\ 
   BOHI & 0.0021 & -0.0403 & -0.0153 \\ 
-  SCSC & -0.0064 & -0.0066 & -0.0270 \\ 
+  SCSC & -0.0064 & -0.0066 & -0.0271 \\ 
    \hline
@@
   \hline
-BOCU & -1.1584 & -0.0242 & -0.0524 \\ 
-  BOHI & 0.3387 & -0.8620 & 0.0123 \\ 
-  SCSC & 0.0615 & -0.0308 & -0.9543 \\ 
+BOCU & -1.1544 & -0.0312 & -0.0643 \\ 
+  BOHI & 0.3410 & -0.8681 & 0.0041 \\ 
+  SCSC & 0.0619 & -0.0336 & -0.9611 \\ 
    \hline
@@
   \hline
-BOGR & -0.0121 & -0.0002 & -0.0229 & -0.0014 \\ 
-  HECO & -0.0023 & -0.0778 & 0.0142 & 0.0061 \\ 
-  PASM & -0.0005 & -0.0039 & -0.0601 & -0.0033 \\ 
-  POSE & 0.0005 & -0.0101 & -0.0404 & -0.1431 \\ 
+BOGR & -0.0124 & -0.0032 & -0.0258 & -0.0001 \\ 
+  HECO & -0.0025 & -0.0976 & 0.0096 & 0.0079 \\ 
+  PASM & -0.0006 & -0.0041 & -0.0695 & -0.0030 \\ 
+  POSE & 0.0006 & -0.0082 & -0.0417 & -0.1410 \\ 
    \hline
@@
   \hline
-BOGR & -0.1280 & 0.0211 & 0.1243 & 0.0173 \\ 
-  HECO & -0.0043 & -0.5661 & 0.1386 & 0.0013 \\ 
-  PASM & 0.0001 & -0.0044 & -1.3046 & 0.0012 \\ 
-  POSE & 0.0010 & 0.0085 & 0.1198 & -1.2434 \\ 
+BOGR & -0.1316 & 0.0076 & 0.1096 & 0.0126 \\ 
+  HECO & -0.0028 & -0.7064 & 0.1423 & -0.0030 \\ 
+  PASM & -0.0012 & -0.0033 & -1.4026 & 0.0024 \\ 
+  POSE & 0.0014 & 0.0082 & 0.1363 & -1.2203 \\ 
    \hline
@@
   \hline
-BOGR & -0.8916 & -0.2522 & -0.1334 & 0.0610 \\ 
-  HECO & 0.0852 & -1.4191 & -0.0685 & -0.0422 \\ 
-  PASM & 0.1252 & 0.3093 & -4.1823 & 0.3032 \\ 
-  POSE & 0.1577 & 0.4906 & 0.0709 & -1.7065 \\ 
+BOGR & -0.9151 & -0.3076 & -0.0984 & 0.0495 \\ 
+  HECO & -0.0785 & -1.4192 & -0.0480 & -0.0468 \\ 
+  PASM & -0.0103 & 0.2773 & -4.3148 & 0.3470 \\ 
+  POSE & 0.1971 & 0.4854 & 0.0623 & -1.7109 \\ 
    \hline
@@
 \centering
-\caption{Interaction coefficients for growth regressions in New Mexico.} 
+\caption{Interaction coefficients for growth regressions in NewMexico.} 
 \begin{tabular}{rrr}
@@
 \centering
-\caption{Interaction coefficients for survival regressions in New Mexico.} 
+\caption{Interaction coefficients for survival regressions in NewMexico.} 
 \begin{tabular}{rrr}
@@
   \hline
-BOER & -0.1575 & -0.0040 \\ 
-  SPFL & -0.0120 & -0.2807 \\ 
+BOER & -0.1576 & -0.0040 \\ 
+  SPFL & -0.0120 & -0.2806 \\ 
    \hline
@@
 \centering
-\caption{Interaction coefficients for recruitment regressions in New Mexico.} 
+\caption{Interaction coefficients for recruitment regressions in NewMexico.} 
 \begin{tabular}{rrr}
@@
   \hline
-BOER & -0.7631 & -0.6198 \\ 
-  SPFL & -0.0684 & -2.5220 \\ 
+BOER & -0.7649 & -0.6230 \\ 
+  SPFL & -0.0691 & -2.5296 \\ 
    \hline

### 2016-07-02 docs/online_supporting_information.tex
final edits for ELE resubmission
R1 to ELE version

@@
 
+\newpage{}
+
 \begin{table}[ht]
@@
 \end{table}
+
+\newpage{}
 
@@
 
+\newpage{}
+
+\subsubsection{Interaction Coefficients for Vital Rates in Arizona}
+
 \begin{table}[ht]
@@
 \end{tabular}
-\end{table}\begin{table}[ht]
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in Arizona.} 
+\begin{tabular}{rrr}
+  \hline
+ & BOER & BORO \\ 
+  \hline
+BOER & -0.2791 & -0.0814 \\ 
+  BORO & -0.0200 & -0.6102 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for recruitment regressions in Arizona.} 
+\begin{tabular}{rrr}
+  \hline
+ & BOER & BORO \\ 
+  \hline
+BOER & -0.5184 & -0.3290 \\ 
+  BORO & -0.5779 & -2.3235 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\subsubsection{Interaction Coefficients for Vital Rates in Idaho}
+
+\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}\begin{table}[ht]
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in Idaho.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & ARTR & HECO & POSE & PSSP \\ 
+  \hline
+ARTR & -0.0236 & -0.0199 & 0.0314 & 0.0158 \\ 
+  HECO & -0.0005 & -1.0206 & -0.0051 & 0.0125 \\ 
+  POSE & 0.0001 & 0.0006 & -0.9088 & 0.0121 \\ 
+  PSSP & 0.0009 & -0.0013 & 0.0248 & -0.1547 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for recruitment regressions in Idaho.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & ARTR & HECO & POSE & PSSP \\ 
+  \hline
+ARTR & -0.4003 & 0.2243 & 0.0117 & 0.2360 \\ 
+  HECO & -0.5681 & -1.7594 & -0.1437 & -0.3195 \\ 
+  POSE & -0.2505 & -0.0131 & -2.2928 & -0.1420 \\ 
+  PSSP & -0.2405 & -0.1389 & 0.0347 & -1.7751 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\subsubsection{Interaction Coefficients for Vital Rates in Kansas}
+
+\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}\begin{table}[ht]
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in Kansas.} 
+\begin{tabular}{rrrr}
+  \hline
+ & BOCU & BOHI & SCSC \\ 
+  \hline
+BOCU & -0.1084 & 0.0006 & 0.0010 \\ 
+  BOHI & 0.0021 & -0.0403 & -0.0153 \\ 
+  SCSC & -0.0064 & -0.0066 & -0.0271 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for recruitment regressions in Kansas.} 
+\begin{tabular}{rrrr}
+  \hline
+ & BOCU & BOHI & SCSC \\ 
+  \hline
+BOCU & -1.1544 & -0.0312 & -0.0643 \\ 
+  BOHI & 0.3410 & -0.8681 & 0.0041 \\ 
+  SCSC & 0.0619 & -0.0336 & -0.9611 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\subsubsection{Interaction Coefficients for Vital Rates in Montana}
+
+\begin{table}[ht]
 \centering
@@
   \hline
-BOGR & -0.0121 & -0.0002 & -0.0229 & -0.0014 \\ 
-  HECO & -0.0023 & -0.0778 & 0.0142 & 0.0061 \\ 
-  PASM & -0.0005 & -0.0039 & -0.0601 & -0.0033 \\ 
-  POSE & 0.0005 & -0.0101 & -0.0404 & -0.1431 \\ 
+BOGR & -0.0124 & -0.0032 & -0.0258 & -0.0001 \\ 
+  HECO & -0.0025 & -0.0976 & 0.0096 & 0.0079 \\ 
+  PASM & -0.0006 & -0.0041 & -0.0695 & -0.0030 \\ 
+  POSE & 0.0006 & -0.0082 & -0.0417 & -0.1410 \\ 
    \hline
 \end{tabular}
-\end{table}\begin{table}[ht]
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for survival regressions in Montana.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & BOGR & HECO & PASM & POSE \\ 
+  \hline
+BOGR & -0.1316 & 0.0076 & 0.1096 & 0.0126 \\ 
+  HECO & -0.0028 & -0.7064 & 0.1423 & -0.0030 \\ 
+  PASM & -0.0012 & -0.0033 & -1.4026 & 0.0024 \\ 
+  POSE & 0.0014 & 0.0082 & 0.1363 & -1.2203 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Interaction coefficients for recruitment regressions in Montana.} 
+\begin{tabular}{rrrrr}
+  \hline
+ & BOGR & HECO & PASM & POSE \\ 
+  \hline
+BOGR & -0.9151 & -0.3076 & -0.0984 & 0.0495 \\ 
+  HECO & -0.0785 & -1.4192 & -0.0480 & -0.0468 \\ 
+  PASM & -0.0103 & 0.2773 & -4.3148 & 0.3470 \\ 
+  POSE & 0.1971 & 0.4854 & 0.0623 & -1.7109 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\subsubsection{Interaction Coefficients for Vital Rates in New Mexico}
+
+\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in Arizona.} 
-\begin{tabular}{rrr}
-  \hline
- & BOER & BORO \\ 
-  \hline
-BOER & -0.2791 & -0.0816 \\ 
-  BORO & -0.0200 & -0.6101 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in Idaho.} 
-\begin{tabular}{rrrrr}
-  \hline
- & ARTR & HECO & POSE & PSSP \\ 
-  \hline
-ARTR & -0.0236 & -0.0202 & 0.0314 & 0.0158 \\ 
-  HECO & -0.0005 & -1.0206 & -0.0052 & 0.0125 \\ 
-  POSE & 0.0001 & 0.0006 & -0.9091 & 0.0121 \\ 
-  PSSP & 0.0009 & -0.0013 & 0.0246 & -0.1543 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in Kansas.} 
-\begin{tabular}{rrrr}
-  \hline
- & BOCU & BOHI & SCSC \\ 
-  \hline
-BOCU & -0.1087 & 0.0006 & 0.0010 \\ 
-  BOHI & 0.0021 & -0.0403 & -0.0153 \\ 
-  SCSC & -0.0064 & -0.0066 & -0.0270 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for survival regressions in Montana.} 
-\begin{tabular}{rrrrr}
-  \hline
- & BOGR & HECO & PASM & POSE \\ 
-  \hline
-BOGR & -0.1280 & 0.0211 & 0.1243 & 0.0173 \\ 
-  HECO & -0.0043 & -0.5661 & 0.1386 & 0.0013 \\ 
-  PASM & 0.0001 & -0.0044 & -1.3046 & 0.0012 \\ 
-  POSE & 0.0010 & 0.0085 & 0.1198 & -1.2434 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
+\end{table}
+
+\begin{table}[ht]
 \centering
@@
   \hline
-BOER & -0.1575 & -0.0040 \\ 
-  SPFL & -0.0120 & -0.2807 \\ 
+BOER & -0.1576 & -0.0040 \\ 
+  SPFL & -0.0120 & -0.2806 \\ 
    \hline
 \end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for recruitment regressions in Arizona.} 
-\begin{tabular}{rrr}
-  \hline
- & BOER & BORO \\ 
-  \hline
-BOER & -0.5304 & -0.3063 \\ 
-  BORO & -0.5798 & -2.3237 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for recruitment regressions in Idaho.} 
-\begin{tabular}{rrrrr}
-  \hline
- & ARTR & HECO & POSE & PSSP \\ 
-  \hline
-ARTR & -0.4054 & 0.2132 & 0.0120 & 0.2389 \\ 
-  HECO & -0.5674 & -1.7722 & -0.1476 & -0.3184 \\ 
-  POSE & -0.2499 & -0.0328 & -2.2873 & -0.1448 \\ 
-  PSSP & -0.2580 & -0.1687 & 0.0365 & -1.7675 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for recruitment regressions in Kansas.} 
-\begin{tabular}{rrrr}
-  \hline
- & BOCU & BOHI & SCSC \\ 
-  \hline
-BOCU & -1.1584 & -0.0242 & -0.0524 \\ 
-  BOHI & 0.3387 & -0.8620 & 0.0123 \\ 
-  SCSC & 0.0615 & -0.0308 & -0.9543 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
-\centering
-\caption{Interaction coefficients for recruitment regressions in Montana.} 
-\begin{tabular}{rrrrr}
-  \hline
- & BOGR & HECO & PASM & POSE \\ 
-  \hline
-BOGR & -0.8916 & -0.2522 & -0.1334 & 0.0610 \\ 
-  HECO & 0.0852 & -1.4191 & -0.0685 & -0.0422 \\ 
-  PASM & 0.1252 & 0.3093 & -4.1823 & 0.3032 \\ 
-  POSE & 0.1577 & 0.4906 & 0.0709 & -1.7065 \\ 
-   \hline
-\end{tabular}
-\end{table}\begin{table}[ht]
+\end{table}
+
+\begin{table}[ht]
 \centering

### 2016-07-05 docs/ms_communitySynchrony_PREPRINT.tex
edits peer review disclosure to preprint
updates preprint for bioRxiv

@@
 \pagestyle{fancy}
-\chead{\color{red}THIS MANUSCRIPT HAS NOT UNDERGONE PEER REVIEW}
+\chead{\color{red}\textsf{THIS MANUSCRIPT HAS NOT UNDERGONE PEER REVIEW}}
 \lhead{}
@@
 \begin{abstract}
-Temporal asynchrony among species is an important mechanism through which diversity can stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
-Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific competition, and demographic stochasticity.
+Temporal asynchrony among species \textcolor{blue}{helps}\footnote{\textcolor{blue}{Changes and insertions unique to this version are shown in blue text.}} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \textcolor{blue}{interactions}, and demographic stochasticity.
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-Simulation experiments showed that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
-Removing interspecific competition and demographic stochasticity had small effects on synchrony.
+\textcolor{blue}{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Removing interspecific \textcolor{blue}{interactions} and demographic stochasticity had small effects on synchrony.
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
@@
 2010). The stabilizing effect of species richness stems from individual
-species responding to environmental fluctuations (environmental
-stochasticity), or fluctuating asynchronously because of random chance
-events (demographic stochasticity) (Isbell et al. 2009, Hector et al.
-2010, {{de Mazancourt}} et al. 2013). Species richness affects synchrony
-because larger species pools are more likely to contain species that
-respond disimilarly to environmental conditions (Yachi and Loreau 1999),
+species responding \textcolor{blue}{in different ways} to environmental
+fluctuations (environmental stochasticity), or fluctuating
+asynchronously because of random chance events (demographic
+stochasticity) (Isbell et al. 2009, Hector et al. 2010, {{de
+Mazancourt}} et al. 2013). Species richness affects synchrony because
+larger species pools are more likely to contain species that respond
+disimilarly to environmental conditions (Yachi and Loreau 1999),
 implying that species losses will reduce ecosystem stability. Even
@@
 the Office of Research and Graduate Studies at Utah State University are
-gratefully acknowledged.
-
-\subsection{STATEMENT OF AUTHORSHIP}
-ATT, CdM, and PBA designed the research.
-CdM and ML designed and analyzed the theoretical model.
-ATT and PBA performed the empirical analysis.
-All authors contributed to the interpretation and presentation of results.
-ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
-
-\setlength{\parindent}{0ex} \newpage{}
+gratefully acknowledged. We also thank three anonymous reviewers who
+provided critical feedback that improved the manuscript.
+
+\setlength{\parindent}{0ex}
 
@@
 92:1703.
+
+Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
+Functional Similarity, and Top-Down Control Drive Synchronization and
+the Reliability of Ecosystem Function. The American Naturalist
+183:394--409.
 
@@
 5:524--533.
+
+Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
+Stability of Food Webs in Metacommunities. The American Naturalist
+175:E16--E34.
 
@@
 knowledge and future challenges. Science 294:804--808.
+
+McGill, B. J., R. S. Etienne, J. S. Gray, D. Alonso, M. J. Anderson, H.
+K. Benecha, M. Dornelas, B. J. Enquist, J. L. Green, F. He, A. H.
+Hurlbert, A. E. Magurran, P. A. Marquet, B. A. Maurer, A. Ostling, C. U.
+Soykan, K. I. Ugland, and E. P. White. 2007. Species abundance
+distributions: Moving beyond single prediction theories to integration
+within an ecological framework. Ecology Letters 10:995--1015.
 
@@
 
+Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
+and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
+
 Tilman, D. 1988. Plant strategies and the dynamics and structure of
@@
 
+Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
+stabilize food web dynamics by increasing synchrony. Ecology Letters
+10:1066--1074.
+
 Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beauséjour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf. 2013.
-Global meta-analysis reveals no net change in local-scale plant
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
 biodiversity over time. Proceedings of the National Academy of Sciences
@@
 
+Yenni, G., P. B. Adler, and S. K. {Morgan Ernest}. 2012. Strong
+self-limitation promotes the persistence of rare species. Ecology
+93:456--461.
+
 Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
@@
 
-\subsection{Table Titles}\label{tables}
-TABLE 1: Site descriptions and focal species.
-
-TABLE 2: Observed synchrony among species' per capita growth rates ($\phi_{R}$), observed synchrony among species' percent cover ($\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community.
-Species richness values reflect the number of species analyzed from the community, not the actual richness.
-
-\newpage{}
-
 \subsection{Tables}\label{tables}
+
 \singlespacing
@@
 
-\subsection{Figures Captions}
-\doublespacing
-
-FIGURE 1: Community-wide species synchrony of per capita growth rates from model simulation experiments.
-Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
-+ No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
-+ No E.S.'').
-Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
-Error bars represent the upper and lower 95\% quantiles from model simulations.
-All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
-Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.
-
-FIGURE 2: Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
-Only'') and when environmental stochasticity is also present removed (``D.S.
-+ E.S.'').
-The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
-The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
-Theoretically, ``D.S.
-Only'' simulations should remain constant across landscape size, whereas ``D.S.
-+ E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-Error bars represent the upper and lower 95\% quantiles from model simulations.
-
-FIGURE 3: Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions).
-We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
-The dashed line is the line of equality.
-Simulation results in this figure are analogous to ``No Comp.
-+ No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.
-1, but here we control the time series of random year effects.
-
-\newpage{}
-
-\singlespacing
-
 \subsection{Figures}\label{figures}
@@
   \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
-Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific competition present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific competition removed (``No Comp.''), interspecific competition and demographic stochasticity removed (``No Comp.
-+ No D.S.''), and interspecific competition and environmental stochasticity removed (``No Comp.
+Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific interactions present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific interactions removed (``No Comp.''), interspecific interactions and demographic stochasticity removed (``No Comp.
++ No D.S.''), and interspecific interactions and environmental stochasticity removed (``No Comp.
 + No E.S.'').
 Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
-Error bars represent the upper and lower 95\% quantiles from model simulations.
-All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
-Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.}
+\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.} All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
 \end{figure}
@@
   \centering
-      \includegraphics[width=4in]{./components/formatted_figures/formatted_figure3.png}
-  \caption{Synchrony of species per capita growth rates simulated in large monocultures (IPM without species interactions) versus synchrony in large polycultures (IPM with species interactions).
-We used the same sequence of random year effects for both simulations (monoculture and polyculture) to mimic biodiversity-ecosystem functioning experiments.
+      \includegraphics[width=3in]{./components/formatted_figures/formatted_figure3.png}
+  \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).
+We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.
 The dashed line is the line of equality.
-Simulation results in this figure are analogous to ``No Comp.
-+ No D.S.'' (monocultures) and ``No D.S.'' (polycultures) in Fig.
+Simulation results in this figure are analagous to ``No Comp.
++ No D.S.'' (species in isolation) and ``No D.S.'' (species in competition) in Fig.
 1, but here we control the time series of random year effects.}
@@
 
-\pagebreak{}
-
 \end{document}

### 2016-08-12 docs/ms_communitySynchrony.Rmd
changing to bw figs

--- 2016-06-24 docs/ms_communitySynchrony.Rmd
@@
 
-\textbf{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
-
-\vspace{2em}
-
-\textbf{Authorship}: ATT, CdM, and PBA designed the research.
-CdM and ML designed and analyzed the theoretical model.
-ATT and PBA performed the empirical analysis.
-All authors contributed to the interpretation and presentation of results.
-ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
-
-\textbf{Running Title}: Drivers of species synchrony in natural plant communities
-
-\textbf{Article Type}: Letter
-
-\textbf{Number of Words}: Abstract = 138; Main Text = 4929 (including section titles and in-line citations)
-
-\textbf{Number of References}: 46
-
-\textbf{Number of Tables and Figures}:  2 Tables;  3 Figures
+\textbf{Running Head}: Drivers of species synchrony in natural plant communities
 
@@
 \begin{abstract}
-Temporal asynchrony among species \hl{helps} diversity \hl{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
-Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \hl{interactions}, and demographic stochasticity.
+Temporal asynchrony among species helps diversity to stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific interactions, and demographic stochasticity.
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-\hl{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
-Removing interspecific \hl{interactions} and demographic stochasticity had small effects on synchrony.
+We found that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Removing interspecific interactions and demographic stochasticity had small effects on synchrony.
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+
+\vspace{1em}
+
+\emph{Key words: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability}
+
 \end{abstract}
 
-\newpage{}
 <!-- THIS HOW TO COMMENT-OUT -->
@@
 Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
-The stabilizing effect of species richness stems from individual species responding \hl{in different ways} to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
+The stabilizing effect of species richness stems from individual species responding in different way} to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
 Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
 Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
-\hl{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}
+The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.
 
-Theory identifies three main determinants of species synchrony: environmental stochasticity, demographic stochasticity, and \hl{interspecific interactions} [@Loreau2008a; @Gonzalez2009; @Loreau2013].
-For example, in a community composed of large \hl{populations} (no demographic stochasticity) \hl{with weak interspecific interations}, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
-However, \hl{this} prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
-\hl{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities} [@Mutshinda2009; @Thibaut2012], \hl{and they did not consider demographic stochasiticity.}
+Theory identifies three main determinants of species synchrony: environmental stochasticity, demographic stochasticity, and interspecific interactions [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+For example, in a community composed of large populations (no demographic stochasticity) with weak interspecific interations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
+However, this prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
+Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities [@Mutshinda2009; @Thibaut2012], and they did not consider demographic stochasiticity.
 
@@
 Even beyond grassland studies, whether competition or environmental responses drive compensatory dynamics remains controversial [reviewed in @Gonzalez2009].
-In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from \hl{the covariance matrix of species abundances} [e.g., @Houlahan2007] is impossible.
+In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from the covariance matrix of species abundances [e.g., @Houlahan2007] is impossible.
 This is because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
@@
 
-The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} is to remove them one-by-one, and in combination.
+The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific interactions is to remove them one-by-one, and in combination.
 In principle, this could be done in an extremely controlled laboratory setting, but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
@@
 They used a statistical technique to decompose temporal variation into competition and environmental components, and found that positively correlated environmental responses among species determined community dynamics.
-@Thibaut2012 \hl{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
+@Thibaut2012 used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.
 While a major step forward, Mutshinda et al.'s (2009) and Thibaut et al.'s (2012) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Furthermore, although both studies quantified the relative importance of environmental stochasticity and \hl{interspecific interactions} to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
+Furthermore, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
 
 Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony.
-Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} on species synchrony in natural plant communities.
+Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific interactions on species synchrony in natural plant communities.
 To achieve these objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
 We then confront our theoretical predictions with simulations from the empirically-based population models.
-Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific \hl{interactions} on community-wide species synchrony.
+Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific interactions on community-wide species synchrony.
 Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, and demographic and environmental stochasticity), and that these models successfully reproduce observed community dynamics [@Chu2015], perturbing the models can reveal the processes that determine species synchrony in our focal grassland communities.
@@
 However, we can derive analytical predictions for species synchrony under special limiting cases.
-\hl{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
+The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).
 We focus on synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
@@
 To derive analytical predictions we solved a first-order approximation of Equation 2 [@DeMazancourt2013a and Online Supporting Information].
-\hl{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos} [@Loreau2008a].
-\hl{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small} (e.g., $r < 2$).
+Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos [@Loreau2008a].
+Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small (e.g., $r < 2$).
 
@@
 \noindent{} where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
-Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) in \hl{growing in isolation} (no interspecific interactions) to calculate the covariance of species' environmental responses.
+Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) growing in isolation (no interspecific interactions) to calculate the covariance of species' environmental responses.
 To estimate environmental responses in natural communities, we turn to our population models built using long-term demographic data.
@@
 We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
-\hl{Crowding was treated as a fixed effect without a temporal component because the 95\% credible intervals for random year effects on crowding broadly overlapped zero.}
+Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.
 Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
@@
 #### Simulation experiments
-We performed simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific \hl{interactions}) were removed one-by-one and in combination.
+We performed simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific interactions) were removed one-by-one and in combination.
 To remove interspecific interactions, we set the off-diagonals of the interaction matrix for each vital rate regression to zero.
-This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are \hl{growing in isolation}.
-\hl{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
+This retains intraspecific interactions, and thus density-dependence, and results in simulations where species are growing in isolation.
+We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.
 To remove the effect of a fluctuating environment, we removed the temporal (interannual) random effects from the regression equations.
@@
 We conducted the following six simulation experiments:
-(1) IBM: All drivers (environmental stochasticity, demographic stochasticity, or \hl{interspecific interactions}) present;
+(1) IBM: All drivers (environmental stochasticity, demographic stochasticity, or interspecific interactions) present;
 (2) IPM: Demographic stochasticity removed;
 (3) IBM: Environmental stochasticity removed;
-(4) IBM: \hl{Interspecific interactions} removed;
-(5) IPM: \hl{Interspecific interactions} and demographic stochasticity removed;
-(6) IBM: \hl{Interspecific interactions} and environmental stochasticity removed.
+(4) IBM: Interspecific interactions removed;
+(5) IPM: Interspecific interactions and demographic stochasticity removed;
+(6) IBM: Interspecific interactions and environmental stochasticity removed.
 We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
@@
 Results from our simulation experiments also allow us to test our theoretical predictions.
-First, in the absence of \hl{interspecific interactions} and demographic stochasticity, \hl{populations can only fluctuate in response to the environment}.
+First, in the absence of interspecific interactions and demographic stochasticity, populations can only fluctuate in response to the environment.
 Therefore, we can use results from simulation (5) to estimate the covariance of species' responses to the environment ($\text{cov}(u_{ie}, u_{je})$) and parameterize Equation 4.
@@
 Synchrony is lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
-\hl{Synchrony of per capita growth rates and CV of percent cover are positively correlated} (Pearson’s $\rho$ = 0.72).
+Synchrony of per capita growth rates and CV of percent cover are positively correlated (Pearson’s $\rho$ = 0.72).
 For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
@@
 Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S3).
-Species interactions are weak in these communities [Table S4 and @Chu2015], and removing \hl{interspecific interactions} had little effect on synchrony (Fig. 1; "No Comp." simulations).
-Removing \hl{interspecific interactions} caused, at most, a 5% change in synchrony (Fig. 1).
+Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific interactions had little effect on synchrony (Fig. 1; "No Comp." simulations).
+Removing interspecific interactions caused, at most, a 5% change in synchrony (Fig. 1).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
@@
 Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
-\hl{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
+Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.
 
@@
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
-\hl{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
-However, the theoretical predictions \hl{closely matched results from} simulations of population models fit to long-term data from natural plant communities (Table 3).
-Such \hl{strong agreement} between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony \hl{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.
+For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.
+However, the theoretical predictions closely matched results from simulations of population models fit to long-term data from natural plant communities (Table 3).
+Such strong agreement between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory.
 
@@
 2), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity was not strong enough to \hl{compensate} the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
+Even in small populations, however, demographic stochasticity was not strong enough to compensate the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
 2).
@@
 Only in the Idaho community does synchrony of per capita growth rates approach $\mathcal{M}_D$ in a non-constant environment (Fig. 2).
-This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a \hl{2}5 $\text{m}^2$ quadrat would only contain a few individuals of that species.
+This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 25 $\text{m}^2$ quadrat would only contain a few individuals of that species.
 
@@
 A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
-\hl{Most communities are dominated by few common species and many rare species}, \hl{meaning that the low number of common species in our focal commiunities is not unusual.} 
-\hl{Rather, the generality of our results hinges upon the influence of rare species.
+Most communities are dominated by few common species and many rare species [@McGill2007], meaning that the low number of common species in our focal commiunities is not unusual. 
+Rather, the generality of our results hinges upon the influence of rare species.
 Rare species could be kept at low densities by competitive suppression due to strong interspecific interactions.
 If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
-However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species} [@Yenni2012], \hl{in which case synchrony would remain predominantly driven by environmental responses.}
+However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species [@Yenni2012], in which case synchrony would remain predominantly driven by environmental responses.
 Rare species could also be limited by generalist natural enemies whose effects are density-independent but alter competitive hierarchies [@Mordecai2011].
@@
 We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
-The lack of an effect of \hl{interspecific interactions} on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
+The lack of an effect of interspecific interactions on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
 
@@
 
-\hl{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
-Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap} [@Vasseur2007], \hl{dispersal} [@Gouhier2010], \hl{and the strength of top-down control} [@Bauer2014] \hl{are all likely to affect community synchrony.}
+Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
+Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap [@Vasseur2007], dispersal [@Gouhier2010], and the strength of top-down control [@Bauer2014] are all likely to affect community synchrony.
 

### 2016-08-12 docs/ms_communitySynchrony.tex
changing to bw figs

@@
 \usepackage{longtable,booktabs}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \ifxetex
@@
 
-\textbf{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
-
-\vspace{2em}
-
-\textbf{Authorship}: ATT, CdM, and PBA designed the research.
-CdM and ML designed and analyzed the theoretical model.
-ATT and PBA performed the empirical analysis.
-All authors contributed to the interpretation and presentation of results.
-ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.
-
-\textbf{Running Title}: Drivers of species synchrony in natural plant communities
-
-\textbf{Article Type}: Letter
-
-\textbf{Number of Words}: Abstract = 136; Main Text = 4992
-
-\textbf{Number of References}: 46
-
-\textbf{Number of Tables and Figures}:  2 Tables;  3 Figures
+\textbf{Running Head}: Drivers of species synchrony in natural plant communities
 
@@
 \begin{abstract}
-Temporal asynchrony among species \textcolor{blue}{helps} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
-Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \textcolor{blue}{interactions}, and demographic stochasticity.
+Temporal asynchrony among species helps diversity to stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific interactions, and demographic stochasticity.
 We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-\textcolor{blue}{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
-Removing interspecific \textcolor{blue}{interactions} and demographic stochasticity had small effects on synchrony.
+We found that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Removing interspecific interactions and demographic stochasticity had small effects on synchrony.
 In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+
+\vspace{1em}
+
+\emph{Key words: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability}
+
 \end{abstract}
 
-\newpage{} \setlength{\parindent}{5ex}
+\setlength{\parindent}{5ex}
 
@@
 2010). The stabilizing effect of species richness stems from individual
-species responding \textcolor{blue}{in different ways} to environmental
-fluctuations (environmental stochasticity), or fluctuating
-asynchronously because of random chance events (demographic
-stochasticity) (Isbell et al. 2009, Hector et al. 2010, {{de
-Mazancourt}} et al. 2013). Species richness affects synchrony because
-larger species pools are more likely to contain species that respond
-disimilarly to environmental conditions (Yachi and Loreau 1999),
-implying that species losses will reduce ecosystem stability. Even
-without species losses, abiotic homogenization can weaken compensatory
-dynamics and, in turn, decrease temporal stability of ecosystem
-functioning (Hautier et al. 2014).
-\textcolor{blue}{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}
+species responding in different way\} to environmental fluctuations
+(environmental stochasticity), or fluctuating asynchronously because of
+random chance events (demographic stochasticity) (Isbell et al. 2009,
+Hector et al. 2010, {{de Mazancourt}} et al. 2013). Species richness
+affects synchrony because larger species pools are more likely to
+contain species that respond disimilarly to environmental conditions
+(Yachi and Loreau 1999), implying that species losses will reduce
+ecosystem stability. Even without species losses, abiotic homogenization
+can weaken compensatory dynamics and, in turn, decrease temporal
+stability of ecosystem functioning (Hautier et al. 2014). The link
+between synchrony and stability means that a mechanistic understanding
+of synchrony can help us predict the impacts of global change on
+ecosystem stability.
 
@@
 environmental stochasticity, demographic stochasticity, and
-\textcolor{blue}{interspecific interactions} (Loreau and {{de
-Mazancourt}} 2008, 2013, Gonzalez and Loreau 2009). For example, in a
-community composed of large \textcolor{blue}{populations} (no
-demographic stochasticity)
-\textcolor{blue}{with weak interspecific interations}, community-wide
-species synchrony should be determined by the covariance of species'
-responses to the environment (Loreau and {{de Mazancourt}} 2008).
-However, \textcolor{blue}{this} prediction relies on a relatively simple
-population model and only holds under two assumptions: (i) species'
-responses to the environment are similar in magnitude and (ii) all
-species have similar growth rates.
-\textcolor{blue}{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities}
-(Mutshinda et al. 2009, Thibaut et al. 2012),
-\textcolor{blue}{and they did not consider demographic stochasiticity.}
+interspecific interactions (Loreau and {{de Mazancourt}} 2008, 2013,
+Gonzalez and Loreau 2009). For example, in a community composed of large
+populations (no demographic stochasticity) with weak interspecific
+interations, community-wide species synchrony should be determined by
+the covariance of species' responses to the environment (Loreau and {{de
+Mazancourt}} 2008). However, this prediction relies on a relatively
+simple population model and only holds under two assumptions: (i)
+species' responses to the environment are similar in magnitude and (ii)
+all species have similar growth rates. Whether such theoretical
+predictions hold in natural communities where species differences are
+unlikely to be symmetrical is unkown because few studies have explicitly
+tested theory on the drivers of species synchrony in natural communities
+(Mutshinda et al. 2009, Thibaut et al. 2012), and they did not consider
+demographic stochasiticity.
 
@@
 controversy remains because quantifying the relative strengths of each
-driver on the degree of synchrony from
-\textcolor{blue}{the covariance matrix of species abundances} (e.g.,
-Houlahan et al. 2007) is impossible. This is because an unbiased null
-expectation for synchrony does not exist (Loreau and {{de Mazancourt}}
-2008) and observed synchrony can arise from non-unique combinations of
-factors (Ranta et al. 2008). For example, weak synchrony of population
-abundances could reflect positive environmental correlations
-(synchronizing effect) offset by strong competition (desynchronizing
-effect), or negative environmental correlations and weak competition.
+driver on the degree of synchrony from the covariance matrix of species
+abundances (e.g., Houlahan et al. 2007) is impossible. This is because
+an unbiased null expectation for synchrony does not exist (Loreau and
+{{de Mazancourt}} 2008) and observed synchrony can arise from non-unique
+combinations of factors (Ranta et al. 2008). For example, weak synchrony
+of population abundances could reflect positive environmental
+correlations (synchronizing effect) offset by strong competition
+(desynchronizing effect), or negative environmental correlations and
+weak competition.
 
 The best way to quantify the effects of environmental stochasticity,
-demographic stochasticity, and interspecific
-\textcolor{blue}{interactions} is to remove them one-by-one, and in
-combination. In principle, this could be done in an extremely controlled
-laboratory setting, but empirically-based models of interacting
-populations, fit with data sets from natural communities, offer a
-practical alternative. For example, Mutshinda et al. (2009) fit a
-dynamic population model to several community time series of insect and
-bird abundances. They used a statistical technique to decompose temporal
-variation into competition and environmental components, and found that
-positively correlated environmental responses among species determined
-community dynamics. Thibaut et al. (2012)
-\textcolor{blue}{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
-While a major step forward, Mutshinda et al.'s (2009) and Thibaut et
-al.'s (2012) modeling technique relied on abundance data that may or may
-not reliably capture competitive interactions that occur at the
-individual level. Furthermore, although both studies quantified the
-relative importance of environmental stochasticity and
-\textcolor{blue}{interspecific interactions} to explain the observed
+demographic stochasticity, and interspecific interactions is to remove
+them one-by-one, and in combination. In principle, this could be done in
+an extremely controlled laboratory setting, but empirically-based models
+of interacting populations, fit with data sets from natural communities,
+offer a practical alternative. For example, Mutshinda et al. (2009) fit
+a dynamic population model to several community time series of insect
+and bird abundances. They used a statistical technique to decompose
+temporal variation into competition and environmental components, and
+found that positively correlated environmental responses among species
+determined community dynamics. Thibaut et al. (2012) used a similar
+approach for reef fish and came to a similar conclusion: environmental
+responses determine synchrony. While a major step forward, Mutshinda et
+al.'s (2009) and Thibaut et al.'s (2012) modeling technique relied on
+abundance data that may or may not reliably capture competitive
+interactions that occur at the individual level. Furthermore, although
+both studies quantified the relative importance of environmental
+stochasticity and interspecific interactions to explain the observed
 variation of species synchrony, they did not use the model to quantify
@@
 relative influence of environmental stochasticity, demographic
-stochasticity, and interspecific \textcolor{blue}{interactions} on
-species synchrony in natural plant communities. To achieve these
-objectives, we first refine theory that has been used to predict the
-effects of species richness on ecosystem stability ({{de Mazancourt}} et
-al. 2013) and species synchrony (Loreau and {{de Mazancourt}} 2008) to
-generate predictions of community-wide species synchrony under two
-limiting cases derived from the dynamics of individual species in
-monoculture. We then confront our theoretical predictions with
-simulations from the empirically-based population models. Second, we use
-the multi-species population models to perform simulation experiments
-that isolate the effects of environmental stochasticity, demographic
-stochasticity, and interspecific \textcolor{blue}{interactions} on
-community-wide species synchrony. Given that our population models
+stochasticity, and interspecific interactions on species synchrony in
+natural plant communities. To achieve these objectives, we first refine
+theory that has been used to predict the effects of species richness on
+ecosystem stability ({{de Mazancourt}} et al. 2013) and species
+synchrony (Loreau and {{de Mazancourt}} 2008) to generate predictions of
+community-wide species synchrony under two limiting cases derived from
+the dynamics of individual species in monoculture. We then confront our
+theoretical predictions with simulations from the empirically-based
+population models. Second, we use the multi-species population models to
+perform simulation experiments that isolate the effects of environmental
+stochasticity, demographic stochasticity, and interspecific interactions
+on community-wide species synchrony. Given that our population models
 capture the essential features of community dynamics important to
@@
 derive analytical predictions for species synchrony under special
-limiting cases.
-\textcolor{blue}{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
-We focus on synchrony of per capita growth rates, rather than
-abundances, because growth rates represent the instantaneous response of
-species to the environment and competition, and are less susceptible to
-the legacy effects of drift and disturbance (Loreau and {{de
-Mazancourt}} 2008). We present equivalent results for synchrony of
-species abundances in the Online Supporting Information, and show that
-they lead to the same overall conclusions as synchrony of per capita
-growth rates. Following Loreau and {{de Mazancourt}} (2008) and {{de
-Mazancourt}} et al. (2013), we define population growth, ignoring
-observation error, as
+limiting cases. The limiting case predictions we derive serve as
+baselines to help us interpret results from empirically-based
+simulations (described below). We focus on synchrony of per capita
+growth rates, rather than abundances, because growth rates represent the
+instantaneous response of species to the environment and competition,
+and are less susceptible to the legacy effects of drift and disturbance
+(Loreau and {{de Mazancourt}} 2008). We present equivalent results for
+synchrony of species abundances in the Online Supporting Information,
+and show that they lead to the same overall conclusions as synchrony of
+per capita growth rates. Following Loreau and {{de Mazancourt}} (2008)
+and {{de Mazancourt}} et al. (2013), we define population growth,
+ignoring observation error, as
 
@@
 \emph{t}, and \(r_i(t)\) is its population growth rate in year \emph{t}.
-\(r_{mi}\) is species \emph{i}'s intrinsic rate of increase, \(K_i\) is
+\(r_{mi}\) is species \emph{i}`s intrinsic rate of increase, \(K_i\) is
 its carrying capacity, and \(\alpha_{ij}\) is the interspecific
@@
 analytical predictions we solved a first-order approximation of Equation
-2 ({{de Mazancourt}} et al. 2013 and Online Supporting Information).
-\textcolor{blue}{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos}
-(Loreau and {{de Mazancourt}} 2008).
-\textcolor{blue}{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small}
-(e.g., \(r < 2\)).
+2 ({{de Mazancourt}} et al. 2013 and Online Supporting Information). Due
+to the linear approximation approach, our analytical predictions will
+likely fail in communities where species exhibit large fluctuations due
+to limit cycles and chaos (Loreau and {{de Mazancourt}} 2008). Indeed,
+one of the advantages of focusing on growth rates rather than abundances
+is that growth rates are more likely to be well-regulated around an
+equilibrium value, if the long-term average of a species' growth rate is
+relatively small (e.g., \(r < 2\)).
 
@@
 represents a community where dynamics are driven by demographic
-stochasticity alone. Our prediction
-\textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{D}$, $\phi_{R,\mathcal{M}_{D}}$, is}
+stochasticity alone. Our prediction for the synchrony of per capita
+growth rates for \(\mathcal{M}_{D}\), \(\phi_{R,\mathcal{M}_{D}}\), is
 
@@
 intrinsic growth rates are identical, and environmental variance is
-identical for all species. Our prediction
-\textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{E}$, $\phi_{R,\mathcal{M}_{E}}$, is }
+identical for all species. Our prediction for the synchrony of per
+capita growth rates for \(\mathcal{M}_{E}\),
+\(\phi_{R,\mathcal{M}_{E}}\), is
 
@@
 requires estimates of species dynamics of large populations (no
-demographic stochasticity) \textcolor{blue}{growing in isolation} (no
-interspecific interactions) to calculate the covariance of species'
-environmental responses. To estimate environmental responses in natural
-communities, we turn to our population models built using long-term
-demographic data.
+demographic stochasticity) growing in isolation (no interspecific
+interactions) to calculate the covariance of species' environmental
+responses. To estimate environmental responses in natural communities,
+we turn to our population models built using long-term demographic data.
 
@@
 via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
-2013).
-\textcolor{blue}{Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.}
-Spatial (quadrat groupings) variation was treated as a random effect on
-the intercept and temporal (interannual) variation was treated as random
-effects on the intercept and the effect of genet size in the previous
-year (Online Supporting Information).
+2013). Crowding was treated as a fixed effect without a temporal
+component because most 95\% credible intervals for random year effects
+on crowding broadly overlapped zero and, in a test case, including
+yearly crowding effects did not change our results. Spatial (quadrat
+groupings) variation was treated as a random effect on the intercept and
+temporal (interannual) variation was treated as random effects on the
+intercept and the effect of genet size in the previous year (Online
+Supporting Information).
 
@@
 We performed simulation experiments where drivers (environmental
-stochasticity, demographic stochasticity, or interspecific
-\textcolor{blue}{interactions}) were removed one-by-one and in
-combination. To remove interspecific interactions, we set the
-off-diagonals of the interaction matrix for each vital rate regression
-to zero. This retains intraspecific interactions, and thus
-density-dependence, and results in simulations where species are
-\textcolor{blue}{growing in isolation}.
-\textcolor{blue}{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
-To remove the effect of a fluctuating environment, we removed the
-temporal (interannual) random effects from the regression equations. To
-remove the effect of demographic stochasticity, we use the IPM rather
-than the IBM because the IPM does not include demographic stochasticity
+stochasticity, demographic stochasticity, or interspecific interactions)
+were removed one-by-one and in combination. To remove interspecific
+interactions, we set the off-diagonals of the interaction matrix for
+each vital rate regression to zero. This retains intraspecific
+interactions, and thus density-dependence, and results in simulations
+where species are growing in isolation. We cannot definitively rule out
+the effects of species interactions on all parameters, meaning that a
+true monoculture could behave differently than our simulations of a
+population growing without interspecific competitors. To remove the
+effect of a fluctuating environment, we removed the temporal
+(interannual) random effects from the regression equations. To remove
+the effect of demographic stochasticity, we use the IPM rather than the
+IBM because the IPM does not include demographic stochasticity
 (demographic stochasticity cannot be removed from the IBM). Since the
@@
 drivers (environmental stochasticity, demographic stochasticity, or
-\textcolor{blue}{interspecific interactions}) present; (2) IPM:
-Demographic stochasticity removed; (3) IBM: Environmental stochasticity
-removed; (4) IBM: \textcolor{blue}{Interspecific interactions} removed;
-(5) IPM: \textcolor{blue}{Interspecific interactions} and demographic
-stochasticity removed; (6) IBM:
-\textcolor{blue}{Interspecific interactions} and environmental
-stochasticity removed. We ran IPM simulations for 2,000 time steps,
-after an initial 500 iteration burn-in period. This allowed species time
-to reach their stable size distribution. We then calculated the
-synchrony of species' per capita growth rates over 100 randomly selected
-contiguous 50 time-step sections. We ran IBM simulations for 100 time
-steps, and repeated the simulations 100 times for each simulation
-experiment. From those, we retained only the simulations in which no
-species went extinct due to demographic stochasticity. Synchrony was
-calculated over the 100 time steps for each no extinction run within a
-model experiment. To explore the effect of demographic stochasticity in
-different sized populations, we ran simulations (1) and (6) on plot
-sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM simulations
-were run on a 25 \(\text{m}^2\) landscape.
+interspecific interactions) present; (2) IPM: Demographic stochasticity
+removed; (3) IBM: Environmental stochasticity removed; (4) IBM:
+Interspecific interactions removed; (5) IPM: Interspecific interactions
+and demographic stochasticity removed; (6) IBM: Interspecific
+interactions and environmental stochasticity removed. We ran IPM
+simulations for 2,000 time steps, after an initial 500 iteration burn-in
+period. This allowed species time to reach their stable size
+distribution. We then calculated the synchrony of species' per capita
+growth rates over 100 randomly selected contiguous 50 time-step
+sections. We ran IBM simulations for 100 time steps, and repeated the
+simulations 100 times for each simulation experiment. From those, we
+retained only the simulations in which no species went extinct due to
+demographic stochasticity. Synchrony was calculated over the 100 time
+steps for each no extinction run within a model experiment. To explore
+the effect of demographic stochasticity in different sized populations,
+we ran simulations (1) and (6) on plot sizes of 1, 4, 9, 16, and 25
+\(\text{m}^2\). All other IBM simulations were run on a 25
+\(\text{m}^2\) landscape.
 
 Results from our simulation experiments also allow us to test our
-theoretical predictions. First, in the absence of
-\textcolor{blue}{interspecific interactions} and demographic
-stochasticity,
-\textcolor{blue}{populations can only fluctuate in response to the environment}.
-Therefore, we can use results from simulation (5) to estimate the
-covariance of species' responses to the environment
-(\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation 4.
-Parameterizing Equation 3 does not require simulation output because the
-only parameters are the species' relative abundances. Second,
-simulations (5) and (6) represent the simulated version of our limiting
-case theoretical predictions. Thus, we directly test the theoretical
-predictions by comparing them to observed synchrony and simulated
-synchrony.
+theoretical predictions. First, in the absence of interspecific
+interactions and demographic stochasticity, populations can only
+fluctuate in response to the environment. Therefore, we can use results
+from simulation (5) to estimate the covariance of species' responses to
+the environment (\(\text{cov}(u_{ie}, u_{je})\)) and parameterize
+Equation 4. Parameterizing Equation 3 does not require simulation output
+because the only parameters are the species' relative abundances.
+Second, simulations (5) and (6) represent the simulated version of our
+limiting case theoretical predictions. Thus, we directly test the
+theoretical predictions by comparing them to observed synchrony and
+simulated synchrony.
 
@@
 Idaho, the only data set that includes two life forms: a shrub and three
-perennial grasses (Table 1).
-\textcolor{blue}{Synchrony of per capita growth rates and CV of percent cover are positively correlated}
-(Pearson's \(\rho\) = 0.72). For all five communities, species synchrony
-from IPM and IBM simulations closely approximated observed synchrony
-(Fig. S1). IBM-simulated synchrony is consistently, but only slightly,
-lower than IPM-simulated synchrony (Fig. S1), likely due to the
-desynchronizing effect of demographic stochasticity.
+perennial grasses (Table 1). Synchrony of per capita growth rates and CV
+of percent cover are positively correlated (Pearson's \(\rho\) = 0.72).
+For all five communities, species synchrony from IPM and IBM simulations
+closely approximated observed synchrony (Fig. S1). IBM-simulated
+synchrony is consistently, but only slightly, lower than IPM-simulated
+synchrony (Fig. S1), likely due to the desynchronizing effect of
+demographic stochasticity.
 
@@
 Species interactions are weak in these communities (Table S4 and Chu and
-Adler 2015), and removing \textcolor{blue}{interspecific interactions}
-had little effect on synchrony (Fig. 1; ``No Comp.'' simulations).
-Removing \textcolor{blue}{interspecific interactions} caused, at most, a
-5\% change in synchrony (Fig. 1). Removing demographic stochasticity
-(``No D.S.'' simulations) caused synchrony to increase slightly in all
-communities (Fig. 1), with an average 6\% increase over synchrony from
-IBM simulations on a 25\(\text{m}^2\) area.
+Adler 2015), and removing interspecific interactions had little effect
+on synchrony (Fig. 1; ``No Comp.'' simulations). Removing interspecific
+interactions caused, at most, a 5\% change in synchrony (Fig. 1).
+Removing demographic stochasticity (``No D.S.'' simulations) caused
+synchrony to increase slightly in all communities (Fig. 1), with an
+average 6\% increase over synchrony from IBM simulations on a
+25\(\text{m}^2\) area.
 
@@
 species synchrony from abundance time series, as expected by theory
-(Loreau and {{de Mazancourt}} 2008).
-\textcolor{blue}{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
+(Loreau and {{de Mazancourt}} 2008). Observed synchrony of per capita
+growth rates was positively correlated with the variability of percent
+cover across our focal communities, which confirms that we are
+investigating an important process underlying ecosystem stability.
 
@@
 population dynamics and required several simplifying assumptions,
-raising questions about their relevance to natural communities.
-\textcolor{blue}{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
-However, the theoretical predictions
-\textcolor{blue}{closely matched results from} simulations of population
-models fit to long-term data from natural plant communities (Table 3).
-Such \textcolor{blue}{strong agreement} between our analytical
-predictions and the simulation results should inspire confidence in the
-ability of simple models to inform our understanding of species
-synchrony
-\textcolor{blue}{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.
+raising questions about their relevance to natural communities. For
+example, the species in our communities do not have equivalent
+environmental and demographic variances (Figs. S4-S7), as required by
+our predictions. However, the theoretical predictions closely matched
+results from simulations of population models fit to long-term data from
+natural plant communities (Table 3). Such strong agreement between our
+analytical predictions and the simulation results should inspire
+confidence in the ability of simple models to inform our understanding
+of species synchrony even in complex natural communities, and allows us
+to place our simulation results within the context of contemporary
+theory.
 
@@
 small populations, however, demographic stochasticity was not strong
-enough to \textcolor{blue}{compensate} the synchronizing effects of
-environmental fluctuations and match the analytical prediction where
-only demographic stochasticity is operating (\(\mathcal{M}_D\) in Fig.
-2). These results confirm the theoretical argument by Loreau and {{de
-Mazancourt}} (2008) that independent fluctuations among interacting
-species in a non-constant environment should be rare. Only in the Idaho
-community does synchrony of per capita growth rates approach
-\(\mathcal{M}_D\) in a non-constant environment (Fig. 2). This is most
-likely due to the strong effect of demographic stochasticity on the
-shrub \emph{Artemisia tripartita} since even a \textcolor{blue}{2}5
-\(\text{m}^2\) quadrat would only contain a few individuals of that
-species.
+enough to compensate the synchronizing effects of environmental
+fluctuations and match the analytical prediction where only demographic
+stochasticity is operating (\(\mathcal{M}_D\) in Fig. 2). These results
+confirm the theoretical argument by Loreau and {{de Mazancourt}} (2008)
+that independent fluctuations among interacting species in a
+non-constant environment should be rare. Only in the Idaho community
+does synchrony of per capita growth rates approach \(\mathcal{M}_D\) in
+a non-constant environment (Fig. 2). This is most likely due to the
+strong effect of demographic stochasticity on the shrub \emph{Artemisia
+tripartita} since even a 25 \(\text{m}^2\) quadrat would only contain a
+few individuals of that species.
 
@@
 A challenge to the generality of our results is that we were only able
-to model common, co-occuring species (see Chu and Adler 2015).
-\textcolor{blue}{Most communities are dominated by few common species and many rare species}
-(McGill et al. 2007),
-\textcolor{blue}{meaning that the low number of common species in our focal commiunities is not unusual.}
-\textcolor{blue}{Rather, the generality of our results hinges upon the influence of rare species.
-Rare species could be kept at low densities by competitive suppression due to strong interspecific interactions.
-If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
-However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species}
-(Yenni et al. 2012),
-\textcolor{blue}{in which case synchrony would remain predominantly driven by environmental responses.}
-Rare species could also be limited by generalist natural enemies whose
-effects are density-independent but alter competitive hierarchies
-(Mordecai 2011). Under such conditions, synchrony will depend on the
-degree of pathogen-induced fitness differences and the pathogen's
-response to environmental conditions. Neither our model nor current
-theory explicitly considers the effects of pathogens on species
+to model common, co-occuring species (see Chu and Adler 2015). Most
+communities are dominated by few common species and many rare species
+(McGill et al. 2007), meaning that the low number of common species in
+our focal commiunities is not unusual. Rather, the generality of our
+results hinges upon the influence of rare species. Rare species could be
+kept at low densities by competitive suppression due to strong
+interspecific interactions. If we had been able to model rare species,
+we might have found a larger effect of competition on community
+synchrony. However, recent theory predicts that persistent rare species
+may actually experience even weaker interspecific competition, and thus
+stronger niche differences, than common species (Yenni et al. 2012), in
+which case synchrony would remain predominantly driven by environmental
+responses. Rare species could also be limited by generalist natural
+enemies whose effects are density-independent but alter competitive
+hierarchies (Mordecai 2011). Under such conditions, synchrony will
+depend on the degree of pathogen-induced fitness differences and the
+pathogen's response to environmental conditions. Neither our model nor
+current theory explicitly considers the effects of pathogens on species
 synchrony, and this highlights the need for theoretical work on the
@@
 between simulations with and without interspecific interactions (Fig.
-2). The lack of an effect of
-\textcolor{blue}{interspecific interactions} on synchrony is in contrast
-to a large body of theoretical work that predicts a strong role for
-competition in creating compensatory dynamics (Tilman 1988) and a recent
-empirical analysis (Gross et al. 2014).
+2). The lack of an effect of interspecific interactions on synchrony is
+in contrast to a large body of theoretical work that predicts a strong
+role for competition in creating compensatory dynamics (Tilman 1988) and
+a recent empirical analysis (Gross et al. 2014).
 
@@
 
-\textcolor{blue}{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
-Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap}
-(Vasseur and Fox 2007), \textcolor{blue}{dispersal} (Gouhier et al.
-2010), \textcolor{blue}{and the strength of top-down control} (Bauer et
-al. 2014)
-\textcolor{blue}{are all likely to affect community synchrony.}
+Our conclusion that species interactions have little impact on synchrony
+only applies to single trophic level communities. Species interactions
+almost certainly play a strong role in multi-trophic communities where
+factors such as resource overlap (Vasseur and Fox 2007), dispersal
+(Gouhier et al. 2010), and the strength of top-down control (Bauer et
+al. 2014) are all likely to affect community synchrony.
 
@@
 
-\setlength{\parindent}{0ex}
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
-Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
-perennial plants: An embarrassment of niches. Ecology Letters
-13:1019--1029.
-
-Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
-mapped quadrats from Kansas prairie: demographic information for
-herbaceous plants. Ecology 88:2673.
-
-Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
-of semi-desert grassland plants in permanent quadrats mapped from 1915
-to 1947. Ecology 93:1492--1492.
-
-Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
-mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
-92:1703.
-
-Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
-Functional Similarity, and Top-Down Control Drive Synchronization and
-the Reliability of Ecosystem Function. The American Naturalist
-183:394--409.
-
-Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
-P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
-Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
-and its impact on humanity. Nature 489:59--67.
-
-Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
-recruitment stage to stabilize grassland coexistence. Ecological
-Monographs 85:373--392.
-
-{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
-Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
-Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
-Loreau. 2013. Predicting ecosystem stability from community composition
-and biodiversity. Ecology Letters 16:617--625.
-
-Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
-Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
-biodiversity change but not systematic loss. Science 344:296--9.
-
-Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
-Compensatory Dynamics in Ecological Communities. Annual Review of
-Ecology, Evolution, and Systematics 40:393--414.
-
-Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
-variability in space and time. Methods in Ecology and Evolution
-5:524--533.
-
-Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
-Stability of Food Webs in Metacommunities. The American Naturalist
-175:E16--E34.
-
-Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
-Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
-temporal stability of biomass production: a new analysis of recent
-biodiversity experiments. The American naturalist 183:1--12.
-
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
-
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
-
-Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
-on ecosystem functioning: a consensus of current knowledge. Ecological
-Monographs 75:3--35.
-
-Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
-Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
-Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
-Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
-Compensatory dynamics are rare in natural ecological communities.
-Proceedings of the National Academy of Sciences of the United States of
-America 104:3273--3277.
-
-Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
-productivity and the temporal stability of productivity: Patterns and
-processes. Ecology Letters 12:443--451.
-
-Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
-species interaction strength in assembled theoretical competition
-communities. Ecology Letters 2:70--74.
-
-Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
-dynamics in ecology and conservation.
-
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
-
-Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
-for a New Ecological Synthesis.
-
-Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
-drivers: neutral and nonneutral community dynamics in fluctuating
-environments. The American Naturalist 172:E48--E66.
-
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
-
-Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
-Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
-and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
-knowledge and future challenges. Science 294:804--808.
-
-McGill, B. J., R. S. Etienne, J. S. Gray, D. Alonso, M. J. Anderson, H.
-K. Benecha, M. Dornelas, B. J. Enquist, J. L. Green, F. He, A. H.
-Hurlbert, A. E. Magurran, P. A. Marquet, B. A. Maurer, A. Ostling, C. U.
-Soykan, K. I. Ugland, and E. P. White. 2007. Species abundance
-distributions: Moving beyond single prediction theories to integration
-within an ecological framework. Ecology Letters 10:995--1015.
-
-Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
-theory, concepts, and empirical work. Ecological Monographs 81:429--441.
-
-Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
-community dynamics? Proceedings of the Royal Society B: Biological
-Sciences 276:2923--2929.
-
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
-
-R Core Team. 2013. R: A language and environment for statistical
-computing.
-
-Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
-O'Hara. 2008. Detecting compensatory dynamics in competitive communities
-under environmental forcing. Oikos 117:1907--1911.
-
-Rees, M., and S. P. Ellner. 2009. Integral projection models for
-populations in temporally varying environments. Ecological Monographs
-79:575--594.
-
-Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
-projection models: A user's guide.
-
-Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
-species have similar dynamics. Journal of Ecology 99:1453--1459.
-
-Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
-A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
-Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
-C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
-Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
-Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
-Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
-Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
-Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
-Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
-Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
-operating space for humanity. Nature 461:472--475.
-
-Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
-for Latent Gaussian Models Using Integrated Nested Laplace
-Approximations. Journal of the Royal Statistical Society, Series B
-71:319--392.
-
-Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
-Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
-analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
-
-Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
-189:732--735.
-
-Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
-and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
-
-Tilman, D. 1988. Plant strategies and the dynamics and structure of
-plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
-
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
-
-Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
-2001. Diversity and productivity in a long-term grassland experiment.
-Science 294:843--845.
-
-Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
-Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
-W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
-2014. Synchronous dynamics of zooplankton competitors prevail in
-temperate lake ecosystems. Proceedings. Biological sciences / The Royal
-Society 281:20140633.
-
-Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
-stabilize food web dynamics by increasing synchrony. Ecology Letters
-10:1066--1074.
-
-Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
-2013. Global meta-analysis reveals no net change in local-scale plant
-biodiversity over time. Proceedings of the National Academy of Sciences
-of the United States of America 110:19456--9.
-
-Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
-Human Domination of Earth' s Ecosystems. Science 277:494--499.
-
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
-
-Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
-
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
-
-Yenni, G., P. B. Adler, and S. K. {Morgan Ernest}. 2012. Strong
-self-limitation promotes the persistence of rare species. Ecology
-93:456--461.
-
-Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
-sagebrush steppe: long-term data for analyzing demographic rates and
-plant--plant interactions. Ecology 91:3427.
-
-\newpage{}
+\setlength{\parindent}{0ex} \newpage{}
 
@@
 Species richness values reflect the number of species analyzed from the community, not the actual richness.}
-{\normalsize
+\begingroup\normalsize
 \begin{tabular}{lrrrr}
@@
 \end{tabular}
-}
+\endgroup
 \end{table}
@@
 Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
-\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.} All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.
+All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
 Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
@@
 + E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}}
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
 \end{figure}
@@
 
+\pagebreak{}
+
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
+perennial plants: An embarrassment of niches. Ecology Letters
+13:1019--1029.
+
+Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
+mapped quadrats from Kansas prairie: demographic information for
+herbaceous plants. Ecology 88:2673.
+
+Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
+of semi-desert grassland plants in permanent quadrats mapped from 1915
+to 1947. Ecology 93:1492--1492.
+
+Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
+mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
+92:1703.
+
+Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
+Functional Similarity, and Top-Down Control Drive Synchronization and
+the Reliability of Ecosystem Function. The American Naturalist
+183:394--409.
+
+Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
+P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
+Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
+and its impact on humanity. Nature 489:59--67.
+
+Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
+recruitment stage to stabilize grassland coexistence. Ecological
+Monographs 85:373--392.
+
+{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
+Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
+Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
+Loreau. 2013. Predicting ecosystem stability from community composition
+and biodiversity. Ecology Letters 16:617--625.
+
+Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
+Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
+biodiversity change but not systematic loss. Science 344:296--9.
+
+Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. Annual Review of
+Ecology, Evolution, and Systematics 40:393--414.
+
+Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
+variability in space and time. Methods in Ecology and Evolution
+5:524--533.
+
+Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
+Stability of Food Webs in Metacommunities. The American Naturalist
+175:E16--E34.
+
+Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
+Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
+temporal stability of biomass production: a new analysis of recent
+biodiversity experiments. The American naturalist 183:1--12.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
+on ecosystem functioning: a consensus of current knowledge. Ecological
+Monographs 75:3--35.
+
+Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
+Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
+Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
+Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
+Compensatory dynamics are rare in natural ecological communities.
+Proceedings of the National Academy of Sciences of the United States of
+America 104:3273--3277.
+
+Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
+productivity and the temporal stability of productivity: Patterns and
+processes. Ecology Letters 12:443--451.
+
+Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
+species interaction strength in assembled theoretical competition
+communities. Ecology Letters 2:70--74.
+
+Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
+dynamics in ecology and conservation.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
+for a New Ecological Synthesis.
+
+Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
+drivers: neutral and nonneutral community dynamics in fluctuating
+environments. The American Naturalist 172:E48--E66.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
+Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
+and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
+knowledge and future challenges. Science 294:804--808.
+
+McGill, B. J., R. S. Etienne, J. S. Gray, D. Alonso, M. J. Anderson, H.
+K. Benecha, M. Dornelas, B. J. Enquist, J. L. Green, F. He, A. H.
+Hurlbert, A. E. Magurran, P. A. Marquet, B. A. Maurer, A. Ostling, C. U.
+Soykan, K. I. Ugland, and E. P. White. 2007. Species abundance
+distributions: Moving beyond single prediction theories to integration
+within an ecological framework. Ecology Letters 10:995--1015.
+
+Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
+theory, concepts, and empirical work. Ecological Monographs 81:429--441.
+
+Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
+community dynamics? Proceedings of the Royal Society B: Biological
+Sciences 276:2923--2929.
+
+Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
+Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
+3rd international workshop on distributed statistical computing (dSC
+2003). march.
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+
+R Core Team. 2013. R: A language and environment for statistical
+computing.
+
+Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
+O'Hara. 2008. Detecting compensatory dynamics in competitive communities
+under environmental forcing. Oikos 117:1907--1911.
+
+Rees, M., and S. P. Ellner. 2009. Integral projection models for
+populations in temporally varying environments. Ecological Monographs
+79:575--594.
+
+Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
+projection models: A user's guide.
+
+Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
+species have similar dynamics. Journal of Ecology 99:1453--1459.
+
+Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
+A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
+Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
+C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
+Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
+Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
+Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
+Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
+Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
+Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
+Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
+operating space for humanity. Nature 461:472--475.
+
+Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
+for Latent Gaussian Models Using Integrated Nested Laplace
+Approximations. Journal of the Royal Statistical Society, Series B
+71:319--392.
+
+Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
+Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
+analysis of latent Gaussian models using Integrated Nested Laplace
+Approximaxion.
+
+Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
+189:732--735.
+
+Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
+and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
+
+Tilman, D. 1988. Plant strategies and the dynamics and structure of
+plant communities. Plant strategies and the dynamics and structure of
+plant communities.:360 pp.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
+2001. Diversity and productivity in a long-term grassland experiment.
+Science 294:843--845.
+
+Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
+Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
+W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
+2014. Synchronous dynamics of zooplankton competitors prevail in
+temperate lake ecosystems. Proceedings. Biological sciences / The Royal
+Society 281:20140633.
+
+Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
+stabilize food web dynamics by increasing synchrony. Ecology Letters
+10:1066--1074.
+
+Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
+biodiversity over time. Proceedings of the National Academy of Sciences
+of the United States of America 110:19456--9.
+
+Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
+Human Domination of Earth' s Ecosystems. Science 277:494--499.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
+across scales in metacommunities. Ecology Letters.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+Yenni, G., P. B. Adler, and S. K. {Morgan Ernest}. 2012. Strong
+self-limitation promotes the persistence of rare species. Ecology
+93:456--461.
+
+Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
+sagebrush steppe: long-term data for analyzing demographic rates and
+plant--plant interactions. Ecology 91:3427.
+
 \end{document}

### 2016-08-12 runscripts/make_ms_plots.R

@@ -118,14 +118,15 @@ colnames(theoretical_preds_and_obs) <- c("site", "envonly_pred", "demonly_pred",
 theoretical_preds_and_obs$site <- site_labels_order
 
 ##  Make the main (all sims) plot -----
-pointocols <- rep("grey25",3)
+pointocols <- rep("black",3)
+exp_colors <- c("black", "grey25", "grey45", "grey65", "grey85", "white")
 ggplot(data=plot_df)+
-  geom_bar(data=plot_df, aes(x=simulation, y=synchrony, fill=site, color=site),
-           stat="identity")+
-  geom_errorbar(data=plot_df, aes(x=simulation, y=synchrony, fill=site, color=site, 
+  geom_bar(data=plot_df, aes(x=simulation, y=synchrony, fill=simulation),
+           stat="identity", color="black")+
+  geom_errorbar(data=plot_df, aes(x=simulation, y=synchrony, fill=simulation, color=site, 
                              ymin=lower_synch, ymax=upper_synch), 
                 color="white", size=1, width=0.25)+
-  geom_errorbar(data=plot_df, aes(x=simulation, y=synchrony, fill=site, color=site, 
+  geom_errorbar(data=plot_df, aes(x=simulation, y=synchrony,
                              ymin=lower_synch, ymax=upper_synch), width=0.25)+
   geom_point(data=theoretical_preds_and_obs, aes(x=3, y=obs), size=3.5, color="white", shape=15)+
   geom_point(data=theoretical_preds_and_obs, aes(x=3, y=obs), size=3, color=pointocols[1], shape=15)+
@@ -134,8 +135,8 @@ ggplot(data=plot_df)+
   geom_point(data=theoretical_preds_and_obs, aes(x=6, y=demonly_pred), size=3.5, color="white", shape=17)+
   geom_point(data=theoretical_preds_and_obs, aes(x=6, y=demonly_pred), size=3, color=pointocols[3], shape=17)+
   facet_wrap("site", nrow=1)+
-  scale_fill_manual(values=site_colors, labels=site_labels, name="")+
-  scale_color_manual(values=site_colors, labels=site_labels, name="")+
+  scale_fill_manual(values=exp_colors, labels=site_labels, name="")+
+  scale_color_manual(values=exp_colors, labels=site_labels, name="")+
[…]

### 2016-08-12 docs/ms_communitySynchrony.Rmd
submission version for Ecology

@@
 - \usepackage{color, soul}
+- \usepackage{times} 
+- \usepackage[document]{ragged2e}
 layout: 12pt
@@
 
-<!--    - \usepackage{times} 
-   - \usepackage[document]{ragged2e} -->
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+
+<!--    
+    -->
 
@@
 Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
-The stabilizing effect of species richness stems from individual species responding in different way} to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
+The stabilizing effect of species richness stems from individual species responding in different ways to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
 Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
@@
 We focus on synchrony of per capita growth rates, rather than abundances, because growth rates represent the instantaneous response of species to the environment and competition, and are less susceptible to the legacy effects of drift and disturbance [@Loreau2008a].
-We present equivalent results for synchrony of species abundances in the Online Supporting Information, and show that they lead to the same overall conclusions as synchrony of per capita growth rates.
+We present equivalent results for synchrony of species abundances in the Appendix S1, and show that they lead to the same overall conclusions as synchrony of per capita growth rates.
 Following @Loreau2008a and @DeMazancourt2013a, we define population growth, ignoring observation error, as
@@
 $\sigma_{di}^2$ is the demographic variance and $u_{di}(t)$ are independent normal variables with zero mean and unit variance.
-To derive analytical predictions we solved a first-order approximation of Equation 2 [@DeMazancourt2013a and Online Supporting Information].
+To derive analytical predictions we solved a first-order approximation of Equation 2 [@DeMazancourt2013a and Appendix S1].
 Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos [@Loreau2008a].
@@
 Vital rate regressions are the building blocks of our dynamic models: an integral projection model (IPM) and an individual-based model (IBM).
-We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment (see Online Supporting Information for full details).
+We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment (see Appendix S1 for full details).
 We modeled survival probability of each genet as function of genet size, temporal variation among years, permanent spatial variation among groups of quadrats, and local neighborhood crowding from conspecific and heterospecific genets.
@@
 Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.
-Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Online Supporting Information).
+Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Appendix S1).
 
@@
 For all five communities, the synchrony of species' growth rates when growing in isolation almost perfectly matched species synchrony in polyculture (Fig. 3).
-Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Online Supporting Information, Figs.
+Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Appendix S1, Figs.
 S2-S3).
@@
 We also found that analyses based on synchrony of species' percent cover, rather than growth rates, were uninformative (Figs.
-S2-S3) since the linear approximation required for analytical predictions is a stronger assumption for abundance than growth rates, especially given relatively short time-series (Online Supporting Information).
+S2-S3) since the linear approximation required for analytical predictions is a stronger assumption for abundance than growth rates, especially given relatively short time-series (Appendix S1).
 Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].

### 2016-08-12 docs/ms_communitySynchrony.tex
submission version for Ecology

@@
 \usepackage{color, soul}
+\usepackage{times}
+\usepackage[document]{ragged2e}
 
@@
 \maketitle
+
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 
@@
 2010). The stabilizing effect of species richness stems from individual
-species responding in different way\} to environmental fluctuations
+species responding in different ways to environmental fluctuations
 (environmental stochasticity), or fluctuating asynchronously because of
@@
 (Loreau and {{de Mazancourt}} 2008). We present equivalent results for
-synchrony of species abundances in the Online Supporting Information,
-and show that they lead to the same overall conclusions as synchrony of
-per capita growth rates. Following Loreau and {{de Mazancourt}} (2008)
-and {{de Mazancourt}} et al. (2013), we define population growth,
-ignoring observation error, as
+synchrony of species abundances in the Appendix S1, and show that they
+lead to the same overall conclusions as synchrony of per capita growth
+rates. Following Loreau and {{de Mazancourt}} (2008) and {{de
+Mazancourt}} et al. (2013), we define population growth, ignoring
+observation error, as
 
@@
 analytical predictions we solved a first-order approximation of Equation
-2 ({{de Mazancourt}} et al. 2013 and Online Supporting Information). Due
-to the linear approximation approach, our analytical predictions will
-likely fail in communities where species exhibit large fluctuations due
-to limit cycles and chaos (Loreau and {{de Mazancourt}} 2008). Indeed,
-one of the advantages of focusing on growth rates rather than abundances
-is that growth rates are more likely to be well-regulated around an
-equilibrium value, if the long-term average of a species' growth rate is
-relatively small (e.g., \(r < 2\)).
+2 ({{de Mazancourt}} et al. 2013 and Appendix S1). Due to the linear
+approximation approach, our analytical predictions will likely fail in
+communities where species exhibit large fluctuations due to limit cycles
+and chaos (Loreau and {{de Mazancourt}} 2008). Indeed, one of the
+advantages of focusing on growth rates rather than abundances is that
+growth rates are more likely to be well-regulated around an equilibrium
+value, if the long-term average of a species' growth rate is relatively
+small (e.g., \(r < 2\)).
 
@@
 followed the approach of Chu and Adler (2015) to fit statistical models
-for survival, growth, and recruitment (see Online Supporting Information
-for full details). We modeled survival probability of each genet as
-function of genet size, temporal variation among years, permanent
-spatial variation among groups of quadrats, and local neighborhood
-crowding from conspecific and heterospecific genets. Regression
-coefficients for the effect of crowding by each species can be
-considered a matrix of interaction coefficients whose diagonals
-represent intraspecific interactions and whose off-diagonals represent
-interspecific interactions (Adler et al. 2010). These interaction
-coefficients can take positive (facilitative) or negative (competitive)
-values. We modeled growth as the change in size of a genet from one year
-to the next, which depends on the same factors as the survival model. We
-fit the survival and growth regressions using INLA (Rue et al. 2014), a
+for survival, growth, and recruitment (see Appendix S1 for full
+details). We modeled survival probability of each genet as function of
+genet size, temporal variation among years, permanent spatial variation
+among groups of quadrats, and local neighborhood crowding from
+conspecific and heterospecific genets. Regression coefficients for the
+effect of crowding by each species can be considered a matrix of
+interaction coefficients whose diagonals represent intraspecific
+interactions and whose off-diagonals represent interspecific
+interactions (Adler et al. 2010). These interaction coefficients can
+take positive (facilitative) or negative (competitive) values. We
+modeled growth as the change in size of a genet from one year to the
+next, which depends on the same factors as the survival model. We fit
+the survival and growth regressions using INLA (Rue et al. 2014), a
 statistical package for fitting generalized linear mixed effects models
@@
 temporal (interannual) variation was treated as random effects on the
-intercept and the effect of genet size in the previous year (Online
-Supporting Information).
+intercept and the effect of genet size in the previous year (Appendix
+S1).
 
@@
 less consistent with analytical predictions and observed synchrony
-(Online Supporting Information, Figs. S2-S3).
+(Appendix S1, Figs. S2-S3).
 
@@
 assumption for abundance than growth rates, especially given relatively
-short time-series (Online Supporting Information). Thus, our results
-provide further evidence that it is difficult to decipher mechanisms of
-species synchrony from abundance time series, as expected by theory
-(Loreau and {{de Mazancourt}} 2008). Observed synchrony of per capita
-growth rates was positively correlated with the variability of percent
-cover across our focal communities, which confirms that we are
-investigating an important process underlying ecosystem stability.
+short time-series (Appendix S1). Thus, our results provide further
+evidence that it is difficult to decipher mechanisms of species
+synchrony from abundance time series, as expected by theory (Loreau and
+{{de Mazancourt}} 2008). Observed synchrony of per capita growth rates
+was positively correlated with the variability of percent cover across
+our focal communities, which confirms that we are investigating an
+important process underlying ecosystem stability.
 

### 2016-08-12 docs/online_supporting_information.Rmd
submission version for Ecology

@@
    - \usepackage{morefloats}
+   - \usepackage{times} 
+   - \usepackage[document]{ragged2e}
 bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
@@
 ---
+
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 

### 2016-08-12 runscripts/make_ms_plots.R

@@ -234,6 +234,34 @@ ggplot(ibm_demo_rms, aes(x=(expansion^2), y=avg_synch, color=experiment))+
         axis.text.y = element_text(size=16))
 
 
+##  Make monoculture v. polyculture plot ----
+polymono <- ipm_synch[,c("site", "typesynch", "experiment", "mean_synch")]
+polymono_wide <- dcast(polymono, site+typesynch~experiment, value.var = "mean_synch")
+plot_df <- subset(polymono_wide,typesynch=="pgr_synch")
+# plot_df$site <- c("2Arizona", "5Idaho", "3Kansas", "4Montana", "1New Mexico")
+
+ggplot(plot_df, aes(x=ENVNOINTER, y=ENVINTER))+
+  geom_abline(aes(intercept=0, slope=1), linetype=3)+
+  geom_point(size=5)+
+  # geom_text(aes(label=site), hjust = 0, nudge_x = -0.13)+
+  scale_y_continuous(limits=c(0,1))+
+  scale_x_continuous(limits=c(0,1))+
+  ylab("Species synchrony in competition")+
+  xlab("Species synchrony in isolation")+
+  # scale_shape_discrete(name="Temporal Variable", labels=c("Per capita growth rate", "Percent cover"))+
+  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"),
+                     name = "",
+                     labels = c("New Mexico", "Arizona", "Kansas", "Montana", "Idaho"))+
+  theme_few()+
+  theme(axis.text.x = element_text(size=14),
+        axis.text.y = element_text(size=14),
+        axis.title=element_text(size=16))
+
+ggsave("../docs/components/poly_vs_mono_synch.png", width = 4.5, height = 4, units = "in")
+
+
+
+
 ####
 ####  SUPPLEMENTAL FIGURES 2-3
 ####
@@ -308,14 +336,15 @@ colnames(theoretical_preds_and_obs) <- c("site", "envonly_pred", "demonly_pred",
 theoretical_preds_and_obs$site <- site_labels_order
 
 ##  Make the main (all sims) plot -----
-pointocols <- rep("grey25",3)
+pointocols <- rep("black",3)
+exp_colors <- c("black", "grey25", "grey45", "grey65", "grey85", "white")
 ggplot(data=plot_df)+
-  geom_bar(data=plot_df, aes(x=simulation, y=synchrony, fill=site, color=site),
[…]

### 2016-10-31 docs/ms_communitySynchrony.Rmd
initial edits; minor comments

@@
     fig_caption: yes
-    keep_tex: yes
+    keep_tex: no
 urlcolor: black
@@
 \begin{center}
-\large{\textbf{Environmental responses, not species interactions, determine species synchrony in natural plant communities}}
+\large{\textbf{Environmental responses, not species interactions, determine \textcolor{blue}{synchrony of dominant species in semiarid grasslands}}}
 
@@
 
-\textbf{Running Head}: Drivers of species synchrony in natural plant communities
+\textbf{Running Head}: Drivers of species synchrony in \textcolor{blue}{grasslands}
 
@@
 ### The model
-While existing theory has identified the factors driving synchrony, we do not have a simple expression to predict synchrony in a particular community with all factors operating simultaneously. 
+While existing theory has identified \textcolor{blue}{environmental responses, demographic stochasticity, and species interactions as the drivers of the temporal synchrony of species' per capita growth rates, and thus abundances,} we do not have a simple expression to predict synchrony in a particular community with all factors operating simultaneously.
 However, we can derive analytical predictions for species synchrony under special limiting cases.
@@
 $r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*.
-Environmental stochasticity is incorporated as $\sigma_{ei}u_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $u_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated between species.
+Environmental stochasticity is incorporated as $\sigma_{ei}u_{ei}(t)$, where $\sigma_{ei}^2$ is the \textcolor{blue}{temporal variance of environmental conditions}.
+\textcolor{blue}{Species-specific environmental resposnes are included as} $u_{ei}$, which are normal random variables with zero mean and unit variance that are independent through time but may be correlated between species.
 Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first-order, normal approximation [@Lande2003; @DeMazancourt2013a].
@@
 
-\noindent where $\sigma_{r_{i}}$ is the temporal variance of species *i'*s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
+\noindent where $\sigma_{r_{i}}$ is the temporal \textcolor{blue}{standard deviation} of species *i'*s per capita population growth rate ($r_i$) and $\sigma^{2}_{T}$ is the temporal variance of the aggregate community-level growth rate.
 $\phi$ ranges from 0 at perfect asynchrony to 1 at perfect synchrony [@Loreau2008a].
@@
 (6) IBM: Interspecific interactions and environmental stochasticity removed.
+\textcolor{blue}{We do not include a simulation with only interspecific interactions because our population models run to determinstic equilibriums in the absence of environmental or demographic perturbations.}
 We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
-This allowed species time to reach their stable size distribution.
+This allowed species \textcolor{red}{\st{time}} to reach their stable size distribution.
 We then calculated the synchrony of species' per capita growth rates over 100 randomly selected contiguous 50 time-step sections.
@@
 
-A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
+\textcolor{red}{\st{A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
 Most communities are dominated by few common species and many rare species [@McGill2007], meaning that the low number of common species in our focal commiunities is not unusual. 
@@
 Under such conditions, synchrony will depend on the degree of pathogen-induced fitness differences and the pathogen's response to environmental conditions.
-Neither our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].
+Neither our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].}}
 
@@
 We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a].
-We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
+We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. \textcolor{blue}{1}).
 The lack of an effect of interspecific interactions on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
@@
 
-However, our results do contrast with a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives species synchrony in grasslands [@Gross2014].
+However, our results do contrast with a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives \textcolor{blue}{the degree of} species synchrony in grasslands [@Gross2014].
 The apparent inconsistency between our results and those of @Gross2014 may be explained by the differences between our studies.
@@
 Such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
-In contrast, synchrony may be driven by environmental fluctuations in older communities.
+
+In contrast, \textcolor{blue}{the degree of} synchrony may be driven by environmental fluctuations in older communities. 
 In other words, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
@@
 Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
-When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2).
+When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. \textcolor{blue}{1}).
 A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
@@
   \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
-Only'') and when environmental stochasticity is also present removed (``D.S.
+Only'') and when environmental stochasticity is also present \textcolor{red}{\st{removed}} (``D.S.
 + E.S.'').
@@
   \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).
-We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.
-The dashed line is the line of equality.
+\textcolor{red}{\st{We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.}} The dashed line is the line of equality.
 Simulation results in this figure are analagous to ``No Comp.

### 2016-11-02 docs/ms_communitySynchrony.Rmd
adds overview section to ms

@@
 Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, and demographic and environmental stochasticity), and that these models successfully reproduce observed community dynamics [@Chu2015], perturbing the models can reveal the processes that determine species synchrony in our focal grassland communities.
+
+OVERVIEW OF OUR APPROACH
+------------------------
+\textcolor{blue}{This paper has two related goals.
+First, we want to refine previous theory on the drivers of species synchrony} [@Loreau2008a].
+\textcolor{blue}{We do not develop new theory, but we do derive new quantitative predictions from previous theory that can be tested with empirical data.
+With explicit predictions in hand, we can (i) test the generality of a relatively simple quantitative model and (ii) place our empirical results in a broader theoretical context.
+Second, we want to use multispecies population models to run simulation experiments that allow us to quantify the relative influence of theoretical drivers of species synchrony.
+With empirically-based simulations in hand, we can (i) determine the most important drivers of species synchrony in our focal grassland communities and (ii) compare simulation results to theoretical predictions.}
+
+\textcolor{blue}{Note that throughout this paper we refer to ``synchrony'' as a continuous variable.
+Dynamics can be more or less synchronous or asynchronous, but synchrony and asynchrony are not dichotomous states.
+Thus, our goal is to determine which processes (environmental responses, demographic stochasticity, and interspecific interactions) most affect the value of community-wide species synchrony, regardless of whether those processes make synchrony increase or decrease.}
 

### 2016-11-03 docs/ms_communitySynchrony.Rmd
more revisions; figure updates

@@
 - \usepackage[document]{ragged2e}
+- \usepackage{titlesec}
+- \titleformat{\section}{\Large\sc\centering}{\thesection}{1em}{}
 layout: 12pt
@@
 
-INTRODUCTION
-------------
+\section{Introduction}
 Ecosystems are being transformed by species extinctions [@Cardinale2012], changes in community composition [@Dornelas2014a; @Vellend2013], and anthropogenic environmental change [@Vitousek1997], impacting the provisioning and stability of ecosystem services [@Loreau2001a; @Hooper2005; @Rockstrom2009].
 Experiments have provided compelling evidence that decreases in species richness will decrease productivity [@Tilman2001] and the temporal stability of productivity [@Tilman2006; @Hector2010].
-The stabilizing effect of species richness stems from individual species responding in different ways to environmental fluctuations (environmental stochasticity), or fluctuating asynchronously because of random chance events (demographic stochasticity) [@Hector2010; @Isbell2009; @DeMazancourt2013a].
-Species richness affects synchrony because larger species pools are more likely to contain species that respond disimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
-Even without species losses, abiotic homogenization can weaken compensatory dynamics and, in turn, decrease temporal stability of ecosystem functioning [@Hautier2014].
+\textcolor{blue}{The stabilizing effect of species richness arises from a combination of selection effects and complementarity} [@Loreau2001c].
+\textcolor{blue}{Selection effects occur when a dominant species has higher (lower) than average temporal variability, which generates a negative (positive) effect on ecosystem stability} [e.g., @Grman2010].
+\textcolor{blue}{Complementarity occurs when species have unique dynamics through time, causing their abundances to fluctuate asynchronously.}
+\textcolor{blue}{All else being equal, lower synchrony of species dynamics through time results in greater ecosystem stability} [@Loreau2013].
 The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.
+
+\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], result from individual species responding in different ways to environmental fluctuations, independent species' dynamics because of random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
+Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
+\textcolor{blue}{More species rich communities may also exhibit lower synchrony, and thus higher stability, due to a greater chance of strong species interactions} [@Gross2014].
+\textcolor{blue}{Determining whether synchrony is primarily an outcome of species-specific responses to environmental conditions or interspecific competition is thus an important step toward understanding how species losses and abiotic changes will impact ecosystem stability} [@Hautier2014].
 
@@
 The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific interactions is to remove them one-by-one, and in combination.
-In principle, this could be done in an extremely controlled laboratory setting, but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
+In principle, this could be done in an extremely controlled laboratory setting [e.g., @Venail2013], but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
 For example, @Mutshinda2009 fit a dynamic population model to several community time series of insect and bird abundances.
@@
 
-OVERVIEW OF OUR APPROACH
-------------------------
+\section{Overview of Our Approach}
 \textcolor{blue}{This paper has two related goals.
@@
 
-THEORETICAL MODEL
------------------
+\textcolor{blue}{Our empirical analysis relies on modeling the dynamics of dominant species in five natural semiarid grasslands.
+Thus, our focal communities are subsets of the full community.
+While this limits our inference to the dynamics of dominant species, the value of our study lies in our ability to explicitly quantify the relative influence of different drivers on species synchrony} [e.g., @Venail2013].
+
+\section{Theoretical Model}
 ### The model
@@
 We present equivalent results for synchrony of species abundances in the Appendix S1, and show that they lead to the same overall conclusions as synchrony of per capita growth rates.
+
 Following @Loreau2008a and @DeMazancourt2013a, we define population growth, ignoring observation error, as
@@
 $\sigma_{di}^2$ is the demographic variance and $u_{di}(t)$ are independent normal variables with zero mean and unit variance.
+
 To derive analytical predictions we solved a first-order approximation of Equation 2 [@DeMazancourt2013a and Appendix S1].
@@
 \noindent{} where $\text{cov}(u_{ei},u_{ej})$ is the standardized covariance of environmental responses between species *i* and species *j*.
+
 Confronting our theoretical predictions with data requires estimates of species dynamics of large populations (no demographic stochasticity) growing in isolation (no interspecific interactions) to calculate the covariance of species' environmental responses.
@@
 
-EMPIRICAL ANALYSIS
-------------------
+\section{Empirical Analysis}
 ### Materials and methods
@@
 We then project the model forward by using the survival regression to determine whether a genet lives or dies, the growth regression to calculate changes in genet size, and the recruitment regression to add new individuals that are distributed randomly in space.
-Crowding is directly calculated at each time step since each genet is spatially referenced (as in the observed data).
+Crowding is directly calculated at each time step since each genet is spatially referenced.
 Environmental stochasticity is not an inherent feature of IBMs, but is easily included since we fit year-specific temporal random effects for each vital rate regression.
@@
 
-The IPM uses the same vital rate regressions as the IBM [@Rees2009; @Rees2014], but it is spatially implicit and does not include demographic stochasticity.
+The IPM uses the same vital rate regressions as the IBM, but it is spatially implicit and does not include demographic stochasticity.
 Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step (Supporting Online Information).
@@
 plots) of different size.
-Indeed, results from an IBM with infinite population size would converge on results from the IPM.
+Results from an IBM with infinite population size would converge on results from the IPM.
 Given computational constraints, the largest landscape we simulate is a 25 $\text{m}^2$ plot.
@@
 We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
-This allowed species \textcolor{red}{\st{time}} to reach their stable size distribution.
-We then calculated the synchrony of species' per capita growth rates over 100 randomly selected contiguous 50 time-step sections.
+This allowed species to reach their stable size distribution.
+We then calculated yearly per capita growth rates from the simlated time series, and then caluclated the synchrony of species' per capita growth rates over 100 randomly selected contiguous 50 time-step sections.
+
 We ran IBM simulations for 100 time steps, and repeated the simulations 100 times for each simulation experiment.
 From those, we retained only the simulations in which no species went extinct due to demographic stochasticity.
-Synchrony was calculated over the 100 time steps for each no extinction run within a model experiment.
-To explore the effect of demographic stochasticity in different sized populations, we ran simulations (1) and (6) on plot sizes of 1, 4, 9, 16, and 25 $\text{m}^2$.
-All other IBM simulations were run on a 25 $\text{m}^2$ landscape.
+Synchrony of per capita growth rates was calculated over the 100 time steps for each no extinction run within a model experiment.
+\textcolor{blue}{In the IBM simulations the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
+To explore this effect}, we ran simulations \textcolor{blue}{(4)} and (6) on plot sizes of 1, 4, 9, 16, and 25 $\text{m}^2$.
+All other IBM simulations were run on a 25 $\text{m}^2$ landscape \textcolor{blue}{to most closely match the implicit spatial scale of the IPM simulations}.
 
+\textcolor{blue}{Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulation-based values of community-wide species synchrony.}
 Results from our simulation experiments also allow us to test our theoretical predictions.
@@
 ### Results
-Synchrony of species' per capita growth rates at our study sites range from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
-Synchrony tends to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability in percent cover (e.g., Montana).
-Synchrony is lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
-Synchrony of per capita growth rates and CV of percent cover are positively correlated (Pearson’s $\rho$ = 0.72).
+\textcolor{blue}{Observed} synchrony of species' per capita growth rates at our study sites range from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
+Synchrony of per capita growth rates and CV of percent cover are positively correlated (Pearson’s $\rho$ = 0.72, *N* = 5).
 For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
@@
 
+<!--
+Synchrony tends to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability in percent cover (e.g., Montana).
+Synchrony is lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
+-->
+
 Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Fig. 1 and Table S1).
-The correlation between our analytical predictions and simulated synchrony was 0.97 for $\phi_{R,\mathcal{M}_D}$ and 0.997 for $\phi_{R,\mathcal{M}_E}$.
+The correlation between our analytical predictions and simulated synchrony was 0.97 for $\phi_{R,\mathcal{M}_D}$ (*N* = 5) and 0.997 for $\phi_{R,\mathcal{M}_E}$ (*N* = 5).
 The largest difference between predicted and simulated synchrony was 0.05 in New Mexico for $\phi_{R,\mathcal{M}_D}$ (Table S1).
@@
 Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 1), but the effect was small (9% increase; Table S2).
-Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S3).
+Overall, species' temporal random effects in the statistical vital rate models are positively, but not perfectly, correlated (Table S3).
 Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific interactions had little effect on synchrony (Fig. 1; "No Comp." simulations).
@@
 
-The desynchronizing effect of demographic stochasticity, which increases as population size decreases, modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
+In IBM simulations, the desynchronizing effect of demographic stochasticity, which increases as population size decreases, modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
 2).
 In the largest, 25 $\text{m}^2$ plots, synchrony was driven by environmental stochasticity (e.g., $\mathcal{M}_E$).
-At 1 $\text{m}^2$, synchrony reflected demographic stochasticity and environmental stochasticity (e.g., between $\mathcal{M}_E$ and $\mathcal{M}_D$).
+At 1 $\text{m}^2$, synchrony reflected \textcolor{blue}{the combined effects of} demographic stochasticity and environmental stochasticity (e.g., \textcolor{blue}{simulated synhcrony fell} between $\mathcal{M}_E$ and $\mathcal{M}_D$).
 For context, population sizes increased from an average of 17 individuals per community in 1 $\text{m}^2$ IBM simulations to an average of 357 individuals per community in 25 $\text{m}^2$ IBM simulations.
 
+<!--
 For all five communities, the synchrony of species' growth rates when growing in isolation almost perfectly matched species synchrony in polyculture (Fig. 3).
+-->
 Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Appendix S1, Figs.
@@
 
-DISCUSSION
-----------
+\section{Discussion}
 Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, though not perfectly, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.
@@
 For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.
-However, the theoretical predictions closely matched results from simulations of population models fit to long-term data from natural plant communities (Table 3).
+However, the theoretical predictions closely matched results from simulations of population models fit to long-term data from natural plant communities (Table 2).
 Such strong agreement between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory.
 
-### Demographic stochasticity decreases synchrony
-Demographic stochasticity partially counteracted the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations were small (Fig.
-2), in agreement with theory [@Loreau2008a].
-Even in small populations, however, demographic stochasticity was not strong enough to compensate the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
+### Demographic stochasticity decreases synchrony \textcolor{blue}{in small populations}
+In large populations, removing demographic stochasticity had no effect on species synchrony (Fig. 1).
+In small populations, demographic stochasticity partially counteracted the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, in agreement with theory [@Loreau2008a].
+\textcolor{blue}{This is shown in Fig.
+2, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity, but the differences between simulations are smaller at lower population sizes.}
+Even in small populations \textcolor{blue}{(e.g., 1 $\text{m}^2$ lanscapes)}, however, demographic stochasticity was not strong enough to compensate the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
 2).
@@
 First, in our simulation experiments, removing interspecific interactions resulted in no discernible change in community-wide species synchrony of per capita growth rates (Fig. 1).
-Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig.
-1), and the synchrony of species in isolation was a very strong predictor of synchrony of species in polyculture (Fig.
-3).
+Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig. 1).
 Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony in the communities we studied.
@@
 
-\textcolor{red}{\st{A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
+<!--
+A challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
 Most communities are dominated by few common species and many rare species [@McGill2007], meaning that the low number of common species in our focal commiunities is not unusual. 
@@
 Under such conditions, synchrony will depend on the degree of pathogen-induced fitness differences and the pathogen's response to environmental conditions.
-Neither our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].}}
+Neither our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].
+-->
 
@@
 
-However, our results do contrast with a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition drives \textcolor{blue}{the degree of} species synchrony in grasslands [@Gross2014].
+<!--
+However, our results do contrast with a recent analysis of several biodiversity-ecosystem functioning experiments showing that competition \textcolor{blue}{determines the degree of} species synchrony in grasslands [@Gross2014].
 The apparent inconsistency between our results and those of @Gross2014 may be explained by the differences between our studies.
@@
 One way to test this prediction is to continue collecting data from biodiversity-ecoystem functioning experiments and conduct the analysis of @Gross2014 every few years.
+-->
 
@@
 
-CONCLUSIONS
------------
+\section{Conclusions}
 Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
@@
 
-ACKNOWLEDGMENTS
----------------
+\section{Acknowledgements}
 This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
@@
 
+<!--
+
 \begin{figure}[!ht]

### 2016-11-03 devel/plot_timeseries.R

@@ -7,5 +7,25 @@ rm(list=ls())
 par(mfrow=c(1,2))
 ibms <- readRDS("../results/ibm_sims/ibm_Kansas_constnointerexpand1.RDS")
 matplot(ibms[,c(4:5)]*100, type="l")
-ibms <- readRDS("../results/ibm_sims/ibm_Kansas_constnointerexpand5.RDS")
-matplot(ibms[,c(4:5)]*100, type="l")
+ibms <- readRDS("../results/ibm_sims/ibm_Montana_fluctnointerexpand5.RDS")
+matplot(ibms[101:200,c(4:7)]*100, type="l", lty = 1, lwd=2)
+
+# caclulate observed growth rates
+# create lagged data frame to only get observed yearly transitions
+ts_mat <- as.data.frame(ibms[1:100,c(2,4:7)])
+lag_df <- ts_mat
+lag_df$lagyear <- lag_df$time+1
+num_spp=4
+species_list <- c("Cov.BOGR","Cov.HECO","Cov.PASM","Cov.POSE")
+colnames(lag_df)[2:(num_spp+1)] <- paste(colnames(lag_df)[2:(num_spp+1)],"_t0", sep="") 
+# merge the lag df with observed
+rm_col <- which(colnames(lag_df)=="time")
+merged_df <- merge(ts_mat, lag_df[,-rm_col], by.x = "time", by.y="lagyear")
+transitions <- nrow(merged_df)
+obs_gr <- matrix(nrow=transitions, ncol=num_spp)
+name_ids1 <- which(colnames(merged_df) %in% species_list)
+name_ids2 <- which(colnames(merged_df) %in% paste(species_list, "_t0", sep=""))
+for(i in 1:transitions){
+  obs_gr[i,] <- as.numeric(log(merged_df[i,name_ids1]/merged_df[i,name_ids2]))
+}
+matplot(obs_gr, type="l", lty = 1, lwd=2)

### 2016-11-04 docs/ms_communitySynchrony.Rmd
updates synchrony function to calc synch by plots then average

@@
 For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.
-However, the theoretical predictions closely matched results from simulations of population models fit to long-term data from natural plant communities (Table 2).
+However, the theoretical predictions closely matched results from simulations of population models fit to long-term data from natural plant communities (Table S1).
 Such strong agreement between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory.

### 2016-11-04 docs/ms_communitySynchrony.Rmd
more revisoions

@@
 Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific interactions, and demographic stochasticity.
-We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
+We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors \textcolor{blue}{on the synchrony of dominant species in five semiarid grasslands.}
 We found that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
 Removing interspecific interactions and demographic stochasticity had small effects on synchrony. 
-In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+\textcolor{blue}{For the dominant species in these plant communities}, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
 
@@
 
-\emph{Key words: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability}
+\emph{Key words: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability, grassland}
 
@@
 Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony.
-Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific interactions on species synchrony in natural plant communities.
-To achieve these objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
+Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific interactions on \textcolor{blue}{the synchrony of dominant species} in natural plant communities.
+\textcolor{blue}{Our focus is limited to dominant species due to data constraints, but previous work indicates that the dynamics of dominant species are most important for ecosystem functioning in grasslands} [@Smith2003; @Bai2004; @Grman2010; @Sasaki2011].
+
+To achieve our objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
 We then confront our theoretical predictions with simulations from the empirically-based population models.
 Second, we use the multi-species population models to perform simulation experiments that isolate the effects of environmental stochasticity, demographic stochasticity, and interspecific interactions on community-wide species synchrony.
-Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, and demographic and environmental stochasticity), and that these models successfully reproduce observed community dynamics [@Chu2015], perturbing the models can reveal the processes that determine species synchrony in our focal grassland communities.
+Given that our population models capture the essential features of community dynamics important to synchrony (density-dependence, and demographic and environmental stochasticity), and that these models successfully reproduce observed community dynamics [@Chu2015], perturbing the models can reveal the processes that determine synchrony \textcolor{blue}{of dominant species} in our focal grassland communities.
 

### 2016-11-07 docs/ms_communitySynchrony.Rmd
updates all minor edits

@@
 @Thibaut2012 used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.
-While a major step forward, Mutshinda et al.'s (2009) and Thibaut et al.'s (2012) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Furthermore, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
+
+While a major step forward, Mutshinda et al.'s (2009) and Thibaut et al.'s (2012) modeling technique does have some limitations.
+\textcolor{blue}{First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
+Second, they relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
+Third, fluctuations in abundance may mask the underlying mechanisms that underpin species synchrony [@Loreau2008a] and ecosystem stability} [@Loreau2013] \textcolor{blue}{due to legacy effects and ecological drift.
+Theory indicates that a focus on the fluctuations of species' per capita growth rates better reveals the mechanisms that ultimately determine species synchrony} [@Loreau2008a].
+\textcolor{blue}{Thus, synchrony of per capita growth rates is the focus of this paper.}
 
@@
 Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first-order, normal approximation [@Lande2003; @DeMazancourt2013a].
-$\sigma_{di}^2$ is the demographic variance and $u_{di}(t)$ are independent normal variables with zero mean and unit variance.
+$\sigma_{di}^2$ is the demographic variance (\textcolor{blue}{i.e., the intrinsic demographic stochasticity of species \emph{i}}) and $u_{di}(t)$ are independent normal variables with zero mean and unit variance \textcolor{blue}{that allow demographic stochasiticity to vary through time.}
 
@@
 We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
-Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.
 Spatial (quadrat groupings) variation was treated as a random effect on the intercept and temporal (interannual) variation was treated as random effects on the intercept and the effect of genet size in the previous year (Appendix S1).
+
+\textcolor{blue}{Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
+In the latter case, important yearly crowding effects would indicate an interaction between environmental conditions and species interactions because we assume that the yearly random effects encompass environmental variability, generally.
+We tested for environment-species interactions by comparing models with and without yearly random effects on crowding.
+In the end, crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.}
 
@@
 Overall, species' temporal random effects in the statistical vital rate models are positively, but not perfectly, correlated (Table S3).
+\textcolor{blue}{Our approach assumes that these temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.}
+
 Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific interactions had little effect on synchrony (Fig. 1; "No Comp." simulations).
@@
 Such strong agreement between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory.
+\textcolor{blue}{However, whether the theoretical model adequately represents more complex communities remains unknown because we had to focus on dominant species.}
 
@@
 The perennial grasses we studied do have similar responses to the environment (Table S2), which will tend to synchronize dynamics.
-However, if community-wide species synchrony is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
+However, if community-wide synchrony \textcolor{blue}{among dominant species} is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
+\textcolor{blue}{Even ecologically-similar species, which are sometimes aggregated into to functional groups, have environmental responses that are dissimilar enough to cause asynchrony.
+Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.}
+
+<!--
 In agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
@@
 Future research on the influence of community assembly on synchrony could reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.
+-->
 
@@
 Compute, storage, and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.
-We also thank three anonymous reviewers who provided critical feedback that improved the manuscript.
+We also thank several anonymous reviewers who provided critical feedback that improved the manuscript.
 

### 2016-11-07 docs/ms_communitySynchrony.Rmd
adds functional group discussion paragraph

@@
 Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig. 1).
-Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony in the communities we studied.
+Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony \textcolor{blue}{among the dominant species} we studied.
 Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural plant communities.
@@
 The interaction of facilitation and competition is clearly capable of having a large effect on species synchrony, and future theoretical efforts should aim to include a wider range of species interactions.
+
+\textcolor{blue}{Environmental responses synchronized dynamics relative to a null expectation of independent species interactions (e.g., "No Comp.
++ No E.S." simulations in Fig.
+1), but observed and simulated synchrony was still less than one in all cases (Fig.
+1).
+This means that there is sufficient temporal niche differentiation among these species to create asynchronous dynamics.
+Many studies of ecosystem stability in semiarid grasslands focus on tradoffs among dominant functional types, such as grasses versus forbs} [@Sasaki2011] \textcolor{blue}{or among many grouped lifeforms} [@Bai2004].
+\textcolor{blue}{Such groupings are based on the idea that ecologically-similar species will have similar environmental responses through time, a valid assumption without evidence to the contrary.
+At first glance our results may appear to support the grouping of perennial grasses because their environmental responses were correlated.
+However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 1).
+The subtle differences among ecologically-similar dominant species does impact species synchrony and, ultimately, ecosystem stability.
+Ignoring such difference may, though not always, mask important dynamics that underpin ecosystem functioning.}
 

### 2016-11-07 docs/ms_communitySynchrony.Rmd
small updates

@@
 With explicit predictions in hand, we can (i) test the generality of a relatively simple quantitative model and (ii) place our empirical results in a broader theoretical context.
-Second, we want to use multispecies population models to run simulation experiments that allow us to quantify the relative influence of theoretical drivers of species synchrony.
+Second, we want to use multispecies population models to run simulation experiments that allow us to quantify the relative influence of theoretical drivers of species synchrony (see Fig.
+S1 for workflow).
 With empirically-based simulations in hand, we can (i) determine the most important drivers of species synchrony in our focal grassland communities and (ii) compare simulation results to theoretical predictions.}
@@
 ### Materials and methods
+\textcolor{blue}{A workflow diagram of our empirical analysis is shown in Fig. S1.}
+
 #### Data
@@
 Synchrony of per capita growth rates and CV of percent cover are positively correlated (Pearson’s $\rho$ = 0.72, *N* = 5).
-For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
-IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to the desynchronizing effect of demographic stochasticity.
+For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S2).
+IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. S2), likely due to the desynchronizing effect of demographic stochasticity.
 
@@
 Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific interactions had little effect on synchrony (Fig. 1; "No Comp." simulations).
-Removing interspecific interactions caused, at most, a 5% change in synchrony (Fig. 1).
+Removing interspecific interactions caused, at most, a 5% change in synchrony (Fig. 1 and Table S2).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
@@
 Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Appendix S1, Figs.
-S2-S3).
+S3-S4).
 
@@
 We also found that analyses based on synchrony of species' percent cover, rather than growth rates, were uninformative (Figs.
-S2-S3) since the linear approximation required for analytical predictions is a stronger assumption for abundance than growth rates, especially given relatively short time-series (Appendix S1).
+S3-S4) since the linear approximation required for analytical predictions is a stronger assumption for abundance than growth rates, especially given relatively short time-series (Appendix S1).
 Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
@@
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
-For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.
+For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S5-S8), as required by our predictions.
 However, the theoretical predictions closely matched results from simulations of population models fit to long-term data from natural plant communities (Table S1).
-Such strong agreement between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory.
+Strong agreement between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory.
 \textcolor{blue}{However, whether the theoretical model adequately represents more complex communities remains unknown because we had to focus on dominant species.}
@@
 tripartita* from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
-S8).
+S9).
 Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.

### 2016-11-08 docs/ms_communitySynchrony.Rmd
adds PBA edits to R1

@@
 Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific interactions, and demographic stochasticity.
-We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors \textcolor{blue}{on the synchrony of dominant species in five semiarid grasslands.}
+We then conduct simulation experiments with empirical population models to quantify the relative \textcolor{blue}{influence} of these factors \textcolor{blue}{on the synchrony of dominant species in five semiarid grasslands.}
 We found that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
@@
 \textcolor{blue}{The stabilizing effect of species richness arises from a combination of selection effects and complementarity} [@Loreau2001c].
-\textcolor{blue}{Selection effects occur when a dominant species has higher (lower) than average temporal variability, which generates a negative (positive) effect on ecosystem stability} [e.g., @Grman2010].
+\textcolor{blue}{Selection effects occur when a dominant species has higher than average temporal variability, which generates a negative effect on ecosystem stability} [e.g., @Grman2010].
 \textcolor{blue}{Complementarity occurs when species have unique dynamics through time, causing their abundances to fluctuate asynchronously.}
@@
 
-\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], result from individual species responding in different ways to environmental fluctuations, independent species' dynamics because of random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
-Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions [@Yachi1999], implying that species losses will reduce ecosystem stability.
+\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], result from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
+Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions, reducing synchrony [@Yachi1999].
 \textcolor{blue}{More species rich communities may also exhibit lower synchrony, and thus higher stability, due to a greater chance of strong species interactions} [@Gross2014].
@@
 While a major step forward, Mutshinda et al.'s (2009) and Thibaut et al.'s (2012) modeling technique does have some limitations.
-\textcolor{blue}{First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change when each factor is removed.
+\textcolor{blue}{First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change if each factor were removed.
 Second, they relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Third, fluctuations in abundance may mask the underlying mechanisms that underpin species synchrony [@Loreau2008a] and ecosystem stability} [@Loreau2013] \textcolor{blue}{due to legacy effects and ecological drift.
-Theory indicates that a focus on the fluctuations of species' per capita growth rates better reveals the mechanisms that ultimately determine species synchrony} [@Loreau2008a].
-\textcolor{blue}{Thus, synchrony of per capita growth rates is the focus of this paper.}
+Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony [@Loreau2008a] and ecosystem stability} [@Loreau2013] \textcolor{blue}{due to legacy effects and ecological drift.
+Theory indicates that a focus on the fluctuations of species' per capita growth rates better reveals the mechanisms that ultimately determine species synchrony} [@Loreau2008a],
+\textcolor{blue}{which is why we focus per capita growth rates.}
 
@@
 1).
-This means that there is sufficient temporal niche differentiation among these species to create asynchronous dynamics.
+Asynchronous dynamics result because the responses of theses species to environmental conditions are sufficiently different.
 Many studies of ecosystem stability in semiarid grasslands focus on tradoffs among dominant functional types, such as grasses versus forbs} [@Sasaki2011] \textcolor{blue}{or among many grouped lifeforms} [@Bai2004].

### 2016-11-08 docs/ms_communitySynchrony.Rmd
fixes labeling in figure 1

@@
 - \usepackage{titlesec}
-- \titleformat{\section}{\Large\sc\centering}{\thesection}{1em}{}
+- \titleformat{\section}{\large\sc\centering}{\thesection}{1em}{}
 layout: 12pt
@@
 Theory indicates that a focus on the fluctuations of species' per capita growth rates better reveals the mechanisms that ultimately determine species synchrony} [@Loreau2008a],
-\textcolor{blue}{which is why we focus per capita growth rates.}
+\textcolor{blue}{which is why we focus on per capita growth rates.}
 
@@
 
+<!--
 \section{Overview of Our Approach}
@@
 While this limits our inference to the dynamics of dominant species, the value of our study lies in our ability to explicitly quantify the relative influence of different drivers on species synchrony} [e.g., @Venail2013].
+-->
 

### 2016-11-08 docs/ms_communitySynchrony.Rmd
updates ms with workflow diagram

@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-
-<!--    
-    -->
 
@@
 
-Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony.
+Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony (Fig. 1).
 Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific interactions on \textcolor{blue}{the synchrony of dominant species} in natural plant communities.
@@
 ### Materials and methods
-\textcolor{blue}{A workflow diagram of our empirical analysis is shown in Fig. S1.}
 
@@
 Synchrony of per capita growth rates and CV of percent cover are positively correlated (Pearson’s $\rho$ = 0.72, *N* = 5).
-For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S2).
-IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. S2), likely due to the desynchronizing effect of demographic stochasticity.
+For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
+IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to the desynchronizing effect of demographic stochasticity.
 
@@
 
-Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Fig. 1 and Table S1).
+Across the five communities, our limiting case predictions closely matched synchrony from the corresponding simulation experiment (Fig. 2 and Table S1).
 The correlation between our analytical predictions and simulated synchrony was 0.97 for $\phi_{R,\mathcal{M}_D}$ (*N* = 5) and 0.997 for $\phi_{R,\mathcal{M}_E}$ (*N* = 5).
@@
 Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out of five communities (Fig.
-1).
+2).
 Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
-Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 1), but the effect was small (9% increase; Table S2).
+Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S2).
 Overall, species' temporal random effects in the statistical vital rate models are positively, but not perfectly, correlated (Table S3).
@@
 
-Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific interactions had little effect on synchrony (Fig. 1; "No Comp." simulations).
-Removing interspecific interactions caused, at most, a 5% change in synchrony (Fig. 1 and Table S2).
+Species interactions are weak in these communities [Table S4 and @Chu2015], and removing interspecific interactions had little effect on synchrony (Fig. 2; "No Comp." simulations).
+Removing interspecific interactions caused, at most, a 5% change in synchrony (Fig. 2 and Table S2).
 Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig.
-1), with an average 6% increase over synchrony from IBM simulations on a 25$\text{m}^2$ area.
+2), with an average 6% increase over synchrony from IBM simulations on a 25$\text{m}^2$ area.
 
 In IBM simulations, the desynchronizing effect of demographic stochasticity, which increases as population size decreases, modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig.
-2).
+3).
 In the largest, 25 $\text{m}^2$ plots, synchrony was driven by environmental stochasticity (e.g., $\mathcal{M}_E$).
@@
 Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Appendix S1, Figs.
-S3-S4).
+S2-S3).
 
@@
 We also found that analyses based on synchrony of species' percent cover, rather than growth rates, were uninformative (Figs.
-S3-S4) since the linear approximation required for analytical predictions is a stronger assumption for abundance than growth rates, especially given relatively short time-series (Appendix S1).
+S2-S3) since the linear approximation required for analytical predictions is a stronger assumption for abundance than growth rates, especially given relatively short time-series (Appendix S1).
 Thus, our results provide further evidence that it is difficult to decipher mechanisms of species synchrony from abundance time series, as expected by theory [@Loreau2008a].
@@
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
-For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S5-S8), as required by our predictions.
+For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.
 However, the theoretical predictions closely matched results from simulations of population models fit to long-term data from natural plant communities (Table S1).
@@
 ### Demographic stochasticity decreases synchrony \textcolor{blue}{in small populations}
-In large populations, removing demographic stochasticity had no effect on species synchrony (Fig. 1).
+In large populations, removing demographic stochasticity had no effect on species synchrony (Fig. 2).
 In small populations, demographic stochasticity partially counteracted the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, in agreement with theory [@Loreau2008a].
 \textcolor{blue}{This is shown in Fig.
-2, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity, but the differences between simulations are smaller at lower population sizes.}
+3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity, but the differences between simulations are smaller at lower population sizes.}
 Even in small populations \textcolor{blue}{(e.g., 1 $\text{m}^2$ lanscapes)}, however, demographic stochasticity was not strong enough to compensate the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
-2).
+3).
 These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
-Only in the Idaho community does synchrony of per capita growth rates approach $\mathcal{M}_D$ in a non-constant environment (Fig. 2).
+Only in the Idaho community does synchrony of per capita growth rates approach $\mathcal{M}_D$ in a non-constant environment (Fig. 3).
 This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 25 $\text{m}^2$ quadrat would only contain a few individuals of that species.
@@
 Our analysis of how demographic stochasticity affects synchrony demonstrates that synchrony depends on the observation area.
-As the observation area increases, population size increases and the desynchronizing effect of demographic stochasticity lessens (Fig. 2).
+As the observation area increases, population size increases and the desynchronizing effect of demographic stochasticity lessens (Fig. 3).
 Thus, our results suggest that community-wide species synchrony will increase as the observation area increases, rising from $\mathcal{M}_D$ to $\mathcal{M}_E$.
@@
 Two lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
-First, in our simulation experiments, removing interspecific interactions resulted in no discernible change in community-wide species synchrony of per capita growth rates (Fig. 1).
-Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig. 1).
+First, in our simulation experiments, removing interspecific interactions resulted in no discernible change in community-wide species synchrony of per capita growth rates (Fig. 2).
+Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig. 2).
 Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony \textcolor{blue}{among the dominant species} we studied.
@@
 tripartita* from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig.
-S9).
+S8).
 Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.
@@
 + No E.S." simulations in Fig.
-1), but observed and simulated synchrony was still less than one in all cases (Fig.
-1).
+2), but observed and simulated synchrony was still less than one in all cases (Fig.
+2).
 Asynchronous dynamics result because the responses of theses species to environmental conditions are sufficiently different.
@@
 At first glance our results may appear to support the grouping of perennial grasses because their environmental responses were correlated.
-However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 1).
+However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 2).
 The subtle differences among ecologically-similar dominant species does impact species synchrony and, ultimately, ecosystem stability.
@@
 We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a].
-We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. \textcolor{blue}{1}).
+We found that community synchrony was virtually indistinguishable between simulations with and without interspecific interactions (Fig. 2).
 The lack of an effect of interspecific interactions on synchrony is in contrast to a large body of theoretical work that predicts a strong role for competition in creating compensatory dynamics [@Tilman1988] and a recent empirical analysis [@Gross2014].
@@
 Our results do not contradict the idea that competition can lead to compensatory dynamics, but they do highlight the fact that interspecific competition must be relatively strong to influence species synchrony.
-The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables S1, S3-S17).
+The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables S1 and S3-S17).
 Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap decreases.
@@
 Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
-When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. \textcolor{blue}{1}).
+When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2).
 A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
@@
 -------
-```{r time_series_fig, echo=FALSE, include=TRUE, fig.width=8.5, fig.height=4, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site.
-Panels are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.", eval=F}
-out_ks$percent_cover$site <- "Kansas"
-out_id$percent_cover$site <- "Idaho"
-out_mt$percent_cover$site <- "Montana"
-out_nm$percent_cover$site <- "New Mexico"
-out_az$percent_cover$site <- "Arizona"
-cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
-                    out_mt$percent_cover, out_nm$percent_cover, 
-                    out_az$percent_cover)
 
-out_ks$growth_rates$site <- "Kansas"
-out_id$growth_rates$site <- "Idaho"
-out_mt$growth_rates$site <- "Montana"
-out_nm$growth_rates$site <- "New Mexico"
-out_az$growth_rates$site <- "Arizona"
-pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
-                  out_mt$growth_rates, out_nm$growth_rates, 
-                  out_az$growth_rates)
-all_spp <- unique(pgr_data$species)
-color_match <- data.frame(species = all_spp,
-                          color = brewer.pal(12,"Paired"))
-pgr_data <- merge(pgr_data, color_match, by = "species")
-cover_data <- merge(cover_data, color_match, by="species")
-
-### Create labelling dataframe
-site_list <- unique(cover_data$site)
-textdf <- data.frame(species=NA, year=NA, tot_cover=NA, num_obs=NA, site=NA, color=NA)
-for(dosite in site_list){
-  tmp <- subset(cover_data, site==dosite)
-  if(dosite=="Kansas"){
-    maxyr <- min(unique(tmp$year))
-    textdf <- rbind(textdf,tmp[tmp$year == maxyr, ])
-  }
-  if(dosite!="Kansas"){
-    maxyr <- max(unique(tmp$year))
-    textdf <- rbind(textdf,tmp[tmp$year == maxyr, ])
-    textdf$year <- textdf$year+2
-  }
-}
-textdf <- textdf[2:nrow(textdf),]
-
-### Change site to two-letter codes
-site_list <- sort(unique(cover_data$site))
-sitecode <- c("AZ", "ID", "KS", "MT", "NM")
-for(i in 1:length(site_list)){
-  cover_data[cover_data$site==site_list[i], "site"] <- sitecode[i]
-  pgr_data[pgr_data$site==site_list[i], "site"] <- sitecode[i]
-}
-
-cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
-                geom_line(alpha=0.25)+
-                geom_point(size=2)+
-                facet_grid(site~., scales = "free_y")+
-                guides(color=FALSE)+
-                ylab("Mean Percent Cover")+
-                xlab("Year")+
-                scale_color_manual(values = brewer.pal(12,"Paired"))+
-                theme_bw()
-
-pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
-              geom_line(alpha=0.25)+
-                geom_point(size=2)+
-              facet_grid(site~.)+
-              guides(color=FALSE)+
-              ylab("Per Capita Growth Rate")+
-              xlab("Year")+
-              scale_color_manual(values = brewer.pal(12,"Paired"))+
-              theme_bw()
-full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
-```
-
-```{r concept_fig, echo=FALSE, fig.cap="Conceptual figure outlining the effects of interspecific interactions and environmental stochasticity in large populations where demographic stochasticity is effectively absent (A) and in populations of varying size where demographic stochasticity is present (B).", eval=F}
-library(png)
-library(grid)
-img <- readPNG("./components/synchrony_conceptual_fig.png")
-grid.raster(img)
-```
-
-```{r read_data, echo=FALSE, eval=T}
-output_list <- readRDS("../results/ipm_comp_nocomp_sims.RDS")
-mlist <- melt(output_list)
-```
-
-```{r figure_1, echo=FALSE, fig.cap="Observed (vertical dashed lines) and simulated (solid density curves) synchrony of species per capita growth rates at each site from the IBM (top panels) and the IPM (bottom panels).
-IPM density curves come from 100 random contiguous sections from 2,000 iteration IPM runs, where the length of each randomly selected section is equal to the numnber of observation years for each data set.
-IBM density curves come from 100 replicate simulations of 75 iterations each.
-Synchrony values come from simulations where environmental stochasticity and interspecific interactions are present.
-The IBM was run on a 5 by 5 meter landscape to reduce the effect of demographic stochasticity.", fig.width=6, fig.height=6, fig.width=6, eval=F}
-nyrs <- c(max(out_az$stability$num_observations),
-          max(out_id$stability$num_observations),
-          max(out_ks$stability$num_observations),
-          max(out_mt$stability$num_observations),
-          max(out_nm$stability$num_observations))
-
-boots <- 100
-sites <- sort(table_data$Site)
-sites[5] <- "NewMexico"
-synch <- matrix(NA, boots, length(sites))
-
-for(do_site in 1:length(sites)){
-  full_ipm_outs <- subset(mlist, L2=="ENVINTER")
-  site_ipm <- subset(full_ipm_outs, L1==sites[do_site])
-  colnames(site_ipm) <- c("year", "species", "cover", "L1", "L2")
-  # if(sites[do_site]=="Idaho") site_ipm <- subset(site_ipm, Var2!="ARTR")
-  species_list <- unique(site_ipm$species)
-  num_spp <- length(species_list)
-  for(i in 1:boots){
-    begin_year <- sample(x = 1:(max(site_ipm$year)-nyrs[do_site]), 1)
-    end_year <- begin_year+nyrs[do_site]
-    tmp <- subset(site_ipm, year %in% begin_year:end_year)
-    tmp2 <- dcast(tmp, year~species, value.var = "cover")
-
-    # caclulate observed growth rates
-    # create lagged data frame to only get observed yearly transitions
-    lag_df <- tmp2
-    lag_df$lagyear <- lag_df$year+1
-    colnames(lag_df)[2:(num_spp+1)] <- paste0(colnames(lag_df)[2:(num_spp+1)],"_t0")
-
-    # merge the lag df with observed
-    rm_col <- which(colnames(lag_df)=="year")
-    merged_df <- merge(tmp2, lag_df[,-rm_col], by.x = "year", by.y="lagyear")
-    transitions <- nrow(merged_df)
-    obs_gr <- matrix(nrow=transitions, ncol=num_spp)
-    name_ids1 <- which(colnames(merged_df) %in% species_list)
-    name_ids2 <- which(colnames(merged_df) %in% paste(species_list, "_t0", sep=""))
-    for(ii in 1:transitions){
-      obs_gr[ii,] <- as.numeric(log(merged_df[ii,name_ids1]/merged_df[ii,name_ids2]))
-    }
-    synch[i,do_site] <- as.numeric(community.sync(obs_gr)[1])
-  }
-}
-
-colnames(synch) <- sites
-synch_df <- melt(synch)
-
-obs_pgr_df <- c(as.numeric(out_az$pgr_synchrony[1]),
-                as.numeric(out_id$pgr_synchrony[1]),
-                as.numeric(out_ks$pgr_synchrony[1]),
-                as.numeric(out_mt$pgr_synchrony[1]),
-                as.numeric(out_nm$pgr_synchrony[1]))
-obs_pgr_df <- data.frame(pgr_synch=obs_pgr_df,
-                         Var2=sites)
-
-### Get IBM synchs
-synch_df$type <- "IPM"
-synch_df <- synch_df[,2:ncol(synch_df)]
-ibm_poly <- readRDS("../devel/mono_poly_ibm.RDS")
-ibm_poly <- subset(ibm_poly, typesynch=="Per capita growth rate")
-ibm_poly$type <- "IBM"
-ibm_df <- ibm_poly[,c("site","synch","type")]
-names(ibm_df) <- names(synch_df)
-synch_df <- rbind(synch_df, ibm_df)
-
-g1 <- ggplot()+
-  geom_line(data=subset(synch_df, type=="IPM"), aes(x=value, color=Var2), 
-            stat="density", adjust=2)+
-  geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
-  facet_wrap("Var2", ncol=1, scales="free_y")+
-  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
-  guides(color=FALSE)+
-  xlab("Community Synchrony")+
-  ylab("Estimated Probability Density")+
-  ggtitle("IPM")+
-  scale_x_continuous(limits=c(0,1))+
-  theme_bw()
-
-g2 <- ggplot()+
-  geom_line(data=subset(synch_df, type=="IBM"), aes(x=value, color=Var2), 
-            stat="density", adjust=2)+
-  geom_vline(data=obs_pgr_df, aes(xintercept=pgr_synch, color=Var2), linetype=2)+
-  facet_wrap("Var2", ncol=1, scales="free_y")+
-  scale_color_manual(values = c("grey45", "steelblue", "slateblue4", "darkorange", "purple"))+
-  guides(color=FALSE)+
-  xlab("Community Synchrony")+
-  ylab("Estimated Probability Density")+
-  ggtitle("IBM")+
-  scale_x_continuous(limits=c(0,1))+
-  theme_bw()
-
-grid.arrange(g1,g2,ncol=2)
-```
-
-<!--
 \begin{figure}[!ht]
   \centering
-      \includegraphics[width=3in]{./components/formatted_fig1.png}
-  \caption{Observed versus predicted synchrony of growth rates under our two limiting case predictions.
-The observed synchrony of per capita growth rates was best predicted by our theoretical prediction in the limiting case of only environmental stochasticity operating ($\phi_R$; green line and points in left panel): it was the only regression line whose slope and intercept did not significantly differ from zero or one, respectively.}
+      \includegraphics[width=6in]{./components/formatted_figures/synchrony_flowchart_bw.png}
+  \caption{Diagram of our coupled theoretical-empirical approach.}
 \end{figure}
@@
 \pagebreak{}
--->
 
@@
       \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
-  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
-Only'') and when environmental stochasticity is also present \textcolor{red}{\st{removed}} (``D.S.
-+ E.S.'').
+  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
++ No E.S.
+(D.S.
+Only)'') and when environmental stochasticity is also present (``No Comp.
+(D.S.
++ E.S.)'').
 The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
 The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
-Theoretically, ``D.S.
-Only'' simulations should remain constant across landscape size, whereas ``D.S.
-+ E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
+Theoretically, ``No Comp.
++ No E.S.
+(D.S.
+Only)'' simulations should remain constant across landscape size, whereas ``No Comp.
+(D.S.
++ E.S.)'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
 Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
@@
 
-<!--
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/formatted_figures/formatted_figure3.png}
-  \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).
-\textcolor{red}{\st{We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.}} The dashed line is the line of equality.
-Simulation results in this figure are analagous to ``No Comp.
-+ No D.S.'' (species in isolation) and ``No D.S.'' (species in competition) in Fig.
-1, but here we control the time series of random year effects.}
-\end{figure}
-
-\pagebreak{}
--->
-
-```{r time_series, eval=FALSE, echo=FALSE, fig.cap="Example time series from the simulations.", eval=F}
-ggplot(mlist, aes(x=year, y=(cover*100), color=species))+
-  geom_line()+
-  facet_grid(L1~L2, scale="free")
-```
-
 References

### 2016-11-14 docs/ms_communitySynchrony.Rmd
adds in Claire and Michel's comments; sent to PBA

@@
 - \usepackage{titlesec}
-- \titleformat{\section}{\large\sc\centering}{\thesection}{1em}{}
+- \titleformat{\section}{\normalsize\sc\centering}{\thesection}{1em}{}
+- \titleformat{\subsection}{\normalsize\it\centering}{\thesubsection}{1em}{}
+- \titleformat*{\paragraph}{\normalsize\it}
 layout: 12pt
@@
 ---
-
+<!--\definecolor{blue}{gray}{0}-->
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
@@
 \textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], result from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
-Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions, reducing synchrony [@Yachi1999].
-\textcolor{blue}{More species rich communities may also exhibit lower synchrony, and thus higher stability, due to a greater chance of strong species interactions} [@Gross2014].
+Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions, reducing synchrony and increasing stability [@Yachi1999].
+\textcolor{blue}{More species rich communities may also exhibit lower synchrony due to a greater chance of strong species interactions, but interspecific competition generally decreases stability of total community biomass} [@Loreau2013; but see @Gross2014].
 \textcolor{blue}{Determining whether synchrony is primarily an outcome of species-specific responses to environmental conditions or interspecific competition is thus an important step toward understanding how species losses and abiotic changes will impact ecosystem stability} [@Hautier2014].
@@
 Second, they relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony [@Loreau2008a] and ecosystem stability} [@Loreau2013] \textcolor{blue}{due to legacy effects and ecological drift.
-Theory indicates that a focus on the fluctuations of species' per capita growth rates better reveals the mechanisms that ultimately determine species synchrony} [@Loreau2008a],
-\textcolor{blue}{which is why we focus on per capita growth rates.}
+Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony, or at least make them more difficult to identify.
+The synchrony of species' abundances ultimately determines the stability of total community biomass, but the processes that drive species synchrony are most tightly linked to each species' immediate response to environmental conditions and competition} [@Loreau2008a].
+\textcolor{blue}{Therefore, we focus on per capita growth rates, which represent a species' immediate response to interannual fluctuations.}
 
@@
 \section{Theoretical Model}
-### The model
+\subsection{The model}
 While existing theory has identified \textcolor{blue}{environmental responses, demographic stochasticity, and species interactions as the drivers of the temporal synchrony of species' per capita growth rates, and thus abundances,} we do not have a simple expression to predict synchrony in a particular community with all factors operating simultaneously.
@@
 $r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*.
-Environmental stochasticity is incorporated as $\sigma_{ei}u_{ei}(t)$, where $\sigma_{ei}^2$ is the \textcolor{blue}{temporal variance of environmental conditions}.
-\textcolor{blue}{Species-specific environmental resposnes are included as} $u_{ei}$, which are normal random variables with zero mean and unit variance that are independent through time but may be correlated between species.
+Environmental stochasticity is incorporated as $\sigma_{ei}u_{ei}(t)$, where $\sigma_{ei}^2$ is the \textcolor{blue}{temporal variance of species \emph{i}'s response to the environment and $u_{ei}(t)$ is a normal variable with unit variance that is independent among species but may be correlated.}
+\textcolor{blue}{The product, $\sigma_{ei}u_{ei}(t)$, is species \emph{i}'s environmental response.}
 Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first-order, normal approximation [@Lande2003; @DeMazancourt2013a].
@@
 
-### Predictions
+\subsection{Predictions}
 Our first prediction assumes no interspecific interactions, no environmental stochasticity, identical intrinsic growth rates, and that demographic stochasticity is operating but all species have identical demographic variances.
@@
 \section{Empirical Analysis}
-### Materials and methods
+\subsection{Materials and methods}
 
-#### Data
+#### Data.--
 We use long-term demographic data from five semiarid grasslands in the western United States [described in detail by @Chu2015].
@@
 
-#### Calculating observed synchrony
+#### Calculating observed synchrony.--
 The data consist of records for individual plant size in quadrats for each year.
@@
 
-#### Fitting statistical models
+#### Fitting statistical models.--
 Vital rate regressions are the building blocks of our dynamic models: an integral projection model (IPM) and an individual-based model (IBM).
@@
 
-#### Building dynamic multi-species models
+#### Building dynamic multi-species models.--
 Once we have fit the vital rate statistical models, building the population models is straightforward.
@@
 
-#### Simulation experiments
+#### Simulation experiments.--
 We performed simulation experiments where drivers (environmental stochasticity, demographic stochasticity, or interspecific interactions) were removed one-by-one and in combination.
@@
 
-### Results
+\subsection{Results}
 \textcolor{blue}{Observed} synchrony of species' per capita growth rates at our study sites range from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
@@
 
-### Simulations support theoretical predictions
+\subsection{Simulations support theoretical predictions}
 Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
@@
 
-### Demographic stochasticity decreases synchrony \textcolor{blue}{in small populations}
+\subsection{Demographic stochasticity decreases synchrony \textcolor{blue}{in small populations}}
 In large populations, removing demographic stochasticity had no effect on species synchrony (Fig. 2).
 In small populations, demographic stochasticity partially counteracted the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, in agreement with theory [@Loreau2008a].
-\textcolor{blue}{This is shown in Fig.
-3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity, but the differences between simulations are smaller at lower population sizes.}
+\textcolor{blue}{This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
+The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, thus reducing the relative effect of environmental forcing.}
 Even in small populations \textcolor{blue}{(e.g., 1 $\text{m}^2$ lanscapes)}, however, demographic stochasticity was not strong enough to compensate the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
@@
 
-### Environmental fluctuations drive community synchrony
+\subsection{Environmental fluctuations drive community synchrony}
 In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations (Equation 4).
@@
 
-### Interspecific interactions had little impact on community synchrony
+\subsection{Interspecific interactions had little impact on community synchrony}
 We expected community synchrony of per capita growth rates to decrease when we removed interspecific interactions [@Loreau2008a].
@@
 
-Tables
-------
+\section{Tables}
 
@@
 
-Figures
--------
+\section{Figures}
 

### 2016-11-21 docs/ms_communitySynchrony.Rmd
adds in pba edits

@@
 \textcolor{blue}{The stabilizing effect of species richness arises from a combination of selection effects and complementarity} [@Loreau2001c].
-\textcolor{blue}{Selection effects occur when a dominant species has higher than average temporal variability, which generates a negative effect on ecosystem stability} [e.g., @Grman2010].
-\textcolor{blue}{Complementarity occurs when species have unique dynamics through time, causing their abundances to fluctuate asynchronously.}
-\textcolor{blue}{All else being equal, lower synchrony of species dynamics through time results in greater ecosystem stability} [@Loreau2013].
-The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.
+\textcolor{blue}{Selection effects occur when a dominant species has lower than average temporal variability, which generates a positive effect on ecosystem stability} [e.g., @Grman2010].
+\textcolor{blue}{Complementarity occurs when species have unique temporal dynamics, causing their abundances to fluctuate asynchronously, which leads an increase in ecosystem stability.}
+\textcolor{blue}{Understanding the mechanisms driving species' temporal dynamics, and resulting (a)synchrony, can help us predict the impacts of global change on ecosystem stability.}
 
-\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], result from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
+\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], \textcolor{blue} occur whenever species synchrony is not perfect} and results from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
 Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions, reducing synchrony and increasing stability [@Yachi1999].
-\textcolor{blue}{More species rich communities may also exhibit lower synchrony due to a greater chance of strong species interactions, but interspecific competition generally decreases stability of total community biomass} [@Loreau2013; but see @Gross2014].
+\textcolor{blue}{Species richness can also affect synchrony if the strength of species interactions varies systematically with richness, because competition generally increases synchrony and reduces stability} [@Loreau2013; but see @Gross2014].
+
 \textcolor{blue}{Determining whether synchrony is primarily an outcome of species-specific responses to environmental conditions or interspecific competition is thus an important step toward understanding how species losses and abiotic changes will impact ecosystem stability} [@Hautier2014].
+\textcolor{blue}{If responses to environment are important, then environmental change could alter synchrony and stability.
+If competition is important, then the direct effects of environmental change may not affect synchrony and stability, but species gains/losses will.
 
-Theory identifies three main determinants of species synchrony: environmental stochasticity, demographic stochasticity, and interspecific interactions [@Loreau2008a; @Gonzalez2009; @Loreau2013].
-For example, in a community composed of large populations (no demographic stochasticity) with weak interspecific interations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
-However, this prediction relies on a relatively simple population model and only holds under two assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
-Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities [@Mutshinda2009; @Thibaut2012], and they did not consider demographic stochasiticity.
-
-In grasslands, most empirical studies have focused on whether species synchrony is primarily an outcome of species-specific responses to environmental conditions [@Hautier2014] or competition [@Gross2014].
-Even beyond grassland studies, whether competition or environmental responses drive compensatory dynamics remains controversial [reviewed in @Gonzalez2009].
+Whether competition or environmental responses drive compensatory dynamics \textcolor{blue}{in natural plant communities} remains controversial [reviewed in @Gonzalez2009].
 In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from the covariance matrix of species abundances [e.g., @Houlahan2007] is impossible.
@@
 For example, weak synchrony of population abundances could reflect positive environmental correlations (synchronizing effect) offset by strong competition (desynchronizing effect), or negative environmental correlations and weak competition.
+
+Theory can help us resolve this empirical question.} 
+Theory identifies three main determinants of species synchrony: environmental stochasticity, interspecific interactions, and demographic stochasticity [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+\textcolor{blue}{This theory has been developed by focusing on simple limiting cases in which only one of these three drivers operates.}
+For example, in a community composed of large populations (no demographic stochasticity) with weak interspecific interations, community-wide species synchrony should be determined by the covariance of species' responses to the environment [@Loreau2008a].
+However, this prediction relies on a relatively simple population model and \textcolor{blue}{depends on two restrictive} assumptions: (i) species' responses to the environment are similar in magnitude and (ii) all species have similar growth rates.
+Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities [@Mutshinda2009; @Thibaut2012], and they did not consider demographic stochasiticity.
 
@@
 \textcolor{blue}{First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change if each factor were removed.
-Second, they relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
-Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony, or at least make them more difficult to identify.
+Second, they relied on popluation abundance data that may or may not reliably capture competitive interactions occuring at the individual level.
+Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony.
 The synchrony of species' abundances ultimately determines the stability of total community biomass, but the processes that drive species synchrony are most tightly linked to each species' immediate response to environmental conditions and competition} [@Loreau2008a].
@@
 
-Here, we use multi-species population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony (Fig. 1).
-Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, demographic stochasticity, and interspecific interactions on \textcolor{blue}{the synchrony of dominant species} in natural plant communities.
-\textcolor{blue}{Our focus is limited to dominant species due to data constraints, but previous work indicates that the dynamics of dominant species are most important for ecosystem functioning in grasslands} [@Smith2003; @Bai2004; @Grman2010; @Sasaki2011].
+Here, we use multispecies population models fit to long-term demographic data from five semi-arid plant communities to test theory on the drivers of species synchrony (Fig. 1).
+Our objectives are to (1) derive and test theoretical predictions of species synchrony and (2) determine the relative influence of environmental stochasticity, interspecific interactions, and demographic stochasticity on \textcolor{blue}{the synchrony of dominant species} in natural plant communities.
+\textcolor{blue}{While our focus is limited to dominant species due to data constraints, previous work indicates that the dynamics of dominant species determine ecosystem functioning in grasslands} [@Smith2003; @Bai2004; @Grman2010; @Sasaki2011].
 
-To achieve our objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases derived from the dynamics of individual species in monoculture.
+To achieve our objectives, we first refine theory that has been used to predict the effects of species richness on ecosystem stability [@DeMazancourt2013a] and species synchrony [@Loreau2008a] to generate predictions of community-wide species synchrony under two limiting cases.
 We then confront our theoretical predictions with simulations from the empirically-based population models.
@@
 \subsection{The model}
-While existing theory has identified \textcolor{blue}{environmental responses, demographic stochasticity, and species interactions as the drivers of the temporal synchrony of species' per capita growth rates, and thus abundances,} we do not have a simple expression to predict synchrony in a particular community with all factors operating simultaneously.
+While existing theory has identified \textcolor{blue}{environmental responses, species interactions, and demographic stochasticity as the drivers of the temporal synchrony,} we do not have a simple expression to predict synchrony in a particular community with all factors operating simultaneously.
 However, we can derive analytical predictions for species synchrony under special limiting cases.
@@
 \textcolor{blue}{Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
-In the latter case, important yearly crowding effects would indicate an interaction between environmental conditions and species interactions because we assume that the yearly random effects encompass environmental variability, generally.
-We tested for environment-species interactions by comparing models with and without yearly random effects on crowding.
-In the end, crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.}
+Strong year-to-year variation in these crowding effects would indicate a statistical interaction between environmental conditions and species interactions.
+We tested for such a dynamic by comparing models with and without random year effects on crowding.
+Based on the results, we decided to treat crowding as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including interannual variation in crowding did not change our results.}
 
@@
 (6) IBM: Interspecific interactions and environmental stochasticity removed.
-\textcolor{blue}{We do not include a simulation with only interspecific interactions because our population models run to determinstic equilibriums in the absence of environmental or demographic perturbations.}
+\textcolor{blue}{We did not include a simulation with only interspecific interactions because our population models run to deterministic equilibriums in the absence of environmental or demographic stochasticity.}
 We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
@@
 Synchrony of per capita growth rates was calculated over the 100 time steps for each no extinction run within a model experiment.
-\textcolor{blue}{In the IBM simulations the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
+\textcolor{blue}{In the IBM simulations, the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
 To explore this effect}, we ran simulations \textcolor{blue}{(4)} and (6) on plot sizes of 1, 4, 9, 16, and 25 $\text{m}^2$.
-All other IBM simulations were run on a 25 $\text{m}^2$ landscape \textcolor{blue}{to most closely match the implicit spatial scale of the IPM simulations}.
+All other IBM simulations were run on a 25 $\text{m}^2$ landscape \textcolor{blue}{to most closely match the implicit, large spatial scale of the IPM simulations}.
 
-\textcolor{blue}{Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulation-based values of community-wide species synchrony.}
-Results from our simulation experiments also allow us to test our theoretical predictions.
+\textcolor{blue}{Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulated values of community-wide species synchrony.}
+The simulation experiments also allow us to test our theoretical predictions.
 First, in the absence of interspecific interactions and demographic stochasticity, populations can only fluctuate in response to the environment.
@@
 Second, simulations (5) and (6) represent the simulated version of our limiting case theoretical predictions.
-Thus, we directly test the theoretical predictions by comparing them to observed synchrony and simulated synchrony.
+\textcolor{blue}{This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalent species' environmental and demographic variances.
+Our empirically-based models do not make these assumptions.
+Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions are valid.}
 
@@
 Overall, species' temporal random effects in the statistical vital rate models are positively, but not perfectly, correlated (Table S3).
-\textcolor{blue}{Our approach assumes that these temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.}
+\textcolor{blue}{These temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.}
 
@@
 Strong agreement between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory.
-\textcolor{blue}{However, whether the theoretical model adequately represents more complex communities remains unknown because we had to focus on dominant species.}
+\textcolor{blue}{However, whether the theoretical model adequately represents more complex communities remains unknown because our analysis was restricted to dominant species.}
 
@@
 \textcolor{blue}{This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
-The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, thus reducing the relative effect of environmental forcing.}
+The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, reducing the relative effect of environmental forcing.}
 Even in small populations \textcolor{blue}{(e.g., 1 $\text{m}^2$ lanscapes)}, however, demographic stochasticity was not strong enough to compensate the synchronizing effects of environmental fluctuations and match the analytical prediction where only demographic stochasticity is operating ($\mathcal{M}_D$ in Fig.
@@
 2).
-Asynchronous dynamics result because the responses of theses species to environmental conditions are sufficiently different.
-Many studies of ecosystem stability in semiarid grasslands focus on tradoffs among dominant functional types, such as grasses versus forbs} [@Sasaki2011] \textcolor{blue}{or among many grouped lifeforms} [@Bai2004].
-\textcolor{blue}{Such groupings are based on the idea that ecologically-similar species will have similar environmental responses through time, a valid assumption without evidence to the contrary.
-At first glance our results may appear to support the grouping of perennial grasses because their environmental responses were correlated.
+Synchrony was far from complete because of differences in species' responses to interannual environmental variation.
+Many studies of ecosystem stability in semiarid grasslands focus on trad-offs among dominant functional types} [@Bai2004; @Sasaki2011].
+\textcolor{blue}{Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
+At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were correlated.
 However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 2).
-The subtle differences among ecologically-similar dominant species does impact species synchrony and, ultimately, ecosystem stability.
-Ignoring such difference may, though not always, mask important dynamics that underpin ecosystem functioning.}
+The subtle differences among ecologically-similar dominant species do impact species synchrony and, ultimately, ecosystem stability.
+Ignoring such differences could mask important dynamics that underpin ecosystem functioning.}
 
@@
 The perennial grasses we studied do have similar responses to the environment (Table S2), which will tend to synchronize dynamics.
-However, if community-wide synchrony \textcolor{blue}{among dominant species} is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
-\textcolor{blue}{Even ecologically-similar species, which are sometimes aggregated into to functional groups, have environmental responses that are dissimilar enough to cause asynchrony.
+However, if community-wide synchrony \textcolor{blue}{among dominant species} is less than perfect, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
+\textcolor{blue}{Even ecologically-similar species, which are sometimes aggregated into to functional groups, have environmental responses that are dissimilar enough to limit synchrony.
 Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.}

### 2016-11-21 docs/ms_communitySynchrony.Rmd
updates synchrony flowchart figure

@@
 
-\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], \textcolor{blue} occur whenever species synchrony is not perfect} and results from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
+\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], \textcolor{blue}{occur whenever species synchrony is not perfect} and results from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
 Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions, reducing synchrony and increasing stability [@Yachi1999].
@@
 \textcolor{blue}{If responses to environment are important, then environmental change could alter synchrony and stability.
-If competition is important, then the direct effects of environmental change may not affect synchrony and stability, but species gains/losses will.
+If competition is important, then the direct effects of environmental change may not affect synchrony and stability, but species gains/losses will.}
 

### 2016-11-21 devel/plot_timeseries.R

@@ -4,11 +4,17 @@
 
 rm(list=ls())
 
-par(mfrow=c(1,2))
+# par(mfrow=c(1,2))
 ibms <- readRDS("../results/ibm_sims/ibm_Kansas_constnointerexpand1.RDS")
 matplot(ibms[,c(4:5)]*100, type="l")
+
+
+mycols <- c("#BACB72", "#1E4B58", "#3997B3", "#57A880")
+
 ibms <- readRDS("../results/ibm_sims/ibm_Montana_fluctnointerexpand5.RDS")
-matplot(ibms[101:200,c(4:7)]*100, type="l", lty = 1, lwd=2)
+matplot(ibms[101:200,c(4:7)]*100, type="l", lty = 1, lwd=2, col = mycols, bty="n", xaxt='n', yaxt='n', ann=FALSE)
+axis(1, labels = FALSE)
+axis(2, labels = FALSE)
 
 # caclulate observed growth rates
 # create lagged data frame to only get observed yearly transitions
@@ -28,4 +34,6 @@ name_ids2 <- which(colnames(merged_df) %in% paste(species_list, "_t0", sep=""))
 for(i in 1:transitions){
   obs_gr[i,] <- as.numeric(log(merged_df[i,name_ids1]/merged_df[i,name_ids2]))
 }
-matplot(obs_gr, type="l", lty = 1, lwd=2)
+matplot(obs_gr, type="l", lty = 1, lwd=2, col = mycols, bty="n", xaxt='n', yaxt='n', ann=FALSE)
+axis(1, labels = FALSE)
+axis(2, labels = FALSE)

### 2016-11-22 docs/ms_communitySynchrony.Rmd
final pba edits for ecolR1 submission

@@
 \textcolor{blue}{Selection effects occur when a dominant species has lower than average temporal variability, which generates a positive effect on ecosystem stability} [e.g., @Grman2010].
-\textcolor{blue}{Complementarity occurs when species have unique temporal dynamics, causing their abundances to fluctuate asynchronously, which leads an increase in ecosystem stability.}
-\textcolor{blue}{Understanding the mechanisms driving species' temporal dynamics, and resulting (a)synchrony, can help us predict the impacts of global change on ecosystem stability.}
+\textcolor{blue}{Complementarity occurs when species have unique temporal dynamics, causing their abundances to fluctuate asynchronously, increasing ecosystem stability.}
+\textcolor{blue}{The premise of this paper is that understanding the mechanisms driving species' temporal dynamics, and resulting (a)synchrony, is necessary to predict the impacts of global change on ecosystem stability.}
 
-\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], \textcolor{blue}{occur whenever species synchrony is not perfect} and results from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
+\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics} [@Gonzalez2009], \textcolor{blue}{occur whenever species synchrony is not perfect} and result from individual species responding in different ways to environmental fluctuations, random chance events, and/or competitive interactions [@Hector2010; @Isbell2009; @DeMazancourt2013a; @Gross2014].
 Species richness affects \textcolor{blue}{the degree of synchrony in a community} because larger species pools are more likely to contain species that respond dissimilarly to environmental conditions, reducing synchrony and increasing stability [@Yachi1999].
@@
 
-\textcolor{blue}{Determining whether synchrony is primarily an outcome of species-specific responses to environmental conditions or interspecific competition is thus an important step toward understanding how species losses and abiotic changes will impact ecosystem stability} [@Hautier2014].
+\textcolor{blue}{The effects of environmental change and species losses on ecosystem stability will depend on whether synchrony is driven by species-specific responses to environmental conditions or interspecific competition} [@Hautier2014].
 \textcolor{blue}{If responses to environment are important, then environmental change could alter synchrony and stability.
@@
 
-Whether competition or environmental responses drive compensatory dynamics \textcolor{blue}{in natural plant communities} remains controversial [reviewed in @Gonzalez2009].
-In part, controversy remains because quantifying the relative strengths of each driver on the degree of synchrony from the covariance matrix of species abundances [e.g., @Houlahan2007] is impossible.
-This is because an unbiased null expectation for synchrony does not exist [@Loreau2008a] and observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
+\textcolor{blue}{The relative role of environmental responses and competition in driving synchrony  in natural plant communities} remains controversial [reviewed in @Gonzalez2009].
+One source of the controversy is that quantifying the relative strengths of each driver based on the covariance matrix of species abundances [e.g., @Houlahan2007] is impossible.
+This is because observed synchrony can arise from non-unique combinations of factors [@Ranta2008].
 For example, weak synchrony of population abundances could reflect positive environmental correlations (synchronizing effect) offset by strong competition (desynchronizing effect), or negative environmental correlations and weak competition.
 
-Theory can help us resolve this empirical question.} 
-Theory identifies three main determinants of species synchrony: environmental stochasticity, interspecific interactions, and demographic stochasticity [@Loreau2008a; @Gonzalez2009; @Loreau2013].
+\textcolor{blue}{Theory can help us resolve this empirical question.} 
+Recent theoretical work has identified three determinants of species synchrony: environmental stochasticity, interspecific interactions, and demographic stochasticity [@Loreau2008a; @Gonzalez2009; @Loreau2013].
 \textcolor{blue}{This theory has been developed by focusing on simple limiting cases in which only one of these three drivers operates.}
@@
 
-The best way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific interactions is to remove them one-by-one, and in combination.
+An intuitive way to quantify the effects of environmental stochasticity, demographic stochasticity, and interspecific interactions is to remove them one-by-one, and in combination.
+\textcolor{blue}{If synchrony changes more when we remove environmental stochasticity than when we remove interspecific interactions, we would conclude that environmental fluctuations are the more important driver.}
 In principle, this could be done in an extremely controlled laboratory setting [e.g., @Venail2013], but empirically-based models of interacting populations, fit with data sets from natural communities, offer a practical alternative.
@@
 Second, simulations (5) and (6) represent the simulated version of our limiting case theoretical predictions.
-\textcolor{blue}{This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalent species' environmental and demographic variances.
+\textcolor{blue}{This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalence in species' growth rates, environmental response variances, and demographic variances.
 Our empirically-based models do not make these assumptions.
-Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions are valid.}
+Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions hold in the natural community our model represents.}
 
@@
 -->
-Results for synchrony of percent cover are qualitatively similar, but simulation results were more variable and less consistent with analytical predictions and observed synchrony (Appendix S1, Figs.
-S2-S3).
+Results for synchrony of percent cover are qualitatively similar, but more variable and less consistent with analytical predictions and observed synchrony (Appendix S1, Figs. S2-S3).
 
@@
 \subsection{Simulations support theoretical predictions}
-Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
+Our theoretical predictions were derived from a simple model of population dynamics and required several restrictive assumptions, raising questions about their relevance to natural communities.
 For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.
@@
 3).
-These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
+These results confirm the theoretical argument of @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
 Only in the Idaho community does synchrony of per capita growth rates approach $\mathcal{M}_D$ in a non-constant environment (Fig. 3).
@@
 \textcolor{blue}{Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
-At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were correlated.
+At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were positively correlated.
 However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 2).
@@
 However, if community-wide synchrony \textcolor{blue}{among dominant species} is less than perfect, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
-\textcolor{blue}{Even ecologically-similar species, which are sometimes aggregated into to functional groups, have environmental responses that are dissimilar enough to limit synchrony.
+\textcolor{blue}{Even ecologically-similar species, which are sometimes aggregated into functional groups, have environmental responses that are dissimilar enough to limit synchrony.
 Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.}
@@
       \includegraphics[width=6in]{./components/formatted_figures/synchrony_flowchart_bw.png}
-  \caption{Diagram of our coupled theoretical-empirical approach.}
+  \caption{Diagram of our coupled theoretical-empirical approach. We followed this workflow for each of our five focal communities.}
 \end{figure}

### 2016-11-22 docs/ms_communitySynchrony.Rmd
final small edits for R1 submission

@@
 Compute, storage, and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.
-We also thank several anonymous reviewers who provided critical feedback that improved the manuscript.
+We also thank Patrick Venail and several anonymous reviewers who provided critical feedback that improved the manuscript.
 

### 2016-11-22 docs/ms_communitySynchrony.tex
slight changes for Ecology submission
final small edits for R1 submission

@@
 \usepackage[document]{ragged2e}
+\usepackage{titlesec}
+\titleformat{\section}{\normalsize\sc\centering}{\thesection}{1em}{}
+\titleformat{\subsection}{\normalsize\it\centering}{\thesubsection}{1em}{}
+\titleformat*{\paragraph}{\normalsize\it}
 
@@
 Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific interactions, and demographic stochasticity.
-We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
+We then conduct simulation experiments with empirical population models to quantify the relative \textcolor{blue}{influence} of these factors \textcolor{blue}{on the synchrony of dominant species in five semiarid grasslands.}
 We found that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
 Removing interspecific interactions and demographic stochasticity had small effects on synchrony. 
-In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+\textcolor{blue}{For the dominant species in these plant communities}, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
 
@@
 
-\subsection{INTRODUCTION}\label{introduction}
+\section{Introduction}
 
@@
 temporal stability of productivity (Tilman et al. 2006, Hector et al.
-2010). The stabilizing effect of species richness stems from individual
-species responding in different ways to environmental fluctuations
-(environmental stochasticity), or fluctuating asynchronously because of
-random chance events (demographic stochasticity) (Isbell et al. 2009,
-Hector et al. 2010, {{de Mazancourt}} et al. 2013). Species richness
-affects synchrony because larger species pools are more likely to
-contain species that respond disimilarly to environmental conditions
-(Yachi and Loreau 1999), implying that species losses will reduce
-ecosystem stability. Even without species losses, abiotic homogenization
-can weaken compensatory dynamics and, in turn, decrease temporal
-stability of ecosystem functioning (Hautier et al. 2014). The link
-between synchrony and stability means that a mechanistic understanding
-of synchrony can help us predict the impacts of global change on
-ecosystem stability.
-
-Theory identifies three main determinants of species synchrony:
-environmental stochasticity, demographic stochasticity, and
-interspecific interactions (Loreau and {{de Mazancourt}} 2008, 2013,
-Gonzalez and Loreau 2009). For example, in a community composed of large
-populations (no demographic stochasticity) with weak interspecific
-interations, community-wide species synchrony should be determined by
-the covariance of species' responses to the environment (Loreau and {{de
-Mazancourt}} 2008). However, this prediction relies on a relatively
-simple population model and only holds under two assumptions: (i)
+2010).
+\textcolor{blue}{The stabilizing effect of species richness arises from a combination of selection effects and complementarity}
+(Loreau and Hector 2001).
+\textcolor{blue}{Selection effects occur when a dominant species has lower than average temporal variability, which generates a positive effect on ecosystem stability}
+(e.g., Grman et al. 2010).
+\textcolor{blue}{Complementarity occurs when species have unique temporal dynamics, causing their abundances to fluctuate asynchronously, increasing ecosystem stability.}
+\textcolor{blue}{The premise of this paper is that understanding the mechanisms driving species' temporal dynamics, and resulting (a)synchrony, is necessary to predict the impacts of global change on ecosystem stability.}
+
+\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics}
+(Gonzalez and Loreau 2009),
+\textcolor{blue}{occur whenever species synchrony is not perfect} and
+result from individual species responding in different ways to
+environmental fluctuations, random chance events, and/or competitive
+interactions (Isbell et al. 2009, Hector et al. 2010, {{de Mazancourt}}
+et al. 2013, Gross et al. 2014). Species richness affects
+\textcolor{blue}{the degree of synchrony in a community} because larger
+species pools are more likely to contain species that respond
+dissimilarly to environmental conditions, reducing synchrony and
+increasing stability (Yachi and Loreau 1999).
+\textcolor{blue}{Species richness can also affect synchrony if the strength of species interactions varies systematically with richness, because competition generally increases synchrony and reduces stability}
+(Loreau and {{de Mazancourt}} 2013, but see Gross et al. 2014).
+
+\textcolor{blue}{The effects of environmental change and species losses on ecosystem stability will depend on whether synchrony is driven by species-specific responses to environmental conditions or interspecific competition}
+(Hautier et al. 2014).
+\textcolor{blue}{If responses to environment are important, then environmental change could alter synchrony and stability.
+If competition is important, then the direct effects of environmental change may not affect synchrony and stability, but species gains/losses will.}
+
+\textcolor{blue}{The relative role of environmental responses and competition in driving synchrony  in natural plant communities}
+remains controversial (reviewed in Gonzalez and Loreau 2009). One source
+of the controversy is that quantifying the relative strengths of each
+driver based on the covariance matrix of species abundances (e.g.,
+Houlahan et al. 2007) is impossible. This is because observed synchrony
+can arise from non-unique combinations of factors (Ranta et al. 2008).
+For example, weak synchrony of population abundances could reflect
+positive environmental correlations (synchronizing effect) offset by
+strong competition (desynchronizing effect), or negative environmental
+correlations and weak competition.
+
+\textcolor{blue}{Theory can help us resolve this empirical question.}
+Recent theoretical work has identified three determinants of species
+synchrony: environmental stochasticity, interspecific interactions, and
+demographic stochasticity (Loreau and {{de Mazancourt}} 2008, 2013,
+Gonzalez and Loreau 2009).
+\textcolor{blue}{This theory has been developed by focusing on simple limiting cases in which only one of these three drivers operates.}
+For example, in a community composed of large populations (no
+demographic stochasticity) with weak interspecific interations,
+community-wide species synchrony should be determined by the covariance
+of species' responses to the environment (Loreau and {{de Mazancourt}}
+2008). However, this prediction relies on a relatively simple population
+model and \textcolor{blue}{depends on two restrictive} assumptions: (i)
 species' responses to the environment are similar in magnitude and (ii)
@@
 
-In grasslands, most empirical studies have focused on whether species
-synchrony is primarily an outcome of species-specific responses to
-environmental conditions (Hautier et al. 2014) or competition (Gross et
-al. 2014). Even beyond grassland studies, whether competition or
-environmental responses drive compensatory dynamics remains
-controversial (reviewed in Gonzalez and Loreau 2009). In part,
-controversy remains because quantifying the relative strengths of each
-driver on the degree of synchrony from the covariance matrix of species
-abundances (e.g., Houlahan et al. 2007) is impossible. This is because
-an unbiased null expectation for synchrony does not exist (Loreau and
-{{de Mazancourt}} 2008) and observed synchrony can arise from non-unique
-combinations of factors (Ranta et al. 2008). For example, weak synchrony
-of population abundances could reflect positive environmental
-correlations (synchronizing effect) offset by strong competition
-(desynchronizing effect), or negative environmental correlations and
-weak competition.
-
-The best way to quantify the effects of environmental stochasticity,
+An intuitive way to quantify the effects of environmental stochasticity,
 demographic stochasticity, and interspecific interactions is to remove
-them one-by-one, and in combination. In principle, this could be done in
-an extremely controlled laboratory setting, but empirically-based models
-of interacting populations, fit with data sets from natural communities,
+them one-by-one, and in combination.
+\textcolor{blue}{If synchrony changes more when we remove environmental stochasticity than when we remove interspecific interactions, we would conclude that environmental fluctuations are the more important driver.}
+In principle, this could be done in an extremely controlled laboratory
+setting (e.g., Venail et al. 2013), but empirically-based models of
+interacting populations, fit with data sets from natural communities,
 offer a practical alternative. For example, Mutshinda et al. (2009) fit
@@
 approach for reef fish and came to a similar conclusion: environmental
-responses determine synchrony. While a major step forward, Mutshinda et
-al.'s (2009) and Thibaut et al.'s (2012) modeling technique relied on
-abundance data that may or may not reliably capture competitive
-interactions that occur at the individual level. Furthermore, although
-both studies quantified the relative importance of environmental
-stochasticity and interspecific interactions to explain the observed
-variation of species synchrony, they did not use the model to quantify
-how much synchrony would change when each factor is removed.
-
-Here, we use multi-species population models fit to long-term
-demographic data from five semi-arid plant communities to test theory on
-the drivers of species synchrony. Our objectives are to (1) derive and
-test theoretical predictions of species synchrony and (2) determine the
-relative influence of environmental stochasticity, demographic
-stochasticity, and interspecific interactions on species synchrony in
-natural plant communities. To achieve these objectives, we first refine
-theory that has been used to predict the effects of species richness on
-ecosystem stability ({{de Mazancourt}} et al. 2013) and species
-synchrony (Loreau and {{de Mazancourt}} 2008) to generate predictions of
-community-wide species synchrony under two limiting cases derived from
-the dynamics of individual species in monoculture. We then confront our
-theoretical predictions with simulations from the empirically-based
-population models. Second, we use the multi-species population models to
-perform simulation experiments that isolate the effects of environmental
+responses determine synchrony.
+
+While a major step forward, Mutshinda et al.'s (2009) and Thibaut et
+al.'s (2012) modeling technique does have some limitations.
+\textcolor{blue}{First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change if each factor were removed.
+Second, they relied on popluation abundance data that may or may not reliably capture competitive interactions occuring at the individual level.
+Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony.
+The synchrony of species' abundances ultimately determines the stability of total community biomass, but the processes that drive species synchrony are most tightly linked to each species' immediate response to environmental conditions and competition}
+(Loreau and {{de Mazancourt}} 2008).
+\textcolor{blue}{Therefore, we focus on per capita growth rates, which represent a species' immediate response to interannual fluctuations.}
+
+Here, we use multispecies population models fit to long-term demographic
+data from five semi-arid plant communities to test theory on the drivers
+of species synchrony (Fig. 1). Our objectives are to (1) derive and test
+theoretical predictions of species synchrony and (2) determine the
+relative influence of environmental stochasticity, interspecific
+interactions, and demographic stochasticity on
+\textcolor{blue}{the synchrony of dominant species} in natural plant
+communities.
+\textcolor{blue}{While our focus is limited to dominant species due to data constraints, previous work indicates that the dynamics of dominant species determine ecosystem functioning in grasslands}
+(Smith and Knapp 2003, Bai et al. 2004, Grman et al. 2010, Sasaki and
+Lauenroth 2011).
+
+To achieve our objectives, we first refine theory that has been used to
+predict the effects of species richness on ecosystem stability ({{de
+Mazancourt}} et al. 2013) and species synchrony (Loreau and {{de
+Mazancourt}} 2008) to generate predictions of community-wide species
+synchrony under two limiting cases. We then confront our theoretical
+predictions with simulations from the empirically-based population
+models. Second, we use the multi-species population models to perform
+simulation experiments that isolate the effects of environmental
 stochasticity, demographic stochasticity, and interspecific interactions
@@
 community dynamics (Chu and Adler 2015), perturbing the models can
-reveal the processes that determine species synchrony in our focal
-grassland communities.
-
-\subsection{THEORETICAL MODEL}\label{theoretical-model}
-
-\subsubsection{The model}\label{the-model}
-
-While existing theory has identified the factors driving synchrony, we
-do not have a simple expression to predict synchrony in a particular
+reveal the processes that determine synchrony
+\textcolor{blue}{of dominant species} in our focal grassland
+communities.
+
+\section{Theoretical Model}\subsection{The model}
+
+While existing theory has identified
+\textcolor{blue}{environmental responses, species interactions, and demographic stochasticity as the drivers of the temporal synchrony,}
+we do not have a simple expression to predict synchrony in a particular
 community with all factors operating simultaneously. However, we can
@@
 lead to the same overall conclusions as synchrony of per capita growth
-rates. Following Loreau and {{de Mazancourt}} (2008) and {{de
-Mazancourt}} et al. (2013), we define population growth, ignoring
-observation error, as
+rates.
+
+Following Loreau and {{de Mazancourt}} (2008) and {{de Mazancourt}} et
+al. (2013), we define population growth, ignoring observation error, as
 
@@
 \emph{t}, and \(r_i(t)\) is its population growth rate in year \emph{t}.
-\(r_{mi}\) is species \emph{i}`s intrinsic rate of increase, \(K_i\) is
+\(r_{mi}\) is species \emph{i}'s intrinsic rate of increase, \(K_i\) is
 its carrying capacity, and \(\alpha_{ij}\) is the interspecific
@@
 species \emph{i}. Environmental stochasticity is incorporated as
-\(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the environmental
-variance and \(u_{ei}\) are normal random variables with zero mean and
-unit variance that are independent through time but may be correlated
-between species. Demographic stochasticity arises from variations in
-births and deaths among individuals (e.g., same states, different
-fates), and is included in the model as a first-order, normal
-approximation (Lande et al. 2003, {{de Mazancourt}} et al. 2013).
-\(\sigma_{di}^2\) is the demographic variance and \(u_{di}(t)\) are
-independent normal variables with zero mean and unit variance. To derive
-analytical predictions we solved a first-order approximation of Equation
-2 ({{de Mazancourt}} et al. 2013 and Appendix S1). Due to the linear
-approximation approach, our analytical predictions will likely fail in
-communities where species exhibit large fluctuations due to limit cycles
-and chaos (Loreau and {{de Mazancourt}} 2008). Indeed, one of the
-advantages of focusing on growth rates rather than abundances is that
-growth rates are more likely to be well-regulated around an equilibrium
-value, if the long-term average of a species' growth rate is relatively
-small (e.g., \(r < 2\)).
-
-\subsubsection{Predictions}\label{predictions}
+\(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the
+\textcolor{blue}{temporal variance of species \emph{i}'s response to the environment and $u_{ei}(t)$ is a normal variable with unit variance that is independent among species but may be correlated.}
+\textcolor{blue}{The product, $\sigma_{ei}u_{ei}(t)$, is species \emph{i}'s environmental response.}
+Demographic stochasticity arises from variations in births and deaths
+among individuals (e.g., same states, different fates), and is included
+in the model as a first-order, normal approximation (Lande et al. 2003,
+{{de Mazancourt}} et al. 2013). \(\sigma_{di}^2\) is the demographic
+variance
+(\textcolor{blue}{i.e., the intrinsic demographic stochasticity of species \emph{i}})
+and \(u_{di}(t)\) are independent normal variables with zero mean and
+unit variance
+\textcolor{blue}{that allow demographic stochasiticity to vary through time.}
+
+To derive analytical predictions we solved a first-order approximation
+of Equation 2 ({{de Mazancourt}} et al. 2013 and Appendix S1). Due to
+the linear approximation approach, our analytical predictions will
+likely fail in communities where species exhibit large fluctuations due
+to limit cycles and chaos (Loreau and {{de Mazancourt}} 2008). Indeed,
+one of the advantages of focusing on growth rates rather than abundances
+is that growth rates are more likely to be well-regulated around an
+equilibrium value, if the long-term average of a species' growth rate is
+relatively small (e.g., \(r < 2\)).
+
+\subsection{Predictions}
 
@@
 covariance of environmental responses between species \emph{i} and
-species \emph{j}. Confronting our theoretical predictions with data
-requires estimates of species dynamics of large populations (no
-demographic stochasticity) growing in isolation (no interspecific
-interactions) to calculate the covariance of species' environmental
-responses. To estimate environmental responses in natural communities,
-we turn to our population models built using long-term demographic data.
-
-\subsection{EMPIRICAL ANALYSIS}\label{empirical-analysis}
-
-\subsubsection{Materials and methods}\label{materials-and-methods}
-
-\paragraph{Data}\label{data}
+species \emph{j}.
+
+Confronting our theoretical predictions with data requires estimates of
+species dynamics of large populations (no demographic stochasticity)
+growing in isolation (no interspecific interactions) to calculate the
+covariance of species' environmental responses. To estimate
+environmental responses in natural communities, we turn to our
+population models built using long-term demographic data.
+
+\section{Empirical Analysis}\subsection{Materials and methods}
+
+\paragraph{Data.--}\label{data.}
 
@@
 \paragraph{Calculating observed
-synchrony}\label{calculating-observed-synchrony}
+synchrony.--}\label{calculating-observed-synchrony.}
 
@@
 
-\noindent where \(\sigma_{r_{i}}\) is the temporal variance of species
-\emph{i'}s per capita population growth rate (\(r_i\)) and
-\(\sigma^{2}_{T}\) is the temporal variance of the aggregate
-community-level growth rate. \(\phi\) ranges from 0 at perfect
-asynchrony to 1 at perfect synchrony (Loreau and {{de Mazancourt}}
-2008). We use the same equation to calculate observed synchrony of
-species' percent cover, which we present to relate our results to
-previous findings, even though we focus on synchrony of growth rates in
-our model simulations (see below).
-
-\paragraph{Fitting statistical models}\label{fitting-statistical-models}
+\noindent where \(\sigma_{r_{i}}\) is the temporal
+\textcolor{blue}{standard deviation} of species \emph{i'}s per capita
+population growth rate (\(r_i\)) and \(\sigma^{2}_{T}\) is the temporal
+variance of the aggregate community-level growth rate. \(\phi\) ranges
+from 0 at perfect asynchrony to 1 at perfect synchrony (Loreau and {{de
+Mazancourt}} 2008). We use the same equation to calculate observed
+synchrony of species' percent cover, which we present to relate our
+results to previous findings, even though we focus on synchrony of
+growth rates in our model simulations (see below).
+
+\paragraph{Fitting statistical
+models.--}\label{fitting-statistical-models.}
 
@@
 via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
-2013). Crowding was treated as a fixed effect without a temporal
-component because most 95\% credible intervals for random year effects
-on crowding broadly overlapped zero and, in a test case, including
-yearly crowding effects did not change our results. Spatial (quadrat
-groupings) variation was treated as a random effect on the intercept and
-temporal (interannual) variation was treated as random effects on the
-intercept and the effect of genet size in the previous year (Appendix
-S1).
+2013). Spatial (quadrat groupings) variation was treated as a random
+effect on the intercept and temporal (interannual) variation was treated
+as random effects on the intercept and the effect of genet size in the
+previous year (Appendix S1).
+
+\textcolor{blue}{Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
+Strong year-to-year variation in these crowding effects would indicate a statistical interaction between environmental conditions and species interactions.
+We tested for such a dynamic by comparing models with and without random year effects on crowding.
+Based on the results, we decided to treat crowding as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including interannual variation in crowding did not change our results.}
 
@@
 \paragraph{Building dynamic multi-species
-models}\label{building-dynamic-multi-species-models}
+models.--}\label{building-dynamic-multi-species-models.}
 
@@
 randomly in space. Crowding is directly calculated at each time step
-since each genet is spatially referenced (as in the observed data).
-Environmental stochasticity is not an inherent feature of IBMs, but is
-easily included since we fit year-specific temporal random effects for
-each vital rate regression. To include temporal environmental variation,
-at each time step we randomly choose a set of estimated survival,
-growth, and recruitment parameters specific to one observation year. For
-all simulations, we ignore the spatial random effect associated with
-variation among quadrat groups, so our simulations represent an average
-quadrat for each site.
-
-The IPM uses the same vital rate regressions as the IBM (Rees and Ellner
-2009, Rees et al. 2014), but it is spatially implicit and does not
-include demographic stochasticity. Following Chu and Adler (2015), we
-use a mean field approximation that captures the essential features of
-spatial patterning to define the crowding index at each time step
-(Supporting Online Information). Temporal variation is included in
-exactly the same way as for the IBM. For full details on the IPMs we
-use, see Chu and Adler (2015).
-
-\paragraph{Simulation experiments}\label{simulation-experiments}
+since each genet is spatially referenced. Environmental stochasticity is
+not an inherent feature of IBMs, but is easily included since we fit
+year-specific temporal random effects for each vital rate regression. To
+include temporal environmental variation, at each time step we randomly
+choose a set of estimated survival, growth, and recruitment parameters
+specific to one observation year. For all simulations, we ignore the
+spatial random effect associated with variation among quadrat groups, so
+our simulations represent an average quadrat for each site.
+
+The IPM uses the same vital rate regressions as the IBM, but it is
+spatially implicit and does not include demographic stochasticity.
+Following Chu and Adler (2015), we use a mean field approximation that
+captures the essential features of spatial patterning to define the
+crowding index at each time step (Supporting Online Information).
+Temporal variation is included in exactly the same way as for the IBM.
+For full details on the IPMs we use, see Chu and Adler (2015).
+
+\paragraph{Simulation experiments.--}\label{simulation-experiments.}
 
@@
 demographic stochasticity by simulating the IBM on areas (e.g.~plots) of
-different size. Indeed, results from an IBM with infinite population
-size would converge on results from the IPM. Given computational
-constraints, the largest landscape we simulate is a 25 \(\text{m}^2\)
-plot.
+different size. Results from an IBM with infinite population size would
+converge on results from the IPM. Given computational constraints, the
+largest landscape we simulate is a 25 \(\text{m}^2\) plot.
 
@@
 and demographic stochasticity removed; (6) IBM: Interspecific
-interactions and environmental stochasticity removed. We ran IPM
-simulations for 2,000 time steps, after an initial 500 iteration burn-in
-period. This allowed species time to reach their stable size
-distribution. We then calculated the synchrony of species' per capita
-growth rates over 100 randomly selected contiguous 50 time-step
-sections. We ran IBM simulations for 100 time steps, and repeated the
-simulations 100 times for each simulation experiment. From those, we
-retained only the simulations in which no species went extinct due to
-demographic stochasticity. Synchrony was calculated over the 100 time
-steps for each no extinction run within a model experiment. To explore
-the effect of demographic stochasticity in different sized populations,
-we ran simulations (1) and (6) on plot sizes of 1, 4, 9, 16, and 25
-\(\text{m}^2\). All other IBM simulations were run on a 25
-\(\text{m}^2\) landscape.
-
-Results from our simulation experiments also allow us to test our
-theoretical predictions. First, in the absence of interspecific
-interactions and demographic stochasticity, populations can only
-fluctuate in response to the environment. Therefore, we can use results
-from simulation (5) to estimate the covariance of species' responses to
-the environment (\(\text{cov}(u_{ie}, u_{je})\)) and parameterize
-Equation 4. Parameterizing Equation 3 does not require simulation output
-because the only parameters are the species' relative abundances.
-Second, simulations (5) and (6) represent the simulated version of our
-limiting case theoretical predictions. Thus, we directly test the
-theoretical predictions by comparing them to observed synchrony and
-simulated synchrony.
-
-\subsubsection{Results}\label{results}
-
-Synchrony of species' per capita growth rates at our study sites range
-from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to
-0.92 (Table 2). Synchrony tends to be higher in communities with few
-species (Arizona and New Mexico) and/or with relatively high temporal
-variability in percent cover (e.g., Montana). Synchrony is lowest in
-Idaho, the only data set that includes two life forms: a shrub and three
-perennial grasses (Table 1). Synchrony of per capita growth rates and CV
-of percent cover are positively correlated (Pearson's \(\rho\) = 0.72).
-For all five communities, species synchrony from IPM and IBM simulations
-closely approximated observed synchrony (Fig. S1). IBM-simulated
-synchrony is consistently, but only slightly, lower than IPM-simulated
-synchrony (Fig. S1), likely due to the desynchronizing effect of
-demographic stochasticity.
+interactions and environmental stochasticity removed.
+\textcolor{blue}{We did not include a simulation with only interspecific interactions because our population models run to deterministic equilibriums in the absence of environmental or demographic stochasticity.}
+We ran IPM simulations for 2,000 time steps, after an initial 500
+iteration burn-in period. This allowed species to reach their stable
+size distribution. We then calculated yearly per capita growth rates
+from the simlated time series, and then caluclated the synchrony of
+species' per capita growth rates over 100 randomly selected contiguous
+50 time-step sections.
+
+We ran IBM simulations for 100 time steps, and repeated the simulations
+100 times for each simulation experiment. From those, we retained only
+the simulations in which no species went extinct due to demographic
+stochasticity. Synchrony of per capita growth rates was calculated over
+the 100 time steps for each no extinction run within a model experiment.
+\textcolor{blue}{In the IBM simulations, the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
+To explore this effect}, we ran simulations \textcolor{blue}{(4)} and
+(6) on plot sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM
+simulations were run on a 25 \(\text{m}^2\) landscape
+\textcolor{blue}{to most closely match the implicit, large spatial scale of the IPM simulations}.
+
+\textcolor{blue}{Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulated values of community-wide species synchrony.}
+The simulation experiments also allow us to test our theoretical
+predictions. First, in the absence of interspecific interactions and
+demographic stochasticity, populations can only fluctuate in response to
+the environment. Therefore, we can use results from simulation (5) to
+estimate the covariance of species' responses to the environment
+(\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation 4.
+Parameterizing Equation 3 does not require simulation output because the
+only parameters are the species' relative abundances. Second,
+simulations (5) and (6) represent the simulated version of our limiting
+case theoretical predictions.
+\textcolor{blue}{This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalence in species' growth rates, environmental response variances, and demographic variances.
+Our empirically-based models do not make these assumptions.
+Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions hold in the natural community our model represents.}
+
+\subsection{Results}
+
+\textcolor{blue}{Observed} synchrony of species' per capita growth rates
+at our study sites range from 0.36 to 0.89 and synchrony of percent
+cover ranged from 0.15 to 0.92 (Table 2). Synchrony of per capita growth
+rates and CV of percent cover are positively correlated (Pearson's
+\(\rho\) = 0.72, \emph{N} = 5). For all five communities, species
+synchrony from IPM and IBM simulations closely approximated observed
+synchrony (Fig. S1). IBM-simulated synchrony is consistently, but only
+slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to
+the desynchronizing effect of demographic stochasticity.
 
 Across the five communities, our limiting case predictions closely
-matched synchrony from the corresponding simulation experiment (Fig. 1
+matched synchrony from the corresponding simulation experiment (Fig. 2
 and Table S1). The correlation between our analytical predictions and
-simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) and 0.997
-for \(\phi_{R,\mathcal{M}_E}\). The largest difference between predicted
-and simulated synchrony was 0.05 in New Mexico for
-\(\phi_{R,\mathcal{M}_D}\) (Table S1).
+simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) (\emph{N} =
+5) and 0.997 for \(\phi_{R,\mathcal{M}_E}\) (\emph{N} = 5). The largest
+difference between predicted and simulated synchrony was 0.05 in New
+Mexico for \(\phi_{R,\mathcal{M}_D}\) (Table S1).
 
@@
 has the largest impact on synchrony, leading to a reduction in synchrony
-of species growth rates in four out of five communities (Fig. 1).
+of species growth rates in four out of five communities (Fig. 2).
 Removing environmental fluctuations (``No E.S'' simulations) decreased
@@
 in New Mexico. Only in Idaho did removing environmental fluctuations
-cause an increase in synchrony (Fig. 1), but the effect was small (9\%
+cause an increase in synchrony (Fig. 2), but the effect was small (9\%
 increase; Table S2). Overall, species' temporal random effects in the
-statistical vital rate models are positively correlated (Table S3).
+statistical vital rate models are positively, but not perfectly,
+correlated (Table S3).
+\textcolor{blue}{These temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.}
+
 Species interactions are weak in these communities (Table S4 and Chu and
 Adler 2015), and removing interspecific interactions had little effect
-on synchrony (Fig. 1; ``No Comp.'' simulations). Removing interspecific
-interactions caused, at most, a 5\% change in synchrony (Fig. 1).
-Removing demographic stochasticity (``No D.S.'' simulations) caused
-synchrony to increase slightly in all communities (Fig. 1), with an
-average 6\% increase over synchrony from IBM simulations on a
+on synchrony (Fig. 2; ``No Comp.'' simulations). Removing interspecific
+interactions caused, at most, a 5\% change in synchrony (Fig. 2 and
+Table S2). Removing demographic stochasticity (``No D.S.'' simulations)
+caused synchrony to increase slightly in all communities (Fig. 2), with
+an average 6\% increase over synchrony from IBM simulations on a
 25\(\text{m}^2\) area.
 
-The desynchronizing effect of demographic stochasticity, which increases
-as population size decreases, modestly counteracted the synchronizing
-force of the environment, but not enough to lower synchrony to the level
-observed when only demographic stochasticity is operating (Fig. 2). In
-the largest, 25 \(\text{m}^2\) plots, synchrony was driven by
-environmental stochasticity (e.g., \(\mathcal{M}_E\)). At 1
-\(\text{m}^2\), synchrony reflected demographic stochasticity and
-environmental stochasticity (e.g., between \(\mathcal{M}_E\) and
+In IBM simulations, the desynchronizing effect of demographic
+stochasticity, which increases as population size decreases, modestly
+counteracted the synchronizing force of the environment, but not enough
+to lower synchrony to the level observed when only demographic
+stochasticity is operating (Fig. 3). In the largest, 25 \(\text{m}^2\)
+plots, synchrony was driven by environmental stochasticity (e.g.,
+\(\mathcal{M}_E\)). At 1 \(\text{m}^2\), synchrony reflected
+\textcolor{blue}{the combined effects of} demographic stochasticity and
+environmental stochasticity (e.g.,
+\textcolor{blue}{simulated synhcrony fell} between \(\mathcal{M}_E\) and
 \(\mathcal{M}_D\)). For context, population sizes increased from an
@@
 
-For all five communities, the synchrony of species' growth rates when
-growing in isolation almost perfectly matched species synchrony in
-polyculture (Fig. 3). Results for synchrony of percent cover are
-qualitatively similar, but simulation results were more variable and
-less consistent with analytical predictions and observed synchrony
-(Appendix S1, Figs. S2-S3).
-
-\subsection{DISCUSSION}\label{discussion}
+Results for synchrony of percent cover are qualitatively similar, but
+more variable and less consistent with analytical predictions and
+observed synchrony (Appendix S1, Figs. S2-S3).
+
+\section{Discussion}
 
@@
 
-\subsubsection{Simulations support theoretical
-predictions}\label{simulations-support-theoretical-predictions}
+\subsection{Simulations support theoretical predictions}
 
 Our theoretical predictions were derived from a simple model of
-population dynamics and required several simplifying assumptions,
+population dynamics and required several restrictive assumptions,
 raising questions about their relevance to natural communities. For
@@
 results from simulations of population models fit to long-term data from
-natural plant communities (Table 3). Such strong agreement between our
+natural plant communities (Table S1). Strong agreement between our
 analytical predictions and the simulation results should inspire
@@
 theory.
-
-\subsubsection{Demographic stochasticity decreases
-synchrony}\label{demographic-stochasticity-decreases-synchrony}
-
-Demographic stochasticity partially counteracted the synchronizing
-effects of environmental fluctuations and interspecific interactions on
-per capita growth rates, but only when populations were small (Fig. 2),
-in agreement with theory (Loreau and {{de Mazancourt}} 2008). Even in
-small populations, however, demographic stochasticity was not strong
-enough to compensate the synchronizing effects of environmental
-fluctuations and match the analytical prediction where only demographic
-stochasticity is operating (\(\mathcal{M}_D\) in Fig. 2). These results
-confirm the theoretical argument by Loreau and {{de Mazancourt}} (2008)
-that independent fluctuations among interacting species in a
-non-constant environment should be rare. Only in the Idaho community
-does synchrony of per capita growth rates approach \(\mathcal{M}_D\) in
-a non-constant environment (Fig. 2). This is most likely due to the
-strong effect of demographic stochasticity on the shrub \emph{Artemisia
-tripartita} since even a 25 \(\text{m}^2\) quadrat would only contain a
-few individuals of that species.
+\textcolor{blue}{However, whether the theoretical model adequately represents more complex communities remains unknown because our analysis was restricted to dominant species.}
+
+\subsection{Demographic stochasticity decreases synchrony \textcolor{blue}{in small populations}}
+
+In large populations, removing demographic stochasticity had no effect
+on species synchrony (Fig. 2). In small populations, demographic
+stochasticity partially counteracted the synchronizing effects of
+environmental fluctuations and interspecific interactions on per capita
+growth rates, in agreement with theory (Loreau and {{de Mazancourt}}
+2008).
+\textcolor{blue}{This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
+The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, reducing the relative effect of environmental forcing.}
+Even in small populations
+\textcolor{blue}{(e.g., 1 $\text{m}^2$ lanscapes)}, however, demographic
+stochasticity was not strong enough to compensate the synchronizing
+effects of environmental fluctuations and match the analytical
+prediction where only demographic stochasticity is operating
+(\(\mathcal{M}_D\) in Fig. 3). These results confirm the theoretical
+argument of Loreau and {{de Mazancourt}} (2008) that independent
+fluctuations among interacting species in a non-constant environment
+should be rare. Only in the Idaho community does synchrony of per capita
+growth rates approach \(\mathcal{M}_D\) in a non-constant environment
+(Fig. 3). This is most likely due to the strong effect of demographic
+stochasticity on the shrub \emph{Artemisia tripartita} since even a 25
+\(\text{m}^2\) quadrat would only contain a few individuals of that
+species.
 
@@
 observation area increases, population size increases and the
-desynchronizing effect of demographic stochasticity lessens (Fig. 2).
+desynchronizing effect of demographic stochasticity lessens (Fig. 3).
 Thus, our results suggest that community-wide species synchrony will
@@
 
-\subsubsection{Environmental fluctuations drive community
-synchrony}\label{environmental-fluctuations-drive-community-synchrony}
+\subsection{Environmental fluctuations drive community synchrony}
 
@@
 discernible change in community-wide species synchrony of per capita
-growth rates (Fig. 1). Second, removing environmental fluctuations from
-simulations consistently reduced synchrony (Fig. 1), and the synchrony
-of species in isolation was a very strong predictor of synchrony of
-species in polyculture (Fig. 3). Our results lead us to conclude that
-environmental fluctuations, not species interactions, are the primary
-driver of community-wide species synchrony in the communities we
-studied. Given accumulating evidence that niche differences in natural
-communities are large (reviewed in Chu and Adler 2015), and thus species
-interactions are likely to be weak, our results may be general in
-natural plant communities.
+growth rates (Fig. 2). Second, removing environmental fluctuations from
+simulations consistently reduced synchrony (Fig. 2). Our results lead us
+to conclude that environmental fluctuations, not species interactions,
+are the primary driver of community-wide species synchrony
+\textcolor{blue}{among the dominant species} we studied. Given
+accumulating evidence that niche differences in natural communities are
+large (reviewed in Chu and Adler 2015), and thus species interactions
+are likely to be weak, our results may be general in natural plant
+communities.
 
@@
 
-A challenge to the generality of our results is that we were only able
-to model common, co-occuring species (see Chu and Adler 2015). Most
-communities are dominated by few common species and many rare species
-(McGill et al. 2007), meaning that the low number of common species in
-our focal commiunities is not unusual. Rather, the generality of our
-results hinges upon the influence of rare species. Rare species could be
-kept at low densities by competitive suppression due to strong
-interspecific interactions. If we had been able to model rare species,
-we might have found a larger effect of competition on community
-synchrony. However, recent theory predicts that persistent rare species
-may actually experience even weaker interspecific competition, and thus
-stronger niche differences, than common species (Yenni et al. 2012), in
-which case synchrony would remain predominantly driven by environmental
-responses. Rare species could also be limited by generalist natural
-enemies whose effects are density-independent but alter competitive
-hierarchies (Mordecai 2011). Under such conditions, synchrony will
-depend on the degree of pathogen-induced fitness differences and the
-pathogen's response to environmental conditions. Neither our model nor
-current theory explicitly considers the effects of pathogens on species
-synchrony, and this highlights the need for theoretical work on the
-interaction between mechanisms of species coexistence and mechanisms of
-species synchrony (Loreau 2010).
-
-\subsubsection{Interspecific interactions had little impact on community
-synchrony}\label{interspecific-interactions-had-little-impact-on-community-synchrony}
+\textcolor{blue}{Environmental responses synchronized dynamics relative to a null expectation of independent species interactions (e.g., "No Comp.
++ No E.S." simulations in Fig.
+2), but observed and simulated synchrony was still less than one in all cases (Fig.
+2).
+Synchrony was far from complete because of differences in species' responses to interannual environmental variation.
+Many studies of ecosystem stability in semiarid grasslands focus on trad-offs among dominant functional types}
+(Bai et al. 2004, Sasaki and Lauenroth 2011).
+\textcolor{blue}{Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
+At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were positively correlated.
+However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 2).
+The subtle differences among ecologically-similar dominant species do impact species synchrony and, ultimately, ecosystem stability.
+Ignoring such differences could mask important dynamics that underpin ecosystem functioning.}
+
+\subsection{Interspecific interactions had little impact on community synchrony}
 
@@
 niche overlap (Chu and Adler 2015) and weak interspecific interactions
-(Tables S1, S3-S17). Mechanistic consumer-resource models (Lehman and
+(Tables S1 and S3-S17). Mechanistic consumer-resource models (Lehman and
 Tilman 2000) and phenomenological Lotka-Volterra models (Lehman and
@@
 
-However, our results do contrast with a recent analysis of several
-biodiversity-ecosystem functioning experiments showing that competition
-drives species synchrony in grasslands (Gross et al. 2014). The apparent
-inconsistency between our results and those of Gross et al. (2014) may
-be explained by the differences between our studies. Gross et al.'s
-results are based on rapidly assembling experimental communities that
-are, at most, 11 years old. The natural, relatively undisturbed
-communities we studied consist of species that have co-occurred for
-decades (Chu and Adler 2015) and represent a much later stage of
-community assembly. In theory, species interactions should weaken
-through time as community assembly proceeds (Kokkoris et al. 1999),
-meaning that in biodiversity-ecosystem functioning experiments the
-processes driving competitive exclusion are still operating. Such
-communities may be highly influenced by transient, but strong,
-interspecific competition that can mask the synchronizing effects of the
-environment. In contrast, synchrony may be driven by environmental
-fluctuations in older communities. In other words, the dominant driver
-of community synchrony should shift from competition to environmental
-fluctuations through time. One way to test this prediction is to
-continue collecting data from biodiversity-ecoystem functioning
-experiments and conduct the analysis of Gross et al. (2014) every few
-years.
-
 Our conclusion that species interactions have little impact on synchrony
@@
 
-\subsection{CONCLUSIONS}\label{conclusions}
+\section{Conclusions}
 
@@
 studied do have similar responses to the environment (Table S2), which
-will tend to synchronize dynamics. However, if community-wide species
-synchrony is less than 1, as it is in all our focal communities, some
-degree of compensatory dynamics must be present (Loreau and {{de
-Mazancourt}} 2008). In agreement with other studies (Rocha et al. 2011,
-Vasseur et al. 2014), we find that environmental responses are primarily
-responsible for the degree of synchrony among ecologically-similar
-species. This result contrasts with a recent analysis of several
-biodiversity-ecosystem functioning experiments showing that competition
-drives community synchrony (Gross et al. 2014). Recently assembled
-communities, such as experimental plots, may exhibit strong species
-interactions that will weaken over time (Kokkoris et al. 1999). Future
-research on the influence of community assembly on synchrony could
-reconcile inconsistent observations on the biotic and abiotic forces
-that jointly drive community dynamics.
-
-\subsection{ACKNOWLEDGMENTS}\label{acknowledgments}
+will tend to synchronize dynamics. However, if community-wide synchrony
+\textcolor{blue}{among dominant species} is less than perfect, as it is
+in all our focal communities, some degree of compensatory dynamics must
+be present (Loreau and {{de Mazancourt}} 2008).
+\textcolor{blue}{Even ecologically-similar species, which are sometimes aggregated into functional groups, have environmental responses that are dissimilar enough to limit synchrony.
+Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.}
+
+\section{Acknowledgements}
 
@@
 the Office of Research and Graduate Studies at Utah State University are
-gratefully acknowledged. We also thank three anonymous reviewers who
-provided critical feedback that improved the manuscript.
+gratefully acknowledged. We also thank Patrick Venail and several
+anonymous reviewers who provided critical feedback that improved the
+manuscript.
 
@@
 
-\subsection{Tables}\label{tables}
+\section{Tables}
 
@@
 
-\subsection{Figures}\label{figures}
+\section{Figures}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/synchrony_flowchart_bw.png}
+  \caption{Diagram of our coupled theoretical-empirical approach. We followed this workflow for each of our five focal communities.}
+\end{figure}
+
+\pagebreak{}
 
@@
 All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
-Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.}
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
 \end{figure}
@@
       \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
-  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
-Only'') and when environmental stochasticity is also present removed (``D.S.
-+ E.S.'').
+  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
++ No E.S.
+(D.S.
+Only)'') and when environmental stochasticity is also present (``No Comp.
+(D.S.
++ E.S.)'').
 The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
 The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
-Theoretically, ``D.S.
-Only'' simulations should remain constant across landscape size, whereas ``D.S.
-+ E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
+Theoretically, ``No Comp.
++ No E.S.
+(D.S.
+Only)'' simulations should remain constant across landscape size, whereas ``No Comp.
+(D.S.
++ E.S.)'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
 Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
-\end{figure}
-
-\pagebreak{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/formatted_figures/formatted_figure3.png}
-  \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).
-We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.
-The dashed line is the line of equality.
-Simulation results in this figure are analagous to ``No Comp.
-+ No D.S.'' (species in isolation) and ``No D.S.'' (species in competition) in Fig.
-1, but here we control the time series of random year effects.}
 \end{figure}
@@
 92:1703.
+
+Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
+and compensatory effects in the Inner Mongolia grassland. Nature
+431:181--184.
 
@@
 175:E16--E34.
+
+Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
+Mechanisms contributing to stability in ecosystem function depend on the
+environmental context. Ecology Letters 13:1400--1410.
 
@@
 
-Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
-species interaction strength in assembled theoretical competition
-communities. Ecology Letters 2:70--74.
-
 Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
@@
 
-Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
-for a New Ecological Synthesis.
-
 Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
@@
 16:106--115.
+
+Loreau, M., and a Hector. 2001. Partitioning selection and
+complementarity in biodiversity experiments. Nature 412:72--6.
 
@@
 
-McGill, B. J., R. S. Etienne, J. S. Gray, D. Alonso, M. J. Anderson, H.
-K. Benecha, M. Dornelas, B. J. Enquist, J. L. Green, F. He, A. H.
-Hurlbert, A. E. Magurran, P. A. Marquet, B. A. Maurer, A. Ostling, C. U.
-Soykan, K. I. Ugland, and E. P. White. 2007. Species abundance
-distributions: Moving beyond single prediction theories to integration
-within an ecological framework. Ecology Letters 10:995--1015.
-
-Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
-theory, concepts, and empirical work. Ecological Monographs 81:429--441.
-
 Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
@@
 under environmental forcing. Oikos 117:1907--1911.
-
-Rees, M., and S. P. Ellner. 2009. Integral projection models for
-populations in temporally varying environments. Ecological Monographs
-79:575--594.
-
-Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
-projection models: A user's guide.
-
-Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
-species have similar dynamics. Journal of Ecology 99:1453--1459.
 
@@
 
+Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
+diversity, regulates temporal stability of plant communities. Oecologia
+166:761--768.
+
+Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
+function with non-random species loss. Ecology Letters 6:509--517.
+
 Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
@@
 Science 294:843--845.
-
-Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
-Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
-W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
-2014. Synchronous dynamics of zooplankton competitors prevail in
-temperate lake ecosystems. Proceedings. Biological sciences / The Royal
-Society 281:20140633.
 
@@
 
+Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
+2013. Shared ancestry influences community stability by altering
+competitive interactions: evidence from a laboratory microcosm
+experiment using freshwater green algae. Proceedings of the Royal
+Society of London B: Biological Sciences 280.
+
 Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
@@
 
-Yenni, G., P. B. Adler, and S. K. {Morgan Ernest}. 2012. Strong
-self-limitation promotes the persistence of rare species. Ecology
-93:456--461.
-
 Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in

### 2017-01-25 docs/ms_communitySynchrony_ForECOLOGY.tex
final version of docs for Ecology

--- 2016-11-22 docs/ms_communitySynchrony.tex
@@
 
-\usepackage{lineno}
-\linenumbers
+%%\usepackage{lineno}
+%%\linenumbers
 \usepackage{setspace}
@@
 
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+%%\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 
@@
 \begin{center}
-\large{\textbf{Environmental responses, not species interactions, determine \textcolor{blue}{synchrony of dominant species in semiarid grasslands}}}
+\large{\textbf{Environmental responses, not species interactions, determine synchrony of dominant species in semiarid grasslands}}
 
@@
 
-\textbf{Running Head}: Drivers of species synchrony in \textcolor{blue}{grasslands}
+\textbf{Running Head}: Species synchrony in grasslands
 
@@
 
+\renewcommand{\abstractname}{\vspace{-\baselineskip}}
+
 \begin{abstract}
-Temporal asynchrony among species helps diversity to stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+\normalsize
+\emph{Abstract.}\hspace{1em}Temporal asynchrony among species helps diversity to stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
 Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific interactions, and demographic stochasticity.
-We then conduct simulation experiments with empirical population models to quantify the relative \textcolor{blue}{influence} of these factors \textcolor{blue}{on the synchrony of dominant species in five semiarid grasslands.}
+We then conduct simulation experiments with empirical population models to quantify the relative influence of these factors on the synchrony of dominant species in five semiarid grasslands.
 We found that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
 Removing interspecific interactions and demographic stochasticity had small effects on synchrony. 
-\textcolor{blue}{For the dominant species in these plant communities}, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+For the dominant species in these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
 
@@
 2010).
-\textcolor{blue}{The stabilizing effect of species richness arises from a combination of selection effects and complementarity}
-(Loreau and Hector 2001).
-\textcolor{blue}{Selection effects occur when a dominant species has lower than average temporal variability, which generates a positive effect on ecosystem stability}
-(e.g., Grman et al. 2010).
-\textcolor{blue}{Complementarity occurs when species have unique temporal dynamics, causing their abundances to fluctuate asynchronously, increasing ecosystem stability.}
-\textcolor{blue}{The premise of this paper is that understanding the mechanisms driving species' temporal dynamics, and resulting (a)synchrony, is necessary to predict the impacts of global change on ecosystem stability.}
-
-\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics}
-(Gonzalez and Loreau 2009),
-\textcolor{blue}{occur whenever species synchrony is not perfect} and
+The stabilizing effect of species richness arises from a combination of selection effects and complementarity (Loreau and Hector 2001).
+Selection effects occur when a dominant species has lower than average temporal variability, which generates a positive effect on ecosystem stability (e.g., Grman et al. 2010).
+Complementarity occurs when species have unique temporal dynamics, causing their abundances to fluctuate asynchronously, increasing ecosystem stability.
+The premise of this paper is that understanding the mechanisms driving species' temporal dynamics, and resulting (a)synchrony, is necessary to predict the impacts of global change on ecosystem stability.
+
+Asynchronous dynamics, also known as compensatory dynamics (Gonzalez and Loreau 2009),
+occur whenever species synchrony is not perfect and
 result from individual species responding in different ways to
@@
 et al. 2013, Gross et al. 2014). Species richness affects
-\textcolor{blue}{the degree of synchrony in a community} because larger
+the degree of synchrony in a community because larger
 species pools are more likely to contain species that respond
@@
 increasing stability (Yachi and Loreau 1999).
-\textcolor{blue}{Species richness can also affect synchrony if the strength of species interactions varies systematically with richness, because competition generally increases synchrony and reduces stability}
+Species richness can also affect synchrony if the strength of species interactions varies systematically with richness, because competition generally increases synchrony and reduces stability
 (Loreau and {{de Mazancourt}} 2013, but see Gross et al. 2014).
 
-\textcolor{blue}{The effects of environmental change and species losses on ecosystem stability will depend on whether synchrony is driven by species-specific responses to environmental conditions or interspecific competition}
+The effects of environmental change and species losses on ecosystem stability will depend on whether synchrony is driven by species-specific responses to environmental conditions or interspecific competition
 (Hautier et al. 2014).
-\textcolor{blue}{If responses to environment are important, then environmental change could alter synchrony and stability.
-If competition is important, then the direct effects of environmental change may not affect synchrony and stability, but species gains/losses will.}
-
-\textcolor{blue}{The relative role of environmental responses and competition in driving synchrony  in natural plant communities}
+If responses to environment are important, then environmental change could alter synchrony and stability.
+If competition is important, then the direct effects of environmental change may not affect synchrony and stability, but species gains/losses will.
+
+The relative role of environmental responses and competition in driving synchrony  in natural plant communities
 remains controversial (reviewed in Gonzalez and Loreau 2009). One source
@@
 
-\textcolor{blue}{Theory can help us resolve this empirical question.}
+Theory can help us resolve this empirical question.
 Recent theoretical work has identified three determinants of species
@@
 Gonzalez and Loreau 2009).
-\textcolor{blue}{This theory has been developed by focusing on simple limiting cases in which only one of these three drivers operates.}
+This theory has been developed by focusing on simple limiting cases in which only one of these three drivers operates.
 For example, in a community composed of large populations (no
@@
 2008). However, this prediction relies on a relatively simple population
-model and \textcolor{blue}{depends on two restrictive} assumptions: (i)
+model and depends on two restrictive assumptions: (i)
 species' responses to the environment are similar in magnitude and (ii)
@@
 them one-by-one, and in combination.
-\textcolor{blue}{If synchrony changes more when we remove environmental stochasticity than when we remove interspecific interactions, we would conclude that environmental fluctuations are the more important driver.}
+If synchrony changes more when we remove environmental stochasticity than when we remove interspecific interactions, we would conclude that environmental fluctuations are the more important driver.
 In principle, this could be done in an extremely controlled laboratory
@@
 al.'s (2012) modeling technique does have some limitations.
-\textcolor{blue}{First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change if each factor were removed.
-Second, they relied on popluation abundance data that may or may not reliably capture competitive interactions occuring at the individual level.
+First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change if each factor were removed.
+Second, they relied on population abundance data that may or may not reliably capture competitive interactions occurring at the individual level.
 Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony.
-The synchrony of species' abundances ultimately determines the stability of total community biomass, but the processes that drive species synchrony are most tightly linked to each species' immediate response to environmental conditions and competition}
+The synchrony of species' abundances ultimately determines the stability of total community biomass, but the processes that drive species synchrony are most tightly linked to each species' immediate response to environmental conditions and competition
 (Loreau and {{de Mazancourt}} 2008).
-\textcolor{blue}{Therefore, we focus on per capita growth rates, which represent a species' immediate response to interannual fluctuations.}
+Therefore, we focus on per capita growth rates, which represent a species' immediate response to interannual fluctuations.
 
@@
 interactions, and demographic stochasticity on
-\textcolor{blue}{the synchrony of dominant species} in natural plant
+the synchrony of dominant species in natural plant
 communities.
-\textcolor{blue}{While our focus is limited to dominant species due to data constraints, previous work indicates that the dynamics of dominant species determine ecosystem functioning in grasslands}
+While our focus is limited to dominant species due to data constraints, previous work indicates that the dynamics of dominant species determine ecosystem functioning in grasslands
 (Smith and Knapp 2003, Bai et al. 2004, Grman et al. 2010, Sasaki and
@@
 reveal the processes that determine synchrony
-\textcolor{blue}{of dominant species} in our focal grassland
+of dominant species in our focal grassland
 communities.
@@
 While existing theory has identified
-\textcolor{blue}{environmental responses, species interactions, and demographic stochasticity as the drivers of the temporal synchrony,}
+environmental responses, species interactions, and demographic stochasticity as the drivers of the temporal synchrony,
 we do not have a simple expression to predict synchrony in a particular
@@
 \(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the
-\textcolor{blue}{temporal variance of species \emph{i}'s response to the environment and $u_{ei}(t)$ is a normal variable with unit variance that is independent among species but may be correlated.}
-\textcolor{blue}{The product, $\sigma_{ei}u_{ei}(t)$, is species \emph{i}'s environmental response.}
+temporal variance of species \emph{i}'s response to the environment and $u_{ei}(t)$ is a normal variable with unit variance that is independent among species but may be correlated.
+The product, $\sigma_{ei}u_{ei}(t)$, is species \emph{i}'s environmental response.
 Demographic stochasticity arises from variations in births and deaths
@@
 variance
-(\textcolor{blue}{i.e., the intrinsic demographic stochasticity of species \emph{i}})
+(i.e., the intrinsic demographic stochasticity of species \emph{i})
 and \(u_{di}(t)\) are independent normal variables with zero mean and
-unit variance
-\textcolor{blue}{that allow demographic stochasiticity to vary through time.}
+unit variance that allow demographic stochasticity to vary through time.
 
@@
 using the metric of Loreau and {{de Mazancourt}} (2008) in the
-`synchrony' package (Gouhier and Guichard 2014) in R (R Core Team 2013).
+\texttt{synchrony} package (Gouhier and Guichard 2014) in \texttt{R} (R Core Team 2013).
 Specifically, we calculated synchrony as
@@
 \noindent where \(\sigma_{r_{i}}\) is the temporal
-\textcolor{blue}{standard deviation} of species \emph{i'}s per capita
+standard deviation of species \emph{i'}s per capita
 population growth rate (\(r_i\)) and \(\sigma^{2}_{T}\) is the temporal
@@
 
-\textcolor{blue}{Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
+Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
 Strong year-to-year variation in these crowding effects would indicate a statistical interaction between environmental conditions and species interactions.
 We tested for such a dynamic by comparing models with and without random year effects on crowding.
-Based on the results, we decided to treat crowding as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including interannual variation in crowding did not change our results.}
+Based on the results, we decided to treat crowding as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including interannual variation in crowding did not change our results.
 
@@
 hierarchical Bayesian approach implemented in JAGS (Plummer 2003) using
-the `rjags' package (Plummer 2014) in R (R Core Team 2013). Again,
+the \texttt{rjags} package (Plummer 2014) in \texttt{R} (R Core Team 2013). Again,
 temporal and spatial variation were treated as random effects.
@@
 interactions and environmental stochasticity removed.
-\textcolor{blue}{We did not include a simulation with only interspecific interactions because our population models run to deterministic equilibriums in the absence of environmental or demographic stochasticity.}
+We did not include a simulation with only interspecific interactions because our population models run to deterministic equilibriums in the absence of environmental or demographic stochasticity.
 We ran IPM simulations for 2,000 time steps, after an initial 500
@@
 size distribution. We then calculated yearly per capita growth rates
-from the simlated time series, and then caluclated the synchrony of
+from the simulated time series, and then calculated the synchrony of
 species' per capita growth rates over 100 randomly selected contiguous
@@
 the 100 time steps for each no extinction run within a model experiment.
-\textcolor{blue}{In the IBM simulations, the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
-To explore this effect}, we ran simulations \textcolor{blue}{(4)} and
+In the IBM simulations, the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
+To explore this effect, we ran simulations (4) and
 (6) on plot sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM
 simulations were run on a 25 \(\text{m}^2\) landscape
-\textcolor{blue}{to most closely match the implicit, large spatial scale of the IPM simulations}.
-
-\textcolor{blue}{Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulated values of community-wide species synchrony.}
+to most closely match the implicit, large spatial scale of the IPM simulations.
+
+Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulated values of community-wide species synchrony.
 The simulation experiments also allow us to test our theoretical
@@
 case theoretical predictions.
-\textcolor{blue}{This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalence in species' growth rates, environmental response variances, and demographic variances.
+This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalence in species' growth rates, environmental response variances, and demographic variances.
 Our empirically-based models do not make these assumptions.
-Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions hold in the natural community our model represents.}
+Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions hold in the natural community our model represents.
+
+All \texttt{R} code necessary to reproduce our results has been deposited on Figshare (\url{http://doi.org/10.6084/m9.figshare.4587073}) and released on \textsc{GitHub} (\url{http://github.com/atredennick/community_synchrony/releases/tag/v1}).
 
@@
 
-\textcolor{blue}{Observed} synchrony of species' per capita growth rates
+Observed synchrony of species' per capita growth rates
 at our study sites range from 0.36 to 0.89 and synchrony of percent
@@
 correlated (Table S3).
-\textcolor{blue}{These temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.}
+These temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.
 
@@
 \(\mathcal{M}_E\)). At 1 \(\text{m}^2\), synchrony reflected
-\textcolor{blue}{the combined effects of} demographic stochasticity and
+the combined effects of demographic stochasticity and
 environmental stochasticity (e.g.,
-\textcolor{blue}{simulated synhcrony fell} between \(\mathcal{M}_E\) and
+simulated synchrony fell between \(\mathcal{M}_E\) and
 \(\mathcal{M}_D\)). For context, population sizes increased from an
@@
 Our study produced four main findings that were generally consistent
-across five natural plant communities: (1) limiting-case predictions
+across five natural grassland plant communities: (1) limiting-case predictions
 from the theoretical model were well-supported by simulations from the
@@
 theory.
-\textcolor{blue}{However, whether the theoretical model adequately represents more complex communities remains unknown because our analysis was restricted to dominant species.}
-
-\subsection{Demographic stochasticity decreases synchrony \textcolor{blue}{in small populations}}
+However, whether the theoretical model adequately represents more complex communities remains unknown because our analysis was restricted to dominant species.
+
+\subsection{Demographic stochasticity decreases synchrony in small populations}
 
@@
 2008).
-\textcolor{blue}{This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
-The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, reducing the relative effect of environmental forcing.}
+This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
+The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, reducing the relative effect of environmental forcing.
 Even in small populations
-\textcolor{blue}{(e.g., 1 $\text{m}^2$ lanscapes)}, however, demographic
+(e.g., 1 $\text{m}^2$ lanscapes), however, demographic
 stochasticity was not strong enough to compensate the synchronizing
@@
 are the primary driver of community-wide species synchrony
-\textcolor{blue}{among the dominant species} we studied. Given
+among the dominant species we studied. Given
 accumulating evidence that niche differences in natural communities are
@@
 
-\textcolor{blue}{Environmental responses synchronized dynamics relative to a null expectation of independent species interactions (e.g., "No Comp.
+Environmental responses synchronized dynamics relative to a null expectation of independent species interactions (e.g., "No Comp.
 + No E.S." simulations in Fig.
@@
 Synchrony was far from complete because of differences in species' responses to interannual environmental variation.
-Many studies of ecosystem stability in semiarid grasslands focus on trad-offs among dominant functional types}
+Many studies of ecosystem stability in semiarid grasslands focus on trad-offs among dominant functional types
 (Bai et al. 2004, Sasaki and Lauenroth 2011).
-\textcolor{blue}{Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
+Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
 At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were positively correlated.
@@
 The subtle differences among ecologically-similar dominant species do impact species synchrony and, ultimately, ecosystem stability.
-Ignoring such differences could mask important dynamics that underpin ecosystem functioning.}
+Ignoring such differences could mask important dynamics that underpin ecosystem functioning.
 
@@
 will tend to synchronize dynamics. However, if community-wide synchrony
-\textcolor{blue}{among dominant species} is less than perfect, as it is
+among dominant species is less than perfect, as it is
 in all our focal communities, some degree of compensatory dynamics must
 be present (Loreau and {{de Mazancourt}} 2008).
-\textcolor{blue}{Even ecologically-similar species, which are sometimes aggregated into functional groups, have environmental responses that are dissimilar enough to limit synchrony.
-Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.}
+Even ecologically-similar species, which are sometimes aggregated into functional groups, have environmental responses that are dissimilar enough to limit synchrony.
+Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.
 
@@
 
-\setlength{\parindent}{0ex} \newpage{}
+\setlength{\parindent}{0ex}
+
+\section{Literature Cited}
+
+Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
+perennial plants: An embarrassment of niches. Ecology Letters
+13:1019--1029.
+
+Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
+mapped quadrats from Kansas prairie: demographic information for
+herbaceous plants. Ecology 88:2673.
+
+Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
+of semi-desert grassland plants in permanent quadrats mapped from 1915
+to 1947. Ecology 93:1492--1492.
+
+Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
+mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
+92:1703.
+
+Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
+and compensatory effects in the Inner Mongolia grassland. Nature
+431:181--184.
+
+Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
+Functional Similarity, and Top-Down Control Drive Synchronization and
+the Reliability of Ecosystem Function. The American Naturalist
+183:394--409.
+
+Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
+P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
+Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
+and its impact on humanity. Nature 489:59--67.
+
+Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
+recruitment stage to stabilize grassland coexistence. Ecological
+Monographs 85:373--392.
+
+{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
+Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
+Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
+Loreau. 2013. Predicting ecosystem stability from community composition
+and biodiversity. Ecology Letters 16:617--625.
+
+Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
+Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
+biodiversity change but not systematic loss. Science 344:296--9.
+
+Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. Annual Review of
+Ecology, Evolution, and Systematics 40:393--414.
+
+Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
+variability in space and time. Methods in Ecology and Evolution
+5:524--533.
+
+Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
+Stability of Food Webs in Metacommunities. The American Naturalist
+175:E16--E34.
+
+Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
+Mechanisms contributing to stability in ecosystem function depend on the
+environmental context. Ecology Letters 13:1400--1410.
+
+Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
+Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
+temporal stability of biomass production: a new analysis of recent
+biodiversity experiments. The American naturalist 183:1--12.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
+on ecosystem functioning: a consensus of current knowledge. Ecological
+Monographs 75:3--35.
+
+Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
+Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
+Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
+Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
+Compensatory dynamics are rare in natural ecological communities.
+Proceedings of the National Academy of Sciences of the United States of
+America 104:3273--3277.
+
+Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
+productivity and the temporal stability of productivity: Patterns and
+processes. Ecology Letters 12:443--451.
+
+Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
+dynamics in ecology and conservation.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
+drivers: neutral and nonneutral community dynamics in fluctuating
+environments. The American Naturalist 172:E48--E66.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+Loreau, M., and a Hector. 2001. Partitioning selection and
+complementarity in biodiversity experiments. Nature 412:72--6.
+
+Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
+Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
+and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
+knowledge and future challenges. Science 294:804--808.
+
+Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
+community dynamics? Proceedings of the Royal Society B: Biological
+Sciences 276:2923--2929.
+
+Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
+Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
+3rd international workshop on distributed statistical computing (dSC
+2003). march.
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+
+R Core Team. 2013. R: A language and environment for statistical
+computing.
+
+Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
+O'Hara. 2008. Detecting compensatory dynamics in competitive communities
+under environmental forcing. Oikos 117:1907--1911.
+
+Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
+A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
+Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
+C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
+Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
+Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
+Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
+Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
+Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
+Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
+Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
+operating space for humanity. Nature 461:472--475.
+
+Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
+for Latent Gaussian Models Using Integrated Nested Laplace
+Approximations. Journal of the Royal Statistical Society, Series B
+71:319--392.
+
+Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
+Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
+analysis of latent Gaussian models using Integrated Nested Laplace
+Approximaxion.
+
+Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
+diversity, regulates temporal stability of plant communities. Oecologia
+166:761--768.
+
+Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
+function with non-random species loss. Ecology Letters 6:509--517.
+
+Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
+189:732--735.
+
+Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
+and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
+
+Tilman, D. 1988. Plant strategies and the dynamics and structure of
+plant communities. Plant strategies and the dynamics and structure of
+plant communities.:360 pp.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
+2001. Diversity and productivity in a long-term grassland experiment.
+Science 294:843--845.
+
+Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
+stabilize food web dynamics by increasing synchrony. Ecology Letters
+10:1066--1074.
+
+Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
+biodiversity over time. Proceedings of the National Academy of Sciences
+of the United States of America 110:19456--9.
+
+Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
+2013. Shared ancestry influences community stability by altering
+competitive interactions: evidence from a laboratory microcosm
+experiment using freshwater green algae. Proceedings of the Royal
+Society of London B: Biological Sciences 280.
+
+Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
+Human Domination of Earth' s Ecosystems. Science 277:494--499.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
+across scales in metacommunities. Ecology Letters.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
+sagebrush steppe: long-term data for analyzing demographic rates and
+plant--plant interactions. Ecology 91:3427.
+
+\newpage{}
 
@@
 
-\section{Figures}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{./components/formatted_figures/synchrony_flowchart_bw.png}
-  \caption{Diagram of our coupled theoretical-empirical approach. We followed this workflow for each of our five focal communities.}
-\end{figure}
-
-\pagebreak{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure1.png}
-  \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
+\doublespacing
+
+\section{Figure Legends}
+
+\textbf{FIGURE 1:} Diagram of our coupled theoretical-empirical approach. We followed this workflow for each of our five focal communities.
+
+\textbf{FIGURE 2:} Community-wide species synchrony of per capita growth rates from model simulation experiments.
 Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific interactions present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific interactions removed (``No Comp.''), interspecific interactions and demographic stochasticity removed (``No Comp.
@@
 All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
-Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
-\end{figure}
-
-\pagebreak{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
-  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.
+
+\textbf{FIGURE 3:} Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
 + No E.S.
@@
 + E.S.)'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
-\end{figure}
-
-\pagebreak{}
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
-Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
-perennial plants: An embarrassment of niches. Ecology Letters
-13:1019--1029.
-
-Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
-mapped quadrats from Kansas prairie: demographic information for
-herbaceous plants. Ecology 88:2673.
-
-Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
-of semi-desert grassland plants in permanent quadrats mapped from 1915
-to 1947. Ecology 93:1492--1492.
-
-Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
-mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
-92:1703.
-
-Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
-and compensatory effects in the Inner Mongolia grassland. Nature
-431:181--184.
-
-Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
-Functional Similarity, and Top-Down Control Drive Synchronization and
-the Reliability of Ecosystem Function. The American Naturalist
-183:394--409.
-
-Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
-P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
-Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
-and its impact on humanity. Nature 489:59--67.
-
-Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
-recruitment stage to stabilize grassland coexistence. Ecological
-Monographs 85:373--392.
-
-{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
-Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
-Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
-Loreau. 2013. Predicting ecosystem stability from community composition
-and biodiversity. Ecology Letters 16:617--625.
-
-Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
-Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
-biodiversity change but not systematic loss. Science 344:296--9.
-
-Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
-Compensatory Dynamics in Ecological Communities. Annual Review of
-Ecology, Evolution, and Systematics 40:393--414.
-
-Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
-variability in space and time. Methods in Ecology and Evolution
-5:524--533.
-
-Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
-Stability of Food Webs in Metacommunities. The American Naturalist
-175:E16--E34.
-
-Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
-Mechanisms contributing to stability in ecosystem function depend on the
-environmental context. Ecology Letters 13:1400--1410.
-
-Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
-Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
-temporal stability of biomass production: a new analysis of recent
-biodiversity experiments. The American naturalist 183:1--12.
-
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
-
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
-
-Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
-on ecosystem functioning: a consensus of current knowledge. Ecological
-Monographs 75:3--35.
-
-Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
-Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
-Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
-Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
-Compensatory dynamics are rare in natural ecological communities.
-Proceedings of the National Academy of Sciences of the United States of
-America 104:3273--3277.
-
-Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
-productivity and the temporal stability of productivity: Patterns and
-processes. Ecology Letters 12:443--451.
-
-Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
-dynamics in ecology and conservation.
-
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
-
-Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
-drivers: neutral and nonneutral community dynamics in fluctuating
-environments. The American Naturalist 172:E48--E66.
-
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
-
-Loreau, M., and a Hector. 2001. Partitioning selection and
-complementarity in biodiversity experiments. Nature 412:72--6.
-
-Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
-Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
-and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
-knowledge and future challenges. Science 294:804--808.
-
-Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
-community dynamics? Proceedings of the Royal Society B: Biological
-Sciences 276:2923--2929.
-
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
-
-R Core Team. 2013. R: A language and environment for statistical
-computing.
-
-Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
-O'Hara. 2008. Detecting compensatory dynamics in competitive communities
-under environmental forcing. Oikos 117:1907--1911.
-
-Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
-A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
-Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
-C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
-Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
-Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
-Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
-Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
-Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
-Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
-Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
-operating space for humanity. Nature 461:472--475.
-
-Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
-for Latent Gaussian Models Using Integrated Nested Laplace
-Approximations. Journal of the Royal Statistical Society, Series B
-71:319--392.
-
-Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
-Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
-analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
-
-Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
-diversity, regulates temporal stability of plant communities. Oecologia
-166:761--768.
-
-Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
-function with non-random species loss. Ecology Letters 6:509--517.
-
-Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
-189:732--735.
-
-Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
-and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
-
-Tilman, D. 1988. Plant strategies and the dynamics and structure of
-plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
-
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
-
-Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
-2001. Diversity and productivity in a long-term grassland experiment.
-Science 294:843--845.
-
-Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
-stabilize food web dynamics by increasing synchrony. Ecology Letters
-10:1066--1074.
-
-Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
-2013. Global meta-analysis reveals no net change in local-scale plant
-biodiversity over time. Proceedings of the National Academy of Sciences
-of the United States of America 110:19456--9.
-
-Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
-2013. Shared ancestry influences community stability by altering
-competitive interactions: evidence from a laboratory microcosm
-experiment using freshwater green algae. Proceedings of the Royal
-Society of London B: Biological Sciences 280.
-
-Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
-Human Domination of Earth' s Ecosystems. Science 277:494--499.
-
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
-
-Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
-
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
-
-Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
-sagebrush steppe: long-term data for analyzing demographic rates and
-plant--plant interactions. Ecology 91:3427.
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.
 

### 2017-01-26 docs/ms_communitySynchrony.tex
just committing things to this branch to get out

@@
 
-\setlength{\parindent}{0ex} \newpage{}
+\setlength{\parindent}{0ex}
+
+\section{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
+perennial plants: An embarrassment of niches. Ecology Letters
+13:1019--1029.
+
+Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
+mapped quadrats from Kansas prairie: demographic information for
+herbaceous plants. Ecology 88:2673.
+
+Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
+of semi-desert grassland plants in permanent quadrats mapped from 1915
+to 1947. Ecology 93:1492--1492.
+
+Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
+mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
+92:1703.
+
+Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
+and compensatory effects in the Inner Mongolia grassland. Nature
+431:181--184.
+
+Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
+Functional Similarity, and Top-Down Control Drive Synchronization and
+the Reliability of Ecosystem Function. The American Naturalist
+183:394--409.
+
+Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
+P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
+Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
+and its impact on humanity. Nature 489:59--67.
+
+Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
+recruitment stage to stabilize grassland coexistence. Ecological
+Monographs 85:373--392.
+
+{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
+Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
+Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
+Loreau. 2013. Predicting ecosystem stability from community composition
+and biodiversity. Ecology Letters 16:617--625.
+
+Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
+Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
+biodiversity change but not systematic loss. Science 344:296--9.
+
+Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. Annual Review of
+Ecology, Evolution, and Systematics 40:393--414.
+
+Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
+variability in space and time. Methods in Ecology and Evolution
+5:524--533.
+
+Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
+Stability of Food Webs in Metacommunities. The American Naturalist
+175:E16--E34.
+
+Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
+Mechanisms contributing to stability in ecosystem function depend on the
+environmental context. Ecology Letters 13:1400--1410.
+
+Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
+Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
+temporal stability of biomass production: a new analysis of recent
+biodiversity experiments. The American naturalist 183:1--12.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
+on ecosystem functioning: a consensus of current knowledge. Ecological
+Monographs 75:3--35.
+
+Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
+Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
+Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
+Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
+Compensatory dynamics are rare in natural ecological communities.
+Proceedings of the National Academy of Sciences of the United States of
+America 104:3273--3277.
+
+Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
+productivity and the temporal stability of productivity: Patterns and
+processes. Ecology Letters 12:443--451.
+
+Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
+dynamics in ecology and conservation.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
+drivers: neutral and nonneutral community dynamics in fluctuating
+environments. The American Naturalist 172:E48--E66.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+Loreau, M., and a Hector. 2001. Partitioning selection and
+complementarity in biodiversity experiments. Nature 412:72--6.
+
+Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
+Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
+and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
+knowledge and future challenges. Science 294:804--808.
+
+Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
+community dynamics? Proceedings of the Royal Society B: Biological
+Sciences 276:2923--2929.
+
+Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
+Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
+3rd international workshop on distributed statistical computing (dSC
+2003). march.
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+
+R Core Team. 2013. R: A language and environment for statistical
+computing.
+
+Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
+O'Hara. 2008. Detecting compensatory dynamics in competitive communities
+under environmental forcing. Oikos 117:1907--1911.
+
+Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
+A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
+Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
+C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
+Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
+Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
+Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
+Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
+Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
+Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
+Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
+operating space for humanity. Nature 461:472--475.
+
+Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
+for Latent Gaussian Models Using Integrated Nested Laplace
+Approximations. Journal of the Royal Statistical Society, Series B
+71:319--392.
+
+Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
+Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
+analysis of latent Gaussian models using Integrated Nested Laplace
+Approximaxion.
+
+Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
+diversity, regulates temporal stability of plant communities. Oecologia
+166:761--768.
+
+Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
+function with non-random species loss. Ecology Letters 6:509--517.
+
+Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
+189:732--735.
+
+Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
+and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
+
+Tilman, D. 1988. Plant strategies and the dynamics and structure of
+plant communities. Plant strategies and the dynamics and structure of
+plant communities.:360 pp.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
+2001. Diversity and productivity in a long-term grassland experiment.
+Science 294:843--845.
+
+Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
+stabilize food web dynamics by increasing synchrony. Ecology Letters
+10:1066--1074.
+
+Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
+biodiversity over time. Proceedings of the National Academy of Sciences
+of the United States of America 110:19456--9.
+
+Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
+2013. Shared ancestry influences community stability by altering
+competitive interactions: evidence from a laboratory microcosm
+experiment using freshwater green algae. Proceedings of the Royal
+Society of London B: Biological Sciences 280.
+
+Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
+Human Domination of Earth' s Ecosystems. Science 277:494--499.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
+across scales in metacommunities. Ecology Letters.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
+sagebrush steppe: long-term data for analyzing demographic rates and
+plant--plant interactions. Ecology 91:3427.
+
+\newpage{}
 
 \section{Tables}
-
-\singlespacing
 
@@
 \pagebreak{}
+\section{Figure Captions}
+FIGURE 1: Diagram of our coupled theoretical-empirical approach. We followed this workflow for each of our five focal communities.
+
+FIGURE 2: Community-wide species synchrony of per capita growth rates from model simulation experiments.
+Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific interactions present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific interactions removed (``No Comp.''), interspecific interactions and demographic stochasticity removed (``No Comp.
++ No D.S.''), and interspecific interactions and environmental stochasticity removed (``No Comp.
++ No E.S.'').
+Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation.
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.
+All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.
+
+FIGURE 3: Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
++ No E.S.
+(D.S.
+Only)'') and when environmental stochasticity is also present (``No Comp.
+(D.S.
++ E.S.)'').
+The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
+The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
+Theoretically, ``No Comp.
++ No E.S.
+(D.S.
+Only)'' simulations should remain constant across landscape size, whereas ``No Comp.
+(D.S.
++ E.S.)'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.
+
+\pagebreak{}
 
@@
 
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
-Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
-perennial plants: An embarrassment of niches. Ecology Letters
-13:1019--1029.
-
-Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
-mapped quadrats from Kansas prairie: demographic information for
-herbaceous plants. Ecology 88:2673.
-
-Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
-of semi-desert grassland plants in permanent quadrats mapped from 1915
-to 1947. Ecology 93:1492--1492.
-
-Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
-mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
-92:1703.
-
-Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
-and compensatory effects in the Inner Mongolia grassland. Nature
-431:181--184.
-
-Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
-Functional Similarity, and Top-Down Control Drive Synchronization and
-the Reliability of Ecosystem Function. The American Naturalist
-183:394--409.
-
-Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
-P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
-Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
-and its impact on humanity. Nature 489:59--67.
-
-Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
-recruitment stage to stabilize grassland coexistence. Ecological
-Monographs 85:373--392.
-
-{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
-Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
-Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
-Loreau. 2013. Predicting ecosystem stability from community composition
-and biodiversity. Ecology Letters 16:617--625.
-
-Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
-Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
-biodiversity change but not systematic loss. Science 344:296--9.
-
-Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
-Compensatory Dynamics in Ecological Communities. Annual Review of
-Ecology, Evolution, and Systematics 40:393--414.
-
-Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
-variability in space and time. Methods in Ecology and Evolution
-5:524--533.
-
-Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
-Stability of Food Webs in Metacommunities. The American Naturalist
-175:E16--E34.
-
-Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
-Mechanisms contributing to stability in ecosystem function depend on the
-environmental context. Ecology Letters 13:1400--1410.
-
-Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
-Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
-temporal stability of biomass production: a new analysis of recent
-biodiversity experiments. The American naturalist 183:1--12.
-
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
-
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
-
-Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
-on ecosystem functioning: a consensus of current knowledge. Ecological
-Monographs 75:3--35.
-
-Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
-Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
-Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
-Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
-Compensatory dynamics are rare in natural ecological communities.
-Proceedings of the National Academy of Sciences of the United States of
-America 104:3273--3277.
-
-Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
-productivity and the temporal stability of productivity: Patterns and
-processes. Ecology Letters 12:443--451.
-
-Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
-dynamics in ecology and conservation.
-
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
-
-Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
-drivers: neutral and nonneutral community dynamics in fluctuating
-environments. The American Naturalist 172:E48--E66.
-
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
-
-Loreau, M., and a Hector. 2001. Partitioning selection and
-complementarity in biodiversity experiments. Nature 412:72--6.
-
-Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
-Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
-and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
-knowledge and future challenges. Science 294:804--808.
-
-Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
-community dynamics? Proceedings of the Royal Society B: Biological
-Sciences 276:2923--2929.
-
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
-
-R Core Team. 2013. R: A language and environment for statistical
-computing.
-
-Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
-O'Hara. 2008. Detecting compensatory dynamics in competitive communities
-under environmental forcing. Oikos 117:1907--1911.
-
-Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
-A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
-Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
-C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
-Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
-Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
-Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
-Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
-Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
-Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
-Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
-operating space for humanity. Nature 461:472--475.
-
-Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
-for Latent Gaussian Models Using Integrated Nested Laplace
-Approximations. Journal of the Royal Statistical Society, Series B
-71:319--392.
-
-Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
-Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
-analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
-
-Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
-diversity, regulates temporal stability of plant communities. Oecologia
-166:761--768.
-
-Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
-function with non-random species loss. Ecology Letters 6:509--517.
-
-Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
-189:732--735.
-
-Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
-and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
-
-Tilman, D. 1988. Plant strategies and the dynamics and structure of
-plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
-
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
-
-Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
-2001. Diversity and productivity in a long-term grassland experiment.
-Science 294:843--845.
-
-Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
-stabilize food web dynamics by increasing synchrony. Ecology Letters
-10:1066--1074.
-
-Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
-2013. Global meta-analysis reveals no net change in local-scale plant
-biodiversity over time. Proceedings of the National Academy of Sciences
-of the United States of America 110:19456--9.
-
-Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
-2013. Shared ancestry influences community stability by altering
-competitive interactions: evidence from a laboratory microcosm
-experiment using freshwater green algae. Proceedings of the Royal
-Society of London B: Biological Sciences 280.
-
-Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
-Human Domination of Earth' s Ecosystems. Science 277:494--499.
-
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
-
-Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
-
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
-
-Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
-sagebrush steppe: long-term data for analyzing demographic rates and
-plant--plant interactions. Ecology 91:3427.
-
 \end{document}

### 2017-01-26 docs/ms_communitySynchrony_PREPRINT.tex
just committing things to this branch to get out

@@
-\documentclass[12pt,]{article}
+\documentclass[11pt,]{article}
 \usepackage{lmodern}
@@
 \pagestyle{fancy}
-\chead{\color{red}\textsf{THIS MANUSCRIPT HAS NOT UNDERGONE PEER REVIEW}}
+\chead{\color{red}\textsf{preprint}}
+\usepackage{titlesec}
+\titleformat{\section}{\normalsize\sc\centering}{\thesection}{1em}{}
+\titleformat{\subsection}{\normalsize\it\centering}{\thesubsection}{1em}{}
+\titleformat*{\paragraph}{\normalsize\it}
 \lhead{}
@@
 \thispagestyle{fancy}
+
+\definecolor{blue}{gray}{0}
+
 \begin{singlespace}
@@
 
-\normalsize{Andrew T.
-Tredennick\textsuperscript{1}\footnote{Correspondence: Email: atredenn@gmail.com}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B.
-Adler\textsuperscript{1}}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
 
@@
 \begin{abstract}
-Temporal asynchrony among species \textcolor{blue}{helps}\footnote{\textcolor{blue}{Changes and insertions unique to this version are shown in blue text.}} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
-Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \textcolor{blue}{interactions}, and demographic stochasticity.
-We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
-\textcolor{blue}{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
-Removing interspecific \textcolor{blue}{interactions} and demographic stochasticity had small effects on synchrony.
-In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
-
-\vspace{2em}
-
-\noindent{}\textbf{Keywords}\\
-\noindent{}synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability
+Temporal asynchrony among species helps diversity to stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
+Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific interactions, and demographic stochasticity.
+We then conduct simulation experiments with empirical population models to quantify the relative \textcolor{blue}{influence} of these factors \textcolor{blue}{on the synchrony of dominant species in five semiarid grasslands.}
+We found that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
+Removing interspecific interactions and demographic stochasticity had small effects on synchrony. 
+\textcolor{blue}{For the dominant species in these plant communities}, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+
+\vspace{1em}
+
+\emph{Key words: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability, grassland}
+
 \end{abstract}
@@
 temporal stability of productivity (Tilman et al. 2006, Hector et al.
-2010). The stabilizing effect of species richness stems from individual
-species responding \textcolor{blue}{in different ways} to environmental
-fluctuations (environmental stochasticity), or fluctuating
-asynchronously because of random chance events (demographic
-stochasticity) (Isbell et al. 2009, Hector et al. 2010, {{de
-Mazancourt}} et al. 2013). Species richness affects synchrony because
-larger species pools are more likely to contain species that respond
-disimilarly to environmental conditions (Yachi and Loreau 1999),
-implying that species losses will reduce ecosystem stability. Even
-without species losses, abiotic homogenization can weaken compensatory
-dynamics and, in turn, decrease temporal stability of ecosystem
-functioning (Hautier et al. 2014).
-\textcolor{blue}{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}
-
-Theory identifies three main determinants of species synchrony:
-environmental stochasticity, demographic stochasticity, and
-\textcolor{blue}{interspecific interactions} (Loreau and {{de
-Mazancourt}} 2008, 2013, Gonzalez and Loreau 2009). For example, in a
-community composed of large \textcolor{blue}{populations} (no
-demographic stochasticity)
-\textcolor{blue}{with weak interspecific interations}, community-wide
-species synchrony should be determined by the covariance of species'
-responses to the environment (Loreau and {{de Mazancourt}} 2008).
-However, \textcolor{blue}{this} prediction relies on a relatively simple
-population model and only holds under two assumptions: (i) species'
-responses to the environment are similar in magnitude and (ii) all
-species have similar growth rates.
-\textcolor{blue}{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities}
-(Mutshinda et al. 2009, Thibaut et al. 2012),
-\textcolor{blue}{and they did not consider demographic stochasiticity.}
-
-In grasslands, most empirical studies have focused on whether species
-synchrony is primarily an outcome of species-specific responses to
-environmental conditions (Hautier et al. 2014) or competition (Gross et
-al. 2014). Even beyond grassland studies, whether competition or
-environmental responses drive compensatory dynamics remains
-controversial (reviewed in Gonzalez and Loreau 2009). In part,
-controversy remains because quantifying the relative strengths of each
-driver on the degree of synchrony from
-\textcolor{blue}{the covariance matrix of species abundances} (e.g.,
-Houlahan et al. 2007) is impossible. This is because an unbiased null
-expectation for synchrony does not exist (Loreau and {{de Mazancourt}}
-2008) and observed synchrony can arise from non-unique combinations of
-factors (Ranta et al. 2008). For example, weak synchrony of population
-abundances could reflect positive environmental correlations
-(synchronizing effect) offset by strong competition (desynchronizing
-effect), or negative environmental correlations and weak competition.
-
-The best way to quantify the effects of environmental stochasticity,
-demographic stochasticity, and interspecific
-\textcolor{blue}{interactions} is to remove them one-by-one, and in
-combination. In principle, this could be done in an extremely controlled
-laboratory setting, but empirically-based models of interacting
-populations, fit with data sets from natural communities, offer a
-practical alternative. For example, Mutshinda et al. (2009) fit a
-dynamic population model to several community time series of insect and
-bird abundances. They used a statistical technique to decompose temporal
-variation into competition and environmental components, and found that
-positively correlated environmental responses among species determined
-community dynamics. Thibaut et al. (2012)
-\textcolor{blue}{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
+2010).
+\textcolor{blue}{The stabilizing effect of species richness arises from a combination of selection effects and complementarity}
+(Loreau and Hector 2001).
+\textcolor{blue}{Selection effects occur when a dominant species has lower than average temporal variability, which generates a positive effect on ecosystem stability}
+(e.g., Grman et al. 2010).
+\textcolor{blue}{Complementarity occurs when species have unique temporal dynamics, causing their abundances to fluctuate asynchronously, increasing ecosystem stability.}
+\textcolor{blue}{The premise of this paper is that understanding the mechanisms driving species' temporal dynamics, and resulting (a)synchrony, is necessary to predict the impacts of global change on ecosystem stability.}
+
+\textcolor{blue}{Asynchronous dynamics, also known as compensatory dynamics}
+(Gonzalez and Loreau 2009),
+\textcolor{blue}{occur whenever species synchrony is not perfect} and
+result from individual species responding in different ways to
+environmental fluctuations, random chance events, and/or competitive
+interactions (Isbell et al. 2009, Hector et al. 2010, {{de Mazancourt}}
+et al. 2013, Gross et al. 2014). Species richness affects
+\textcolor{blue}{the degree of synchrony in a community} because larger
+species pools are more likely to contain species that respond
+dissimilarly to environmental conditions, reducing synchrony and
+increasing stability (Yachi and Loreau 1999).
+\textcolor{blue}{Species richness can also affect synchrony if the strength of species interactions varies systematically with richness, because competition generally increases synchrony and reduces stability}
+(Loreau and {{de Mazancourt}} 2013, but see Gross et al. 2014).
+
+\textcolor{blue}{The effects of environmental change and species losses on ecosystem stability will depend on whether synchrony is driven by species-specific responses to environmental conditions or interspecific competition}
+(Hautier et al. 2014).
+\textcolor{blue}{If responses to environment are important, then environmental change could alter synchrony and stability.
+If competition is important, then the direct effects of environmental change may not affect synchrony and stability, but species gains/losses will.}
+
+\textcolor{blue}{The relative role of environmental responses and competition in driving synchrony  in natural plant communities}
+remains controversial (reviewed in Gonzalez and Loreau 2009). One source
+of the controversy is that quantifying the relative strengths of each
+driver based on the covariance matrix of species abundances (e.g.,
+Houlahan et al. 2007) is impossible. This is because observed synchrony
+can arise from non-unique combinations of factors (Ranta et al. 2008).
+For example, weak synchrony of population abundances could reflect
+positive environmental correlations (synchronizing effect) offset by
+strong competition (desynchronizing effect), or negative environmental
+correlations and weak competition.
+
+\textcolor{blue}{Theory can help us resolve this empirical question.}
+Recent theoretical work has identified three determinants of species
+synchrony: environmental stochasticity, interspecific interactions, and
+demographic stochasticity (Loreau and {{de Mazancourt}} 2008, 2013,
+Gonzalez and Loreau 2009).
+\textcolor{blue}{This theory has been developed by focusing on simple limiting cases in which only one of these three drivers operates.}
+For example, in a community composed of large populations (no
+demographic stochasticity) with weak interspecific interations,
+community-wide species synchrony should be determined by the covariance
+of species' responses to the environment (Loreau and {{de Mazancourt}}
+2008). However, this prediction relies on a relatively simple population
+model and \textcolor{blue}{depends on two restrictive} assumptions: (i)
+species' responses to the environment are similar in magnitude and (ii)
+all species have similar growth rates. Whether such theoretical
+predictions hold in natural communities where species differences are
+unlikely to be symmetrical is unkown because few studies have explicitly
+tested theory on the drivers of species synchrony in natural communities
+(Mutshinda et al. 2009, Thibaut et al. 2012), and they did not consider
+demographic stochasiticity.
+
+An intuitive way to quantify the effects of environmental stochasticity,
+demographic stochasticity, and interspecific interactions is to remove
+them one-by-one, and in combination.
+\textcolor{blue}{If synchrony changes more when we remove environmental stochasticity than when we remove interspecific interactions, we would conclude that environmental fluctuations are the more important driver.}
+In principle, this could be done in an extremely controlled laboratory
+setting (e.g., Venail et al. 2013), but empirically-based models of
+interacting populations, fit with data sets from natural communities,
+offer a practical alternative. For example, Mutshinda et al. (2009) fit
+a dynamic population model to several community time series of insect
+and bird abundances. They used a statistical technique to decompose
+temporal variation into competition and environmental components, and
+found that positively correlated environmental responses among species
+determined community dynamics. Thibaut et al. (2012) used a similar
+approach for reef fish and came to a similar conclusion: environmental
+responses determine synchrony.
+
 While a major step forward, Mutshinda et al.'s (2009) and Thibaut et
-al.'s (2012) modeling technique relied on abundance data that may or may
-not reliably capture competitive interactions that occur at the
-individual level. Furthermore, although both studies quantified the
-relative importance of environmental stochasticity and
-\textcolor{blue}{interspecific interactions} to explain the observed
-variation of species synchrony, they did not use the model to quantify
-how much synchrony would change when each factor is removed.
-
-Here, we use multi-species population models fit to long-term
-demographic data from five semi-arid plant communities to test theory on
-the drivers of species synchrony. Our objectives are to (1) derive and
-test theoretical predictions of species synchrony and (2) determine the
-relative influence of environmental stochasticity, demographic
-stochasticity, and interspecific \textcolor{blue}{interactions} on
-species synchrony in natural plant communities. To achieve these
-objectives, we first refine theory that has been used to predict the
-effects of species richness on ecosystem stability ({{de Mazancourt}} et
-al. 2013) and species synchrony (Loreau and {{de Mazancourt}} 2008) to
-generate predictions of community-wide species synchrony under two
-limiting cases derived from the dynamics of individual species in
-monoculture. We then confront our theoretical predictions with
-simulations from the empirically-based population models. Second, we use
-the multi-species population models to perform simulation experiments
-that isolate the effects of environmental stochasticity, demographic
-stochasticity, and interspecific \textcolor{blue}{interactions} on
-community-wide species synchrony. Given that our population models
+al.'s (2012) modeling technique does have some limitations.
+\textcolor{blue}{First, although both studies quantified the relative importance of environmental stochasticity and interspecific interactions to explain the observed variation of species synchrony, they did not use the model to quantify how much synchrony would change if each factor were removed.
+Second, they relied on popluation abundance data that may or may not reliably capture competitive interactions occuring at the individual level.
+Third, fluctuations in abundance may mask the mechanisms that underpin species synchrony.
+The synchrony of species' abundances ultimately determines the stability of total community biomass, but the processes that drive species synchrony are most tightly linked to each species' immediate response to environmental conditions and competition}
+(Loreau and {{de Mazancourt}} 2008).
+\textcolor{blue}{Therefore, we focus on per capita growth rates, which represent a species' immediate response to interannual fluctuations.}
+
+Here, we use multispecies population models fit to long-term demographic
+data from five semi-arid plant communities to test theory on the drivers
+of species synchrony (Fig. 1). Our objectives are to (1) derive and test
+theoretical predictions of species synchrony and (2) determine the
+relative influence of environmental stochasticity, interspecific
+interactions, and demographic stochasticity on
+\textcolor{blue}{the synchrony of dominant species} in natural plant
+communities.
+\textcolor{blue}{While our focus is limited to dominant species due to data constraints, previous work indicates that the dynamics of dominant species determine ecosystem functioning in grasslands}
+(Smith and Knapp 2003, Bai et al. 2004, Grman et al. 2010, Sasaki and
+Lauenroth 2011).
+
+To achieve our objectives, we first refine theory that has been used to
+predict the effects of species richness on ecosystem stability ({{de
+Mazancourt}} et al. 2013) and species synchrony (Loreau and {{de
+Mazancourt}} 2008) to generate predictions of community-wide species
+synchrony under two limiting cases. We then confront our theoretical
+predictions with simulations from the empirically-based population
+models. Second, we use the multi-species population models to perform
+simulation experiments that isolate the effects of environmental
+stochasticity, demographic stochasticity, and interspecific interactions
+on community-wide species synchrony. Given that our population models
 capture the essential features of community dynamics important to
@@
 community dynamics (Chu and Adler 2015), perturbing the models can
-reveal the processes that determine species synchrony in our focal
-grassland communities.
-
-\subsection{THEORETICAL MODEL}\label{theoretical-model}
-
-\subsubsection{The model}\label{the-model}
-
-While existing theory has identified the factors driving synchrony, we
-do not have a simple expression to predict synchrony in a particular
+reveal the processes that determine synchrony
+\textcolor{blue}{of dominant species} in our focal grassland
+communities.
+
+\section{Theoretical Model}\subsection{The model}
+
+While existing theory has identified
+\textcolor{blue}{environmental responses, species interactions, and demographic stochasticity as the drivers of the temporal synchrony,}
+we do not have a simple expression to predict synchrony in a particular
 community with all factors operating simultaneously. However, we can
 derive analytical predictions for species synchrony under special
-limiting cases.
-\textcolor{blue}{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
-We focus on synchrony of per capita growth rates, rather than
-abundances, because growth rates represent the instantaneous response of
-species to the environment and competition, and are less susceptible to
-the legacy effects of drift and disturbance (Loreau and {{de
-Mazancourt}} 2008). We present equivalent results for synchrony of
-species abundances in the Online Supporting Information, and show that
-they lead to the same overall conclusions as synchrony of per capita
-growth rates. Following Loreau and {{de Mazancourt}} (2008) and {{de
-Mazancourt}} et al. (2013), we define population growth, ignoring
-observation error, as
+limiting cases. The limiting case predictions we derive serve as
+baselines to help us interpret results from empirically-based
+simulations (described below). We focus on synchrony of per capita
+growth rates, rather than abundances, because growth rates represent the
+instantaneous response of species to the environment and competition,
+and are less susceptible to the legacy effects of drift and disturbance
+(Loreau and {{de Mazancourt}} 2008). We present equivalent results for
+synchrony of species abundances in the Appendix S1, and show that they
+lead to the same overall conclusions as synchrony of per capita growth
+rates.
+
+Following Loreau and {{de Mazancourt}} (2008) and {{de Mazancourt}} et
+al. (2013), we define population growth, ignoring observation error, as
 
@@
 species \emph{i}. Environmental stochasticity is incorporated as
-\(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the environmental
-variance and \(u_{ei}\) are normal random variables with zero mean and
-unit variance that are independent through time but may be correlated
-between species. Demographic stochasticity arises from variations in
-births and deaths among individuals (e.g., same states, different
-fates), and is included in the model as a first-order, normal
-approximation (Lande et al. 2003, {{de Mazancourt}} et al. 2013).
-\(\sigma_{di}^2\) is the demographic variance and \(u_{di}(t)\) are
-independent normal variables with zero mean and unit variance. To derive
-analytical predictions we solved a first-order approximation of Equation
-2 ({{de Mazancourt}} et al. 2013 and Online Supporting Information).
-\textcolor{blue}{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos}
-(Loreau and {{de Mazancourt}} 2008).
-\textcolor{blue}{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small}
-(e.g., \(r < 2\)).
-
-\subsubsection{Predictions}\label{predictions}
+\(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the
+\textcolor{blue}{temporal variance of species \emph{i}'s response to the environment and $u_{ei}(t)$ is a normal variable with unit variance that is independent among species but may be correlated.}
+\textcolor{blue}{The product, $\sigma_{ei}u_{ei}(t)$, is species \emph{i}'s environmental response.}
+Demographic stochasticity arises from variations in births and deaths
+among individuals (e.g., same states, different fates), and is included
+in the model as a first-order, normal approximation (Lande et al. 2003,
+{{de Mazancourt}} et al. 2013). \(\sigma_{di}^2\) is the demographic
+variance
+(\textcolor{blue}{i.e., the intrinsic demographic stochasticity of species \emph{i}})
+and \(u_{di}(t)\) are independent normal variables with zero mean and
+unit variance
+\textcolor{blue}{that allow demographic stochasiticity to vary through time.}
+
+To derive analytical predictions we solved a first-order approximation
+of Equation 2 ({{de Mazancourt}} et al. 2013 and Appendix S1). Due to
+the linear approximation approach, our analytical predictions will
+likely fail in communities where species exhibit large fluctuations due
+to limit cycles and chaos (Loreau and {{de Mazancourt}} 2008). Indeed,
+one of the advantages of focusing on growth rates rather than abundances
+is that growth rates are more likely to be well-regulated around an
+equilibrium value, if the long-term average of a species' growth rate is
+relatively small (e.g., \(r < 2\)).
+
+\subsection{Predictions}
 
@@
 covariance of environmental responses between species \emph{i} and
-species \emph{j}. Confronting our theoretical predictions with data
-requires estimates of species dynamics of large populations (no
-demographic stochasticity) \textcolor{blue}{growing in isolation} (no
-interspecific interactions) to calculate the covariance of species'
-environmental responses. To estimate environmental responses in natural
-communities, we turn to our population models built using long-term
-demographic data.
-
-\subsection{EMPIRICAL ANALYSIS}\label{empirical-analysis}
-
-\subsubsection{Materials and methods}\label{materials-and-methods}
-
-\paragraph{Data}\label{data}
+species \emph{j}.
+
+Confronting our theoretical predictions with data requires estimates of
+species dynamics of large populations (no demographic stochasticity)
+growing in isolation (no interspecific interactions) to calculate the
+covariance of species' environmental responses. To estimate
+environmental responses in natural communities, we turn to our
+population models built using long-term demographic data.
+
+\section{Empirical Analysis}\subsection{Materials and methods}
+
+\paragraph{Data.--}\label{data.}
 
@@
 followed the approach of Chu and Adler (2015) to fit statistical models
-for survival, growth, and recruitment (see Online Supporting Information
-for full details). We modeled survival probability of each genet as
-function of genet size, temporal variation among years, permanent
-spatial variation among groups of quadrats, and local neighborhood
-crowding from conspecific and heterospecific genets. Regression
-coefficients for the effect of crowding by each species can be
-considered a matrix of interaction coefficients whose diagonals
-represent intraspecific interactions and whose off-diagonals represent
-interspecific interactions (Adler et al. 2010). These interaction
-coefficients can take positive (facilitative) or negative (competitive)
-values. We modeled growth as the change in size of a genet from one year
-to the next, which depends on the same factors as the survival model. We
-fit the survival and growth regressions using INLA (Rue et al. 2014), a
+for survival, growth, and recruitment (see Appendix S1 for full
+details). We modeled survival probability of each genet as function of
+genet size, temporal variation among years, permanent spatial variation
+among groups of quadrats, and local neighborhood crowding from
+conspecific and heterospecific genets. Regression coefficients for the
+effect of crowding by each species can be considered a matrix of
+interaction coefficients whose diagonals represent intraspecific
+interactions and whose off-diagonals represent interspecific
+interactions (Adler et al. 2010). These interaction coefficients can
+take positive (facilitative) or negative (competitive) values. We
+modeled growth as the change in size of a genet from one year to the
+next, which depends on the same factors as the survival model. We fit
+the survival and growth regressions using INLA (Rue et al. 2014), a
 statistical package for fitting generalized linear mixed effects models
 via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
-2013).
-\textcolor{blue}{Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.}
-Spatial (quadrat groupings) variation was treated as a random effect on
-the intercept and temporal (interannual) variation was treated as random
-effects on the intercept and the effect of genet size in the previous
-year (Online Supporting Information).
+2013). Spatial (quadrat groupings) variation was treated as a random
+effect on the intercept and temporal (interannual) variation was treated
+as random effects on the intercept and the effect of genet size in the
+previous year (Appendix S1).
+
+\textcolor{blue}{Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
+Strong year-to-year variation in these crowding effects would indicate a statistical interaction between environmental conditions and species interactions.
+We tested for such a dynamic by comparing models with and without random year effects on crowding.
+Based on the results, we decided to treat crowding as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including interannual variation in crowding did not change our results.}
 
@@
 randomly in space. Crowding is directly calculated at each time step
-since each genet is spatially referenced (as in the observed data).
-Environmental stochasticity is not an inherent feature of IBMs, but is
-easily included since we fit year-specific temporal random effects for
-each vital rate regression. To include temporal environmental variation,
-at each time step we randomly choose a set of estimated survival,
-growth, and recruitment parameters specific to one observation year. For
-all simulations, we ignore the spatial random effect associated with
-variation among quadrat groups, so our simulations represent an average
-quadrat for each site.
-
-The IPM uses the same vital rate regressions as the IBM (Rees and Ellner
-2009, Rees et al. 2014), but it is spatially implicit and does not
-include demographic stochasticity. Following Chu and Adler (2015), we
-use a mean field approximation that captures the essential features of
-spatial patterning to define the crowding index at each time step
-(Supporting Online Information). Temporal variation is included in
-exactly the same way as for the IBM. For full details on the IPMs we
-use, see Chu and Adler (2015).
-
-\paragraph{Simulation experiments}\label{simulation-experiments}
+since each genet is spatially referenced. Environmental stochasticity is
+not an inherent feature of IBMs, but is easily included since we fit
+year-specific temporal random effects for each vital rate regression. To
+include temporal environmental variation, at each time step we randomly
+choose a set of estimated survival, growth, and recruitment parameters
+specific to one observation year. For all simulations, we ignore the
+spatial random effect associated with variation among quadrat groups, so
+our simulations represent an average quadrat for each site.
+
+The IPM uses the same vital rate regressions as the IBM, but it is
+spatially implicit and does not include demographic stochasticity.
+Following Chu and Adler (2015), we use a mean field approximation that
+captures the essential features of spatial patterning to define the
+crowding index at each time step (Supporting Online Information).
+Temporal variation is included in exactly the same way as for the IBM.
+For full details on the IPMs we use, see Chu and Adler (2015).
+
+\paragraph{Simulation experiments.--}\label{simulation-experiments.}
 
 We performed simulation experiments where drivers (environmental
-stochasticity, demographic stochasticity, or interspecific
-\textcolor{blue}{interactions}) were removed one-by-one and in
-combination. To remove interspecific interactions, we set the
-off-diagonals of the interaction matrix for each vital rate regression
-to zero. This retains intraspecific interactions, and thus
-density-dependence, and results in simulations where species are
-\textcolor{blue}{growing in isolation}.
-\textcolor{blue}{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
-To remove the effect of a fluctuating environment, we removed the
-temporal (interannual) random effects from the regression equations. To
-remove the effect of demographic stochasticity, we use the IPM rather
-than the IBM because the IPM does not include demographic stochasticity
+stochasticity, demographic stochasticity, or interspecific interactions)
+were removed one-by-one and in combination. To remove interspecific
+interactions, we set the off-diagonals of the interaction matrix for
+each vital rate regression to zero. This retains intraspecific
+interactions, and thus density-dependence, and results in simulations
+where species are growing in isolation. We cannot definitively rule out
+the effects of species interactions on all parameters, meaning that a
+true monoculture could behave differently than our simulations of a
+population growing without interspecific competitors. To remove the
+effect of a fluctuating environment, we removed the temporal
+(interannual) random effects from the regression equations. To remove
+the effect of demographic stochasticity, we use the IPM rather than the
+IBM because the IPM does not include demographic stochasticity
 (demographic stochasticity cannot be removed from the IBM). Since the
@@
 drivers (environmental stochasticity, demographic stochasticity, or
-\textcolor{blue}{interspecific interactions}) present; (2) IPM:
-Demographic stochasticity removed; (3) IBM: Environmental stochasticity
-removed; (4) IBM: \textcolor{blue}{Interspecific interactions} removed;
-(5) IPM: \textcolor{blue}{Interspecific interactions} and demographic
-stochasticity removed; (6) IBM:
-\textcolor{blue}{Interspecific interactions} and environmental
-stochasticity removed. We ran IPM simulations for 2,000 time steps,
-after an initial 500 iteration burn-in period. This allowed species time
-to reach their stable size distribution. We then calculated the
-synchrony of species' per capita growth rates over 100 randomly selected
-contiguous 50 time-step sections. We ran IBM simulations for 100 time
-steps, and repeated the simulations 100 times for each simulation
-experiment. From those, we retained only the simulations in which no
-species went extinct due to demographic stochasticity. Synchrony was
-calculated over the 100 time steps for each no extinction run within a
-model experiment. To explore the effect of demographic stochasticity in
-different sized populations, we ran simulations (1) and (6) on plot
-sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM simulations
-were run on a 25 \(\text{m}^2\) landscape.
-
-Results from our simulation experiments also allow us to test our
-theoretical predictions. First, in the absence of
-\textcolor{blue}{interspecific interactions} and demographic
-stochasticity,
-\textcolor{blue}{populations can only fluctuate in response to the environment}.
-Therefore, we can use results from simulation (5) to estimate the
-covariance of species' responses to the environment
+interspecific interactions) present; (2) IPM: Demographic stochasticity
+removed; (3) IBM: Environmental stochasticity removed; (4) IBM:
+Interspecific interactions removed; (5) IPM: Interspecific interactions
+and demographic stochasticity removed; (6) IBM: Interspecific
+interactions and environmental stochasticity removed.
+\textcolor{blue}{We did not include a simulation with only interspecific interactions because our population models run to deterministic equilibriums in the absence of environmental or demographic stochasticity.}
+We ran IPM simulations for 2,000 time steps, after an initial 500
+iteration burn-in period. This allowed species to reach their stable
+size distribution. We then calculated yearly per capita growth rates
+from the simlated time series, and then caluclated the synchrony of
+species' per capita growth rates over 100 randomly selected contiguous
+50 time-step sections.
+
+We ran IBM simulations for 100 time steps, and repeated the simulations
+100 times for each simulation experiment. From those, we retained only
+the simulations in which no species went extinct due to demographic
+stochasticity. Synchrony of per capita growth rates was calculated over
+the 100 time steps for each no extinction run within a model experiment.
+\textcolor{blue}{In the IBM simulations, the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
+To explore this effect}, we ran simulations \textcolor{blue}{(4)} and
+(6) on plot sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM
+simulations were run on a 25 \(\text{m}^2\) landscape
+\textcolor{blue}{to most closely match the implicit, large spatial scale of the IPM simulations}.
+
+\textcolor{blue}{Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulated values of community-wide species synchrony.}
+The simulation experiments also allow us to test our theoretical
+predictions. First, in the absence of interspecific interactions and
+demographic stochasticity, populations can only fluctuate in response to
+the environment. Therefore, we can use results from simulation (5) to
+estimate the covariance of species' responses to the environment
 (\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation 4.
@@
 simulations (5) and (6) represent the simulated version of our limiting
-case theoretical predictions. Thus, we directly test the theoretical
-predictions by comparing them to observed synchrony and simulated
-synchrony.
-
-\subsubsection{Results}\label{results}
-
-Synchrony of species' per capita growth rates at our study sites range
-from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to
-0.92 (Table 2). Synchrony tends to be higher in communities with few
-species (Arizona and New Mexico) and/or with relatively high temporal
-variability in percent cover (e.g., Montana). Synchrony is lowest in
-Idaho, the only data set that includes two life forms: a shrub and three
-perennial grasses (Table 1).
-\textcolor{blue}{Synchrony of per capita growth rates and CV of percent cover are positively correlated}
-(Pearson's \(\rho\) = 0.72). For all five communities, species synchrony
-from IPM and IBM simulations closely approximated observed synchrony
-(Fig. S1). IBM-simulated synchrony is consistently, but only slightly,
-lower than IPM-simulated synchrony (Fig. S1), likely due to the
-desynchronizing effect of demographic stochasticity.
+case theoretical predictions.
+\textcolor{blue}{This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalence in species' growth rates, environmental response variances, and demographic variances.
+Our empirically-based models do not make these assumptions.
+Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions hold in the natural community our model represents.}
+
+\subsection{Results}
+
+\textcolor{blue}{Observed} synchrony of species' per capita growth rates
+at our study sites range from 0.36 to 0.89 and synchrony of percent
+cover ranged from 0.15 to 0.92 (Table 2). Synchrony of per capita growth
+rates and CV of percent cover are positively correlated (Pearson's
+\(\rho\) = 0.72, \emph{N} = 5). For all five communities, species
+synchrony from IPM and IBM simulations closely approximated observed
+synchrony (Fig. S1). IBM-simulated synchrony is consistently, but only
+slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to
+the desynchronizing effect of demographic stochasticity.
 
 Across the five communities, our limiting case predictions closely
-matched synchrony from the corresponding simulation experiment (Fig. 1
+matched synchrony from the corresponding simulation experiment (Fig. 2
 and Table S1). The correlation between our analytical predictions and
-simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) and 0.997
-for \(\phi_{R,\mathcal{M}_E}\). The largest difference between predicted
-and simulated synchrony was 0.05 in New Mexico for
-\(\phi_{R,\mathcal{M}_D}\) (Table S1).
+simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) (\emph{N} =
+5) and 0.997 for \(\phi_{R,\mathcal{M}_E}\) (\emph{N} = 5). The largest
+difference between predicted and simulated synchrony was 0.05 in New
+Mexico for \(\phi_{R,\mathcal{M}_D}\) (Table S1).
 
@@
 in New Mexico. Only in Idaho did removing environmental fluctuations
-cause an increase in synchrony (Fig. 1), but the effect was small (9\%
+cause an increase in synchrony (Fig. 2), but the effect was small (9\%
 increase; Table S2). Overall, species' temporal random effects in the
-statistical vital rate models are positively correlated (Table S3).
+statistical vital rate models are positively, but not perfectly,
+correlated (Table S3).
+\textcolor{blue}{These temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.}
+
 Species interactions are weak in these communities (Table S4 and Chu and
-Adler 2015), and removing \textcolor{blue}{interspecific interactions}
-had little effect on synchrony (Fig. 1; ``No Comp.'' simulations).
-Removing \textcolor{blue}{interspecific interactions} caused, at most, a
-5\% change in synchrony (Fig. 1). Removing demographic stochasticity
-(``No D.S.'' simulations) caused synchrony to increase slightly in all
-communities (Fig. 1), with an average 6\% increase over synchrony from
-IBM simulations on a 25\(\text{m}^2\) area.
-
-The desynchronizing effect of demographic stochasticity, which increases
-as population size decreases, modestly counteracted the synchronizing
-force of the environment, but not enough to lower synchrony to the level
-observed when only demographic stochasticity is operating (Fig. 2). In
-the largest, 25 \(\text{m}^2\) plots, synchrony was driven by
-environmental stochasticity (e.g., \(\mathcal{M}_E\)). At 1
-\(\text{m}^2\), synchrony reflected demographic stochasticity and
-environmental stochasticity (e.g., between \(\mathcal{M}_E\) and
+Adler 2015), and removing interspecific interactions had little effect
+on synchrony (Fig. 2; ``No Comp.'' simulations). Removing interspecific
+interactions caused, at most, a 5\% change in synchrony (Fig. 2 and
+Table S2). Removing demographic stochasticity (``No D.S.'' simulations)
+caused synchrony to increase slightly in all communities (Fig. 2), with
+an average 6\% increase over synchrony from IBM simulations on a
+25\(\text{m}^2\) area.
+
+In IBM simulations, the desynchronizing effect of demographic
+stochasticity, which increases as population size decreases, modestly
+counteracted the synchronizing force of the environment, but not enough
+to lower synchrony to the level observed when only demographic
+stochasticity is operating (Fig. 3). In the largest, 25 \(\text{m}^2\)
+plots, synchrony was driven by environmental stochasticity (e.g.,
+\(\mathcal{M}_E\)). At 1 \(\text{m}^2\), synchrony reflected
+\textcolor{blue}{the combined effects of} demographic stochasticity and
+environmental stochasticity (e.g.,
+\textcolor{blue}{simulated synhcrony fell} between \(\mathcal{M}_E\) and
 \(\mathcal{M}_D\)). For context, population sizes increased from an
@@
 
-For all five communities, the synchrony of species' growth rates when
-growing in isolation almost perfectly matched species synchrony in
-polyculture (Fig. 3). Results for synchrony of percent cover are
-qualitatively similar, but simulation results were more variable and
-less consistent with analytical predictions and observed synchrony
-(Online Supporting Information, Figs. S2-S3).
-
-\subsection{DISCUSSION}\label{discussion}
+Results for synchrony of percent cover are qualitatively similar, but
+more variable and less consistent with analytical predictions and
+observed synchrony (Appendix S1, Figs. S2-S3).
+
+\section{Discussion}
 
@@
 assumption for abundance than growth rates, especially given relatively
-short time-series (Online Supporting Information). Thus, our results
-provide further evidence that it is difficult to decipher mechanisms of
-species synchrony from abundance time series, as expected by theory
-(Loreau and {{de Mazancourt}} 2008).
-\textcolor{blue}{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}
-
-\subsubsection{Simulations support theoretical
-predictions}\label{simulations-support-theoretical-predictions}
+short time-series (Appendix S1). Thus, our results provide further
+evidence that it is difficult to decipher mechanisms of species
+synchrony from abundance time series, as expected by theory (Loreau and
+{{de Mazancourt}} 2008). Observed synchrony of per capita growth rates
+was positively correlated with the variability of percent cover across
+our focal communities, which confirms that we are investigating an
+important process underlying ecosystem stability.
+
+\subsection{Simulations support theoretical predictions}
 
 Our theoretical predictions were derived from a simple model of
-population dynamics and required several simplifying assumptions,
-raising questions about their relevance to natural communities.
-\textcolor{blue}{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
-However, the theoretical predictions
-\textcolor{blue}{closely matched results from} simulations of population
-models fit to long-term data from natural plant communities (Table 3).
-Such \textcolor{blue}{strong agreement} between our analytical
-predictions and the simulation results should inspire confidence in the
-ability of simple models to inform our understanding of species
-synchrony
-\textcolor{blue}{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.
-
-\subsubsection{Demographic stochasticity decreases
-synchrony}\label{demographic-stochasticity-decreases-synchrony}
-
-Demographic stochasticity partially counteracted the synchronizing
-effects of environmental fluctuations and interspecific interactions on
-per capita growth rates, but only when populations were small (Fig. 2),
-in agreement with theory (Loreau and {{de Mazancourt}} 2008). Even in
-small populations, however, demographic stochasticity was not strong
-enough to \textcolor{blue}{compensate} the synchronizing effects of
-environmental fluctuations and match the analytical prediction where
-only demographic stochasticity is operating (\(\mathcal{M}_D\) in Fig.
-2). These results confirm the theoretical argument by Loreau and {{de
-Mazancourt}} (2008) that independent fluctuations among interacting
-species in a non-constant environment should be rare. Only in the Idaho
-community does synchrony of per capita growth rates approach
-\(\mathcal{M}_D\) in a non-constant environment (Fig. 2). This is most
-likely due to the strong effect of demographic stochasticity on the
-shrub \emph{Artemisia tripartita} since even a \textcolor{blue}{2}5
+population dynamics and required several restrictive assumptions,
+raising questions about their relevance to natural communities. For
+example, the species in our communities do not have equivalent
+environmental and demographic variances (Figs. S4-S7), as required by
+our predictions. However, the theoretical predictions closely matched
+results from simulations of population models fit to long-term data from
+natural plant communities (Table S1). Strong agreement between our
+analytical predictions and the simulation results should inspire
+confidence in the ability of simple models to inform our understanding
+of species synchrony even in complex natural communities, and allows us
+to place our simulation results within the context of contemporary
+theory.
+\textcolor{blue}{However, whether the theoretical model adequately represents more complex communities remains unknown because our analysis was restricted to dominant species.}
+
+\subsection{Demographic stochasticity decreases synchrony \textcolor{blue}{in small populations}}
+
+In large populations, removing demographic stochasticity had no effect
+on species synchrony (Fig. 2). In small populations, demographic
+stochasticity partially counteracted the synchronizing effects of
+environmental fluctuations and interspecific interactions on per capita
+growth rates, in agreement with theory (Loreau and {{de Mazancourt}}
+2008).
+\textcolor{blue}{This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
+The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, reducing the relative effect of environmental forcing.}
+Even in small populations
+\textcolor{blue}{(e.g., 1 $\text{m}^2$ lanscapes)}, however, demographic
+stochasticity was not strong enough to compensate the synchronizing
+effects of environmental fluctuations and match the analytical
+prediction where only demographic stochasticity is operating
+(\(\mathcal{M}_D\) in Fig. 3). These results confirm the theoretical
+argument of Loreau and {{de Mazancourt}} (2008) that independent
+fluctuations among interacting species in a non-constant environment
+should be rare. Only in the Idaho community does synchrony of per capita
+growth rates approach \(\mathcal{M}_D\) in a non-constant environment
+(Fig. 3). This is most likely due to the strong effect of demographic
+stochasticity on the shrub \emph{Artemisia tripartita} since even a 25
 \(\text{m}^2\) quadrat would only contain a few individuals of that
@@
 
-A challenge to the generality of our results is that we were only able
-to model common, co-occuring species (see Chu and Adler 2015).
-\textcolor{blue}{Most communities are dominated by few common species and many rare species}
-(McGill et al. 2007),
-\textcolor{blue}{meaning that the low number of common species in our focal commiunities is not unusual.}
-\textcolor{blue}{Rather, the generality of our results hinges upon the influence of rare species.
-Rare species could be kept at low densities by competitive suppression due to strong interspecific interactions.
-If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
-However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species}
-(Yenni et al. 2012),
-\textcolor{blue}{in which case synchrony would remain predominantly driven by environmental responses.}
-Rare species could also be limited by generalist natural enemies whose
-effects are density-independent but alter competitive hierarchies
-(Mordecai 2011). Under such conditions, synchrony will depend on the
-degree of pathogen-induced fitness differences and the pathogen's
-response to environmental conditions. Neither our model nor current
-theory explicitly considers the effects of pathogens on species
-synchrony, and this highlights the need for theoretical work on the
-interaction between mechanisms of species coexistence and mechanisms of
-species synchrony (Loreau 2010).
-
-\subsubsection{Interspecific interactions had little impact on community
-synchrony}\label{interspecific-interactions-had-little-impact-on-community-synchrony}
+\textcolor{blue}{Environmental responses synchronized dynamics relative to a null expectation of independent species interactions (e.g., "No Comp.
++ No E.S." simulations in Fig.
+2), but observed and simulated synchrony was still less than one in all cases (Fig.
+2).
+Synchrony was far from complete because of differences in species' responses to interannual environmental variation.
+Many studies of ecosystem stability in semiarid grasslands focus on trad-offs among dominant functional types}
+(Bai et al. 2004, Sasaki and Lauenroth 2011).
+\textcolor{blue}{Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
+At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were positively correlated.
+However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 2).
+The subtle differences among ecologically-similar dominant species do impact species synchrony and, ultimately, ecosystem stability.
+Ignoring such differences could mask important dynamics that underpin ecosystem functioning.}
+
+\subsection{Interspecific interactions had little impact on community synchrony}
 
@@
 
-However, our results do contrast with a recent analysis of several
-biodiversity-ecosystem functioning experiments showing that competition
-drives species synchrony in grasslands (Gross et al. 2014). The apparent
-inconsistency between our results and those of Gross et al. (2014) may
-be explained by the differences between our studies. Gross et al.'s
-results are based on rapidly assembling experimental communities that
-are, at most, 11 years old. The natural, relatively undisturbed
-communities we studied consist of species that have co-occurred for
-decades (Chu and Adler 2015) and represent a much later stage of
-community assembly. In theory, species interactions should weaken
-through time as community assembly proceeds (Kokkoris et al. 1999),
-meaning that in biodiversity-ecosystem functioning experiments the
-processes driving competitive exclusion are still operating. Such
-communities may be highly influenced by transient, but strong,
-interspecific competition that can mask the synchronizing effects of the
-environment. In contrast, synchrony may be driven by environmental
-fluctuations in older communities. In other words, the dominant driver
-of community synchrony should shift from competition to environmental
-fluctuations through time. One way to test this prediction is to
-continue collecting data from biodiversity-ecoystem functioning
-experiments and conduct the analysis of Gross et al. (2014) every few
-years.
-
-\textcolor{blue}{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
-Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap}
-(Vasseur and Fox 2007), \textcolor{blue}{dispersal} (Gouhier et al.
-2010), \textcolor{blue}{and the strength of top-down control} (Bauer et
-al. 2014)
-\textcolor{blue}{are all likely to affect community synchrony.}
-
-\subsection{CONCLUSIONS}\label{conclusions}
+Our conclusion that species interactions have little impact on synchrony
+only applies to single trophic level communities. Species interactions
+almost certainly play a strong role in multi-trophic communities where
+factors such as resource overlap (Vasseur and Fox 2007), dispersal
+(Gouhier et al. 2010), and the strength of top-down control (Bauer et
+al. 2014) are all likely to affect community synchrony.
+
+\section{Conclusions}
 
@@
 the Office of Research and Graduate Studies at Utah State University are
-gratefully acknowledged. We also thank three anonymous reviewers who
-provided critical feedback that improved the manuscript.
-
-\setlength{\parindent}{0ex}
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
-Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
-perennial plants: An embarrassment of niches. Ecology Letters
-13:1019--1029.
-
-Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
-mapped quadrats from Kansas prairie: demographic information for
-herbaceous plants. Ecology 88:2673.
-
-Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
-of semi-desert grassland plants in permanent quadrats mapped from 1915
-to 1947. Ecology 93:1492--1492.
-
-Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
-mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
-92:1703.
-
-Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
-Functional Similarity, and Top-Down Control Drive Synchronization and
-the Reliability of Ecosystem Function. The American Naturalist
-183:394--409.
-
-Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
-P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
-Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
-and its impact on humanity. Nature 489:59--67.
-
-Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
-recruitment stage to stabilize grassland coexistence. Ecological
-Monographs 85:373--392.
-
-{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
-Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
-Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
-Loreau. 2013. Predicting ecosystem stability from community composition
-and biodiversity. Ecology Letters 16:617--625.
-
-Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
-Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
-biodiversity change but not systematic loss. Science 344:296--9.
-
-Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
-Compensatory Dynamics in Ecological Communities. Annual Review of
-Ecology, Evolution, and Systematics 40:393--414.
-
-Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
-variability in space and time. Methods in Ecology and Evolution
-5:524--533.
-
-Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
-Stability of Food Webs in Metacommunities. The American Naturalist
-175:E16--E34.
-
-Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
-Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
-temporal stability of biomass production: a new analysis of recent
-biodiversity experiments. The American naturalist 183:1--12.
-
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
-
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
-
-Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
-on ecosystem functioning: a consensus of current knowledge. Ecological
-Monographs 75:3--35.
-
-Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
-Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
-Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
-Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
-Compensatory dynamics are rare in natural ecological communities.
-Proceedings of the National Academy of Sciences of the United States of
-America 104:3273--3277.
-
-Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
-productivity and the temporal stability of productivity: Patterns and
-processes. Ecology Letters 12:443--451.
-
-Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
-species interaction strength in assembled theoretical competition
-communities. Ecology Letters 2:70--74.
-
-Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
-dynamics in ecology and conservation.
-
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
-
-Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
-for a New Ecological Synthesis.
-
-Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
-drivers: neutral and nonneutral community dynamics in fluctuating
-environments. The American Naturalist 172:E48--E66.
-
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
-
-Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
-Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
-and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
-knowledge and future challenges. Science 294:804--808.
-
-McGill, B. J., R. S. Etienne, J. S. Gray, D. Alonso, M. J. Anderson, H.
-K. Benecha, M. Dornelas, B. J. Enquist, J. L. Green, F. He, A. H.
-Hurlbert, A. E. Magurran, P. A. Marquet, B. A. Maurer, A. Ostling, C. U.
-Soykan, K. I. Ugland, and E. P. White. 2007. Species abundance
-distributions: Moving beyond single prediction theories to integration
-within an ecological framework. Ecology Letters 10:995--1015.
-
-Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
-theory, concepts, and empirical work. Ecological Monographs 81:429--441.
-
-Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
-community dynamics? Proceedings of the Royal Society B: Biological
-Sciences 276:2923--2929.
-
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
-
-R Core Team. 2013. R: A language and environment for statistical
-computing.
-
-Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
-O'Hara. 2008. Detecting compensatory dynamics in competitive communities
-under environmental forcing. Oikos 117:1907--1911.
-
-Rees, M., and S. P. Ellner. 2009. Integral projection models for
-populations in temporally varying environments. Ecological Monographs
-79:575--594.
-
-Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
-projection models: A user's guide.
-
-Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
-species have similar dynamics. Journal of Ecology 99:1453--1459.
-
-Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
-A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
-Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
-C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
-Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
-Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
-Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
-Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
-Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
-Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
-Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
-operating space for humanity. Nature 461:472--475.
-
-Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
-for Latent Gaussian Models Using Integrated Nested Laplace
-Approximations. Journal of the Royal Statistical Society, Series B
-71:319--392.
-
-Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
-Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
-analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
-
-Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
-189:732--735.
-
-Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
-and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
-
-Tilman, D. 1988. Plant strategies and the dynamics and structure of
-plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
-
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
-
-Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
-2001. Diversity and productivity in a long-term grassland experiment.
-Science 294:843--845.
-
-Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
-Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
-W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
-2014. Synchronous dynamics of zooplankton competitors prevail in
-temperate lake ecosystems. Proceedings. Biological sciences / The Royal
-Society 281:20140633.
-
-Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
-stabilize food web dynamics by increasing synchrony. Ecology Letters
-10:1066--1074.
-
-Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
-2013. Global meta-analysis reveals no net change in local-scale plant
-biodiversity over time. Proceedings of the National Academy of Sciences
-of the United States of America 110:19456--9.
-
-Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
-Human Domination of Earth' s Ecosystems. Science 277:494--499.
-
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
-
-Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
-
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
-
-Yenni, G., P. B. Adler, and S. K. {Morgan Ernest}. 2012. Strong
-self-limitation promotes the persistence of rare species. Ecology
-93:456--461.
-
-Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
-sagebrush steppe: long-term data for analyzing demographic rates and
-plant--plant interactions. Ecology 91:3427.
-
-\newpage{}
-
-\subsection{Tables}\label{tables}
+gratefully acknowledged. We also thank Patrick Venail and several
+anonymous reviewers who provided critical feedback that improved the
+manuscript.
+
+\setlength{\parindent}{0ex} \newpage{}
+
+\section{Tables}
 
@@
       \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
-  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S.
-Only'') and when environmental stochasticity is also present removed (``D.S.
-+ E.S.'').
+  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
++ No E.S.
+(D.S.
+Only)'') and when environmental stochasticity is also present (``No Comp.
+(D.S.
++ E.S.)'').
 The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line).
 The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase.
-Theoretically, ``D.S.
-Only'' simulations should remain constant across landscape size, whereas ``D.S.
-+ E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-\textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}}
+Theoretically, ``No Comp.
++ No E.S.
+(D.S.
+Only)'' simulations should remain constant across landscape size, whereas ``No Comp.
+(D.S.
++ E.S.)'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
 \end{figure}
@@
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/formatted_figures/formatted_figure3.png}
-  \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions).
-We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments.
-The dashed line is the line of equality.
-Simulation results in this figure are analagous to ``No Comp.
-+ No D.S.'' (species in isolation) and ``No D.S.'' (species in competition) in Fig.
-1, but here we control the time series of random year effects.}
-\end{figure}
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
+perennial plants: An embarrassment of niches. Ecology Letters
+13:1019--1029.
+
+Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
+mapped quadrats from Kansas prairie: demographic information for
+herbaceous plants. Ecology 88:2673.
+
+Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
+of semi-desert grassland plants in permanent quadrats mapped from 1915
+to 1947. Ecology 93:1492--1492.
+
+Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
+mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
+92:1703.
+
+Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
+and compensatory effects in the Inner Mongolia grassland. Nature
+431:181--184.
+
+Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
+Functional Similarity, and Top-Down Control Drive Synchronization and
+the Reliability of Ecosystem Function. The American Naturalist
+183:394--409.
+
+Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
+P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
+Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
+and its impact on humanity. Nature 489:59--67.
+
+Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
+recruitment stage to stabilize grassland coexistence. Ecological
+Monographs 85:373--392.
+
+{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
+Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
+Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
+Loreau. 2013. Predicting ecosystem stability from community composition
+and biodiversity. Ecology Letters 16:617--625.
+
+Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
+Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
+biodiversity change but not systematic loss. Science 344:296--9.
+
+Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. Annual Review of
+Ecology, Evolution, and Systematics 40:393--414.
+
+Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
+variability in space and time. Methods in Ecology and Evolution
+5:524--533.
+
+Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
+Stability of Food Webs in Metacommunities. The American Naturalist
+175:E16--E34.
+
+Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
+Mechanisms contributing to stability in ecosystem function depend on the
+environmental context. Ecology Letters 13:1400--1410.
+
+Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
+Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
+temporal stability of biomass production: a new analysis of recent
+biodiversity experiments. The American naturalist 183:1--12.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
+on ecosystem functioning: a consensus of current knowledge. Ecological
+Monographs 75:3--35.
+
+Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
+Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
+Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
+Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
+Compensatory dynamics are rare in natural ecological communities.
+Proceedings of the National Academy of Sciences of the United States of
+America 104:3273--3277.
+
+Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
+productivity and the temporal stability of productivity: Patterns and
+processes. Ecology Letters 12:443--451.
+
+Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
+dynamics in ecology and conservation.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
+drivers: neutral and nonneutral community dynamics in fluctuating
+environments. The American Naturalist 172:E48--E66.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+Loreau, M., and a Hector. 2001. Partitioning selection and
+complementarity in biodiversity experiments. Nature 412:72--6.
+
+Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
+Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
+and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
+knowledge and future challenges. Science 294:804--808.
+
+Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
+community dynamics? Proceedings of the Royal Society B: Biological
+Sciences 276:2923--2929.
+
+Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
+Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
+3rd international workshop on distributed statistical computing (dSC
+2003). march.
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+
+R Core Team. 2013. R: A language and environment for statistical
+computing.
+
+Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
+O'Hara. 2008. Detecting compensatory dynamics in competitive communities
+under environmental forcing. Oikos 117:1907--1911.
+
+Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
+A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
+Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
+C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
+Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
+Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
+Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
+Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
+Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
+Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
+Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
+operating space for humanity. Nature 461:472--475.
+
+Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
+for Latent Gaussian Models Using Integrated Nested Laplace
+Approximations. Journal of the Royal Statistical Society, Series B
+71:319--392.
+
+Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
+Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
+analysis of latent Gaussian models using Integrated Nested Laplace
+Approximaxion.
+
+Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
+diversity, regulates temporal stability of plant communities. Oecologia
+166:761--768.
+
+Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
+function with non-random species loss. Ecology Letters 6:509--517.
+
+Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
+189:732--735.
+
+Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
+and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
+
+Tilman, D. 1988. Plant strategies and the dynamics and structure of
+plant communities. Plant strategies and the dynamics and structure of
+plant communities.:360 pp.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
+2001. Diversity and productivity in a long-term grassland experiment.
+Science 294:843--845.
+
+Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
+stabilize food web dynamics by increasing synchrony. Ecology Letters
+10:1066--1074.
+
+Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
+biodiversity over time. Proceedings of the National Academy of Sciences
+of the United States of America 110:19456--9.
+
+Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
+2013. Shared ancestry influences community stability by altering
+competitive interactions: evidence from a laboratory microcosm
+experiment using freshwater green algae. Proceedings of the Royal
+Society of London B: Biological Sciences 280.
+
+Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
+Human Domination of Earth' s Ecosystems. Science 277:494--499.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
+across scales in metacommunities. Ecology Letters.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
+sagebrush steppe: long-term data for analyzing demographic rates and
+plant--plant interactions. Ecology 91:3427.
 

### 2017-02-02 docs/ms_communitySynchrony_ForECOLOGY_PREPRINT.tex
more preprint stuff

--- 2017-01-25 docs/ms_communitySynchrony_ForECOLOGY.tex
@@
 %%\linenumbers
-\usepackage{setspace}
 \usepackage{todonotes}
-\doublespacing
 \usepackage{rotating}
@@
 
-\begin{singlespace}
-
 \begin{center}
@@
 
-\vspace{2em}
-
-\textbf{Running Head}: Species synchrony in grasslands
-
-\textbf{Corresponding Author}: \\
-\hspace{2em}Andrew Tredennick \\  
-\hspace{2em}Department of Wildland Resources and the Ecology Center\\
-\hspace{2em}Utah State University \\
-\hspace{2em}5230 Old Main Hill \\
-\hspace{2em}Logan, Utah 84322 USA \\
-\hspace{2em}Phone: +1-970-443-1599 \\
-\hspace{2em}Fax: +1-435-797-3796 \\
-\hspace{2em}Email: atredenn@gmail.com
-
-\end{singlespace}\renewcommand*{\thefootnote}{\arabic{footnote}}
+\renewcommand*{\thefootnote}{\arabic{footnote}}
 
 \setcounter{footnote}{0}
-
-\newpage{}
 
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/synchrony_flowchart_bw.png}
+  \caption{Diagram of our coupled theoretical-empirical approach. We followed this workflow for each of our five focal communities.}
+\end{figure}
+
 To achieve our objectives, we first refine theory that has been used to
@@
 
-\paragraph{Calculating observed
-synchrony.--}\label{calculating-observed-synchrony.}
-
-The data consist of records for individual plant size in quadrats for
-each year. To obtain estimates of percent cover for each focal species
-in each year, we summed the individual-level data within quadrats and
-then averaged percent cover, by species, over all quadrats. We
-calculated per capita growth rates as
-\(\text{log}(x_t) - \text{log}(x_{t-1})\), where \emph{x} is species'
-percent cover in year \emph{t}. Using the community time series of per
-capita growth rates or percent cover, we calculated community synchrony
-using the metric of Loreau and {{de Mazancourt}} (2008) in the
-\texttt{synchrony} package (Gouhier and Guichard 2014) in \texttt{R} (R Core Team 2013).
-Specifically, we calculated synchrony as
-
-\begin{equation}
-\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
-\end{equation}
-
-\noindent where \(\sigma_{r_{i}}\) is the temporal
-standard deviation of species \emph{i'}s per capita
-population growth rate (\(r_i\)) and \(\sigma^{2}_{T}\) is the temporal
-variance of the aggregate community-level growth rate. \(\phi\) ranges
-from 0 at perfect asynchrony to 1 at perfect synchrony (Loreau and {{de
-Mazancourt}} 2008). We use the same equation to calculate observed
-synchrony of species' percent cover, which we present to relate our
-results to previous findings, even though we focus on synchrony of
-growth rates in our model simulations (see below).
-
-\paragraph{Fitting statistical
-models.--}\label{fitting-statistical-models.}
-
-Vital rate regressions are the building blocks of our dynamic models: an
-integral projection model (IPM) and an individual-based model (IBM). We
-followed the approach of Chu and Adler (2015) to fit statistical models
-for survival, growth, and recruitment (see Appendix S1 for full
-details). We modeled survival probability of each genet as function of
-genet size, temporal variation among years, permanent spatial variation
-among groups of quadrats, and local neighborhood crowding from
-conspecific and heterospecific genets. Regression coefficients for the
-effect of crowding by each species can be considered a matrix of
-interaction coefficients whose diagonals represent intraspecific
-interactions and whose off-diagonals represent interspecific
-interactions (Adler et al. 2010). These interaction coefficients can
-take positive (facilitative) or negative (competitive) values. We
-modeled growth as the change in size of a genet from one year to the
-next, which depends on the same factors as the survival model. We fit
-the survival and growth regressions using INLA (Rue et al. 2014), a
-statistical package for fitting generalized linear mixed effects models
-via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
-2013). Spatial (quadrat groupings) variation was treated as a random
-effect on the intercept and temporal (interannual) variation was treated
-as random effects on the intercept and the effect of genet size in the
-previous year (Appendix S1).
-
-Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
-Strong year-to-year variation in these crowding effects would indicate a statistical interaction between environmental conditions and species interactions.
-We tested for such a dynamic by comparing models with and without random year effects on crowding.
-Based on the results, we decided to treat crowding as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including interannual variation in crowding did not change our results.
-
-We modeled recruitment at the quadrat scale, rather than the individual
-scale, because the original data do not attribute new genets to specific
-parents (Chu and Adler 2015). Our recruitment model assumes that the
-number of recruits produced in each year follows a negative binomial
-distribution with the mean dependent on the cover of the parent species,
-permanent spatial variation among groups, temporal variation among
-years, and inter- and intraspecific interactions as a function of total
-species' cover in the quadrat. We fit the recruitment model using a
-hierarchical Bayesian approach implemented in JAGS (Plummer 2003) using
-the \texttt{rjags} package (Plummer 2014) in \texttt{R} (R Core Team 2013). Again,
-temporal and spatial variation were treated as random effects.
-
-\paragraph{Building dynamic multi-species
-models.--}\label{building-dynamic-multi-species-models.}
-
-Once we have fit the vital rate statistical models, building the
-population models is straightforward. For the IBM, we initialize the
-model by randomly assigning plants spatial coordinates, sizes, and
-species identities until each species achieves a density representative
-of that observed in the data. We then project the model forward by using
-the survival regression to determine whether a genet lives or dies, the
-growth regression to calculate changes in genet size, and the
-recruitment regression to add new individuals that are distributed
-randomly in space. Crowding is directly calculated at each time step
-since each genet is spatially referenced. Environmental stochasticity is
-not an inherent feature of IBMs, but is easily included since we fit
-year-specific temporal random effects for each vital rate regression. To
-include temporal environmental variation, at each time step we randomly
-choose a set of estimated survival, growth, and recruitment parameters
-specific to one observation year. For all simulations, we ignore the
-spatial random effect associated with variation among quadrat groups, so
-our simulations represent an average quadrat for each site.
-
-The IPM uses the same vital rate regressions as the IBM, but it is
-spatially implicit and does not include demographic stochasticity.
-Following Chu and Adler (2015), we use a mean field approximation that
-captures the essential features of spatial patterning to define the
-crowding index at each time step (Supporting Online Information).
-Temporal variation is included in exactly the same way as for the IBM.
-For full details on the IPMs we use, see Chu and Adler (2015).
-
-\paragraph{Simulation experiments.--}\label{simulation-experiments.}
-
-We performed simulation experiments where drivers (environmental
-stochasticity, demographic stochasticity, or interspecific interactions)
-were removed one-by-one and in combination. To remove interspecific
-interactions, we set the off-diagonals of the interaction matrix for
-each vital rate regression to zero. This retains intraspecific
-interactions, and thus density-dependence, and results in simulations
-where species are growing in isolation. We cannot definitively rule out
-the effects of species interactions on all parameters, meaning that a
-true monoculture could behave differently than our simulations of a
-population growing without interspecific competitors. To remove the
-effect of a fluctuating environment, we removed the temporal
-(interannual) random effects from the regression equations. To remove
-the effect of demographic stochasticity, we use the IPM rather than the
-IBM because the IPM does not include demographic stochasticity
-(demographic stochasticity cannot be removed from the IBM). Since the
-effect of demographic stochasticity on population dynamics depends on
-population size (Lande et al. 2003), we can control the strength of
-demographic stochasticity by simulating the IBM on areas (e.g.~plots) of
-different size. Results from an IBM with infinite population size would
-converge on results from the IPM. Given computational constraints, the
-largest landscape we simulate is a 25 \(\text{m}^2\) plot.
-
-We conducted the following six simulation experiments: (1) IBM: All
-drivers (environmental stochasticity, demographic stochasticity, or
-interspecific interactions) present; (2) IPM: Demographic stochasticity
-removed; (3) IBM: Environmental stochasticity removed; (4) IBM:
-Interspecific interactions removed; (5) IPM: Interspecific interactions
-and demographic stochasticity removed; (6) IBM: Interspecific
-interactions and environmental stochasticity removed.
-We did not include a simulation with only interspecific interactions because our population models run to deterministic equilibriums in the absence of environmental or demographic stochasticity.
-We ran IPM simulations for 2,000 time steps, after an initial 500
-iteration burn-in period. This allowed species to reach their stable
-size distribution. We then calculated yearly per capita growth rates
-from the simulated time series, and then calculated the synchrony of
-species' per capita growth rates over 100 randomly selected contiguous
-50 time-step sections.
-
-We ran IBM simulations for 100 time steps, and repeated the simulations
-100 times for each simulation experiment. From those, we retained only
-the simulations in which no species went extinct due to demographic
-stochasticity. Synchrony of per capita growth rates was calculated over
-the 100 time steps for each no extinction run within a model experiment.
-In the IBM simulations, the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
-To explore this effect, we ran simulations (4) and
-(6) on plot sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM
-simulations were run on a 25 \(\text{m}^2\) landscape
-to most closely match the implicit, large spatial scale of the IPM simulations.
-
-Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulated values of community-wide species synchrony.
-The simulation experiments also allow us to test our theoretical
-predictions. First, in the absence of interspecific interactions and
-demographic stochasticity, populations can only fluctuate in response to
-the environment. Therefore, we can use results from simulation (5) to
-estimate the covariance of species' responses to the environment
-(\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation 4.
-Parameterizing Equation 3 does not require simulation output because the
-only parameters are the species' relative abundances. Second,
-simulations (5) and (6) represent the simulated version of our limiting
-case theoretical predictions.
-This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalence in species' growth rates, environmental response variances, and demographic variances.
-Our empirically-based models do not make these assumptions.
-Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions hold in the natural community our model represents.
-
-All \texttt{R} code necessary to reproduce our results has been deposited on Figshare (\url{http://doi.org/10.6084/m9.figshare.4587073}) and released on \textsc{GitHub} (\url{http://github.com/atredennick/community_synchrony/releases/tag/v1}).
-
-\subsection{Results}
-
-Observed synchrony of species' per capita growth rates
-at our study sites range from 0.36 to 0.89 and synchrony of percent
-cover ranged from 0.15 to 0.92 (Table 2). Synchrony of per capita growth
-rates and CV of percent cover are positively correlated (Pearson's
-\(\rho\) = 0.72, \emph{N} = 5). For all five communities, species
-synchrony from IPM and IBM simulations closely approximated observed
-synchrony (Fig. S1). IBM-simulated synchrony is consistently, but only
-slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to
-the desynchronizing effect of demographic stochasticity.
-
-Across the five communities, our limiting case predictions closely
-matched synchrony from the corresponding simulation experiment (Fig. 2
-and Table S1). The correlation between our analytical predictions and
-simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) (\emph{N} =
-5) and 0.997 for \(\phi_{R,\mathcal{M}_E}\) (\emph{N} = 5). The largest
-difference between predicted and simulated synchrony was 0.05 in New
-Mexico for \(\phi_{R,\mathcal{M}_D}\) (Table S1).
-
-Simulation experiments revealed that removing environmental fluctuations
-has the largest impact on synchrony, leading to a reduction in synchrony
-of species growth rates in four out of five communities (Fig. 2).
-Removing environmental fluctuations (``No E.S'' simulations) decreased
-synchrony by 33\% in Arizona, 48\% in Kansas, 39\% in Montana, and 40\%
-in New Mexico. Only in Idaho did removing environmental fluctuations
-cause an increase in synchrony (Fig. 2), but the effect was small (9\%
-increase; Table S2). Overall, species' temporal random effects in the
-statistical vital rate models are positively, but not perfectly,
-correlated (Table S3).
-These temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.
-
-Species interactions are weak in these communities (Table S4 and Chu and
-Adler 2015), and removing interspecific interactions had little effect
-on synchrony (Fig. 2; ``No Comp.'' simulations). Removing interspecific
-interactions caused, at most, a 5\% change in synchrony (Fig. 2 and
-Table S2). Removing demographic stochasticity (``No D.S.'' simulations)
-caused synchrony to increase slightly in all communities (Fig. 2), with
-an average 6\% increase over synchrony from IBM simulations on a
-25\(\text{m}^2\) area.
-
-In IBM simulations, the desynchronizing effect of demographic
-stochasticity, which increases as population size decreases, modestly
-counteracted the synchronizing force of the environment, but not enough
-to lower synchrony to the level observed when only demographic
-stochasticity is operating (Fig. 3). In the largest, 25 \(\text{m}^2\)
-plots, synchrony was driven by environmental stochasticity (e.g.,
-\(\mathcal{M}_E\)). At 1 \(\text{m}^2\), synchrony reflected
-the combined effects of demographic stochasticity and
-environmental stochasticity (e.g.,
-simulated synchrony fell between \(\mathcal{M}_E\) and
-\(\mathcal{M}_D\)). For context, population sizes increased from an
-average of 17 individuals per community in 1 \(\text{m}^2\) IBM
-simulations to an average of 357 individuals per community in 25
-\(\text{m}^2\) IBM simulations.
-
-Results for synchrony of percent cover are qualitatively similar, but
-more variable and less consistent with analytical predictions and
-observed synchrony (Appendix S1, Figs. S2-S3).
-
-\section{Discussion}
-
-Our study produced four main findings that were generally consistent
-across five natural grassland plant communities: (1) limiting-case predictions
-from the theoretical model were well-supported by simulations from the
-empirical models; (2) demographic stochasticity decreased community
-synchrony, as expected by theory, and its effect was largest in small
-populations; (3) environmental fluctuations increased community
-synchrony relative to simulations in constant environments because
-species-specific responses to the environment were positively, though
-not perfectly, correlated; and (4) interspecific interactions were weak
-and therefore had little impact on community synchrony. We also found
-that analyses based on synchrony of species' percent cover, rather than
-growth rates, were uninformative (Figs. S2-S3) since the linear
-approximation required for analytical predictions is a stronger
-assumption for abundance than growth rates, especially given relatively
-short time-series (Appendix S1). Thus, our results provide further
-evidence that it is difficult to decipher mechanisms of species
-synchrony from abundance time series, as expected by theory (Loreau and
-{{de Mazancourt}} 2008). Observed synchrony of per capita growth rates
-was positively correlated with the variability of percent cover across
-our focal communities, which confirms that we are investigating an
-important process underlying ecosystem stability.
-
-\subsection{Simulations support theoretical predictions}
-
-Our theoretical predictions were derived from a simple model of
-population dynamics and required several restrictive assumptions,
-raising questions about their relevance to natural communities. For
-example, the species in our communities do not have equivalent
-environmental and demographic variances (Figs. S4-S7), as required by
-our predictions. However, the theoretical predictions closely matched
-results from simulations of population models fit to long-term data from
-natural plant communities (Table S1). Strong agreement between our
-analytical predictions and the simulation results should inspire
-confidence in the ability of simple models to inform our understanding
-of species synchrony even in complex natural communities, and allows us
-to place our simulation results within the context of contemporary
-theory.
-However, whether the theoretical model adequately represents more complex communities remains unknown because our analysis was restricted to dominant species.
-
-\subsection{Demographic stochasticity decreases synchrony in small populations}
-
-In large populations, removing demographic stochasticity had no effect
-on species synchrony (Fig. 2). In small populations, demographic
-stochasticity partially counteracted the synchronizing effects of
-environmental fluctuations and interspecific interactions on per capita
-growth rates, in agreement with theory (Loreau and {{de Mazancourt}}
-2008).
-This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
-The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, reducing the relative effect of environmental forcing.
-Even in small populations
-(e.g., 1 $\text{m}^2$ lanscapes), however, demographic
-stochasticity was not strong enough to compensate the synchronizing
-effects of environmental fluctuations and match the analytical
-prediction where only demographic stochasticity is operating
-(\(\mathcal{M}_D\) in Fig. 3). These results confirm the theoretical
-argument of Loreau and {{de Mazancourt}} (2008) that independent
-fluctuations among interacting species in a non-constant environment
-should be rare. Only in the Idaho community does synchrony of per capita
-growth rates approach \(\mathcal{M}_D\) in a non-constant environment
-(Fig. 3). This is most likely due to the strong effect of demographic
-stochasticity on the shrub \emph{Artemisia tripartita} since even a 25
-\(\text{m}^2\) quadrat would only contain a few individuals of that
-species.
-
-Our analysis of how demographic stochasticity affects synchrony
-demonstrates that synchrony depends on the observation area. As the
-observation area increases, population size increases and the
-desynchronizing effect of demographic stochasticity lessens (Fig. 3).
-Thus, our results suggest that community-wide species synchrony will
-increase as the observation area increases, rising from
-\(\mathcal{M}_D\) to \(\mathcal{M}_E\). Such a conclusion assumes,
-however, that species richness remains constant as observation area
-increases, which is unlikely (Taylor 1961). Recent theoretical work has
-begun to explore the linkage between ecosystem stability and spatial
-scale (Wang and Loreau 2014, 2016), and our results suggest that
-including demographic stochasticity in theoretical models of
-metacommunity dynamics may be important for understanding the role of
-species synchrony in determining ecosystem stability across spatial
-scales.
-
-\subsection{Environmental fluctuations drive community synchrony}
-
-In large populations where interspecific interactions are weak,
-synchrony is expected to be driven exclusively by environmental
-fluctuations (Equation 4). Under such conditions community synchrony
-should approximately equal the synchrony of species' responses to the
-environment (Loreau and {{de Mazancourt}} 2008). Two lines of evidence
-lead us to conclude that environmental fluctuations drive species
-synchrony in our focal plant communities. First, in our simulation
-experiments, removing interspecific interactions resulted in no
-discernible change in community-wide species synchrony of per capita
-growth rates (Fig. 2). Second, removing environmental fluctuations from
-simulations consistently reduced synchrony (Fig. 2). Our results lead us
-to conclude that environmental fluctuations, not species interactions,
-are the primary driver of community-wide species synchrony
-among the dominant species we studied. Given
-accumulating evidence that niche differences in natural communities are
-large (reviewed in Chu and Adler 2015), and thus species interactions
-are likely to be weak, our results may be general in natural plant
-communities.
-
-In the Idaho community, removing environmental fluctuations did not
-cause a large decrease in synchrony. However, that result appears to be
-an artifact. Removing environmental variation results in a negative
-invasion growth rate for \emph{A. tripartita}. Although we only analyzed
-IBM runs in which \emph{A. tripartita} had not yet gone extinct, it was
-at much lower abundance than in the other simulation runs. When we
-removed \emph{A. tripartita} from all simulations, the Idaho results
-conformed with results from all other sites: removing environmental
-stochasticity caused a significant reduction in species synchrony (Fig.
-S8). Our main results for Idaho (Fig. 2), with \emph{A. tripartita}
-included, demonstrate how the processes that determine species synchrony
-interact in complex ways. \emph{A. tripartita} has a facilitative effect
-on each grass species across all vital rates, except for a small
-competitive effect on \emph{H. comata}'s survival probability (Tables
-S8-S10). At the same time, all the perennial grasses have negative
-effects on each other for each vital rate (Tables S8-S10). We know
-synchrony is affected by interspecific competition (Loreau and {{de
-Mazancourt}} 2008), but how facilitative effects manifest themselves is
-unknown. The interaction of facilitation and competition is clearly
-capable of having a large effect on species synchrony, and future
-theoretical efforts should aim to include a wider range of species
-interactions.
-
-Environmental responses synchronized dynamics relative to a null expectation of independent species interactions (e.g., "No Comp.
-+ No E.S." simulations in Fig.
-2), but observed and simulated synchrony was still less than one in all cases (Fig.
-2).
-Synchrony was far from complete because of differences in species' responses to interannual environmental variation.
-Many studies of ecosystem stability in semiarid grasslands focus on trad-offs among dominant functional types
-(Bai et al. 2004, Sasaki and Lauenroth 2011).
-Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
-At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were positively correlated.
-However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 2).
-The subtle differences among ecologically-similar dominant species do impact species synchrony and, ultimately, ecosystem stability.
-Ignoring such differences could mask important dynamics that underpin ecosystem functioning.
-
-\subsection{Interspecific interactions had little impact on community synchrony}
-
-We expected community synchrony of per capita growth rates to decrease
-when we removed interspecific interactions (Loreau and {{de Mazancourt}}
-2008). We found that community synchrony was virtually indistinguishable
-between simulations with and without interspecific interactions (Fig.
-2). The lack of an effect of interspecific interactions on synchrony is
-in contrast to a large body of theoretical work that predicts a strong
-role for competition in creating compensatory dynamics (Tilman 1988) and
-a recent empirical analysis (Gross et al. 2014).
-
-Our results do not contradict the idea that competition can lead to
-compensatory dynamics, but they do highlight the fact that interspecific
-competition must be relatively strong to influence species synchrony.
-The communities we analyzed are composed of species with very little
-niche overlap (Chu and Adler 2015) and weak interspecific interactions
-(Tables S1 and S3-S17). Mechanistic consumer-resource models (Lehman and
-Tilman 2000) and phenomenological Lotka-Volterra models (Lehman and
-Tilman 2000, Loreau and {{de Mazancourt}} 2013) both confirm that the
-effect of competition on species synchrony diminishes as niche overlap
-decreases. In that sense, our results are not surprising: interspecific
-interactions are weak, so of course removing them does not affect
-synchrony.
-
-Our conclusion that species interactions have little impact on synchrony
-only applies to single trophic level communities. Species interactions
-almost certainly play a strong role in multi-trophic communities where
-factors such as resource overlap (Vasseur and Fox 2007), dispersal
-(Gouhier et al. 2010), and the strength of top-down control (Bauer et
-al. 2014) are all likely to affect community synchrony.
-
-\section{Conclusions}
-
-Species-specific responses to temporally fluctuating environmental
-conditions is an important mechanism underlying asynchronous population
-dynamics and, in turn, ecosystem stability (Loreau and {{de Mazancourt}}
-2013). When we removed environmental variation, we found that synchrony
-decreased in four out of the five grassland communities we studied (Fig.
-2). A tempting conclusion is that our study confirms that compensatory
-dynamics are rare in natural communities, and that ecologically-similar
-species will exhibit synchronous dynamics (e.g., Houlahan et al. 2007).
-Such a conclusion misses an important subtlety. The perennial grasses we
-studied do have similar responses to the environment (Table S2), which
-will tend to synchronize dynamics. However, if community-wide synchrony
-among dominant species is less than perfect, as it is
-in all our focal communities, some degree of compensatory dynamics must
-be present (Loreau and {{de Mazancourt}} 2008).
-Even ecologically-similar species, which are sometimes aggregated into functional groups, have environmental responses that are dissimilar enough to limit synchrony.
-Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.
-
-\section{Acknowledgements}
-
-This work was funded by the National Science Foundation through a
-Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
-CAREER award to PBA (DEB-1054040). CdM and ML were supported by the
-TULIP Laboratory of Excellence (ANR-10-LABX-41). We thank the original
-mappers of the permanent quadrats at each site and the digitizers in the
-Adler lab, without whom this work would not have been possible. Compute,
-storage, and other resources from the Division of Research Computing in
-the Office of Research and Graduate Studies at Utah State University are
-gratefully acknowledged. We also thank Patrick Venail and several
-anonymous reviewers who provided critical feedback that improved the
-manuscript.
-
-\setlength{\parindent}{0ex}
-
-\section{Literature Cited}
-
-Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
-perennial plants: An embarrassment of niches. Ecology Letters
-13:1019--1029.
-
-Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
-mapped quadrats from Kansas prairie: demographic information for
-herbaceous plants. Ecology 88:2673.
-
-Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
-of semi-desert grassland plants in permanent quadrats mapped from 1915
-to 1947. Ecology 93:1492--1492.
-
-Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
-mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
-92:1703.
-
-Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
-and compensatory effects in the Inner Mongolia grassland. Nature
-431:181--184.
-
-Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
-Functional Similarity, and Top-Down Control Drive Synchronization and
-the Reliability of Ecosystem Function. The American Naturalist
-183:394--409.
-
-Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
-P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
-Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
-and its impact on humanity. Nature 489:59--67.
-
-Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
-recruitment stage to stabilize grassland coexistence. Ecological
-Monographs 85:373--392.
-
-{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
-Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
-Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
-Loreau. 2013. Predicting ecosystem stability from community composition
-and biodiversity. Ecology Letters 16:617--625.
-
-Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
-Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
-biodiversity change but not systematic loss. Science 344:296--9.
-
-Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
-Compensatory Dynamics in Ecological Communities. Annual Review of
-Ecology, Evolution, and Systematics 40:393--414.
-
-Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
-variability in space and time. Methods in Ecology and Evolution
-5:524--533.
-
-Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
-Stability of Food Webs in Metacommunities. The American Naturalist
-175:E16--E34.
-
-Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
-Mechanisms contributing to stability in ecosystem function depend on the
-environmental context. Ecology Letters 13:1400--1410.
-
-Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
-Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
-temporal stability of biomass production: a new analysis of recent
-biodiversity experiments. The American naturalist 183:1--12.
-
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
-
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
-
-Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
-on ecosystem functioning: a consensus of current knowledge. Ecological
-Monographs 75:3--35.
-
-Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
-Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
-Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
-Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
-Compensatory dynamics are rare in natural ecological communities.
-Proceedings of the National Academy of Sciences of the United States of
-America 104:3273--3277.
-
-Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
-productivity and the temporal stability of productivity: Patterns and
-processes. Ecology Letters 12:443--451.
-
-Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
-dynamics in ecology and conservation.
-
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
-
-Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
-drivers: neutral and nonneutral community dynamics in fluctuating
-environments. The American Naturalist 172:E48--E66.
-
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
-
-Loreau, M., and a Hector. 2001. Partitioning selection and
-complementarity in biodiversity experiments. Nature 412:72--6.
-
-Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
-Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
-and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
-knowledge and future challenges. Science 294:804--808.
-
-Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
-community dynamics? Proceedings of the Royal Society B: Biological
-Sciences 276:2923--2929.
-
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
-
-R Core Team. 2013. R: A language and environment for statistical
-computing.
-
-Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
-O'Hara. 2008. Detecting compensatory dynamics in competitive communities
-under environmental forcing. Oikos 117:1907--1911.
-
-Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
-A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
-Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
-C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
-Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
-Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
-Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
-Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
-Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
-Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
-Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
-operating space for humanity. Nature 461:472--475.
-
-Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
-for Latent Gaussian Models Using Integrated Nested Laplace
-Approximations. Journal of the Royal Statistical Society, Series B
-71:319--392.
-
-Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
-Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
-analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
-
-Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
-diversity, regulates temporal stability of plant communities. Oecologia
-166:761--768.
-
-Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
-function with non-random species loss. Ecology Letters 6:509--517.
-
-Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
-189:732--735.
-
-Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
-and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
-
-Tilman, D. 1988. Plant strategies and the dynamics and structure of
-plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
-
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
-
-Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
-2001. Diversity and productivity in a long-term grassland experiment.
-Science 294:843--845.
-
-Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
-stabilize food web dynamics by increasing synchrony. Ecology Letters
-10:1066--1074.
-
-Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
-Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
-2013. Global meta-analysis reveals no net change in local-scale plant
-biodiversity over time. Proceedings of the National Academy of Sciences
-of the United States of America 110:19456--9.
-
-Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
-2013. Shared ancestry influences community stability by altering
-competitive interactions: evidence from a laboratory microcosm
-experiment using freshwater green algae. Proceedings of the Royal
-Society of London B: Biological Sciences 280.
-
-Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
-Human Domination of Earth' s Ecosystems. Science 277:494--499.
-
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
-
-Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
-
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
-
-Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
-sagebrush steppe: long-term data for analyzing demographic rates and
-plant--plant interactions. Ecology 91:3427.
-
-\newpage{}
-
-\section{Tables}
-
-\singlespacing
+\vspace{2em}
 
 Table 1: Site descriptions and focal species. \footnotesize
-
 \begin{longtable}[c]{@{}lllll@{}}
@@
 
-\pagebreak{}
+\paragraph{Calculating observed
+synchrony.--}\label{calculating-observed-synchrony.}
+
+The data consist of records for individual plant size in quadrats for
+each year. To obtain estimates of percent cover for each focal species
+in each year, we summed the individual-level data within quadrats and
+then averaged percent cover, by species, over all quadrats. We
+calculated per capita growth rates as
+\(\text{log}(x_t) - \text{log}(x_{t-1})\), where \emph{x} is species'
+percent cover in year \emph{t}. Using the community time series of per
+capita growth rates or percent cover, we calculated community synchrony
+using the metric of Loreau and {{de Mazancourt}} (2008) in the
+\texttt{synchrony} package (Gouhier and Guichard 2014) in \texttt{R} (R Core Team 2013).
+Specifically, we calculated synchrony as
+
+\begin{equation}
+\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
+\end{equation}
+
+\noindent where \(\sigma_{r_{i}}\) is the temporal
+standard deviation of species \emph{i'}s per capita
+population growth rate (\(r_i\)) and \(\sigma^{2}_{T}\) is the temporal
+variance of the aggregate community-level growth rate. \(\phi\) ranges
+from 0 at perfect asynchrony to 1 at perfect synchrony (Loreau and {{de
+Mazancourt}} 2008). We use the same equation to calculate observed
+synchrony of species' percent cover, which we present to relate our
+results to previous findings, even though we focus on synchrony of
+growth rates in our model simulations (see below).
+
+\paragraph{Fitting statistical
+models.--}\label{fitting-statistical-models.}
+
+Vital rate regressions are the building blocks of our dynamic models: an
+integral projection model (IPM) and an individual-based model (IBM). We
+followed the approach of Chu and Adler (2015) to fit statistical models
+for survival, growth, and recruitment (see Appendix S1 for full
+details). We modeled survival probability of each genet as function of
+genet size, temporal variation among years, permanent spatial variation
+among groups of quadrats, and local neighborhood crowding from
+conspecific and heterospecific genets. Regression coefficients for the
+effect of crowding by each species can be considered a matrix of
+interaction coefficients whose diagonals represent intraspecific
+interactions and whose off-diagonals represent interspecific
+interactions (Adler et al. 2010). These interaction coefficients can
+take positive (facilitative) or negative (competitive) values. We
+modeled growth as the change in size of a genet from one year to the
+next, which depends on the same factors as the survival model. We fit
+the survival and growth regressions using INLA (Rue et al. 2014), a
+statistical package for fitting generalized linear mixed effects models
+via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
+2013). Spatial (quadrat groupings) variation was treated as a random
+effect on the intercept and temporal (interannual) variation was treated
+as random effects on the intercept and the effect of genet size in the
+previous year (Appendix S1).
+
+Interspecific and intraspecific crowding, which represent species interactions, can be included as fixed effects or as random effects that vary each year.
+Strong year-to-year variation in these crowding effects would indicate a statistical interaction between environmental conditions and species interactions.
+We tested for such a dynamic by comparing models with and without random year effects on crowding.
+Based on the results, we decided to treat crowding as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including interannual variation in crowding did not change our results.
+
+We modeled recruitment at the quadrat scale, rather than the individual
+scale, because the original data do not attribute new genets to specific
+parents (Chu and Adler 2015). Our recruitment model assumes that the
+number of recruits produced in each year follows a negative binomial
+distribution with the mean dependent on the cover of the parent species,
+permanent spatial variation among groups, temporal variation among
+years, and inter- and intraspecific interactions as a function of total
+species' cover in the quadrat. We fit the recruitment model using a
+hierarchical Bayesian approach implemented in JAGS (Plummer 2003) using
+the \texttt{rjags} package (Plummer 2014) in \texttt{R} (R Core Team 2013). Again,
+temporal and spatial variation were treated as random effects.
+
+\paragraph{Building dynamic multi-species
+models.--}\label{building-dynamic-multi-species-models.}
+
+Once we have fit the vital rate statistical models, building the
+population models is straightforward. For the IBM, we initialize the
+model by randomly assigning plants spatial coordinates, sizes, and
+species identities until each species achieves a density representative
+of that observed in the data. We then project the model forward by using
+the survival regression to determine whether a genet lives or dies, the
+growth regression to calculate changes in genet size, and the
+recruitment regression to add new individuals that are distributed
+randomly in space. Crowding is directly calculated at each time step
+since each genet is spatially referenced. Environmental stochasticity is
+not an inherent feature of IBMs, but is easily included since we fit
+year-specific temporal random effects for each vital rate regression. To
+include temporal environmental variation, at each time step we randomly
+choose a set of estimated survival, growth, and recruitment parameters
+specific to one observation year. For all simulations, we ignore the
+spatial random effect associated with variation among quadrat groups, so
+our simulations represent an average quadrat for each site.
+
+The IPM uses the same vital rate regressions as the IBM, but it is
+spatially implicit and does not include demographic stochasticity.
+Following Chu and Adler (2015), we use a mean field approximation that
+captures the essential features of spatial patterning to define the
+crowding index at each time step (Supporting Online Information).
+Temporal variation is included in exactly the same way as for the IBM.
+For full details on the IPMs we use, see Chu and Adler (2015).
+
+\paragraph{Simulation experiments.--}\label{simulation-experiments.}
+
+We performed simulation experiments where drivers (environmental
+stochasticity, demographic stochasticity, or interspecific interactions)
+were removed one-by-one and in combination. To remove interspecific
+interactions, we set the off-diagonals of the interaction matrix for
+each vital rate regression to zero. This retains intraspecific
+interactions, and thus density-dependence, and results in simulations
+where species are growing in isolation. We cannot definitively rule out
+the effects of species interactions on all parameters, meaning that a
+true monoculture could behave differently than our simulations of a
+population growing without interspecific competitors. To remove the
+effect of a fluctuating environment, we removed the temporal
+(interannual) random effects from the regression equations. To remove
+the effect of demographic stochasticity, we use the IPM rather than the
+IBM because the IPM does not include demographic stochasticity
+(demographic stochasticity cannot be removed from the IBM). Since the
+effect of demographic stochasticity on population dynamics depends on
+population size (Lande et al. 2003), we can control the strength of
+demographic stochasticity by simulating the IBM on areas (e.g.~plots) of
+different size. Results from an IBM with infinite population size would
+converge on results from the IPM. Given computational constraints, the
+largest landscape we simulate is a 25 \(\text{m}^2\) plot.
+
+We conducted the following six simulation experiments: (1) IBM: All
+drivers (environmental stochasticity, demographic stochasticity, or
+interspecific interactions) present; (2) IPM: Demographic stochasticity
+removed; (3) IBM: Environmental stochasticity removed; (4) IBM:
+Interspecific interactions removed; (5) IPM: Interspecific interactions
+and demographic stochasticity removed; (6) IBM: Interspecific
+interactions and environmental stochasticity removed.
+We did not include a simulation with only interspecific interactions because our population models run to deterministic equilibriums in the absence of environmental or demographic stochasticity.
+We ran IPM simulations for 2,000 time steps, after an initial 500
+iteration burn-in period. This allowed species to reach their stable
+size distribution. We then calculated yearly per capita growth rates
+from the simulated time series, and then calculated the synchrony of
+species' per capita growth rates over 100 randomly selected contiguous
+50 time-step sections.
+
+We ran IBM simulations for 100 time steps, and repeated the simulations
+100 times for each simulation experiment. From those, we retained only
+the simulations in which no species went extinct due to demographic
+stochasticity. Synchrony of per capita growth rates was calculated over
+the 100 time steps for each no extinction run within a model experiment.
+In the IBM simulations, the strength of demographic stochasticity should weaken as population size increases, meaning that synchrony should be less influenced by demographic stochasticity in large populations compared to small populations.
+To explore this effect, we ran simulations (4) and
+(6) on plot sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM
+simulations were run on a 25 \(\text{m}^2\) landscape
+to most closely match the implicit, large spatial scale of the IPM simulations.
+
+Our simulations allow us to quantify the relative importance of environmental responses, species interactions, and demographic stochasticity by comparing the simulated values of community-wide species synchrony.
+The simulation experiments also allow us to test our theoretical
+predictions. First, in the absence of interspecific interactions and
+demographic stochasticity, populations can only fluctuate in response to
+the environment. Therefore, we can use results from simulation (5) to
+estimate the covariance of species' responses to the environment
+(\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation 4.
+Parameterizing Equation 3 does not require simulation output because the
+only parameters are the species' relative abundances. Second,
+simulations (5) and (6) represent the simulated version of our limiting
+case theoretical predictions.
+This approach to testing theoretical predictions may seem circular, but recall we derived the predictions using strict assumptions about equivalence in species' growth rates, environmental response variances, and demographic variances.
+Our empirically-based models do not make these assumptions.
+Comparisons between parameterized predictions and simulated synchrony reveal whether the assumptions we must make to derive analytical predictions hold in the natural community our model represents.
+
+All \texttt{R} code necessary to reproduce our results has been deposited on Figshare (\url{http://doi.org/10.6084/m9.figshare.4587073}) and released on \textsc{GitHub} (\url{http://github.com/atredennick/community_synchrony/releases/tag/v1}).
+
+\subsection{Results}
+
+Observed synchrony of species' per capita growth rates
+at our study sites range from 0.36 to 0.89 and synchrony of percent
+cover ranged from 0.15 to 0.92 (Table 2). Synchrony of per capita growth
+rates and CV of percent cover are positively correlated (Pearson's
+\(\rho\) = 0.72, \emph{N} = 5). For all five communities, species
+synchrony from IPM and IBM simulations closely approximated observed
+synchrony (Fig. S1). IBM-simulated synchrony is consistently, but only
+slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to
+the desynchronizing effect of demographic stochasticity.
 
@@
 
-\pagebreak{}
-
-\doublespacing
-
-\section{Figure Legends}
-
-\textbf{FIGURE 1:} Diagram of our coupled theoretical-empirical approach. We followed this workflow for each of our five focal communities.
-
-\textbf{FIGURE 2:} Community-wide species synchrony of per capita growth rates from model simulation experiments.
+Across the five communities, our limiting case predictions closely
+matched synchrony from the corresponding simulation experiment (Fig. 2
+and Table S1). The correlation between our analytical predictions and
+simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) (\emph{N} =
+5) and 0.997 for \(\phi_{R,\mathcal{M}_E}\) (\emph{N} = 5). The largest
+difference between predicted and simulated synchrony was 0.05 in New
+Mexico for \(\phi_{R,\mathcal{M}_D}\) (Table S1).
+
+Simulation experiments revealed that removing environmental fluctuations
+has the largest impact on synchrony, leading to a reduction in synchrony
+of species growth rates in four out of five communities (Fig. 2).
+Removing environmental fluctuations (``No E.S'' simulations) decreased
+synchrony by 33\% in Arizona, 48\% in Kansas, 39\% in Montana, and 40\%
+in New Mexico. Only in Idaho did removing environmental fluctuations
+cause an increase in synchrony (Fig. 2), but the effect was small (9\%
+increase; Table S2). Overall, species' temporal random effects in the
+statistical vital rate models are positively, but not perfectly,
+correlated (Table S3).
+These temporal random effects represent environmental responses, meaning that positively correlated temporal random effects indicate positively correlated environmental responses.
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure1.png}
+  \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments.
 Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific interactions present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific interactions removed (``No Comp.''), interspecific interactions and demographic stochasticity removed (``No Comp.
@@
 All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape.
-Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediction.
-
-\textbf{FIGURE 3:} Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
+Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical prediciton.}
+\end{figure}
+
+Species interactions are weak in these communities (Table S4 and Chu and
+Adler 2015), and removing interspecific interactions had little effect
+on synchrony (Fig. 2; ``No Comp.'' simulations). Removing interspecific
+interactions caused, at most, a 5\% change in synchrony (Fig. 2 and
+Table S2). Removing demographic stochasticity (``No D.S.'' simulations)
+caused synchrony to increase slightly in all communities (Fig. 2), with
+an average 6\% increase over synchrony from IBM simulations on a
+25\(\text{m}^2\) area.
+
+In IBM simulations, the desynchronizing effect of demographic
+stochasticity, which increases as population size decreases, modestly
+counteracted the synchronizing force of the environment, but not enough
+to lower synchrony to the level observed when only demographic
+stochasticity is operating (Fig. 3). In the largest, 25 \(\text{m}^2\)
+plots, synchrony was driven by environmental stochasticity (e.g.,
+\(\mathcal{M}_E\)). At 1 \(\text{m}^2\), synchrony reflected
+the combined effects of demographic stochasticity and
+environmental stochasticity (e.g.,
+simulated synchrony fell between \(\mathcal{M}_E\) and
+\(\mathcal{M}_D\)). For context, population sizes increased from an
+average of 17 individuals per community in 1 \(\text{m}^2\) IBM
+simulations to an average of 357 individuals per community in 25
+\(\text{m}^2\) IBM simulations.
+
+Results for synchrony of percent cover are qualitatively similar, but
+more variable and less consistent with analytical predictions and
+observed synchrony (Appendix S1, Figs. S2-S3).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
+  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochasticity is present (``No Comp.
 + No E.S.
@@
 + E.S.)'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing.
-Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.
+Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}
+\end{figure}
+
+\section{Discussion}
+
+Our study produced four main findings that were generally consistent
+across five natural grassland plant communities: (1) limiting-case predictions
+from the theoretical model were well-supported by simulations from the
+empirical models; (2) demographic stochasticity decreased community
+synchrony, as expected by theory, and its effect was largest in small
+populations; (3) environmental fluctuations increased community
+synchrony relative to simulations in constant environments because
+species-specific responses to the environment were positively, though
+not perfectly, correlated; and (4) interspecific interactions were weak
+and therefore had little impact on community synchrony. We also found
+that analyses based on synchrony of species' percent cover, rather than
+growth rates, were uninformative (Figs. S2-S3) since the linear
+approximation required for analytical predictions is a stronger
+assumption for abundance than growth rates, especially given relatively
+short time-series (Appendix S1). Thus, our results provide further
+evidence that it is difficult to decipher mechanisms of species
+synchrony from abundance time series, as expected by theory (Loreau and
+{{de Mazancourt}} 2008). Observed synchrony of per capita growth rates
+was positively correlated with the variability of percent cover across
+our focal communities, which confirms that we are investigating an
+important process underlying ecosystem stability.
+
+\subsection{Simulations support theoretical predictions}
+
+Our theoretical predictions were derived from a simple model of
+population dynamics and required several restrictive assumptions,
+raising questions about their relevance to natural communities. For
+example, the species in our communities do not have equivalent
+environmental and demographic variances (Figs. S4-S7), as required by
+our predictions. However, the theoretical predictions closely matched
+results from simulations of population models fit to long-term data from
+natural plant communities (Table S1). Strong agreement between our
+analytical predictions and the simulation results should inspire
+confidence in the ability of simple models to inform our understanding
+of species synchrony even in complex natural communities, and allows us
+to place our simulation results within the context of contemporary
+theory.
+However, whether the theoretical model adequately represents more complex communities remains unknown because our analysis was restricted to dominant species.
+
+\subsection{Demographic stochasticity decreases synchrony in small populations}
+
+In large populations, removing demographic stochasticity had no effect
+on species synchrony (Fig. 2). In small populations, demographic
+stochasticity partially counteracted the synchronizing effects of
+environmental fluctuations and interspecific interactions on per capita
+growth rates, in agreement with theory (Loreau and {{de Mazancourt}}
+2008).
+This is shown in Fig. 3, where IBM simulations with environmental forcing and demographic stochasticity have higher synchrony than simulations with only demographic stochasticity.
+The differences between simulations are smaller at lower population sizes because as the number of individuals decreases with area, the strength of demographic stochasticity increases, reducing the relative effect of environmental forcing.
+Even in small populations
+(e.g., 1 $\text{m}^2$ lanscapes), however, demographic
+stochasticity was not strong enough to compensate the synchronizing
+effects of environmental fluctuations and match the analytical
+prediction where only demographic stochasticity is operating
+(\(\mathcal{M}_D\) in Fig. 3). These results confirm the theoretical
+argument of Loreau and {{de Mazancourt}} (2008) that independent
+fluctuations among interacting species in a non-constant environment
+should be rare. Only in the Idaho community does synchrony of per capita
+growth rates approach \(\mathcal{M}_D\) in a non-constant environment
+(Fig. 3). This is most likely due to the strong effect of demographic
+stochasticity on the shrub \emph{Artemisia tripartita} since even a 25
+\(\text{m}^2\) quadrat would only contain a few individuals of that
+species.
+
+Our analysis of how demographic stochasticity affects synchrony
+demonstrates that synchrony depends on the observation area. As the
+observation area increases, population size increases and the
+desynchronizing effect of demographic stochasticity lessens (Fig. 3).
+Thus, our results suggest that community-wide species synchrony will
+increase as the observation area increases, rising from
+\(\mathcal{M}_D\) to \(\mathcal{M}_E\). Such a conclusion assumes,
+however, that species richness remains constant as observation area
+increases, which is unlikely (Taylor 1961). Recent theoretical work has
+begun to explore the linkage between ecosystem stability and spatial
+scale (Wang and Loreau 2014, 2016), and our results suggest that
+including demographic stochasticity in theoretical models of
+metacommunity dynamics may be important for understanding the role of
+species synchrony in determining ecosystem stability across spatial
+scales.
+
+\subsection{Environmental fluctuations drive community synchrony}
+
+In large populations where interspecific interactions are weak,
+synchrony is expected to be driven exclusively by environmental
+fluctuations (Equation 4). Under such conditions community synchrony
+should approximately equal the synchrony of species' responses to the
+environment (Loreau and {{de Mazancourt}} 2008). Two lines of evidence
+lead us to conclude that environmental fluctuations drive species
+synchrony in our focal plant communities. First, in our simulation
+experiments, removing interspecific interactions resulted in no
+discernible change in community-wide species synchrony of per capita
+growth rates (Fig. 2). Second, removing environmental fluctuations from
+simulations consistently reduced synchrony (Fig. 2). Our results lead us
+to conclude that environmental fluctuations, not species interactions,
+are the primary driver of community-wide species synchrony
+among the dominant species we studied. Given
+accumulating evidence that niche differences in natural communities are
+large (reviewed in Chu and Adler 2015), and thus species interactions
+are likely to be weak, our results may be general in natural plant
+communities.
+
+In the Idaho community, removing environmental fluctuations did not
+cause a large decrease in synchrony. However, that result appears to be
+an artifact. Removing environmental variation results in a negative
+invasion growth rate for \emph{A. tripartita}. Although we only analyzed
+IBM runs in which \emph{A. tripartita} had not yet gone extinct, it was
+at much lower abundance than in the other simulation runs. When we
+removed \emph{A. tripartita} from all simulations, the Idaho results
+conformed with results from all other sites: removing environmental
+stochasticity caused a significant reduction in species synchrony (Fig.
+S8). Our main results for Idaho (Fig. 2), with \emph{A. tripartita}
+included, demonstrate how the processes that determine species synchrony
+interact in complex ways. \emph{A. tripartita} has a facilitative effect
+on each grass species across all vital rates, except for a small
+competitive effect on \emph{H. comata}'s survival probability (Tables
+S8-S10). At the same time, all the perennial grasses have negative
+effects on each other for each vital rate (Tables S8-S10). We know
+synchrony is affected by interspecific competition (Loreau and {{de
+Mazancourt}} 2008), but how facilitative effects manifest themselves is
+unknown. The interaction of facilitation and competition is clearly
+capable of having a large effect on species synchrony, and future
+theoretical efforts should aim to include a wider range of species
+interactions.
+
+Environmental responses synchronized dynamics relative to a null expectation of independent species interactions (e.g., "No Comp.
++ No E.S." simulations in Fig.
+2), but observed and simulated synchrony was still less than one in all cases (Fig.
+2).
+Synchrony was far from complete because of differences in species' responses to interannual environmental variation.
+Many studies of ecosystem stability in semiarid grasslands focus on trad-offs among dominant functional types
+(Bai et al. 2004, Sasaki and Lauenroth 2011).
+Such groupings are based on the idea that ecologically-similar species will have similar responses to environmental fluctuations.
+At first glance our results may appear to support the grouping of perennial grasses in one functional type because their environmental responses were positively correlated.
+However, even though environmental responses among the dominant species we studied were similar, they were dissimilar enough to cause synchrony to be less than perfect (Fig. 2).
+The subtle differences among ecologically-similar dominant species do impact species synchrony and, ultimately, ecosystem stability.
+Ignoring such differences could mask important dynamics that underpin ecosystem functioning.
+
+\subsection{Interspecific interactions had little impact on community synchrony}
+
+We expected community synchrony of per capita growth rates to decrease
+when we removed interspecific interactions (Loreau and {{de Mazancourt}}
+2008). We found that community synchrony was virtually indistinguishable
+between simulations with and without interspecific interactions (Fig.
+2). The lack of an effect of interspecific interactions on synchrony is
+in contrast to a large body of theoretical work that predicts a strong
+role for competition in creating compensatory dynamics (Tilman 1988) and
+a recent empirical analysis (Gross et al. 2014).
+
+Our results do not contradict the idea that competition can lead to
+compensatory dynamics, but they do highlight the fact that interspecific
+competition must be relatively strong to influence species synchrony.
+The communities we analyzed are composed of species with very little
+niche overlap (Chu and Adler 2015) and weak interspecific interactions
+(Tables S1 and S3-S17). Mechanistic consumer-resource models (Lehman and
+Tilman 2000) and phenomenological Lotka-Volterra models (Lehman and
+Tilman 2000, Loreau and {{de Mazancourt}} 2013) both confirm that the
+effect of competition on species synchrony diminishes as niche overlap
+decreases. In that sense, our results are not surprising: interspecific
+interactions are weak, so of course removing them does not affect
+synchrony.
+
+Our conclusion that species interactions have little impact on synchrony
+only applies to single trophic level communities. Species interactions
+almost certainly play a strong role in multi-trophic communities where
+factors such as resource overlap (Vasseur and Fox 2007), dispersal
+(Gouhier et al. 2010), and the strength of top-down control (Bauer et
+al. 2014) are all likely to affect community synchrony.
+
+\section{Conclusions}
+
+Species-specific responses to temporally fluctuating environmental
+conditions is an important mechanism underlying asynchronous population
+dynamics and, in turn, ecosystem stability (Loreau and {{de Mazancourt}}
+2013). When we removed environmental variation, we found that synchrony
+decreased in four out of the five grassland communities we studied (Fig.
+2). A tempting conclusion is that our study confirms that compensatory
+dynamics are rare in natural communities, and that ecologically-similar
+species will exhibit synchronous dynamics (e.g., Houlahan et al. 2007).
+Such a conclusion misses an important subtlety. The perennial grasses we
+studied do have similar responses to the environment (Table S2), which
+will tend to synchronize dynamics. However, if community-wide synchrony
+among dominant species is less than perfect, as it is
+in all our focal communities, some degree of compensatory dynamics must
+be present (Loreau and {{de Mazancourt}} 2008).
+Even ecologically-similar species, which are sometimes aggregated into functional groups, have environmental responses that are dissimilar enough to limit synchrony.
+Subtle differences among dominant species ultimately determine ecosystem stability and should not be ignored.
+
+\section{Acknowledgements}
+
+This work was funded by the National Science Foundation through a
+Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
+CAREER award to PBA (DEB-1054040). CdM and ML were supported by the
+TULIP Laboratory of Excellence (ANR-10-LABX-41). We thank the original
+mappers of the permanent quadrats at each site and the digitizers in the
+Adler lab, without whom this work would not have been possible. Compute,
+storage, and other resources from the Division of Research Computing in
+the Office of Research and Graduate Studies at Utah State University are
+gratefully acknowledged. We also thank Patrick Venail and several
+anonymous reviewers who provided critical feedback that improved the
+manuscript.
+
+\setlength{\parindent}{0ex}
+
+\section{Literature Cited}
+
+Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
+perennial plants: An embarrassment of niches. Ecology Letters
+13:1019--1029.
+
+Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
+mapped quadrats from Kansas prairie: demographic information for
+herbaceous plants. Ecology 88:2673.
+
+Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
+of semi-desert grassland plants in permanent quadrats mapped from 1915
+to 1947. Ecology 93:1492--1492.
+
+Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
+mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
+92:1703.
+
+Bai, Y., X. Han, J. Wu, Z. Chen, and L. Li. 2004. Ecosystem stability
+and compensatory effects in the Inner Mongolia grassland. Nature
+431:181--184.
+
+Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
+Functional Similarity, and Top-Down Control Drive Synchronization and
+the Reliability of Ecosystem Function. The American Naturalist
+183:394--409.
+
+Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
+P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
+Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
+and its impact on humanity. Nature 489:59--67.
+
+Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
+recruitment stage to stabilize grassland coexistence. Ecological
+Monographs 85:373--392.
+
+{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
+Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
+Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
+Loreau. 2013. Predicting ecosystem stability from community composition
+and biodiversity. Ecology Letters 16:617--625.
+
+Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
+Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
+biodiversity change but not systematic loss. Science 344:296--9.
+
+Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. Annual Review of
+Ecology, Evolution, and Systematics 40:393--414.
+
+Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
+variability in space and time. Methods in Ecology and Evolution
+5:524--533.
+
+Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
+Stability of Food Webs in Metacommunities. The American Naturalist
+175:E16--E34.
+
+Grman, E., J. A. Lau, D. R. Schoolmaster, and K. L. Gross. 2010.
+Mechanisms contributing to stability in ecosystem function depend on the
+environmental context. Ecology Letters 13:1400--1410.
+
+Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
+Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
+temporal stability of biomass production: a new analysis of recent
+biodiversity experiments. The American naturalist 183:1--12.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
+on ecosystem functioning: a consensus of current knowledge. Ecological
+Monographs 75:3--35.
+
+Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
+Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
+Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
+Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
+Compensatory dynamics are rare in natural ecological communities.
+Proceedings of the National Academy of Sciences of the United States of
+America 104:3273--3277.
+
+Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
+productivity and the temporal stability of productivity: Patterns and
+processes. Ecology Letters 12:443--451.
+
+Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
+dynamics in ecology and conservation.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
+drivers: neutral and nonneutral community dynamics in fluctuating
+environments. The American Naturalist 172:E48--E66.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+Loreau, M., and a Hector. 2001. Partitioning selection and
+complementarity in biodiversity experiments. Nature 412:72--6.
+
+Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
+Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
+and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
+knowledge and future challenges. Science 294:804--808.
+
+Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
+community dynamics? Proceedings of the Royal Society B: Biological
+Sciences 276:2923--2929.
+
+Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
+Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
+3rd international workshop on distributed statistical computing (dSC
+2003). march.
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+
+R Core Team. 2013. R: A language and environment for statistical
+computing.
+
+Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
+O'Hara. 2008. Detecting compensatory dynamics in competitive communities
+under environmental forcing. Oikos 117:1907--1911.
+
+Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
+A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
+Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
+C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
+Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
+Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
+Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
+Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
+Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
+Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
+Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
+operating space for humanity. Nature 461:472--475.
+
+Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
+for Latent Gaussian Models Using Integrated Nested Laplace
+Approximations. Journal of the Royal Statistical Society, Series B
+71:319--392.
+
+Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
+Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
+analysis of latent Gaussian models using Integrated Nested Laplace
+Approximaxion.
+
+Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
+diversity, regulates temporal stability of plant communities. Oecologia
+166:761--768.
+
+Smith, M. D., and A. K. Knapp. 2003. Dominant species maintain ecosystem
+function with non-random species loss. Ecology Letters 6:509--517.
+
+Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
+189:732--735.
+
+Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
+and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.
+
+Tilman, D. 1988. Plant strategies and the dynamics and structure of
+plant communities. Plant strategies and the dynamics and structure of
+plant communities.:360 pp.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
+2001. Diversity and productivity in a long-term grassland experiment.
+Science 294:843--845.
+
+Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
+stabilize food web dynamics by increasing synchrony. Ecology Letters
+10:1066--1074.
+
+Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
+Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
+2013. Global meta-analysis reveals no net change in local-scale plant
+biodiversity over time. Proceedings of the National Academy of Sciences
+of the United States of America 110:19456--9.
+
+Venail, P. A., M. A. Alexandrou, T. H. Oakley, and B. J. Cardinale.
+2013. Shared ancestry influences community stability by altering
+competitive interactions: evidence from a laboratory microcosm
+experiment using freshwater green algae. Proceedings of the Royal
+Society of London B: Biological Sciences 280.
+
+Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
+Human Domination of Earth' s Ecosystems. Science 277:494--499.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
+across scales in metacommunities. Ecology Letters.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
+sagebrush steppe: long-term data for analyzing demographic rates and
+plant--plant interactions. Ecology 91:3427.
 

### 2017-03-10 docs/ms_communitySynchrony_ForECOLOGY_PREPRINT.tex
adds proof corrections to preprint version for EU posting

@@
 \usepackage{color, soul}
-\usepackage{times}
+%\usepackage{times}
 \usepackage[document]{ragged2e}
@@
 \titleformat*{\paragraph}{\normalsize\it}
+ \usepackage{bm,mathrsfs}
+  \usepackage{mathptmx}
 
@@
 \begin{center}
-\large{\textbf{Environmental responses, not species interactions, determine synchrony of dominant species in semiarid grasslands}}
+\Large{Environmental responses, not species interactions, determine synchrony of dominant species in semiarid grasslands}
 
@@
 Removing interspecific interactions and demographic stochasticity had small effects on synchrony. 
-For the dominant species in these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
+For the dominant species in these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species' responses to the environment.
 
@@
 baselines to help us interpret results from empirically-based
-simulations (described below). We focus on synchrony of per capita
+simulations (see \textsc{Empirical Analysis}). We focus on synchrony of per capita
 growth rates, rather than abundances, because growth rates represent the
@@
 results to previous findings, even though we focus on synchrony of
-growth rates in our model simulations (see below).
+growth rates in our model simulations (see \emph{Simulation experiments}).
 
@@
 effect on the intercept and temporal (interannual) variation was treated
-as random effects on the intercept and the effect of genet size in the
+as random effects on both the intercept and the effect of genet size in the
 previous year (Appendix S1).
@@
 evidence that it is difficult to decipher mechanisms of species
-synchrony from abundance time series, as expected by theory (Loreau and
+synchrony from abundance time series, as predicted by theory (Loreau and
 {{de Mazancourt}} 2008). Observed synchrony of per capita growth rates
@@
 CAREER award to PBA (DEB-1054040). CdM and ML were supported by the
-TULIP Laboratory of Excellence (ANR-10-LABX-41). We thank the original
+TULIP Laboratory of Excellence (ANR-10-LABX-41) and by the BIOSTASES 
+Advanced Grant, funded by the European Research Council under the European 
+Union's Horizon 2020 research and innovation programme (grant agreement No 666971).
+We thank the original
 mappers of the permanent quadrats at each site and the digitizers in the
@@
 
-Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
-Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
-3rd international workshop on distributed statistical computing (dSC
-2003). march.
-
-Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.
+Plummer, M.
+2003.
+JAGS: a program for analysis of Bayesian graphical models using Gibbs sampling.
+Proceedings of the Third International Workshop on Distributed Statistical Computing, 20?22 March 2003, Vienna, Austria.
+http://www.r-project.org/conferences/DSC-2003/Proceedings/Plummer.pdf
+
+Plummer, M. 2014. rjags: Bayesian graphical models using MCMC. R package version 4-6. http://cran.r-project.org/package=rjags
 
 R Core Team. 2013. R: A language and environment for statistical
-computing.
+computing. https://www.r-project.org/
 
@@
 analysis of latent Gaussian models using Integrated Nested Laplace
-Approximaxion.
+Approximaxion. R package version 0.0-1483776432. http://www.r-inla.org/home.
 
@@
 plant communities. Plant strategies and the dynamics and structure of
-plant communities.:360 pp.
+plant communities. Princeton University Press, Princeton, New Jersey.
 
@@
 experiment using freshwater green algae. Proceedings of the Royal
-Society of London B: Biological Sciences 280.
+Society of London B: Biological Sciences 280:20131548.
 
@@
 Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
-across scales in metacommunities. Ecology Letters.
+across scales in metacommunities. Ecology Letters 19:510--518.
 
